### 2020-11-03 avalanche.tex
add initial source and 'fr' make target

\subsection{Avalanche: Adding a DAG}%\tronly{}{\vspace{-0.5em}}

Avalanche consists of multiple single-decree Snowball instances instantiated as a multi-decree protocol that
maintains a dynamic, append-only directed acyclic graph (DAG) of all known transactions.
The DAG has a single sink that is the \emph{genesis vertex}.
Maintaining a DAG provides two significant benefits.
First, it improves efficiency, because a single vote on a DAG vertex implicitly votes for all transactions on the path to the genesis vertex.
Second, it also improves security, because the DAG intertwines the fate of transactions, similar to the Bitcoin blockchain.
This renders past decisions difficult to undo without the approval of correct nodes.

% Avalanche's DAG embodies all the transactions that have been proposed to the system, where each transaction is represented as a vertex.
% The DAG is rooted at a protocol-defined, well-known \emph{genesis} vertex.
When a client creates a transaction, it names one or more \emph{parents}, which are included inseparably in the transaction and form the edges of the DAG\@.
The parent-child relationships encoded in the DAG may, but do not need to, correspond to application-specific dependencies; for instance, a child transaction need not spend or have any relationship with the funds received in the parent transaction.
% These additional relationships entangle the fate of previous decisions made by the system.
We use the term \emph{ancestor set} to refer to all transactions reachable via parent edges back in history, and \emph{progeny} to refer to all children transactions and their offspring.

The central challenge in the maintenance of the DAG is to choose among \emph{conflicting transactions}.
The notion of conflict is application-defined and transitive, forming an equivalence relation.
In our cryptocurrency application, transactions that spend the same funds (\emph{double-spends}) conflict, and form a \emph{conflict set}
%(Figure~\tronly{\ref{fig:dag-conflict-set}}{\ref{fig:dag-cd}})
(shaded regions in Figure~\ref{fig:dag-cd}), out of which only a single one can be accepted.
Note that the conflict set of a virtuous transaction is always a singleton.
%Note that the graph structure of transactions that spend depend on each other, also known as the UTXO graph, is completely independent of the DAG that Avalanche maintains. As a result, any two vertices may be in conflict. Figure~\ref{fig:dag-conflicting-set} shows an example.

%\tronly{
%\begin{figure}\begin{center}
%    \input{figures/dag-conflict-set.tex}
%    \captionof{figure}{DAG vertices partitioned by conflict sets. At most one vertex in each shaded region will be accepted.}\label{fig:dag-conflict-set}
%\end{center}
%\end{figure}
%}{}

Avalanche instantiates a Snowball instance for each conflict set.
Whereas Snowball uses repeated queries and multiple counters to capture the amount of confidence built in conflicting transactions (colors),
Avalanche takes advantage of the DAG structure and uses a transaction's progeny.
Specifically, when a transaction $T$ is queried, all transactions reachable from $T$ by following the DAG edges are implicitly part of the query.
A node will only respond positively to the query if $T$ and its entire ancestry are currently the preferred option in their respective conflict sets.
If more than a threshold of responders vote positively, the transaction is said to collect a \emph{chit}.
Nodes then compute their \emph{confidence} as the total number of chits in the progeny of that transaction.
They query a transaction just once and rely on new vertices and possible chits, added to the progeny, to build up their confidence.
Ties are broken by an initial preference for first-seen transactions.
Note that chits are decoupled from the DAG structure, making the protocol immune to attacks where
the attacker generates large, padded subgraphs.

\subsection{Avalanche: Specification}%\tronly{}{\vspace{-0.5em}}
\label{subsection:specification}

\newcommand{\codeedges}{\mathit{edges}}
\newcommand{\codedata}{\mathit{data}}
\begin{figure}
\begin{center}
\small
\begin{algorithmic}[1]
    \Procedure{init}{}
        \State $\mathcal{T} \assign \emptyset$ \hspace{1ex}\textrm{// the set of known transactions}
        \State $\mathcal{Q} \assign \emptyset$ \hspace{1ex}\textrm{// the set of queried transactions}
    \EndProcedure
    \Procedure{onGenerateTx}{$\codedata$}
        \State $\codeedges \assign \{T' \gets T: T' \in \Call{parentSelection}{\mathcal{T}}\}$
        \State $T \assign \Call{Tx}{\codedata, \codeedges}$
        \State \Call{onReceiveTx}{$T$}
    \EndProcedure
    \Procedure{onReceiveTx}{$T$}
        \If{$T \notin \mathcal{T}$}
            \If{$\mathcal{P}_T = \emptyset$}
                \State $\mathcal{P}_T \assign \{T\}$, $\mathcal{P}_T\mathit{.pref} \assign T$
                \State $\mathcal{P}_T\mathit{.last} \assign T, \mathcal{P}_T\mathit{.cnt} \assign 0$
            \Else$\ \mathcal{P}_T \assign \mathcal{P}_T \cup \{T\}$
            \EndIf
            \State $\mathcal{T} \assign \mathcal{T} \cup \{T\}$, $c_T \assign 0$.
        \EndIf
    \EndProcedure
    \captionof{figure}{Avalanche: transaction generation.}\label{fig:gossipchain-ongen}
\end{algorithmic}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\small
\begin{algorithmic}[1]
    \Procedure{avalancheLoop}{}
        \While {\codetrue}
            \State\textrm{find  $T$ that satisfies }%\vspace*{-.6\baselineskip}
            $T \in \mathcal{T} \land T \notin \mathcal{Q}$
            %\begin{align*}\hspace{2ex}
            %    T \in \mathcal{T} &\land T \notin \mathcal{Q} \\
            %            &\land (\forall T', T' \gets T: T' \in \mathcal{Q})
            %\end{align*}
            \State $\mathcal{K} \assign \Call{sample}{\mathcal{N}\backslash u, k}$
            \State $P \assign \sum_{v \in \mathcal{K}}\Call{query}{v, T}$
            \If{$P \ge \alpha$}
                \State $c_T \assign 1$
            \State\textrm{// update the preference for ancestors}
            \For{$T' \in \mathcal{T} : T' \stackrel{*}{\gets} T$}
                \If{$d(T') > d(\mathcal{P}_{T'}\mathit{.pref})$}
                    \State $\mathcal{P}_{T'}\mathit{.pref} \assign T'$
                \EndIf
                \If{$T'\neq \mathcal{P}_{T'}\mathit{.last}$}
                    \State $\mathcal{P}_{T'}\mathit{.last} \assign T'$, $\mathcal{P}_{T'}\mathit{.cnt} \assign 1$
                \Else
                    \State \texttt{++}$\mathcal{P}_{T'}\mathit{.cnt}$
                \EndIf
            \EndFor
            \Else
            \For{$T' \in \mathcal{T} : T' \stackrel{*}{\gets} T$}
                    \State$\mathcal{P}_{T'}\mathit{.cnt} \assign 0$
            \EndFor
            \EndIf
            \State\textrm{// otherwise, }$c_T$\textrm{ remains 0 forever}
            \State $\mathcal{Q} \assign \mathcal{Q} \cup \{T\}$ \hspace {1ex} \textrm{// mark T as queried}
        \EndWhile
    \EndProcedure
    \captionof{figure}{Avalanche: the main loop.}\label{fig:gossipchain-main}
\end{algorithmic}
\end{center}
\end{figure}

% TODO IMPORTANT The isPreferred must imply that T is queried, but current writing can be not queried. Also, do you do max over only queried transactions?
\begin{figure}[t]
\begin{center}
\small
\begin{algorithmic}[1]
    \Function{isPreferred}{$T$}
        %\State \Return $d(T) = \max_{T' \in \mathcal{P}_T} d(T')$
        \State \Return $T = \mathcal{P}_T\mathit{.pref}$
    \EndFunction
    \Function{isStronglyPreferred}{$T$}
        \State \Return $\forall T'\in\mathcal{T}, T' \stackrel{*}{\gets} T: \Call{isPreferred}{T'}$
    \EndFunction
    \Function{isAccepted}{$T$}
        \State\Return
            \vspace*{-.5\baselineskip}
        \begin{align*}
            (&(\forall T' \in \mathcal{T}, T' \gets T: \Call{isAccepted}{T'}) \\
                &\land |\mathcal{P}_T| = 1 \land \mathcal{P}_T\mathit{.cnt} \ge \beta_1) \texttt{\hspace{.1in}// safe early commitment} \\
            %& \left(\frac{d(T)}{d'} > \gamma \land d' \ge \beta_2\right)
            \lor &(%\mathcal{P}_T\textrm{.pref} = \mathcal{P}_T\textrm{.last} \land
            \mathcal{P}_T\mathit{.cnt} \ge \beta_2)\texttt{\hspace{.1in}// consecutive counter}
        \end{align*}
    \EndFunction
    \Procedure{onQuery}{$j, T$}
        \State \Call{onReceiveTx}{$T$}
        \State \Call{respond}{$j, \textsc{isStronglyPreferred}(T)$}
    \EndProcedure
    \captionof{figure}{Avalanche: voting and decision primitives.}\label{fig:gossipchain-onquery}
\end{algorithmic}
\end{center}
\end{figure}

Each correct node $u$
keeps track of all transactions it has learned about in set $\mathcal{T}_u$,
partitioned into mutually exclusive conflict sets $\mathcal{P}_T$, $T \in \mathcal{T}_u$.
% Transactions are determined to be conflicting based on a deterministic function known to every node.
Since conflicts are transitive, if $T_i$ and $T_j$ are conflicting, then they belong to the same conflict set, i.e. $\mathcal{P}_{T_i} = \mathcal{P}_{T_j}$. This relation may sound counter-intuitive: conflicting transitions have the \emph{equivalence} relation, because they are equivocations spending the \emph{same} funds.
%\Jon{By $=$ surely you mean $\neq$?}

We write $T' \gets T$ if $T$ has a parent edge to transaction $T'$,
The ``$\stackrel{*}{\gets}$''-relation is its reflexive transitive closure, indicating a path from $T$ to $T'$.
DAGs built by different nodes are guaranteed to be compatible, though at any one time, the two nodes may not have a complete view of all vertices in the system.
Specifically, if $T' \gets T$, then every node in the system that has $T$ will also have $T'$ and the same relation $T' \gets T$; and conversely, if $T' \cancel{\gets} T$, then no nodes will end up with $T' \gets T$.

Each node $u$ can compute a confidence value, $d_u(T)$, from the progeny as follows:
\[ d_u(T) = \sum_{T' \in \mathcal{T}_u, T \stackrel{*}{\gets} T'}c_{uT'}\]
where $c_{uT'}$ stands for the chit value of $T'$ for node $u$. Each transaction initially has a chit value of $0$ before the node gets
the query results. If the node collects a threshold of $\alpha$ yes-votes after the query, the value $c_{uT'}$ is set to 1, otherwise remains $0$ forever.
Therefore, a chit value reflects the result from the one-time query of its associated transaction and becomes immutable afterwards, while $d(T)$ can increase as the DAG grows by collecting more chits in its progeny.
Because $c_T \in \{0, 1\}$, confidence values are monotonic.
%\Jon{the notation $c_{uT}$ shows up before it's definet. I guess it's the number of chits counted for $T$ by participant $u$?}

In addition, node $u$ maintains its own local list of known nodes $\mathcal{N}_u \subseteq \mathcal{N}$ that comprise the system.
For simplicity, we assume for now $\mathcal{N}_u = \mathcal{N}$, and elide subscript $u$ in contexts without ambiguity.
%
% TODO talk to ted and ask how to rephrase this. "the data structure" is unclear, and so is immutability
% This is because a transaction is immutable in the sense that its identity is uniquely determined by both the application data and parent links it embodies.
%Here, notation ``$T' \gets T$'' means $T'$ is one of $T$'s immediate parents, and
%``$\stackrel{*}{\gets}$''-relation is the reflexive transitive closure of ``$\gets$''-relation; namely,
%if $T' \stackrel{*}{\gets} T$, there exists a path from $T$ to $T'$. Specially, ``$T \stackrel{*}{\gets} T$''.

Each node implements an event-driven state machine, centered around a query that serves both to solicit votes on each transaction and to notify other nodes of the existence of newly discovered transactions.
In particular, when node $u$ discovers a transaction $T$ through a query, it starts a one-time query process by sampling $k$ random peers and sending a message to them, after $T$ is delivered via $\textsc{onReceiveTx}$.
% determining the set of conflicting transactions $\mathcal{P}_T$,

Node $u$ answers a query by checking whether each $T'$ such that $T' \stackrel{*}{\gets} T$ is currently preferred among competing transactions $\forall T'' \in \mathcal{P}_{T'}$.
If every single ancestor $T'$ fulfills this criterion, the transaction is said to be \emph{strongly preferred}, and receives a yes-vote (1). A failure of this criterion at any $T'$ yields a no-vote (0).
When $u$ accumulates $k$ responses, it checks whether there are $\alpha$ yes-votes for $T$, and if so grants the chit (chit value $c_T \assign 1$) for $T$.
The above process will yield a labeling of the DAG with a chit value and associated confidence for each transaction~$T$.

\begin{figure}
\begin{center}
    %\tronly{\input{figures/dag-cd}}{\input{figures/dag-cd-compact}}
    \input{figures/dag-cd-compact}
    \captionof{figure}{Example of $\langle \textrm{chit}, \textrm{confidence}\rangle$ values.  Darker boxes indicate transactions with higher confidence values. At most one transaction in each shaded region will be accepted.}
    \label{fig:dag-cd}
\end{center}
\end{figure}

Figure~\ref{fig:dag-cd} illustrates a sample DAG built by Avalanche.
Similar to Snowball, sampling in Avalanche will create a positive feedback for the preference of a single transaction in its conflict set.
For example, because $T_2$ has larger confidence than $T_3$, its descendants are more likely collect chits in the future compared to $T_3$.

\tronly{Similar to Bitcoin, Avalanche leaves determining the acceptance point of a transaction to the application. An application supplies an \textsc{isAccepted} predicate that can take into account the value at risk in the transaction and the chances of a decision being reverted to determine when to decide.}{}

Committing a transaction can be performed through a \emph{safe early commitment}. For virtuous transactions, $T$ is accepted when it is the only transaction in its conflict set and has a confidence not less than threshold $\beta_1$.
As in Snowball, $T$ can also be accepted after a $\beta_2$ number of consecutive successful queries.
If a virtuous transaction fails to get accepted due to a problem
with parents, it could be accepted if reissued with different parents.
Figure~\ref{fig:gossipchain-ongen} shows how Avalanche entangles transactions. Because transactions that consume and generate the same UTXO do not conflict with each other, any transaction can be reissued with different parents.

\tronly{
Figure~\ref{fig:gossipchain-main} illustrates the protocol main loop
executed by each node.
In each iteration, the node attempts to select a transaction $T$ that has not
yet been queried.  If no such
transaction exists, the loop will stall until a new transaction is
added to $\mathcal{T}$.
It then selects $k$ peers and queries those peers.
If more than $\alpha$ of those peers return a positive response, the chit value is set to~1.
After that, it updates the preferred transaction of each conflict set
of the transactions in its ancestry.
Next, $T$ is added to the set $\mathcal{Q}$
so it will never be queried again by the node.
The code that selects additional peers if some of the $k$ peers are
unresponsive is omitted for simplicity.

Figure~\ref{fig:gossipchain-onquery} shows what happens when a node
receives a query for transaction $T$ from peer $j$.
First it adds $T$ to $\mathcal{T}$, unless it already has it.
Then it determines if $T$ is currently strongly preferred.
If so, the node returns a positive response to peer $j$.
Otherwise, it returns a negative response.
Notice that in the pseudocode, we assume when a node knows $T$, it also
recursively knows the entire ancestry of $T$. This can be achieved by
postponing the delivery of $T$ until its entire ancestry is recursively
fetched.
In practice, an additional gossip process that disseminates
transactions is used in parallel, but is not shown in pseudocode for simplicity.
}{}

### 2020-11-03 figures/analysis1.tex
add initial source and 'fr' make target

\usetikzlibrary{decorations.pathreplacing}
\begin{tikzpicture}[x=1.12cm, scale=0.8, every node/.append style={transform shape}]
    \draw[line width=0.2ex, line cap=round] (0,0) -- (7,0); %edit here for the axis
    \foreach \x in {0,1,3,5,7} % edit here for the vertical lines
    \draw[line width=0.2ex, line cap=round, shift={(\x,0)},color=black] (0pt,3pt) -- (0pt,-3pt);
    \foreach \x in {0.5,1.5} % edit here for the vertical lines
    \draw[line width=0.15ex, line cap=round, shift={(\x,0)},color=black] (0pt,2pt) -- (0pt,-2pt);
    \foreach \x/\y/\z in {
        0/$s_{c/2}$/a,
        1/$s_{ps}$/b,
        3/$s_{c/2+\Delta}$/c,
        5/$s_c$/d,
        7/$s_n$/e}
    \draw[line width=0.2ex, shift={(\x,0)},color=black] (0pt,0pt) -- (0pt,-3pt) node[below] (\z) {\y};
    \foreach \x/\y/\z in {
        0.5/\small$s_{ps-1}$/f,
        1//h,
        1.5/\small$s_{ps+1}$/g}
    \draw[line width=0.2ex, shift={(\x,0)},color=black] (0pt,0pt) -- (0pt,2pt) node[above] (\z) {\y};

    \begin{scope}[line width=0.15ex]
        %\path[->] (c) edge[out=-90, in=-85, looseness=0.7] node[above] {} (d);
        \path[->] (c) edge[out=-90, in=-90, looseness=0.7] node[above] {$\epsilon$} (b);
        \path[->] (h) edge[out=90, in=90, looseness=2] node[above,xshift=-0.5in] {\small$P(s_{ps} \to s_{ps-1})$} (f);
        \path[->] (h) edge[out=90, in=90, looseness=2] node[above,xshift=+0.4in] {\small$P(s_{ps} \to s_{ps+1})$} (g);
    \end{scope}
    %\draw [thick,decoration={brace,raise=1ex, amplitude=2ex},decorate] (5, 0) -- (7, 0) node[midway,above,yshift=3.5ex] {$b$};
\end{tikzpicture}

### 2020-11-03 figures/analysis2.tex
add initial source and 'fr' make target

--- 2020-11-03 figures/analysis1.tex
@@
 \begin{tikzpicture}[x=1.12cm, scale=0.8, every node/.append style={transform shape}]
-    \draw[line width=0.2ex, line cap=round] (0,0) -- (7,0); %edit here for the axis
-    \foreach \x in {0,1,3,5,7} % edit here for the vertical lines
+    \draw[line width=0.2ex, line cap=round] (0,0) -- (8,0); %edit here for the axis
+    \foreach \x in {0,4,6.5,8} % edit here for the vertical lines
     \draw[line width=0.2ex, line cap=round, shift={(\x,0)},color=black] (0pt,3pt) -- (0pt,-3pt);
-    \foreach \x in {0.5,1.5} % edit here for the vertical lines
-    \draw[line width=0.15ex, line cap=round, shift={(\x,0)},color=black] (0pt,2pt) -- (0pt,-2pt);
-    \foreach \x/\y/\z in {
-        0/$s_{c/2}$/a,
-        1/$s_{ps}$/b,
-        3/$s_{c/2+\Delta}$/c,
-        5/$s_c$/d,
-        7/$s_n$/e}
+    \foreach \x/\y/\z in {%
+        0/$0$/a,
+        4/${c/2}$/b,
+        6.5/$$/c,
+        8/$c$/d}
     \draw[line width=0.2ex, shift={(\x,0)},color=black] (0pt,0pt) -- (0pt,-3pt) node[below] (\z) {\y};
-    \foreach \x/\y/\z in {
-        0.5/\small$s_{ps-1}$/f,
-        1//h,
-        1.5/\small$s_{ps+1}$/g}
-    \draw[line width=0.2ex, shift={(\x,0)},color=black] (0pt,0pt) -- (0pt,2pt) node[above] (\z) {\y};
 
     \begin{scope}[line width=0.15ex]
-        %\path[->] (c) edge[out=-90, in=-85, looseness=0.7] node[above] {} (d);
-        \path[->] (c) edge[out=-90, in=-90, looseness=0.7] node[above] {$\epsilon$} (b);
-        \path[->] (h) edge[out=90, in=90, looseness=2] node[above,xshift=-0.5in] {\small$P(s_{ps} \to s_{ps-1})$} (f);
-        \path[->] (h) edge[out=90, in=90, looseness=2] node[above,xshift=+0.4in] {\small$P(s_{ps} \to s_{ps+1})$} (g);
+        \path[->] (c) edge[out=-90, in=-90, looseness=0.4] node[above] {$\le \varepsilon$} (a);
     \end{scope}
-    %\draw [thick,decoration={brace,raise=1ex, amplitude=2ex},decorate] (5, 0) -- (7, 0) node[midway,above,yshift=3.5ex] {$b$};
+    \draw [thick,decoration={brace,raise=1ex, amplitude=2ex},decorate] (4, 0) -- (6.5, 0) node[midway,above,yshift=3.5ex] {$\delta$};
 \end{tikzpicture}

### 2020-11-03 figures/cash-system-a.tex
add initial source and 'fr' make target

\definecolor{lightgray}{HTML}{dddddd}
\definecolor{medgray}{HTML}{cccccc}
\definecolor{medgray2}{HTML}{bbbbbb}
\definecolor{darkgray}{HTML}{aaaaaa}
\definecolor{lightgp}{HTML}{ddddee}
\begin{tikzpicture}[x=1.12cm, scale=0.7, every node/.append style={transform shape}]
    \begin{scope}[all/.style={draw,fill=white,minimum height=0.5cm, minimum width=0.5cm},line width=0.08ex]

        \node[all, label={[anchor=east]left:$\texttt{TX}_g$},
        rectangle,minimum width=0.5in, minimum height=0.4in, line width=0.1ex] (a0) at (0, 0) {};
        \node[all, label={[anchor=east]left:$\texttt{TX}_a$}, rectangle,minimum width=1in, minimum height=0.55in, line width=0.1ex]
        (a1) at (-1.45, -2) {};
        \node[all, label={[anchor=west]right:$\texttt{TX}_b$}, rectangle,minimum width=1in, minimum height=0.55in, line width=0.1ex]
        (a2) at (1.45, -2) {};
        \node[all, label={[anchor=east]left:$\texttt{TX}_c$}, rectangle,minimum width=1in, minimum height=0.55in, line width=0.1ex]
        (a3) at (-1.45, -4) {};

        \node[fill=lightgp, circle,minimum width=1cm, opacity=0.8] (pt1) at (-2, -2) {};
        \node[fill=lightgp, circle,minimum width=1cm, opacity=0.8] (pt4) at (2, -2) {};
        \node[fill=lightgp, circle,minimum width=1cm, opacity=0.8] (pt5) at (-2, -4) {};
        \node[fill=lightgp, circle,minimum width=1cm, opacity=0.8] (pt6) at (-0.9, -4) {};
        \fill[color=lightgp, opacity=0.8] plot[smooth cycle] coordinates { (1.4, -2) (1.1, -2.4) (-1.1, -2.4) (-1.4, -2) (-1.1, -1.6) (1.1, -1.6)} node (pt2) {};
        %\fill[color=lightgp] plot[smooth cycle] coordinates {(2.1, -1.75) (2, -2.4) (0.6, -3.4) (0, -3.5) (-0.3, -3.3)(-0.3, -2.8) (0.1, -2.4) (1.2, -1.65) (1.7, -1.5) } node (pt3) {};

        %\node[all, draw] (b0) at (0, 0) {$\texttt{In}_a$};
        \node[all, draw] (b1) at (-2, -2) {$\texttt{In}_{a1}$};
        \node[all, draw] (b2) at (-0.9, -2) {$\texttt{In}_{a2}$};
        \node[all, draw] (b3) at (0.9, -2) {$\texttt{In}_{b1}$};
        \node[all, draw] (b4) at (2, -2) {$\texttt{In}_{b2}$};

        \node[all, draw] (b5) at (-2, -4) {$\texttt{In}_{c1}$};
        \node[all, draw] (b6) at (-0.9, -4) {$\texttt{In}_{c2}$};

        \begin{scope}[line width=0.15ex]
        \path[->] (a1) edge[out=90, in=-110] node[sloped,above] {} (a0) ;
        \path[->] (a2) edge[out=100, in=-70] node[sloped,above] {} (a0) ;
        \path[->] (a3) edge[out=100, in=-90] node[sloped,above] {} (a1) ;
        \end{scope}
        %\node[] (p1) at (3, 0.5) {\footnotesize$\mathcal{P}_{T_1}$};
        %\node[] (p2) at (3, 0) {\footnotesize$\mathcal{P}_{T_2} = \mathcal{P}_{T_3}$};
        %\node[] (p3) at (3, -0.5) {\footnotesize$\mathcal{P}_{T_9} = \mathcal{P}_{T_6} = \mathcal{P}_{T_7}$};
        \begin{scope}[color=darkgray,dashed,line width=0.2ex]
        %\path[->] (p1) edge[out=180, in=0] node[sloped,above] {} (pt1) ;
        %\path[->] (p2) edge[out=180, in=60] node[sloped,above] {} (pt2) ;
        %\path[->] (p3) edge[out=-90, in=30] node[sloped,above] {} (pt3) ;
        \end{scope}
    \end{scope}
\end{tikzpicture}

### 2020-11-03 figures/cash-system-b.tex
add initial source and 'fr' make target

--- 2020-11-03 figures/cash-system-a.tex
@@
 \definecolor{medgray2}{HTML}{bbbbbb}
-\definecolor{darkgray}{HTML}{aaaaaa}
+\definecolor{darkgray}{HTML}{999999}
 \definecolor{lightgp}{HTML}{ddddee}
@@
     \begin{scope}[all/.style={draw,fill=white,minimum height=0.5cm, minimum width=0.5cm},line width=0.08ex]
-
-        \node[all, label={[anchor=east]left:$\texttt{TX}_g$},
-        rectangle,minimum width=0.5in, minimum height=0.4in, line width=0.1ex] (a0) at (0, 0) {};
-        \node[all, label={[anchor=east]left:$\texttt{TX}_a$}, rectangle,minimum width=1in, minimum height=0.55in, line width=0.1ex]
+    \begin{scope}[color=darkgray,dashed,line width=0.2ex]
+        \node[all, label={[anchor=east, font=\fontsize{14}{14}]left:$\texttt{TX}_g$},
+        rectangle,minimum width=0.5in, minimum height=0.4in, line width=0.15ex] (a0) at (0, 0) {};
+        \node[all, label={[anchor=east, font=\fontsize{14}{14}]left:$\texttt{TX}_a$}, rectangle,minimum width=1in, minimum height=0.55in, line width=0.15ex]
         (a1) at (-1.45, -2) {};
-        \node[all, label={[anchor=west]right:$\texttt{TX}_b$}, rectangle,minimum width=1in, minimum height=0.55in, line width=0.1ex]
+        \node[all, label={[anchor=west, font=\fontsize{14}{14}]right:$\texttt{TX}_b$}, rectangle,minimum width=1in, minimum height=0.55in, line width=0.15ex]
         (a2) at (1.45, -2) {};
-        \node[all, label={[anchor=east]left:$\texttt{TX}_c$}, rectangle,minimum width=1in, minimum height=0.55in, line width=0.1ex]
+        \node[all, label={[anchor=east, font=\fontsize{14}{14}]left:$\texttt{TX}_c$}, rectangle,minimum width=1in, minimum height=0.55in, line width=0.15ex]
         (a3) at (-1.45, -4) {};
+    \end{scope}
 
@@
         %\fill[color=lightgp] plot[smooth cycle] coordinates {(2.1, -1.75) (2, -2.4) (0.6, -3.4) (0, -3.5) (-0.3, -3.3)(-0.3, -2.8) (0.1, -2.4) (1.2, -1.65) (1.7, -1.5) } node (pt3) {};
+
+        \node[rectangle, fill=lightgray, draw, minimum width=1ex](ain) at (-1.45, -1.5) {};
+        \node[rectangle, fill=lightgray, draw, minimum width=1ex](bin) at (1.45, -1.5) {};
+        \node[rectangle, fill=lightgray, draw, minimum width=1ex](aout) at (-1.45, -2.5) {};
+        \node[rectangle, fill=lightgray, draw, minimum width=1ex](cin) at (-1.45, -3.5) {};
+        \node[rectangle, fill=lightgray, draw, minimum width=1ex](gout) at (0, -0.3) {};
 
@@
         \begin{scope}[line width=0.15ex]
-        \path[->] (a1) edge[out=90, in=-110] node[sloped,above] {} (a0) ;
-        \path[->] (a2) edge[out=100, in=-70] node[sloped,above] {} (a0) ;
-        \path[->] (a3) edge[out=100, in=-90] node[sloped,above] {} (a1) ;
+        \path[->] (ain) edge[out=90, in=-110] node[sloped,above] {} (gout) ;
+        \path[->] (bin) edge[out=100, in=-70] node[sloped,above] {} (gout) ;
+        \path[->] (cin) edge[out=100, in=-90] node[sloped,above] {} (aout) ;
+        \path[->] (b1) edge[out=100, in=180] node[sloped,above] {} (ain);
+        \path[->] (b2) edge[out=100, in=0] node[sloped,above] {} (ain);
+        \path[->] (b3) edge[out=100, in=180] node[sloped,above] {} (bin);
+        \path[->] (b4) edge[out=100, in=0] node[sloped,above] {} (bin);
+        \path[->] (b5) edge[out=100, in=180] node[sloped,above] {} (cin);
+        \path[->] (b6) edge[out=100, in=0] node[sloped,above] {} (cin);
+        \path[->] (aout) edge[out=180, in=-90, looseness=1.3] node[sloped,above] {} (b1);
+        \path[->] (aout) edge[out=0, in=-90, looseness=1.3] node[sloped,above] {} (b2);
         \end{scope}

### 2020-11-03 figures/dag-cd-compact.tex
add initial source and 'fr' make target

\definecolor{lightgray}{HTML}{dddddd}
\definecolor{medgray}{HTML}{cccccc}
\definecolor{medgray2}{HTML}{bbbbbb}
\definecolor{darkgray}{HTML}{aaaaaa}
\definecolor{lightgp}{HTML}{ddddee}
\begin{tikzpicture}[x=1.12cm, scale=0.9, every node/.append style={transform shape}]
    \begin{scope}[all/.style={draw, minimum height=0.5cm, minimum width=0.5cm},line width=0.08ex]
        \node[fill=lightgp, circle,minimum width=1cm] (pt1) at (0, 0) {};
        \node[fill=lightgp, circle,minimum width=1cm] (pt4) at (-1.5, -2) {};
        \node[fill=lightgp, circle,minimum width=1cm] (pt5) at (-0.5, -2) {};
        \node[fill=lightgp, circle,minimum width=1cm] (pt6) at (-2, -3) {};
        \fill[color=lightgp] plot[smooth cycle] coordinates { (1.2, -1) (0.9, -1.4) (-0.9, -1.4) (-1.2, -1) (-0.9, -0.6) (0.9, -0.6)} node (pt2) {};
        \fill[color=lightgp] plot[smooth cycle] coordinates {(2.1, -1.75) (2, -2.4) (0.6, -3.4) (0, -3.5) (-0.3, -3.3)(-0.3, -2.8) (0.1, -2.4) (1.2, -1.65) (1.7, -1.5) } node (pt3) {};
        \node[all, draw,fill=darkgray] (b0) at (0, 0) {$T_1$};
        \node[all, draw,fill=darkgray] (b1) at (-0.8, -1) {$T_2$};
        \node[all, draw,fill=white] (b2) at (0.8, -1) {$T_3$};
        \node[all, draw,fill=medgray] (b3) at (-1.5, -2) {$T_4$};
        \node[all, draw,fill=medgray2] (b4) at (-0.5, -2) {$T_5$};
        \node[all, draw,fill=white] (b5) at (0.9, -2.5) {$T_6$};
        \node[all, draw,fill=white] (b6) at (1.7, -2) {$T_7$};
        \node[all, draw,fill=lightgray] (b7) at (-2, -3) {$T_8$};
        \node[all, draw,fill=lightgray] (b8) at (0.2, -3) {$T_9$};
        \begin{scope}[line width=0.2ex]
        \path[->] (b1) edge[out=90, in=-110] node[sloped,above] {} (b0) ;
        \path[->] (b2) edge[out=100, in=-70] node[sloped,above] {} (b0) ;
        \path[->] (b3) edge[out=90, in=-110] node[sloped,above] {} (b1) ;
        \path[->] (b4) edge[out=90, in=-70] node[sloped,above] {} (b1) ;
        \path[->] (b5) edge[out=100, in=-110] node[sloped,above] {} (b2) ;
        \path[->] (b6) edge[out=120, in=-70] node[sloped,above] {} (b2) ;
        \path[->] (b7) edge[out=90, in=-90] node[sloped,above] {} (b3) ;
        \path[->] (b7) edge[out=0, in=-110] node[sloped,above] {} (b4) ;
        \path[->] (b8) edge[out=90, in=-70] node[sloped,above] {} (b4) ;
        \end{scope}
        \node[] (p1) at (3, 0.5) {\footnotesize$\mathcal{P}_{T_1}$} ;
        \node[] (p2) at (3, 0) {\footnotesize$\mathcal{P}_{T_2} = \mathcal{P}_{T_3}$};
        \node[] (p3) at (3, -0.5) {\footnotesize$\mathcal{P}_{T_9} = \mathcal{P}_{T_6} = \mathcal{P}_{T_7}$};
        \begin{scope}[color=darkgray,dashed,line width=0.2ex]
        \path[->] (p1) edge[out=180, in=0] node[sloped,above] {} (pt1) ;
        \path[->] (p2) edge[out=180, in=60] node[sloped,above] {} (pt2) ;
        \path[->] (p3) edge[out=-90, in=30] node[sloped,above] {} (pt3) ;
        \end{scope}

        \node[anchor=south, yshift=0.1cm] at (b0.north) {\footnotesize$\langle \mathbf{c_{T_1}}, d(T_1) \rangle = \langle\mathbf{1}, 6\rangle$};
        \node[anchor=east,xshift=-0.1cm] at (b1.west) {\footnotesize$\langle \mathbf{1}, 5 \rangle$};
        \node[anchor=west,xshift=0.1cm] at (b2.east) {\footnotesize$\langle \mathbf{0}, 0 \rangle$};
        \node[anchor=east,xshift=-0.1cm] at (b3.west) {\footnotesize$\langle \mathbf{1}, 2 \rangle$};
        \node[anchor=west,xshift=0.1cm] at (b4.east) {\footnotesize$\langle \mathbf{1}, 3 \rangle$};
        \node[anchor=north west,xshift=0.1cm] at (b5.south east) {\footnotesize$\langle \mathbf{0}, 0 \rangle$};
        \node[anchor=north west] at (b6.south east) {\footnotesize$\langle \mathbf{0}, 0 \rangle$};
        \node[anchor=north west] at (b8.south east) {\footnotesize$\langle \mathbf{1}, 1 \rangle$};
        \node[anchor=north,yshift=-0.2cm] at (b7.south) {\footnotesize$\langle \mathbf{1}, 1 \rangle$};

    \end{scope}
\end{tikzpicture}

### 2020-11-03 figures/dag-conflict-set.tex
add initial source and 'fr' make target

--- 2020-11-03 figures/dag-cd-compact.tex
@@
 \definecolor{lightgp}{HTML}{ddddee}
-\begin{tikzpicture}[x=1.12cm, scale=0.9, every node/.append style={transform shape}]
-    \begin{scope}[all/.style={draw, minimum height=0.5cm, minimum width=0.5cm},line width=0.08ex]
+\begin{tikzpicture}[x=1.12cm]
+    \begin{scope}[all/.style={draw,fill=white,minimum height=0.5cm, minimum width=0.5cm},line width=0.08ex]
         \node[fill=lightgp, circle,minimum width=1cm] (pt1) at (0, 0) {};
@@
         \fill[color=lightgp] plot[smooth cycle] coordinates {(2.1, -1.75) (2, -2.4) (0.6, -3.4) (0, -3.5) (-0.3, -3.3)(-0.3, -2.8) (0.1, -2.4) (1.2, -1.65) (1.7, -1.5) } node (pt3) {};
-        \node[all, draw,fill=darkgray] (b0) at (0, 0) {$T_1$};
-        \node[all, draw,fill=darkgray] (b1) at (-0.8, -1) {$T_2$};
-        \node[all, draw,fill=white] (b2) at (0.8, -1) {$T_3$};
-        \node[all, draw,fill=medgray] (b3) at (-1.5, -2) {$T_4$};
-        \node[all, draw,fill=medgray2] (b4) at (-0.5, -2) {$T_5$};
-        \node[all, draw,fill=white] (b5) at (0.9, -2.5) {$T_6$};
-        \node[all, draw,fill=white] (b6) at (1.7, -2) {$T_7$};
-        \node[all, draw,fill=lightgray] (b7) at (-2, -3) {$T_8$};
-        \node[all, draw,fill=lightgray] (b8) at (0.2, -3) {$T_9$};
+        \node[all, draw] (b0) at (0, 0) {$T_1$};
+        \node[all, draw] (b1) at (-0.8, -1) {$T_2$};
+        \node[all, draw] (b2) at (0.8, -1) {$T_3$};
+        \node[all, draw] (b3) at (-1.5, -2) {$T_4$};
+        \node[all, draw] (b4) at (-0.5, -2) {$T_5$};
+        \node[all, draw] (b5) at (0.9, -2.5) {$T_6$};
+        \node[all, draw] (b6) at (1.7, -2) {$T_7$};
+        \node[all, draw] (b7) at (-2, -3) {$T_8$};
+        \node[all, draw] (b8) at (0.2, -3) {$T_9$};
         \begin{scope}[line width=0.2ex]
@@
         \end{scope}
-        \node[] (p1) at (3, 0.5) {\footnotesize$\mathcal{P}_{T_1}$} ;
+        \node[] (p1) at (3, 0.5) {\footnotesize$\mathcal{P}_{T_1}$};
         \node[] (p2) at (3, 0) {\footnotesize$\mathcal{P}_{T_2} = \mathcal{P}_{T_3}$};
@@
         \end{scope}
-
-        \node[anchor=south, yshift=0.1cm] at (b0.north) {\footnotesize$\langle \mathbf{c_{T_1}}, d(T_1) \rangle = \langle\mathbf{1}, 6\rangle$};
-        \node[anchor=east,xshift=-0.1cm] at (b1.west) {\footnotesize$\langle \mathbf{1}, 5 \rangle$};
-        \node[anchor=west,xshift=0.1cm] at (b2.east) {\footnotesize$\langle \mathbf{0}, 0 \rangle$};
-        \node[anchor=east,xshift=-0.1cm] at (b3.west) {\footnotesize$\langle \mathbf{1}, 2 \rangle$};
-        \node[anchor=west,xshift=0.1cm] at (b4.east) {\footnotesize$\langle \mathbf{1}, 3 \rangle$};
-        \node[anchor=north west,xshift=0.1cm] at (b5.south east) {\footnotesize$\langle \mathbf{0}, 0 \rangle$};
-        \node[anchor=north west] at (b6.south east) {\footnotesize$\langle \mathbf{0}, 0 \rangle$};
-        \node[anchor=north west] at (b8.south east) {\footnotesize$\langle \mathbf{1}, 1 \rangle$};
-        \node[anchor=north,yshift=-0.2cm] at (b7.south) {\footnotesize$\langle \mathbf{1}, 1 \rangle$};
-
     \end{scope}

### 2020-11-03 figures/dag.tex
add initial source and 'fr' make target

--- 2020-11-03 figures/dag-cd-compact.tex
@@
 \definecolor{lightgp}{HTML}{ddddee}
-\begin{tikzpicture}[x=1.12cm, scale=0.9, every node/.append style={transform shape}]
+\begin{tikzpicture}[x=1.12cm]
     \begin{scope}[all/.style={draw, minimum height=0.5cm, minimum width=0.5cm},line width=0.08ex]
@@
         \end{scope}
-        \node[] (p1) at (3, 0.5) {\footnotesize$\mathcal{P}_{T_1}$} ;
-        \node[] (p2) at (3, 0) {\footnotesize$\mathcal{P}_{T_2} = \mathcal{P}_{T_3}$};
-        \node[] (p3) at (3, -0.5) {\footnotesize$\mathcal{P}_{T_9} = \mathcal{P}_{T_6} = \mathcal{P}_{T_7}$};
+        \node[anchor=south, yshift=0.1cm] at (b0.north) {\footnotesize$(c, d) = (\textbf{1}, 5)$};
+        \node[anchor=east,xshift=-0.1cm] at (b1.west) {\footnotesize$(\textbf{1}, 5)$};
+        \node[anchor=west,xshift=0.1cm] at (b2.east) {\footnotesize$(\textbf{0}, 0)$};
+        \node[anchor=east,xshift=-0.1cm] at (b3.west) {\footnotesize$(\textbf{1}, 2)$};
+        \node[anchor=west,xshift=0.1cm] at (b4.east) {\footnotesize$(\textbf{1}, 3)$};
+        \node[anchor=north west,xshift=0.1cm] at (b5.south east) {\footnotesize$(\textbf{0}, 0)$};
+        \node[anchor=north west] at (b6.south east) {\footnotesize$(\textbf{0}, 0)$};
+        \node[anchor=north west] at (b8.south east) {\footnotesize$(\textbf{1}, 1)$};
+        \node[anchor=north,yshift=-0.2cm] at (b7.south) {\footnotesize$(\textbf{1}, 1)$};
+        \node[] (p1) at (3, 0.5) {\footnotesize$\textrm{Parition 1}$};
+        \node[] (p2) at (3, 0) {\footnotesize$\textrm{Parition 2}$};
+        \node[] (p3) at (3, -0.5) {\footnotesize$\textrm{Parition 3}$};
+        \node[] (p4) at (3, -1) {\footnotesize$\textrm{Parition 4}$};
+        \node[] (p5) at (3, -1.5) {\footnotesize$\textrm{Parition 5}$};
+        \node[] (p6) at (3, -2) {\footnotesize$\textrm{Parition 6}$};
         \begin{scope}[color=darkgray,dashed,line width=0.2ex]
@@
         \path[->] (p2) edge[out=180, in=60] node[sloped,above] {} (pt2) ;
-        \path[->] (p3) edge[out=-90, in=30] node[sloped,above] {} (pt3) ;
+        \path[->] (p3) edge[out=180, in=30] node[sloped,above] {} (pt3) ;
+        \path[->] (p4) edge[out=170, in=30] node[sloped,above] {} (pt4) ;
+        \path[->] (p5) edge[out=160, in=20] node[sloped,above] {} (pt5) ;
+        \path[->] (p6) edge[out=270, in=-30] node[sloped,above] {} (pt6) ;
         \end{scope}
-
-        \node[anchor=south, yshift=0.1cm] at (b0.north) {\footnotesize$\langle \mathbf{c_{T_1}}, d(T_1) \rangle = \langle\mathbf{1}, 6\rangle$};
-        \node[anchor=east,xshift=-0.1cm] at (b1.west) {\footnotesize$\langle \mathbf{1}, 5 \rangle$};
-        \node[anchor=west,xshift=0.1cm] at (b2.east) {\footnotesize$\langle \mathbf{0}, 0 \rangle$};
-        \node[anchor=east,xshift=-0.1cm] at (b3.west) {\footnotesize$\langle \mathbf{1}, 2 \rangle$};
-        \node[anchor=west,xshift=0.1cm] at (b4.east) {\footnotesize$\langle \mathbf{1}, 3 \rangle$};
-        \node[anchor=north west,xshift=0.1cm] at (b5.south east) {\footnotesize$\langle \mathbf{0}, 0 \rangle$};
-        \node[anchor=north west] at (b6.south east) {\footnotesize$\langle \mathbf{0}, 0 \rangle$};
-        \node[anchor=north west] at (b8.south east) {\footnotesize$\langle \mathbf{1}, 1 \rangle$};
-        \node[anchor=north,yshift=-0.2cm] at (b7.south) {\footnotesize$\langle \mathbf{1}, 1 \rangle$};
-
     \end{scope}

### 2020-11-03 figures/dag2.tex
add initial source and 'fr' make target

--- 2020-11-03 figures/dag.tex
@@
         \node[anchor=north,yshift=-0.2cm] at (b7.south) {\footnotesize$(\textbf{1}, 1)$};
-        \node[] (p1) at (3, 0.5) {\footnotesize$\textrm{Parition 1}$};
-        \node[] (p2) at (3, 0) {\footnotesize$\textrm{Parition 2}$};
-        \node[] (p3) at (3, -0.5) {\footnotesize$\textrm{Parition 3}$};
-        \node[] (p4) at (3, -1) {\footnotesize$\textrm{Parition 4}$};
-        \node[] (p5) at (3, -1.5) {\footnotesize$\textrm{Parition 5}$};
-        \node[] (p6) at (3, -2) {\footnotesize$\textrm{Parition 6}$};
-        \begin{scope}[color=darkgray,dashed,line width=0.2ex]
-        \path[->] (p1) edge[out=180, in=0] node[sloped,above] {} (pt1) ;
-        \path[->] (p2) edge[out=180, in=60] node[sloped,above] {} (pt2) ;
-        \path[->] (p3) edge[out=180, in=30] node[sloped,above] {} (pt3) ;
-        \path[->] (p4) edge[out=170, in=30] node[sloped,above] {} (pt4) ;
-        \path[->] (p5) edge[out=160, in=20] node[sloped,above] {} (pt5) ;
-        \path[->] (p6) edge[out=270, in=-30] node[sloped,above] {} (pt6) ;
-        \end{scope}
+
     \end{scope}

### 2020-11-03 obsolete/extra.tex
add initial source and 'fr' make target

\section{Core Snowball Network Overlay Protocol}
In the discussion of the Snowball protocol we initially assumed a logically fully-connected network of nodes. 
Later, we relaxed that assumption and showed how the network view set intersection between correct nodes can have disjoint subsets and still maintain safety and liveness.

In this section, we introduce the concrete details of the Snowball network overlay protocol, \textit{SNO}, whose function is to allow the network nodes to maintain a highly-up-to-date view of all active nodes in the system, i.e. allow for churn.
Beyond that, SNO also provides other useful features, including partial detection and prevention of Byzantine behavior, as well as network participation incentivization, which becomes important in the instantiation of the Snowball-based payment system \textit{SnoCoin}.

\subsection{Background}
SNO makes usage of the core ideas of the Fireflies\cite{johansen2006fireflies} network overlay protocol, but modifies certain operational details. 
The Fireflies protocol forms a ring-based pseudorandom mesh from a network of $\mathcal{N}$ nodes, such that each node has a total of $\mathcal{O}(\log{|\mathcal{N}|})$ neighbors, and such that even if up to half of the neighbors are Byzantine, each correct node will maintain a highly up-to-date view of the entire set of active nodes.

As before, we still assume for simplicity a logically fully connected network of nodes $\mathcal{N}$.
Each correct node, however, will only need to maintain $\mathcal{O}(\log{|\mathcal{N}|})$ network links.
Let each node $u \in \mathcal{N}$ have some unique identifier, represented by $u$.id, and let \texttt{H} be a secure cryptographic hash function that is pre-image and collision resistant. 
We generate a set of $\log |\mathcal{N}|$ ``rings'', where each ring -- indexed by $i$ such that $1 \leq i \leq 2*|\mathcal{B}| + 1$ -- is an ordering of the nodes based on the output of \texttt{H}($u$.id $||$ $i$), i.e. 
\[
    u.\text{\texttt{pos}}[i] = \text{\texttt{H}}(u.\text{\texttt{id}} || i); \forall u \in \mathcal{N}, 1 \leq i \leq 2|\mathcal{B}| + 1
\]
Nodes have two neighbors (two closest ordered nodes) on each ring, and all nodes can publicly verify the neighbors of other nodes. The first and last nodes wrap around and are always neighbors with each other on every ring.

The Fireflies overlay guarantees, with high probability, that there exists a path between any two nodes and that this path is pseudorandom. 
Nodes periodically ping other nodes using the pseudorandom mesh generated by the rings structure, and based on the responses can accuse other nodes of failure. 
If a node is accused falsely, then the node may provide evidence of liveness. 
Using this mechanism, each node will maintain a reasonably up-to-date view of all the active nodes in the network and consequently weed out any failed nodes.  
We point the reader to the Fireflies paper for formal proofs.

\subsection{Snowball Network Overlay}
We use the network overlay construction in the Fireflies protocol, but modify it to act as the carrier of messages in the network, rather than just ping nodes periodically.
Every node $u$ chooses the set of $k$ random solicited nodes exactly as before (i.e. from the set of all nodes $\mathcal{N}$). 
This means that the semantics and interface of the Snowball core protocol remain exactly unmodified. 
However, rather than communicate with the solicited nodes directly, the solicitor queries along one of the random paths induced by the random communication mesh.

For example, suppose that node $u$ chooses a random sample $\mathcal{R} \subset \mathcal{N}_u$ of size $k$ of nodes to solicit, and suppose that the shortest path to node $v \in \mathcal{R}$ is through node $g \in \mathcal{N}_u$. 
$u$ then transmits the message first to $g$, which will queue and forward it to $v$.

On top of merely forwarding messages, $u$ and $g$ will keep track of the response by $v$. Untimely responses by $v$ will raise accusations by both $u$ and $g$. Enough accusations will result in all correct nodes removing $v$ from their respective network view.

{\color{blue} Kevin: Expand the subsection here with a more formal protocol description. It should basically be the same as the Fireflies protocol. }

\section{Instantiation}
We now instantiate a payment system with Snowball, called SnoCoin. Before discussing the technical details, we discuss challenges and introduce the solutions adopted in this paper.

\subsection{Sybils}
\label{subsection:sybil}
Any open and unprotected network can trivially become a victim to a Sybil attack~\cite{douceur2002sybil}, wherein an adversary can cheaply generate a large number of identities and take down the network. 
In Bitcoin, Sybils are deterred through the protocol-level usage of cryptographic puzzles, whose mathematical properties ensure that participation in decision making in the network is costly. 
Since Snowball does not make usage of proofs-of-work, alternative measures must be adopted. 
There are several techniques that may be used. 
Fundamentally, these techniques rely on trading one resource for authorization to participate in the system, but there are big practical differences between the actual resource chosen.
\begin{itemize}
\item IPv4 Scarcity: Many existing peer-to-peer systems, including Bittorrent and Bitcoin, employ Sybil deterrence at their network layer based on the scarcity of IPv4 addresses. 
Specifically, they observe that the IPv4 address space is densely allocated, and that it is difficult for an adversary to amass large numbers of addresses across the Internet.
Bitcoin's main mechanism against eclipse attacks coupled with low-difficulty chain construction is to ensure that a node peers with a diverse set of peers from multiple IP prefixes. 
One possibility, inspired by this mechanism, is to limit the number of participants per prefix. 
Coupled with an IANA-provided map of IP prefix allocations, this may help limit the number of Sybil nodes that an attacker, who has access to a limited number of machines and networks, can bring into the system.
\item Proof-of-Bond: A bond, denominated in the Snowball currency itself to eliminate any dependency on external resources, may be used to raise the costs to launch an attack. 
In this scheme, entry into the system requires the nodes to put up a bond for the duration of their stay in the system. 
This ensures that an attacker has to tie up significant amounts of cash to mount an attack.
Note that this is quite different from traditional proof-of-stake protocols in that it does not suffer from nothing-at-stake, the participants cannot retroactively rewrite history using stale keys, and the bond is not at risk of being usurped by the system.
\item Proof-of-Uniqueness: It is possible to require participants in Snowball to provide a proof-of-uniqueness provided by a secure hardware mechanism. 
One possibility is for every node to run a small, predetermined program in an SGX-like enclave and obtain a certificate that they executed the code on a genuine secure processor. 
This program can then issue just a single random number, derived from the on-chip secret.
\item External Credentials: In domains where the participants are issued a unique credential by an external entity, this credential can be used to authorize the node's entry into the system.
For instance, in settings where participants are issued unique certificates, they may join the network with a join transaction that takes the credential as an input. 
The network can then ensure that any given credential is used by at most one node.

\end{itemize}

SnoCoin adopts a combination of the first two measures to deter participants. We envision SNO to provide a platform for global remittance with the least amount of friction to participation. We envision potentially hundreds of thousands to millions of Internet-connected low-powered devices (such as mobile devices), transacting at peak capacity. Plainly speaking, a combination of the two provides the least amount of resistance for individual honest clients, but provides a reasonable deterrence to an adversary willing to spawn off millions of clusters around the world.

% \subsection{View and Identity Management}
% {\color{blue}{Kevin: This section is of crucial importance, and needs to be precisely described.}}

% As discussed previously, an inconsistent view amongst correct nodes increases the ability of Byzantine nodes to affect both liveness and safety of the system. 
% Several mechanisms can be used to ensure a highly consistent view, but in SNO we make usage of a two-phase mechanism. 
% In the first phase, we use the proof-of-bond for providing membership into the network. 
% All nodes can authenticate the membership of all other nodes by simply verifying that the bond has been posted. 
% In the second phase, we make usage of Fireflies\cite{johansen2006fireflies}, an intrusion-tolerant network overlay protocol, which provides correct nodes a reasonably fresh view of the set of active nodes in the network, whilst tolerating a Byzantine presence with up to half of all active nodes.
% Constructing such a network overlay provides several benefits. 
% First, it allows correct nodes to efficiently prune out inactive nodes, increasing efficiency of polling in the Snowball protocol. 
% Second, it provides a clean means of rewarding good behavior. 
% Correct nodes that participate and stay active in the network may request rewards in the system in the form of coin minting in SNO.

\subsection{Failure Detection}

\subsection{Network Congestion Control}
In our analysis, we assumed a Byzantine adversary with powers that, in reality, are extremely difficult to achieve. A large body of work on network congestion control is highly applicable to any instantiation of Snowball. Several relevant techniques are:
\paragraph{Randomized backoff} Whenever a correct node is queried, the node can randomly choose a wait time before responding to the querying node. Although this simple technique makes no distinction between nodes, it is still extremely efficient in reducing the ability of Byzantine nodes to mount attacks. 
\paragraph{Targetted backoff} Additionally, a correct node may keep track of the frequency of requests for each neighbor, and respond to requests accordingly.

% \subsection{Network View}
% \label{subsection:networkview}
% As in any active open network, it is hard to guarantee that the network view is consistent amongst all correct nodes. 
% In Snowball, we make usage of intrusion tolerant network overlays in order to

\section{Extra Text}

The advantage of linking the transactions together rather than running
repetitive voting process for each isolated one is two-fold. On the one hand,
the proper policy will create a positive feedback that once a transaction is
supported by most nodes out of its conflict sets, it will become increasingly
likely to be supported in the future.

On the other hand, it makes the system scalable regarding the total number of
transactions, because for each node in the network, every transaction only need
to be voted once.  When the node asks for votes for a newly discovered
transaction, the voting result transitively contributes to all the ancestor
transactions on the DAG\@.  Without the DAG, conflict sets are free-floating with
their transactions, so one has to keep polling for the votes for transactions
in all conflict sets, even the confidence values of oldest ones are rocksolid to
reach a consensus. The reason we have to repeat voting in absence of a DAG is
for a practical system, especially an electronic cash system, any node
should be able to recover from incorrect temporary state seamlessly
to achieve an eventual consistency. This means the consensus protocol is
theoretically always running, but the probabilistic safety guarantee gets
improved over time, so the decision threshold is just an observation from the
current process of consensus and guides the business decisions (e.g.\ whether
to render the service or deliver the goods).  Bitcoin uses the longest chain
rule to represent the current state (decisions) of the transactions, and when a
block is appended to the chain, it enhances the belief on all the preceding
blocks.

Gossip-based protocols are usually designed to work in settings where the nodes
are benign. However, in our system, where Byzantine behavior could present,
some extra care is taken to improve the security given the attackers could push
around incorrect information and disrupt decision making.

The core protocol is mainly pull-based. That is, the only ``pushed''
information is the transaction data, which do not affect the decision process.
Transaction data are neutral, and transactions themselves do not play a role in
making decisions until they are polled. A node only runs the protocol based on
the local information gathered from voting summary sampled at random and will
never need to trust others. The only external information other than
transaction data is the polling result from the sampled nodes. But because the
sampling is controlled by the poller, the attackers could not affect the voting
result if they are not asked, nor could they force the set of nodes to be
polled. There is no exchange of voting result or confidence value between the
nodes. The belief of the chosen transactions are autonomously built by active
sampling and polling, on behalf of the node itself.

2.

The extensive past work on Byzantine consensus can be divided into two families: permissioned and permissionless consensus protocols.

\paragraph{Permissioned BFT Consensus}
The seminal protocol PBFT~\cite{castro1999practical} protocol was the first Byzantine consensus protocol to guarantee low transaction latency and safety for a network of $3|\mathcal{B}|+1$ nodes.  
PBFT entails quadratic $\mathcal{O}(|\mathcal{N}|^2)$ message complexity and its guarantees break down non-gracefully with bound violation. 
Abd-El-Malek et al.~\cite{abd2005fault} and Cowling et al.~\cite{cowling2006hq} use a quorum-based approach to optimize for contention-free cases of operation to achieve consensus in only a single round of communication. 
However, although these protocols improve on performance, they degrade very poorly under contention. 
Zyzzyva~\cite{kotla2007zyzzyva} couples BFT with speculative execution to improve the failure-free operation case. 
Guerraoui et al.~\cite{guerraoui2010next} introduce a protocol with optimized performance under several -- rather than just one -- cases of execution. 
In contrast, Clement et al.~\cite{clement2009making} sacrificed some performance to tolerate worst-case degradation, providing a more uniform execution profile. 
This work, in particular, sacrificed failure-free optimizations to provide consistent throughput even at high number of failures.

Other past work in permissioned BFT systems typically requires at least $2|\mathcal{B}|+1$ replicas tolerance. 
However, work in 2012 by Kapitza et al.~\cite{kapitza2012cheapbft} showed how to leverage trusted hardware components to construct a protocol that uses $|\mathcal{B}|+1$ replicas.

On the other end of the spectrum, other work attempts to introduce new protocols under redefinitions and relaxations of the BFT model. 
Rodrigues et al.~\cite{rodrigues2007large} modify PBFT to allow for arbitrary choice of number of replicas and failure threshold, providing a probabilistic guarantee of liveness for some failure ratio but protecting safety with high probability. 
Such relaxation of the BFT protocol is a core design principle of Snowball, where we postulate that for a payment system such as SnoCoin, safety is far more important than liveness. 
In another form of relaxation, the work of Singh et al.~\cite{singh2009zeno} introduced a BFT state machine replication protocol that trades consistency for high availability. 
More specifically, this paper guarantees eventual consistency rather than linearizability, meaning that participants can be inconsistent but eventually agree once the network stabilizes. 
By providing an even weaker consistency guarantee, namely fork-join-causal consistency, in 2011 Mahajan et al.\cite{mahajan2011depot} described a protocol that guarantees safety under $|\mathcal{B}|+1$ replicas.

\paragraph{Permissionless BFT Consensus}
Bitcoin~\cite{nakamoto2008bitcoin} was the first protocol to show the possibility of consensus under no predefined set of identities. The core insight is the usage of proof-of-work (PoW) to replace voting by identities with voting by computational power. While techniques based on proof-of-work~\cite{dwork1992pricing,aspnes2005exposing}, and even cryptocurrencies with minting based on proof-of-work~\cite{vishnumurthy2003karma,rivest1997payword}, had been explored before, Bitcoin was the first to incorporate PoW into its consensus process.

Bitcoin, however, has at least two major weaknesses: performance and scale. 
Bitcoin incurs high confirmation latency at around 10 minutes, is hard-capped at 7 transactions/sec maximum throughput, and exhibits significant short-term variance in both of these metrics due to its mining protocol -- it is not uncommon for transactions to take much more than an hour to confirm. 
These fundamental protocol limits pose problems for the protocol's scale~\cite{croman2016scaling}. 
Further, nodes can take up to a few days to start up, as validating the blockchain involves vetting every single transaction since the deployment of the system. 
The MimbleWimble~\cite{mimblewimble} proposal promises to save on disk storage by pruning spent transactions through homomorphic cryptography.

In the last two years, a flurry of protocols that tackle scale and efficiency of Bitcoin have been proposed. 
Bitcoin-NG~\cite{eyal2016bitcoin} refactors the creation of blocks such that block dissemination is no longer a bottleneck.
In a similar fashion, both Pass et al.~\cite{pass2016hybrid} and ByzCoin~\cite{kokoris2016enhancing} make use of the underlying PoW mechanism to choose quorums to process transactions in each epoch, using PBFT.
Luu et al.~\cite{luusharding2016} similarly use PoW to create logarithmic-sized shards upon which protocols like PBFT can be run.
All of these systems still rely crucially on proof of work, and thus do not tackle a crucial inefficiency.

Proof of stake (PoS) protocols, which employ variants of PBFT among a cohort of nodes whose desired behavior is incentivized through a financial bond, have been suggested as a way to replace computationally wasteful proofs of work. 
While there have been many PoS proposals~\cite{peercoin,NXT,steem,casper,bentovpoa}, few of them have any associated proof of correctness.
Algorand~\cite{micali2017algorand}, Snow White~\cite{bentovsnow} and Ouroboros~\cite{kiayias2016provably} are some of the earliest provably secure proof-of-stake (PoS) protocols. Of these, Snow White provides the strongest guarantees, as it provides both liveness and safety for a system of at most half adversarial nodes.

{\color{blue}Kevin: further discussions need to go here about related works. There is some stuff in the related works section that maybe doesn't even really belong all that much, and definitely some stuff that's kind of missing, especially in the latest wave of work.}

------------------------------------------------------------------------------------------------

\section{Future Work}
\begin{itemize}
    \item \textbf{Network Model}: we conjecture this protocol family also works
        under a partially synchronous network assumption. This is because the
        sampling and querying process makes the system self-synchronized,
        and our current randomized global scheduler model is weaker
        than the round-based model used by strongly synchronized protocols.
        A new model or model transformation is expected.
    \item \textbf{Minting}: in order to build a decentralized cryptocurrency,
        we need a minting mechanism that incentivizes correct behavior of nodes.
    \item \textbf{Membership}: to build a permissionless system, a complete
        staking and unstaking mechanism needs to be implemented, which may fusion into
        the existing Avalanche protocol to benefit from its safety guarantees.
    \item \textbf{Joining}: since Avalanche has relatively high throughput,
        it may take very long time for a new node to catch up the entire history of DAG using
        the existing mechanism.
\end{itemize}

### 2020-11-03 obsolete/oldtheta1.tex
add initial source and 'fr' make target

% -------------
% DO NOT DELETE.
% DO NOT DELETE.
% DO NOT DELETE.
% -------------
Every correct node is chosen by the scheduler at most a fixed number of times, $\delta$, after which the value of every correct node is collected and counted.
Note another (and final) condition: if a node is selected and that node has seen $\phi$ consecutive majorities for the same value, it no longer updates its value, regardless of the output of the polling.

\subsubsection{Safety of \Theta_1}
The expression for the hypergeometric experiment is reasonably complex, and it becomes hard to reason about the different ways that each of the variables inside the expression affects the outcomes. In particular, it is not clear how network size affects the experiment. This is of interest because if the effect of network size has little bearing on the outcomes of the experiment, it has important consequences for the scalability of the $\Theta_1$ protocol. To see how network size affects the system, we can fortunately make usage of the

\subsubsection{Adversarial Optimality}
We will first try to determine the optimal strategy that the Byzantine
adversary can play.  Once we have this strategy, we will analyze how
effective this strategy is.
Recall that we do not care about violating liveness for conflicting
transactions, so the adversary's objective is to violate
safety, that is, it wants some correct node to decide red while another
correct node decides blue.

As a thought experiment, consider that most correct nodes lean red
while only one correct node leans blue.  Given our sampling protocol,
it is clear that if there are not too many Byzantine nodes, the red-leaning
nodes will continue to lean red, while the blue-leaning node will
start increasing its confidence in red and eventually become red-leaning
as well.  It thus stand to reason that a primary objective of the adversary
is to have half of the correct nodes lean red, while the other half lean
blue, and to keep it this way.

However, the adversary faces a conflict.  Consider that the adversary
has only limited time ($\phi$ rounds) to accomplish its goal of
violating safety.  When half the nodes are leaning blue, and half
the nodes are leaning red, their confidence in either blue or red
will not grow rapidly, and they likely will not decide in the limited
amount of time.  Also, if some node $u$ is close to deciding blue
and a small majority of nodes are leaning blue, then if $u$ polls
a Byzantine node it is not clear whether the Byzantine node should
return red or blue.  If the Byzantine node returns blue, it may
cause $u$ to decide blue but still give the adversary a chance to
get other nodes to decide red.  If the Byzantine node returns red,
it may help restore the balance between red- and blue-leaning nodes.

This conflict for the adversary illustrates exactly the strength of
the Snowball protocol.  While it is relatively easy for the adversary
to prevent deciding on conflicting transactions, it is difficult for
the adversary to cause two correct nodes to decide different transactions.

\begin{comment}
If this is the situation, then if a correct
node samples other nodes, it will, by expectation, receive the same number of
red and blue votes from the correct nodes that it samples.

The Byzantine strategy then is straightforward:
if there are as many red-leaning nodes as blue-leaning nodes and
a red-leaning node samples from a Byzantine node, the Byzantine node
returns red, and vice versa.
If the node has equal confidence in red or blue, it does not matter
which color the Byzantine node returns and can, say, flip a coin.
If, however, the system is not in balance, and
there are more red than blue nodes, then a Byzantine node will always
``push back'' by returning blue (and vice versa).
We will later demonstrate that this strategy is optimal.

Next, we need to show that although optimal, the strategy cannot
cause a violation of safety with high probability.
We will show that whp no 
correct node ever decides (if there are conflicting values), and in particular
no two correct nodes ever disagree.
If the Byzantine strategy is non-optimal, correct nodes may decide, but
they will decide the same value whp.
\end{comment}

\begin{comment}
We define \emph{skew} as the maximum among the distances between the
confidence vectors (interpreted as points in a Euclidean space)
of correct node pairs.  For example, if there are three correct nodes
with confidence vectors $\langle 0, 3 \rangle$, $\langle 1, 3 \rangle$,
and $\langle 4, 0 \rangle$, then the skew is~5 (the distance between
the first and last confidence vector endpoints).
The optimal strategy for the Byzantine nodes is to maximize skew.

% TODO: Expand in sentences the notation written about maximal policy below. 
We want to characterize \textit{drift} in the differences of values between any two correct nodes. Specifically, letting the absolute difference between the values of any correct node correspond to some confidence differential, represented by $\max_{\Delta}(u) = u[0]-u[1]$ if $|u[0]| > u[1]$ and $\max_{\Delta}(u) = u[1]-u[0]$ if $u[1] > |u[0]|$, we desire to precisely quantify bounds on the maximum drifts between any two correct nodes, and for all possible execution instances $\mathcal{I}_i$. In other words, what is the execution path $\mathcal{I}_{opt}$ such that
\[
    \mathcal{I}_{opt}(\mathcal{C}^{t_0}) = \max_{\forall i}\{\mathcal{I}_i(\mathcal{C}^{t_0})\}
\]
where
\[
    \mathcal{I}_{i}(\mathcal{C}^{t_0}) = \max_{\forall u, v \in \mathcal{C}^{t_\phi}}\{\max_{\Delta}(u)-\max_{\Delta}(v)\}
\]
To exactly characterize the probability of a safety failure of the optimal execution path, we would need to calculate the probability distribution over all possible permutations of Byzantine answers, which is a function of the choices of execution of the correct nodes as well as the distribution of the choices of the neighbors.
In other words, over all possible execution paths (i.e. there are $c$ nodes to choose at each time step, for a total of $\phi^{c}$ ways of choosing over $\mathcal{C}$ nodes) and over all possible choices of the $k$-sampled neighbors (for a total of ${n \choose k}$ such choices, per node, per execution round), we would need to keep track of all of the possible tuples of the Byzantine value updates. 
Generating such a stochastic model would, plainly speaking, be infeasible. 
\end{comment}

\paragraph{Optimality Condition}
We can simplify the analysis greatly if we fix the Byzantine strategy to some optimal policy, $\mathcal{A}_{\pi}$. 
Let $s$ be a three-tuple where the first entry is a color configuration of the set of nodes in $\mathcal{C}$ (i.e. the set $\lbrace(u_1[\mathtt{R}], u_1[\mathtt{B}]), (u_2[\mathtt{R}], u_2[\mathtt{B}]), \dots, (u_c[\mathtt{R}], u_c[\mathtt{B}])\rbrace$, where $u[\mathtt{R}], u[\mathtt{B}] \in [0, \phi]$), the second entry is the chosen node at some round $u \in \mathcal{C}$ in round $t_i$, and the third entry is the round itself $t_i$. Let $\mathtt{S}$ be the set of all such tuples $s$. The elements of $\mathtt{S}$ are the necessary pieces of information needed for the Byzantine strategy. 
For each state $s \in \mathtt{S}$, the set of allowable actions for the Byzantine nodes is either $\mathtt{R}$ or $\mathtt{B}$. Let $P(s' | s, {\mathtt{R}})$ be the probability of transitioning from state $s$ to $s'$ given that the Byzantine strategy responds with $\mathtt{R}$, and conversely for $\mathtt{B}$. Let $\mathtt{Z}(s)$ be the reward for entering state $s$. Let the policy $\mathcal{A}_{opt}(s)$ be the optimal policy which returns the color chosen by the Byzantine node at state $s$, given round $i$. 
The utility $\mathcal{U}$ of $s$ (and the optimal policy) are then defined as
\begin{equation}
\begin{split}
\mathcal{U}(\mathcal{A}_{opt}, s) =\ \mathtt{Z}(s)\ +& \sum_{s' \in S} P(s' | s, \mathcal{A}_{opt}(s))\\
\times &\ \mathcal{U}(\mathcal{A}_{opt}, s')
\end{split}
\end{equation}
where 
\begin{equation}
\begin{split}
    \mathcal{A}_{opt}(s) &= \argmax_{a \in \lbrace \mathtt{R}, \mathtt{B} \rbrace} \left \lbrace \sum_{s' \in S}P(s'|s, a)\mathcal{U}(\mathcal{A}_{opt}, s') \right \rbrace
\end{split}
\end{equation}
These are the equations that follow from Bellman's optimality principle. Note that we do not include a discount factor. Further, we assign a reward of 1 to the states where there exists two nodes $u, v$ such that $\gamma_u \geq \gamma$ and $\gamma_v \leq 1/\gamma$.

Since the state space is computationally infeasible, we solve for the optimal policy inductively, rather than iteratively. 
Let $count(\mathcal{C}^{t_i}, \mathtt{R})$ be defined as: 
\begin{equation}
\begin{split}
    count(\mathcal{C}^{t_i}, \mathtt{R}) = \sum_{u \in \mathcal{C}^{t_i}} \mathbbm{1}(u, \mathtt{R})
\end{split}
\end{equation}
where $\mathbbm{1}$ is the indicator function defined as:
\begin{equation}
\begin{cases}
    \mathbbm{1}(u, \mathtt{R}) = 1\ \text{\textbf{if}}\ u[\mathtt{R}] > u[\mathtt{B}]\\
    \mathbbm{1}(u, \mathtt{R}) = 0\ \text{\textbf{if}}\ u[\mathtt{R}] < u[\mathtt{B}]\\
    \mathbbm{1}(u, \mathtt{R}) = \{1, 0\}\ \text{\textbf{if}}\ u[\mathtt{R}] == u[\mathtt{B}]\ \text{\textbf{and} $u$ lastly}\\
    \text{preferred $\{\mathtt{R}, \mathtt{B}\}$}
\end{cases}
\end{equation}
and conversely for $count(t_i, \mathtt{B})$.

Let $\mathcal{C}_{\mathtt{R}}$ be the set of correct nodes that the Byzantine nodes color as red at the cold start phase, and $\mathcal{C}_\mathtt{R}$ the converse. It is easy to see that a coloring of equal split between red and blue in the cold-start phase is the optimal configuration for a single iteration of the $\Theta_1$ protocol. It is not clear, however, how this configuration must evolve (if at all) as time progresses.

Let $\mathcal{A}_\pi$ be the chosen Byzantine strategy described formally in Figure \ref{figure:mathcal_a_pi}. 
\begin{figure}[h!]
\label{figure:mathcal_a_pi}
\fbox{
    \parbox{0.95\linewidth}{
        \begin{center}
            \textbf{The $\mathcal{A}_\pi$ policy}
        \end{center}
        Let the Byzantine nodes set $\mathcal{C}_\mathtt{R}$ to be exactly half of the correct nodes in the network, and $\mathcal{C}_\mathtt{B}$ the other half at the cold-start phase. We initialize a local memory $\mathds{M}$, which stores various information about the history of execution up to step $t_i$. Let $\{\mathcal{C}', u, t_i\} \leftarrow s$, where $\mathcal{C}^{t_i} = \mathcal{C}'$ (note that there are multiple valid versions of $\mathcal{C}^{t_i}$, nevertheless in this case we refer to the specific set $\mathcal{C}$ extracted from state $s$). 
        \begin{itemize}
            \item If $count(\mathcal{C}^{t_i}, \mathtt{R}) = c/2$:
            \subitem In this case, the set of correct nodes is equally split. If $u \in \mathcal{C}_\mathtt{R}$, then all Byzantine nodes are colored red. Conversely for blue. Add $s$ to $\mathds{M}$, and return. 
            \item If $count(\mathcal{C}^{t_i}, \mathtt{R}) < c/2$:
            \subitem In this state, the set of correct nodes has shifted to a majority blue. This implies that at some time step $t_j < t_i$, a node $v \in \mathcal{C}_\mathtt{R}$ gained at least 1 extra confidence for blue than for red. Locate this node $v$ by searching the history stored in memory $\mathds{M}$. Move $v$ to the corresponding color assignment, i.e. set $\mathcal{C}_\mathtt{R} = \mathcal{C}_\mathtt{R}\backslash v$ and $\mathcal{C}_\mathtt{B} = \mathcal{C}_\mathtt{B} \cup v$. Search in $\mathcal{C}^{t_i}$ for all nodes $v'$, s.t. $\mathbbm{1}(v', \mathtt{B}) = 1$, and find the $v'$ with the minimum skew between blue and red, i.e. $\min \{v'[\mathtt{B}] - v'[\mathtt{B}]\} \forall v'$. Move the target $v'$ to the opposite color assignment, i.e. set $\mathcal{C}_\mathtt{B} = \mathcal{C}_\mathtt{B}\backslash v'$ and $\mathcal{C}_\mathtt{R} = \mathcal{C}_\mathtt{R} \cup v'$. Finally, if $u \in \mathcal{C}_\mathtt{R}$, then all Byzantine nodes are colored red. Conversely for blue. Add $s$ to $\mathds{M}$, and return. 
        \end{itemize}
    }
}
\end{figure}

\begin{comment}
    \begin{algorithm}
    \begin{algorithmic}[1]
    \Procedure{$\mathcal{A}_\pi$}{$s$}
    \State $\{\mathcal{C}', u, t_i\} \leftarrow s$
    \If{$\mathbbm{1}(u, \mathtt{R})$} \Comment{$u$ prefers $\mathtt{R}$}
    \If{$u \in \mathcal{C}_\mathtt{R}$}
    \State
    \EndIf
    \EndIf
    \State $\mathcal{A}(u, \mathcal{C})$ \Comment{Deterministic Byzantine strategy}
    \State $s \xleftarrow{\$^k} \mathcal{N}$\textbackslash $u$
    \If{count($s, \mathtt{R}$) $\geq$ $\alpha k$}
    \State $u[\mathtt{R}]~ += 1$
    \EndIf
    \If{count($s, \mathtt{B}$) $\geq$ $\alpha k$}
    \State $u[\mathtt{B}]~ += 1$
    \EndIf
    \State update($\mathcal{C}$, $u$)
    \EndProcedure
    \end{algorithmic}
    \end{algorithm}
\end{comment}

\begin{theorem}
The Byzantine strategy $\mathcal{A}_\pi$ is optimal, i.e. $\mathcal{A}_\pi = \mathcal{A}_{opt}$.
\end{theorem}
\begin{proof}
We inductively prove the claim. 
The base case is when the nodes are equally split in half, i.e. $count(\mathcal{C}^{t_i}, \mathtt{R}) = c/2$. Let $\mathcal{H}_{\mathcal{N}, c/2 + b + j}^k/\mathcal{H}_{\mathcal{N}, c/2 - j}^k$ represent the expected ratio growth of $\gamma_u$, where $u \in \mathcal{C}^\mathtt{R}$. 
Naturally, this ratio grows as $j$ grows (i.e. as the number of red-preferring correct nodes increases). 
However, the converse is also true, meaning that as $j$ increases, the $\gamma_v$ ratio for $v \in \mathcal{C}_\mathtt{B}$ decreases at the same rate. 
In particular, this rate is poly-exponential, meaning that even small changes to $j$ incur large benefits (and conversely penalties) on the confidence ratios. 
Therefore, the optimal split between correct nodes that maximizes the probability of $\gamma_u > \gamma_v$ and $\gamma_v > 1/\gamma$ where $u$ prefers red and $v$ prefers blue is when $j = 0$. 
In the usual case, the Byzantine nodes initially label the correct nodes equally between red and blue, and respond with red whenever a red-labelled correct node samples, and blue whenever a blue-labelled correct node samples. 
However, in the case when an initially red-labelled node (i.e. $u \in \mathcal{C}_\mathtt{R}$) suddenly switches confidence to blue, the Byzantine node must update it's strategy. 
The Byzantine nodes have three options: either keep responding red to the newly switched node, update responses to blue, or choose a new node and update its colors. 
Since every selection by the scheduler is uniformly at random, the expected number of times a node will be selected starting from time $t_i$ is equal amongst all correct nodes and is exactly $\phi - i / c - 1$. 
Therefore, in expectation, there is no difference between $u$ and a node $v$ which prefers the same color and which has no bigger skew than $u$. 
Sticking to the prior color labelling is clearly suboptimal since the growth of $\gamma_*$ of the new majority color will be larger than the minority color. Therefore, the optimal strategy is to simply pick the node of same color as the newly flipped node that has the \textit{minimum} skew (i.e. confidence) in its color. 
\end{proof}

Under no adversarial presence, the progress made by correct nodes is equal, in expectation, amongst all honest nodes. More formally, let $\mathcal{B} = \emptyset$. Then, $\forall u^{t_i}, v^{t_i} \in \mathcal{C}^{t_i}$, $i \geq 0$, 
\[
    \mathbb{E}[u^{t_i}[\mathtt{R}]] = \mathbb{E}[v^{t_i}[\mathtt{R}]]
\]
\[
    \mathbb{E}[u^{t_i}[\mathtt{B}]] = \mathbb{E}[v^{t_i}[\mathtt{B}]]
\]
At any time-step, the probability of each correct node being selected and subsequently adding either a red ball or blue ball is a function of the underlying number of red and blue preferences, but otherwise equal and uniform among all nodes. In Lemma \ref{lemma:expected_absorption}, we computed the expected number of steps to reach the state where all nodes prefer red. Upon reaching this state, every new added color will be the same, and growth of the opposing color will be zero.

If Byzantine nodes are introduced, the expectation changes as a function of the underlying Byzantine strategy. Let $p_q^l(\mathtt{R}, \mathcal{S}, k, \alpha)$ be the probability of achieving at least $l$ consecutive successes for the color red over a sequence of $q$ trials given that $\mathcal{S}$ nodes in the network $\mathcal{N}$ prefer red. We let $\epsilon < 1$ be a security parameter, and after fixing $q$, we solve for the minimal value of $l$ such that 
\[
    p_q^l(\mathtt{R}, c, k, \alpha) \leq \epsilon
\]
In other words, we compute the minimal number of consecutive support for red until, with probability no more than $\epsilon$, at least $c$ nodes in the network also prefer red.

It is easy to see that the probability is equal between the opposing colors as well as maximal when $\mathcal{S} = c/2 + b$. However, this probability is clearly much smaller than $\mathcal{O}(\epsilon^2)$. Suppose that the Byzantine nodes deviate from the optimal strategy in order to render the probability of committing at least one color to some value greater than $\epsilon$. In that case, the Byzantine nodes must allow the correct nodes to swing away from an equal network split.

At this point in the deviation of the strategy, the ``snowball'' effect of $\Theta_1$ becomes apparent and highly compounded. Let $c/2 + \delta$ represent the expansion of the number of correct nodes towards red, and $c/2 - \delta$ the contraction of the number of correct nodes towards blue. Using the result from a prior lemma, the ratio of the speed of absorption will now be $g(c/2 + \delta)$.

\begin{comment}
========= END PHASE 1

Fortunately, we can make usage of a useful observation: the probability of two independent successes for \texttt{opposing} values for any pair of hypergeometric experiments is maximum only when the frequency of the values in the network is exactly equal. 
To prove this, we simply need to solve for $i$ in the following expression:

\begin{equation}
\max \{P(\mathcal{H}_{n, i}^k \geq \alpha k)*P(\mathcal{H}_{\mathcal{N}, n-i}^k \geq \alpha k)\}, 0 \geq i \geq n
\end{equation}

This probability is increases exponentially quickly as $i \rightarrow n/2$, with a maximum at $n/2$, and an exponential dropoff from $i = n/2$ to $n$. 
This observation is tremendously helpful because it allows us to fix the Byzantine adversary to a single optimal strategy: force the correct nodes into two equally split partitions, and -- throughout all $\delta$ executions -- put all the Byzantine nodes towards answering the corresponding value to each partition. 
Any other split, including non-fixed dynamic Byzantine strategies, would incur lower probability of success.

Following on our observation, we can now fix the Byzantine strategy to the following: at time $t_0$, the Byzantine nodes choose two equal partitions of correct nodes, and for each subsequent time step, every Byzantine node will answer $1$ whenever the first partition queries and $2$ whenever the second partition queries.

The probability that we receive $\phi$ or more successes over $\delta$ trials is a complicated expression, whose exact value requires a recursive expression. We leave the reader to Feller\cite{feller1} (Vol 1. Sec. XIII.7) for more information. Let $\mathcal{U}_{\delta}(\mathcal{H}_{\mathcal{N}, C^{t_i}}^k)$ be the random variable that outputs the number of consecutive successes of the hypergeometric experiment over $\delta$ trials. 
Then, the probability that a specific correct node receives $\phi$ or more consecutive successes is then:

\begin{equation}
P_{consecutive} = \frac{1}{|\mathcal{C}|}P(\mathcal{U}_{\delta}(\mathcal{H}_{\mathcal{N}, C^{t_i}}^k) \geq \phi)
\end{equation}

The probabilities are symmetric across all correct nodes. Therefore, the probability of safety failure ($P_{sf}$) over all possible $|\mathcal{C}/2|^2$ pairs of correct nodes is then:

\begin{equation}
    P_{sf} = |\mathcal{C}/2|^2 * P_{consecutive}^2
\end{equation}

The probabilities of safety failure are shown in Figure \ref{fig:safety_failure}.

\begin{figure}[h!]
\centering
\includegraphics[width=\linewidth]{./figures/safety_failure.pdf}
\caption{Probability of safety failure, $\phi = 20$, $\delta = 100$, $|\mathcal{N}| = 500$, and $k = 10$, for various size of Byzantine nodes and $\alpha k$.}
\label{fig:safety_failure}
\end{figure}

% \subsubsection{Liveness of \Theta_1}
% A key metric related to liveness is convergence latency, the amount of time between the introduction of a value proposal to the time every correct node agress on the same value.
% Described in terms of network configurations, this convergence latency calculates the expected number of rounds from any bivalent network configuration until reaching a state where all correct nodes vote for the same value. 
% We first show that convergence latency for a Byzantine-free network is logarithmic in the size of the network. Later, we calculate bounds on the maximum adversarial size that still guarantees expected logarithmic convergence. 
% Let the system be initialized to state $\psi = \{|\mathcal{C}_1^{t_0}|, |\mathcal{C}_2^{t_0}|\}$. 
% We perform a Markovian analysis on the probabilities of transitioning to any of all the possible states arising from $\psi$. 
% The state space $\mathbb{S}$ of all possible configurations arising from $\psi$ is all possible bivalent network splits between values $1$ and $2$, such that $|\mathcal{C}_1^{t_0}| + |\mathcal{C}_2^{t_0}| = |\mathcal{N}|$. 
% As an example, suppose the total number of correct nodes is 100. Then, all possible states of the system are $\{0, 100\}, \{1, 99\}, \dots, \{100, 0\}$.
% We can now construct the transition probability matrix $M$ of size $|\mathbb{S}| \times |\mathbb{S}|$. The expression for generating the values inside $M$ is highly verbose, however to provide the reader some intuition on how it is build, we given an example. Suppose the system is at state $\{50, 50\}$. Then, the system can only move to states $\{49, 51\}$, $\{51, 49\}$, or itself. The probability of moving to $\{49, 51\}$ is the probability of selecting a node that votes value $2$ and having that node switch to value $1$. The probability of staying in place is the probability that any selected node does not achieve the required $\alpha k$ threshold, or it achieves a threshold for its own value. $M$ will be sparse since most states will have only a transition to a small number of states.

% Given any starting state $s \in \mathbb{S}$, the hitting time $\mu$ is the expected number of transitions to reach states $A = \{\{0, |C|\}, \{|C|, 0\}\}$. 
% To calculate the hitting time, we first construct the system of linear equations shown in equation \ref{equation:minimalsolution} for all states in $\mathbb{S}$.

% \begin{equation}
% \label{equation:minimalsolution}
% \begin{cases}
% \mu_{sA} = 0, & \forall s \in A\\
% \mu_{sA} = 1 +\sum_{r\notin A}^{}{p_{sr}\mu_{rA}}, & \forall s \notin A
% \end{cases}
% \end{equation}

% The minimal non-negative solution vector $\mu_{A}~=~(\mu_{sA}, s \in \mathbb{S})$ to the system of linear equations in equation \ref{equation:minimalsolution}, yields the expected number of rounds to reach one of the convergence states. In other words, the solution gives the expected number of rounds for each node before all nodes agree on the same value.

% The convergence latency grows logarithmically with the network size, as shown in Table \ref{table:growth_worst_case}. 
% At 10k nodes, for example, the equal-network split expected time to converge is around 10 rounds. 
% \begin{table}[h!]
% 	\centering
% 	\begin{tabular}{llllll}
% 		Net. Size   & 600    & 1200   & 2400   & 4800   & 9600    \\ \hline
% 		Exp. Conv. & 7.18 & 7.94 & 8.69 & 9.44 & 10.19
% 	\end{tabular}
% 	\caption{Expected number of per-node-iterations to convergence starting at worst-case (equal) network split. Network sizes grow exponentially, whereas worst-case convergence grows logarithmically. $k = 10,\ \alpha=6$.}
% 	\label{table:growth_worst_case}
% \end{table}

\paragraph{Adversarial Progress} 
Without loss of generality, let $\lambda$ be chosen such that the expected frequency of $v_1$ and $v_2$ amongst the $\lambda$ chosen neighbors is nearly the same as $cn(v_1)$ and $cn_(v_2)$. 
In order to prevent progress, the number of adversarial nodes must then be larger than the absolute difference between the total number of honest nodes that vote for $v_1$ and $v_2$. This absolute difference denotes the total number of nodes that the adversary must have in its control in order to maintain an equal split between honest nodes. This adversarial size denotes the minimal adversarial size to prevent convergence in expected logarithmic number of rounds.

Conversely, we can look at the case when the adversary pushes honest nodes to commit one value, and then equivocates. In this case, As long as the adversary is greater than ($1 - \beta/\lambda$), honest nodes that have not yet committed will no longer be able to commit. Any and all solicitations from any honest node will result in expected $(\lambda-\beta)/\lambda$ or more equivocations, which is insufficient to make progress.

\label{section:analysis}
\subsection{Snowball Safety Analysis}
% \label{subsec:safetyanalysis}
% \subsubsection{Terminology and Definitions}
% \label{subsubsec:termanddef}
% In this subsection, we refresh the reader of some of the old notations and as well introduce new notation, terminology, and definitions. 
%Finally, let $T_i$ and $T_j$ be two conflicting transactions belonging to the same conflict set.

In order to analyze the safety of our protocol, we turn our focus exclusively towards two transactions of interest, $T_i$ and $T_j$, which belong to the same conflict set. 
Although the system is comprised of many conflicting transactions, we can safely ignore transactions that are not in the ancestry or progeny sub-graph of $T_i$ and $T_j$, since -- by construction -- those transactions are completely independent from $T_i$ and $T_j$ and thus do not contribute to the evolution of preference for either of the transactions of interest. 
Lastly, we note that unless explicitly stated, the analysis is through a ``God's eye'' view of the network.

In the specification section (S. \ref{subsection:specification}), we introduced the general definition of \textit{preference} for a transaction $T$. 
Using that definition, we construct the notion of \textit{network support} for $T$, denoted as $\mathcal{S}_{u}^{t}(T)$, which represents the fraction of nodes in $\mathcal{N}$ that prefer $T$ at time $t$ as perceived by node $u$. 
To illustrate, suppose that the correct nodes are equally split in preference between $T_i$ and $T_j$. Furthermore, let $\mathcal{B}_{u}^{t}(T_i), \mathcal{B}_{u}^{t}(T_j) \subseteq \mathcal{B}$, be the subsets of the Byzantine nodes that vote $T_i$ and $T_j$ respectively as perceived by node $u$, which are mutually exclusive since according to $u$'s perspective the same Byzantine node cannot prefer two equivocating transactions simultaneously. 
A correct node $u$ would then observe that the network support for $T_{\{i,j\}}$ must be $\mathcal{S}_{u}^{t}(T_{\{i,j\}}) = (|\mathcal{B}_{T_{\{i,j\}}}| + |\mathcal{C}|/2)/|\mathcal{N}|$).

Lastly, we call a \textit{solicitation} the act of $u$ querying a set of $k$ neighbors, chosen uniformly at random, to discover if they prefer transaction $T_i$. 
Note that necessarily $k \leq |\mathcal{N}|$, but otherwise the choice of size of $k$ is independent of the size of $\mathcal{N}$, as discussed previously.

% \subsubsection{Network Model}
\paragraph{Network Model}
We initially assume that every node has full view of every other node. 
Later, we relax this requirement to allow for looser views. 
Each node maintains a logical round counter (time) that starts at 0 at its initial solicitation and is incremented by one for each solicitation of a child transaction. 
Importantly, we remind that there is no notion of global system rounds, unless explicitly stated.

For simplicity, we assume that each node receives all $k$ responses of a solicitation within the same round. 
Later, we relax this requirement to interleave solicitation responses.

% \subsubsection{Byzantine Adversary}
% \label{subsec:byzantineadversary}
\paragraph{Byzantine Adversary}
The adversary, whose nodes are in set $\mathcal{B}$, is round-adaptive and can both read all messages and view the state of all nodes in the network.
Round-adaptive means that the adversary can choose a new strategy for each of its nodes after seeing the messages exchanged by correct nodes.
In other words, whenever a correct node solicits its neighbors, the adversary can correspondingly instruct nodes under its ownership before the correct node has the chance to receive any responses. 
Such an adversary is only prohibited from modifying the internal state of a correct node and the in-transit responses between any two correct nodes.

\subsubsection{Analysis}
\label{subsec:analysis}
% \paragraph{Statistical Errors and Parameters}
% We begin the analysis by first extracting some statistical bounds on protocol parameters. 
% We extract these bounds by isolating a single correct node $u$, and analyzing the results of a single isolated network polling experiment. Basically, we freeze the network at some specific point in time $t$ and then compute the statistical errors between two nodes $u$ and $v$.

% For notational ease, we represent $\mathcal{S}_{u}^{t}(T_i)$ as $\mathcal{S}'$. 
% Let $\mathcal{H}_{N, \mathcal{S}'}^k \rightarrow [0, k] \in \mathbb{N}$ be a random variable which represents the total count of the $k$ randomly chosen neighbors that prefer transaction $T_i$ at round $t$ as queried by $u$. $\mathcal{H}$ is parameterized by the sample size $k$, the network $\mathcal{N}$, and the network support of $T_i$, $\mathcal{S}'$. For some constant $1/2 \leq \alpha \leq 1$, we are interested in the probability that a majority or more, $\alpha k \in \mathbb{N}$, of the sampled $k$ neighbors prefer $T_i$. The probability can be calculated as follows, 
% \begin{equation}
% P(\mathcal{H}_{\mathcal{N}, \mathcal{S}'}^k \geq \alpha k) = \sum_{j = \alpha k}^{k} \frac{{\mathcal{S}'|\mathcal{N}| \choose j}{|\mathcal{N}| - \mathcal{S}'|\mathcal{N}| \choose k - j}}{{|\mathcal{N}| \choose k}}
% \label{eq:hypergeometric}
% \end{equation}
% where the expected value of $\mathcal{H}_{\mathcal{N}, \mathcal{S}'}^k$ is exactly $k\mathcal{S}'$. From now on, if $\mathcal{H}_{N, \mathcal{S}'}^k \geq \alpha k$, we say that the solicitation was successful. 
% %Intuitively, the meaning behind this term is that some threshold majority $\alpha k$ of the solicited neighbors indeed support the transaction in question.

% We use the Hoeffding \cite{hoeffding1963probability} inequality to calculate the probabilities of deviation from the expected result of the polling experiment. Hoeffding's tail bounds then show that -- given a fixed constant $\psi$ -- the probability $\mathcal{H}$ deviating by $\psi$ from the expected value is given by, 
% \begin{equation}
% \begin{split}
%     P(\mathcal{H}_{N, \mathcal{S}'}^k \leq (\mathcal{S}'-\psi)k) %&\leq e^{-k\mathcal{D}(\mathcal{S}'-\psi, \mathcal{S}')} %\\
%     %&
%     \leq e^{-2\psi^2k}
% \end{split}
% \end{equation}
% % where $\mathcal{D}(\mathcal{S}'-\psi, \mathcal{S}')$ is the Kullback-Leibler divergence, measured as
% % \begin{equation}
% % \begin{split}
% %     \mathcal{D}(\mathcal{S}'-\psi, \mathcal{S}') &= (\mathcal{S}'-\psi) \log (\frac{\mathcal{S}'-\psi}{\mathcal{S}'}) \\
% %     &+ (1 - \mathcal{S}' + \psi) \log \frac{1 - \mathcal{S}' + \psi}{1 - \mathcal{S}'}
% % \end{split}
% % \end{equation}
% The very first thing to notice is that the probabilities are independent of network size, meaning that network size doesn't affect the performance of the protocol.

% The second thing to notice is that the the upper bound $e^{-2\psi^2k}$ indicates that the probability of success of a solicitation is an inverse exponential parametrized by the sample size $k$. 
% For $k$, every +1 increase means an \textit{exponential decrease} in network view error.

% The third and last thing to notice is the implication for the value $\psi$, which has a quadratic effect on the exponential decrease in network view error. The network split amongst the correct nodes that maximizes the tail bound probability is exactly when half of the correct nodes are partitioned to initially prefer $T_i$ and half to initially prefer $T_j$. If the Byzantine nodes partition the network at some other ratio, the probability the Byzantine adversary successfully forces $u$ to prefer $T_i$ and $v$ to prefer $T_j$ for one solicitation is suboptimal. However, although this conclusion is correct over a single polling experiment, it may not hold over multiple experiments. This is because it would need to account for changes in the underlying graph structure. To that end, we now begin to analyze the protocol as it evolves over time.

% % The implication of this result is that the optimal strategy for the Byzantine nodes is to minimize the deviation of $\psi$ (i.e. $\psi, k \rightarrow 0$), by splitting the network equally into two conflicting halves.

% %The structure of the dag confers an important property. 
% %Due to the invariant imposed by the structure of the dag which states that every transaction on the path to a preferred child must be preferred in turn, every successful solicitation of a transaction implies a successful solicitation of its parent. 
% % Let $T' \xleftarrow{*} T$. At any time $t$, we necessarily have 
% % \begin{equation}
% % P(\mathcal{H}_{N, \mathcal{S}_{u}^{t}(T')}^k \geq \alpha k) \geq P(\mathcal{H}_{N, \mathcal{S}_{u}^{t}(T)}^k \geq \alpha k)
% % \end{equation}
% % We can now substitute the net total solicitations of the child transactions with an equivalent set of repeated solicitations for the parent transaction. 
% % This simplification allows us to focus solely on $T_i$, and safely ignore all other transactions except direct ancestors of $T_i$.

% % Let $\mathcal{P}_b$ = $\{T_i, T_j\}$ be a set of two conflicting transactions. 
% % Let $u, v \in \mathcal{C}$ be any two correct nodes, and let $c_{\{u, v\}}(T_{\{i,j\}}, n)$ represent the confidence (i.e. the number of successful solicitations) for each of the transactions by each of the two correct nodes after sampling $n$ children transactions at rounds $t_0, t_1, \dots, t_n$, respectively.

% % The expectation of the confidence of transaction $T_i$ after $n$ solicitations is equal to the probability of a successful solicitation for $T_i$ summed independently over every one of the $n$ experiments. More precisely, 
% % \begin{equation}
% % \begin{split}
% %     \mathbb{E}[c_{u}(T_{i}, n)] &=\ \sum_{j = 0}^{n}P(\mathcal{H}_{N, \mathcal{S}_{u}^{t_j}(T_i)}^k \geq \alpha k) \\
% % \end{split}
% % \end{equation}
\paragraph{Evolution Over Time}
The Byzantine nodes have the freedom to attack the system using any strategy. 
In order to maximize damage, the Byzantine nodes must pick the path of execution that ensures that the confidence of the two equivocating transactions is maximized between two correct nodes.
The adversary has access to a large set of attack options. On the one extreme end, the adversary may launch a ``bifurcation attack'', where the attacker equivocates at the very beginning and attempts to split the network equally in half, hoping that part of the network slowly prefers one transaction and the other half prefers the other. On the other extreme end, the adversary can launch a ``double-spend'' attack, where the adversary behaves correctly at the beginning, but after some large amount of time equivocates.

% To calculate these expectations, and consequently define the most damaging Byzantine strategy, we must first extract some tail bounds on $\mathcal{H}$. 
%Figure \ref{fig:hypersplits} shows the probability of a single successful experiment over various network splits. 
%The probability of a single successful iteration decreases poly-exponentially with increasing difference between $\mathcal{S}N$ and $\alpha k$. 
% The most relevant tail bound can be extracted from Hoeffding \cite{hoeffding1963probability}. Given a network support $\mathcal{S}'$, the expected value of $\mathcal{H}_{N, \mathcal{S}'}^k$ is exactly $k \mathcal{S}'$. Hoeffding's inequality then shows that -- given a fixed constant $\psi$ -- the probability $\mathcal{H}$ deviating by $\psi$ from the expected value is given by, 
% \begin{equation}
% \begin{split}
%     P(\mathcal{H}_{N, \mathcal{S}'}^k \leq (\mathcal{S}'-\psi)k) %&\leq e^{-k\mathcal{D}(\mathcal{S}'-\psi, \mathcal{S}')} %\\
%     %&
%     \leq e^{-2\psi^2k}
% \end{split}
% \end{equation}
% % where $\mathcal{D}(\mathcal{S}'-\psi, \mathcal{S}')$ is the Kullback-Leibler divergence, measured as
% % \begin{equation}
% % \begin{split}
% %     \mathcal{D}(\mathcal{S}'-\psi, \mathcal{S}') &= (\mathcal{S}'-\psi) \log (\frac{\mathcal{S}'-\psi}{\mathcal{S}'}) \\
% %     &+ (1 - \mathcal{S}' + \psi) \log \frac{1 - \mathcal{S}' + \psi}{1 - \mathcal{S}'}
% % \end{split}
% % \end{equation}
% The very first thing to notice is that the probabilities are independent of network size, and are only a function of the sampled neighbor size $k$. This means that the network size doesn't affect performance of the protocol.

% The second thing to notice is that the the upper bound $e^{-2\psi^2k}$ indicates that the probability of success of a solicitation is an inverse exponential parametrized by the squared value of $\psi$. 
% The implication of this result is that the optimal strategy for the Byzantine nodes is to minimize the deviation of $\psi$ (i.e. $\psi, k \rightarrow 0$), by splitting the network equally into two conflicting halves. 
%Correct nodes can be fooled by Byzantine nodes by tuning parameters such that $e^{-2\psi^2k}$ is maximized. 
%This occurs when the deviation and sample size approach zero (i.e. $\psi, k \rightarrow 0$). 
% Since the Byzantine nodes are limited in their network presence, the most damaging strategy must necessarily be when all correct nodes are equally split.

We start calculating the evolution of confidence between $T_i$ and $T_j$ by first computing the network split between the two transactions without introducing any children transactions. Let $T_*$.\texttt{sum} be the total sum of all chits for transaction $T_*$ amongst all correct nodes in the network. Let $\mathcal{C}_1$ be the subset of correct nodes that prefer $T_i$ at time $t = 0$, and $\mathcal{C}_2$ the corresponding set for $T_j$. Lastly, let $\mathcal{Z}_1$ be the subset of correct nodes that the Byzantine nodes have chosen to force to prefer $T_i$, and $\mathcal{Z}_2$ be the corresponding set of nodes for $T_j$. At $t = 1$ (the base case), the we can compute the new values of $T_*$.\texttt{sum} as follows:

\begin{equation}
    \begin{split}
        & T_1.\text{\texttt{sum}} = 0;\ T_2.\text{\texttt{sum}} = 0;\\
        t_1:\ \ \ &\gamma_{1}^{t_1} \leftarrow \frac{\mathcal{Z}_1}{\mathcal{C}} P(\mathcal{H}(\mathcal{N}, \mathcal{C}_1 + \mathcal{B}, k, \alpha) \geq \alpha k)
        % = \sum_{i = \alpha}^{k} {k \choose i}\bigg(\frac{\mathcal{C}_1 + \mathcal{B}}{\mathcal{N}}\bigg)^i \bigg(1 - \frac{\mathcal{C}_1 + \mathcal{B}}{\mathcal{N}}\bigg)^{k - i}
        \\
        & \gamma_{2}^{t_1} \leftarrow \frac{\mathcal{Z}_2}{\mathcal{C}} P(\mathcal{H}(\mathcal{N}, \mathcal{C}_1, k, \alpha) \geq \alpha k) 
        %= \sum_{i = \alpha}^{k} {k \choose i}\bigg(\frac{\mathcal{C}_1}{\mathcal{N}}\bigg)^i \bigg(1 - \frac{\mathcal{C}_1}{\mathcal{N}}\bigg)^{k - i}
        \\
        & T_1.\text{\texttt{sum}} = T_1.\text{\texttt{sum}} + \gamma_{1}^{t_1} + \gamma_{2}^{t_1}\\
        &\xi_{1}^{t_1} \leftarrow \frac{\mathcal{Z}_1}{\mathcal{C}} P(\mathcal{H}(\mathcal{N}, \mathcal{C}_2, k, \alpha) \geq \alpha k)\\
        & \xi_{2}^{t_1} \leftarrow \frac{\mathcal{Z}_2}{\mathcal{C}} P(\mathcal{H}(\mathcal{N}, \mathcal{C}_2 + \mathcal{B}, k, \alpha) \geq \alpha k)\\
        & T_2.\text{\texttt{sum}} = T_2.\text{\texttt{sum}} + \xi_{1}^{t_1} + \xi_{2}^{t_1}\\
    \end{split}
\end{equation}

In other words, the new expected value of $T_1$.\texttt{sum} after a single time-step is the probability that a correct node in $\mathcal{Z}_1$ is chosen and that node successfully samples at least $\alpha k$ neighbors that also prefer $T_i$ plus the probability that a correct node in $\mathcal{Z}_2$ is chosen and that node successfully samples at least $\alpha k$ neighbors that also prefer $T_i$. In the next time-step, we must include the newly-updated transaction preference sums as part of the computation. For the rest of the time-steps, we compute as follows:

\begin{equation}
    \begin{split}
        t_2:\ \ \ &\gamma_{1}^{t_2} \leftarrow \frac{\mathcal{Z}_1}{\mathcal{C}} P(\mathcal{H}(\mathcal{N}, \mathcal{C}_1 + \mathcal{B} + \gamma_{2}^{t_1} - \xi_{1}^{t_1}, k, \alpha) \\&\geq \alpha k)\\
        & \gamma_{2}^{t_2} \leftarrow \frac{\mathcal{Z}_2}{\mathcal{C}} P(\mathcal{H}(\mathcal{N}, \mathcal{C}_1 + \gamma_{2}^{t_1} - \xi_{1}^{t_1}, k, \alpha) \geq \alpha k)\\
        & T_1.\text{\texttt{sum}} = T_1.\text{\texttt{sum}} + \gamma_{1}^{t_2} + \gamma_{2}^{t_2}\\
        &\xi_{1}^{t_2} \leftarrow \frac{\mathcal{Z}_1}{\mathcal{C}} P(\mathcal{H}(\mathcal{N}, \mathcal{C}_2 + \xi_{1}^{t_1} - \gamma_{2}^{t_1}, k, \alpha) \geq \alpha k)\\
        & \xi_{2}^{t_2} \leftarrow \frac{\mathcal{Z}_2}{\mathcal{C}} P(\mathcal{H}(\mathcal{N}, \mathcal{C}_2 + \mathcal{B} + \xi_{1}^{t_1} - \gamma_{2}^{t_1}, k, \alpha) \geq \alpha k)\\
        & T_2.\text{\texttt{sum}} = T_2.\text{\texttt{sum}} + \xi_{1}^{t_2} + \xi_{2}^{t_2}\\
        t_3:\ \ \ &\gamma_{1}^{t_3} \leftarrow \frac{\mathcal{Z}_1}{\mathcal{C}} P(\mathcal{H}(\mathcal{N}, \mathcal{C}_1 + \mathcal{B} + \gamma_{2}^{t_1} + \gamma_{2}^{t_2} - \xi_{1}^{t_1} - \xi_{1}^{t_2}, k, \alpha) \geq \alpha k)\\
        & \gamma_{2}^{t_3} \leftarrow \frac{\mathcal{Z}_2}{\mathcal{C}} P(\mathcal{H}(\mathcal{N}, \mathcal{C}_1  + \gamma_{2}^{t_1} + \gamma_{2}^{t_2} - \xi_{1}^{t_1} - \xi_{1}^{t_2}, k, \alpha) \geq \alpha k)\\
        & T_1.\text{\texttt{sum}} = T_1.\text{\texttt{sum}} + \gamma_{1}^{t_3} + \gamma_{2}^{t_3}\\
        &\xi_{1}^{t_3} \leftarrow \frac{\mathcal{Z}_1}{\mathcal{C}} P(\mathcal{H}(\mathcal{N}, \mathcal{C}_2 + \xi_{1}^{t_1} + \xi_{1}^{t_2} - \gamma_{2}^{t_1} - \gamma_{2}^{t_2}, k, \alpha) \geq \alpha k)\\
        & \xi_{2}^{t_3} \leftarrow \frac{\mathcal{Z}_2}{\mathcal{C}} P(\mathcal{H}(\mathcal{N}, \mathcal{C}_2 + \mathcal{B} + \xi_{1}^{t_1} + \xi_{1}^{t_2} - \gamma_{2}^{t_1} - \gamma_{2}^{t_2}, k, \alpha) \geq \alpha k)\\
        & T_2.\text{\texttt{sum}} = T_2.\text{\texttt{sum}} + \xi_{1}^{t_3} + \xi_{2}^{t_3}\\
        t_4:\ \ \ & \dots
    \end{split}
\end{equation}

Calculating this complex expression, we arrive at the following results shown in Figures \ref{fig:variousalpha}, \ref{fig:variousk}, and \ref{fig:variousb}.

\begin{figure}[h!]
\centering
\includegraphics[width=\linewidth]{./various_alpha.png}
\caption{Various $\alpha$}
\label{fig:variousalpha}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=\linewidth]{./various_k.png}
\caption{Various $k$}
\label{fig:variousk}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=\linewidth]{./various_b.png}
\caption{Various $\mathcal{B}$}
\label{fig:variousb}
\end{figure}

$<<$TODO: DISCUSS IMPLICATIONS OF THE GRAPH, AND HOW THEY SUPPORT PRIOR CLAIMS.$>>$

At $t_0$, the Byzantine nodes partitioned the correct nodes according to some strategy, and we computed the expected outcomes of the network preferences after all correct nodes had a chance to poll. We now continue computing the evolution of the preferences in the network by introducing children transactions. For simplicity, we first assume that the children transactions all belong to partitions of size one, even if the children transactions are generated by Byzantine nodes.

Due to the invariant imposed by the structure of the dag which states that every transaction on the path to a preferred child must be preferred in turn, every successful solicitation of a transaction implies a successful solicitation of its parent. Moreover, and more importantly, the network support for every immediate child transaction is exactly equal to the probability of the parent being preferred. 
Let $T' \xleftarrow{} T$. At any time $t$, we necessarily have 
\begin{equation}
P(\mathcal{H}_{N, \mathcal{S}_{u}^{t}(T')}^k \geq \alpha k) \geq P(\mathcal{H}_{N, \mathcal{S}_{u}^{t}(T)}^k \geq \alpha k)
\end{equation}
and, more importantly, the 
\begin{equation}
\mathcal{S}_{u}^{t_b}(T) = \mathbb{E}[\mathcal{H}_{N, \mathcal{S}_{u}^{t_a}(T')}^k]/k
\end{equation}
where $t_b$ is larger than $t_a$. 
The expectation of confidence between two conflicting transaction in a permanently split network after $n$ rounds is then
\begin{equation}
    \begin{split}
        \mathbb{E}[c_{u}(T_{i}, n)] &=\ \mathbb{E}[c_{v}(T_{j}, n)]\\
         &=\ nP(\mathcal{H}_{\mathcal{N}, |\mathcal{H}|/2 + |\mathcal{B}|}^k \geq \alpha k)\\
\end{split}
\end{equation}

Before moving on, we can now extract some numerical values. 
Let $|\mathcal{N}| = 10000$, where 25\% of nodes are Byzantine.
Let half of the correct nodes prefer $T_i$ such that $d(T_i) = 1$ and $d(T_j) = 0$, and the other half prefer $T_j$ such that $d(T_i) = 0$ and $d(T_j) = 1$.
Let $u$ be a correct node that prefers $T_i$ and $v$ a correct node that prefers $T_j$.
Let $\alpha k$ = 180, and $k$ = 200. 
Then, 
\begin{equation}
    \begin{split}
    P(\mathcal{H}_{\mathcal{N}, \mathcal{S}_{u}^{t_1}(T_i)}^k \geq \alpha k) &= P(\mathcal{H}_{\mathcal{N}, \mathcal{S}_{v}^{t_1}(T_j)}^k \geq \alpha k)\\
    &= 5.616\times10^{-19}
    \end{split}
\end{equation}
as calculated from equation \ref{eq:hypergeometric}. 
If compounded over all nodes in the network, then the probability that any two nodes from the two preferrence subsets successfully increase confidence in their respective preference is $4.435\times10^{-30}$. 
% Other splits are shown in Figure \ref{fig:hypersplits}. 
%The probability that specifically $u$ and $v$ prefer $T_i$ and $T_j$ for only a preference difference of 1 is $3.154\times10^{-37}$. 
%If we compound over all pairs of partitioned correct nodes, then the probability increases to $4.435\times10^{-30}$. 
Notice that this is for a single round. If we assume a throughput of 1000 solicitations/second, then the probability of success over a second increases to $4.435\times10^{-27}$. Therefore, the expected time before any two correct nodes prefer two equivocating transactions can be calculated through the mean of the geometric distribution and is equal to $2.25\times10^{26}$ seconds, or about $7.14\times10^{18}$ years.

Notice that the calculation above is the optimal probability. Other splits in network result in far lower probabilities and far longer times. Furthermore, notice that the above Byzantine attack assumes that Byzantine nodes can arbitrarily generate transactions \textit{at will}. In reality, simple anti-transaction-flooding measures would severely cripple the Byzantine adversary.

Let's break down the intuition of the results in our analysis. 
The tail bounds indicate that as long as a reasonably large $k$ sample size is picked, the only way for the Byzantine nodes to fool any correct node is to actually build enough network confidence for the transaction. 
The threshold $\alpha$ can be chosen as a function of the Byzantine size, where the higher the Byzantine presence, the larger the threshold should be set.

The consequence of a higher $\alpha$, however, is that it is possible for the Byzantine nodes to deny service to the network. Since liveness is decoupled from safety, we later introduce non-protocol specific mechanisms to thwart such behavior, ensuring that rogue transactions die out quickly and virtuous transactions get accepted quickly.

\subsubsection{Network View}
The analysis up to now assumes that the network view is fully consistent among all nodes. 
In any real open distributed system, such an assumption invariably breaks downs. 
Therefore, we must quantify the amount of tolerance in network view discrepancy between any two correct nodes.

Safety may be violated using a subnetwork partition attack. 
The Byzantine nodes are able to partition the network into several subnetworks such that each subnetwork shares only some set of correct nodes. 
To make this clear, let $\mathcal{E}_{1}, \mathcal{E}_{2} \subset \mathcal{N}$, where $\mathcal{E}_1 \cup \mathcal{E}_2 = \mathcal{N}$, be two subsets of the entire network of nodes, where all members of each subset have -- at minimum -- all other members in the same subset in their view. 
Lastly, let the nodes in the intersection $\mathcal{E}_1 \cap \mathcal{E}_2$ have a view of all nodes in $\mathcal{E}_1$ and $\mathcal{E}_2$. 
We can then view each subnetwork independently of each other, and apply the same analysis as before. The difference, however, is that the network of nodes from the perspective of each subnetwork is not $\mathcal{N}$, but rather just $\mathcal{E}_*$. Therefore, if the Byzantine presence with a fully consistent view was only $\mathcal{B}/\mathcal{N}$, it now becomes $\mathcal{B}/\mathcal{E}_*$. In the worst case scenario, if $|\mathcal{E}_1 \cap \mathcal{E}_2| < (k - \alpha k)$, then safety can be violated without any Byzantine presence. However, in practice, although churn exists, such large network partition is difficult to achieve.

In later sections, we introduce the Snowball network overlay protocol, which guarantees with high probability that the network doesn't suffer from large partitioning.

% \begin{figure}
% \centering
% \includegraphics[width=\linewidth]{./hypersplits.pdf}
% \caption{Probability of single successful sample over different network splits.}
% \label{fig:hypersplits}
% \end{figure}
\end{comment}

### 2020-11-03 paper_fr.tex
add initial source and 'fr' make target

\documentclass[letterpaper,twocolumn,10pt]{article}
\usepackage{usenix2019_v3}
%\documentclass[conference]{IEEEtran}
%\newcommand{\subparagraph}{}
%\usepackage[letterpaper, total={6.5in, 9in}, columnsep=0.25in]{geometry}
%\usepackage[letterpaper, total={7in, 9in}, columnsep=0.33in]{geometry}
%\usepackage[letterpaper, total={7.15in, 9.5in}, columnsep=0.15in]{geometry}
\newcommand{\tronly}[2]{#1}
\usepackage{newtxtext}
\usepackage{appendix}
\tronly{}{%
%\usepackage{microtype}
\usepackage[subtle]{savetrees}
}
\usepackage{comment}
\usepackage{mathtools}
\usepackage{fancyvrb}
\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
\DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}

\setlength{\parskip}{0pt}
\setlength{\abovedisplayskip}{3pt}
\setlength{\belowdisplayskip}{3pt}
\setlength{\textfloatsep}{2pt plus 1.0pt minus 1.0pt}
\setlength{\floatsep}{2pt plus 1.0pt minus 1.0pt}
%\setlength{\textfloatsep}{1pt plus 0.0pt minus 1.0pt}
%\setlength{\floatsep}{1pt plus 0.0pt minus 1.0pt}

\usepackage{cite}
\usepackage{paralist}
%\usepackage[hyphens]{url}
\usepackage[compact]{titlesec}
\usepackage{enumitem}
\setdefaultleftmargin{0em}{}{}{}{}{}
\setdefaultitem{{}}{}{}{}
\setlist[itemize]{leftmargin=*}
\setlist[enumerate]{leftmargin=*}
%\expandafter\def\expandafter\UrlBreaks\expandafter{\UrlBreaks%  save the current one
%\do\a\do\b\do\c\do\d\do\e\do\f\do\g\do\h\do\i\do\j%
%\do\k\do\l\do\m\do\n\do\o\do\p\do\q\do\r\do\s\do\t%
%\do\u\do\v\do\w\do\x\do\y\do\z\do\A\do\B\do\C\do\D%
%\do\E\do\F\do\G\do\H\do\I\do\J\do\K\do\L\do\M\do\N%
%\do\Oh\do\P\do\Q\do\R\do\S\do\T\do\U\do\V\do\W\do\X%
%\do\Y\do\Z}
\usepackage{caption}
% \captionsetup[figure]{font=small,labelfont=small,skip=3pt}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{bbm}
\usepackage{thmtools, thm-restate}
\usepackage{mathrsfs}
\usepackage{amssymb}
\usepackage{tikz}
\usepackage{dsfont}
\usepackage{capt-of}
\usepackage{algorithmicx}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{amsthm}
\usepackage[makeroom]{cancel}
\usepackage{datetime}
\renewcommand\ttdefault{cmtt}
\graphicspath{{./figures}}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}
\newcommand{\Oh}[1]{O(#1)}
\let\oldemptyset\emptyset
\let\emptyset\varnothing
% New definitions
\algnewcommand\algorithmicswitch{\textbf{switch}}
\algnewcommand\algorithmiccase{\textbf{case}}
\algnewcommand\algorithmicassert{\texttt{assert}}
\algnewcommand\Assert[1]{\State \algorithmicassert(#1)}%
% New "environments"
\algdef{SE}[SWITCH]{Switch}{EndSwitch}[1]{\algorithmicswitch\ #1\ \algorithmicdo}{\algorithmicend\ \algorithmicswitch}%
\algdef{SE}[CASE]{Case}{EndCase}[1]{\algorithmiccase\ #1}{\algorithmicend\ \algorithmiccase}%
\algtext*{EndSwitch}%
\algtext*{EndCase}%⇧
\algnewcommand\algorithmiccontinue{\textbf{continue}}
\algnewcommand\algorithmicbreak{\textbf{break}}
\algnewcommand\Continue{\algorithmiccontinue}
\algnewcommand\Break{\algorithmicbreak}
\newcommand{\editremove}[1]{{\color{red}\sout{#1}}}
\newcommand{\editinsert}[1]{{\color{blue}#1}}
\newcommand{\editchange}[1]{{\color{orange}#1}}
\newcommand\ddfrac[2]{\frac{\displaystyle #1}{\displaystyle #2}}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{conjecture}[theorem]{Conjecture}
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\newcommand\blfootnote[1]{%
  \begingroup
  \renewcommand\thefootnote{}\footnote{#1}%
  \addtocounter{footnote}{-1}%
  \endgroup
}

\settimeformat{hhmmsstime}
\mmddyyyydate
\title{
\Large\bf Scalable and Probabilistic Leaderless BFT Consensus through Metastability}
\author{%
\tronly{%
Team~Rocket,
Maofan~Yin,
Kevin~Sekniqi,
Robbert~van~Renesse,
Emin~G\"un~Sirer\\
Cornell University\textsuperscript{*}
}{%
Anonymous Submission 134
}
}
\date{}

\tronly{%
\usepackage[hang]{footmisc}
\setlength\footnotemargin{0em}
}

\begin{document}
\maketitle
\tronly{%
\blfootnote{%
\textsuperscript{*}\emph{Blasts off at the speed of light!} --- Team Rocket\\
An earlier version of this paper published on May 16th 2018, IPFS, was titled \emph{Snowflake to Avalanche: A Novel Metastable Consensus Protocol Family for Cryptocurrencies}.

%Proof of Team Rocket: \verb|SHA256("avalanche ted yin kevin sekniqi robbert van |\\
%\verb|renesse emin gun sirer\n") = 0x 8a5d 2d32 e68b c500 36e4 |\\
%\verb|d086 0446 17fe 4a0a 0296 b274 999b a568 ea92 da46 d533|
}
}{}

\begin{abstract}
This paper introduces a family of leaderless Byzantine fault tolerance protocols, built around a metastable mechanism via network subsampling.
These protocols provide a strong probabilistic safety guarantee in the presence of Byzantine adversaries while their concurrent and leaderless nature enables them to achieve high throughput and scalability.
Unlike blockchains that rely on proof-of-work, they are quiescent and green.
Unlike traditional consensus protocols where one or more nodes typically process linear bits in the number of total nodes per decision, no node processes more than logarithmic bits. It does not require accurate knowledge of all participants and exposes new possible tradeoffs and improvements in safety and liveness for building consensus protocols.
%Finally, unlike traditional consensus protocols and similar to longest-chain protocols, our protocols provide increase in both liveness and safety guarantees when the observed adversarial presence decreases below the maximal expected bound.

%where $k$ is dependent on the security requirements of the system.
% require $O(n^{2})$
% communication, their communication complexity ranges from
% $O(kn\log n)$
% to
% $O(kn)$
% for some security parameter $k \ll n$. \Jon{This sentence led me to assuem that the two ``best'' protocols were both going to be important; that there would be a reason to use the $O(kn\log n)$ version despite its poorer aymptotics. But when I got to later pages, I see you're only discussing Avalanche after all; the other algorithms are purely steps on the pedagogical garden path, right? Maybe remove the reference to the intermediate steps from the abstract?}

The paper describes the Snow protocol family, analyzes its guarantees, and describes how it can be used to construct the core of an internet-scale electronic payment system called Avalanche, which is evaluated in a large scale deployment.
Experiments demonstrate that the system can achieve high throughput (3400 tps), provide low confirmation latency (1.35 sec), and scale well compared to existing systems that deliver similar functionality. For our implementation and setup, the bottleneck of the system is in transaction verification.
\end{abstract}

\section{Introduction}

Achieving agreement among a set of distributed hosts lies at the core of countless applications, ranging from Internet-scale services that serve billions of people~\cite{Burrows06,HuntKJR10} to cryptocurrencies worth billions of dollars~\cite{marketcapcryptocurrency}.
To date, there have been two main families of solutions to this problem.
Traditional consensus protocols %, exemplified by PBFT~\cite{castro1999practical},
rely on all-to-all communication to ensure that all correct nodes reach the same decisions with absolute certainty.
Because they usually require quadratic communication overhead and accurate knowledge of membership, they have been difficult to scale
to large numbers of participants.
On the other hand, Nakamoto consensus protocols~\cite{nakamoto2008bitcoin,GarayKL15, PassSS17, SompolinskyZ15, SompolinskyLZ16, SompolinskyZ18, BentovHMN17, EyalGSR16,Kokoris-KogiasJ16,PassS16a, PassS18} have become popular with the rise of Bitcoin.
These protocols provide a probabilistic safety guarantee: Nakamoto consensus decisions may revert with some probability $\varepsilon$.
A protocol parameter allows this probability to be rendered arbitrarily small, enabling high-value financial systems to be constructed on this foundation.
This family is a natural fit for open, permissionless settings where any node can join the system at any time.
Yet, these protocols are costly, wasteful, and limited in performance.
By construction, they cannot quiesce: their security relies on constant participation by miners, even when there are no decisions to be made.
Bitcoin currently consumes around 63.49 TWh/year~\cite{bitcoinpower}, about twice as all of Denmark~\cite{denmarkpower}.
Moreover, these protocols suffer from an inherent scalability bottleneck that is difficult to overcome through simple reparameterization~\cite{CromanDEGJKMSSS16}. % \Jon{I wanted the paper to expand on this point, but I missed it.}

This paper introduces a new family of consensus protocols called Snow.
Inspired by gossip algorithms, this family gains its properties through a deliberately metastable mechanism.
Specifically, the system operates by repeatedly sampling the network at random, and steering correct nodes towards a common outcome.
Analysis shows that this metastable mechanism is powerful: it can move a large network to an irreversible state quickly, where the irreversibility implies that a sufficiently large portion of the network has accepted a proposal and a conflicting proposal will not be accepted with any higher than negligible ($\varepsilon$) probability. 
%{\color{red} We should clarify this.  The current statement is too strong.}
% though it is not always guaranteed to do so.
% \Jon{Not universal in that some applications require a hard guarantee?}

Similar to Nakamoto consensus, the Snow protocol family provides a probabilistic safety guarantee, using a tunable security parameter that can render the possibility of a consensus failure arbitrarily small.
Unlike Nakamoto consensus, the protocols are green, quiescent and efficient; they do not rely on proof-of-work~\cite{DworkN92} and do not consume energy when there are no decisions to be made.
The efficiency of the protocols stems partly from removing the leader bottleneck: each node requires $\Oh{1}$ communication overhead per round and $\Oh{\log{n}}$ rounds in expectation, whereas classical consensus protocols have one or more nodes that require $\Oh{n}$ communication per round (phase).
Further, the Snow family tolerates discrepancies in knowledge of membership, as we discuss later. In contrast, classical consensus protocols require the full and accurate knowledge of $n$ as its safety foundation.

Snow's subsampled voting mechanism has two additional properties that improve on previous approaches for consensus. 
Whereas the safety of quorum-based approaches breaks down immediately when the predetermined threshold $f$ is exceeded,
Snow's probabilistic safety guarantee degrades smoothly when Byzantine participants exceed $f$.
This makes it easier to pick the critical threshold $f$.
It also exposes new tradeoffs between safety and liveness: the Snow family is more efficient when the fraction of Byzantine nodes is small, and it can be parameterized to tolerate more than a third of the Byzantine
nodes by trading off liveness.

To demonstrate the potential of this protocol family, we illustrate a practical
peer-to-peer payment system, Avalanche. In effect, Avalanche executes multiple Snowball (one from the Snow family) instances with the aid of a Directed Acyclic Graph (DAG). The DAG serves to piggyback multiple instances, reducing the cost from $\Oh{\log{n}}$ to $\Oh{1}$ per node and streamlining the path where there are no conflicting transactions.
%This combination of the best features of traditional and Nakamoto consensus involves one significant tradeoff: liveness for \editchange{equivocating proposals}. %conflicting transactions.
%
%\editchange{Specifically, the consensus protocol from the Snowball family does not guarantee liveness in all cases. It does, however, guarantee logarithmic expected rounds for termination given a propoer initialization heuristic (Section~\ref{sec:liveness}). It also guarantees fast termination when the proposed value is unanimous. As the core part of Avalanche, it guarantees liveness only for \emph{virtuous} transactions, i.e.\ those issued by correct clients and thus guaranteed not to conflict with other transactions.}

Overall, the main contribution of this paper is to introduce a brand new family
of consensus protocols, based on randomized sampling and metastable decision.
The next section provides the model, goals and necessary assumptions for the new protocols.
Section~\ref{sec:protocol} gives intuition behind the protocols, followed by their full specification,
%Section~\ref{sec:attack} briefly discusses typical attack vectors and their possible defense,
Section~\ref{sec:analysis} provides methodology used
by our formal analysis of safety and liveness in Appendix~\ref{sec:full-analysis},
%Section~\ref{sec:implementation} describes a Bitcoin-like payment system,  Section~\ref{sec:evaluation} evaluates the system,
Section~\ref{sec:implementation} describes Avalanche, a Bitcoin-like payment system,
Section~\ref{sec:evaluation} evaluates Avalanche,
Section~\ref{sec:related-work} presents related work, and finally, Section~\ref{sec:conclusions} summarizes our contributions.

\section{Model and Goals}
\label{sec:model_and_goals}
% \editchange{Because our Snowball family provides strong safety guarantee and liveness guarantee},
% it is possible to build other applications involving large-scale probabilistic consensus. We focus on a cryptocurrency application because of many challenges it poses.
%Nevertheless, Snowball does provide liveness guarantee for conflicting values, so it is possible to extend this single-decree protocol for other possible applications involving large-scale probabilistic consensus.
%\Jon{Are there other applications where this consensus model would be valuable? It's not a BFT RSM, but it would be very intriguing if there were other additional applications.}

\tronly{\paragraph{Key Guarantees}}{}

% XXXKEVIN NOT CLEAR THIS PARAGRAPH BELOW BELONGS IN THIS SECTION
% \editinsert{In Avalanche}, we adopt what is commonly known as
% Bitcoin's unspent transaction output (UTXO) model.
% In this model,
% \emph{clients} are authenticated and issue cryptographically signed
% transactions that fully consume an existing UTXO and issue new UTXOs.
% Unlike nodes, clients do not participate in the decision process, but only
% supply transactions to the \emph{nodes} running the consensus protocol.
% Two transactions \emph{conflict} if
% they consume the same UTXO and yield different outputs.
% Correct clients never issue conflicting transactions, nor is it possible
% for Byzantine clients to forge conflicts with transactions
% issued by correct clients.
% However, Byzantine clients can issue multiple transactions that conflict
% with one another, and correct clients can only consume one of
% those transactions.
% \editchange{To make use of Snowball consensus in Avalanche}, then, is to \emph{accept} a
% set of non-conflicting transactions in the presence of Byzantine behavior.
% Each client can be considered a replicated state machine whose
% transitions are defined by a totally ordered list of accepted transactions.

\paragraph{Safety} Unlike classical consensus protocols, and similar to longest-chain-based consensus protocols such as Nakamoto consensus~\cite{nakamoto2008bitcoin}, we adopt an $\varepsilon$-safety guarantee that is probabilistic. 
In practice, this probabilistic guarantee is as strong as traditional safety guarantees, since appropriately small choices of $\varepsilon$ can render consensus failure negligible, lower than the probability of hardware failure due to random events.
Figure~\ref{fig:fandepsilon} shows how the portion ($f/n$) of misbehaving participants (or computation power) affects the probability of system safety failure (decision of two conflicting proposals), given a choice of finality.
% The maximum tolerated Byzantine presence $f$, is dependent on the choice of $\varepsilon$. Similar to Nakamoto consensus, given an observed Byzantine presence $f'$, where $f' \leq f$, the safety failure probability of our protocols maintains the property that $\varepsilon' \leq \varepsilon$. On the other hand, for classical consensus protocols, the failure probability is simply zero for all $f' \leq f$, and almost sure (i.e. one) for $f' > f$.

\begin{figure}[h]
    \includegraphics[width=\linewidth]{figures/safety-f-eps.pdf}
    \caption{Classical BFT protocols that tolerate $f$ failures will encounter total safety failure when the threshold is exceeded even by one additional node. The Bitcoin curve shows a typical finality choice for Bitcoin where a block is considered final when it is ``buried'' in a branch having 6 additional blocks compared to any other competing forks. Snowflake belongs to the Snow family, and it is configured with $k=10$, $\beta=150$. Snowflake-7,8 uses $\alpha=7$ and $\alpha=8$ respectively (see Section~\ref{sec:protocol} for the definition of $k$, $\alpha$ and $\beta$.}
    \label{fig:fandepsilon}
\end{figure}

\tronly{%
\begin{figure}[h]
    \includegraphics[width=\linewidth]{figures/safety-f-eps-log.pdf}
    \caption{Figure~\ref{fig:fandepsilon} with log-scaled y-axis.}
    \label{fig:fandepsilonlog}
\end{figure}
}{}

\paragraph{Liveness} All our protocols provide a non-zero probability guarantee of termination within a bounded amount of time. 
This bounded guarantee is similar to various protocols such as Ben-Or~\cite{ben1983another} and longest-chain protocols.
In particular, for Nakamoto consensus, the number of required blocks for a transaction increases exponentially with the number of adversarial nodes, with an asymptote at $f = n/2$ wherein the number is infinite.
In other words, the time required for finality approaches $\infty$ as $f$ approaches $n/2$\tronly{ (Figure~\ref{fig:livenessproperties}).}{.}
Furthermore, the required number of rounds is calculable ahead of time, as to allow the system designer to tune liveness at the expense of safety. Lastly, unlike traditional consensus protocols and similar to Nakamoto, our protocols benefit from lower adversarial presence, as discussed in property P3 below.

\tronly{%
\begin{figure}[h!]
    \includegraphics[width=\linewidth]{figures/liveness-f-time.pdf}
    \caption{The relation between $f/n$ and the convergence speed, given $\varepsilon = 10^{-20}$. The left figure shows the expected number of blocks to guarantee $\varepsilon$ in Bitcoin, which, counter to commonly accepted folk wisdom, is not a constant $6$, but depends on adversary size to withhold the same $\varepsilon$. The right figure shows the maximum number of rounds required by Snowflake, where being different from Bitcoin, the asymptote is below $0.5$ and varies by the choice of parameters.}
    \label{fig:livenessproperties}
\end{figure}
}

\emph{Formal Guarantees}: Let the system be parameterized for an $\varepsilon$ safety failure probability under a maximum expected $f$ number of adversarial nodes. Let $\Oh{\log n} < t_{max} < \infty$ be the upper bound of the execution of the protocols. The Snow protocols then provide the following guarantees:
\begin{compactitem}
% RVR: this property is unnecessary because the liveness property already ensures non-triviality
%    \item \textbf{Non-triviality}: To ensure that the protocol carries out real work, decisions must include values proposed by external clients;

    \item \textbf{P1. Safety.} When decisions are made by any two correct nodes, they decide on conflicting transactions with negligible probability ($\leq \varepsilon$).
    % \item \textbf{P2. Liveness.} Transactions will be accepted by every correct node in $\Oh{\log n}$ rounds if either the observed Byzantine presence $f'$, where $f' < f$, is sufficiently small, or the network is initialized in a univalent state.
    \item \textbf{P2. Liveness (Upper Bound).} Snow protocols terminate with a strictly positive probability within $t_{max}$ rounds.  
    \item \textbf{P3. Liveness (Strong Form).} If $f \leq \Oh{\sqrt{n}}$, then the Snow protocols terminate with high probability ($\geq 1 - \varepsilon$) in $\Oh{\log{n}}$ rounds. 
\end{compactitem}
% Instead of unconditional agreement, our approach guarantees that safety will be upheld with probability $1-\varepsilon$, where the choice of the security parameter $\varepsilon$ is under the control of the system designer and applications.

\paragraph{Network} 
In the standard definition of asynchrony~\cite{ben1983another}, message transmission time is finite, but the distribution is unspecified (and thus the delivery time can be unbounded for some messages). This implies that the scheduling of message transmission itself could behave arbitrarily, and potentially even maliciously (with full asynchrony).
We use a modified version of this model, which is well-accepted~\cite{banerjee2014epidemic, ganesh2005effect, draief2006thresholds, keeling2011modeling, liggett1997stochastic} in the analysis of epidemic networks and gossip-based stochastic systems. In particular, we fix the distribution of message delay to that of the exponential distribution.
We note that, just like in the standard asynchronous model, there is a strictly non-zero probability that any correct node may execute its next local round only after an arbitrarily large amount of time has passed.
Furthermore, we also note that scheduling only applies to correct nodes, and the adversary may execute arbitrarily, as discussed later.

\paragraph{Achieving Liveness}
Classical consensus that works with asynchrony does not get stuck in a single phase of voting because the vote initiator always polls votes from all known participants and waits for $n - f$ responses.
%The consequence of these two assumptions is therefore that progress cannot be indefinitely delayed by the adversary since the correct nodes (whose number is sufficient for voting) will eventually deliver the required messages.
In our system, however, nodes operate via subsampling, hence it is possible for a single sample to select a majority of adversarial nodes, and therefore the node gets stuck waiting for the responses. To ensure liveness, a node should be able to wait with some timeout. Therefore, our protocols are synchronous in order to guarantee liveness. Lastly, it is worth noting that Nakamoto consensus is synchronous, in which the required difficulty of proof-of-work is dependent on the maximum network delay~\cite{PassSS17}. %Also, other protocols that rely on gossip [Citation here] are also essentially synchronous due to the synchronous nature of gossip protocols. \editinsert{RvR pointed this out, we need to carefully word it though.

\paragraph{Adversary}
The adversarial nodes execute under their own internal scheduler, which is unbounded in speed, meaning that all adversarial nodes can execute at any infinitesimally small point in time, unlike correct nodes. 
The adversary can view the state of every honest node at all times and can instantly modify the state of all adversarial nodes. 
It cannot, however, schedule or modify communication between correct nodes.
Finally, we make zero assumptions about the behavior of the adversary, meaning that it can choose any execution strategy of its liking.
In short, the adversary is computationally bounded (it cannot forge digital signatures) but otherwise is point-to-point informationally unbounded (knows all state) and round-adaptive (can modify its strategy at any time).

\paragraph{Sybil Attacks}
Consensus protocols provide their guarantees based on assumptions that only a fraction of participants are adversarial.
These bounds could be violated if the network is naively left open to arbitrary participants.
In particular, a Sybil attack~\cite{douceur2002sybil}, wherein a large number of identities are generated by an adversary, could be used to exceed the bounds.

A long line of work, including PBFT~\cite{castro1999practical}, treats the Sybil problem separately from consensus, and rightfully so, as Sybil control mechanisms are distinct from the underlying, more complex agreement protocol\footnote{This is not to imply that every consensus protocol can be coupled/decoupled with every Sybil control mechanism.}.
In fact, to our knowledge, only Nakamoto-style consensus has ``baked-in'' Sybil prevention as part of its consensus, made possible by chained proof-of-work~\cite{aspnes2005exposing}, which requires miners to continuously stake a hardware investment.
Other protocols, discussed in Section~\ref{sec:related-work}, rely on proof-of-stake (by economic argument) or proof-of-authority (by administrative argument that makes the system ``permissioned'').
The consensus protocols presented in this paper can adopt any
Sybil control mechanism, although proof-of-stake is most aligned with their quiescent operation.
One can use an already established proof-of-stake based mechanism~\cite{GiladHMVZ17}.
The full deployment of an autonomous P2P payment system incorporating staking mechanism is beyond the scope of this paper, whose focus is on a novel design paradigm of the core consensus algorithm.

% Sybil control mechanisms are orthogonal and separate from the consensus protocols. All such protocols, excluding Nakamoto-style consensus protocols, require a mechanism for network identity establishment. In a real implementation of Avalanche, the Sybil problem would be solved through a staking mechanism.

\paragraph{Flooding Attacks}
Flooding/spam attacks are a problem for any distributed system. 
Without a protection mechanism, an attacker can generate large numbers of transactions and flood protocol data structures, consuming storage.
% Although there is no proper, protocol-level methodology to prevent such an attack, there are very effective heuristics that we can employ.
There are a multitude of techniques to deter such attacks, including network-layer protection, proof-of-authority, local proof-of-work and economic mechanisms. 
In Avalanche, we use transaction fees, making such attacks costly even if the attacker is sending money back to addresses under its control.

\paragraph{Additional Assumptions}
We do not assume that all members of the network are known to all participants, but rather may temporarily have some discrepancies in network view.
We quantify the bounds on the discrepancy in Appendix~\ref{sec:full-analysis-churn}.
We assume a safe bootstrapping mechanism, similar to that of Bitcoin, that enables a node to connect with sufficiently many correct nodes to acquire a statistically unbiased view of the network.
We do not assume a PKI\@.
Finally, we make standard cryptographic assumptions related to digital signatures and hash functions.
%\editinsert{TODO: Talk about synchrony assumption, and justify it by mentioning Bitcoin.}
%\editinsert{TODO: Talk about we don't address Sybil attack, and justify it by mentioning Algorand (we also use Algorand as a comparison in Evaluation, so it will totally make sense here.}

\section{Protocol Design}\label{sec:protocol}
We start with a non-BFT protocol called Slush and progressively build up to Snowflake and Snowball, all based on the same common majority-based metastable voting mechanism.
These protocols are single-decree consensus protocols of increasing robustness.
We provide full specifications for the protocols in this section, and defer the analysis to the next section, and present formal proofs in the appendix.

\subsection{Slush: Introducing Metastability}
The core of our approach is a single-decree consensus protocol,
inspired by epidemic or gossip protocols.
% In this approach, a state in which all possible outcomes are equally
% probable is metastable and disturbances are likely to cause the
% protocol to converge to one of multiple stable states.
% We will call such consensus protocols metastable themselves.
%We can model a preference for a decision as a kind of infection: a node catches a disease, and converts other nodes to carry the same disease in epidemic fashion.
%If a threshold of its interactions are with other infected hosts, it flips its color to the competing agent and joins their ranks.
%
The simplest protocol, Slush, is the foundation of this family, shown in Figure~\ref{fig:slush-loop}.
Slush is \emph{not} tolerant to Byzantine faults, only crash-faults (CFT), but serves as an illustration for the BFT protocols that follow.
For ease of exposition, we will describe the operation of Slush using a decision between two conflicting colors, red and blue.
% , though the technique generalizes to any number of competing proposed values.
% Slush operates as follows.

In Slush, a node starts out initially in an uncolored state.
Upon receiving a transaction from a client, an uncolored node updates its own color to the one carried in the transaction and initiates a query.
To perform a query, a node picks a small, constant sized ($k$) sample of the network uniformly at random, and sends a query message.
Upon receiving a query, an uncolored node adopts the color in the query, responds with that color, and initiates its own query, whereas a colored node simply responds with its current color.
% If $k$ responses are not received within a time bound, the node picks an additional sample from the remaining nodes uniformly at random and queries them until it collects all responses.
Once the querying node collects $k$ responses, it checks if a fraction $\ge\alpha$ are for the same color, where $\alpha > \floor{k/2}$ is a protocol parameter.
If the $\alpha$ threshold is met and the sampled color differs from the node's own color, the node flips to that color.
It then goes back to the query step, and initiates a subsequent round of query, for a total of $m$ rounds.
Finally, the node decides the color it ended up with at time $m$.

\algnewcommand{\IIf}[1]{\State\algorithmicif\ #1\ \algorithmicthen}
\algnewcommand{\EndIIf}{\unskip\ \algorithmicend\ \algorithmicif}

\newcommand{\codecolor}{\mathit{col}}
\newcommand{\codecount}{\mathit{cnt}}
\newcommand{\codelastcol}{\mathit{lastcol}}
\newcommand{\assign}{\coloneqq}
\begin{figure}
    \small
\begin{algorithmic}[1]
    \Procedure{onQuery}{$v, \codecolor'$}
        \IIf{$\codecolor = \bot$} $\codecolor \assign \codecolor'$
        \State\Call{respond}{$v, \codecolor$}
    \EndProcedure
    \Procedure{slushLoop}{$u, \codecolor_0 \in \{\texttt{R}, \texttt{B}, \bot\}$}
        \State $\codecolor \assign \codecolor_0$ \textrm{// initialize with a color}
        \For{$r \in \{1\ldots m\}$}
            \State \textrm{// if $\bot$, skip until \textsc{onQuery} sets the color}
            \If{$\codecolor = \bot$}
                \Continue
            \EndIf
            \State \textrm{// randomly sample from the known nodes}
            \State $\mathcal{K} \assign \Call{sample}{\mathcal{N}\backslash u, k}$
            \State $P \assign \texttt{[}\Call{query}{v, \codecolor}\quad\textbf{for}\ v \in \mathcal{K}\texttt{]}$
            \For{$\codecolor' \in \{\texttt{R}, \texttt{B}\}$}
                \If{$P.\Call{count}{\codecolor'} \ge \alpha$}
                    \State $\codecolor \assign \codecolor'$
                \EndIf
            \EndFor
        \EndFor
        \State \Call{accept}{$\codecolor$}
    \EndProcedure
    \captionof{figure}{Slush protocol. Timeouts elided for readability.}\label{fig:slush-loop}
\end{algorithmic}
\end{figure}

Slush has a few properties of interest. 
First, it is almost \emph{memoryless}: a node retains no state between rounds other than its current color, and in particular maintains no history of interactions with other peers.
Second, unlike traditional consensus protocols that query every participant, every round involves sampling just a small, constant-sized slice of the network at random.
Third, Slush makes progress under any network configuration (even fully bivalent state, i.e. 50/50 split between colors), since random perturbations in sampling will cause one color to gain a slight edge and repeated samplings afterwards will build upon and amplify that imbalance.
Finally, if $m$ is chosen high enough, Slush ensures that all nodes will be colored identically with high probability (whp).
Each node has a constant, predictable communication overhead per round, and $m$ grows logarithmically with $n$.

\tronly{
The Slush protocol does not provide a strong safety guarantee in the presence of Byzantine nodes.
In particular, if the correct nodes develop a preference for one color, a Byzantine adversary can attempt to flip nodes
to the opposite so as to keep the network in balance, preventing a decision.
We address this in our first BFT protocol that introduces more state storage at the nodes.
}{
We next examine how to extend Slush to tolerate Byzantine behavior.
}

\subsection{Snowflake: BFT}

Snowflake augments Slush with a single counter that captures the strength of a node's conviction in its current color.
This per-node counter stores how many consecutive samples of the network by that node have all yielded the same color.
A node accepts the current color when its counter reaches $\beta$, another security parameter.
Figure~\ref{fig:snowflake-loop} shows the amended protocol, which includes
the following modifications:

\begin{compactenum}
	\item Each node maintains a counter $\mathit{cnt}$;
    \item Upon every color change, the node resets $\mathit{cnt}$ to 0;
    \item Upon every successful query that yields $\ge \alpha$ responses for the same color as the node, the node increments $\mathit{cnt}$.
\end{compactenum}

\newcommand{\codemaj}{\mathit{maj}}
\newcommand{\codefalse}{\texttt{false}}
\newcommand{\codetrue}{\texttt{true}}
\begin{figure}
    \small
\begin{algorithmic}[1]
    \Procedure{snowflakeLoop}{$u, \codecolor_0 \in \{\texttt{R}, \texttt{B}, \bot\}$}
        \State $\codecolor \assign \codecolor_0$, $\codecount \assign 0$
        \While{\textrm{undecided}}
            \If{$\codecolor = \bot$}
                \Continue
            \EndIf
            \State $\mathcal{K} \assign \Call{sample}{\mathcal{N}\backslash u, k}$
            \State $P \assign \texttt{[}\Call{query}{v, \codecolor}\quad\textbf{for}\ v \in \mathcal{K}\texttt{]}$
            \State $\codemaj \assign \codefalse$
            \For{$\codecolor' \in \{\texttt{R}, \texttt{B}\}$}
            \If{$P.\Call{count}{\codecolor'} \ge \alpha$}
                \State $\codemaj \assign \codetrue$
                \If{$\codecolor' \neq \codecolor$}
                    \State $\codecolor \assign \codecolor'$, $\codecount \assign 1$
                \Else\hspace{1ex}$\codecount\texttt{++}$
                \EndIf
                \IIf {$\codecount \ge \beta$} \Call{accept}{$\codecolor'$}
            \EndIf
            \EndFor
            \IIf{$\codemaj = \codefalse$} $\codecount \assign 0$
        \EndWhile
    \EndProcedure
    \captionof{figure}{Snowflake.}\label{fig:snowflake-loop}
\end{algorithmic}
\end{figure}

\begin{figure}[t]
    \small
\begin{algorithmic}[1]
    \Procedure{snowballLoop}{$u, \codecolor_0 \in \{\texttt{R}, \texttt{B}, \bot\}$}
        \State $\codecolor \assign \codecolor_0$, $\codelastcol \assign \codecolor_0$, $\codecount \assign 0$
        \State $d[\texttt{R}] \assign 0$, $d[\texttt{B}] \assign 0$
        \While{\textrm{undecided}}
            \IIf{$\codecolor = \bot$} \Continue
            \State $\mathcal{K} \assign \Call{sample}{\mathcal{N}\backslash u, k}$
            \State $P \assign \texttt{[}\Call{query}{v, \codecolor}\quad\textbf{for}\ v \in \mathcal{K}\texttt{]}$
            \State $\codemaj \assign \codefalse$
            \For{$\codecolor' \in \{\texttt{R}, \texttt{B}\}$}
            \If{$P.\Call{count}{\codecolor'} \ge \alpha$}
                \State $\codemaj \assign \codetrue$
                \State $d[\codecolor']\texttt{++}$
                \If {$d[\codecolor'] > d[\codecolor]$}
                        \State$\codecolor \assign \codecolor'$
                \EndIf
                \If{$\codecolor' \neq \codelastcol$}
                    \State$\codelastcol \assign \codecolor'$, $\codecount \assign 1$
                \Else\hspace{1ex}$\codecount\texttt{++}$
                \EndIf
                \IIf {$\codecount \ge \beta$} \Call{accept}{$\codecolor'$}
            \EndIf
            \EndFor
            \IIf{$\codemaj = \codefalse$} $\codecount \assign 0$
        \EndWhile
    \EndProcedure
    \captionof{figure}{Snowball.}\label{fig:snowball-loop}
\end{algorithmic}
\end{figure}

When the protocol is correctly parameterized for a given threshold of Byzantine nodes and a desired $\varepsilon$-guarantee, it can ensure both safety (P1) and liveness (P2, P3).
As we later show, there exists an irreversible state after which a decision is inevitable. Correct nodes begin to commit past the irreversible state to adopt the same color, whp. For additional intuition, which we do not expand in this paper, there also exists a phase-shift point, where the Byzantine nodes lose ability to keep network in a bivalent state.
%there exists a phase-shift point after which correct nodes are more likely to tend towards a decision than a bivalent state.
%Further, there exists a point-of-no-return after which a decision is inevitable.
%The Byzantine nodes lose control past the phase shift, and the correct nodes begin to commit past the point-of-no-return, to adopt the same color, whp.

\subsection{Snowball: Adding Confidence}

Snowflake's notion of state is ephemeral: the counter gets reset with every color flip.
% While, theoretically, the protocol is able to make strong guarantees with
% minimal state, we will now improve the protocol to make it harder to
% attack by incorporating a more permanent notion of belief.
Snowball augments Snowflake with \emph{confidence counters} that capture the number of queries that have yielded a threshold result for their corresponding color (Figure ~\ref{fig:snowball-loop}).
A node decides if it gets $\beta$ consecutive chits for a color. However, it only changes preference based on the total accrued confidence.
The differences between Snowflake and Snowball are as follows:
\begin{compactenum}
    \item Upon every successful query, the node increments its confidence counter for that color.
    \item A node switches colors when the confidence in its current color becomes lower than the confidence value of the new color.
\end{compactenum}
% Snowball is not only harder to attack than Snowflake, but is more easily generalized to multi-decree protocols.

\section{Analysis}
\label{sec:analysis}
Due to space limits, we move some core details to Appendix~\ref{sec:full-analysis}, where we show that under certain independent and distinct assumptions, the Snow family of consensus protocols provide safety (P1) and liveness (P2, P3) properties.
In this section, we summarize our core results and provide some proof sketches.

\paragraph{Notation} Let the network consist of a set of $n$ nodes (represented by set $\mathcal{N}$), where $c$ are correct nodes (represented by set $\mathcal{C}$) and $f$ are Byzantine nodes (represented by set $\mathcal{B}$). 
Let $u, v \in \mathcal{C}$ refer to any two correct nodes in the network. Let $k, \alpha, \beta \in \mathbb{Z}^+$ be positive integers where $\alpha > \floor{k/2}$. From now on, $k$ will always refer to the network sample size, where $k \leq n$, and $\alpha$ will be the majority threshold required to consider the voting experiment a ``success''. In general, we will refer to $\mathcal{S}$ as the state (or configuration) of the network at any given time.

\paragraph{Modelling Framework} To formally model our protocols, we use continuous-time Markov processes (CTMC). 
The state space is enumerable (and finite), and state transitions occur in continuous time. 
CTMCs naturally model our protocols since state transitions do not occur in epochs and in lockstep for every node (at the end of every time unit) but rather occur at any time and independently of each other. 
% CMTCs can also trivially encapsulate synchronous time protocols through the underlying jump chain, whose probabilities of transitions can be extracted from the underlying generator. 
% This modelling framework, in fact, is the primary way by which stochastic processes that describe systems of interacting particles (epidemic networks, biology, physics, etc) are analyzed. 
% Analyzing the protocols introduced in this paper
%differ from classical as well as Nakamoto consensus-style protocols. 
%Even simple, non-linear dynamical systems often incur complicated analyses as they may not admit analytical solutions.
% In particular, since the state of a node depends on randomly sampling other nodes, possibly inducing nondeterministic dependency cycles, analyzing all possible state executions is computationally infeasible. 
% Here, we employ stochastic techniques used in the analysis of epidemic networks, which may also be of independent interest to the distributed systems community. 
%Here, we employ 
% requires complex stochastic techniques to reason about the state of the network in aggregate.

We focus on binary consensus, although the safety results generalize to more than two values. We can think of the network as a set of nodes either colored red or blue, and we will refer to this configuration at time $t$ as $\mathcal{S}_t$. 
We model our protocols through a continuous-time process with two absorbing states, where either all nodes are red or all nodes are blue. 
The state space $\mathcal{S}$ of the stochastic process is a condensed version of the full configuration space, where each state $\{0, \dots, n\}$ represents the total number of blue nodes in the system. 
% At any time, a set of nodes either change their color, or increase their confidence in their latest color. 
% The dynamics of the system resemble those of epidemic networks, where nodes update their state based on the state of their neighbors, using a function chosen over some probability distribution. 
% It would be infeasible to keep track of all possible execution paths, since at every round, the possible number of branching executions is equal to the number of possible correct nodes times all possible $k$-sample choices of outcomes that the node may sample. Furthermore, the difficulty of this task is greatly amplified due to the Byzantine adversary, whose optimal attack strategy is chosen over an unknown function, possibly itself nondeterministic.

The simplification that allows us to analyze this system is to obviate the need to keep track of all of the execution paths, as well as all possible adversarial strategies, and rather focus entirely on a single state of interest, without regards to how we achieve this state. 
More specifically, the core extractable insight of our analysis is in identifying the \textit{irreversibility} state of the system, the state upon which so many correct nodes have usurped either red or blue that reverting back to the minority color is highly unlikely.

\subsection{Safety}

\paragraph{Slush} 
We assume that
%$\mathcal{L}(u) = \mathcal{N}$ for all $u \in \mathcal{N}$.
all nodes share the same $\mathcal{N}$, and in
Appendix~\ref{sec:full-analysis-churn}, we sketch how to relax the requirement of the membership knowledge.
We model the dynamics of the system through a continuous-time process where two states are absorbing, namely the all-red or all-blue state. Let $\{X_{t \geq 0}\}$ be the random variable that describes the state of the system at time $t$, where $X_0 \in \{0, \dots, c\}$.
We begin by immediately discussing the most important result of the safety dynamics of our processes: the \emph{reversibility} probabilities of the \textbf{Slush} process. All the other formal results in this paper are, informally speaking, intuitive derivations and augmentations of this result.

\begin{theorem}
Let the configuration of the system at time $t$ be $\mathcal{S}_t = n/2 + \delta$, meaning that the network has drifted to $2\delta$ more blue nodes than red nodes ($\delta = 0$ means that red and blue are equal). Let $\xi_\delta$ be the probability of absorption to the all-red state (minority). Then, for all $0 \leq \delta \leq n/2$, we have 
\begin{equation}
\begin{split}
    \xi_\delta &\leq \left(\dfrac{1/2 - \delta/n}{\alpha/k}\right)^{\alpha}\left(\dfrac{1/2 + \delta/n}{1- \alpha/k}\right)^{k-\alpha}\\
    &\leq e^{-2((\alpha/k) - (1/2) + (\delta/n))^2 k}
\end{split}
\end{equation}
\end{theorem}

\begin{proof}
This bound follows from the Hoeffding-derived tail bounds of the hypergeometric distribution by Chvatal~\cite{chvatal1979tail}. 
\end{proof}

We note that Chvatal's bounds are introduced for simplicity of exposition and are extremely weak.  
We leave the full closed-form expression in Theorem~\ref{theorem:slush_prob_convergence_minority} to the appendix, which is also significantly stronger than the Chvatal bound. 
Nonetheless, using the loose Chvatal bound, we make the key observation that as the drift $\delta$ increases, given fixed $\alpha$ and $k$, the probability of moving towards the minority value decreases \emph{exponentially fast} (in fact, even faster, since there is a quadratic term in the inverse exponent). Additionally, the same result holds for increasing $\alpha$ given a fixed $k$.

The outcomes of this theorem demonstrate a key property: once the network loses full bivalency (i.e. $\delta > 0$), it tends to topple and converge rapidly towards the majority color, unable to revert back to the minority with significant probability. This is the fundamental property exploited by our protocols, and what makes them secure despite only sampling a small, constant-sized set of the network. The core result that follows for the safety guarantees in Snowflake is in finding regions (given specific parameter choices) where the reversibility holds with no higher than $\varepsilon$ probability even under adversarial presence.

\paragraph{Snowflake} 
For Snowflake, we assume that some fraction of nodes are adversarial. In Slush, once the network gains significant majority support for one proposal (e.g., the color blue), it becomes unlikely for a minority proposal (e.g., the color red) to ever become decided in the future (irreversibility). Furthermore, in Slush nodes simply have to execute the protocol for a deterministic number of rounds, $m$, which is known ahead of protocol execution. When introducing adversarial nodes with arbitrary strategies, however, nodes cannot simply execute the protocol for a deterministic number of rounds, since the adversary may nondeterministically affect the value of $m$. Instead, correct nodes must implement a mechanism to explicitly detect that irreversibility has been reached. To that end, in Snowflake, every correct node implements a decision function, $\mathcal{D}(u, \mathcal{S}_t, blue) \rightarrow \{0, 1\}$, which is a random variable that outputs $1$ if node $u$ detects that the network has reached an irreversibility state at time $t$ for blue. The decision mechanism is probabilistic, meaning that it can fail, although it is designed to do so with negligible probability. We now sketch the proof of Snowflake.

\noindent \emph{Proof Sketch}. We define safety failure to be the event wherein any two correct nodes $u$ and $v$ decide on blue and red, i.e. $\mathcal{D}(u, \mathcal{S}_t, blue) \rightarrow 1$ and $\mathcal{D}(v, \mathcal{S}_{t'}, red) \rightarrow 1$, for any two times $t$ and $t'$. We again model the system as a continuous time random process. The state space is defined the same way as in Slush. However, we note some critical subtleties. First, even if all correct nodes accept a color, the Byzantine nodes may revert. Second, we also have to consider the decision mechanism $\mathcal{D}(*)$. To analyze, we obviate the need to keep track of all network configurations under all adversarial strategies and assume that a node $u$ first decides on blue. Then, conditioned on the state of the network upon $u$ deciding, we calculate the probability that another node $v$ decides red, which is a function of both the probability that the network reverts towards a minority blue state and that $v$ decides at that state. 
We show that under appropriate choices of $k$, $\alpha$, and $\beta$, we can construct highly secure instances of Snowflake (i.e. safety failure with probability $\leq \varepsilon$) when the network reaches some bias of $\delta$, as shown in Figure~\ref{fig:states_feasible_solutions}. A concrete example is provided in Figure~\ref{fig:fandepsilon}.

\begin{figure}[h]
\begin{center}
\input{figures/analysis2}
\caption{Representation of the irreversibility state, which exists when -- even under $f$ Byzantine nodes -- the number of blue correct nodes exceeds that of red correct nodes by more than $2\delta$.
}
\label{fig:states_feasible_solutions}
\end{center}
\end{figure}

% An important insight is that there exists an irreversible state, or \textit{point of no return}, after which the system will converge to an absorbing state whp. Furthermore a correct node only decides when the system is beyond the point of no return. Composing these two guarantees together, the probability of a safety violation is strictly less than $\varepsilon$, which can be configured as desired. 
% Unsurprisingly, there is an inherent tension between safety and liveness, but suitable parameters can be found that are practical.
% Larger values of $k$ obtain higher levels of security for correct nodes, at the expense of slower convergence.

\paragraph{Snowball}
Snowball is an improvement in security over Snowflake, where random perturbations in network samples are reduced by introducing a limited form of history, which we refer to as confidence. 
% Modeling Snowball with a Markov chain is difficult because of a state space explosion problem. In particular, it is not sufficient to simply keep track of color preferences of each correct node, the analysis must also maintain information about their confidence.

\noindent \emph{Proof Sketch}. We apply martingale concentration inequalities to prove that once the system has reached the irreversibility state, then the growth of the confidence of the majority decided color will perpetually grow and drift further away from those of the minority color, effectively rendering reversibility less likely over time. If the drifts ever revert, then reversibility analysis becomes identical to that of Snowflake.

\subsection{Liveness}

We assume that the observed adversarial presence $0 \leq f' \leq n(k - \alpha - \psi)/k \leq f$, where we refer to $\psi$ as the buffer zone. 
The bigger $\psi$, the quicker the ability of the decision mechanism to finalize a value. 
If, of course, $\psi$ approaches zero or becomes negative, then we violate the upper bound of adversarial tolerance for the parameterized system, and thus the adversary can, with high probability, stall termination by simply choosing to not respond, although the safety guarantees may still hold.

Assuming that $\psi$ is strictly positive, termination is strictly finite under all network configurations where a proposal has at least $\alpha$ support. Furthermore, not only is termination finite with probability one, we also have a strictly positive probability of termination within any bounded amount of time $t_{max}$, as discussed in Lemma~\ref{lemma:finitetermination}, which follows from Theorem~\ref{theorem:mean-convergence-time}. This captures liveness property P2.

\noindent\emph{Proof Sketch.} Using the construction of the system to prove irreversibility, we characterize the distribution of the average time spent (sojourn times) at each state before the system terminates execution by absorption at either absorbing state. The termination time is then a union of these times.

For non-conflicting transactions, since the adversary is unable to forge a conflict, the time to decision is simply the mixing time , which is $\Oh{\log{n}}$.
Liveness guarantees under a fully bivalent network configuration reduce to an optimal convergence time of $\Oh{\log{n}}$ rounds if the adversary is at most $\Oh{\sqrt{n}}$, for $\alpha = \floor{k/2} + 1$. We leave additional detains to Lemma~\ref{lemma:centrallimit}.
% This result is independently supported by Ben-Or~\cite{ben1983another} and Doerr et al~\cite{doerr2011stabilizing}. 
When the adversary surpasses $\Oh{\sqrt{n}}$ nodes, the worst-case number of rounds increases polynomially, and as $f$ approaches $n/2$ it approaches exponential convergence rates.

\noindent\emph{Proof Sketch.} We modify Theorem~\ref{theorem:mean-convergence-time} to include the adversary, which reverts any imbalances in the network by keeping network fully bivalent.

\paragraph{Multi-Value Consensus}
Our binary consensus protocol could support multi-value consensus by running logarithmic binary instances, one for each bit of the proposed value. However, such theoretical reduction might not be efficient in practice. Instead, we could directly incorporate multi-values as multi-colors in the protocol, where safety analysis could still be generalized.

For liveness, we sketch a leaderless initialization mechanism, which in expectation uses $\Oh{\log{n}}$ rounds under the assumption that the network is synchronized in the Appendix~\ref{sec:sync-heuristic}. 
While the design of initialization mechanisms is interesting, note that it is not necessary for a decentralized payment system, as we show in Section~\ref{sec:implementation}.
%We leave additional research into the initialization mechanisms to future work. 
%However, we finally note that when building a decentralized payment system, an initialization mechanism is not necessary, as we show in the implementation of Avalanche in Section~\ref{sec:implementation}.
Finally, in the Appendix~\ref{sec:full-analysis-churn}, we discuss churn and view discrepancies.

% \subsection{Additional Insights}
% Two interesting insights arise from the way our protocols are designed and used.
% First, the protocols discussed in this work lead to both safety and liveness guarantees whose underlying function is smooth, rather than a step function. 
% In many other consensus protocols, safety is guaranteed with up to a fixed threshold number (e.g.~1/3) of adversarial nodes, beyond which no guarantees are provided. In our protocols, the guarantees degrade gracefully with an adversarial percentage beyond the pre-established bound. For example, optimal system parameters can be chosen to tolerate precisely 1/5 adversarial presence with failure probability $\varepsilon$. However, if the system faces an adversarial presence greater than 1/5, then the probability of failure degrades to slightly above $\varepsilon$, rather than immediately to $1$. 
% Second, these protocols externalize the safety and liveness tradeoffs. 
% The system designer may choose to guarantee safety even under catastrophic events, such as an adversarial presence beyond $1/2$, at the expense of liveness.  
% This presents a powerful knob not available in classical or Nakamoto-based consensus protocols.

% \begin{compactitem}
% \item\textbf{Eventually good ancestry.} Virtuous transactions can be retried by picking new parents, selected from a set that is more likely to be preferred.
% Ultimately, one can always attach a transaction to decided parents to completely mitigate this risk.
% \Jon{So why don't we always do this? Later the paper clarifies that moving up the chain reduces the efficiency
%     and increases latency for the proposed transaction. I got pretty lost in the last two paragraphs of section 4}
% A simple technique for parent selection is to select new parents for a virtuous transaction at successively lower heights in the DAG, proceeding towards the genesis vertex.
% This procedure is guaranteed to terminate with uncontested, decided parents, ensuring that the transaction cannot suffer liveness failure due to rogue transactions.

% \item\textbf{Sufficient chits.} A secondary mechanism is necessary to ensure that virtuous transactions with decided ancestry will receive sufficient chits.
% To ensure this, correct nodes examine the DAG for virtuous non-nop transactions that lack sufficient progeny and emit nop transactions to help increase their confidence.
% A nop transaction has just one parent and no application side-effects, and can be issued by any node. They cannot be abused by Byzantine nodes because, even though nops trigger new queries, they do not automatically grant chits.

% With these two mechanisms in place, it is easy to see that, at worst, Avalanche will degenerate into separate instances of Snowball, and thus provide the same liveness guarantee for virtuous transactions.
% \end{compactitem}

%XXX discuss how no-ops can be implemented

% \subsection{Communication Complexity}\tronly{}{\vspace{-0.5em}}
% Since liveness is not guaranteed for rogue transactions, we focus our message complexity analysis solely for the case of virtuous transactions.
% For the case of virtuous transactions, Snowflake and Snowball are both guaranteed to terminate after $O(kn\log n)$ messages.
% This follows from the well-known results related to epidemic algorithms\cite{demers1987epidemic}, and is confirmed by Table~\ref{table:growth_worst_case} in Appendix~\ref{sec:full-analysis}.

\section{Peer-to-Peer Payment System}
\label{sec:implementation}

%\label{sec:evaluation}

Using Snowball consensus, we have implemented a bare-bones payment system, Avalanche, which supports Bitcoin transactions. In this section, we describe the design and sketch how the implementation can support the value transfer primitive at the center of cryptocurrencies.
Deploying a full cryptocurrency involves bootstrapping, minting, staking, unstaking,
and inflation control. While we have solutions for these issues, their full
discussion is beyond the scope of this paper, whose focus is centered on the
novel Snow consensus protocol family.
%In this section,
%we focus on how Avalanche can support the value transfer primitive at the center of cryptocurrencies.
%\paragraph{XXX: the original implementation paragraph in Model}

%The Snow family is presented in the form of single-decree, binary consensus for its own theoretical value and simplicity.
%On the other hand, we show later in Section~\ref{sec:implementation} our practical payment system, Avalanche, utilizes multiple Snowball instances to support standard Bitcoin-like transactions. %, by humbly cheating in two ways.
%First, instead of a single replicated state machine (RSM) model, where the system determines a sequence of totally-ordered transactions $T_0, T_1, T_2, \ldots$ issued by any client, we adopt a \emph{parallel consensus model}, where each client interacts independently with its own RSMs and optionally transfers ownership of its RSM to another client. The system establishes only a partial order between dependent transactions.
%Second,
In a cryptocurrency setting, cryptographic signatures enforce that only a key owner is able to create a transaction that spends a particular coin. Since correct clients follow the protocol as prescribed and never double spend coins, in Avalanche, they are guaranteed both safety and liveness for their \emph{virtuous} transactions. In contrast, liveness is not guaranteed for \emph{rogue} transactions, submitted by Byzantine clients, which conflict with one another. Such decisions may stall in the network, but have no safety impact on virtuous transactions.
We show that this is a sensible tradeoff, and that the resulting system is sufficient for building complex payment systems.

\input{avalanche}

\subsection{Multi-Input UTXO Transactions}
In addition to the DAG structure in Avalanche, an \emph{unspent transaction output} (UTXO)~\cite{nakamoto2008bitcoin} graph that captures
spending dependency is used to realize the ledger for the payment system. To
avoid ambiguity, we denote the transactions that encode the data for money
transfer \emph{transactions}, while we call the
transactions ($T \in \mathcal{T}$) in Avalanche's DAG \emph{vertices}.

%\tronly
%{
%Each transaction represents a money transfer that takes several inputs from
%source accounts and generates several outputs to destinations. As a UTXO-based system
%that keeps a decentralized ledger, balances are kept by the \emph{unspent outputs}
%of transactions.
%
%More specifically, a transaction $\texttt{TX}_a$ maintains
%a list of inputs: $\texttt{In}_{a1}$, $\texttt{In}_{a2}$, $\cdots$. Each input
%has two fields: the reference to an unspent transaction output and a spend script.
%The unspent transaction output uniquely refers to an output of a previously made transaction.
%The script snippet will be prepended to the script from the referred output forming a complete computation.
%It could typically be a cryptographic proof of validity, but could also be any Turing-complete
%computation in general.
%Each output $\texttt{Out}_{a1}$, $\texttt{Out}_{a2}$, $\cdots$,
%contains some amount of money and a script which typically contains cryptographic verification that takes the proof from the future input and verifies validity.
%
%In our payment system, there are \emph{addresses} representing different
%accounts by cryptographic keys. The public
%key is used as the identity for recipients in the output scripts, while
%the private key is for creating signatures in the
%input scripts, spending the available funds. Only the key owner is able spend the unspent output by creating an input with
%the signature in a new transaction.
%
%Due to the possibility of double spending by the private key owner,
%cryptocurrencies such as Bitcoin use a blockchain as the linear log to reject the
%transaction that comes later in the log and tries to spend some output twice.
%Instead, in our payment system, we use Avalanche to resolve the double-spend
%conflicts in each conflict set, without maintaining a linear log.
%
%If we could assume each transaction can only have one single input, the
%initialization of Avalanche would be straightforward. We let each vertex on the
%underlying DAG be one transaction.  The conflict set in Avalanche is the set of
%transactions that try to spend the same unspent output. The conflict sets are
%disjoint because each transaction only has one input spending one
%unspent output, and thus belongs to exactly one set.
%
%%If we support transactions with multiple inputs using the same
%%method as in single-input case, we will violate the disjoint conflict set model assumed by
%%Avalanche, because the double spending relation between transactions with multiple
%%inputs is not transitive: $\texttt{TX}_a$ may double spend the same output with
%%$\texttt{TX}_b$ due to the same input value $\texttt{In}_{a1} =
%%\texttt{In}_{b1}$, and $\texttt{TX}_b$ might double spend with $\texttt{TX}_c$
%%due to $\texttt{In}_{b2} = \texttt{In}_{c1}$, but $\texttt{TX}_a$ may not
%%double spend with $\texttt{TX}_c$ at all.
%
%Multi-input transactions consume multiple UTXOs, and in Avalanche, may appear in multiple conflict sets.
%To account for these correctly, we represent \emph{transaction-input} pairs (e.g. $\texttt{In}_{a1}$) as an Avalanche vertex,
%and use the conjunction of \textsc{isAccepted} for all inputs of a transaction to ensure that no transaction will be accepted unless all its inputs are accepted.
%Since the acceptance for each pair is meaningful for
%the payment system only if all pairs from the same transaction are accepted, we can tie the fate of these
%pairs from the same transaction together by a single, bundled query: the queried
%node will only answer ``yes'' if all of the pairs are strongly preferred
%according to the DAG\@. This more conservative predicate will not undermine safety because merely introducing transactions that gather no chits will not increase confidence value in the
%protocol.
%
%\begin{figure}[t]
%\begin{center}
%    \input{figures/cash-system-a}
%    \captionof{figure}{The actual DAG implementation at transition granularity.}
%    \label{fig:cash-system-a}
%\end{center}
%\end{figure}
%
%Figure~\ref{fig:cash-system-a} demonstrates the actual implementation where the DAG is built at transaction granularity, whereas Figure~\ref{fig:cash-system-b} shows the equivalent logic of the underlying protocol, where vertices are at transaction-input granularity.
%}
%{
We inherit the transaction and address mechanisms from Bitcoin. At their simplest, transactions consist of multiple inputs and outputs, with corresponding redeem scripts.
Addresses are identified by the hash of their public keys, and signatures are generated by corresponding private keys.
The full scripting language is used to ensure that a redeem script is authenticated to spend a UTXO\@.
UTXOs are fully consumed by a valid transaction, and may generate new UTXOs spendable by named recipients.
Multi-input transactions consume multiple UTXOs, and in Avalanche, may appear in multiple conflict sets.
To account for these correctly, we represent \emph{transaction-input} pairs (e.g. $\texttt{In}_{a1}$) as Avalanche vertices.
The conflict relation of transaction-input pairs is transitive because of each pair only spends one unspent output.
Then, we use the conjunction of \textsc{isAccepted} for all inputs of a transaction to ensure that no transaction will be accepted unless all its inputs are accepted (Figure~\ref{fig:cash-system-b}). In other words, a transaction is accepted only if all its transaction-input pairs are accepted in their respective Snowball conflict sets.
Following this idea, we finally implement the DAG of transaction-input pairs such that multiple
transactions can be batched together per query.

%}

%However, supporting only a single input is both inconvenient and inefficient.
%One has to issue multiple transactions to merge the unspent balance in order to
%make larger payments. Moreover, with more than one output, the growth of unspent output
%set (i.e. UTXO set) size could be exponential, causing
%trouble for maintenance. However, if we also require a single output for
%all transactions, the granularity of coins cannot be changed, and one
%separated transaction is needed for transfering each unit of the currency.

\begin{figure}[t]
\begin{center}
    \input{figures/cash-system-b}
    \captionof{figure}{The underlying logical DAG structure used by Avalanche.
    The tiny squares with shades are dummy vertices which just help form the
    DAG topology for the purpose of clarity, and can be replaced by direct
    edges. The rounded gray regions are the conflict sets.}\label{fig:cash-system-b}
\end{center}
\end{figure}

%\noindent\textbf{Parent Selection.}
%The goal of the parent selection algorithm is to yield a well-structured DAG
%that maximizes the likelihood that virtuous transactions will be quickly
%accepted by the network. While this algorithm does not affect the safety of the
%protocol, it affects liveness and plays a crucial role in determining the shape of the
%DAG\@. A good parent selection algorithm grows the DAG
%in depth with a roughly steady ``width.'' The DAG should not diverge like a
%tree or converge to a chain, but instead should provide concurrency so nodes can work on multiple fronts.
%
%There are inherent trade-offs in the parent selection algorithm: selecting well-accepted parents makes it
%more likely for a transaction to find support, but can lead to vote dilution. Further, selecting more
%recent parents at the frontier of the DAG can lead to stuck transactions, as the parents may turn out
%to be rogue and remain unsupported.  In the following discussion, we illustrate this dilemma.
%We assume that every transaction will select a small number, $p$, of parents.
%We focus on the selection of eligible parent set, from which a subset of size $p$ can be chosen arbitrarily.
%%\Jon{Why is it important to choose parents randomly? What's the attack?}
%
%\tronly{
%Perhaps the simplest idea is to issue a fresh transaction with parents picked uniformly at random
%among those transactions that are currently strongly preferred. Specifically, we can adopt the predicate
%used in the voting rule to determine eligible parents on which a node would vote positively, as follows:
%\[
%    \mathcal{E} = \{T: \forall T\in\mathcal{T}, \textsc{isStronglyPreferred}(T)\}.
%\]
%
%But this strategy will yield large sets of eligible parents, consisting mostly of historical, old transactions.
%When a node samples the transactions uniformly from $\mathcal{E}$, the resulting DAG
%will have large, ever-increasing fan-out. Because new transactions will have scarce progenies,
%the voting process will take a long time to build the required confidence on any given new transaction.
%
%In contrast, efforts to reduce fan-out and control the shape of the DAG by selecting the recent transactions at
%the decision frontier suffer from another problem. Most recent transactions will have very low confidence,
%simply because they do not have enough descendants. Further, their conflict sets may not be well-distributed
%and well-known across the network, leading to a parent attachment under a transaction that will never be supported.
%This means the best parent candidates lie somewhere near the frontier, but not too far deep in history.
%
%The adaptive parent selection algorithm chooses parents by starting at the DAG frontier and retreating towards the
%genesis vertex until finding an eligible parent.
%}{
%The frontier transactions are those that currently do not have any children in the node's
%$\mathcal{T}$.  The set of strongly preferred transactions $\mathcal{E}$ is
%filtered by an additional criterion that the confidence should be larger than
%zero if the transaction has known conflicts. This filters out 0-confidence transactions from
%unstable conflict sets.  If there exist $p$ transactions
%in the filtered $\mathcal{E}$, then the adaptive selection will yield these transactions.
%
%Otherwise, the algorithm tries the parents of the transactions in $\mathcal{E}$,
%thus increasing the chance of finding more stabilized transactions as it
%retreats.
%}
%%XXX issue no-ops after a timeout
%The retreating search is guaranteed to terminate when it reaches the genesis
%vertex.  Formally, the selected parents in this adaptive selection algorithm
%is:
%
%\begin{center}
%\begin{algorithmic}[1]
%    \small
%    \Function{parentSelection}{$\mathcal{T}$}
%    \tronly{}{\State $\mathcal{E} = \{T: \forall T\in\mathcal{T}, \textsc{isStronglyPreferred}(T)\}$.}
%        \State $\mathcal{E}' \assign \{T: \forall T\in\mathcal{E}, |\mathcal{P}_T| = 1 \lor d(T) > 0\}$.
%        \State \Return $\{T: T \in \mathcal{E}' \land \forall T'\in\mathcal{T}, T \gets T', T' \notin \mathcal{E}'\}$.
%    \EndFunction
%    \captionof{figure}{Adaptive parent selection.}
%\end{algorithmic}
%\end{center}

\paragraph{Optimizations}
We implement some optimizations to help the system scale.
First, we use \emph{lazy updates} to the DAG, because the recursive definition for confidence may otherwise require a costly DAG traversal.
We maintain the current $d(T)$ value for each active vertex on the DAG, and update it only when a descendant vertex gets a chit.
Since the search path can be pruned at accepted vertices, the cost for an update is constant if the rejected vertices have a limited number of descendants and the undecided region of
the DAG stays at constant size.
Second, the conflict set could be large in practice, because a rogue client can generate a large volume of conflicting transactions.
Instead of keeping a container data structure for each conflict set, we create a mapping from each UTXO to the preferred transaction that stands as the representative for the entire conflict set.
This enables a node to quickly determine future conflicts, and the appropriate response to queries.
Finally, we speed up the query process by terminating early as soon as the $\alpha$ threshold is met, without waiting for $k$ responses.

%The safety guarantees of Snowball can be mapped to those of Avalanche, which is a concrete instantiation of Snowball using a directed acyclic graph to amortize cost. 
%We note that the structure of the Avalanche DAG itself does not correspond to votes, which is a subtle difference between other consensus protocols that make usage of a DAG. The DAG is merely a performance optimization, and is itself entirely orthogonal to the consensus process.

\paragraph{DAG} Compared to Snowball, Avalanche introduces a DAG structure that entangles the fate of unrelated conflict sets, each of which is a single-decree instance.
This entanglement embodies a tension: attaching a virtuous transaction to undecided parents helps propel transactions towards a decision, while it puts transactions at risk of suffering liveness failures when parents turn out to be rogue.
We can resolve this tension and provide a liveness guarantee with the aid of two mechanisms.

First we adopt an adaptive parent selection strategy, where transactions are attached at the live edge of the DAG, and are retried with new parents closer to the genesis vertex. This procedure is guaranteed to terminate with uncontested, decided parents, ensuring that a transaction cannot suffer liveness failure due to contested, rogue transactions.
A secondary mechanism ensures that virtuous transactions with decided ancestry will receive sufficient chits. Correct nodes examine the DAG for virtuous transactions that lack sufficient progeny and emit no-op transactions to help increase their confidence.
With these two mechanisms in place, it is easy to see that, at worst, Avalanche will degenerate into separate instances of Snowball, and thus provide the same liveness guarantee for virtuous transactions.

Unlike other cryptocurrencies~\cite{IOTA} that use graph vertices
directly as votes, Avalanche only uses a DAG for the purpose of batching queries
in the underlying Snowball instances.
Because confidence is built by collected chits, and not by just the presence of
a vertex, simply flooding the network with vertices attached to the rejected
side of a subgraph will not subvert the protocol.

\subsection{Communication Complexity}
Let the DAG induced by Avalanche have an expected branching factor of $p$, corresponding to the width of the DAG, and determined by the parent selection algorithm.
Given the $\beta_1$ and $\beta_2$ decision threshold, a transaction that has just reached the point of decision will have an associated progeny $\mathcal{Y}$.
Let $m$ be the expected depth of $\mathcal{Y}$.
If we were to let the Avalanche network make progress and then freeze the DAG at a depth $y$,
then it will have roughly $py$ vertices/transactions, of which $p(y - m)$ are decided in expectation.
Only $pm$ recent transactions would lack the progeny required for a decision.
For each node, each query requires $k$ samples, and therefore the total message cost per transaction is in expectation $(pky) / (p(y - m)) = ky/(y-m)$.
Since $m$ is a constant determined by the undecided region of the DAG as the system constantly makes progress, message complexity per node is $O(k)$, while the total complexity is $O(kn)$.

\section{Evaluation}
\label{sec:evaluation}
\input{sections/evaluation}

\section{Related Work}
\label{sec:related-work}
\input{sections/related-work}

\section{Conclusion}
\label{sec:conclusions}
This paper introduced a novel family of consensus protocols, coupled with the appropriate mathematical tools for analyzing them.
\tronly{These protocols are highly efficient and robust, combining the best features of classical and Nakamoto consensus.}
They scale well, achieve high throughput and quick finality, work without precise membership knowledge, and degrade gracefully under catastrophic adversarial attacks.

There is much work to do to improve this line of research. \tronly{
One such improvement could be the introduction of an adversarial network scheduler.
Another}{One} improvement would be to characterize the system's guarantees under an adversary whose powers are realistically limited, whereupon performance would improve even further. \tronly{Finally, more}{More} sophisticated initialization mechanisms would bear fruitful in improving liveness of multi-value consensus.
Overall, we hope that the protocols and analysis techniques presented here add to the arsenal of the distributed system developers and provide a foundation for new lightweight and scalable mechanisms.

%\section*{Acknowledgments}\tronly{}{\vspace{-0.5em}}
%\label{sec:related-work}
%We are grateful to the many individuals who provided feedback on earlier drafts. 
% Bobby Kleinberg, Vitalik Buterin, ...

%latex here should be the name of your bibtex file minus '.bib'
%\clearpage
\bibliographystyle{acm}
\bibliography{paper}
%\newpage
\begin{appendices}
\section{Analysis}
\label{sec:full-analysis}
In this appendix, we provide an analysis of Slush, Snowflake and Snowball.
%For lack of space, we leave additional details to an accompanying report.

\subsection{Preliminaries}
We assume the network model as discussed in Section~\ref{sec:model_and_goals}. We let $\mathtt{R}$ (``red'') and $\mathtt{B}$ (``blue'') represent two generic conflicting choices.
Without loss of generality, we focus our attention on counts of $\mathtt{B}$, i.e.\ the total number of nodes that prefer blue.

\paragraph{Hypergeometric Distribution} Each network query of $k$ peers corresponds to a sample without replacement out of a network of $n$ nodes, also referred to as a hypergeometric sample.
We let the random variable $\mathcal{H}(\mathcal{N}, x, k) \rightarrow \{0, \dots, k\}$ denote the resulting counts of $\mathtt{B}$ in the sample (unless otherwise stated), where $x$ is the total count of $\mathtt{B}$ in the population. The probability that the query achieves the required threshold of $\alpha$ or more votes is given by:
\begin{equation}
\small
P(\mathcal{H}(\mathcal{N}, x, k) \geq \alpha) = \left.\sum_{j = \alpha}^{k} {x \choose j} {n - x \choose k - j} \middle/ {n \choose k}\right.
\label{eq:hypergeometric}
\end{equation}
For ease of notation, we overload $\mathcal{H}(*)$ by implicitly referring to $P(\mathcal{H}(\mathcal{N}, x, k) \geq \alpha)$ as $\mathcal{H}(\mathcal{N}, x, k, \alpha)$.

\paragraph{Tail Bounds On Hypergeometric Distribution} We can reduce some of the complexity in Equation~\ref{eq:hypergeometric} by introducing a bound on the hypergeometric distribution induced by $\mathcal{H}^k_{\mathcal{N},x}$.
Let $p=x/n$ be the ratio of support for $\mathtt{B}$ in the population.
The expectation of $\mathcal{H}(\mathcal{N}, x, k)$ is exactly $kp$.
Then, the probability that $\mathcal{H}(\mathcal{N}, x, k)$ will deviate from the mean by more than some small constant $\psi$ is given by the Hoeffding tail bound \cite{hoeffding1963probability}, as follows,
\begin{equation}
\begin{split}
    P(\mathcal{H}(\mathcal{C}, x, k) \leq (p-\psi)k) &\leq e^{-k\mathcal{D}(p-\psi, p)}\\
    &\leq e^{-2(p-\psi)^2k}
\end{split}
\end{equation}
where $\mathcal{D}(p-\psi, p)$ is the Kullback-Leibler divergence, measured as
\begin{equation}
\begin{split}
    \mathcal{D}(a, b) &= a \log \frac{a}{b} + (1 - a) \log \frac{1 - a}{1 - b}
\end{split}
\end{equation}

\paragraph{Concentration of Sub-Martingales}
Let $\{X_{\{t \geq 0\}}\}$ be a sub-martingale and $|X_t - X_{t-1}| < c_t$ almost surely. Then, for all positive reals $\psi$ and all positive integers $t$, 
\begin{equation}
\begin{split}
P(X_t \geq X_0 + \psi) \leq e^{-\psi^2 / 2\sum_{i = 1}^{t} c_t^2}
\end{split}
\label{eq:submartingale}
\end{equation}

\subsection{Slush}
%\label{sec:analysis}
Slush operates in a non-Byzantine setting; that is, $f = 0$, $c = n$.
In this section, we will characterize the irreversibility properties of Slush (which appear in Snowflake and Snowball), as well as the precise converge rate distribution. The distribution of of both safety and liveness of Slush translate well to the Byzantine setting.

% \begin{figure}
%     \small
% \begin{algorithmic}[1]
%     \State initialize $u.\codecolor \in \{\texttt{R},\texttt{B}\}$ for all $u\in\mathcal{N}$
%     \For{$t = 1$ to $\phi$}
%         \State $u \assign \Call{sample}{\mathcal{N}, 1}$
%         \State $\mathcal{K} \assign \Call{sample}{\mathcal{N}\backslash u, k}$
%         \State $P \assign \texttt{[}v.\codecolor\quad\textbf{for}\ v \in \mathcal{K}\texttt{]}$
%         \For{$\codecolor' \in \{\texttt{R}, \texttt{B}\}$}
%             \If{$P.\Call{count}{\codecolor'} \ge \alpha$}
%                 \State $u.\codecolor \assign \codecolor'$
%             \EndIf
%         \EndFor
%     \EndFor
% \captionof{figure}{Slush run by a global scheduler.}\label{fig:slush_protocol_simulator}
% \end{algorithmic}
% \end{figure}

The procedural version of Slush in Figure~\ref{fig:slush-loop} made use of a parameter $m$, the number of rounds that a node executes Slush queries. 
% To derive this parameter, we transform the protocol execution from a procedural and concurrent version to one carried out by a scheduler, shown in Figure~\ref{fig:slush_protocol_simulator}.
What we ultimately want to extract is the total number of rounds $\phi$ that the scheduler will need to execute in order to guarantee that the entire network is the same color, whp.

We analyze the system mainly using a continuous time process. Let $\{X_{\{t \geq 0\}}\}$ be a CTMC.
The state space $\mathcal{S}$ of the stochastic process is a condensed version of the full configuration space, where each state $\{0, \dots, n\}$ represents the total number of blue nodes in the system.

Let $\mathcal{F}_{X_s}$ be the filtration, or the history pertaining to the process, up to time $s$. This process is Markovian and time-homogeneous, conforming to 
\[
    P\{X_t = j | \mathcal{F}_{X_s}\} = P\{X_t = j | X_s\} = P\{X_t = j | X_0\}    
\]
Throughout the paper, we use $Q \equiv (q_{ij}, i, j \in \mathcal{S})$ notation to refer to the infinitesimal generator of the process, where death ($i \rightarrow i-1$) and birth ($i \rightarrow i+1$) rates of configuration transitions are denoted via $\mu_i$ and $\lambda_i$ ($\lambda_ i$ is distinct from the clock parameter $\lambda$, and will be clear from context). These rates are 
\[
    \begin{cases}
        \mu_i = i\ \mathcal{H}(\mathcal{N}, c-i, k, \alpha), & \text{for } i \rightarrow i - 1 \\
        \lambda_i = (c-i)\ \mathcal{H}(\mathcal{N}, i, k, \alpha), & \text{for } j \rightarrow i + 1 \\
    \end{cases}
\]
for $1\leq i\leq c-1$, and where $i = 0$ and $i = c$ are absorbing. Let $p_{ij}(t)$ refer to the probability of transitioning from state $i$ to $j$ at time $t$. 
We always assume that 
\[
    p_{ij}(t) = 
    \begin{cases}
      \lambda_it + o(t), & \text{for } j = i + 1 \\
      \mu_it + o(t), & \text{for } j = i - 1 \\
      1 - (\lambda_i + \mu_i)t + o(t), & \text{for } j = i \\
      o(t), & \text{otherwise }\\
    \end{cases}
\]
where all $o(t)$ are uniform in $i$.

\paragraph{Irreversibility}
%We now discuss the core results of irreversibility properties of our family of protocols.
In Section~\ref{sec:analysis}, we discussed the loose Chvatal bound which provided intuitive understanding into the strong irreversibility dynamics of our core subsampling mechanism. In particular, once the network drifts to some majority value, it tends to revert back with only an exponentially small probability. We compute the closed-form expression for reversibility, and show that it is exponentially small.
\begin{theorem}
\label{theorem:slush_prob_convergence_minority}
Let $\xi_\delta$ be the probability of absorption into the all-red state ($s_0$), starting from a drift of $\delta$ (i.e. $\delta$ drift away from $n/2$). Then, assuming $\delta > 1$, 
\begin{equation}
\xi_\delta = 1 - \ddfrac{\sum_{l = 1}^{\delta} \prod_{i = 1}^{l-1} \mu_i^2 \prod_{j = l}^{n-l}\lambda_j}{2\sum_{l = 1}^{n/2}\prod_{i=1}^{l-1}\mu_i^2\prod_{j=l}^{n-l}\mu_j}
\end{equation}
and
% \begin{equation}
% \xi_\delta - \xi_{\delta+1} = \ddfrac{\mu_1^2\dots\mu_{\delta}^2\lambda_{\delta+1}\dots\lambda_{n-\delta}}{\Phi(\alpha, k, n)}
% \end{equation}
% where $\Phi(\alpha, k, n)$ is a constant dependent on $\alpha$, $k$, and $n$, but otherwise independent of starting configuration. 
\begin{equation}
\begin{split}
\ddfrac{\xi_{\delta} - \xi_{\delta+1}}{\xi_{\delta+1} - \xi_{\delta+2}} &= \mathcal{u}_{\delta+1} = \ddfrac{\lambda_{\delta+1}}{\mu_{\delta+1}} \\
&\approx \ddfrac{n-\delta-1 \sum_{j = \alpha}^{k}\ddfrac{(n-\delta-1)^k (\delta+1)^{k-j}}{n^{2k - j}}}{\delta+1 \sum_{j = \alpha}^{k}\ddfrac{(\delta+1)^k (n-\delta-1)^{k-j}}{n^{2k - j}}}
\end{split}
\end{equation}
where from now on we refer to $\mathcal{u}_{\delta+1}$ as the drift of the process. 
\end{theorem}

\begin{proof}
Our results are derived based on constructions from Tan~\cite{tan1976absorption}. We construct a sub-matrix of $Q$, denoted $B$, as shown in Figure~\ref{fig:matrixB}.
\begin{figure*}
\[B = 
\begin{bmatrix}
    -(\lambda_1 + \mu_1) & \lambda_1 & 0 & \cdots & \cdots & 0 \\
    \mu_2 & -(\lambda_2 + \mu_2) & \lambda_2 & 0 & \cdots & 0\\
    0 & \mu_3 & -(\lambda_3 + \mu_3) & \lambda_3 & \cdots & 0\\
    \vdots & \vdots & \ddots & \ddots & \ddots & \vdots\\
    \vdots & \vdots & \mu_{n-3} & -(\lambda_{n-2} + \mu_{n-2}) & \lambda_{n-3} & 0\\
    \vdots & \dots & 0 & \mu_{n-1} & -(\lambda_{n-2} + \mu_{n-2}) & \lambda_{n-2}\\
    0 & \dots & 0 & 0 & \mu_{n-1} & -(\lambda_{n-1} + \mu_{n-1})
\end{bmatrix}
\]
\caption{Matrix $B$.}
\label{fig:matrixB}
\end{figure*}
Let $W_1'$ = $(\mu_1, 0, \dots, 0)$, $W_{n-1}'$ = $(0, \dots, 0, \lambda_{n-1})$. Then, we can express $Q$ as 
\[
    Q =
    \begin{bmatrix}
        0 & \dots & 0\\
        W_1 & B & W_{n-1}\\
        0 & \dots & 0
    \end{bmatrix}
\]
As a reminder, the stationary distribution can be found via $\lim_{t \rightarrow \infty} P(t) = e^{Qt}$, where we have
\[
    e^{Qt} = \sum_{i = 0}^{\infty} \frac{t^i}{i!} Q^i = \sum_{i = 0}^{\infty} \frac{t^i}{i!}
    \begin{bmatrix}
        0 & \dots & 0\\
        B^{i-1}W_1 & B^i & B^{i-1}W_{n-1}\\
        0 & \dots & 0
    \end{bmatrix}
\]
As Tan (eq. 2.3) shows, we have
\[
    \xi(t) = B^{-1}\left[\sum_{i = 0}^{\infty} B^i  - \mathbb{I}_{n-1} \right] W_1
\]
Since we want the ultimate probabilities, we have that 
\[
    \xi = \lim_{t \rightarrow \infty} \xi(t) = -B^{-1}W_1
\]
We can explicitly compute $\xi_\delta$ in terms of our rates $\mu_i$ and $\lambda_i$, getting 
\[
    \xi_\delta = \ddfrac{\sum_{l = 1}^{n-\delta}\prod_{i = 1}^{n-l}\mu_i \prod_{j = n-l+1}^{n-1}\lambda_j}{\sum_{l = 1}^{n}\prod_{i = 1}^{n-l}\mu_i \prod_{j = n-l+1}^{n-1}\lambda_j}
\]
However, we note that $u_{i} = \lambda_{n-i}$. Algebraic manipulation from this observation leads to the two equations in the theorem. This expression is strictly lower than the Chvatal bounds used in Section~\ref{sec:analysis}.
\end{proof}

Using the construction for the absorption (and (ir)reversibility) probabilities as discussed previously, a natural follow up computation is in regards to \emph{mean convergence time}. 
Let $T_{z}(t) = \inf \{t \geq 0 : X_t = \{0, n\} | X_0 = z\}$, and let $\tau_z = \mathbb{E}[T_{z}(t)]$. $\tau_z$ is the mean time to reach either absorbing state, starting from state $z$, which corresponds to the mean convergence time. The next theorem characterizes this distribution.

\begin{theorem}
\label{theorem:mean-convergence-time}
Let $\tau_z$ be the expected time to convergence, starting from state $z > n/2$, to any of the two converging states in the network (all-red or all-blue). Then, 
\begin{equation}
\tau_z = \ddfrac{\sum_{d = 1}^{n-1}x(d)y(d)}{2\sum_{l = 1}^{n/2}\prod_{i=1}^{l-1}\mu_i^2\prod_{j=l}^{n-l}\mu_j}
\end{equation}
where $x(d)$ and $y(d)$ are
\begin{equation}
\begin{split}
x(d) &= \sum_{l = 1}^{\min(z, d)} \prod_{i=1}^{l-1} \mu_i \prod_{j = l}^{d-1} \lambda_j\\
y(d) &= \sum_{l = 1}^{n - d - \max(z-d, 0)} \prod_{i = d+1}^{n-l} \mu_i \prod_{j = n - l + 1}^{n - 1} \lambda_j
\end{split}
\end{equation}
\end{theorem}

\begin{proof}
Following the calculations from before, $-B^{-1}$ at row $z$ provides the number of traversals to each other state starting from $z$. Calculating their sum, we have our result. The above equation is the full expression of the matrix row sum. 
\end{proof}

Theorem~\ref{theorem:mean-convergence-time} leads to the next lemma that captures property P2, under the assumption that at the beginning of the protocol, one proposal has at least $\alpha$ support in the network. 
\begin{lemma}
Slush reaches an absorbing state in finite time almost surely.
\label{lemma:finitetermination}
\end{lemma}

\begin{proof}
Starting from any non-absorbing, transient state, there is a non-zero probability of being absorbed. Additionally, since termination is finite and everywhere differentiable, Theorem~\ref{theorem:mean-convergence-time} also implies that the probability of termination of any network configuration where a proposal has $\geq \alpha$ support in bounded time $t_{max}$ is strictly positive. 
\end{proof}

\subsection{Snowflake}
\label{subsection:appendix_snowflake}
In Snowflake, the sampled set of nodes includes Byzantine nodes.
We introduce the decision function $\mathcal{D}(*)$, which is constructed by having each node also keep track of the total number of consecutive times it has sampled a majority of the same color ($\beta$). 
Finally, we introduce a function called $\mathcal{A}(\mathcal{S}_t)$, the adversarial strategy, that takes as parameters the entire configuration of the network at time $t$, as well as the next set of nodes chosen by the scheduler to execute, and as a side-effect, modifies the set of nodes $\mathcal{B}$ to some arbitrary configuration of colors.

In order for our prior framework to apply to Snowflake, we must deal with a key subtlety. 
Unlike in Slush, where it is clear that once the network has reached one of the converging states and therefore may not revert back, this no longer applies to Snowflake, since any adversary $f \geq \alpha$ has strictly positive probability of reverting the system, albeit this probability may be infinitesimally small. 
The CTMC is flexible enough to deal with a system where there is only one absorbing state, but the long-term behavior of the system is no longer meaningful since, after an infinite amount of time, the system is guaranteed to revert, violating safety. 
We could trivially bound the amount of time, and show safety using this bounded time assumption by simply characterizing the distribution of $e^{tQ}$, where $Q$ is the generator. 
However, we can make the following observation: if the probability of going from state $c$ (all-blue) to $c-1$ is exponentially small, then it will take the attacker exponential time (in expectation; note, this is a lower bound, and in reality it will take much longer) to succeed in reverting the system. 
Hence, we can assume that once all correct nodes are the same color, the attack from the adversary will terminate since it is impractical to continue an attack. 
In fact, under reasonably bounded timeframes, the variational distance between the exact approach and the approximation is very small. 
We leave details to the accompanying paper, but we briefly discuss how analysis proceeds for Snowflake.

As stated in Section~\ref{sec:analysis}, the way to analyze the adversary using the same construction as in Slush is to condition reversibility on the first node $u$ deciding on blue, which can happen at any state (as specified by $\mathcal{D}(*)$). At that point, the adversarial strategy collapses to a single function, which is to continually vote for red. The probabilities of reversibility, for all states $\{1, \dots, c-1\}$ must encode the probability that additional blue nodes commit, and the single function of the adversary. The birth and death rates are transformed as follows:
\[
    \begin{cases}
        \mu_i = &i(1 - \mathbb{I}[\mathcal{D}(*, i, \mathbb{B})])\ \mathcal{H}(\mathcal{N}, c-i + f, k, \alpha)\\
        \lambda_i = &(c-i)(1 - \mathbb{I}[\mathcal{D}(*, c-i, \mathbb{R})])\ \mathcal{H}(\mathcal{N}, i, k, \alpha)\\
    \end{cases}
\]
From here on, the analysis is the same as in Slush. Under various $k$ and $\beta$, we can find the minimal $\alpha$ that provides the system strong irreversibility properties.

The next lemma captures P3, and the proof follows from central limit theorem. 
\begin{lemma}
If $f < \Oh{\sqrt{n}}$, and $\alpha = \floor{k/2} + 1$, then Snowflake terminates in $\Oh{\log n}$ rounds with high probability. 
\label{lemma:centrallimit}
\end{lemma}
\begin{proof}
The results follows from central limit theorem, wherein for $\alpha = \floor{k/2} + 1$, the expected bias in the network after sampling will be $\Oh{\sqrt{n}}$. An adversary smaller than this bias will be unable to keep the network in a fully-bivalent state for more than a constant number of rounds. The logarithmic factor remains from the mixing time lower bound. 
\end{proof}

\subsection{Snowball}
We make the following observation: if the confidences between red and blue are equal, then the adversary has the same identical leverage in the irreversibility of the system as in Snowflake, regardless of network configuration. In fact, Snowflake can be viewed as Snowball but where drifts in confidences never exceed one. The same analysis applies to Snowball as in Snowflake, with the additional requirement of bounding the long-term behavior of the confidences in the network. To that end, analysis follows using martingale concentration inequalities, in particular the one introduced in Equation~\ref{eq:submartingale}. Snowball can be viewed as a two-urn system, where each urn is a sub-martingale. The guarantees that can be extracted hereon are that the confidences of the majority committed value (in our frame of reference is always blue), grow always more than those of the minority value, with high probability, drifting away as $t \rightarrow t_{max}$. 
\subsection{Safe Early Commitment}
As we reasoned previously, each conflict set in Avalanche can be viewed as an instance of Snowball, where each progeny instance iteratively votes for the entire path of the ancestry.
This feature provides various benefits; however, it also can lead to some virtuous transactions that depend on a rogue transaction to suffer the fate of the latter.
%has the drawback that it can entangle the fate of some unfortunate virtuous transactions with the fate of rogue ones.
In particular, rogue transactions can interject in-between virtuous transactions and reduce the ability of the virtuous transactions to ever reach the required $\textsc{isAccepted}$ predicate.
As a thought experiment, suppose that a transaction $T_i$ names a set of parent transactions that are all decided, as per local view.
If $T_i$ is sampled over a large enough set of successful queries without discovering any conflicts, then, since by assumption the entire ancestry of $T_i$ is decided, it must be the case (probabilistically) that we have achieved irreversibility.

To then statistically measure the assuredness that $T_i$ has been accepted by a large percentage of correct nodes without any conflicts, we make use of a one-way birth process, where a birth occurs when a new correct node discovers the conflict of $T_i$. Necessarily, deaths cannot exist in this model, because a conflicting transaction cannot be unseen once a correct node discovers it. 
% Let $t = 0$ be the time when $T_j$, which conflicts with $T_i$, is introduced to a single correct node $u$.
% Let $s_x$, for $x = 1$ to $c$, be the state where the number of correct nodes that know about $T_j$ is $x$, and let $p(s_x)$ be the probability of birth at state $s_x$. Then, we have:
Our births are as follows:
\begin{equation}
    \lambda_i = \frac{c - i}{c} \left(1 - \frac{{n - i \choose k}}{{n \choose k}}\right)
\end{equation}
Solving for the expected time to reach the final birth state provides a lower bound to the $\beta_1$ parameter in the $\textsc{isAccepted}$ fast-decision branch. The table below shows an example of the analysis for $n = 2000$, $\alpha = 0.8k$, and various $k$, where $\varepsilon \ll 10^{-9}$, and where $\beta$ is the minimum required value before deciding.
\begin{table}[h!]
    \small
	\centering
	\begin{tabular}{llllll}
		$k$   & 10 & 20 & 30 & 40 \\ \hline
		$\beta$ & 10.87625 & 10.50125 & 10.37625 & 10.25125
	\end{tabular}
	\label{table:fast-path-beta}
\end{table}
% \begin{figure}
%     \includegraphics[width=\linewidth]{figures/fast-path-beta.pdf}
%     \captionof{figure}{An example of $\beta$ solutions for different $k$, with $n = 2000, \alpha = 0.8k$.}
%     \label{fig:fast-path-beta}
% \end{figure}
\noindent Overall, a very small number of iterations are sufficient for the safe early commitment predicate. This supports the choice of $\beta$ in our evaluation.

\subsection{Initialization Heuristic}
\label{sec:sync-heuristic}
To improve liveness, we can use strong synchrony assumptions. The heuristic works as follows. Every node operates in two phases: in the first phase, it gossips and collects proposals for $\Oh{\log{n}}$ rounds, where each round lasts for the maximum message delay, which ensures the proposal from a correct node will propagate to almost all other correct nodes; in the second phase, each node stops collecting proposals, and gossips the existing proposals for an additional $\Oh{\log{n}}$ rounds so that every correct node ends up with approximately the same set of proposals. Finally, each node samples for the proposals it knows of locally, checking for those that have an $\alpha$ majority, ordered deterministically, such as by hash values. It then selects the first value by the order as its initial state when it starts the actual consensus protocol.
In a cryptocurrency setting, the deterministic ordering function would incorporate fees paid out for every new proposal, which means that the adversary is financially limited in its ability to launch a fairness attack against the initialization.

\subsection{Churn and View Discrepancy}\label{sec:full-analysis-churn}

Realistic systems need to accommodate the departure and arrival of nodes.
Up to now, we simplified our analysis by assuming a precise knowledge of network membership, i.e. $\mathcal{L}(u) = \mathcal{N}$.
We now demonstrate that correct nodes can admit a well-characterized amount of churn, by showing how to pick parameters such that Avalanche nodes can differ in their view of the network and still safely make decisions.

{\color{black}

To characterize churn we use a generalized set intersection construction that allows us to make arguments about worst-case network view splits. Before formalizing it, we provide the intuition: suppose we split the network into two entirely independent, but fully connected, subsets. Clearly, the Byzantine adversary wins with probability one since it can send two conflicting transactions to the two independent networks respectively, and they would finalize the transactions immediately. This represents the worst case view split. 
\begin{figure}
\centering
\includegraphics[width=0.8\linewidth]{figures/network_view.pdf}
\caption{Changes in network view based on $S_d$'s size. All prior proofs up to now represent the variant where $S_a = S_b = \emptyset$. With the new construction, the probability of safety violation is simply a direct application of the prior sets of proofs under the new subsets.}
\label{fig:network_view}
\end{figure}
We can generalize this to arbitrary network splits, which can even be applied recursively in each subset. The proofs of safety then are a matter of characterizing the probability of red and blue committing into the two (possibly independent) subsets.

Suppose we divide the set of correct nodes into three subsets, $S_a$, $S_d$, $S_b$. We overload $\mathcal{L}(S_{*})$ to represent the views of any node within the input set. The view of all nodes in $S_a$ is $\mathcal{L}(S_a) = S_a \cup S_d \cup \mathcal{B}$, the view of all nodes in $S_b$ is $\mathcal{L}(S_b) = S_b \cup S_d \cup \mathcal{B}$, and the view of $S_d$ is $\mathcal{N}$. We assume the worst case, which means that adversarial nodes are common to all subsets. When $S_d = \emptyset$, this represents a division of the network into two equal subsets where $|S_a| = |S_b| = n/2$~\footnote{$|S_a|$ and $|S_b|$ do not have to be equal, we assume so as a demonstration.}. If $S_d$ is all correct nodes, then $|S_a| = |S_b| = n$. This construction is visually demonstrated in Figure~\ref{fig:network_view}.

\begin{lemma}
Let $\tau \in \mathbb{Z}^{+}$. Let $|S_d| = n - \tau$, and thus $|S_{\{a, b\}}| = \tau/2$. There exists some maximal size of $\tau$ such that probability of any two nodes $u, v \in S_a,\ S_b,\ S_d$ finalizing equivocating transactions is less than $\epsilon$. 
\end{lemma}
\begin{proof}
We assume that the adversary has full control of the network view splits, meaning that they choose in full how to create $S_a,\ S_b,\ S_d$. To prove safety, we simply reuse the same exact construction as in Subsection~\ref{subsection:appendix_snowflake}, but we replace the original set $\mathcal{N}$ with a new set of interest, namely $S_a \cup S_d \cup \mathcal{B}$ (i.e., we exclude $S_b$)~\footnote{Sets are symmetric in this example.}. To thus find the maximal $\tau$, we simply replace $u_i$ and $\lambda_i$ with 
\begin{equation}
    \begin{cases}
        \bar \mu_i = i\ \mathcal{H}(S_a \cup S_d \cup \mathcal{B}, c-(\tau/2)-i, k, \alpha), & \text{for } i \rightarrow i - 1 \\
        \lambda_i = (c-i)\ \mathcal{H}(S_a \cup S_d \cup \mathcal{B}, i, k, \alpha), & \text{for } j \rightarrow i + 1 \\
    \end{cases}
\end{equation}
where
\begin{equation}
\begin{split}
P(\mathcal{H}&(S_a \cup S_d \cup \mathcal{B}, x, k) \geq \alpha)\\
&= \left.\sum_{j = \alpha}^{k} {x \choose j}{c-(\tau/2)-x \choose k-j}\middle/{c - (\tau/2) \choose k}\right.\\
\end{split}
\end{equation}
and apply the same construction as in Subsection~\ref{subsection:appendix_snowflake}. As $\tau$ increases, it is clear that $S_a$ (and conversely, $S_b$), become more independent and less reliant on the values proposed by members in $S_d$, thus incrementing the ability of the adversaries. 
\end{proof}

}

% Consider a network whose operation is divided into epochs of length $\tau$, and a view update from epoch $t$ to $t+1$ during which $\gamma$ nodes join the network and $\bar \gamma$ nodes depart.
% Under our static construction, the state space $\mathcal{S}_t$ of the network had a key parameter $\Delta^{t}$ at time $t$, induced by $c^{t}, f^{t}, n^{t}$ and the chosen security parameters.
% Churn can, at worst, impact the network by adding $\gamma$ nodes of color $\mathtt{B}$, and remove $\bar \gamma$ nodes of color $\mathtt{R}$.
% At time $t+1$, $n^{t+1} = n^{t} + \gamma - {\bar \gamma}$, while $f^{t+1}$ and $c^{t+1}$ will be modified by an amount $\leq \gamma - {\bar \gamma}$, and thus induce a new $\Delta^{t+1}$ for the chosen security parameters.
% This new $\Delta^{t+1}$ has to be chosen such that the probability of reversibility from state $c^{t+1}/2 + \Delta^{t+1} - \gamma$ is $\leq \varepsilon$, which ensures that the system will converge under the previous pessimal assumptions. The system designer can easily do this by picking an upper bound on $\gamma, \bar \gamma$.

% The final step in assuring the correctness of a view change is to account for a mix of nodes that straddle the $\tau$ boundary. We would like the network to avoid an unsafe state no matter which nodes are using the old and the new views.
% The easiest way to do this is to determine $\Delta^t$ and $\Delta^{t+1}$ for desired bounds on $\gamma, \bar \gamma$, and then to use the conservative value $\Delta^{t+1}$ during epoch $t$. In essence, this ensures that no commitments are made in configuration $\mathcal{S}_t$ unless they conservatively fulfill the safety criteria in state space $\mathcal{S}_{t+1}$. As a result, there is no possibility of a node deciding red at time $t$, the network going through an epoch change and finding itself to the left of the new irreversibility state $\Delta^{t+1}$.

% This approach trades off some of the feasibility space, to add the ability to accommodate $\gamma, \bar \gamma$ node churn per epoch.
% Overall, if $\tau$ is in excess of the time required for a decision (on the order of minutes to hours), and nodes are loosely synchronized,
% they can add or drop up to $\gamma, \bar \gamma$ nodes in each epoch using the conservative process described above.
% % Due to space reasons,
% We leave the precise method of entering and exiting the network by staking and unstaking to a subsequent paper, and instead
% rely on a membership oracle that acts as a sequencer and $\gamma$-rate-limiter, using technologies like Fireflies~\cite{JohansenRVJ15}.
% and Ethereum~\cite{wood2014ethereum}.
\end{appendices}
\end{document}

### 2020-11-03 paper_old_analysis.tex
add initial source and 'fr' make target

\documentclass[letterpaper,twocolumn,10pt]{article}
%\usepackage[letterpaper, total={6.5in, 9in}, columnsep=0.25in]{geometry}
\usepackage[letterpaper, total={7in, 9in}, columnsep=0.33in]{geometry}
\newcommand{\tronly}[2]{#1}
\usepackage{newtxtext}
\usepackage{appendix}
\tronly{}{
\usepackage{microtype}
\usepackage[leadingfraction=1,moderate]{savetrees}
}
\usepackage{comment}

\setlength{\parskip}{0pt}
\setlength{\abovedisplayskip}{3pt}
\setlength{\belowdisplayskip}{3pt}
\setlength{\textfloatsep}{2pt plus 1.0pt minus 2.0pt}
\setlength{\floatsep}{2pt plus 1.0pt minus 2.0pt}
\usepackage{cite}
\usepackage{paralist}
\usepackage[hyphens]{url}
\usepackage[compact]{titlesec}
\usepackage{enumitem}
\setdefaultleftmargin{0em}{}{}{}{}{}
\setdefaultitem{{}}{}{}{}
\setlist[itemize]{leftmargin=*}
\setlist[enumerate]{leftmargin=*}
%\expandafter\def\expandafter\UrlBreaks\expandafter{\UrlBreaks%  save the current one
%\do\a\do\b\do\c\do\d\do\e\do\f\do\g\do\h\do\i\do\j%
%\do\k\do\l\do\m\do\n\do\o\do\p\do\q\do\r\do\s\do\t%
%\do\u\do\v\do\w\do\x\do\y\do\z\do\A\do\B\do\C\do\D%
%\do\E\do\F\do\G\do\H\do\I\do\J\do\K\do\L\do\M\do\N%
%\do\O\do\P\do\Q\do\R\do\S\do\T\do\U\do\V\do\W\do\X%
%\do\Y\do\Z}
\usepackage{caption}
% \captionsetup[figure]{font=small,labelfont=small,skip=3pt}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{bbm}
\usepackage{thmtools, thm-restate}
\usepackage{mathrsfs}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{tikz}
\usepackage{dsfont}
\usepackage{capt-of}
\usepackage{algorithmicx}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{amsthm}
\usepackage[makeroom]{cancel}
\usepackage{datetime}
\renewcommand\ttdefault{cmtt}
\graphicspath{{./figures}}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}
\let\oldemptyset\emptyset
\let\emptyset\varnothing
% New definitions
\algnewcommand\algorithmicswitch{\textbf{switch}}
\algnewcommand\algorithmiccase{\textbf{case}}
\algnewcommand\algorithmicassert{\texttt{assert}}
\algnewcommand\Assert[1]{\State \algorithmicassert(#1)}%
% New "environments"
\algdef{SE}[SWITCH]{Switch}{EndSwitch}[1]{\algorithmicswitch\ #1\ \algorithmicdo}{\algorithmicend\ \algorithmicswitch}%
\algdef{SE}[CASE]{Case}{EndCase}[1]{\algorithmiccase\ #1}{\algorithmicend\ \algorithmiccase}%
\algtext*{EndSwitch}%
\algtext*{EndCase}%⇧
\algnewcommand\algorithmiccontinue{\textbf{continue}}
\algnewcommand\algorithmicbreak{\textbf{break}}
\algnewcommand\Continue{\algorithmiccontinue}
\algnewcommand\Break{\algorithmicbreak}
\newcommand{\editremove}[1]{{\color{red}\sout{#1}}}
\newcommand{\editinsert}[1]{{\color{blue}#1}}
\newcommand{\editchange}[1]{{\color{orange}#1}}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{conjecture}[theorem]{Conjecture}
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\setcounter{footnote}{1}
\newcommand{\Jon}[1]{{\color{blue} \textbf{Jon: } #1}}

\settimeformat{hhmmsstime}
\mmddyyyydate
\title{
\vspace{-0.30in}
    \Large \bf \editchange{Snowball}: The Power of Metastable Consensus}
%\tronly{{}
    \author{Team Rocket\footnote{\texttt{\ 0x\ 8a5d 2d32 e68b c500 36e4 d086 0446 17fe 4a0a 0296 b274 999b a568 ea92 da46 d533}
    }\\t-rocket@protonmail.com}
%}{
%\author{Paper \texttt{\#220}}
%}
%\tronly{\date{Revision: \today{} \currenttime{} UTC}}
\date{Revision: \today{} \currenttime{} UTC}
%{\date{\vspace{-2ex}}}
\usepackage[hang]{footmisc}
\setlength\footnotemargin{0em}
\begin{document}
\maketitle

\begin{abstract}
This paper introduces a new family of leaderless Byzantine fault tolerance protocols,
built around a metastable mechanism.
These protocols provide a strong probabilistic safety guarantee in the presence of Byzantine adversaries\tronly{, while their concurrent nature enables them to achieve high throughput and scalability.}{.}
Unlike blockchains that rely on proof-of-work, they are quiescent and green.
Unlike traditional consensus protocols having leader nodes that
process $O(n)$ information (in terms of number of bits exchanged), 
no node processes more than \editchange{$O(k\log n)$ or $O(k)$} for some security parameter~$k$.
% require $O(n^{2})$
% communication, their communication complexity ranges from
% $O(kn\log n)$
% to
% $O(kn)$
% for some security parameter $k \ll n$. \Jon{This sentence led me to assuem that the two ``best'' protocols were both going to be important; that there would be a reason to use the $O(kn\log n)$ version despite its poorer aymptotics. But when I got to later pages, I see you're only discussing Avalanche after all; the other algorithms are purely steps on the pedagogical garden path, right? Maybe remove the reference to the intermediate steps from the abstract?}
\tronly{% new paragraph

The paper describes the \editchange{Snowball protocol family}, analyzes its guarantees, and describes how it can be used to construct the core of an internet-scale electronic payment system.
}{}
\editinsert{A peer-to-peer payment system, Avalanche} is evaluated in a large scale deployment.
Experiments demonstrate that the system can achieve high throughput (3400 tps), provide low confirmation latency (1.35 sec), and scale well compared to existing systems that deliver similar functionality. For our implementation and setup, the bottleneck of the system is in transaction verification.
\end{abstract}

\section{Introduction}\tronly{}{\vspace{-0.5em}}

\tronly{Achieving agreement among a set of distributed hosts lies at the core of countless applications, ranging from Internet-scale services that serve billions of people~\cite{Burrows06,HuntKJR10} to cryptocurrencies worth billions of dollars~\cite{marketcapcryptocurrency}.}{}
To date, there have been two main families of solutions to this problem.
Traditional consensus protocols %, exemplified by PBFT~\cite{castro1999practical},
rely on all-to-all communication to ensure that all correct nodes reach the same decisions with absolute certainty.
Because they require quadratic communication overhead and accurate knowledge of membership, they have been difficult to scale
\tronly{to large numbers of participants.}{.}

On the other hand, Nakamoto consensus protocols~\cite{nakamoto2008bitcoin,GarayKL15, PassSS17, SompolinskyZ15, SompolinskyLZ16, SompolinskyZ18, BentovHMN17, EyalGSR16,Kokoris-KogiasJ16,PassS16a, PassS18} have become popular with the rise of Bitcoin.
These protocols provide a probabilistic safety guarantee: Nakamoto consensus decisions may revert with some probability $\varepsilon$.
A protocol parameter allows this probability to be rendered arbitrarily small\tronly{, enabling high-value financial systems to be constructed on this foundation.}{.}
This family is a natural fit for open, permissionless settings where any node can join the system at any time.
Yet, these protocols are costly, wasteful, and limited in performance.
By construction, they cannot quiesce: their security relies on constant participation by miners, even when there are no decisions to be made.
\tronly{Bitcoin currently consumes around 63.49 TWh/year~\cite{bitcoinpower}, about twice as all of Denmark~\cite{denmarkpower}.}{}
Moreover, these protocols suffer from an inherent scalability bottleneck that is difficult to overcome through simple reparameterization (see~\cite{CromanDEGJKMSSS16} for more details). % \Jon{I wanted the paper to expand on this point, but I missed it.}

This paper introduces a new family of consensus protocols.
Inspired by gossip algorithms, this family gains its safety through a deliberately metastable mechanism.
Specifically, the system operates by repeatedly sampling the network at random, and steering correct nodes towards the same outcome.
Analysis shows that this metastable mechanism is powerful: it can move a large network to an irreversible state quickly.
%{\color{red} We should clarify this.  The current statement is too strong.}
% though it is not always guaranteed to do so.
% \Jon{Not universal in that some applications require a hard guarantee?}

Similar to Nakamoto consensus, this new protocol family provides a probabilistic safety guarantee, using a tunable security parameter that can render the possibility of a consensus failure arbitrarily small.
Unlike Nakamoto consensus, the protocols are green, quiescent and efficient; they do not rely on proof-of-work~\cite{DworkN92} and do not consume energy when there are no decisions to be made.
The efficiency of the protocols stems from two sources: each node requires communication overheads ranging from $O(k \log n)$ to $O(k)$ for some small security parameter $k$ (whereas leader-based consensus protocols have one or more nodes that require $O(n)$ communication), and the protocols establish only a partial order among dependent transactions.

This combination of the best features of traditional and Nakamoto consensus involves one significant tradeoff: liveness for \editchange{equivocating values}. %conflicting transactions.

\editchange{Specifically, the theorecital, binary consensus protocol from the Snowball family guarantees liveness in logarithmic number of rounds in expectation. While, in the demonstrated payment system, Avalanche, the liveness is affected by the diversity of initial proposals and will be guaranteed when a unanimous proposal is given.
It guarantees liveness only for \emph{virtuous} transactions, i.e.\ those issued by correct clients and thus guaranteed not to conflict with other transactions.
However,} in a cryptocurrency setting, cryptographic signatures enforce that only a key owner is able to create a transaction that spends a particular coin.
Since correct clients follow the protocol as prescribed, they are guaranteed both safety and liveness.
In contrast, the protocols do not guarantee liveness for \emph{rogue} transactions, submitted by Byzantine clients, which conflict with one another.
Such decisions may stall in the network, but have no safety impact on virtuous transactions.
We show that this is a sensible tradeoff, and that resulting system is sufficient for building complex payment systems.

Overall, this paper makes one significant contribution: a brand new family of consensus protocols, based on randomized sampling and metastable decision. The protocols provide a strong probabilistic safety guarantee, and \editchange{the liveness guarantee suitable for building a decentralized, peer-to-peer, payment system.}

The next section provides the intuition behind the new protocols, followed by their full specification,
Section~\ref{sec:attack} briefly discusses typical attack vectors and their possible defense,
Section~\ref{sec:analysis} provides methodology used
by our formal analysis of safety and liveness in Appendix~\ref{sec:full-analysis},
%Section~\ref{sec:implementation} describes a Bitcoin-like payment system,  Section~\ref{sec:evaluation} evaluates the system,
Section~\ref{sec:implementation} describes \editchange{Avalanche, a Bitcoin-like payment system,
Section~\ref{sec:evaluation} evaluates Avalanche},
Section~\ref{sec:related-work} presents related work, and finally, Section~\ref{sec:conclusions} summarizes our contributions.

\section{Approach}\tronly{}{\vspace{-0.5em}}
\label{sec:approach}
We start with a non-Byzantine protocol, Slush, and progressively build up \editchange{Snowflake and Snowball}, all based on the same common metastable mechanism.
\tronly{
\editremove{Slush, Snowflake, and Snowball} \editinsert{They} are single-decree binary consensus protocols
of increasing robustness\editremove{---Avalanche extends Snowball into a full
cryptocurrency solution; other extensions for other applications are certainly
possible}.
We provide full specifications for the protocols in this section, and defer the analysis to the next section, and present formal
proofs in the appendix.
}{}

\editremove{Overall, this protocol family achieves its properties by humbly cheating in three different ways.}
\editremove{First, taking} \editinsert{Taking} inspiration from Bitcoin, we adopt a safety guarantee that is probabilistic.
\tronly{This probabilistic guarantee is indistinguishable from traditional safety guarantees in practice, since appropriately small choices of $\varepsilon$ can render consensus failure practically infeasible, less frequent than CPU miscomputations or hash collisions.}{}

\editinsert{On the other hand, our practical payment system, Avalanche (Section~\ref{sec:implementation}), achieves its
properties by humbly cheating in two ways.}
\editremove{Second, }\editinsert{First, }instead of a single replicated state machine (RSM) model, where the system determines a sequence of totally-ordered transactions $T_0, T_1, T_2, \ldots$ issued by any client, we adopt a \emph{parallel consensus model}, where each client interacts independently with its own RSMs and optionally transfers ownership of its RSM to another client. (MOVED to Model Section) The system establishes only a partial order between dependent transactions.
\editremove{Finally, }\editinsert{Second, }for Avalanche, we provide no liveness guarantee for misbehaving clients, but ensure that well-behaved clients will eventually be serviced.
These techniques, in conjunction, enable the system to implement a useful Bitcoin-like cryptocurrency, with drastically better performance and scalability.

\editchange{Because our Snowball family provides strong safety guarantee and liveness guarantee},
it is possible to build other applications involving large-scale probabilistic consensus. We focus on a cryptocurrency application because of many challenges it poses.
%Nevertheless, Snowball does provide liveness guarantee for conflicting values, so it is possible to extend this single-decree protocol for other possible applications involving large-scale probabilistic consensus.
%\Jon{Are there other applications where this consensus model would be valuable? It's not a BFT RSM, but it would be very intriguing if there were other additional applications.}

\subsection{Model, Goals, Threat Model}\tronly{}{\vspace{-0.5em}}

We assume a collection of nodes, $\mathcal{N}$,
composed of correct nodes $\mathcal{C}$ and
Byzantine nodes $\mathcal{B}$, where $n = |\mathcal{N}|$.

\editinsert{In Avalanche}, we adopt what is commonly known as
Bitcoin's unspent transaction output (UTXO) model.
In this model,
\emph{clients} are authenticated and issue cryptographically signed
transactions that fully consume an existing UTXO and issue new UTXOs.
Unlike nodes, clients do not participate in the decision process, but only
supply transactions to the \emph{nodes} running the consensus protocol.
Two transactions \emph{conflict} if
they consume the same UTXO and yield different outputs.
Correct clients never issue conflicting transactions, nor is it possible
for Byzantine clients to forge conflicts with transactions
issued by correct clients.
However, Byzantine clients can issue multiple transactions that conflict
with one another, and correct clients can only consume one of
those transactions.
\editchange{To make use of Snowball consensus in Avalanche}, then, is to \emph{accept} a
set of non-conflicting transactions in the presence of Byzantine behavior.
Each client can be considered a replicated state machine whose
transitions are defined by a totally ordered list of accepted transactions.

Our family of protocols provide the following guarantees with high probability (whp):
\begin{compactitem}
% RVR: this property is unnecessary because the liveness property already ensures non-triviality
%    \item \textbf{Non-triviality}: To ensure that the protocol carries out real work, decisions must include values proposed by external clients;

    \item \textbf{P1. Safety.} No two correct nodes will accept conflicting transactions.
% the following is an assumption, not a property.
%	for some well-characterized $|\mathcal{B}|$.

    \item \textbf{P2. Liveness.} Any transaction issued by a correct client (aka virtuous transaction) will eventually be accepted by every correct node.
\end{compactitem}

Instead of unconditional agreement, our approach guarantees that safety will be upheld with probability $1-\varepsilon$, where the choice of the security parameter $\varepsilon$ is under the control of the system designer and applications.

We assume a powerful adaptive adversary capable of observing the internal state and communications of every node in the network, but not capable of interfering with communication between correct nodes.
Our analysis assumes a synchronous network, while our deployment and evaluation is performed in a partially synchronous setting.
We conjecture that our results hold in partially synchronous networks, but the proof is left to future work.
We do not assume that all members of the network are known to all participants, but rather may temporarily have some discrepancies in network view.
We assume a safe bootstrapping system, similar to that of Bitcoin, that enables a node to connect with sufficiently many correct nodes to acquire a statistically unbiased view of the network.
We do not assume a PKI\@.
We make standard cryptographic assumptions related to public key signatures and hash functions.

\subsection{Slush: Introducing Metastability}\tronly{}{\vspace{-0.5em}}
The core of our approach is a single-decree consensus protocol,
inspired by epidemic or gossip protocols.
% In this approach, a state in which all possible outcomes are equally
% probable is metastable and disturbances are likely to cause the
% protocol to converge to one of multiple stable states.
% We will call such consensus protocols metastable themselves.
%We can model a preference for a decision as a kind of infection: a node catches a disease, and converts other nodes to carry the same disease in epidemic fashion.
%If a threshold of its interactions are with other infected hosts, it flips its color to the competing agent and joins their ranks.
%
The simplest metastable protocol, Slush, is the foundation of this family, shown in Figure~\ref{fig:slush-loop}.
Slush is \emph{not} tolerant to Byzantine faults, but serves as an illustration for the BFT protocols that follow.
For ease of exposition, we will describe the operation of Slush using a decision between two conflicting colors, red and blue.
% , though the technique generalizes to any number of competing proposed values.
% Slush operates as follows.

In Slush, a node starts out initially in an uncolored state.
Upon receiving a transaction from a client, an uncolored node updates its own color to the one carried in the transaction and initiates a query.
To perform a query, a node picks a small, constant sized ($k$) sample of the network uniformly at random, and sends a query message.
Upon receiving a query, an uncolored node adopts the color in the query, responds with that color, and initiates its own query, whereas a colored node simply responds with its current color.
If $k$ responses are not received within a time bound, the node picks an additional sample from the remaining nodes uniformly at random and queries them until it collects all responses.
Once the querying node collects $k$ responses, it checks if a fraction $\ge\alpha k$ are for the same color, where $\alpha > 0.5$ is a protocol parameter.
If the $\alpha k$ threshold is met and the sampled color differs from the node's own color, the node flips to that color.
It then goes back to the query step, and initiates a subsequent round of query, for a total of $m$ rounds.
Finally, the node decides the color it ended up with at time $m$.

\algnewcommand{\IIf}[1]{\State\algorithmicif\ #1\ \algorithmicthen}
\algnewcommand{\EndIIf}{\unskip\ \algorithmicend\ \algorithmicif}

\newcommand{\codecolor}{\textrm{col}}
\newcommand{\codecount}{\textrm{cnt}}
\begin{figure}
    \small
\begin{algorithmic}[1]
    \Procedure{onQuery}{$v, \codecolor'$}
        \IIf{$\codecolor = \bot$} $\codecolor \coloneqq \codecolor'$
        \State\Call{respond}{$v, \codecolor$}
    \EndProcedure
    \Procedure{slushLoop}{$u, \codecolor_0 \in \{\texttt{R}, \texttt{B}, \bot\}$}
        \State $\codecolor \coloneqq \codecolor_0$ \textrm{// initialize with a color}
        \For{$r \in \{1\ldots m\}$}
            \State \textrm{// if $\bot$, skip until \textsc{onQuery} sets the color}
            \If{$\codecolor = \bot$}
                \Continue
            \EndIf
            \State \textrm{// randomly sample from the known nodes}
            \State $\mathcal{K} \coloneqq \Call{sample}{\mathcal{N}\textbackslash u, k}$
            \State $P \coloneqq \texttt{[}\Call{query}{v, \codecolor}\quad\textbf{for}\ v \in \mathcal{K}\texttt{]}$
            \For{$\codecolor' \in \{\texttt{R}, \texttt{B}\}$}
                \If{$P.\Call{count}{\codecolor'} \ge \alpha\cdot k$}
                    \State $\codecolor \coloneqq \codecolor'$
                \EndIf
            \EndFor
        \EndFor
        \State \Call{accept}{$\codecolor$}
    \EndProcedure
    \captionof{figure}{Slush protocol. Timeouts elided for readability.}\label{fig:slush-loop}
\end{algorithmic}
\end{figure}

This simple protocol has some curious properties.
First, it is almost \emph{memoryless}: a node retains no state between rounds other than its current color, and in particular maintains no history of interactions with other peers.
Second, unlike traditional consensus protocols that query every participant, every round involves sampling just a small, constant-sized slice of the network at random.
\tronly{Third, even if the network starts in the metastable state of a 50/50 red-blue split, random perturbations in sampling will cause one color to gain a slight edge and repeated samplings afterwards will build upon and amplify that imbalance.}{}
Finally, if $m$ is chosen high enough, Slush ensures that all nodes will be colored identically whp.
Each node has a constant, predictable communication overhead per round, and $m$ grows logarithmically with $n$. %XXX where do we do this?

\tronly{
%If Slush is deployed in a network with Byzantine nodes, the adversary can interfere with decisions.
%In particular, if the correct nodes develop a preference for one color, the adversary can attempt to flip nodes to the opposite so as to keep the network in balance.
%The Slush protocol lends itself to analysis but does not, by itself, provide a strong safety guarantee in the presence of Byzantine nodes, because the nodes lack state.
%We address this in our first BFT protocol.
% Ittay:
The Slush protocol does not provide a strong safety guarantee in the presence of Byzantine nodes.
In particular, if the correct nodes develop a preference for one color, a Byzantine adversary can attempt to flip nodes
to the opposite so as to keep the network in balance, preventing a decision.
We address this in our first BFT protocol that introduces more state storage at the nodes.
}{
We next examine how to extend Slush to tolerate Byzantine behavior.
}

\subsection{Snowflake: BFT}\tronly{}{\vspace{-0.5em}}

Snowflake augments Slush with a single counter that captures the strength of a node's conviction in its current color.
This per-node counter stores how many consecutive samples of the network by that node have all yielded the same color.
A node accepts the current color when its counter exceeds $\beta$, another security parameter.
Figure~\ref{fig:snowflake-loop} shows the amended protocol, which includes
the following modifications:

\begin{compactenum}
	\item Each node maintains a counter $\mathit{cnt}$;
    \item Upon every color change, the node resets $\mathit{cnt}$ to 0;
    \item Upon every successful query that yields $\ge \alpha k$ responses for the same color as the node, the node increments $\mathit{cnt}$.
\end{compactenum}

\begin{figure}
    \small
\begin{algorithmic}[1]
    \Procedure{snowflakeLoop}{$u, \codecolor_0 \in \{\texttt{R}, \texttt{B}, \bot\}$}
        \State $\codecolor \coloneqq \codecolor_0$, $\codecount \coloneqq 0$
        \While{\textrm{undecided}}
            \If{$\codecolor = \bot$}
                \Continue
            \EndIf
            \State $\mathcal{K} \coloneqq \Call{sample}{\mathcal{N}\textbackslash u, k}$
            \State $P \coloneqq \texttt{[}\Call{query}{v, \codecolor}\quad\textbf{for}\ v \in \mathcal{K}\texttt{]}$
            \For{$\codecolor' \in \{\texttt{R}, \texttt{B}\}$}
            \If{$P.\Call{count}{\codecolor'} \ge \alpha\cdot k$}
                \If{$\codecolor' \neq \codecolor$}
                    \State $\codecolor \coloneqq \codecolor'$, $\codecount \coloneqq 0$
                \Else
                    \IIf {$\texttt{++}\codecount > \beta$} \Call{accept}{$\codecolor$}
                \EndIf
            \EndIf
            \EndFor
        \EndWhile
    \EndProcedure
    \captionof{figure}{Snowflake.}\label{fig:snowflake-loop}
\end{algorithmic}
\end{figure}

When the protocol is correctly parameterized for a given threshold of Byzantine nodes and a desired $\varepsilon$-guarantee, it can ensure both safety (P1) and liveness (P2).
As we later show, there exists a phase-shift point after which correct nodes are more likely to tend towards a decision than a bivalent state.
Further, there exists a point-of-no-return after which a decision is inevitable.
The Byzantine nodes lose control past the phase shift, and the correct nodes begin to commit past the point-of-no-return, to adopt the same color, whp.

\tronly{
\begin{figure}
    \small
\begin{algorithmic}[1]
    \Procedure{snowballLoop}{$u, \codecolor_0 \in \{\texttt{R}, \texttt{B}, \bot\}$}
        \State $\codecolor \coloneqq \codecolor_0$, $\textrm{lastcol} \coloneqq \codecolor_0$, $\codecount \coloneqq 0$
        \State $d[\texttt{R}] \coloneqq 0$, $d[\texttt{B}] \coloneqq 0$
        \While{\textrm{undecided}}
            \IIf{$\codecolor = \bot$} \Continue
            \State $\mathcal{K} \coloneqq \Call{sample}{\mathcal{N}\textbackslash u, k}$
            \State $P \coloneqq \texttt{[}\Call{query}{v, \codecolor}\quad\textbf{for}\ v \in \mathcal{K}\texttt{]}$
            \For{$\codecolor' \in \{\texttt{R}, \texttt{B}\}$}
            \If{$P.\Call{count}{\codecolor'} \ge \alpha\cdot k$}
                \State $d[\codecolor']\texttt{++}$
                \If {$d[\codecolor'] > d[\codecolor]$}
                        \State$\codecolor \coloneqq \codecolor'$
                \EndIf
                \If{$\codecolor' \neq \textrm{lastcol}$}
                    \State$\textrm{lastcol} \coloneqq \codecolor'$, $\codecount \coloneqq 0$
                \Else
                    \IIf {$\texttt{++}\codecount > \beta$} \Call{accept}{$\codecolor$}
                \EndIf
            \EndIf
            \EndFor
        \EndWhile
    \EndProcedure
    \captionof{figure}{Snowball.}\label{fig:snowball-loop}
\end{algorithmic}
\end{figure}
}{}

\subsection{Snowball: Adding Confidence}\tronly{}{\vspace{-0.5em}}

Snowflake's notion of state is ephemeral: the counter gets reset with every color flip.
While, theoretically, the protocol is able to make strong guarantees with
minimal state, we will now improve the protocol to make it harder to
attack by incorporating a more permanent notion of belief.
Snowball augments Snowflake with \emph{confidence counters} that capture the number of queries that have yielded a threshold result for their corresponding color (Figure ~\ref{fig:snowball-loop}).
A node decides if it gets $\beta$ consecutive chits for a color. However, it only changes preference based on the total accrued confidence.
The differences between Snowflake and Snowball are as follows:

\begin{compactenum}
    \item Upon every successful query, the node increments its confidence counter for that color.
    \item A node switches colors when the confidence in its current color becomes lower than the confidence value of the new color.
\end{compactenum}

Snowball is not only harder to attack than Snowflake, but is more easily
generalized to multi-decree protocols.

\editremove{MOVED Avalanche to the implementation section}
\section{Analysis}
\label{sec:analysis}
Due to space limits, we move the full safety analysis to Appendix~\ref{sec:full-analysis}.
There, we show that under certain independent and distinct assumptions, the \editchange{Snowball} family of consensus protocols provide properties P1 (safety) and P2 (liveness) with probability $1 - \varepsilon$, where $\varepsilon$ is a security parameter that can be set according to the needs of the system designer.
In this section, we summarize the methodology used for our analysis and provide some insight into our core results.
\subsection{Safety}
% Analyzing the protocols introduced in this paper
%differ from classical as well as Nakamoto consensus-style protocols. 
%Even simple, non-linear dynamical systems often incur complicated analyses as they may not admit analytical solutions.
% In particular, since the state of a node depends on randomly sampling other nodes, possibly inducing nondeterministic dependency cycles, analyzing all possible state executions is computationally infeasible. 
% Here, we employ stochastic techniques used in the analysis of epidemic networks, which may also be of independent interest to the distributed systems community. 
%Here, we employ 
% requires complex stochastic techniques to reason about the state of the network in aggregate.

\paragraph{Scheduler, Network, and Adversary.} 
We analyze the protocols through a global view. 
Like Bitcoin and other blockchain protocols based on gossip dissemination, we assume that gossip spreads to all correct nodes within a known time bound.  
In other words, we assume a synchronous communication network, where the system makes progress in rounds, and where at the beginning of each round the system scheduler chooses a single correct node $u$ uniformly at random. 
Node $u$ will sample and update its state by the end of the round. 
The Byzantine adversary will be aware of the identity of $u$. Furthermore, the adversary has full knowledge of the internal state of all nodes in the network at all times, and is able to fully update the state of all nodes under its control immediately after $u$ chooses its neighbor sample set. 
In essence, the adversary is only constrained by an inability to directly update the state of correct nodes.

\paragraph{System Dynamics.} We can think of the network as a set of nodes either colored red or blue. Each new round will update the state of one of the nodes, either changing its color, or increasing its confidence in its current color. 
The dynamics of the system resemble those of epidemic networks, where nodes update their state based on the state of their neighbors, using a function chosen over some probability distribution. 
It would be infeasible to keep track of all possible execution paths, since at every round, the possible number of branching executions is equal to the number of possible correct nodes times all possible $k$-sample choices of outcomes that the node may sample. Furthermore, the difficulty of this task is greatly amplified due to the Byzantine adversary, whose optimal attack strategy is chosen over an unknown function, possibly itself nondeterministic.

The simplification that allows us to analyze this system is to obviate the need to keep track of all of the execution paths, as well as all possible adversarial strategies, and rather focus entirely on a single state of interest, without regards to how we achieve this state. 
More specifically, the core extractable insight of our analysis is in identifying the \textit{irreversibility} state of the system, the state upon which so many correct nodes have usurped either red or blue that reverting back to the minority color is highly unlikely.

\paragraph{Slush}
We model the Slush protocol is through a birth-death Markov process with two absorbing states, where either all nodes are red or all nodes are blue. 
In our construction, all states (excluding absorbing states) are transient, meaning that there is a nonzero probability of never returning to that state. 
Our analysis shows that the Slush protocol reaches, in finite time, an absorbing state with non-zero probability.
Furthermore, the probability of convergence to one of the two colors can be precisely measured using only a local round counter. In fact, Slush converges to a decision in close to logarithmic steps whp.

% Another interesting outcome of the protocol is that optimal convergence rate occurs at $k = \alpha = 1$, meaning that the network will reach unanimous decision fastest when every node simply polls a single other neighbor. 
% As we discuss for later protocols, however, there is a tension between convergence rate and safety guarantees. Smaller values of $k$ impede the ability of the correct nodes to protect against an adversarial attack.

\paragraph{Snowflake}
For Snowflake, we relax the assumption that all nodes are correct and assume that some fraction of nodes are adversarial.
An important insight is that there exists an irreversible state, or \textit{point of no return}, after which the system will converge to an absorbing state whp. Furthermore a correct node only decides when the system is beyond the point of no return. Composing these two guarantees together, the probability of a safety violation is strictly less than $\epsilon$, which can be configured as desired. 
Unsurprisingly, there is an inherent tension between safety and liveness, but suitable parameters can be found that are practical.
Larger values of $k$ obtain higher levels of security for correct nodes, at the expense of slower convergence.

\paragraph{Snowball}
Snowball is an improvement over Snowflake, where random perturbations in network samples are reduced by introducing a partial form of history, which we refer to as confidence. 
Modeling Snowball with a Markov chain is difficult because of a state space explosion problem. In particular, it is not sufficient to simply keep track of color preferences of each correct node, the analysis must also maintain information about their confidence. 
To make analysis possible, we structure the mathematical foundations via a game of balls and urns, where each urn represents one of the correct nodes, and the ball count correspond to confidences in either color. 
Using this model, the analysis applies various tail inequalities to prove that the safety guarantees of Snowball are strictly stronger than those of Snowflake. In particular, once the system has reached the point of no return, the probability of reverting is strictly lower than in Snowflake.

\paragraph{Avalanche}
\editremove{%
MOVED to implementation}

\subsection{Liveness}\tronly{}{\vspace{-0.5em}}
% \tronly{
% \noindent\textbf{Slush.}
% Slush is a non-BFT protocol, which terminates whp within a finite number of rounds.
% }{}

\noindent\textbf{Snowflake \& Snowball.}
Both protocols make use of a counter to keep track of consecutive majority support. Since the adversary is unable to forge
a conflict for a virtuous transaction, initially,
all correct nodes will have color red or
$\bot$. A Byzantine node cannot respond to any query with any answer
other than red since it is unable to forge conflicts and $\bot$ is not allowed by protocol. Therefore, the only misbehavior for the Byzantine node is refusing to answer. Since the correct node will re-sample if the query times out, by expected convergence in \tronly{Equation~\ref{equation:expected_convergence}}{\cite{avalanche}} of Appendix~\ref{sec:full-analysis},
 all correct nodes will terminate with the unanimous red value within a finite number of rounds whp.

\editremove{%
\noindent\textbf{Avalanche.}
MOVED to implementation}

\subsection{Additional Insights}
Two interesting insights arise from our analysis.
First, the protocols discussed in this work lead to both safety and liveness guarantees whose underlying function is smooth, rather than a step function. 
In many other consensus protocols, safety is guaranteed with up to a fixed threshold number (e.g.~1/3) of adversarial nodes, beyond which no guarantees are provided. In our protocols, the guarantees degrade gracefully with an adversarial percentage beyond the pre-established bound. For example, optimal system parameters can be chosen to tolerate precisely 1/5 adversarial presence with failure probability $\epsilon$. However, if the system faces an adversarial presence greater than 1/5, then the probability of failure degrades to slightly above $\epsilon$, rather than immediately to $1$. 
Second, these protocols externalize the safety and liveness tradeoffs. 
The system designer may choose to guarantee safety even under catastrophic events, such as an adversarial presence beyond $1/2$, at the expense of liveness.  
This presents a powerful knob not available in classical or Nakamoto-based consensus protocols.

% \begin{compactitem}
% \item\textbf{Eventually good ancestry.} Virtuous transactions can be retried by picking new parents, selected from a set that is more likely to be preferred.
% Ultimately, one can always attach a transaction to decided parents to completely mitigate this risk.
% \Jon{So why don't we always do this? Later the paper clarifies that moving up the chain reduces the efficiency
%     and increases latency for the proposed transaction. I got pretty lost in the last two paragraphs of section 4}
% A simple technique for parent selection is to select new parents for a virtuous transaction at successively lower heights in the DAG, proceeding towards the genesis vertex.
% This procedure is guaranteed to terminate with uncontested, decided parents, ensuring that the transaction cannot suffer liveness failure due to rogue transactions.

% \item\textbf{Sufficient chits.} A secondary mechanism is necessary to ensure that virtuous transactions with decided ancestry will receive sufficient chits.
% To ensure this, correct nodes examine the DAG for virtuous non-nop transactions that lack sufficient progeny and emit nop transactions to help increase their confidence.
% A nop transaction has just one parent and no application side-effects, and can be issued by any node. They cannot be abused by Byzantine nodes because, even though nops trigger new queries, they do not automatically grant chits.

% With these two mechanisms in place, it is easy to see that, at worst, Avalanche will degenerate into separate instances of Snowball, and thus provide the same liveness guarantee for virtuous transactions.
% \end{compactitem}

%XXX discuss how no-ops can be implemented

\subsection{Attack Vectors}
\label{sec:attack}
There are a multitude of attack vectors against consensus protocols.  
We consider the two most important ones.
\paragraph{Sybil Attacks}
Consensus protocols provide their guarantees based on assumptions that only a fraction of participants are adversarial.
These bounds could be violated if the network is naively left open to arbitrary participants.
In particular, a Sybil attack~\cite{douceur2002sybil}, wherein a large number of identities are generated by an adversary, could be used to exceed the adversarial bound.

A long line of work, including PBFT~\cite{castro1999practical}, treats the Sybil problem separately from consensus, and rightfully so, as Sybil control mechanisms are distinct from the underlying, more complex agreement protocol\footnote{This is not to imply that every consensus protocol can be coupled/decoupled with every Sybil control mechanism.}.
Nakamoto consensus, for instance, uses proof-of-work~\cite{aspnes2005exposing} to limit Sybils, which requires miners to continuously stake a hardware investment.
Other protocols, discussed in Section~\ref{sec:related-work}, rely on proof-of-stake or proof-of-authority.
The consensus protocols presented in this paper can adopt any
% established
Sybil control mechanism, although proof-of-stake is most aligned with their quiescent operation.
The full design of a cryptocurrency incorporating staking, unstaking and minting mechanism is beyond the scope of this paper, whose focus is on the core consensus protocol.

% Sybil control mechanisms are orthogonal and separate from the consensus protocols. All such protocols, excluding Nakamoto-style consensus protocols, require a mechanism for network identity establishment. In a real implementation of Avalanche, the Sybil problem would be solved through a staking mechanism.

\paragraph{Flooding Attacks}
Flooding/spam attacks are a problem for any distributed system. 
Without a protection mechanism, an attacker can generate large numbers of transactions and flood the DAG, consuming storage. 
% Although there is no proper, protocol-level methodology to prevent such an attack, there are very effective heuristics that we can employ. 
There are a multitude of techniques to deter such attacks, including network-layer protections, proof-of-authority, or economic mechanisms. 
In Avalanche, we use transaction fees, making such attacks costly even if the attacker is sending money back to addresses under its control.

\subsection{Communication Complexity}\tronly{}{\vspace{-0.5em}}
Since liveness is not guaranteed for rogue transactions, we focus our message complexity analysis solely for the case of virtuous transactions.
For the case of virtuous transactions, Snowflake and Snowball are both guaranteed to terminate after $O(kn\log n)$ messages.
This follows from the well-known results related to epidemic algorithms\cite{demers1987epidemic}, and is confirmed by Table~\ref{table:growth_worst_case} in Appendix~\ref{sec:full-analysis}.

\editremove{Avalanche: MOVED to implementation}

\section{Peer-to-Peer Payment System}\tronly{}{\vspace{-0.5em}}
\label{sec:implementation}

%\label{sec:evaluation}

\editchange{We have implemented a bare-bones payment system, Avalanche, and
fully ported Bitcoin transactions to it.} In this section, \editchange{we describe the design and sketch how the implementation can support the value transfer primitive at the center of cryptocurrencies.} \editremove{and then examine its throughput, scalability, and latency through a large scale deployment on Amazon AWS\@, and finally, provide a comparison to known results from other systems.}

Deploying a full cryptocurrency involves bootstrapping, minting, staking, unstaking,
and inflation control. While we have solutions for these issues, their full discussion is beyond the scope of this paper.
%In this section,
%we focus on how Avalanche can support the value transfer primitive at the center of cryptocurrencies.

\input{avalanche}

\editremove{\noindent\textbf{UTXO Transactions.}}
\editinsert{\subsection{Support UTXO Transactions}}
In addition to the DAG structure in Avalanche, a UTXO graph that captures
spending dependency is used to realize the ledger for the payment system. To
avoid ambiguity, we denote the transactions that encode the data for money
transfer \emph{transactions}, while we call the
transactions ($T \in \mathcal{T}$) in Avalanche's DAG \emph{vertices}.

%\tronly
%{
%Each transaction represents a money transfer that takes several inputs from
%source accounts and generates several outputs to destinations. As a UTXO-based system
%that keeps a decentralized ledger, balances are kept by the \emph{unspent outputs}
%of transactions.
%
%More specifically, a transaction $\texttt{TX}_a$ maintains
%a list of inputs: $\texttt{In}_{a1}$, $\texttt{In}_{a2}$, $\cdots$. Each input
%has two fields: the reference to an unspent transaction output and a spend script.
%The unspent transaction output uniquely refers to an output of a previously made transaction.
%The script snippet will be prepended to the script from the referred output forming a complete computation.
%It could typically be a cryptographic proof of validity, but could also be any Turing-complete
%computation in general.
%Each output $\texttt{Out}_{a1}$, $\texttt{Out}_{a2}$, $\cdots$,
%contains some amount of money and a script which typically contains cryptographic verification that takes the proof from the future input and verifies validity.
%
%In our payment system, there are \emph{addresses} representing different
%accounts by cryptographic keys. The public
%key is used as the identity for recipients in the output scripts, while
%the private key is for creating signatures in the
%input scripts, spending the available funds. Only the key owner is able spend the unspent output by creating an input with
%the signature in a new transaction.
%
%Due to the possibility of double spending by the private key owner,
%cryptocurrencies such as Bitcoin use a blockchain as the linear log to reject the
%transaction that comes later in the log and tries to spend some output twice.
%Instead, in our payment system, we use Avalanche to resolve the double-spend
%conflicts in each conflict set, without maintaining a linear log.
%
%If we could assume each transaction can only have one single input, the
%initialization of Avalanche would be straightforward. We let each vertex on the
%underlying DAG be one transaction.  The conflict set in Avalanche is the set of
%transactions that try to spend the same unspent output. The conflict sets are
%disjoint because each transaction only has one input spending one
%unspent output, and thus belongs to exactly one set.
%
%%If we support transactions with multiple inputs using the same
%%method as in single-input case, we will violate the disjoint conflict set model assumed by
%%Avalanche, because the double spending relation between transactions with multiple
%%inputs is not transitive: $\texttt{TX}_a$ may double spend the same output with
%%$\texttt{TX}_b$ due to the same input value $\texttt{In}_{a1} =
%%\texttt{In}_{b1}$, and $\texttt{TX}_b$ might double spend with $\texttt{TX}_c$
%%due to $\texttt{In}_{b2} = \texttt{In}_{c1}$, but $\texttt{TX}_a$ may not
%%double spend with $\texttt{TX}_c$ at all.
%
%Multi-input transactions consume multiple UTXOs, and in Avalanche, may appear in multiple conflict sets.
%To account for these correctly, we represent \emph{transaction-input} pairs (e.g. $\texttt{In}_{a1}$) as an Avalanche vertex,
%and use the conjunction of \textsc{isAccepted} for all inputs of a transaction to ensure that no transaction will be accepted unless all its inputs are accepted.
%Since the acceptance for each pair is meaningful for
%the payment system only if all pairs from the same transaction are accepted, we can tie the fate of these
%pairs from the same transaction together by a single, bundled query: the queried
%node will only answer ``yes'' if all of the pairs are strongly preferred
%according to the DAG\@. This more conservative predicate will not undermine safety because merely introducing transactions that gather no chits will not increase confidence value in the
%protocol.
%
%\begin{figure}[t]
%\begin{center}
%    \input{figures/cash-system-a}
%    \captionof{figure}{The actual DAG implementation at transition granularity.}
%    \label{fig:cash-system-a}
%\end{center}
%\end{figure}
%
%Figure~\ref{fig:cash-system-a} demonstrates the actual implementation where the DAG is built at transaction granularity, whereas Figure~\ref{fig:cash-system-b} shows the equivalent logic of the underlying protocol, where vertices are at transaction-input granularity.
%}
%{
We inherit the transaction and address mechanisms from Bitcoin. At their simplest, transactions consist of multiple inputs and outputs, with corresponding redeem scripts.
Addresses are identified by the hash of their public keys, and signatures are generated by corresponding private keys.
The full scripting language is used to ensure that a redeem script is authenticated to spend a UTXO\@.
UTXOs are fully consumed by a valid transaction, and may generate new UTXOs spendable by named recipients.
Multi-input transactions consume multiple UTXOs, and in Avalanche, may appear in multiple conflict sets.
To account for these correctly, we represent \emph{transaction-input} pairs (e.g. $\texttt{In}_{a1}$) as Avalanche vertices.
The conflict relation of transaction-input pairs are transitive because of each pair only spends one unspent output.
Then, we use the conjunction of \textsc{isAccepted} for all inputs of a transaction to ensure that no transaction will be accepted unless all its inputs are accepted (Figure~\ref{fig:cash-system-b}). 
Following this idea, we finally implement the DAG of transaction-input pairs such that multiple
transactions can be batched together per query.

%}

%However, supporting only a single input is both inconvenient and inefficient.
%One has to issue multiple transactions to merge the unspent balance in order to
%make larger payments. Moreover, with more than one output, the growth of unspent output
%set (i.e. UTXO set) size could be exponential, causing
%trouble for maintenance. However, if we also require a single output for
%all transactions, the granularity of coins cannot be changed, and one
%separated transaction is needed for transfering each unit of the currency.

\begin{figure}[t]
\begin{center}
    \input{figures/cash-system-b}
    \captionof{figure}{The underlying logical DAG structure used by Avalanche.
    The tiny squares with shades are dummy vertices which just help form the
    DAG topology for the purpose of clarity, and can be replaced by direct
    edges. The rounded gray regions are the conflict sets.}\label{fig:cash-system-b}
\end{center}
\end{figure}

%\noindent\textbf{Parent Selection.}
%The goal of the parent selection algorithm is to yield a well-structured DAG
%that maximizes the likelihood that virtuous transactions will be quickly
%accepted by the network. While this algorithm does not affect the safety of the
%protocol, it affects liveness and plays a crucial role in determining the shape of the
%DAG\@. A good parent selection algorithm grows the DAG
%in depth with a roughly steady ``width.'' The DAG should not diverge like a
%tree or converge to a chain, but instead should provide concurrency so nodes can work on multiple fronts.
%
%There are inherent trade-offs in the parent selection algorithm: selecting well-accepted parents makes it
%more likely for a transaction to find support, but can lead to vote dilution. Further, selecting more
%recent parents at the frontier of the DAG can lead to stuck transactions, as the parents may turn out
%to be rogue and remain unsupported.  In the following discussion, we illustrate this dilemma.
%We assume that every transaction will select a small number, $p$, of parents.
%We focus on the selection of eligible parent set, from which a subset of size $p$ can be chosen arbitrarily.
%%\Jon{Why is it important to choose parents randomly? What's the attack?}
%
%\tronly{
%Perhaps the simplest idea is to issue a fresh transaction with parents picked uniformly at random
%among those transactions that are currently strongly preferred. Specifically, we can adopt the predicate
%used in the voting rule to determine eligible parents on which a node would vote positively, as follows:
%\[
%    \mathcal{E} = \{T: \forall T\in\mathcal{T}, \textsc{isStronglyPreferred}(T)\}.
%\]
%
%But this strategy will yield large sets of eligible parents, consisting mostly of historical, old transactions.
%When a node samples the transactions uniformly from $\mathcal{E}$, the resulting DAG
%will have large, ever-increasing fan-out. Because new transactions will have scarce progenies,
%the voting process will take a long time to build the required confidence on any given new transaction.
%
%In contrast, efforts to reduce fan-out and control the shape of the DAG by selecting the recent transactions at
%the decision frontier suffer from another problem. Most recent transactions will have very low confidence,
%simply because they do not have enough descendants. Further, their conflict sets may not be well-distributed
%and well-known across the network, leading to a parent attachment under a transaction that will never be supported.
%This means the best parent candidates lie somewhere near the frontier, but not too far deep in history.
%
%The adaptive parent selection algorithm chooses parents by starting at the DAG frontier and retreating towards the
%genesis vertex until finding an eligible parent.
%}{
%The frontier transactions are those that currently do not have any children in the node's
%$\mathcal{T}$.  The set of strongly preferred transactions $\mathcal{E}$ is
%filtered by an additional criterion that the confidence should be larger than
%zero if the transaction has known conflicts. This filters out 0-confidence transactions from
%unstable conflict sets.  If there exist $p$ transactions
%in the filtered $\mathcal{E}$, then the adaptive selection will yield these transactions.
%
%Otherwise, the algorithm tries the parents of the transactions in $\mathcal{E}$,
%thus increasing the chance of finding more stabilized transactions as it
%retreats.
%}
%%XXX issue no-ops after a timeout
%The retreating search is guaranteed to terminate when it reaches the genesis
%vertex.  Formally, the selected parents in this adaptive selection algorithm
%is:
%
%\begin{center}
%\begin{algorithmic}[1]
%    \small
%    \Function{parentSelection}{$\mathcal{T}$}
%    \tronly{}{\State $\mathcal{E} = \{T: \forall T\in\mathcal{T}, \textsc{isStronglyPreferred}(T)\}$.}
%        \State $\mathcal{E}' \coloneqq \{T: \forall T\in\mathcal{E}, |\mathcal{P}_T| = 1 \lor d(T) > 0\}$.
%        \State \Return $\{T: T \in \mathcal{E}' \land \forall T'\in\mathcal{T}, T \gets T', T' \notin \mathcal{E}'\}$.
%    \EndFunction
%    \captionof{figure}{Adaptive parent selection.}
%\end{algorithmic}
%\end{center}

\tronly{
\noindent
\editremove{\textbf{Optimizations}}
\editinsert{%
\subsection{Optimizations}}
We implement some optimizations to help the system scale.
First, we use \emph{lazy updates} to the DAG, because the recursive definition for confidence may otherwise require a costly DAG traversal.
We maintain the current $d(T)$ value for each active vertex on the DAG, and update it only when a descendant vertex gets a chit.
Since the search path can be pruned at accepted vertices, the cost for an update is constant if the rejected vertices have limited number of descendants and the undecided region of
the DAG stays at constant size.
Second, the conflict set could be very large in practice, because a rogue client can generate a large volume of conflicting transactions.
Instead of keeping a container data structure for each conflict set, we create a mapping from each UTXO to the preferred transaction that stands as the representative for the entire conflict set.
This enables a node to quickly determine future conflicts, and the appropriate response to queries.
Finally, we speed up the query process by terminating early as soon as the $\alpha k$ threshold is met, without waiting for $k$ responses.
}{}

\editinsert{%
\subsection{Safety}
Lastly, the safety guarantees of Snowball can be mapped to those of Avalanche, which is a concrete instantiation of Snowball using a directed acyclic graph to amortize cost. 
We note that the structure of the Avalanche DAG itself does not correspond to votes, which is a subtle difference between other consensus protocols that make usage of a DAG. The DAG is merely a performance optimization, and is itself entirely orthogonal to the consensus process.
}

\editinsert{
\subsection{Liveness}
Avalanche introduces a DAG structure that entangles the fate of unrelated conflict sets, each of which is a single-decree instance.
This entanglement embodies a tension: attaching a virtuous transaction to undecided parents helps propel transactions towards a decision, while it puts transactions at risk of suffering liveness failures when parents turn out to be rogue.
We can resolve this tension and provide a liveness guarantee with the aid of two mechanisms.

First we adopt an adaptive parent selection strategy, where transactions are attached at the live edge of the DAG, and are retried with new parents closer to the genesis vertex. This procedure is guaranteed to terminate with uncontested, decided parents, ensuring that a transaction cannot suffer liveness failure due to contested, rogue transactions. 
A secondary mechanism ensures that virtuous transactions with decided ancestry will receive sufficient chits. Correct nodes examine the DAG for virtuous transactions that lack sufficient progeny and emit nop transactions to help increase their confidence.
With these two mechanisms in place, it is easy to see that, at worst, Avalanche will degenerate into separate instances of Snowball, and thus provide the same liveness guarantee for virtuous transactions.
}

\editinsert{%
\subsection{Communication Complexity}
Communication complexity for Avalanche is more subtle.
Let the DAG induced by Avalanche have an expected branching factor of $p$, corresponding to the width of the DAG, and determined by the parent selection algorithm.
Given the $\beta$ decision threshold, a transaction that has just reached the point of decision will have an associated progeny $\mathcal{Y}$.
Let $m$ be the expected depth of $\mathcal{Y}$.
If we were to let the Avalanche network make progress and then freeze the DAG at a depth $y$,
then it will have roughly $py$ vertices/transactions, of which $p(y - m)$ are decided in expectation.
Only $pm$ recent transactions would lack the progeny required for a decision.
For each node, each query requires $k$ samples, and therefore the total message cost per transaction is in expectation $(pky) / (p(y - m)) = ky/(y-m)$.
Since $m$ is a constant determined by the undecided region of the DAG as the system constantly makes progress, message complexity per node is $O(k)$, while the total complexity is $O(kn)$.
}

\section{Evaluation}\tronly{}{\vspace{-0.5em}}
\label{sec:evaluation}
\input{sections/evaluation}

\section{Related Work}\tronly{}{\vspace{-0.5em}}
\label{sec:related-work}
\input{sections/related-work}

\section{Conclusion}\tronly{}{\vspace{-0.5em}}
\label{sec:conclusions}
% This paper introduced a novel family of consensus protocols, coupled with the appropriate mathematical tools for analyzing them.
% Unlike classical or Nakamoto-style consensus, these protocols are highly efficient and robust. They scale well, achieve high throughput and low latency, work without precise membership knowledge, as well as degrade gracefully under catastrophic adversarial attacks, beyond current bounds.

% This work opens up new avenues of research. Chief among these is relaxing the synchrony assumption in this paper to accommodate asynchrony. We believe that the protocols and analysis techniques established in this paper will form the foundation for new venues of research.
% %and provide a strong safety guarantee,
% %though they achieve these properties by not guaranteeing liveness for conflicting transactions.
% %We have illustrated how they can be used to implement a Bitcoin-like payment system, that achieves 3400 tps in a geo-replicated setting.

This paper introduced a novel family of consensus protocols, coupled with the appropriate mathematical tools for analyzing them.
These protocols are highly efficient and robust, combining the best features of classical and Nakamoto-style consensus.
They scale well, achieve high throughput and quick finality, work without precise membership knowledge, and degrade gracefully under catastrophic adversarial attacks.

There is much work to do to improve this line of research. Chief among these is relaxing the synchrony assumption in this initial paper. We believe strong guarantees are possible even under asynchrony, but the details remain to be worked out. Another improvement would be to characterize the system's guarantees under an adversary whose powers are realistically limited, whereupon performance would improve even further.
Overall, we hope that the protocols and analysis techniques presented here add to the arsenal of the distributed system developers and provide a foundation for new lightweight and scalable mechanisms.

%\section*{Acknowledgments}\tronly{}{\vspace{-0.5em}}
%\label{sec:related-work}
%We are grateful to the many individuals who provided feedback on earlier drafts. 
% Bobby Kleinberg, Vitalik Buterin, ...

%latex here should be the name of your bibtex file minus '.bib'
\bibliographystyle{acm}
\bibliography{paper}
\newpage
\begin{appendices}
\section{Analysis}\tronly{}{\vspace{-0.5em}}
\label{sec:full-analysis}
In this appendix, we analyze Slush, Snowflake, Snowball, and Avalanche.
\tronly{}{For brevity, we leave details and proofs to an accompanying tech report~\cite{avalanche}.}

\paragraph{Network Model} We assume a synchronous communication network, where at each time step $t$,
a global scheduler chooses a single correct node uniformly at random.

\paragraph{Preliminaries} Let $c = |\mathcal{C}|$ and $b = |\mathcal{B}|$;
let $u \in \mathcal{C}$ be any correct node;
let $k \in \mathbb{N}_+$;
and let $\alpha \in \mathbb{R} = (1/2, 1]$.
We let $\mathtt{R}$ (``red'') and $\mathtt{B}$ (``blue'') represent two generic conflicting choices.
Without loss of generality, we focus our attention on counts of $\mathtt{R}$, i.e.\ the total number of nodes that prefer $\mathtt{R}$.

Each network query of $k$ peers corresponds to a sample without replacement out of a network of $n$ nodes, also referred to as a hypergeometric sample.
We let the random variable $\mathcal{H}_{\mathcal{N}, x}^k \rightarrow [0, k]$ denote the resulting counts of $\mathtt{R}$ (ranging from 0 to $k$), where $x$ is the total count of $\mathtt{R}$ in the population. The probability that the query achieves the required threshold of $\alpha k$ or more votes is given by:

\begin{equation}
\small
P(\mathcal{H}_{\mathcal{N}, x}^k \geq \alpha k) = \sum_{j = \alpha k}^{k} \frac{{x \choose j} {n - x \choose k - j}}{{n \choose k}}
\label{eq:hypergeometric}
\end{equation}

\tronly{
\paragraph{Tail Bound} We can reduce some of the complexity in Equation~\ref{eq:hypergeometric} by introducing a bound on the hypergeometric distribution induced by $\mathcal{H}^k_{\mathcal{C},x}$.
Let $p=x/c$ be the ratio of support for $\mathtt{R}$ in the population.
The expectation of $\mathcal{H}_{\mathcal{C}, x}^k$ is exactly $kp$.
Then, the probability that $\mathcal{H}_{\mathcal{C}, x}^k$ will deviate from the mean by more than some small constant $\psi$ is given by the Hoeffding tail bound \cite{hoeffding1963probability}, as follows,
\begin{equation}
\begin{split}
    P(\mathcal{H}_{\mathcal{C}, x}^k \leq (p-\psi)k) &\leq e^{-k\mathcal{D}(p-\psi, p)}
    % \leq e^{-2\psi^2k}
\end{split}
\end{equation}
where $\mathcal{D}(p-\psi, p)$ is the Kullback-Leibler divergence, measured as
\begin{equation}
\begin{split}
    \mathcal{D}(a, b) &= a \log \frac{a}{b} + (1 - a) \log \frac{1 - a}{1 - b}
\end{split}
\end{equation}}{}

\subsection{Analysis of Slush}\tronly{}{\vspace{-0.5em}}
%\label{sec:analysis}
Slush operates in a non-Byzantine setting; that is, $\mathcal{B} = \emptyset$ and thus $\mathcal{C} = \mathcal{N}$ and $c = n$.
In this section, we will show that
(R1) Slush converges to a state where all nodes agree on the same color in finite time almost surely (i.e.~with probability~1); 
(R2) provide a closed-form expression for speed towards convergence; and 
(R3) characterize the minimum number of steps per node required to reach agreement with probability $\ge 1-\varepsilon$.

\begin{figure}
    \small
\begin{algorithmic}[1]
    \State initialize $u.\codecolor \in \{\texttt{R},\texttt{B}\}$ for all $u\in\mathcal{N}$
    \For{$t = 1$ to $\phi$}
        \State $u \coloneqq \Call{sample}{\mathcal{N}, 1}$
        \State $\mathcal{K} \coloneqq \Call{sample}{\mathcal{N}\textbackslash u, k}$
        \State $P \coloneqq \texttt{[}v.\codecolor\quad\textbf{for}\ v \in \mathcal{K}\texttt{]}$
        \For{$\codecolor' \in \{\texttt{R}, \texttt{B}\}$}
            \If{$P.\Call{count}{\codecolor'} \ge \alpha\cdot k$}
                \State $u.\codecolor \coloneqq \codecolor'$
            \EndIf
        \EndFor
    \EndFor
\captionof{figure}{Slush run by a global scheduler.}\label{fig:slush_protocol_simulator}
\end{algorithmic}
\end{figure}

\tronly{
The procedural version of Slush in Figure~\ref{fig:slush-loop} made use of a parameter $m$, the number of rounds that a node executes Slush queries. To derive this parameter, we transform the protocol execution from a procedural and concurrent version to one carried out by a scheduler, shown in Figure~\ref{fig:slush_protocol_simulator}.
What we ultimately want to extract is the total number of rounds $\phi$ that the scheduler will need to execute in order to guarantee that the entire network is the same color, whp.
}{}

We analyze the system mainly using standard Markov chain techniques. Let $\mathbf{X} = \{\mathit{X_1}, \mathit{X_2}, \dots\, \mathit{X}_\phi\}$ be a discrete-time Markov chain.
Let $s_i$, $i \in [0, c]$, represent a state in this Markov chain, where the count of $\mathtt{R}$ is $i$ and the count of $\mathtt{B}$ is $c - i$.
Let $\mathit{M}$ be the transition probability matrix for the Markov chain.
Let $q(s_i) = \mathit{M}(s_i, s_{i-1})$, $p(s_i) = \mathit{M}(s_i, s_{i+1})$, and $r(s_i) = \mathit{M}(s_i, s_{i})$, where $p(s_0) = q(s_0) = p(s_c) = q(s_c) = 0$.
This is a birth-death process where the number of states is $c$ and the two endpoints, $s_0$ and $s_c$, are absorbing states, where all nodes are red and blue, respectively.

To extract concrete values, we apply the following probabilities to $q(s_i)$ and $p(s_i)$, which capture the probability of picking a red node and successfully sampling for blue, and vice versa.
\begin{equation*}
\begin{split}
&p(s_i) = \frac{c-i}{c}P(\mathcal{H}_{\mathcal{C}, i}^{k} \geq \alpha k)\\
&q(s_i) = \frac{i}{c}P(\mathcal{H}_{\mathcal{C}, c-i}^{k} \geq \alpha k)
\end{split}
\end{equation*}

\begin{lemma}[R1]
Slush reaches an absorbing state in finite time almost surely.
\end{lemma}
\begin{proof}
% This follows in a straightforward way from the construction of $\mathbf{X}$.
% More precisely,
Let $s_i$ be a starting state where $i \leq c - \alpha k$. All such states $s_i$ then have a non-zero probability of reaching absorbing state $s_0$ for all time-steps $t \geq i$. States where $i > c - \alpha$ have no possible threshold majority for $\mathtt{B}$, and all timesteps $t \leq i$ do not allow enough transitions to ever reach $s_0$. 
Under these conditions, the probability of absorption is strictly $> 0$, and therefore Slush converges in finite steps. 
\end{proof}

% First we note that this Markov chain is \emph{absorbing}, that is, every state in the chain can
% reach an absorbing state.  Moreover, because the chain has a finite number of states, the expected
% number of steps of reaching an absorbing state is finite.  Note also that the chain is symmetric,
% and one can think of this Markov chain as a generalization of the Gambler's Ruin Problem, for which
% we have closed form expressions for the expected number of steps.  Finally, note that our Markov
% chain should absorb faster than Gambler's Ruin because the ``pull'' toward an absorbing state increases
% with reduced distance to an absorbing state.

% R2 can be shown using the following lemma, which computes the expected number of steps to reach $s_0$, starting from $s_i$.

% \begin{comment}
% The next two lemmas establish recurrence relations that enable us to compute both the probability of absorbing in the all-red state and the expected time to reach that state.
% \tronly{
% \begin{lemma}
% \textbf{Probability of Deciding $\mathtt{R}$}. Let $\mathscr{A}(s_i)$ be the event of absorption at state $s_0$ starting from state $s_i$.
% Let
% \begin{equation}
%     \small
%     \mathscr{Y}_c(h) = \sum_{l = s_h}^{s_{c-1}}\prod_{j = s_1}^{l}\frac{q(j)}{p(j)}
% \end{equation}
% Then,
% \begin{equation}
%     \small
%     P(\mathscr{A}(s_i)) = \mathscr{Y}_c(i) / \mathscr{Y}_c(0)
% \end{equation}
% % \begin{comment}
% % \begin{equation}
% % \begin{split}
% %     P(\mathscr{A}(s_i)) &= \mathscr{U}_n^{-1}\sum_{l = s_i}^{s_{c-1}}\prod_{j = s_1}^{l}\frac{q(j)}{p(j)}
% % \end{split}
% % \end{equation}
% % where
% % \begin{equation}
% % \begin{split}
% %     \mathscr{U}_n = \sum_{l = s_0}^{s_{c-1}}\prod_{j = s_1}^{l}\frac{q(j)}{p(j)}
% % \end{split}
% % \end{equation}
% % \end{comment}
% \label{lemma:absorption_theta_0}
% \end{lemma}
% \begin{proof}
% For $i \in [1, c)$, the absorbing probability satisfies the recurrence relation
% \begin{align}
%     P(\mathscr{A}(s_i)) =\ q(s_i)P(\mathscr{A}(s_{i-1})) &+ p(s_i)P(\mathscr{A}(s_{i+1})) \nonumber \\
%     &+ r(s_i)P(\mathscr{A}(s_i))
% \end{align}
% with boundary conditions $P(\mathscr{A}(s_0)) = 1$ and $P(\mathscr{A}(s_{c})) = 0$. Solving that recurrence relation yields the result.
% \end{proof}}{}
% \end{comment}

\tronly{
\begin{lemma}[R2]
\label{lemma:expected_absorption}
\textbf{Expected Rounds to Deciding $\mathtt{B}$.} Let $\mathit{U}_{s_i}$ be a random variable that expresses the number of steps to reach $s_0$ from state $s_i$.
Let
\begin{equation}
    \small
    \mathscr{Y}_c(h) = \sum_{l = s_h}^{s_{c-1}}\prod_{j = s_1}^{l}\frac{q(j)}{p(j)}
\end{equation}
Then,
\begin{equation}
\small
\begin{split}
    \mathbb{E}[\mathit{U}_{s_i}] &= \mathbb{E}[\mathit{U}_{s_1}]\sum_{l=s_0}^{s_{i-1}}\prod_{j = s_1}^{l}\frac{q(j)}{p(j)}\\
    &- \sum_{l=s_0}^{s_{i-1}}\sum_{j=s_1}^{l}\frac{1}{p(j)}\prod_{v=j}^{l}\frac{q(v+1)}{p(v+1)}
\end{split}
\label{equation:expected_convergence}
\end{equation}
where
\begin{equation}
\small
\begin{split}
    \mathbb{E}[\mathit{U}_{s_1}] &= \mathscr{Y}_c(0)^{-1}\sum_{l=s_1}^{s_{c-1}}\sum_{j=s_1}^{l}\frac{1}{p(j)}\prod_{v=j}^{l}\frac{q(v+1)}{p(v+1)}
\end{split}
\end{equation}
\end{lemma}
\begin{proof}
The expected time to absorption in $s_0$ can be modeled using the following recurrence
equation:
\begin{align}
        \mathbb{E}[\mathit{U}_{s_i}] = q(s_i)\mathbb{E}[\mathit{U}_{s_{i-1}}] &+ p(s_i)\mathbb{E}[\mathit{U}_{s_{i+1}}]\nonumber \\
        &+ r(s_i)\mathbb{E}[\mathit{U}_{s_{i}}]
\end{align}
for $i \in [1, c-1]$, with boundary condition $\mathbb{E}[\mathit{U}_{s_0}] = 0$. 
Solving for the recurrence relation yields the result. 
\end{proof}
}{}

% \begin{lemma}
% Let $\phi \to \infty$. The Slush protocol enters an absorbing state at $\phi^{'}$, where $\phi^{'}$ is finite.
% \end{lemma}
% \begin{proof}
%     The chain $M$ is irreducible and aperiodic, and therefore converging. \tronly{More precisely, applying $p(s_i)$ and $q(s_i)$ computed above to Lemma~\ref{lemma:expected_absorption}, we see that $\phi^{'}$ is finite over all network configurations where count of $\mathtt{R}$ or count of $\mathtt{B}$ $\geq \alpha k$.}{}
% \end{proof}
% \tronly{

% \begin{lemma}
% \textbf{``Pull'' from State $s_i$.} Let $g(s_i) = q(s_i)/p(s_i)$ represent the ratio of the speed of absorption at $s_0$, starting at state $s_i$. Then, a lower-bound of $g(s_i)$ is
% \begin{equation}
% \small
% (c-i)^{\tau - 1}\ i^{1-\tau} \leq g(s_i)
% \end{equation}
% where $\tau = (2\alpha - 1) k$.
% \label{lemma:speed_absorption}
% \end{lemma}
% \begin{proof}
%     This follows directly from the lower bound of the binomial coefficient, i.e. $(z/k)^k \leq {n \choose k}$, for some $z \geq k \geq 1$, applied to the expression of $g(s_i)$.
% \end{proof}
% This inequality provides some insight into the convergence process.
% The bivalent state $s_{c/2}$, where the network is divided equally between the two colors, poses the worst case in terms of expected rounds to convergence.
% As the network diverges from the 50/50 state $s_{c/2}$, as measured by $i$, the probability of moving further in that direction increases at least polynomially in $i$.
% }{}

\begin{table}[h!]
	\centering
	\begin{tabular}{llllll}
		$\mathcal{C}$ Size   & 600    & 1200   & 2400   & 4800   & 9600    \\ \hline
		Exp. Conv. & 12.66 & 14.39 & 15.30 & 16.43 & 18.61
	\end{tabular}
    \caption{Expected number of per-node-iterations to convergence starting at worst-case (equal) $\mathcal{C}$ network split, in the case of $k = 10$, $\alpha=0.8$. Standard deviation for all samples is $\leq 2.5$. Computed via a Monte Carlo simulation.}
	\label{table:growth_worst_case}
\end{table}
Table~\ref{table:growth_worst_case} shows the latency (i.e. number of timesteps per-node) to reach the $\mathtt{B}$ deciding state starting from the metastable 50/50 split\tronly{, in other words, value $\mathbb{E}[\mathit{U}_{s_{c/2}}]/c$}{}. 
% \Jon{How was the table computed? Monte Carlo or otherwise numerically?}
As the network size $c$ doubles, expected convergence grows only linearly.

Finally, to demonstrate R3, we need to compute the first timestep of the Markov chain $M$ that yields a probability of $\leq \varepsilon$ for all transient states.
% Since the last $\alpha k - 1$ states (i.e. $s_{c-\alpha k}$, $\dots$, $s_{c}$) correspond to regions where nodes can no longer receive a sample of $\alpha k$ or more successes for the opposite color, the first of such states can serve as the desired target \tronly{state for computing the minimal time to $\varepsilon$-convergence.}{.}
This computation has no simple closed-form approximation, therefore we must iteratively raise the transition probability matrix $M$ until the first timestep $t$ that achieves probability $\leq \varepsilon$ for all states $s_i$ where $s_{\alpha k} < s_i < s_{c - \alpha k}$.
% \tronly{Figure~\ref{fig:expected_time_to_epsilon_convergence_slush} shows the expected time to reach $s_{c - \alpha k}$ whp, over various network sizes, with fixed $k = 10, \alpha = 0.8$.

% \begin{figure}
% \centering
% % file: gossip-chain/paper/kevinscripts/expected_time_to_epsilon_convergence_slush.py
% \tronly{\includegraphics[width=\linewidth]{figures/epsilon-convergence-large.pdf}}
%         {\includegraphics[width=\linewidth]{figures/epsilon-convergence.pdf}}
% \caption{Necessary parameter $\phi/c$ required in order to achieve $\varepsilon$-convergence in the Slush protocol, i.e.\ every node has reached the same color whp. We fix $k = 10$, $\alpha = 0.8$.}
% \label{fig:expected_time_to_epsilon_convergence_slush}
% \end{figure}
% }{}

\tronly{
The next two additional results provide some useful intuition to the underlying stochastic processes of Slush and subsequent protocols. 
The probability of a single successful query as a function of the total number of $\mathtt{R}$-supporting nodes in the network is shown in Figure~\ref{fig:varyingk}.
This probability decreases rapidly as the value of $k$ increases.
The network topples over faster if the value of $k$ is smaller due to the larger amount of random query perturbance.
Indeed, for Slush, $k=1$ is optimal.
Later we will see that for the Byzantine case, we will need a larger value for $k$.
}{}

\begin{figure}
\centering
% file: gossip-chain/paper/plots/trend-divergence.py
\tronly{\includegraphics[width=\linewidth]{figures/trend-diver-large.pdf}}
        {\includegraphics[width=\linewidth]{figures/trend-diver.pdf}}
\caption{Probability of successful samples vs.\ network splits.}
\label{fig:trend_divergence}
\end{figure}

Figure~\ref{fig:trend_divergence} shows the divergence of the probabilities of successful samples over various network splits. When the network is evenly split, the probabilities are equal for either color, but very rapidly diverge as the network becomes less balanced.

\begin{comment}
In this section, we have shown three results. First, we showed that Slush convergences to an absorbing state in finite time. Second, we provided a closed-form expression to calculate the expected time to convergence given any starting state, which provides useful information about the effectiveness of Slush to quickly terminate. Third, we discussed the iterative method, as well as some supporting numerical values, to calculate the minimum number of timesteps per node to guarantee absorption with probability $\geq 1 - \varepsilon$. 
\end{comment}

\tronly{
\begin{figure}[h!]
\centering
% file: gossip-chain/paper/kevinscripts/plots/varying_k_theta_0.py
\tronly{\includegraphics[width=\linewidth]{figures/varying-k-theta0-large.pdf}}
        {\includegraphics[width=\linewidth]{figures/varying-k-theta0.pdf}}
\caption{Probability of a single successful majority-red sample with varying $k$, $\alpha = 0.8$, and successes ranging from half the network to full-support.}
\label{fig:varyingk}
\end{figure}
}{}

\subsection{Safety Analysis of Snowflake}\tronly{}{\vspace{-0.5em}}
Snowflake, whose pseudocode is shown in Figure~\ref{fig:snowflake_simulator}, is similar to Slush, but has three key differences.\tronly{}{The scheduler version of Snowflake, omitted for brevity, is analogous to that for Slush, but has three key differences.} 
First, the sampled set of nodes includes Byzantine nodes.
Second, each node also keeps track of the total number of consecutive times it has sampled a majority of the same color.
We also introduce a function called $\mathcal{A}$, the adversarial strategy, that takes as parameters the entire space of correct nodes as well as the chosen node $u$ at time $t$, and as a side-effect, modifies the set of nodes $\mathcal{B}$ to some arbitrary configuration of colors.

The birth-death chain that models Snowflake, shown below, is nearly identical to that of Slush, with the added presence of Byzantine nodes. 
\begin{equation*}
% \small
\begin{split}
&p(s_i)=\frac{c-i}{c}P(\mathcal{H}_{\mathcal{N}, i}^{k} \geq \alpha k),\\
&q(s_i)=\frac{i}{c}P(\mathcal{H}_{\mathcal{N}, c-i+b}^{k} \geq \alpha k)
\end{split}
\end{equation*}
In Slush, the metastable state of the network was a 50/50 split at Markov state $s_{c/2}$. 
In Snowflake, Byzantine nodes have the ability to compensate for deviations from $s_{c/2}$ up to a threshold proportional to their size in the network.
Let the Markov chain state $s_{ps} > s_{c/2}$ be the state after which the probability of entering the $\mathtt{R}$-absorbing state becomes greater than the probability of entering the $\mathtt{B}$-absorbing state given the entire weight of the Byzantine nodes. 
Symmetrically, there exists a phase shift point for \texttt{B} as well. 
All states between these two phase-shift points are Byzantine controlled, whereas all other states are metastable.

In this section, we analyze the two conditions that ensure the safety of Snowflake. The first condition (C1) requires demonstrating that that there exists a point of no return $s_{ps + \Delta}$ after which
the maximum probability under full Byzantine activity of bringing the system back to the phase-shift point is less than $\varepsilon$.

The second condition (C2) requires the construction of a mechanism to ensure that a process can only commit to a color if the system is beyond the minimal point of no return, whp. 
As we show later, satisfying these two conditions is necessary and sufficient to guarantee safety of Snowflake whp.

\begin{figure}
\begin{algorithmic}[1]
    \State initialize $\forall u, u.\codecolor$,
    \State initialize $\forall u, u.\codecount \coloneqq 0$
\For{$t = 1$ to $\phi$}
    \State $u \coloneqq \Call{sample}{\mathcal{C}, 1}$
    % \State $u$.total++
    \State $\mathcal{A}(u, \mathcal{C})$
    \State $\mathcal{K} \coloneqq \Call{sample}{\mathcal{N}\textbackslash u, k}$
    \State $P \coloneqq \texttt{[}v.\codecolor\quad\textbf{for}\ v \in \mathcal{K}\texttt{]}$
    \For{$\codecolor' \in \{\texttt{R}, \texttt{B}\}$}
    \If{$P.\Call{count}{\codecolor'} \ge \alpha\cdot k$}
        \If{$\codecolor' \neq u.\codecolor$}
            \State $u.\codecolor \coloneqq \codecolor'$, $u.\codecount \coloneqq 0$
        \Else
            \State $\texttt{++}u.\codecount$
        \EndIf
    \EndIf
    \EndFor
\EndFor
\captionof{figure}{Snowflake run by a global scheduler.}\label{fig:snowflake_simulator}
\end{algorithmic}
\end{figure}

\paragraph{Properties of Sampling Parameter $k$.} Before constructing and analyzing the two conditions that satisfy the safety of Snowflake, we introduce some important observations of the sample size $k$. 
The first observation of interest is that the hypergeometric distribution approximates the binomial distribution for large-enough network sizes.
If $x$ is a constant fraction of network size $n$, then:

% This is formalized in the following lemma:
% \begin{lemma}
% For $x$ as a fixed ratio of network support with increasing $n$, we have:
\begin{equation}
    \small
    \begin{split}
        \lim_{n \rightarrow \infty} P(\mathcal{H}_{\mathcal{N}, (xn)}^k \geq \alpha k) = \sum_{j=\alpha k}^k {k \choose j}x^k (1 - x)^{k - j}
    \end{split}
\end{equation}
% \end{lemma}
% \begin{comment}
% \begin{proof}
%     The proof follows in a straightforward way by applying the limit as $n \rightarrow \infty$ to the hypergeometric expression. For all $j = [\alpha k, k]$, we have:
%     \begin{equation}
%         \begin{split}
%             & \lim_{n \rightarrow \infty} {x \choose j}{n - x \choose k - j}/{n \choose k}\\
%             = & \lim_{n \rightarrow \infty} {k \choose j} \prod_{l = 1}^{j} \frac{x - j + l}{n - j + l} \prod_{m = 1}^{k - j} \frac{n - x - k + j + m}{n - k + m}\\
%             = & {k \choose j} (x/n)^{j} (1 - (x/n))^{k - j}
%         \end{split}
%     \end{equation}
%     and therefore, summed over all $j$, we have our approximation. 
% \end{proof}
% \end{comment}
\noindent This dictates that for large enough network sizes, the probability of success for any single sample becomes a function dependent only on the values of $k$ and $\alpha$, unaffected by further increases of network size. 
The result has consequences for the scalability of Snowflake and subsequent protocols. For sufficiently large and fixed $k$, the security of the system will be approximately equal for all network sizes $n > k$.
% Kevin: The next result bounds how fast the approximation occurs, but we don't widely make usage of this bound apart from providing intuition. Therefore, I think removing it completely is the sensible thing to do. 
% Following up, the next result bounds this approximation for small network sizes.
% \begin{lemma}
%     The probabilities of the hypergeometric experiment, given a fixed network sample parameter $k$ and $\alpha$ threshold, are independent of network size, up to a constant approximation factor $L$.
%     \label{lemma:lconvergence}
% \end{lemma}
% \tronly{
% \begin{proof}
%     To show this, we fix $k$ and $\alpha$ to two constants, and then compute the deviation of the probabilities of the hypergeometric as the network size (and successes $x$) increases by some factor $\gamma$. More precisely,
%     \begin{equation}
%         \small
%         \begin{split}
%             L = \lim_{\gamma \rightarrow \infty} \left( \sum_{j = \alpha k}^{k} \frac{{x\gamma \choose j} {c\gamma - x\gamma \choose k - j}}{{c\gamma \choose k}} \middle/ \sum_{j = \alpha k}^{k} \frac{{x \choose j} {c - x \choose k - j}}{{c \choose k}} \right)
%         \end{split}
%     \end{equation}
%     Rearranging, and using the smallest factor in the summation ($j = k$), we see that
%     \begin{equation}
%         \small
%         \begin{split}
%             L \geq L^{'} = \xi \lim_{\gamma \rightarrow \infty} \frac{{x\gamma \choose k}}{{c\gamma \choose k}}, \text{where}\ \xi = \frac{{c \choose k}}{{x \choose k}}
%         \end{split}
%     \end{equation}
%     For the lower bound of $x = c/2$ and the upper bound of $x = c$, we have that $2^{-k}\xi \leq L^{'} \leq \xi$.
%     % A perhaps cleaner, but less precise, way of exposing the relationship between $k$ and the network size is by using the Hoeffding tail bound, which shows that $P(\mathcal{H}_{\mathcal{C}, x}^k \leq ((x/c) - \psi)k) \leq e^{-2k((x/c) - \psi)^2}$ \cite{hoeffding1963probability}, where $\psi$ is a constant. As the network size increases, it has no effect on the probability distribution of the outcomes. Contrary to our more precise approximation of the bound above, the Hoeffding bound entirely removes 
% \end{proof}}{}
% \noindent The prior two lemmas indicate that Snowflake scales well with increasing network size. 
% In Figure \ref{fig:plot_k_needed_for_same_security_over_increasing_network} we show how $L$ converges over various network sizes. For networks of size $\geq 1000$, probabilities of various outcomes change minimally for the same $k$ and $\alpha$ parameters.
% \begin{figure}
% % file: gossip-chain/paper/kevinscripts/plot_k_needed_for_same_security_over_increasing_network.py
% \includegraphics[width=\linewidth]{./plot_k_needed_for_same_security_over_increasing_network.pdf}
% \caption{Ratios of the probabilities of successes over various network sizes, with fixed $k = 10$, $\alpha = 0.8$, and $x = 0.8c$. \texttt{RVR: the labels are too small, and we don't need a title above the graph.}}
% \label{fig:plot_k_needed_for_same_security_over_increasing_network}
% \end{figure}

We now examine how the phase shift point behaves with increasing $k$. %At the maximum allowable $k = n$, the phase shift point $s_{ps}^{*}$, reaches its maximal value.
Whereas $k=1$ is optimal for Slush (i.e. the network topples the fastest when $k = 1$), we see that Snowflake's safety is maximally displaced from its ideal position for $k=1$, reducing the number of feasible solutions.
Luckily, small increases in $k$ have an exponential effect on $s_{ps}$, creating a larger set of feasible solutions. More formally: 
\begin{lemma}
$s_{ps}$ approaches $s_{n/2}$ exponentially fast as $k$ approaches $n$, and at $k = n$, $s_{ps} = s_{n/2}$.
\end{lemma}
\begin{proof}
At $s_{ps}$, the ratio $p(s_i)/q(s_i)$ is equal to 1, by definition. Substituting the solution $i = c/2 + b/2$ into this equation yields the result at $k=n$.
Computing the differential in $s_{ps}$ between any chosen $k$ and $k+1$ yields the exponential trend.
\end{proof}

% \noindent Figure~\ref{fig:plt_various_k_phase_shift} shows the exponential relationship between decreasing $k$ and $s_{ps}$. The graph shows that small-but-not-tiny values of $k$ lead to large feasible regions.

% \begin{figure}[h!]
%     \centering
%     % file: gossip-chain/paper/kevinscripts/plots/plot_various_k_phase_shift.py
% \tronly{\includegraphics[width=\linewidth]{figures/k-phase-shift-large.pdf}}
%         {\includegraphics[width=\linewidth]{figures/k-phase-shift.pdf}}
%     \caption{Phase shift state of Snowflake/Snowball vs. $k$ values (x-axis), with $n = 2000$, $\alpha = 0.8$ and $b = 20\%$ of $n$. The phase-shift is indexed starting from $c/2$. \Jon{I really didn't get the story here. It seems like cool graphical representations that would really help me understand the underlying arguments, but I missed something.}}
%     \label{fig:plt_various_k_phase_shift}
% \end{figure}

% To analyze the probability of the Byzantine nodes triggering safety violation, we use the same state space $\forall i\,\,\{s_i\}$ as in Slush, with the difference that the transition probabilities are now dependent on the strategy chosen by the Byzantine nodes, complicating analysis.
% We sidestep this difficulty by conservatively focusing on the states that are irreversible under \emph{any} Byzantine strategy.

% {
% \color{red}
% Let $\Delta$ represent the network bias towards red, represented by state $s_{c/2 + \Delta}$.
% Let $s_{ps} \geq s_{c/2}$ represent the phase-shift state in the system, as defined previously.
% We are interested in determining network parameters $k$ and $\alpha$ for an anticipated maximum number of Byzantine nodes $b$. Then, $\exists \Delta$ where $s_{c/2 + \Delta} > s_{ps}$ s.t.\ the probability of the system reverting back to $s_{ps}$ is less than or equal to $\varepsilon$ for some upper bound $\phi$. Figure~\ref{fig:states_feasible_solutions} summarizes the region of interest,
% where the desired $\Delta$ is the \emph{point of no return} after which the network is not likely to revert whp. We refer to states that meet this condition with regards to $s_{ps}$ as \emph{irreversible}.
% }

\paragraph{Satisfiability of Safety Criteria}
Let $\varepsilon$ be a system security parameter of value $\leq 2^{-32}$, typically referred to as \emph{negligible} probability.

% $\textsc{isAccepted}$ be the function which returns true if $u$ samples the same color successfully for at least $\beta$ consecutive times.
% A crucial property of $\textsc{isAccepted}$ is that it returns true 
% We are interested in the set of outcomes where, at time $\phi$, there exists at least two distinct correct nodes $u$ and $v$ for which the \textsc{isAccepted} predicate returns true for different colors. \tronly{Formally, we are interested in the following game:
% \begin{equation}
%     \begin{cases}
%         \textrm{win} & \textbf{if}\ t = \phi \land \exists u, v, C, C': u \ne v \land C \ne C' \Rightarrow \\
%                      & u.\textsc{isAccepted}(C) \land v.\textsc{isAccepted}(C') \\
%         \textrm{lose} &  \textbf{otherwise}
%     \end{cases}
% \end{equation}}{}
We now formalize and analyze the safety condition C1. We refer to all states $s_i$ that satisfy condition C1 as \emph{states of feasible solutions}. 
For the system to guarantee safety, there must exist some state $s_{c/2 + \Delta} > s_{c/2 + b} > s_{ps}$ that implies $\varepsilon$-irreversibility in the birth-death chain.
In other words, once the system reaches this state, then it reverts back with only negligible probability, under \emph{any} Byzantine strategy. Figure~\ref{fig:states_feasible_solutions} summarizes the region of interest.
Our goal is to determine values for tunable system parameters to achieve this condition, given target network size $n$, maximum Byzantine component $b$, and security parameter $\varepsilon$, fixed by the system designer.

\begin{figure}[h]
    \begin{center}
\input{figures/analysis1}
\caption{Representation of the space of feasible solutions. Outside the phase-shift point, the network builds a bias towards one end. At the $\Delta$ point, this bias is $\varepsilon$-irreversible.
% \Jon{I really didn't get the story here. It seems like cool graphical representations that would really help me understand the underlying arguments, but I missed something.}
}
\label{fig:states_feasible_solutions}
    \end{center}
\end{figure}

\noindent To ensure C1, we must find $\Delta$ such that the following condition holds:
\begin{equation}
\begin{split}
    \exists \Delta, \text{where}\ s_{c/2 + \Delta} &> s_{ps},\ s.t.\ \forall t \leq \phi:\\
    \varepsilon &\geq (V_c[s_{c/2 + \Delta}, s_{ps}])^t
\end{split}
\end{equation}
where $V_c$ is the $(c+1)^2$ probability transition matrix for the system.
In probability theory terminology, the statement above computes the $t$-step hitting probability, where the starting state is $s_{c/2 + \Delta}$ and the hitting state is $s_{ps}$.
Satisfying this first constraint, that is, moving the ensemble of nodes to a decision, is not sufficient. The nodes have to be able to assure \emph{themselves} that they can safely decide, which means they must ensure condition C2.
While there are multiple ways of constructing the predicate that ensures this condition, the one we adopt in Snowflake involves $\beta$ consecutive successful queries, such that $P(\textsc{isAccepted} \,|\, s_i \leq s_{c/2 + \Delta}) \leq \varepsilon$. 
To achieve this threshold probability of failure for C2, we solve for the smallest $\beta$ that meets inequality
$P(\mathcal{G}(\mathtt{p}, \phi/c) \geq \beta) \leq \varepsilon$
where $\mathtt{p} = P(\mathcal{H}_{\mathcal{N}, i}^k \geq \alpha k)$ and $\mathcal{G}(\mathtt{p}, \phi/c)$ is a random variable which counts the largest number of consecutive same-color successes for a run over $\phi/c$ trials.
Solving for the minimum $i$ yields $\beta$.

The task of the system designer, finally, is to derive values for $\alpha$ and $\beta$, given desired $n, b, k, \varepsilon$.
Choice of $\alpha$ immediately follows from $b$ and is $(n - b)/ n$, the maximum allowable size to tolerate $\leq b$ Byzantine nodes.
Solving for $\Delta$ with a closed form expression would be ideal but unfortunately is difficult, hence we employ a numerical, iterative search.
Since $k$ and $\beta$ are mutually dependent, a system designer will typically fix one and search for the other.
Because $\beta$ is a tunable parameter whose consideration is primarily latency vs.\ security, a designer can fix it at a desired value and search for a corresponding $k$.
Since low $k$ are desirable, we perform this by starting at $k=1$, and successively evaluating larger values of $k$ until a feasible solution is found.
Depending on $b$ and the chosen $\varepsilon$, such a solution may not exist. If so, this will be apparent during system design. 
% \Jon{I found myself afloat in a sea of single-letter variable names, and was wishing for a quick-reference key to decode them. (Or even better, for some to be eliminated. :V)}

We finalize our analysis by formally composing C1 and C2: 
\begin{theorem}
    If C1 and C2 are satisfied under appropriately chosen system parameters, then the probability that two nodes $u$ and $v$ decide on $\mathtt{R}$ and $\mathtt{B}$ respectively is strictly $< \varepsilon$ over all timesteps $0 \leq t \leq \phi$. 
\end{theorem}
\begin{proof}
    The proof follows in a straightforward manner from the core guarantees that C1 and C2 provide. Without loss of generality, suppose a single node $u$ decides $\mathtt{R}$ at time $t < \phi$, and suppose that the network is at state $s_j$ at the time of decision. By construction, $\textsc{isAccepted}$ will return true with probability no greater than $\varepsilon$ for all network states $s_i < s_{c/2 + \Delta}$. Therefore, since $u$ decided, it must be the case that $s_j > s_{c/2 + \Delta}$, whp. Lastly, since the system will only revert back to a state with a majority of $\mathtt{B}$ nodes with negligible probability once it is past $s_{c/2 + \Delta}$, the probability that a node $v$ decides on $\mathtt{B}$ is strictly $< \varepsilon$.
\end{proof}

For\ \ illustration purposes, we examine a network with $n = 2000,\ \alpha = 0.8$, $\varepsilon \leq 2^{-32}$, and throughput of 10000tps.
If we allow this system to run for \textasciitilde{}4,000 years (which corresponds to $\phi = 10^{15}$), we choose $\beta$ such that the probability of a node committing
at state $s_{c/2+\Delta - 1}$ is $< \varepsilon$.
A safety failure then occurs when both a node commits $\mathtt{R}$ and the system reverts to $\mathtt{B}$. Figure~\ref{figure:failure_theta_0} shows the probability of failure and the corresponding $\beta$ for different choices of $k$.

\begin{figure}[h!]
\tronly{\includegraphics[width=\linewidth]{figures/consecutive-beta-large.pdf}}
        {\includegraphics[width=\linewidth]{figures/consecutive-beta.pdf}}
    \caption{$\beta$ and failure probability $p_f$ vs. $k$ values (x-axis), with $n = 2000$, $\alpha = 0.8$.}
\label{figure:failure_theta_0}
\end{figure}

\begin{comment}
In this section, we have shown that if the network parameters chosen by the system designer satisfy both conditions C1 and C2, then probability of a safety failure in Snowflake is negligible. 
Finally, there exists one more consequence to the Snowflake protocol.
Namely,
\end{comment}
Note that a larger choice of the $\alpha$ parameter dictates a larger space of feasible solutions, which implies a larger tolerance of Byzantine presence. 
If the system designer chooses to allow a larger presence of Byzantine nodes, she may achieve this goal at the cost of liveness. 
%XXX expand on this counterintuitive idea at some point, it implies that we can put up with very large byzantine coalitions
%XXX how big can $b$ be?
%XXX KEVIN: There are actually many cool things I can be saying additionally here, but what is written up to now is sufficient for tech report. I will soon-enough add up some interesting insights about speed of convergence once the system actually enters s_{c/2 + \Delta}.

\subsection{Safety Analysis of Snowball}\tronly{}{\vspace{-0.5em}}
In this section, we will demonstrate that Snowball provides strictly better
safety than Snowflake.
The key difference in protocol mechanics between Snowflake and Snowball is that correct nodes now monotonically increase their confidence in $\mathtt{R}$ and $\mathtt{B}$.
This limits the effect of random perturbations, as well as the impact of the Byzantine adversary.

To intuitively understand the behavioral differences between Snowball and its predecessor, we illustrate with an example, accompanied by some concrete numerical values.
\begin{comment}
The example will also provide intuition on how Snowball admits a safety guarantee strictly stronger than that of Snowflake. The goal of this section is to formally prove this guarantee.
\end{comment}
Suppose a network of size $n$ where $b = (1/5)n$, $\alpha = (4/5)k$ (i.e. the maximum allowable threshold to tolerate the $(1/5)n$ presence of Byzantine nodes), and where $k$ is chosen large enough to admit a feasible solution. 
Suppose that the network swings to state $s_{c/2 + \Delta}$, such that correct nodes have a non-negligible probability of deciding on $\mathtt{R}$, and suppose, for the sake of simplicity, that $\Delta$ is chosen as the highest possible value, specifically $c/2 + \Delta + b = \alpha n = (4/5)n$, the upper-threshold set up by the system designer.

In this configuration, at least $(4/5)n - b = (3/5)n$ (correct) nodes prefer $\mathtt{R}$. 
These nodes perceive network support for $\mathtt{R}$ of at most $(4/5)n$ if the Byzantine nodes choose to also vote for $\mathtt{R}$, and at least $(3/5)n$ if the Byzantine nodes choose to withhold support for $\mathtt{R}$. 
Conversely, the rest of the $(1/5)n$ (correct) nodes that may prefer $\mathtt{B}$ perceive network support for $\mathtt{B}$ of at least $(1/5)n$ and at most $(2/5)n$, depending on the Byzantine strategy. 
Regardless of Byzantine strategy, however, all correct nodes that prefer $\mathtt{R}$ will have an expected growth of the $\mathtt{R}$ confidence to be strictly greater than the $\mathtt{B}$ confidence for the nodes that prefer $\mathtt{B}$. 
This is because network support for $\mathtt{R}$ is at least $(3/5)n$, which is greater than the maximum network support for $\mathtt{B}$ of $(2/5)n$. 
Therefore, once the network swings to the minimum required state, the expected growth of the $\mathtt{R}$-confidence will be greater than that of $\mathtt{B}$-confidence, for all correct nodes and for all Byzantine strategies.

In the previous section, we provided the two conditions necessary for guaranteeing the safety of Snowflake. The second condition, C2, remains identical in Snowball, and therefore does not need to be modified and re-analyzed. Conversely, condition C1 (i.e. showing that the network is irreversible whp) needs to be re-analyzed. In this section, using the same intuition provided in the example above, we formally show how Snowball achieves irreversibility under strictly higher probability than in Snowflake.

\tronly{
\paragraph{Security of Snowball.}
Figure~\ref{fig:snowball_simulator} illustrates the scheduler-based version of Snowball.
In Snowball, a node $u$ \textit{prefers} $\mathtt{R}$ if $u.d[\mathtt{R}] > u.d[\mathtt{B}]$, and vice versa. At the start of the protocol, these preferences are initialized to $\langle 0, 0 \rangle$. 
This is in contrast to Snowflake, where nodes prefer a color based only on the latest successful sample.
We can model the protocol with a Markov chain similar to that of Snowflake, and derive the parameters and feasibility for the protocol.

\begin{figure}[h!]
\begin{algorithmic}[1]
\State initialize $\forall u, u.\codecolor$
\State initialize $\forall u, u.\textrm{lastcol}$
\State initialize $\forall u, u.\codecount \coloneqq 0$
\State initialize $\forall u, \forall \codecolor' \in \{\texttt{R}, \texttt{B}\}, u.d[\codecolor'] \coloneqq 0$
\For{$t = 1$ to $\phi$}
    \State $u \coloneqq \Call{sample}{\mathcal{C}, 1}$
    \State $\mathcal{A}(u, \mathcal{C})$
    \State $\mathcal{K} \coloneqq \Call{sample}{\mathcal{N}\textbackslash u, k}$
    \State $P \coloneqq \texttt{[}v.\codecolor\quad\textbf{for}\ v \in \mathcal{K}\texttt{]}$
    \For{$c' \in \{\texttt{R}, \texttt{B}\}$}
    \If{$P.\Call{count}{\codecolor'} \ge \alpha\cdot k$}
        \State $u.d[\codecolor']\texttt{++}$
        \If {$u.d[\codecolor'] > u.d[\codecolor]$}
            \State $u.\codecolor \coloneqq \codecolor'$
        \EndIf
        \If{$\codecolor' \neq u.\textrm{lastcol}$}
            \State $u.\textrm{lastcol} \coloneqq \codecolor'$, $u.\codecount \coloneqq 0$
        \Else
            \State $\texttt{++}u.\codecount$
        \EndIf
    \EndIf
    \EndFor
\EndFor
\captionof{figure}{Snowball run by a global scheduler.}
\label{fig:snowball_simulator}
\end{algorithmic}
\end{figure}

We fix a $\Delta$ such that, under the Markovian construction of Snowflake, if the system reaches $s_{c/2 + \Delta}$, then whp it does not revert. We show that this choice of $\Delta$ provides an even stronger guarantee once the network switches protocol to Snowball.
Without loss of generality, we cluster the correct nodes into two groups, and represent each group through two nodes $u$ and $v$, where $u$ represents the set of correct nodes that prefer red and $v$ those that prefer blue. 
Lastly, let the state of the system be initialized at $s_{c/2 + \Delta}$.

% With non-negligible probability, $u$ will decide on $\mathtt{R}$, since the network is biased enough to trigger the decision predicate. 
The best possible confidence-configuration that the Byzantine nodes can attempt to force correct nodes into is where all $\mathtt{R}$-preferring nodes, represented by $u$, are forced to maintain nearly equal confidence between the two colors, and where all of the $\mathtt{B}$-preferring nodes, represented by $v$, gain as much confidence as possible for~$\mathtt{B}$.

Let $u.d[\mathtt{R}] = u.d[\mathtt{B}] + 1$, corresponding to the minimal viable color difference for the red-preferring nodes, and let $v.d[\mathtt{B}] - v.d[\mathtt{R}] = \kappa$ be the difference for the blue-preferring nodes. While $\kappa \geq 0$, then the expectation of preference growths at time $t$ are:
\begin{equation}
    {\scriptsize
\begin{split}
    \mathbb{E}[{u.d}^t[\mathtt{R}]] &= \mathbb{E}[{u.d}^{t-1}[\mathtt{R}]] + \left(\frac{1}{2}+\frac{\Delta}{c}\right)P(\mathcal{H}_{\mathcal{N}, c/2 + \Delta + b}^k  \geq \alpha k)\\
    \mathbb{E}[{u.d}^t[\mathtt{B}]] &= \mathbb{E}[{u.d}^{t-1}[\mathtt{B}]] + \left(\frac{1}{2}+\frac{\Delta}{c}\right)P(\mathcal{H}_{\mathcal{N}, c/2 - \Delta + b}^k  \geq \alpha k)\\
    \mathbb{E}[{v.d}^t[\mathtt{R}]] &= \mathbb{E}[{v.d}^{t-1}[\mathtt{R}]] + \left(\frac{1}{2}-\frac{\Delta}{c}\right)P(\mathcal{H}_{\mathcal{N}, c/2 + \Delta}^k  \geq \alpha k)\\
    \mathbb{E}[{v.d}^t[\mathtt{B}]] &= \mathbb{E}[{v.d}^{t-1}[\mathtt{B}]] + \left(\frac{1}{2}-\frac{\Delta}{c}\right)P(\mathcal{H}_{\mathcal{N}, c/2 - \Delta + b}^k  \geq \alpha k)\\
\end{split}
    }
    \label{equation:expectation_snowball}
\end{equation}
Note that we are implicitly using an indicator function when adding up the probability of success to the expected confidence growth, where the expected value of the indicator value is exactly the probability of success of a sample. 
Substituting for Hoeffding's tail bound, we get the rate at which $\kappa$ decreases per time-step to be
\begin{equation}
    % p = \cfrac{c/2 + \Delta + b}{c+b}
    \begin{split}
    \small
        r = \biggl(\left(\frac{1}{2}-\frac{\Delta}{c}\right) &e^{-k\alpha \log \left(\alpha/\frac{c/2 - \Delta + b}{c+b}\right)}\\
    &e^{-k(1-\alpha) \log \left(\left(1-\alpha\right)/\left(1-\frac{c/2 - \Delta + b}{c+b}\right)\right)}\biggr)\\
        - \biggl(\left(\frac{1}{2}-\frac{\Delta}{c}\right)&e^{-k\alpha \log \left(\alpha/\frac{c/2 + \Delta + b}{c+b}\right)}\\
    &e^{-k(1-\alpha) \log \left(\left(1-\alpha\right)/\left(1-\frac{c/2 + \Delta + b}{c+b}\right)\right)}\biggr)\\
    \end{split}
    \label{equation:rate_decrease}
\end{equation}

Upon reaching $\kappa = -1$, the blue-preferring nodes will have flipped to red.
From then on, all correct nodes are red, and the confidence of red grows significantly faster than that of blue.
We now show that the rate of growth of $\mathtt{R}$ will always be larger than the rate of growth of $\mathtt{B}$, over all correct nodes.

Letting $X({u.d}^t[\mathtt{R}])$ be a random variable that outputs the confidence of red for $u$ at timestep $t$, and letting $\bar X({u.d}^t[\mathtt{R}]) = X({u.d}^1[\mathtt{R}]) + \dots + X({u.d}^t[\mathtt{R}])$, we apply Hoeffding's concentration inequality to get that $P({\bar X({u.d}^t[\mathtt{R}])} - \mathbb{E}[\bar X({u.d}^t[\mathtt{R}])] \geq l) \leq e^{-2l^2t}$.
At time $t$, the expected confidence of red for $u$ is
\begin{equation}
    {u.d}^0[\mathtt{R}] + t\left(\frac{1}{2} + \frac{\Delta}{c}\right)\left(\frac{\frac{c}{2} + \Delta + b}{c+b}\right)
\end{equation}
And conversely the expected confidence of blue for $u$ is
\begin{equation}
    {u.d}^0[\mathtt{B}] + t\left(\frac{1}{2} + \frac{\Delta}{c}\right)\left(\frac{\frac{c}{2} - \Delta + b}{c+b}\right)
\end{equation}
Then, the probability that the expected value of red confidence is close to the expected value of the blue confidence is
\begin{equation}
    e^{-2t^3(1/2 + \Delta/c)^2(2\Delta/(c+b))^2}
    \label{equation:concentration_mean}
\end{equation}
Solving for $t$ that leads to such probability being negligible leads to:
\begin{equation}
    t = \left(\cfrac{\log(\varepsilon)}{-2\cdot(1/2 + \Delta/c)^2(2\Delta/(c+b))^2}\right)^{1/3}
    \label{equation:minimal_time}
\end{equation}
Therefore, at only a small number of timesteps $t$, $u$ has a red-confidence centered around its mean, with probability $\varepsilon$ close to the mean of the blue-confidence. An identical analysis follows for $v$.
In other words, after a small number of time-steps, the red-confidence of $u$ grows large enough such that the probability of this value being close to that of the blue-confidence is negligible. Further timesteps amplify this distance and further decrease the probability of the two confidences being close. 
}

\tronly{}{The details are provided in an accompanying technical report~\cite{avalanche}. The central result is as follows:}
We conclude our result with the following Theorem:
\begin{theorem}
Over all viable network parameters $n,\ b$, and for all appropriately chosen system parameters $k$ and $\alpha$, the probability of violating safety in Snowball is strictly less than the probability of violating safety in Snowflake.
\end{theorem}
\begin{proof}
%    The proof is straightforward.
Under the construction of Snowball in comparison to Snowflake, we see that safety criterion C2 remains the same. In other words, the consecutive-successes decision predicate has the same guarantees. 
On the other hand, the probabilistic guarantees of C1 change, meaning that the reversibility probabilities of the system are different. 
However, as we determined in Equation~\ref{equation:minimal_time}, over a few timesteps
$u$'s confidence difference between $\mathtt{R}$ and $\mathtt{B}$ grows large enough to guarantee that this confidence difference will revert back with only negligible probability.
As time progresses, the probability of being close decreases poly-exponentially, as shown in Equation \ref{equation:concentration_mean}. The same results follow for $v$, but inversely, meaning that the means get closer to each other rather than deviate. This continues until $v$ flips color. 
\end{proof}

\tronly{Snowball has strictly stronger security guarantees than Snowflake, which implies that appropriately chosen parameters chosen for Snowflake automatically apply for Snowball. Using the same techniques as before, the system designer chooses appropriate $k$ and $\beta$ values that ensure the desired $\varepsilon$ safety guarantees.}{}

% ------------------------------------------------------
% OLD STUFF ABOUT THETA 1 AND BYZANTINE OPTIMALITY ADDED
% TO FILE oldtheta1.tex, see line below. DO NOT DELETE!!
% \input{oldtheta1.tex}
% ------------------------------------------------------

\subsection{Safety Analysis of Avalanche}\tronly{}{\vspace{-0.5em}}
The key difference between Avalanche and Snowball is that in Avalanche, queries on the DAG on transaction $T_i$ are used to implicitly query the entire ancestry of $T_i$.
In particular, a transaction $T_i$ is preferred by $u$ \textit{if and only if} all ancestors are also preferred.
Suppose that $T_i$ and $T_j$ are in the same conflict set.
We can now infer two things.
First, we can consider the entire ancestry of $T_i$ and $T_j$ as a single decision instance of Snowball, where the ancestry of $T_i$ can be considered to be the $\mathtt{R}$ decision, and the ancestry of $T_j$ can be considered to be the $\mathtt{B}$ decision.
Second, since $T_i$ must be preferred if a child of $T_i$ is to be preferred, then we can collapse the progeny of $T_i$ into a single urn which repeatedly adds confidence to $\mathtt{R}$ whenever a child of $T_i$ gets a chit.
Consequently, Avalanche maps to an instance of Snowball, with the previously shown safety properties.

We note, however, since a decision on a virtuous transaction is dependent on its parents, Avalanche's liveness guarantees do not mirror that of Snowball. We address this in the next two sub-sections.

\subsection{Safe Early Commitment}\tronly{}{\vspace{-0.5em}}
\tronly{
As we reasoned previously, each conflict set in Avalanche can be viewed as an instance of Snowball, where each progeny instance iteratively votes for the entire path of the ancestry.
This feature provides various benefits; however, it also can lead to some virtuous transactions that depend on a rogue transaction to suffer the fate of the latter.
%has the drawback that it can entangle the fate of some unfortunate virtuous transactions with the fate of rogue ones.
In particular, rogue transactions can interject in-between virtuous transactions and reduce the ability of the virtuous transactions to ever reach the required $\textsc{isAccepted}$ predicate.
}{}
As a thought experiment, suppose that a transaction $T_i$ names a set of parent transactions that are all decided, as per local view.
If $T_i$ is sampled over a large enough set of successful queries without discovering any conflicts, then, since by assumption the entire ancestry of $T_i$ is decided, it must imply that at least $c/2 + \Delta$ correct nodes also vote for $T_i$, achieving irreversibility.

To then statistically measure the assuredness that $T_i$ has been accepted by a large percentage of correct nodes without any conflicts, we make use of a one-way birth process, where a birth occurs when a new correct node discovers the conflict of $T_i$. \tronly{Necessarily, deaths cannot exist in this model, because a conflicting transaction cannot be unseen once a correct node discovers it. Let $t = 0$ be the time when $T_j$, which conflicts with $T_i$, is introduced to a single correct node $u$.
Let $s_x$, for $x = 1$ to $c$, be the state where the number of correct nodes that know about $T_j$ is $x$, and let $p(s_x)$ be the probability of birth at state $s_x$. Then, we have:
\begin{equation}
    p(s_x) = \frac{c - x}{c} \left(1 - \frac{{n - x \choose k}}{{n \choose k}}\right)
\end{equation}
Solving for the expected time to reach the final birth state provides a lower bound to the $\beta_1$ parameter in the $\textsc{isAccepted}$ fast-decision branch.}{We leave details to the tech report.} The table below shows an example of the analysis for $n = 2000$, $\alpha = 0.8$, and various $k$, where $\varepsilon \ll 10^{-9}$, and where $\beta$ is the minimum required size of $d(T_i)$.
\tronly{}{\vspace{-2ex}}
\begin{table}[h!]
    \small
	\centering
	\begin{tabular}{llllll}
		$k$   & 10 & 20 & 30 & 40 \\ \hline
		$\beta$ & 10.87625 & 10.50125 & 10.37625 & 10.25125
	\end{tabular}
	\label{table:fast-path-beta}
\end{table}
\tronly{}{\vspace{-2ex}}
% \begin{figure}
%     \includegraphics[width=\linewidth]{figures/fast-path-beta.pdf}
%     \captionof{figure}{An example of $\beta$ solutions for different $k$, with $n = 2000, \alpha = 0.8$.}
%     \label{fig:fast-path-beta}
% \end{figure}
\noindent Overall, a very small number of iterations are sufficient for the safe early commitment predicate.

\subsection{Churn and View Updates}\tronly{}{\vspace{-0.5em}}
\tronly{Any realistic system needs to accommodate the departure and arrival of nodes.}{}
Up to now, we simplified our analysis by assuming a precise knowledge of network membership.
We now demonstrate that Avalanche nodes \tronly{can admit a well-characterized amount of churn,
by showing how to pick parameters such that Avalanche nodes}{} can differ in their view of the network and still safely make decisions.

Consider a network whose operation is divided into epochs of length $\tau$, and a view update from epoch $t$ to $t+1$ during which $\gamma$ nodes join the network and $\bar \gamma$ nodes depart.
Under our static construction, the state space $s$ of the network had a key parameter $\Delta^{t}$ at time $t$, induced by $c^{t}, b^{t}, n^{t}$ and the chosen security parameters.
This can, at worst, impact the network by adding $\gamma$ nodes of color $\mathtt{B}$, and remove $\bar \gamma$ nodes of color $\mathtt{R}$.
At time $t+1$, $n^{t+1} = n^{t} + \gamma - {\bar \gamma}$, while $b^{t+1}$ and $c^{t+1}$ will be modified by an amount $\leq \gamma - {\bar \gamma}$, and thus induce a new $\Delta^{t+1}$ for the chosen security parameters.
This new $\Delta^{t+1}$ has to be chosen such that $P(s_{c^{t+1}/2 + \Delta^{t+1} - \gamma} \rightarrow s_{ps}) \leq \varepsilon$, to ensure that the system will converge under the previous pessimal assumptions. The system designer can easily do this by picking an upper bound on $\gamma, \bar \gamma$.

The final step in assuring the correctness of a view change is to account for a mix of nodes that straddle the $\tau$ boundary. We would like the network to avoid an unsafe state no matter which nodes are using the old and the new views.
The easiest way to do this is to determine $\Delta^t$ and $\Delta^{t+1}$ for desired bounds on $\gamma, \bar \gamma$, and then to use the conservative value $\Delta^{t+1}$ during epoch $t$. In essence, this ensures that no commitments are made in state space $s^t$ unless they conservatively fulfill the safety criteria in state space $s^{t+1}$. As a result, there is no possibility of a node deciding red at time $t$, the network going through an epoch change and finding itself to the left of the new point of no return $\Delta^{t+1}$.

This approach trades off some of the feasibility space, to add the ability to accommodate $\gamma, \bar \gamma$ node churn per epoch.
\tronly{
Overall, if $\tau$ is in excess of the time required for a decision (on the order of minutes to hours), and nodes are loosely synchronized,
they can add or drop up to $\gamma, \bar \gamma$ nodes in each epoch using the conservative process described above. }{}
% Due to space reasons,
We leave the precise method of determining the next view to a subsequent paper, and instead
rely on a membership oracle that acts as a sequencer and $\gamma$-rate-limiter, using technologies like Fireflies~\cite{JohansenRVJ15}.
% and Ethereum~\cite{wood2014ethereum}.
\end{appendices}
\end{document}

### 2020-11-03 sections/evaluation.tex
add initial source and 'fr' make target

\newcommand{\sysname}{Avalanche}

%We have fully implemented the proposed payment system.  In this section, we examine its throughput, scalability, and latency
%through a large scale deployment on Amazon AWS\@, and provide a comparison to known results
%from other systems.

\subsection{Setup}%\tronly{}{\vspace{-0.25em}}
% hardware setup
We conduct our experiments on Amazon EC2 by running from hundreds (125) to thousands (2000)
of virtual machine instances.  We use \texttt{c5.large} instances, each of
which simulates an individual node. AWS provides
bandwidth of up to 2 Gbps, though the {\sysname} protocol utilizes at most around 100 Mbps.

% the transaction specification
Our implementation supports two versions of transactions: one is the customized UTXO format,
while the other uses the code directly from Bitcoin 0.16. Both supported formats use secp256k1
crypto library from bitcoin and provide the same address format for wallets. All experiments
use the customized format except for the geo-replication, where results for both are given.

We simulate a constant flow of new transactions from users by creating
separate client processes, each of which maintains
separated wallets, generates transactions with new recipient addresses and
sends the requests to {\sysname} nodes.
%\textbf{[RVR: what is a ``full'' node? (fixed)]}
We use several such client
processes to max out the capacity of our system.  The number of recipients
for each transaction is tuned to achieve average transaction sizes of
around 250 bytes (1--2 inputs/outputs per transaction on average and a stable
UTXO size), the current average transaction size of Bitcoin. To utilize the
network efficiently, we batch up to 40 transactions during a query, but
maintain confidence values at individual transaction granularity.

% how do we measure throughput/latency
All reported metrics reflect end-to-end measurements taken from the perspective of all clients.
That is, clients examine the total number of confirmed
transactions per second for throughput, and, for each transaction,
subtract the initiation timestamp
from the confirmation timestamp for latency. Each throughput experiment is
repeated for 5 times and standard deviation is indicated in each figure.
%\tronly{Because we
%saturate the capacity of the system in all runs, some transactions
%will have much higher latency than most, we use the 1.5$\times$IQR rule
%commonly used to filter out outliers. Take the geo-replicated experiment as an example,
%$15.8\%$ data are the outliers having 6.63 second latency on average. They fall
%out of 1.5$\times$IQR (approximately $3\sigma$) range and are filtered out. There are very few outliers
%when the system is not saturated. All reported latencies (including maximum) are those not filtered.
%%\textbf{[RVR: how high are these outliers?  how many are there?  do they still happen if the system is not saturated? (fixed)]}
%}{}
As for security parameters, we pick $k = 10$, $\alpha = 0.8$, $\beta_1 = 11$, $\beta_2 = 150$, which yields an MTTF of \textasciitilde{}$10^{24}$ years.

\subsection{Throughput} %\tronly{}{\vspace{-0.5em}}

We first measure the throughput of the system by saturating it with
transactions and examining the rate at which transactions are confirmed in the
steady state.  For this experiment, we first run {\sysname} on 125 nodes
with 10 client processes, each of which maintains 400 outstanding transactions at
any given time.

As shown by the first group of bars in Figure~\ref{fig:eval-thr}, the system achieves
6851 transactions per second (tps) for a batch size of 20 and above 7002 tps for a batch size of 40.
Our system is saturated by a small batch size comparing to other blockchains with known performance:
Bitcoin batches several thousands of transactions per block, Algorand~\cite{GiladHMVZ17} uses 2--10 Mbyte blocks, i.e., 8.4--41.9K tx/batch and Conflux~\cite{confluxLLXLC18} uses 4 Mbyte blocks, i.e., 16.8K tx/batch. These systems are relatively slow in making a single decision, and thus require a very large batch (block) size for better performance. Achieving high throughput with small batch size implies low latency, as we will show later.

\begin{figure}[h]
\includegraphics[width=\linewidth]{figures/thr-ava.pdf}
\captionof{figure}{Throughput vs. network size. Each pair of bars is produced with batch size of 20 and 40, from left to right.}
%    The slight degradation with doubled network sizes illustrates scalability.}
\label{fig:eval-thr}
\end{figure}

\begin{figure}
\includegraphics[width=\linewidth]{figures/thr-raw.pdf}
\captionof{figure}{Throughput for batch size of 40, with (left) and without (right) signature verification.}
\label{fig:eval-thr-raw}
\end{figure}

\subsection{Scalability}%\tronly{}{\vspace{-0.25em}}

To examine how the system scales in terms of the number of nodes
participating in {\sysname} consensus, we run experiments with identical settings
and vary the number of nodes from 125 up to 2000.

Figure~\ref{fig:eval-thr} shows that overall throughput degrades about $1.34\%$ to 6909 tps when the network grows by a factor of 16 to $n = 2000$.
This degradation is minor compared to the growth of the network size.
Note that the x-axis is logarithmic.
% , and thus throughput degradation is negligible.

Avalanche acquires its scalability from three sources: first,
maintaining a partial order that captures only the spending relations
allows for more concurrency than a classical BFT replicated
log that linearizes all transactions; second, the lack of a leader naturally avoids bottlenecks;
finally, the number of messages each node has to handle per decision is $O(k)$ and does not grow as the network scales up.
% Another way to view this is that {\sysname} implicitly performs similar to sharding with
% the help of the DAG, in that multiple nodes can work on different parts of the
% central data structure concurrently.
%{\sysname} trades off total order for a scalable system.

%(Figure 4: x-axis is the number of client processes, y-axis is the averaged latency)

%\tronly{{}
\subsection{Cryptography Bottleneck}%\tronly{}{\vspace{-0.25em}}

We next examine where bottlenecks lie in our current implementation.
The purple bar on the right of each group in Figure~\ref{fig:eval-thr-raw} shows the throughput of Avalanche with signature verification
disabled. Throughputs get approximately 2.6x higher, compared to the blue bar on the left.
This reveals that cryptographic verification overhead is the current bottleneck of our system implementation.
This bottleneck can be addressed by offloading
transaction verification to a GPU\@. Even without such optimization,
7K tps is far in excess of extant blockchains.
%}{}

\subsection{Latency}%\tronly{}{\vspace{-0.25em}}

\begin{figure}
\includegraphics[width=\linewidth]{figures/lat.pdf}
\captionof{figure}{Transaction latency distribution for $n = 2000$. The x-axis
    is the transaction latency in log-scaled seconds, while the
    y-axis is the portion of transactions that fall into the confirmation
    time (normalized to $1$).  Histogram of all transaction latencies for a client is shown on the left with $100$ bins,
    while its CDF is on the right.}
\label{fig:eval-lat1}
\end{figure}

The latency of a transaction is the time spent from the moment of its
submission until it is confirmed as accepted.
%\textbf{[RVR: this is not quite clear to me.  Where is finality defined? (fixed)  I only see the isAccepted() predicate that checks finality, but I assume finality, whatever it is, is achieved before a client learns about it.]}
Figure~\ref{fig:eval-lat1} tallies
the latency distribution histogram using the same setup as for the throughput
measurements with 2000 nodes. The x-axis is the time in seconds while the y-axis
is the portion of transactions that are finalized within the corresponding
time period. This figure also outlines the Cumulative Distribution Function (CDF)
by accumulating the number of finalized transactions over time.

This experiment shows that most transactions are confirmed within approximately 0.3 seconds.
% The sharp transition that goes from nearly $0\%$ to $100\%$ shows that
The most common latencies are around 206 ms and variance is low,
indicating that nodes converge on the final value as a group around the same time.
The second vertical line shows the maximum latency we observe, which is around 0.4 seconds.
% (XXX: more on the distribution?)
%\textbf{[RVR: why use a log scale for the x-axis?  Is the transition really sharp? (yes) It doesn't look to me that way.  Is the 1.1 second maximum latency before or after removing outliers? (fixed) Is ``confirmation time'' the same as latency, and if so, why use different terms? (fixed)  The histogram is tiny... (yeah, because of cdf..)]}

Figure~\ref{fig:eval-lat2} shows
transaction latencies for different numbers of nodes.
The horizontal edges of boxes represent minimum,
first quartile, median, third quartile and maximum latency respectively, from
bottom to top.  Crucially, the experimental data show that median latency is more-or-less
independent of network size.
%\textbf{[RVR: the x-axis *should* be a log scale here.  It looks like a log scale but it is linear. (fixed)]}

\begin{figure}[h]
\includegraphics[width=\linewidth]{figures/lat2.pdf}
\captionof{figure}{Transaction latency vs. network size. ``b'' indicates batch size and ``raw'' is the run without signature verification.}
\label{fig:eval-lat2}
\end{figure}

\subsection{Misbehaving Clients}%\tronly{}{\vspace{-0.25em}}
\label{sec:evaluation-misbehaving}
\begin{figure}
\includegraphics[width=\linewidth]{figures/lat3.pdf}
\captionof{figure}{Latency vs. ratio of rogue transactions.}
\label{fig:eval-lat3}
\end{figure}

We next examine how rogue transactions issued by misbehaving clients that
double spend unspent outputs can affect latency for virtuous
transactions created by honest clients. We adopt a strategy to
simulate misbehaving clients where a fraction (from $0\%$ to $25\%$) of the
pending transactions conflict with some existing ones.
%\textbf{[RVR: this is confusing, but maybe I should reread Section 2.  It used to be the case that it was impossible to create transactions that conflict with virtuous ones. (they double spend their own unspent outputs)]}
The client processes
achieve this by designating some double spending transaction flows among all
simulated pending transactions and sending the conflicting transactions to
different nodes. We use the same setup with $n = 1000$ as in the previous
experiments, and only measure throughput and latency of confirmed transactions.

{\sysname}'s latency is only slightly affected by misbehaving clients, as shown
in Figure~\ref{fig:eval-lat3}. Surprisingly, maximum latencies drop slightly when
the percentage of rogue transactions increases.  This behavior occurs
because, with the introduction of rogue transactions, the overall
\emph{effective} throughput is reduced and thus alleviates system load.
%\tronly{
This is confirmed by
Figure~\ref{fig:eval-thr2}, which shows that throughput (of virtuous transactions) decreases with the ratio of rogue transactions.
Further, the reduction in throughput appears proportional to the number of misbehaving clients,
that is, there is no leverage provided to the attackers.
%}{
%This is confirmed by throughput decreases with the ratio of rogue transactions.
%Further, the reduction in throughput appears proportional to the number of misbehaving clients,
%that is, there is no leverage provided to the attackers.
%}
%(XXX: is this universally true, or just for this attacker? Is this the strongest attacker we can imagine?)

\begin{figure}
\includegraphics[width=\linewidth]{figures/thr-byz.pdf}
\captionof{figure}{Throughput vs. ratio of rogue transactions.}
\label{fig:eval-thr2}
\end{figure}

\subsection{Geo-replication}%\tronly{}{\vspace{-0.25em}}

\begin{figure}
\includegraphics[width=\linewidth]{figures/lat4.pdf}
\captionof{figure}{Latency histogram/CDF for $n = 2000$ in 20 cities.}
\label{fig:eval-lat4}
\end{figure}

Next experiment shows the system in an emulated geo-replicated scenario, patterned
after the same scenario in prior work~\cite{GiladHMVZ17}.
We selected 20 major cities that appear to be near substantial numbers of reachable
Bitcoin nodes, according to~\cite{bitnodes2018}. The cities cover North America,
Europe, West Asia, East Asia, Oceania, and also cover the top 10 countries with the
highest number of reachable nodes. We use the latency and jitter matrix crawled
from~\cite{wondernetworkping2018} and emulate network packet latency in the Linux kernel using
\texttt{tc} and \texttt{netem}. 2000 nodes are distributed evenly to each
city, with no additional network latency emulated between nodes within the same city.
Like Algorand's evaluation, we also cap our bandwidth per process to 20 Mbps
to simulate internet-scale settings where there are many commodity network links.
We assign a client process to each city, maintaining 400 outstanding transactions per city
at any moment.

In this scenario, Avalanche achieves an average throughput of 3401 tps, with a standard deviation of 39 tps.
As shown in Figure~\ref{fig:eval-lat4}, the median transaction latency
is 1.35 seconds, with a maximum latency of 4.25 seconds. We also support native Bitcoin code
for transactions; in this case, the throughput is 3530 tps, with $\sigma = 92$ tps.

\subsection{Comparison to Other Systems}%\tronly{}{\vspace{-0.25em}}
Though there are seemingly abundant blockchain or cryptocurrency protocols,
most of them only present a sketch of their protocols and do not offer practical implementation or evaluation results.
Moreover, among those who do provide results, most are not evaluated in realistic, large-scale (hundreds to thousands of full nodes participating in consensus) settings.

Therefore, we choose Algorand and Conflux for our comparison. Algorand, Conflux, and Avalanche are all fundamentally different in their
% consensus principle and
design. Algorand's committee-scale consensus algorithm is quorum-based Byzantine agreement, and Conflux extends Nakamoto consensus by a DAG structure to facilitate higher throughput, while Avalanche belongs to a new protocol family based on metastability. Additionally, we use Bitcoin~\cite{nakamoto2008bitcoin} as a baseline.

Both Algorand and Avalanche evaluations use a decision network of size 2000 on EC2.
Our evaluation picked shared \texttt{c5.large} instances, while Algorand used \texttt{m4.2xlarge}\@.
These two platforms are very similar except for a slight CPU clock speed edge for \texttt{c5.large}, which goes largely unused because our process only consumes $30\%$ in these experiments.
The security parameters chosen in our experiments guarantee a safety violation
probability below $10^{-9}$ in the presence of $20\%$ Byzantine nodes, while
Algorand's evaluation guarantees a violation probability below $5 \times 10^{-9}$ with $20\%$ Byzantine nodes.

Neither Algorand nor Conflux evaluations take into account the overhead of cryptographic verification.
Their evaluations use blocks that carry megabytes of dummy data and present the throughput in MB/hour or GB/hour unit. So we use the average size of a Bitcoin transaction, 250 bytes, to derive their throughputs. In contrast, our experiments carry real transactions and fully take all cryptographic overhead into account.

The throughput is 3-7 tps for Bitcoin,
874 tps for Algorand (with 10 Mbyte blocks),
3355 tps for Conflux (in the paper it claims 3.84x Algorand's throughput under the same settings).

In contrast, {\sysname} achieves over 3400 tps
consistently on up to 2000 nodes without committee or proof-of-work.
As for latency, a transaction is confirmed after 10--60
minutes in Bitcoin, around 50 seconds in Algorand,
7.6--13.8 minutes in Conflux, and 1.35 seconds in {\sysname}.

Avalanche performs much better than Algorand in both throughput and latency
because Algorand uses a verifiable random function to elect committees, and
maintains a totally-ordered log while {\sysname} establishes only a partial
order. Although Algorand's leadership is anonymous and changes continuously,
it is still leader-based which could be the bottleneck for scalability, while
{\sysname} is leader-less.

Avalanche has similar throughput to Conflux, but its latency is 337--613x better.
Conflux also uses a DAG structure to amortize the cost for consensus and increase the throughput,
however, it is still rooted in Nakamoto consensus (PoW), making it unable to
have instant confirmation compared to Avalanche.

In a blockchain system, one can usually improve throughput at the cost of latency through batching. The real bottleneck of the performance is the number of decisions the system can make per second, and this is fundamentally limited by either Byzantine Agreement ($\mathrm{BA}^{*}$) in Algorand and Nakamoto consensus in Conflux.

### 2020-11-03 sections/other-crypto.tex
add initial source and 'fr' make target

\subsection{Snowball vs. Blockchain}
Informally speaking, Nakamoto consensus achieved by the Blockchain
algorithm, is a process where the whole system synchronizes their replicated
log by creating a large time delay upon each generation of a block. The
Blockchain system is leaderless because there is no notion of elected leader
that could control the content of multiple slots on the log. However, the peer who is lucky enough
to come up with the block containing a proper nonce will take over the
ownership of the slot, and thus becomes a conceptual temporary ``leader''
only for the slot.  This analogy to traditional BFT reveals the
fundamental inefficiency in Nakamoto consensus, where each commonly accepted
block can be viewed as the outcome of a one-time leader election, the creator
of the block loses leadership immediately after establishment of the block.

There have been work on improving the performance by explicitly stating and
extending the leadership to several slots, such as Bitcoin-NG and Thunderella.
But this makes the protocol leader-based, and may suffer from the issues of
having a single leader in an Internet-scale network environment.

On the other hand, Nakamoto consensus was originally used for Bitcoin, a
peer-to-peer electronic cash system. So each block usually contains hundreds or
thousands of transactions that are mutually compatible (no double spending of
money), and there exists another separated logic, UTXO (Unspent Transaction
Output) graph that is made of all transactions to represent the actual state of
the cash system. On a UTXO graph, a transaction has multiple inputs and
outputs. A valid transaction will provide with signature proofs for all the
inputs that correspond and consume some outputs from some transactions prior in
history, and then offer some dangling unspent outputs which could be taken as
inputs in the future by subsequent transactions. The whole UTXO graph is valid
only if there does not exist two or more transactions that take the same input,
which is usually called a ``double spend''. With that being said,
Nakamoto consensus provides with more power than that is required for
maintaining a double-spend-free UTXO graph, because Blockchain ensures all transactions are
\emph{linearly} ordered, while a valid UTXO graph used for ledger does not
require such linear ordering. For example, suppose there are two transactions
A, B that both have only one input taken from one of the two outputs from
transaction C. So using Blockchain, eventually only a specific linear ordering
of A, B and C will be accepted, which might be unnecessary because both C, A, B
and C, B, A are valid orderings as A and B are independent. The only
necessary dependencies are all exactly captured by UTXO graph itself.

This leads to two possible paths of research: to fully utilize the ability
of keeping a linear log as in state machine replication so Nakamoto consensus is
not wasted, or to construct another weaker consensus that only resolves double
spends, not capable of keeping a linear log but just fits the need for an
electronic cash system. One well-known system for the first approach is
Ethereum, a decentralized application platform that makes use of the linear log
to certify executions and interactions of smart contract logic. The second approach motivates us
to Snowball.

### 2020-11-03 sections/protocol.tex
add initial source and 'fr' make target

Snowball is a leaderless, multi-decree, probabilistic consensus protocol that makes
use of sub-sampled voting. It entangles voting results to render the process
more efficient than independently deciding on each item alone. We first provide
the intuition behind the protocol informally, then present the formal protocol
specification, followed by a proof of safety and discussion on liveness for virtuous
transactions.

\subsection{Intuition}

The protocol operates by building a directed acyclic graph (DAG) of all known transactions.
This DAG is rooted at a protocol-defined \emph{genesis} transaction shared by all nodes.
Each transaction is represented as a vertex in the DAG.
When a transaction is created, it names one or more \emph{parent transactions}, which point to already accepted vertices in the DAG.
The parent-child relationships encoded in the DAG do not correspond to application-specific structure; for instance, a child vertex need not spend or have any relationship with the funds received in the parent transaction.
These relationships exist solely to entangle and enforce the fate of previous decisions made by the system.
For each vertex in the resulting DAG, there exists a static \emph{ancestor set} that contains all transactions reachable via parent edges back in history, and a \emph{progeny} that recursively includes all children transactions and their offspring.

The central challenge in Snowball is to choose among a set of conflicting transactions.
This notion of conflict in Snowball is application-defined.
In a cryptocurrency application, transactions that spend the same funds (\emph{double-spends}) form a \emph{conflict set}, out of which only a single one can be accepted.
Note that the graph structure of transactions that depend on each other, also known as the UTXO graph, is completely independent of the DAG that Snowball maintains. As a result, any two vertices may be in conflict.
Figure~\ref{fig:dag-conflicting-set} shows an example.

\begin{center}
    \input{figures/dag-conflicting-set}
    \captionof{figure}{The DAG structure in Snowball.  Shaded groups of transactions indicate \emph{conflict sets}---transactions that are conflicting.}\label{fig:dag-conflicting-set}
\end{center}

Snowball makes its choice of an accepted transaction from a conflict set using a novel voting protocol.
%Snowball makes its choice of an accepted transaction from a conflict set using a meta-stable, intertwined voting protocol.
The insight behind this protocol is for the correct nodes in the system to topple over, or ``roll snowball'', towards a system-wide decision.
Of course, the FLP impossibility result~\cite{fischer1985impossibility}
implies that there will be a path where the system remains perfectly in balance among a set of conflicting options, but remaining in this state for long periods of time are probabilistically unlikely.
The decision protocol is designed to amplify and reinforce random perturbations encountered during sampling, leading to a decision.

Snowball's voting protocol takes advantage of the DAG structure that the system maintains.
Specifically, every node independently measures the amount of network-wide support for every new transaction it encounters by sampling a subset of its peers and querying them for whether they could accept this transaction, and by extension, its ancestors.
This is, in effect, a one-time sub-sampled vote on this transaction.
If more than a threshold of responders vote positively, this transaction collects a \emph{chit}.
Nodes associate a second metric, called \emph{confidence}, with every transaction, where confidence is the sum of chits in the progeny of that transaction.
In a bit of seemingly circular logic, nodes vote positively on transactions only when they have high confidence for all transactions in the ancestor set.
In practice, the circulatory is broken by an initial arbitrary preference for first-seen transactions, and the genesis transaction as a common ancestor for bootstrapping.

Overall, %YYY
by repetitive voting that yields chits building up confidence,
guiding future voting towards the same direction, a positive
feedback is rapidly established once a perturbation is detected, so the network
topples towards the same decision in collaboration of all correct nodes, which
results in a protocol with probabilistic, efficient guarantee.

\subsection{Specification}
\label{subsection:specification}

% TODO add definitions on "what is compatible", "what is a conflict set". Specification should stand on its own without the intuition

% TODO Remove the conflict set stuff
Let the system be comprised by a set of nodes $\mathcal{N}$, where each node is an independent processing unit that stores some local state. 
All nodes are connected to each other through a network and communicate by exchanging messages.
The local state can be modified by executing transactions
% -- denoted by $T_*$ --
that represent special types of messages that encode a sequence of deterministic commands.

Each node $u \in \mathcal{N}$ keeps track of all transactions it has learned about in set $\mathcal{T}_u$. 
This set is partitioned into mutually exclusive \emph{conflict sets},
% subsets, denoted by $\mathcal{P}_*$,
where each transaction in $\mathcal{T}_u$ belongs to exactly one conflict set. 
Transactions are determined to be conflicting based on a deterministic and known function which is run by every node. 
We denote by $\mathcal{P}_T$ the conflict set that transaction $T$ is in.
If $T_i$ and $T_j$ are conflicting transactions, then
$\mathcal{P}_{T_i} = \mathcal{P}_{T_j}$.
% Let $T_i$ and $T_j$ be two transactions that belong in the same conflict set $\mathcal{P}$.
% subset of $\mathcal{T}_u$, i.e. $T_i, T_j \in \mathcal{P}_b$. 
% We then refer to $T_i$ and $T_j$ as conflicting transactions, and to $\mathcal{P}$ as the partition of $T_i$ and $T_j$. 
% For notational ease, we refer to $\mathcal{P}$ as $\mathcal{P}_{T_i}$ or $\mathcal{P}_{T_j}$. 
% The index of the latter two partitions intuitively refers to the partition that $T_i$ and $T_j$ belong both belong to, respectively, even though ultimately both transactions belong to the same partition $\mathcal{P}_b$.

%All transactions in $\mathcal{T}$ are partitioned by some pre-defined equivalence relation into disjoint conflict sets.
Part of the information encoded in the each transaction $T$ are links to other transactions, called parent transactions. 
We write $T' \gets T$ if $T$ has a parent link to a transaction $T'$.
Additionally, the ``$\stackrel{*}{\gets}$''-relation is the reflexive transitive closure of the ``$\gets$''-relation; 
namely, if $T' \stackrel{*}{\gets} T$, then there exists a path from $T$ to $T'$. 
These links, when aggregated amongst all known transactions, generate a DAG. 
Additionally, every transaction $T$ has associated with it a binary value chit, $c_T$, that stores the results of a one time vote.
The chit is not encoded in the transaction, but rather is computed as part of a voting mechanism. 
Based on this DAG, each node can compute a confidence value, $d(T)$, that sums up the number of chits in $T$'s progeny as follows:
\[ d(T) = \sum_{T' \in \mathcal{T}, T \stackrel{*}{\gets} T'}c_{T'}.\]

Each node $u$ maintains its own local list of known nodes $\mathcal{N}_u \subseteq \mathcal{N}$ that comprise the system.
For simplicity of exposition, we assume that all nodes share the same view, i.e. $\mathcal{N}_u = \mathcal{N}$, and later discuss how this assumption can be relaxed. 
% TODO talk to ted and ask how to rephrase this. "the data structure" is unclear, and so is immutability
Note that although $\mathcal{T}$ can vary between nodes, the inferred DAG will always be compatible between any pair of nodes.
% , because the data structure is immutable.
Specifically, if $T' \gets T$, transaction $T$ has a parent link to $T'$, then every node in the system will eventually see the same relation $T' \gets T$; on the other hand, if $T' \cancel{\gets} T$, transaction $T'$ is not a parent of $T$, then no nodes will end up with $T' \gets T$.
%Here, notation ``$T' \gets T$'' means $T'$ is one of $T$'s immediate parents, and
%``$\stackrel{*}{\gets}$''-relation is the reflexive transitive closure of ``$\gets$''-relation; namely,
%if $T' \stackrel{*}{\gets} T$, there exists a path from $T$ to $T'$. Specially, ``$T \stackrel{*}{\gets} T$''.

Each node implements an event-driven state machine, centered around a ``poll'' event which serves both to solicit votes on a transaction and, simultaneously, notify other nodes of the existence of the newly discovered transaction.

When node $u$ discovers a transaction $T$ for the first time, through a poll message it receives, it will start a one-time polling process by sampling $k$ other nodes uniformly at random.
This polling process starts by adding $T$ to $\mathcal{T}_u$, initializing the chit value $c_T$ to $0$, determining the set of conflicting transactions $\mathcal{P}_T$, and then sending a recursive poll message for $T$ to the selected $k$ sample peers.

When a node receives a poll solicitation for a transaction $T$ that it has seen before, it votes in support or against $T$.
Votes reflect the suitability of transaction $T$ for acceptance, compared to the acceptance status of other transactions in $T$'s conflict set $\mathcal{P}_T$, recursively applied to all transactions in $T$'s ancestry.
They are computed by checking whether the confidence for $T$ is the highest among competing transactions $\forall T' \in \mathcal{P}_*$.
If every single ancestor of $T$ fulfills this criteria, the transaction is said to be \emph{strongly preferred}, and receives a unit (1) vote. A failure of this criteria at $T$ or any of its ancestors yields a null (0) vote.

Node $u$ will tally the responses it receives, and form an opinion on the suitability of the transaction for acceptance based on voting results.
Specifically, if $k$ responses have been received, it computes a new chit value $c_T$ for transaction $T$.
It sets $c_t$ to 1 if a threshold $\alpha k \in \mathbb{N}$ of the responses were positive, and to 0 otherwise.  Once computed, a chit value is never modified again.
If fewer than $k$ responses are returned, the node will retry by sampling additional nodes until it has acquired a total of $k$ responses.
% sampled uniformly at random.

The above process will yield a labeling of the DAG with chits $c_T$ and associated confidence for each transaction $T$.
Because chit values can only increase (from 0 to 1),
confidence values are monotonically increasing.

\begin{center}
    \input{figures/dag-cd}
    \captionof{figure}{Example of chits and confidence values.  Darker boxes mean transactions with greater confidence values.}\label{fig:dag-cd}
\end{center}

In Figure~\ref{fig:dag-cd}, notation $\langle c_T, d(T)\rangle$ is the pair of voting result
$c_T$ and confidence $d(T)$ tagged to each transaction $T$.  Notice $c_T$ is
immutable after voting, while $d(T)$ is subject to change as the DAG grows.
Given the
voting rule, the system will topple over to a single transaction in each
conflict set.  For example, in set~2, $T_2$ has much larger confidence than
$T_3$ so it will likely be voted $c_{T_2} = 1$ in the future.

%YYY
%what I really want to write here is that there exists some equation X, that gives us the probability that
%T will get reverted. As a probabilistic protocol, we cannot provide a clear commit point. So applications
%can at best decide to commit a transaction when the probability of reversion is low enough for their use.
%In this sense, we are exactly like Nakamoto consensus.
%There should be a function $\gamma$ that gives us the probability of reversion.

Finally, unlike deterministic protocols, Snowball leaves the finalization of a
transaction to the application, by providing a predicate associated with
probabilistic guarantee.  The \textsc{isCommitted} predicate becomes true when $T$ is the
preferred transaction in its conflict set, and the confidence $d(T)$ exceeds
the sum of the confidence for all conflicting transactions by a constant
$\beta$ (XXX change here). As a probabilistic protocol, the truth of this predicate could be flipped, though,
with very low probability depending on the security parameter $\beta$. The
application using Snowball achieves the finality by choosing the appropriate
security parameters that deliver the desired probabilistic bound of safety:
given the observation of \textsc{isCommitted} being true, the
probability of all correct nodes eventually stick to true.

%The Snowball protocol relies on two security parameters $\alpha$ and $\beta$.
%XXX describe alpha and beta, and how they are selected based on desired probability of loss
%XXX if transactions carry monetary amounts, one can bound the actual dollar loss by multiplying
%XXX \gamma with the amount of the transaction.

The final outstanding component of the protocol is transaction generation, specifically, the choice of parents for a newly minted transactions.
We note that this choice does not affect the safety guarantees of the protocol, as discussed later in our analysis in Subsection \ref{subsec:safetyanalysis}. 
%Because for a particular transaction in the DAG, its confidence solely depends on the chits in its own progeny. 
%Moreover, since the \textsc{isCommitted} predicate is determined by the difference of confidence between transactions in the same conflict set, the only thing that matters in affecting the difference is which progeny set future transactions attach them to, rather than the microscopic DAG structure within the progeny sets.
However, parent selection can significantly affect the liveness of certain transactions, where transactions with badly chosen parents can remain stuck, forever unable to accrue the confidence required for commitment. 
In Section~\ref{sec:parent}, we provide three heuristics,
\emph{random}, \emph{frontier}, and \emph{adaptive}, and show that adaptive parent selection leads to good liveness properties that the \textsc{isCommitted} predicate will be probabilistically guaranteed to hold eventually for virtuous transactions.
%YYY what is good liveness?
A virtuous transaction is guaranteed to have a singleton conflict set consisting of just itself.

Figure~\ref{fig:gossipchain-ongen} shows the initialization of
a node's variables as well as what happens when a node learns of a
new transaction.  Section~\ref{sec:parent} discusses parent
selection from $\mathcal{T}$ for new transactions (which does not
affect safety).

\begin{figure}
\begin{center}
\begin{algorithmic}[1]
    \small
    \Procedure{init}{}
        \State $\mathcal{T} \coloneqq \emptyset$ \hspace{1ex}\textrm{// the set of known transactions}
        \State $\mathcal{Q} \coloneqq \emptyset$ \hspace{1ex}\textrm{// the set of polled transactions}
    \EndProcedure
    \Procedure{onGenerateTx}{$\mathrm{data}$}
        \State $\mathrm{edges} \coloneqq \{T' \gets T: T' \in \Call{parentSelection}{\mathcal{T}}\}$
        \State $T \coloneqq \texttt{Tx}(\mathrm{data}, \mathrm{edges})$
        \State $\mathcal{T} \coloneqq \mathcal{T} \cup \{T\}$
    \EndProcedure
    \Procedure{onReceiveTx}{$T$}
        \If{$T \notin \mathcal{T}$}
            \State $\mathcal{P}_T \coloneqq \mathcal{P}_T \cup \{T\}$
            \State $\mathcal{T} \coloneqq \mathcal{T} \cup \{T\}$
            \State $c_T \coloneqq 0$.
        \EndIf
    \EndProcedure
    \captionof{figure}{Transaction generation and bookkeeping.}\label{fig:gossipchain-ongen}
\end{algorithmic}
\end{center}
\end{figure}

% In Figure~\ref{fig:dag-conflicting-set}, there are 9 transactions and 6
% conflict sets. Solid edges represent parent relation and purple area stands for
% conflict set grouping.  Transaction $T_8$ has two parents $T_4$ and $T_5$ while
% the others only have one.

Figure~\ref{fig:gossipchain-main} shows the main loop of the core protocol.
In each iteration, the loop attempts to select a transaction $T$ that has not
yet been polled, but whose ancestors have all been polled.  (If no such
transaction exists, the loop will stall until a new transaction is
added to $\mathcal{T}$.)
The loop then selects $k$ peers and polls those peers.
If more than $\alpha k$ of those peers return a positive response,
the chit value is set to~1.  Then $T$ is added to the set $\mathcal{Q}$
so it will never be polled again by the node.
(The code that selects additional peers if some of the $k$ peers are
unresponsive is omitted for simplicity.)

\begin{figure}
\begin{center}
\small
\begin{algorithmic}[1]
    \Procedure{gossipChain}{}
        \While {\texttt{true}}
            \State \textrm{// select Tx with polled ancestors}
            \State\textrm{find  $T$ that satisfies }%\vspace*{-.6\baselineskip}
            $T \in \mathcal{T} \land T \notin \mathcal{Q}$
            %\begin{align*}\hspace{2ex}
            %    T \in \mathcal{T} &\land T \notin \mathcal{Q} \\
            %            &\land (\forall T', T' \gets T: T' \in \mathcal{Q})
            %\end{align*}
            \State \textrm{// randomly sample from the known nodes}
            \State $\mathcal{K} \coloneqq \texttt{sample}(\mathcal{N}, k)$
            \State $P \coloneqq \sum_{i \in \mathcal{K}}\texttt{poll}(i, T)$
            \If{$P > \alpha\cdot k$}
                \State $c_T \coloneqq 1$
            \EndIf
            \State\textrm{// otherwise, }$c_T$\textrm{ remains 0 forever}
            \State $\mathcal{Q} \coloneqq \mathcal{Q} \cup \{T\}$ \hspace {1ex} \textrm{// mark T as polled}
        \EndWhile
    \EndProcedure
    \captionof{figure}{The main gossip loop.}\label{fig:gossipchain-main}
\end{algorithmic}
\end{center}
\end{figure}

% TODO IMPORTANT The isPreferred must imply that T is polled, but current writing can be not polled. Also, do you do max over only polled transactions? 
\begin{figure}
\begin{center}
\small
\begin{algorithmic}[1]
    \Function{isPreferred}{$T$}
        \State \Return $d(T) = \max_{T' \in \mathcal{P}_T} d(T')$
    \EndFunction
    \Function{isStronglyPreferred}{$T$}
        \State \Return $\forall T'\in\mathcal{T}, T' \stackrel{*}{\gets} T: \Call{isPreferred}{T'}$
    \EndFunction
    \Procedure{onPoll}{$j, T$}
        \State \Call{onReceiveTx}{$T$}
        \If{\Call{isStronglyPreferred}{$T$}}
            \State $\Call{respond}{j, 1}$\hspace{2ex}\textrm{// vote yes}
        \Else
            \State $\Call{respond}{j, 0}$\hspace{2ex}\textrm{// vote no}
        \EndIf
    \EndProcedure
    \captionof{figure}{The voting rule.}\label{fig:gossipchain-onpoll}
\end{algorithmic}
\end{center}
\end{figure}

Figure~\ref{fig:gossipchain-onpoll} shows what happens when a node
receives a poll for a transaction $T$ from a peer $j$.
First it adds $T$ to $\mathcal{T}$, unless it already has it.
Then it determines if $T$ is currently strongly preferred.
If so, the node returns a positive response to peer $j$.
If not, the node returns a negative response.

\begin{figure}
\begin{center}
\begin{algorithmic}[1]
    \small
    \Function{isCommitted}{$T$}
        \State\Return
            \vspace*{-.8\baselineskip}
        \begin{align*}
            \Call{isPreferred}{T} &\land d(T) \\
            &- \sum_{T' \in \mathcal{P}_T, T' \neq T}{d(T')} \ge \beta
        \end{align*}
    \EndFunction
    \captionof{figure}{The commit condition.}\label{fig:gossipchain-diff-commit}
\end{algorithmic}
\end{center}
\end{figure}

Finally, Figure~\ref{fig:gossipchain-diff-commit} shows pseudo-code for
the commit decision for a transaction $T$.

\subsection{Parent Selection}
\label{sec:parent}

The goal of the parent selection policy is to yield a well-structured DAG that maximizes
the likelihood that virtuous transactions will be quickly accepted by the network. While this
policy will not affect the safety of the protocol, it plays a crucial role in determining
the shape and structure of the DAG, and affects liveness.

There are inherent trade-offs in the parent selection policy: selecting well-accepted parents make it
more likely for a transaction to find support, but can lead to vote dilution; as well, selecting more
recent parents at the frontier of the DAG can lead to stuck transactions, as the parents may turn out
to be in conflict and remain unsupported.  In the following discussion, we illustrate this dilemma.
We assume that every transaction will select a small number, $p$, of parents, provided as a protocol parameter.
We focus on the the selection of eligible parent set, from which a subset of size $p$ can be chosen at random.

Perhaps the simplest idea is to mint a fresh transactions with parents picked uniformly at random
among those transactions that are already well-accepted. Specifically, we can adopt the predicate
used in the voting rule to determine eligible parents on which a node would vote positively, as follows:
\[
    \mathcal{E} = \{T: \forall T\in\mathcal{T}, \textsc{isStronglyPreferred}(T)\}.
\]

But this strategy has the drawback that it can suffer from vote dilution and lead to starved transactions.
This heuristic will yield to large sets of eligible parents, consisting mostly of historical, old transactions.
When a node samples the transactions uniformly from $\mathcal{E}$, the resulting DAG
will have large, ever-increasing fan-out. Because new transactions will have scarce progenies,
the voting process will take a long time to build the required confidence on any given new transaction.

In contrast, efforts to reduce fan-out and control the shape of DAG by selecting the recent transactions at
decision frontier suffers from another problem. Most recent transactions will have very small (usually zero) $d$ values,
simply because they do not have enough descendants. Further, their conflict sets may not be well-distributed
and well-known across the network, leading to a parent attachment under a transaction that will never be supported.
This means the best transactions to choose lie somewhere near the frontier, but not too far deep in history.

In adaptive selection, a node starts at the frontier and retreats towards the genesis transaction until it finds an eligible parent.
The frontier transactions are those that currently do not have any children in the node's $\mathcal{T}$.
The set of strongly preferred transactions $\mathcal{E}$ is filtered by an additional criterion that the confidence should be larger than zero. This filters out transactions from unstable conflict set having zero confidence.
If there exist $p$ transactions in the filtered $\mathcal{E}$, then the adaptive selection will just yield them.
%YYY why isn't this bad, for the reasons in the previous paragraph?
Otherwise, the policy tries the parents of the transactions in $\mathcal{E}$, thus increasing the chance of finding more stabilized transactions as it retreats.
The retreating search will terminate when it reaches the genesis transaction.
Formally, the selected parents in this adaptive selection policy is precisely:

\begin{center}
\begin{algorithmic}[1]
    \small
    \Function{parentSelection}{$\mathcal{T}$}
        \State $\mathcal{E}' \coloneqq \{T: d(T) > 0, \forall T\in\mathcal{E}\}$.
        \State \Return $\{T: T \in \mathcal{E}' \land \forall T'\in\mathcal{T}, T \gets T', T' \notin \mathcal{E}'\}$.
    \EndFunction
    \captionof{figure}{Adaptive parent selection.}\label{fig:gossipchain-parent}
\end{algorithmic}
\end{center}
%YYY what are you memoizing and why does this help? Also, can't an attacker create a gazillion nodes at the frontier and dilute your efforts?
The formal definition involves checking the modified frontier condition each
time it retreats and reaches a vertex. In fact, each ``forall'' quantifier
demands an $O(|\mathcal{T}|)$ time, so the overall cost is
$O(|\mathcal{T}|^2)$. However, by doing search from the frontier with
memoization, the quadratic cost could be avoided. Moreover, using an
approximation technique in Section~\ref{sec:pruning}, it is possible to replace the
$|\mathcal{T}|$ by a constant number of transactions in practice.

Attackers cannot suddenly create a surge of vertices at the frontier to dilute
the progress of Snowball, because with the above policy, only the transactions
with positive confidence will be considered as parents. The polling happens via
gossiping, and each query has to wait for the response from $k$ randomly chosen
peers. The pace is then coordinated by all correct nodes. So even if an
attacker increases the fan-out of the DAG, those new vertices will not
get much attention unless its has confidence larger than zero, in which the
progress is not diluted.

### 2020-11-03 calculations/hyper.py

@@ -1,0 +1,102 @@
+import multiprocessing
+from decimal import Decimal
+import decimal
+decimal.getcontext = 1000
+import matplotlib.pyplot as plt
+import math
+import numpy as np
+
+def nck(n, k):
+    return Decimal(math.factorial(n))/(Decimal(math.factorial(k))*Decimal(math.factorial(n-k)))
+
+def hyperpmf(population, successes, draws, observed_successes):
+    ret = nck(successes, observed_successes)*(nck(population-successes, draws-observed_successes)/nck(population, draws))
+    return ret
+
+def hyperormore(tup):
+    total = Decimal('0')
+    N = tup[0]
+    K = tup[1]
+    n = tup[2]
+    k = tup[3]
+    for i in range(k, n+1):
+        total += hyperpmf(N, K, n, i)
+    return total
+
+# with multiprocessing.Pool(20) as p:
+#     ranges = range(5000, 8200, 200)
+#     y = p.map(hyperormore, [(10000, i, 100, 80) for i in ranges])
+
+# plt.plot(ranges, y)
+# plt.show()
+
+from scipy.stats import hypergeom
+
+def hyperormore_fast(pop, succ, sample, threshold):
+    rv = hypergeom(pop, succ, sample)
+    return 1 - rv.cdf(threshold) + rv.pmf(threshold)
+
+import pykov
+
+'''
+Attempting to instantiate the Markov process that describes an attack where 
+the Byzantine nodes partition the network into some H1 and H2 sets, and then 
+attempt to constantly try to prefer one trasanaction over the other. 
+'''
+import random
+data = []
+trials = 10000
+ksample = 50
+alpha = 40
+total_nodes = 100000
+total_byz = 20000
+ranges = range(40000, 80000, 50)
+for i in ranges:
+    total_h1 = i
+    total_h2 = total_nodes - total_h1 - total_byz
+    total_honest = total_h1 + total_h2
+    # honest_nodes_t1 = [0]*total_honest
+    # honest_nodes_t2 = [0]*total_honest
+    
+    # for trial in range(trials):
+    #     index = random.randint(0, total_honest-1)
+    #     if index < total_h1:
+    #         x = hyperormore_fast(total_nodes, total_h1 + total_byz, ksample, alpha)
[…]

### 2020-11-03 calculations/loadpltfigure.py

@@ -1,0 +1,5 @@
+import matplotlib.pyplot as plt
+import pickle as pl
+print("TRYING TO LOAD PICKLE")
+fig = pl.load(open('output.pickle','rb'))
+plt.show()

### 2020-11-03 calculations/partitionspread.py

@@ -1,0 +1,143 @@
+import random
+import matplotlib.pyplot as plt
+from scipy.stats import hypergeom
+import math
+from multiprocessing import Pool
+from scipy.special import comb
+from decimal import Decimal
+import decimal
+decimal.getcontext = 100000
+
+# import functools
+# import operator as op
+# def ncr(n, r):
+#     print(n, r)
+#     r = min(r, n-r)
+#     if r == 0: return 1
+#     numer = functools.reduce(op.mul, range(n, n-r, -1))
+#     denom = functools.reduce(op.mul, range(1, r+1))
+#     return numer//denom
+
+def ncr(n, r):
+    return comb(n, r, exact=True)
+
+def hyper(N, succ, sample, x):
+    succi = int(succ)
+    outcome = Decimal(0)
+    for i in range(x, sample+1):
+        # print(":::", succi, i, N-succi, sample-i, sample, N)
+        top = ncr(succi, i) * ncr(N - succi, sample - i)
+        bottom = ncr(N, sample)
+        outcome += (Decimal(top)/Decimal(bottom))
+    return outcome
+    # rv = hypergeom(N, succi, sample)
+    # return decimal.Decimal(1.0) - decimal.Decimal(rv.cdf(x)) + decimal.Decimal(rv.pmf(x))
+
+def exp(t):
+    k = t[0]
+    alpha = t[1]
+    N = t[2]
+    B = t[3]
+    C = N - B
+    C1 = t[4]
+    C2 = C - C1
+    Z1 = C1
+    Z2 = C2
+    T1sum = Decimal(0)
+    T2sum = Decimal(0)
+    # Time 0, base case
+    gamma1 = Decimal(Z1/C)*hyper(N, C1 + B, k, alpha)
+    gamma2 = Decimal(Z2/C)*hyper(N, C1, k, alpha)
+    T1sum += (Decimal(gamma1) + Decimal(gamma2))
+    xi1 = Decimal(Z1/C)*hyper(N, C2, k, alpha)
+    xi2 = Decimal(Z2/C)*hyper(N, C2 + B, k, alpha)
+    T2sum += (Decimal(xi1) + Decimal(xi2))
+    gammas = [Decimal(gamma2)]
+    xis = [Decimal(xi2)]
+    # other cases
+    for i in range(1, C):
+        # print(i, T1sum, T2sum)
[…]

### 2020-11-03 calculations/partitionspread_markov.py

@@ -1,0 +1,61 @@
+import random
+import math
+from multiprocessing import Pool
+
+import numpy as npy
+import mpmath as mpmath
+import matplotlib.pyplot as plt
+from scipy.stats import hypergeom
+from scipy.special import comb
+
+mpmath.mp.dps = 200
+
+def ncr(n, r):
+    return comb(n, r, exact=True)
+
+def hyper(N, succ, sample, x):
+    outcome = mpmath.mpf(0)
+    for i in range(x, sample+1):
+        # print(":::", succi, i, N-succi, sample-i, sample, N)
+        top = ncr(succ, i) * ncr(N - succ, sample - i)
+        bottom = ncr(N, sample)
+        outcome += (mpmath.mpf(top)/mpmath.mpf(bottom))
+    return outcome
+
+def exp(t):
+    k = t[0]
+    alpha = t[1]
+    N = t[2]
+    B = t[3]
+    C = N - B
+    C1 = t[4]
+    C2 = C - C1
+    Z1 = C1
+    Z2 = C2
+    T1sum = mpmath.mpf(0)
+    T2sum = mpmath.mpf(0)
+    # Generate matrix
+    npy.matrix
+    return (t[4], T1sum, T2sum)
+
+def main():
+    k = 40
+    alpha = 32
+    N = 1000
+    B = 200
+    step = 100
+    with Pool(8) as p:
+        r = p.map(exp, [(k, alpha, N, B, i) for i in range(0, int((N - B)/2) + step, step)])
+    # r = [exp(200)]
+    x = [i[0] for i in r]
+    y1 = [0 if math.isnan(i[1]) else i[1] for i in r]
+    y2 = [0 if math.isnan(i[2]) else i[2] for i in r]
+    plt.plot(x, y1, label="C1, k:{}, a:{}, N:{}, B:{}".format(k, alpha, N, B))
+    plt.plot(x, y2, label="C2, k:{}, a:{}, N:{}, B:{}".format(k, alpha, N, B))
+    print(min(y1), max(y1))
+    print(min(y2), max(y2))
+    plt.legend()
+    plt.show()
+
+if __name__ == '__main__':
+    main()

### 2020-11-03 calculations/partitionspread_withchildren.py

@@ -1,0 +1,250 @@
+import random
+import matplotlib.pyplot as plt
+from scipy.stats import hypergeom
+import math
+import mpmath as mpm
+from multiprocessing import Pool
+from scipy.special import comb
+from termcolor import colored
+import sys
+import copy
+import time
+
+mpm.mp.dps = 30
+
+def ncr(n, r):
+    return comb(n, r, exact=True)
+
+def hyper(N, succ, sample, x):
+    outcome = mpm.mpf(0)
+    for i in range(x, sample+1):
+        top = ncr(succ, i) * ncr(N - succ, sample - i)
+        bottom = ncr(N, sample)
+        outcome += (mpm.mpf(top)/mpm.mpf(bottom))
+    return outcome
+
+from scipy.special import erf
+
+def trunc_normal(x, mu, sigma, a, b):
+    return phi((x - mu)/sigma)/(sigma*(cdf((b-mu)/sigma) - cdf((a-mu)/sigma)))
+
+def phi(eps):
+    return (1/math.sqrt(2*math.pi))*math.pow(math.e, -0.5*math.pow(eps, 2))
+
+def cdf(x):
+    return 0.5*(1 + erf(x/math.sqrt(2)))
+
+def normal(x, sigma):
+    return (1/math.sqrt(2*math.pi*math.pow(sigma, 2)))*math.pow(math.e, -0.5*(x**2)*(1/math.pow(sigma, 2)))
+
+def lineary(index, length):
+    slope = -(2/math.pow(length, 2))
+    return (slope*index) + (2/length)
+
+class HISTORY():
+    def __init__(self, length):
+        self.array = []
+        self.probabilities = None
+        self.length = length
+    def prepend(self, C1nodes, C2nodes):
+        self.array.insert(0, [copy.deepcopy(C1nodes), copy.deepcopy(C2nodes)])
+        if len(self.array) == self.length + 1:
+            del self.array[self.length]
+        self.probabilities = [lineary(i+0.5, len(self.array)) for i in range(len(self.array))]
+
+def exp(t):
+    k = t[0]
+    alpha = t[1]
+    N = t[2]
+    B = t[3]
+    C = N - B
+    C1 = t[4]
+    C2 = C - C1
+    Z1 = C1
+    Z2 = C2
+    # T1globalpref = C1
+    # T2globalpref = C2
+    T1sum = mpm.mpf(0)
+    T2sum = mpm.mpf(0)
+    # gammas = []
+    # xis = []
[…]

### 2020-11-03 calculations/partitionspread_withchildren_optimized.py

@@ -1,0 +1,280 @@
+import random
+import matplotlib.pyplot as plt
+from scipy.stats import hypergeom
+import math
+import mpmath as mpm
+from multiprocessing import Pool
+from scipy.special import comb
+from termcolor import colored
+import sys
+import copy
+import time
+import numpy as np
+
+mpm.mp.dps = 30
+
+def ncr(n, r):
+    return comb(n, r, exact=True)
+
+def hyper(N, succ, sample, x):
+    outcome = mpm.mpf(0)
+    for i in range(x, sample+1):
+        top = ncr(succ, i) * ncr(N - succ, sample - i)
+        bottom = ncr(N, sample)
+        outcome += (mpm.mpf(top)/mpm.mpf(bottom))
+    return outcome
+
+from scipy.special import erf
+
+def trunc_normal(x, mu, sigma, a, b):
+    return phi((x - mu)/sigma)/(sigma*(cdf((b-mu)/sigma) - cdf((a-mu)/sigma)))
+
+def phi(eps):
+    return (1/math.sqrt(2*math.pi))*math.pow(math.e, -0.5*math.pow(eps, 2))
+
+def cdf(x):
+    return 0.5*(1 + erf(x/math.sqrt(2)))
+
+def normal(x, sigma):
+    return (1/math.sqrt(2*math.pi*math.pow(sigma, 2)))*math.pow(math.e, -0.5*(x**2)*(1/math.pow(sigma, 2)))
+
+def lineary(index, length):
+    slope = -(2/math.pow(length, 2))
+    return (slope*index) + (2/length)
+
+class HISTORY():
+    def __init__(self, length):
+        self.array = []
+        self.probabilities = None
+        self.length = length
+    def prepend(self, C1nodes, C2nodes):
+        self.array.insert(0, [copy.deepcopy(C1nodes), copy.deepcopy(C2nodes)])
+        if len(self.array) == self.length + 1:
+            del self.array[self.length]
+        self.probabilities = [lineary(i+0.5, len(self.array)) for i in range(len(self.array))]
+
+def exp(t):
+    k = t[0]
+    alpha = t[1]
+    N = t[2]
+    B = t[3]
+    C = N - B
+    C1 = t[4]
+    C2 = C - C1
+    Z1 = C1
+    Z2 = C2
+    T1sum = mpm.mpf(0)
+    T2sum = mpm.mpf(0)
+    gammas = mpm.mpf(0)
+    xis = mpm.mpf(0)
[…]

### 2020-11-03 calculations/partitionspread_withchildren_optimized_varying_partitions.py

@@ -1,0 +1,260 @@
+import random
+import matplotlib.pyplot as plt
+from scipy.stats import hypergeom
+import math
+import mpmath as mpm
+from multiprocessing import Pool
+from scipy.special import comb
+from termcolor import colored
+import sys
+import copy
+import time
+
+mpm.mp.dps = 30
+
+def ncr(n, r):
+    return comb(n, r, exact=True)
+
+def hyper(N, succ, sample, x):
+    outcome = mpm.mpf(0)
+    for i in range(x, sample+1):
+        top = ncr(succ, i) * ncr(N - succ, sample - i)
+        bottom = ncr(N, sample)
+        outcome += (mpm.mpf(top)/mpm.mpf(bottom))
+    return outcome
+
+from scipy.special import erf
+
+def trunc_normal(x, mu, sigma, a, b):
+    return phi((x - mu)/sigma)/(sigma*(cdf((b-mu)/sigma) - cdf((a-mu)/sigma)))
+
+def phi(eps):
+    return (1/math.sqrt(2*math.pi))*math.pow(math.e, -0.5*math.pow(eps, 2))
+
+def cdf(x):
+    return 0.5*(1 + erf(x/math.sqrt(2)))
+
+def normal(x, sigma):
+    return (1/math.sqrt(2*math.pi*math.pow(sigma, 2)))*math.pow(math.e, -0.5*(x**2)*(1/math.pow(sigma, 2)))
+
+def lineary(index, length):
+    slope = -(2/math.pow(length, 2))
+    return (slope*index) + (2/length)
+
+class HISTORY():
+    def __init__(self, length):
+        self.array = []
+        self.probabilities = None
+        self.length = length
+    def prepend(self, C1nodes, C2nodes):
+        self.array.insert(0, [copy.deepcopy(C1nodes), copy.deepcopy(C2nodes)])
+        if len(self.array) == self.length + 1:
+            del self.array[self.length]
+        self.probabilities = [lineary(i+0.5, len(self.array)) for i in range(len(self.array))]
+
+def exp(t):
+    round_counter = t[5]
+    max_round = t[6]
+    k = t[0]
+    alpha = t[1]
+    N = t[2]
+    B = t[3]
+    
+    T1sum = mpm.mpf(0)
+    T2sum = mpm.mpf(0)
+    gammas = mpm.mpf(0)
+    xis = mpm.mpf(0)
+    NUMTX = max_round
+
[…]

### 2020-11-03 calculations/partitionspread_withchildren_simulation.py

@@ -1,0 +1,261 @@
+import random
+import matplotlib.pyplot as plt
+from scipy.stats import hypergeom
+import math
+import mpmath as mpm
+from multiprocessing import Pool
+from scipy.special import comb
+from termcolor import colored
+import sys
+import copy
+import time
+import numpy as np
+
+mpm.mp.dps = 30
+
+def ncr(n, r):
+    return comb(n, r, exact=True)
+
+def hyper(N, succ, sample, x):
+    outcome = mpm.mpf(0)
+    for i in range(x, sample+1):
+        top = ncr(succ, i) * ncr(N - succ, sample - i)
+        bottom = ncr(N, sample)
+        outcome += (mpm.mpf(top)/mpm.mpf(bottom))
+    return outcome
+
+from scipy.special import erf
+
+def trunc_normal(x, mu, sigma, a, b):
+    return phi((x - mu)/sigma)/(sigma*(cdf((b-mu)/sigma) - cdf((a-mu)/sigma)))
+
+def phi(eps):
+    return (1/math.sqrt(2*math.pi))*math.pow(math.e, -0.5*math.pow(eps, 2))
+
+def cdf(x):
+    return 0.5*(1 + erf(x/math.sqrt(2)))
+
+def normal(x, sigma):
+    return (1/math.sqrt(2*math.pi*math.pow(sigma, 2)))*math.pow(math.e, -0.5*(x**2)*(1/math.pow(sigma, 2)))
+
+def lineary(index, length):
+    slope = -(2/math.pow(length, 2))
+    return (slope*index) + (2/length)
+
+
+class HISTORY():
+    def __init__(self, length):
+        self.array = []
+        self.probabilities = []
+        self.probabilities_normalized = []
+        self.length = length
+        
+    def prepend(self, nodes):
+        self.array.insert(0, copy.deepcopy(nodes))
+        if len(self.array) == self.length + 1:
+            del self.array[self.length]
+        if len(self.probabilities) < self.length:
+            self.probabilities = [lineary(i+0.5, len(self.array)) for i in range(len(self.array))]
+            self.probabilities_normalized = [self.probabilities[0]]
+            for i in range(1, len(self.probabilities)):
+                self.probabilities_normalized.append(self.probabilities[i] + self.probabilities_normalized[i-1])
+            self.probabilities_normalized = [int(i*10000) for i in self.probabilities_normalized]
+
+
+    def select_random_history(self):
+        # print("Probabilities:", self.probabilities)
[…]

### 2020-11-03 calculations/partitionspread_withchildren_withprob.py

@@ -1,0 +1,238 @@
+import random
+import matplotlib.pyplot as plt
+from scipy.stats import hypergeom
+import math
+import mpmath as mpm
+from multiprocessing import Pool
+from scipy.special import comb
+from termcolor import colored
+import sys
+import copy
+import time
+import numpy as np
+
+mpm.mp.dps = 30
+
+def ncr(n, r):
+    return comb(n, r, exact=True)
+
+def hyper(N, succ, sample, x):
+    outcome = mpm.mpf(0)
+    for i in range(x, sample+1):
+        top = ncr(succ, i) * ncr(N - succ, sample - i)
+        bottom = ncr(N, sample)
+        outcome += (mpm.mpf(top)/mpm.mpf(bottom))
+    return outcome
+
+def binom(p, k, alpha):
+    outcome = mpm.mpf(0)
+    for i in range(alpha, k+1):
+        outcome += ncr(k, i) * (p**i) * ((1 - p)**(k - i))
+    return outcome
+
+def exp(t):
+    
+    k = t[0]
+    alpha = t[1]
+    N = mpm.mpf(t[2])
+    B = mpm.mpf(t[3])
+    C = mpm.mpf(N - B)
+    C1 = mpm.mpf(t[4])
+    C2 = C - C1
+    
+    T1sum = mpm.mpf(0)
+    T2sum = mpm.mpf(0)
+    gammas = mpm.mpf(0)
+    xis = mpm.mpf(0)
+    
+    NUMTX = int(C*2)
+
+    # Each probs is P(T1>T2),   P(T2>T1),   P(T1==T2)
+    C1Probs = [     mpm.mpf(1), mpm.mpf(0), mpm.mpf(0)]
+    C2Probs = [     mpm.mpf(0), mpm.mpf(1), mpm.mpf(0)]
+
+    # These are the expected sums in each bucket
+    C1Exps = [mpm.mpf(0), mpm.mpf(0)]
+    C2Exps = [mpm.mpf(0), mpm.mpf(0)]
+
+    log = []
+
+    T1globalpref = C1
+    T2globalpref = C2
+
+    T1globalprefhistory = []
+    T2globalprefhistory = []
+
+    CiProbshistory = []
+
+    for i in range(0, 100):
+        print(i, colored([float(_i) for _i in C1Probs], "red"), colored([float(_i) for _i in C2Probs], "green"))
+            # if i % 200 == 0:
+            #     print(i, NUMTX)
+        
[…]

### 2020-11-03 calculations/plotb.py

@@ -1,0 +1,140 @@
+import matplotlib.pyplot as plt
+from mpl_toolkits.mplot3d import Axes3D
+import matplotlib.colors as colors
+from matplotlib import cm
+from scipy.stats import hypergeom
+from matplotlib.colors import LogNorm
+import matplotlib.ticker as ticker
+import numpy as np
+
+def hyper(N, S, k, alpha):
+    rv = hypergeom(N, S, k)
+    return 1 - rv.cdf(alpha) + rv.pmf(alpha)
+
+def probOfStreak(numCoins, minHeads, headProb, saved=None):
+    if saved == None: saved = {}
+    ID = (numCoins, minHeads, headProb)
+    if ID in saved: return saved[ID]
+    else:
+        if minHeads > numCoins or numCoins <= 0:
+            result = 0
+        else:
+            result = headProb**minHeads
+            for firstTail in range(1, minHeads+1):
+                pr = probOfStreak(numCoins-firstTail, minHeads, headProb, saved)
+                result += (headProb**(firstTail-1))*(1-headProb)*pr
+        saved[ID] = result
+        return result
+
+def print_B_vs_alpha_liveness():
+    N = 2000
+    B = list(range(0, int(N/2), 1))
+    C1 = [int((N - i)/2) for i in B]
+    C2 = [N - B[i] - C1[i] for i in range(len(B))]
+    k = 20
+    alpha = list(range(int(k/2)+1, k+1))
+    z = []
+    for i in range(len(B)):
+        print(i)
+        zs = []
+        for j in range(len(alpha)):
+            r = hyper(N, N - B[i], k, alpha[j])
+            # print("{},{},{}".format(B[i], C1[i], a), r)
+            zs.append(r)
+        z.append(zs)
+    fig = plt.figure()
+    ax = fig.add_subplot(111, projection='3d')
+    x, y = np.meshgrid(alpha, B)
+    z = np.array(z)
+    print(x)
+    print(y)
+    print(z)
+    ax.plot_surface(x, y, z, cmap=cm.coolwarm,linewidth=0, antialiased=False)
+    plt.ylabel("Byzantine")
+    plt.xlabel("Alpha")
+    plt.show()
+
+def print_B_vs_alpha_safety():
+    N = 500
+    B = list(range(0, int(N/2), 1))
+    C1 = [int((N - i)/2) for i in B]
[…]

### 2020-11-03 calculations/runremotely.sh

@@ -1,0 +1,3 @@
+ssh -t ks2232@compute17.fractus.cs.cornell.edu python3 partitionspread_withchildren_optimized.py;
+scp ks2232@compute17.fractus.cs.cornell.edu:/home/ks2232/output.pickle output.pickle;
+python3 loadpltfigure.py

### 2020-11-03 calculations/simple_markov.py

@@ -1,0 +1,56 @@
+import pykov
+import matplotlib.pyplot as plt
+from mpl_toolkits.mplot3d import Axes3D
+from matplotlib import cm
+from scipy.stats import hypergeom
+import numpy as np
+
+def hyper(N, S, k, alpha):
+    rv = hypergeom(N, S, k)
+    return 1 - rv.cdf(alpha) + rv.pmf(alpha)
+
+def main():
+    # N = 100
+    # B = list(range(0, int(N/2), 1))
+    # C1 = [int((N - i)/2) for i in B]
+    # C2 = [N - B[i] - C1[i] for i in range(len(B))]
+    # k = 20
+    # alpha = list(range(int(k/2), k+1))
+    N = 100
+    B = 20
+    C1 = 40
+    C2 = 40
+    k = 20
+    alpha = 15
+    states = []
+    for i in range(int((C1+C2)/2), C1+C2+1):
+        states.append((i, C1+C2-i))
+    chain = pykov.Chain()    
+    for i in range(1, len(states)-1):
+        # print(states[i])
+        # Prob of going up is prob choosing red and getting maj blue
+        up =   (states[i][0]/(C1+C2))*hyper(N, states[i][1], k, alpha)
+        # Prob of going down is prob of choosing blue
+        down = (states[i][1]/(C1+C2))*hyper(N, states[i][0], k, alpha)
+        # Prob of staying is prob of choosing blue and getting maj blue 
+        # or choosing red and getting maj red
+        # stay = ((states[i][0]/(C1+C2))*hyper(N, states[i][0], k, alpha))+((states[i][1]/(C1+C2))*hyper(N, states[i][1], k, alpha))
+        stay = 1 - up - down
+        chain[(str(states[i]),str(states[i-1]))] = up
+        chain[(str(states[i]),str(states[i+1]))] = down
+        chain[(str(states[i]),str(states[i]))] = stay
+    chain[(str(states[0]),str(states[0]))] = 1 - (states[0][1]/(C1+C2))*hyper(N, states[0][0], k, alpha)
+    chain[(str(states[0]),str(states[1]))] = (states[0][1]/(C1+C2))*hyper(N, states[0][0], k, alpha)
+    chain[(str(states[len(states)-1]),str(states[len(states)-1]))] = 1 - (states[i][0]/(C1+C2))*hyper(N, states[i][1], k, alpha)
[…]

### 2020-11-03 kevinscripts/calculations.py

@@ -1,0 +1,282 @@
+import matplotlib.pyplot as plt
+import numpy as np
+from sympy import Symbol, Rational, binomial, expand_func
+import mpmath as mp
+import math
+import sys
+from multiprocessing import Pool
+sys.setrecursionlimit(200000)
+mp.mp.dps = 20000
+
+def hyper(N, suc, sample, x):
+    total = mp.mp.mpf(0)
+    bottom = binomial(N, sample)
+    for i in range(x, sample+1):
+        top = binomial(suc, i) * binomial(N-suc, sample-i)
+        total += (top/bottom)
+    return total
+
+def binom(trials, successes, p):
+    total = mp.mp.mpf(0)
+    p_ = mp.mp.mpf(p)
+    for i in range(successes, trials+1):
+        total += (binomial(trials, i) * mp.power(p_, i) * mp.power(1 - p_, trials - i))
+    return total
+
+def test1():
+    n = 1000
+    b = 200
+    k = 10
+    alpha = 8
+    step = 100
+    _range = range(step, n+step, step)
+    trials = 400
+    _successes = [int(i) for i in np.linspace(10, trials, num=5)]
+    for successes in _successes:
+        results = []
+        for i in _range:
+            print(i)
+            h = hyper(n, i, k, alpha)
+            results.append(binom(trials, successes, h))
+        plt.semilogy(_range, results, label=successes)
+    plt.plot(_range, [2**-32 for i in _range])
+    plt.legend()
+    plt.show()
+
+def test2():
+    _range = range(790, 810, 2)
+    for trials in [140]:
+        print(trials)
+        results = []
+        for i in _range:
+            h = hyper(1000, i, 10, 8)
+            results.append(binom(trials, int(trials*0.9), h))
+        plt.semilogy(_range, results, label=trials)
+    plt.plot(_range, [1/(10**math.log(2**(32), 10))]*len(_range))
+    plt.plot(_range, [0.5]*len(_range))
+    plt.legend()
+    plt.show()
+
+def test3():
+    trials = 350
+    results = []
+    _range = [800]
+    for i in _range:
+        print(i)
+        h = hyper(1000, i, 10, 8)
+        results.append(binom(trials, int(0.90*trials), h))
[…]

### 2020-11-03 kevinscripts/calculations2.py

@@ -1,0 +1,377 @@
+import matplotlib.pyplot as plt
+import numpy as np
+from multiprocessing import Pool
+from sympy import binomial, zeros, pprint
+import mpmath as mp
+import math
+import time
+mp.mp.dps = 20
+
+def hyper(N, suc, sample, x):
+    total = mp.mp.mpf(0)
+    bottom = binomial(N, sample)
+    for i in range(x, sample+1):
+        top = binomial(suc, i) * binomial(N-suc, sample-i)
+        total += (top/bottom)
+    return total
+
+def binom(trials, successes, p):
+    total = mp.mp.mpf(0)
+    p_ = mp.mp.mpf(p)
+    for i in range(successes, trials+1):
+        total += (binomial(trials, i) * mp.power(p_, i) * mp.power(1 - p_, trials - i))
+    return total
+
+def test3Helper(t):
+    n, k, alpha, c, b = t
+    for i in range(0, 205):
+        # print(i, ((0.5*c + i)*hyper(n, 0.5*c + i - 1, k, alpha)), ((0.5*c-i)*hyper(n, 0.5*c - i + b - 1, k, alpha)))
+        if (((0.5*c-i)/c)*hyper(n, 0.5*c + i, k, alpha)) >= (((0.5*c+i)/c)*hyper(n, 0.5*c - i + b, k, alpha)):
+            return i
+
+def test3( _range):
+    r = []    
+    with Pool(4) as p:
+        r = p.map(test3Helper, _range)
+    plt.plot([i[1] for i in _range], r, label="alpha = {}*k".format(_range[0][2]/_range[0][1]))
+
+def test4Helper(t):
+    n, k, alpha, c, b, consecutive = t
+    foundI = None
+    for i in range(0, 205):
+        if ((0.5*c+i)*hyper(n, 0.5*c + i - 1, k, alpha)) >= ((0.5*c-i)*hyper(n, 0.5*c - i + b - 1, k, alpha)):
+            print("Found i =", i, ";", n, k, alpha, c, b)
+            foundI = i
+            break
+            # return mp.power(hyper(n, 0.5*c + i - 1, k, alpha), consecutive)
+    for i in range(1, 100):
+        if mp.power(hyper(n, 0.5*c + foundI - 1, k, alpha), i) <= mp.power(2, -32):
+            return i*k
+
+def test4( _range):
+    r = []
+    with Pool(4) as p:
[…]

### 2020-11-03 kevinscripts/expected_extinction_time.py

@@ -1,0 +1,70 @@
+import random
+import math
+import itertools
+from scipy.stats import hypergeom
+from functools import reduce
+import operator
+
+def hyper(N, S, k, a):
+    rv = hypergeom(N, S, k)
+    return 1.0 - rv.cdf(a) + rv.pmf(a)
+
+def prod(iterable):
+    if not len(iterable):
+        return 1
+    return reduce(operator.mul, iterable, 1)
+
+def t1(births, deaths):
+    sigma = 0
+    for k in range(1, len(deaths)):
+        top = prod(births[0:k])
+        bottom = prod(deaths[0:k+1])
+        sigma += (top/bottom)
+    return (1/deaths[0])
+
+
+# def extinction_time(births, deaths):
+#     stays = [1-(births[i]+deaths[i]) for i in range(len(births))]
+#     print(births, deaths, stays)
+#     tm = t1(births, deaths)
+
+def An(n, p, q):
+    sigma = 0
+    for i in range(1, n-1):
+        print(p[1:i+1])
+        print(q[1:i+1])
+        print("-----------")
+        top = prod(q[1:i+1])
+        bottom = prod(p[1:i+1])
+        sigma += top/bottom
+    return sigma
+
+def mn1(n, p, q):
+    a = 1/An(n, p, q)
+    top_sigma = 0
+    for y in range(1, n-1):
+        bottom_sigma = 0
+        for z in range(1, y+1):
+            top = prod(q[z+1:y+1])
+            bottom = prod(p[z:y+1])
+            print("y {}, z {}, top {}, bottom {}".format(y, z, q[z+1:y+1], p[z:y+1]))
+            bottom_sigma += (top/bottom)
+        top_sigma += bottom_sigma
+    print(a * top_sigma)
+    return a * top_sigma
+
+def mn(x, p, q):
+    _mn1 = mn1(len(p), p, q)
+    
+
+# Birth death discrete chain with absorbing endpoints
+def vn(x, p, q):
+    f = 1/An(len(p), p, q)
+    sigma = 0
+    for i in range(x, len(p)-2):
+        top = prod(q[1:i+1])
+        bottom = prod(p[1:i+1])
+        sigma += (top/bottom)
+    return(f*sigma)
+
+print(mn(2, [0, 1/2, 1/2, 1/2, 0], [0, 1/2, 1/2, 1/2, 0]))

### 2020-11-03 kevinscripts/expected_time_to_epsilon_convergence_slush.py

@@ -1,0 +1,10 @@
+'''
+This graph plots the Figure "expected_time_to_epsilon_convergence_slush.pdf"
+'''
+
+import matplotlib.pyplot as plt
+
+results = [42.505, 41.67, 41.2525, 41.002, 41.66833333333334, 41.43, 41.87625, 41.66777777777778, 42.001, 42.27363636363636, 42.08416666666667, 42.308461538461536, 42.50071428571429, 42.66733333333333, 42.500625, 42.64764705882353, 42.778333333333336, 42.89526315789474]
+network_sizes = range(200, 2000, 100)
+
+

### 2020-11-03 kevinscripts/hoeffding_safety_slush.py

@@ -1,0 +1,30 @@
+from scipy.stats import hypergeom
+import matplotlib.pyplot as plt
+import math
+import mpmath as mpm
+import numpy as np
+from scipy.linalg import expm
+
+mpm.mp.dps = 30
+
+n = 100
+alpha = 32
+k = 40
+results_stronger = []
+results_weaker = []
+for delta in range(0, int(n/2)):
+    alpha_over_k = mpm.mpf(alpha)/mpm.mpf(k)
+    delta_over_n = mpm.mpf(delta)/mpm.mpf(n)
+    # Get stronger result first
+    p = mpm.mpf((int(n/2) - delta))/n
+    t = mpm.mpf(alpha/k) - mpm.mpf(0.5) + mpm.mpf(delta/n)
+    print(t)
+    results_stronger.append(mpm.power(mpm.power(p/(p+t), p+t)*mpm.power((1-p)/(1-p-t), 1-p-t), k))
+    # Get weaker results
+    results_weaker.append(mpm.power(mpm.e, -2*mpm.power(alpha_over_k - mpm.mpf(0.5) + delta_over_n, 2)*k))
+
+print(results_stronger)
+plt.semilogy(range(0, int(n/2)), results_stronger, label="strong")
+plt.semilogy(range(0, int(n/2)), results_weaker, label="weak")
+plt.legend()
+plt.show()

### 2020-11-03 kevinscripts/plot.py

@@ -1,0 +1,68 @@
+import numpy as np
+from multiprocessing import Pool
+import math
+from sympy import binomial, zeros
+import mpmath as mp
+mp.mp.dps = 20
+import matplotlib.pyplot as plt
+from matplotlib.colors import LogNorm
+plt.style.use('seaborn-white')
+import numpy as np
+import pickle
+
+def hyper(N, suc, sample, x):
+    total = mp.mp.mpf(0)
+    bottom = binomial(N, sample)
+    for i in range(x, sample+1):
+        top = binomial(suc, i) * binomial(N-suc, sample-i)
+        total += (top/bottom)
+    return total
+
+def birthDeathProbability(n, k, alpha, c, b):
+    P = zeros(c/2, c/2)
+    for i in range(1, int(0.5*c)-1):
+        print(i)
+        birth = ((0.5*c-i)/c)*hyper(n, 0.5*c + i, k, alpha)
+        death = ((0.5*c+i)/c)*hyper(n, 0.5*c - i + b, k, alpha)
+        P[i, i-1] = death
+        P[i, i+1] = birth
+        P[i, i] = mp.mp.mpf(1.0) - birth - death
+    # Half blue half red
+    # Choose blue, get red
+    P[0, 1] = (0.5)*hyper(n, c/2, k, alpha)
+    # Choose blue, get nothing
+    P[0, 0] = mp.mp.mpf(1.0) - (0.5)*hyper(n, c/2, k, alpha)
+    # No blue, all red
+    # Choose red, get blue
+    P[int(0.5*c)-1, int(0.5*c)-2] = hyper(n, b, k, alpha)
+    # Choose red, get nothing
+    P[int(0.5*c)-1, int(0.5*c)-1] = mp.mp.mpf(1.0) - hyper(n, b, k, alpha)
+    powerraise = 50000
+    P = P**powerraise
+    with open("./P_{}_{}_{}_{}_{}_{}.out".format(n,k,alpha,c,b,powerraise), 'wb+') as f:
+        pickle.dump(P, f)
+    # plt.plot(range(0, int(c/2)), P.col(20))
+    P = np.array(P).astype(np.float64)    
+
+    x = range(0, int(c/2))
+    y = range(0, int(c/2))
+
+    X, Y = np.meshgrid(x, y)
+    Z = P
+
[…]

### 2020-11-03 kevinscripts/plot_k_needed_for_same_security_over_increasing_network.py

@@ -1,0 +1,13 @@
+import matplotlib.pyplot as plt
+import numpy as np
+from scipy.stats import hypergeom
+
+def hyper(N, suc, sample, x):                                                                                                   
+    rv = hypergeom(N, suc, sample)                                                                                              
+    return 1.0 - rv.cdf(x) + rv.pmf(x)
+
+plt.semilogx([10**i for i in range(2, 10)], [hyper(10**i, int((10**i)*0.8), 5, 4)/hyper(100, 80, 5, 4) for i in range(2, 10)])
+plt.xlabel("Network size")
+plt.ylabel("Ratio")
+plt.title("Ratios of the probabilities of successes over various network sizes, with fixed k = 10, alpha = 0.8, and x = 0.8c.")
+plt.show()

### 2020-11-03 kevinscripts/plots/beta1.py

@@ -1,0 +1,7 @@
+import matplotlib.pyplot as plt
+
+k = [10, 20, 30, 40]
+beta = [10.87625, 10.50125, 10.37625, 10.25125]
+
+plt.plot(k, beta)
+plt.show()

### 2020-11-03 kevinscripts/plots/plot_various_k_phase_shift.py

@@ -1,0 +1,63 @@
+import matplotlib.pyplot as plt
+import numpy as np
+from sympy import Symbol, Rational, binomial, expand_func
+import mpmath as mp
+import math
+import matplotlib as mpl
+from multiprocessing import Pool
+# mpl.style.use("seaborn-dark-palette")
+
+params = {'legend.fontsize': 40,
+          'figure.figsize': (15, 5),
+         'axes.labelsize': 60,
+         'axes.titlesize': 60,
+         'xtick.labelsize': 60,
+         'ytick.labelsize': 60}
+plt.rcParams.update(params)
+
+mp.mp.dps = 20
+
+def hyper(N, suc, sample, x):
+    total = mp.mp.mpf(0)
+    bottom = binomial(N, sample)
+    for i in range(x, sample+1):
+        top = binomial(suc, i) * binomial(N-suc, sample-i)
+        total += (top/bottom)
+    return total
+
+def test3Helper(t):
+    n, k, alpha, c, b = t
+    print(n, k, alpha, c, b)
+    for i in range(0, 205):
+        # print(i, ((0.5*c + i)*hyper(n, 0.5*c + i - 1, k, alpha)), ((0.5*c-i)*hyper(n, 0.5*c - i + b - 1, k, alpha)))
+        if (((0.5*c-i)/c)*hyper(n, 0.5*c + i, k, alpha)) >= (((0.5*c+i)/c)*hyper(n, 0.5*c - i + b, k, alpha)):
+            print(i)
+            return i
+
+def test3( _range, lab):
+    r = []    
+    with Pool(4) as p:
+        r = p.map(test3Helper, _range)
+    plt.plot([i[1] for i in _range], r, label=r"{}".format(lab))
+
+def main():
+    print(mpl.style.available)
+    plt.rc('text', usetex=True)
+    plt.rc('font', family='serif')
+    step = 2
+    _range = [(1000, i, i, 800, 200) for i in range(1, 100+step, step)]
+    test3(_range, "\alpha = 1")
+    _range = [(1000, i, math.ceil(0.8*i), 800, 200) for i in range(1, 100+step, step)]
+    test3(_range, "\alpha = 0.8")
+    _range = [(1000, i, math.ceil(0.6*i), 800, 200) for i in range(1, 100+step, step)]
+    test3(_range, "\alpha = 0.6")
+
+    plt.xlabel(r'k', fontsize=60)
[…]

### 2020-11-03 kevinscripts/plots/plot_various_k_phase_shift_snowball.py

@@ -1,0 +1,63 @@
+import matplotlib.pyplot as plt
+import numpy as np
+from sympy import Symbol, Rational, binomial, expand_func
+import mpmath as mp
+import math
+import matplotlib as mpl
+from multiprocessing import Pool
+# mpl.style.use("seaborn-dark-palette")
+
+params = {'legend.fontsize': 40,
+          'figure.figsize': (15, 5),
+         'axes.labelsize': 60,
+         'axes.titlesize': 60,
+         'xtick.labelsize': 60,
+         'ytick.labelsize': 60}
+plt.rcParams.update(params)
+
+mp.mp.dps = 20
+
+def hyper(N, suc, sample, x):
+    total = mp.mp.mpf(0)
+    bottom = binomial(N, sample)
+    for i in range(x, sample+1):
+        top = binomial(suc, i) * binomial(N-suc, sample-i)
+        total += (top/bottom)
+    return total
+
+def test3Helper(t):
+    n, k, alpha, c, b = t
+    print(n, k, alpha, c, b)
+    for i in range(0, 205):
+        # print(i, ((0.5*c + i)*hyper(n, 0.5*c + i - 1, k, alpha)), ((0.5*c-i)*hyper(n, 0.5*c - i + b - 1, k, alpha)))
+        if (((0.5*c+i)/c)*hyper(n, 0.5*c + i, k, alpha)) >= (((0.5*c-i)/c)*hyper(n, 0.5*c - i + b, k, alpha)):
+            print(i)
+            return i
+
+def test3( _range, lab):
+    r = []    
+    with Pool(4) as p:
+        r = p.map(test3Helper, _range)
+    plt.plot([i[1] for i in _range], r, label=r"{}".format(lab))
+
+def main():
+    print(mpl.style.available)
+    plt.rc('text', usetex=True)
+    plt.rc('font', family='serif')
+    step = 2
+    # _range = [(1000, i, i, 800, 200) for i in range(1, 100+step, step)]
+    # test3(_range, "\alpha = 1")
+    _range = [(1000, i, math.ceil(0.8*i), 800, 200) for i in range(1, 100+step, step)]
+    test3(_range, "\alpha = 0.8")
+    # _range = [(1000, i, math.ceil(0.6*i), 800, 200) for i in range(1, 100+step, step)]
+    # test3(_range, "\alpha = 0.6")
+
+    plt.xlabel(r'k', fontsize=60)
[…]

### 2020-11-03 kevinscripts/plots/security_over_various_k_theta_0_bar_consecutive_predicate.py

@@ -1,0 +1,89 @@
+import matplotlib.pyplot as plt
+
+font=30
+params = {'legend.fontsize': 40,
+          'figure.figsize': (15, 5),
+         'axes.labelsize': font,
+         'axes.titlesize': font,
+         'xtick.labelsize': font,
+         'ytick.labelsize': font}
+plt.rcParams.update(params)
+plt.rc('text', usetex=True)
+plt.rc('font', family='serif')
+
+# n = 100, b = 20, c = 80
+
+# k = 5, alpha = 4
+# Total trials per node = 10**15
+# Total trials = 10**15 * c
+# Consecutive = 200, to achieve red = 60, blue = 20 (maximal successes 80, since 60 + 20 byz = 80)
+# phase shift = 17 (i.e. c/2 + 17 = 57)
+# Security => 7.1e-9
+
+# k = 10, alpha = 8
+# Total trials per node = 10**15
+# Total trials = 10**15 * c
+# Consecutive = 150, to achieve red = 60, blue = 20 (maximal successes 80, since 60 + 20 byz = 80)
+# phase shift = 13 (i.e. c/2 + 13 = 53)
+# Security => 8.58e-35
+
+# k = 20, alpha = 16
+# Total trials per node = 10**15
+# Total trials = 10**15 * c
+# Consecutive = 120, to achieve red = 60, blue = 20 (maximal successes 80, since 60 + 20 byz = 80)
+# phase shift = 11 (i.e. c/2 + 11 = 51)
+# Security => 1.11e-99
+
+# k = 40, alpha = 32
+# Total trials per node = 10**15
+# Total trials = 10**15 * c
+# Consecutive = 110, to achieve red = 60, blue = 20 (maximal successes 80, since 60 + 20 byz = 80)
+# phase shift = 11 (i.e. c/2 + 11 = 51)
+# Security => 0
+
+# plt.subplot(2,1,1)
+# plt.semilogy([5, 10, 20, 40], [7.1e-9, 8.85e-35, 1.11e-99, 1.0e-300])
+# #plt.title("Varying k")
+# plt.ylabel(r"$p_f$")
+# plt.subplot(2,1,2)
+# plt.plot([5, 10, 20, 40], [200, 150, 120, 110])
+# plt.xlabel(r"$k$")
+# plt.ylabel(r"$l$")
+# plt.show()
+
+
+# n = 1000, b = 200, c = 800
+
+# k = 10, alpha = 8
+# Total trials per node = 10**15
[…]

### 2020-11-03 kevinscripts/plots/varying_k_theta_0.py

@@ -1,0 +1,39 @@
+import matplotlib.pyplot as plt
+import numpy as np
+from sympy import Symbol, Rational, binomial, expand_func
+import mpmath as mp
+import math
+import matplotlib as mpl
+# mpl.style.use("seaborn-dark-palette")
+
+params = {'legend.fontsize': 40,
+          'figure.figsize': (15, 5),
+         'axes.labelsize': 60,
+         'axes.titlesize': 60,
+         'xtick.labelsize': 60,
+         'ytick.labelsize': 60}
+plt.rcParams.update(params)
+
+mp.mp.dps = 20
+
+def hyper(N, suc, sample, x):
+    total = mp.mp.mpf(0)
+    bottom = binomial(N, sample)
+    for i in range(x, sample+1):
+        top = binomial(suc, i) * binomial(N-suc, sample-i)
+        total += (top/bottom)
+    return total
+
+print(mpl.style.available)
+plt.rc('text', usetex=True)
+plt.rc('font', family='serif')
+for k in [1, 5, 10, 20, 40, 80, 160]:
+    print(k)
+    alpha = math.ceil(0.8*k)
+    plt.semilogy(range(400, 800), [hyper(1000, i, k, alpha) for i in range(400, 800)], label=r"k = {}".format(k))
+
+plt.xlabel(r's_i', fontsize=60)
+plt.ylabel(r'p(s_i)',fontsize=60)
+plt.legend()
+
+plt.show()

### 2020-11-03 kevinscripts/simulator_slush.py

@@ -1,0 +1,36 @@
+import random
+import numpy as np
+
+def slush(sizen):
+    k = 10
+    alpha = 8
+    iters = 100
+    trials = 0
+    numbers = []
+    for trial in range(iters):
+        n = int(sizen/2)*[0] + int(sizen/2)*[1]    
+        i = 1
+        while True:
+            rcount = n.count(0)
+            if rcount == sizen or rcount == 0:
+                # print(i)
+                trials +=i
+                numbers.append(i/sizen)
+                break
+            u = random.randint(0, sizen-1)
+            sample = [n[j] for j in random.sample(range(0, sizen), k)]
+            if sample.count(0) >= alpha:
+                n[u] = 0
+            elif sample.count(1) >= alpha:
+                n[u] = 1
+            i += 1
+    print(sizen, (trials/iters)/sizen)
+    print(np.std(numbers))
+    print('-------------------------')
+
+def main():
+    for nsize in [600, 1200, 2400, 4800, 9600]:
+        slush(nsize)
+
+if __name__ == '__main__':
+    main()

### 2020-11-03 kevinscripts/snowflake_complete_network_ctmc_absorption_prob_extended_reversibility_probs.py

@@ -1,0 +1,130 @@
+from scipy.stats import hypergeom
+import matplotlib.pyplot as plt
+import math
+import mpmath as mpm
+from multiprocessing import Pool
+
+mpm.mp.dps = 1000
+
+def hyper(pop, succ, sample, threshold):
+    rv = hypergeom(pop, succ, sample)
+    return mpm.mpf(1.0) - rv.cdf(threshold) + rv.pmf(threshold)
+
+def helper(_input):
+    f, N, k, alpha, beta = _input
+    correct = N - f
+    births = []
+    deaths = []
+
+    '''
+    State space = {alpha-1 blue,  alpha blue,  alpha+1 blue, ...,  N-alpha+1 blue}
+                = {N-alpha+1 red, N-alpha red, N-alpha-1 red, ..., alpha-1 red}
+    '''
+
+    state_space = list(range(alpha-1, correct+1))
+
+    for state in state_space:
+        # Generate absorption states
+        if state == alpha-1:
+            births.append(0)
+            deaths.append(0)
+            continue
+        if state == correct:
+            births.append(0)
+            deaths.append(0)
+            continue
+        # Generate transient states
+        num_blue = state
+        num_red = correct-state
+        expected_blue_commit = num_blue * (hyper(N, num_blue, k, alpha)**beta)
+        expected_red_commit = num_red * (hyper(N, num_red + f, k, alpha)**beta)
+        # print(expected_blue_commit, expected_red_commit)
+        # Birth rate (more blue) = Select red, sample blue majority
+        sample_blue = hyper(N, num_blue, k, alpha)
+        births.append((num_red-expected_red_commit)*sample_blue)
+        # Death rate (less blue) = Select blue, sample red majority
+        sample_red = hyper(N, num_red + f, k, alpha)
+        deaths.append((num_blue-expected_blue_commit)*sample_red)
+        # deaths.append((num_blue-1)*sample_red)
+
+    # print(state_space)
+    normalized_state_space = list(range(0, len(state_space)))
+    # print(normalized_state_space)
+    # print(births)
+    # print(deaths)
+    # exit()
+
+    M = max(normalized_state_space)
+    print(M)
+    results_probs = []
+    for s in range(M):
+        print(s)
+        # Calculate absorption into all-red state from state s
+
+        # Calculate numerator
+        numerator_total = mpm.mpf(0.0)
+
[…]

### 2020-11-03 kevinscripts/snowflake_complete_network_ctmc_tau_various_byz.py

@@ -1,0 +1,137 @@
+'''
+Double absorbing states Snowflake (CTMC), 
+mean-time to absorption to either 0 or N
+using the matrix method (Tan method)
+'''
+
+from scipy.stats import hypergeom
+import matplotlib.pyplot as plt
+import math
+import mpmath as mpm
+import numpy as np
+# np.set_printoptions(threshold=np.nan)
+from scipy.linalg import expm
+from multiprocessing import Pool
+
+mpm.mp.dps = 30
+
+def hyper(pop, succ, sample, threshold):
+    rv = hypergeom(pop, succ, sample)
+    return mpm.mpf(1.0) - rv.cdf(threshold) + rv.pmf(threshold)
+
+def print_matrix(matrix):
+    print(np.array2string(matrix, max_line_width=np.inf))
+
+def helper(tup):
+    f, N, k, alpha, beta = tup
+    # Now plot the prob of commit under bounded rounds (in this case, 
+    # rounds is about c*log(N) where c is some constant)
+    results = []
+    print(f, N//2)
+    correct = N - f
+
+    births = []
+    deaths = []
+
+    '''
+    State space = {alpha-1 blue,  alpha blue,  alpha+1 blue, ...,  N-alpha+1 blue}
+                = {N-alpha+1 red, N-alpha red, N-alpha-1 red, ..., alpha-1 red}
+    '''
+
+    state_space = list(range(alpha-1, correct-alpha+2))
+
+    for state in state_space:
+        # Generate absorption states
+        if state == alpha-1:
+            births.append(0)
+            deaths.append(0)
+            continue
+        # Generate transient states
+        num_blue = state
+        num_red = correct-state
+        expected_blue_commit = num_blue * (hyper(N, num_blue, k, alpha)**beta)
+        expected_red_commit = num_red * (hyper(N, num_red + f, k, alpha)**beta)
+        if state < correct//2:
+            # Birth rate (more blue) = Select red, sample blue majority
+            sample_blue = hyper(N, num_blue+f, k, alpha)
+            assert num_red >= expected_red_commit
+            assert num_blue >= expected_blue_commit
+            births.append((num_red-expected_red_commit)*sample_blue)
+            # Death rate (less blue) = Select blue, sample red majority
+            sample_red = hyper(N, num_red, k, alpha)
+            deaths.append((num_blue-expected_blue_commit)*sample_red)
+        elif state > correct//2:
[…]

### 2020-11-03 kevinscripts/test_colors_in_bins.py

@@ -1,0 +1,107 @@
+'''
+In this test, we initialize a bin with RED and BLUE balls, initially zero. 
+Then, for every step, we check the count, and if there's more RED balls than 
+BLUE balls we add a BLUE ball with some probability p, we add a red with probability q, 
+and add nothing with probability 1 - p - q (vice versa). It must be that p > q
+'''
+
+import random
+import matplotlib.pyplot as plt
+
+def varying_rounds():
+    for pp in [25]:
+        averages = []
+        distances = []
+        rounds = range(50, 1100, 50)
+        for r in rounds:
+            trialsaverages = []
+            trialsdistances = []
+            p = pp
+            q = 50 - p
+            npq = 100 - p - q
+            print(p, q, npq)
+            for trial in range(1000):
+                BINS = [0, 0]
+                for _r in range(r):
+                    if BINS[0] > BINS[1]:
+                        index = random.randint(1, 100)
+                        if 0 < index <= p:
+                            BINS[1] += 1
+                        elif p < index <= p + q:
+                            BINS[0] += 1
+                        else:
+                            continue
+                    elif BINS[1] > BINS[0]:
+                        index = random.randint(1, 100)
+                        if 0 < index <= p:
+                            BINS[0] += 1
+                        elif p < index <= p + q:
+                            BINS[1] += 1
+                        else:
+                            continue
+                    else:
+                        index = random.randint(1, 100)
+                        if index <= 50:
+                            BINS[0] += 1
+                        else:
+                            BINS[1] += 1
+                trialsaverages.append(abs(BINS[0]+BINS[1])/2)
+                trialsdistances.append(abs(BINS[0]-BINS[1]))
+            averages.append(sum(trialsaverages)/len(trialsaverages))
+            distances.append(sum(trialsdistances)/len(trialsdistances))
+        print(distances)
+        plt.plot(rounds, distances, label="{}".format(p))
+        # plt.plot(rounds, averages)
+    plt.legend()
+    plt.show()
+
+def varying_p():
+    averages = []
+    distances = []
+    pp = range(25, 51)
+    for p in pp:
+        trialsaverages = []
+        trialsdistances = []
+        q = 50 - p
+        npq = 100 - p - q
+        print(p, q, npq)
+        for trial in range(50):
+            BINS = [0, 0]
+            for round in range(1000):
+                if BINS[0] > BINS[1]:
+                    index = random.randint(1, 100)
[…]

### 2020-11-03 kevinscripts/test_colors_in_bins_dynamic.py

@@ -1,0 +1,107 @@
+'''
+In this test, we initialize a bin with RED and BLUE balls, initially zero. 
+Then, for every step, we check the count, and if there's more RED balls than 
+BLUE balls we add a BLUE ball with some probability p, we add a red with probability q, 
+and add nothing with probability 1 - p - q (vice versa). It must be that p > q.
+'''
+
+import random
+import matplotlib.pyplot as plt
+
+def varying_rounds():
+    for pp in [25]:
+        averages = []
+        distances = []
+        rounds = range(50, 1100, 50)
+        for r in rounds:
+            trialsaverages = []
+            trialsdistances = []
+            p = pp
+            q = 50 - p
+            npq = 100 - p - q
+            print(p, q, npq)
+            for trial in range(1000):
+                BINS = [0, 0]
+                for _r in range(r):
+                    if BINS[0] > BINS[1]:
+                        index = random.randint(1, 100)
+                        if 0 < index <= p:
+                            BINS[1] += 1
+                        elif p < index <= p + q:
+                            BINS[0] += 1
+                        else:
+                            continue
+                    elif BINS[1] > BINS[0]:
+                        index = random.randint(1, 100)
+                        if 0 < index <= p:
+                            BINS[0] += 1
+                        elif p < index <= p + q:
+                            BINS[1] += 1
+                        else:
+                            continue
+                    else:
+                        index = random.randint(1, 100)
+                        if index <= 50:
+                            BINS[0] += 1
+                        else:
+                            BINS[1] += 1
+                trialsaverages.append(abs(BINS[0]+BINS[1])/2)
+                trialsdistances.append(abs(BINS[0]-BINS[1]))
+            averages.append(sum(trialsaverages)/len(trialsaverages))
+            distances.append(sum(trialsdistances)/len(trialsdistances))
+        print(distances)
+        plt.plot(rounds, distances, label="{}".format(p))
+        # plt.plot(rounds, averages)
+    plt.legend()
+    plt.show()
+
+def varying_p():
+    averages = []
+    distances = []
+    pp = range(25, 51)
+    for p in pp:
+        trialsaverages = []
+        trialsdistances = []
+        q = 50 - p
+        npq = 100 - p - q
+        print(p, q, npq)
+        for trial in range(50):
+            BINS = [0, 0]
+            for round in range(1000):
+                if BINS[0] > BINS[1]:
+                    index = random.randint(1, 100)
[…]

### 2020-11-03 kevinscripts/test_optimal_byzantine.py

@@ -1,0 +1,80 @@
+import random
+import math
+import itertools
+from scipy.stats import hypergeom
+
+def hyper(N, S, k, a):
+    rv = hypergeom(N, S, k)
+    return 1.0 - rv.cdf(a) + rv.pmf(a)
+
+def main():
+    _range = 10
+    print("RANGE:", _range)
+    outputs = []
+    for seq in itertools.product("01", repeat=_range):
+        trials = []
+        distances = []
+        _seq = [int(_s) for _s in seq]
+        for trial in range(20):
+            correct = [-1]*3 + [1]*3
+            byzantine = [-1]*3
+            k = 5
+            a = 3.75
+            choices = [0]*len(correct)
+            indeces = []
+            # _range = int(3*len(correct)*math.log(len(correct), math.e))
+            for i in range(_range):
+                # Set byzantine
+                if seq[i] == "0":
+                    byzantine = [-1]*3
+                elif seq[i] == "1":
+                    byzantine = [1]*3
+                # Select random correct node
+                index = random.randint(0, len(correct)-1)
+                if correct[index] < 0:
+                    indeces.append(0)
+                elif correct[index] > 0:
+                    indeces.append(1)
+                choices[index] += 1
+                correct_copy = correct[:]
+                del correct_copy[index]
+                sample = random.sample(correct_copy+byzantine, k)
+                neg_count = len([i for i in sample if i < 0])
+                pos_count = len([i for i in sample if i > 0])
+                if neg_count >= a:
+                    # print(sample, neg_count, neg_count/len(sample))
+                    if correct[index] < 0:
+                        correct[index] += -1
+                    elif correct[index] > 0:
+                        correct[index] = -1
+                    else:
+                        print("ERROR1")
+                        exit(-1)
+                elif pos_count >= a:
+                    # print(sample, pos_count, pos_count/len(sample))
+                    if correct[index] < 0:
+                        correct[index] = 1
+                    elif correct[index] > 0:
+                        correct[index] += 1
+                    else:
+                        print("ERROR2")
+                        exit(-1)
+                elif pos_count < a and neg_count < a:
+                    continue
+                else:
+                    print("ERROR3")
+                    exit(-1)
+            distance = math.sqrt(sum([(indeces[j]-_seq[j])**2 for j in range(_range)]))
+            trials.append(correct[:])
+            distances.append(distance)
[…]

### 2020-11-03 kevinscripts/test_optimal_byzantine_expectation.py

@@ -1,0 +1,62 @@
+import random
+import math
+import itertools
+from scipy.stats import hypergeom
+
+def hyper(N, S, k, a):
+    rv = hypergeom(N, S, k)
+    return 1.0 - rv.cdf(a) + rv.pmf(a)
+
+def main():
+    _range = 10
+    print("RANGE:", _range)
+    outputs = []
+    for seq in itertools.product("01", repeat=_range):
+        correct = [-1]*3 + [1]*3
+        byzantine = [-1]*3
+        k = 5
+        a = 3.75
+        choices = [0]*len(correct)
+        indeces = []
+        # _range = int(3*len(correct)*math.log(len(correct), math.e))
+        for i in range(_range):
+            # Set byzantine
+            if seq[i] == "0":
+                byzantine = [-1]*3
+            elif seq[i] == "1":
+                byzantine = [1]*3
+            # Select random correct node
+            index = random.randint(0, len(correct)-1)
+            indeces.append(index)
+            choices[index] += 1
+            correct_copy = correct[:]
+            del correct_copy[index]
+            sample = random.sample(correct_copy+byzantine, k)
+            neg_count = len([i for i in sample if i < 0])
+            pos_count = len([i for i in sample if i > 0])
+            if neg_count >= a:
+                # print(sample, neg_count, neg_count/len(sample))
+                if correct[index] < 0:
+                    correct[index] += -1
+                elif correct[index] > 0:
+                    correct[index] = -1
+                else:
+                    print("ERROR1")
+                    exit(-1)
+            elif pos_count >= a:
+                # print(sample, pos_count, pos_count/len(sample))
+                if correct[index] < 0:
+                    correct[index] = 1
+                elif correct[index] > 0:
+                    correct[index] += 1
+                else:
+                    print("ERROR2")
+                    exit(-1)
+            elif pos_count < a and neg_count < a:
+                continue
+            else:
+                print("ERROR3")
+                exit(-1)
+
+if __name__ == '__main__':
+    main()

### 2020-11-03 kevinscripts/test_optimal_byzantine_sanity.py

@@ -1,0 +1,155 @@
+import random
+import math
+import itertools
+from scipy.stats import hypergeom
+
+class CorrectNode(object):
+    def __init__(self, color):
+        self.color = color
+        if self.color = 0:
+            self.red_confidence = 1
+            self.blue_confidence = 0
+        else:
+            self.blue_confidence = 1
+            self.red_confidence = 0
+
+    def get_color(self):
+        return self.color                
+        
+
+def get_color(l):
+    if abs(l[0]) > abs(l[1]):
+        return -1
+    elif abs(l[1] > abs(l[0])):
+        return 1
+    else:
+        c = random.randint(0,1)
+        if c:
+            return 1
+        else:
+            return -1
+
+def main():
+    _range = 7
+    outputs = []
+    num_correct_each_color = 20
+    num_byz = 15
+    num_trials = 20
+    k = 5
+    a = 3.5
+    outputs = []
+    skipper = 0
+    for byz_sequence in itertools.product("01", repeat=_range):
+        for nature_sequence in itertools.product("01", repeat=_range):
+            skipper += 1
+            if skipper % 2 == 0:
+                continue
+            print("".join(byz_sequence), "".join(nature_sequence))
+            trials = []
+            for trial in range(num_trials):
+                correct = [[-1,0]]*num_correct_each_color + [[0,1]]*num_correct_each_color
+                byzantine = [-1]*num_byz            
+                choices = [0]*len(correct)
+                # Iterate over time
+                for i in range(_range):
+                    # print("--------------------------")
+                    # print(byz_sequence)
+                    # print(nature_sequence)
+                    # print("TIME:", i, correct)
+                    # Set byzantine
+                    if byz_sequence[i] == "0":
+                        byzantine = [[-1,0]]*num_byz
+                    elif byz_sequence[i] == "1":
+                        byzantine = [[0,1]]*num_byz
+                    # Correct node color
+                    corr_color = nature_sequence[i]
+                    corr_color_index = None
+                    if corr_color == "0":
+                        # Red node was chosen, go ahead and find all red nodes 
+                        # and select one of them uniformly at random
+                        _indeces_red = []
+                        for _i in range(len(correct)):
+                            if get_color(correct[_i]) == -1:
+                                _indeces_red.append(_i)
+                        if _indeces_red:
+                            # At least one red color node exists
+                            corr_color_index = random.choice(_indeces_red)
+                        else:
[…]

### 2020-11-03 obsolete/plots/minimal_k_varying_b.py

@@ -1,0 +1,53 @@
+import matplotlib.pyplot as plt
+
+font=30
+params = {'legend.fontsize': 40,
+          'figure.figsize': (15, 5),
+         'axes.labelsize': font,
+         'axes.titlesize': font,
+         'xtick.labelsize': font,
+         'ytick.labelsize': font}
+plt.rcParams.update(params)
+plt.rc('text', usetex=True)
+plt.rc('font', family='serif')
+
+# Consecutive predicate used
+# n = 100, b = 0, c = 100
+# Total trials per node = 10**15
+# Total trials = 10**15 * c
+# k = 5, alpha = 4
+
+# Consecutive = 200, to achieve red = 60, blue = 20 (maximal successes 80, since 60 + 20 byz = 80)
+# phase shift = 17 (i.e. c/2 + 17 = 57)
+# Security => 7.1e-9
+
+# k = 10, alpha = 8
+# Total trials per node = 10**15
+# Total trials = 10**15 * c
+# Consecutive = 150, to achieve red = 60, blue = 20 (maximal successes 80, since 60 + 20 byz = 80)
+# phase shift = 13 (i.e. c/2 + 13 = 53)
+# Security => 8.58e-35
+
+# k = 20, alpha = 16
+# Total trials per node = 10**15
+# Total trials = 10**15 * c
+# Consecutive = 120, to achieve red = 60, blue = 20 (maximal successes 80, since 60 + 20 byz = 80)
+# phase shift = 11 (i.e. c/2 + 11 = 51)
+# Security => 1.11e-99
+
+# k = 40, alpha = 32
+# Total trials per node = 10**15
+# Total trials = 10**15 * c
+# Consecutive = 110, to achieve red = 60, blue = 20 (maximal successes 80, since 60 + 20 byz = 80)
+# phase shift = 11 (i.e. c/2 + 11 = 51)
+# Security => 0
+
+plt.subplot(2,1,1)
+plt.semilogy([5, 10, 20, 40], [7.1e-9, 8.85e-35, 1.11e-99, 1.0e-300])
+#plt.title("Varying k")
+plt.ylabel(r"$p_f$")
+plt.subplot(2,1,2)
+plt.plot([5, 10, 20, 40], [200, 150, 120, 110])
+plt.xlabel(r"$k$")
+plt.ylabel(r"$l$")
+plt.show()

### 2020-11-03 obsolete/plots/trend_divergence.py

@@ -1,0 +1,33 @@
+import matplotlib.pyplot as plt
+import numpy as np
+from multiprocessing import Pool
+from sympy import binomial, zeros, pprint
+import mpmath as mp
+import math
+mp.mp.dps = 20
+
+def hyper(N, suc, sample, x):
+    total = mp.mp.mpf(0)
+    bottom = binomial(N, sample)
+    for i in range(x, sample+1):
+        top = binomial(suc, i) * binomial(N-suc, sample-i)
+        total += (top/bottom)
+    return total
+
+def main():
+    n = 1000
+    b = 200
+    k = 10
+    alpha = 8
+    _range = range(0, 1001)
+    plot1 = [hyper(1000, i, k, alpha) for i in _range]
+    plot2 = [hyper(1000, 1000-i, k, alpha) for i in _range]
+    plt.semilogy(_range, plot1, color="red", label="Probability of \mathtt{R} successful sample")
+    plt.semilogy(_range, plot2, color="blue", label="Probability of \mathtt{B} successful sample")
+    plt.xlabel("Number of \mathtt{R}")
+    plt.ylabel("Probability of a success")
+    plt.legend()
+    plt.show()
+
+if __name__ == '__main__':
+    main()

### 2020-11-03 sections/evaluation_fr.tex
add French translation of the evaluation section

\newcommand{\sysname}{Avalanche}

\subsection{Configuration de test}

Nous conduisons nos expériences sur Amazon EC2 en faisant tourner une plage allant d'une
centaine (125) à des milliers (2000) d'instances de machines virtuelles. Nous  utilisons
des instances \texttt{c5.large}, chacune simulant un noeud individuel. AWS fournit une bande passante
allant jusqu'à 2 Gigabits par seconde, sachant que le protocole {\sysname} utilise au maximum 100 Megabits
par seconde environ.

Notre implémentation supporte deux types de transactions: la première est le format d'UTXO
customisé, tandis que l'autre repose directement sur le code de Bitcoin 0.16. Ces deux formats
supportés utilisent la librairie de cryptographie secp256k1 issue du projet Bitcoin et exposent
le même format d'adresse pour les portefeuilles. Toutes nos expériences utilisent le format
customisé à l'exception de la géo-réplication, pour laquelle les résultats sont donnés pour
les deux types de transactions.

Nous simulons un flux continu de nouvelles transactions venant des utilisateurs en créant des
processus client séparés, chacun de ces processus gérant un portefeuille différent, qui génèrent
des transactions à destination de nouvelles adresses de réception puis envoient les requêtes vers
les noeuds {\sysname}.

Nous utilisons plusieurs de ces processus client pour atteindre la limite de capacité de notre
système. Le nombre de récepteurs pour chaque transaction est paramétré pour obtenir une taille
de transaction moyenne d'environ 250 octets (1--2 entrées/sorties par transaction en moyenne et
une taille d'UTXO fixe), la taille de transaction moyenne actuelle sur Bitcoin. Afin d'utiliser
le réseau efficacement, nous regroupons jusqu'à 40 transactions en une requête, tout en gardant
des valeurs de confiance à un niveau de granularité équivalent à une seule transaction.

Tous les indicateurs rapportés ici reflètent les mesures de bout en bout prises depuis le point
de vue de tous les clients. Plus précisément, les clients examinent le nombre total de transactions
confirmées par seconde pour mesurer le débit global, et, pour chaque transaction, ils soustraient
l'horodatage d'initiation de la transaction à l'horodatage de confirmation de celle-ci pour en
mesurer la latence. Chaque expérience de mesure du débit est répétée 5 fois et l'écart type est
mentionné dans chaque graphique.

Pour ce qui est des paramètres de sécurité, nous prenons $k = 10$, $\alpha = 0.8$, $\beta_1 = 11$, $\beta_2 = 150$,
ce qui donne un MTTF de \textasciitilde{}$10^{24}$ années.

\subsection{Débit de transactions}

Nous mesurons d'abord le débit du système en le saturant avec des transactions tout en examinant
la fréquence à laquelle les transactions sont confirmées à l'état nominal. Pour cette expérience, nous lançons
d'abord {\sysname} sur 125 noeuds exécutant chacun 10 processus client qui génèrent un nombre de
400 transactions sortantes à tout moment.

Comme le montre le premier groupe de barres dans la Figure~\ref{fig:eval-thr}, le système atteint 6851
transactions par seconde (tps) pour une taille de batch de 20 et plus de 7002 tps pour une taille de
batch de 40. Notre système sature plus vite avec des petites tailles de batch en comparaison avec d'autres
blockchains dont les performances sont connues: Bitcoin regroupe des batchs de plusieurs milliers de
transactions par bloc, Algorand~\cite{GiladHMVZ17} gère des blocks de 2-10 MégaOctets, i.e., 8.4--41.9K
transactions par batch et Conflux~\cite{confluxLLXLC18} gère des blocs de 4 MégaOctets, i.e., 16.8K tx/batch.
Ces systèmes sont relativement lents pour prendre une simple décision, et demandent de surcroît une taille
de batch (et donc de bloc) importante pour obtenir de meilleures performances. Atteindre un haut niveau de débit
transactionnel avec une petite taille de batch implique une latence réduite, comme nous le démontrons plus loin.

\begin{figure}[h]
\includegraphics[width=\linewidth]{figures/thr-ava.pdf}
\captionof{figure}{Débit transactionnel en fonction de la taille du réseau. Chaque paire de barres est produite avec
une taille de batch de 20 et 40, de gauche à droite.}
\label{fig:eval-thr}
\end{figure}

\begin{figure}
\includegraphics[width=\linewidth]{figures/thr-raw.pdf}
\captionof{figure}{Débit transactionnel pour une taille de batch de 40, avec (gauche) et sans (droite) vérification
de la signature.}
\label{fig:eval-thr-raw}
\end{figure}

\subsection{Mise à l'échelle}

Afin d'observer la manière dont le système passe à l'échelle en terme de nombre de noeuds participant au consensus
{\sysname}, nous effectuons des expériences en gardant des paramètres identiques tout en faisant varier le
nombre de noeuds de 125 jusqu'à 2000.

La Figure~\ref{fig:eval-thr} montre que le débit global se dégrade d'environ $1.34\%$ à 6909 tps lorsque le
réseau augmente d'un facteur 16 pour atteindre $n = 2000$. Cette dégradation est mineure si on la compare au
taux d'augmentation du réseau. Notez que l'axe X est logarithmique.

Avalanche tire sa bonne mise à l'échelle de plusieurs sources: premièrement en maintenant un ordre partiel qui
capture uniquement les relations entre les dépenses ce qui permet une meilleure parallélisation qu'un journal
BFT répliqué classique qui traite toutes les transactions de manière linéaire; deuxièmement l'absence de leader
permet d'éviter les goulots d'étranglement; et finalement le nombre de messages que chaque noeud doit gérer par
décision est $O(k)$ et n'augmente pas au fur et à mesure que le réseau s'agrandit.

\subsection{Engorgement dû à la cryptographie}

Nous examinons ensuite où se situent les goulots d'étranglement dans notre implémentation.
La barre violette à la droite de chaque groupe dans la Figure~\ref{fig:eval-thr-raw} montre le débit
d'Avalanche lorsque la vérification des signatures est désactivée. Les débits s'améliorent d'environ 2.6x, par
par rapport à la barre bleue sur la gauche.
Cela révèle que le surplus de calcul dû à la vérification cryptographique est le principal goulot d'étranglement
de l'implémentation de notre système. Ce problème peut être mitigé en déléguant la vérification des transactions
à un GPU\@. Même en l'absence d'une telle optimisation, 7K tps dépasse de loin ce qui existe en matière de blockchain.

\subsection{Latence}

\begin{figure}
\includegraphics[width=\linewidth]{figures/lat.pdf}
\captionof{figure}{Distribution de la latence de transaction pour $n = 2000$. L'axe X représente la latence
    de transaction en secondes sur une échelle logarithmique, tandis que l'axe Y est la portion des transactions
    qui tombe dans la période de confirmation (normalisée à $1$). L'histogramme de toutes les latences de transaction
    pour un client est visible sur la gauche avec $100$ bins, et son CDF visible sur la droite.}
\label{fig:eval-lat1}
\end{figure}

La latence d'une transaction est le temps écoulé à partir du moment où elle est soumise au réseau jusqu'à ce qu'elle
soit confirmée et donc acceptée par le réseau.

La Figure~\ref{fig:eval-lat1} représente un histogramme montrant la distribution des latences en utilisant la
même configuration que pour les mesures de débits à partir de 2000 noeuds. L'axe X représente le temps en secondes
tandis que l'axe Y est la portion des transactions qui sont finalisées dans la période de temps correspondante.
Cette figure met également en avant la Fonction de Distribution Cumulée (CDF) en accumulant le nombre de
transactions finalisées sur la période.

Cette expérience démontre que la plupart des transactions sont confirmées en moins de 0.3 secondes environ.
La majorité des latences se trouve aux alentours des 206 ms avec une variance relativement faible, ce qui
indique que les noeuds convergent sur la décision finale en tant que groupe environ au même moment. La deuxième
ligne verticale montre la latence maximale observée, qui se situe aux alentours des 0.4 secondes.

La Figure~\ref{fig:eval-lat2} montre la latence de transaction pour un nombre variable de noeuds. Les arêtes
horizontales des boîtes représentent respectivement du bas vers le haut le minimum, le premier quartile, la médiane,
le troisième quartile, et le maximum de latence. En définitive, les données expérimentales montrent que la latence
médiane est plus ou moins indépendante de la taille du réseau.

\begin{figure}[h]
\includegraphics[width=\linewidth]{figures/lat2.pdf}
\captionof{figure}{Latence de transaction en fonction de la taille du réseau. ``b'' indique la taille de batch et
``raw'' est l'occurence sans la vérification des signatures.}
\label{fig:eval-lat2}
\end{figure}

\subsection{Clients malicieux}
\label{sec:evaluation-misbehaving}
\begin{figure}
\includegraphics[width=\linewidth]{figures/lat3.pdf}
\captionof{figure}{Latence en fonction du taux de transactions véreuses.}
\label{fig:eval-lat3}
\end{figure}

Nous évaluons ensuite comment des transactions véreuses générées par des clients malicieux qui effectuent des
double dépenses sur des sorties non dépensées peuvent affecter la latence des transactions légitimes crées par
des clients honnêtes. Nous adoptons une stratégie qui simule des clients malicieux où une fraction (de $0\%$ à $25\%$)
des transactions en attente sont en conflit avec d'autres transactions existantes.

Le processus client effectue ceci en désignant des flux de transactions contenant des double dépenses parmi toutes
les transactions simulées en attente, et envoie les transactions conflictuelles vers différents noeuds. Nous utilisons
pour cela les mêmes paramètres avec $n = 1000$ comme dans les expériences précédentes, et nous mesurons uniquement
le débit et la latence des transactions confirmées.

La latence d'{\sysname} est seulement légèrement influencée par les clients malicieux, comme le montre la
Figure~\ref{fig:eval-lat3}. Etonnement la latence maximale baisse légèrement lorsque le pourcentage de transactions
véreuses augmente. Ce comportement s'explique par le fait que lorsqu'on introduit des transactions véreuses, le débit
\emph{effectif} est réduit et soulage donc la charge du système.

Cette observation est confirmée par la Figure~\ref{fig:eval-thr2}, qui montre que le débit (de transactions légitimes)
decroît avec le taux de transactions véreuses. De plus, la baisse du débit apparaît proportionnelle au nombre de
clients malicieux ce qui démontre que les attaques potentielles ne bénéficient pas d'un effet de levier.

\begin{figure}
\includegraphics[width=\linewidth]{figures/thr-byz.pdf}
\captionof{figure}{Débit de transactions en fonction du taux de transactions véreuses.}
\label{fig:eval-thr2}
\end{figure}

\subsection{Géo-réplication}

\begin{figure}
\includegraphics[width=\linewidth]{figures/lat4.pdf}
\captionof{figure}{Histogramme de latence/CDF pour $n = 2000$ réparties dans 20 villes.}
\label{fig:eval-lat4}
\end{figure}

L'expérience suivante place le système dans l'émulation d'un scénario géo-répliqué, modélisé sur le même scénario
détaillé dans un travail préliminaire~\cite{GiladHMVZ17}.
Nous avons selectionné 20 villes majeures qui semblent être localisées à proximité d'un nombre conséquent de noeuds
Bitcoin joignables, d'après ~\cite{bitnodes2018}. Ces villes couvrent l'Amérique du nord, l'Europe, l'Asie occidentale,
l'Asie orientale, l'Océanie, et incluent également les 10 villes où l'on trouve le plus grand nombre de noeuds
joignables. Nous utilisons les données de latence obtenues depuis~\cite{wondernetworkping2018} et nous émulons
la latence des paquets réseau dans le noyau Linux en utilisant les commandes \texttt{tc} et \texttt{netem}.
2000 noeuds sont distribués équitablement parmi toutes les villes, sans émuler de latence supplémentaire entre les
noeuds d'une même ville. Comme pour l'évaluation d'Algorand, nous limitons aussi la bande passante par processus
à 20 Mégabits par seconde dans le but de simuler des paramètres à l'échelle d'internet d'un réseau comportant de
nombreux liens. Nous assignons ensuite un processus client à chaque ville, qui génère 400 transactions sortantes
par ville à tout moment.

Dans ce scénario, Avalanche atteint un débit de transactions moyen de 3401 tps, avec un écart type de 39 tps. Comme
le montre la Figure~\ref{fig:eval-lat4}, la latence de transaction médiane est de 1.35 secondes, avec une latence
maximale de 4.25 secondes. Nous supportons également le code initial de Bitcoin pour les transactions, qui dans ce
cas donne un débit de 3530 tps avec $\sigma = 92$ tps.

\subsection{Comparaison avec d'autres systèmes}

Même s'il existe une abondance de protocoles blockchain et de cryptomonnaies, la plupart d'entre eux présentent
une ébauche de leur protocole sans proposer d'implémentation pratique ou de résultats d'évaluation. En outre, parmi
ceux qui proposent effectivement des résultats, la plupart ne sont pas évalués dans des conditions réalistes ou à
grande échelle (des centaines de milliers de noeuds complets participant au consensus).

En conséquent, nous choisissons Algorand and Conflux pour notre étude comparative. Algorand, Conflux, et Avalanche
sont tous fondamentalement différents dans leur conception. L'algorithme de consensus à l'échelle d'un comité
d'Algorand repose sur un accord byzantin basé sur un quorum, et Conflux supplémente le consensus de Nakamoto d'une
structure type Graphe Acyclique Dirigé (DAG) pour permettre un meilleur débit, tandis qu'Avalanche appartient à
une nouvelle famille de protocoles basés sur la méta-stabilité. Par ailleurs, nous utilisons
Bitcoin~\cite{nakamoto2008bitcoin} comme base technique.

Les évaluations d'Algorand et Avalanche reposent toutes deux sur un réseau décisionnel de 2000 noeuds sur EC2.
Notre évaluation se fait sur des instances \texttt{c5.large} partagées, tandis qu'Algorand utilise des instances
\texttt{m4.2xlarge}\@. Ces deux plateformes sont très similaires à l'exception d'un léger avantage en terme de
fréquence CPU pour le \texttt{c5.large}, qui au final n'est que partiellement utilisé vu que nos processus ne
consomment que $30\%$ de temps processeur lors de ces expériences. Les paramètres de sécurité choisis pour nos
expériences garantissent une probabilité de violation de sécurité en dessous de $10^{-9}$ en présence de
$20\%$ de noeuds byzantins, en comparaison de l'évaluation Algorand qui garantit une probabilité de violation en dessous
de $5 \times 10^{-9}$ avec $20\%$ noeuds byzantins.

Ni Algorand ni Conflux ne prennent en compte le surplus de calcul lié à la vérification cryptographique dans leurs
études. Leurs évaluations utilisent des blocs qui contiennent des MégaOctets de données factices et présentent les
mesures de débit en MégaOctets ou GigaOctets par heure. Nous utilisons donc la taille moyenne d'une transaction
Bitcoin, 250 octets, pour en déduire leurs débits de transactions. Par contraste nos expériences transportent des
données de transactions réelles et prennent entièrement en compte le temps de calcul des vérifications cryptographiques.

Le débit de transactions est de 3-7 tps pour Bitcoin, 874 tps pour Algorand (avec des blocs de 10 MégaOctets), 3355 tps
pour Conflux (le papier fait mention d'un débit 3.84 fois supérieur au débit d'Algorand dans les mêmes conditions).

En comparaison, {\sysname} obtient plus de 3400 tps de manière consistante sur un réseau allant jusqu'à 2000 noeuds
sans aucun comité de décision ou preuve de travail. Pour ce qui est de la latence, une transaction est confirmée
après 10--60 minutes pour Bitcoin, environ 50 secondes pour Algorand, 7.6--13.8 minutes avec Conflux, et 1.35 secondes
pour {\sysname}.

Avalanche se comporte bien mieux qu'Algorand à la fois en terme de débit et de latence vu qu'Algorand utilise une
fonction aléatoire vérifiable pour élire ses comités, et maintient un journal ordonné pendant qu'{\sysname} établit
un ordre partiel uniquement. Bien que l'entité à la tête d'Algorand soit anonyme et change continuellement, elle reste
néanmoins basée sur un leader unique qui pourrait être un facteur limitant pour la mise à l'échelle, tandis
qu'{\sysname} n'est pas contrôlé par une entité unique.

Avalanche a un débit similaire à Conflux, mais sa latence est 337--613 fois meilleure. Conflux utilise également une
structure DAG pour amortir le coût du consensus et augmenter le débit de transactions, cependant il est toujours ancré
sur un consensus de Nakamoto (PoW) le rendant incapable de confirmer des transactions instantanément comme le fait
Avalanche.

Dans un système à base de blockchain, on ne peut généralement améliorer le débit des transactions qu'aux dépens de la
latence à travers le regroupement par batchs. Le vrai facteur limitant au niveau de la performance est le nombre de
décisions qu'un système est en mesure de prendre en une seconde, et ce paramètre est fondamentalement limité soit par
l'accord byzantin ($\mathrm{BA}^{*}$) chez Algorand soit par le consensus de Nakamoto pour Conflux.

### 2020-11-04 paper_fr.tex
add French translation of the evaluation section
paper.tex - Abstract, Introduction

@@
 \title{
-\Large\bf Scalable and Probabilistic Leaderless BFT Consensus through Metastability}
+%O \Large\bf Scalable and Probabilistic Leaderless BFT Consensus through Metastability}
+\Large\bf Consensus BFT sans dirigeant, extensible et probabiliste, par métastabilité}
 \author{%
@@
 \blfootnote{%
-\textsuperscript{*}\emph{Blasts off at the speed of light!} --- Team Rocket\\
-An earlier version of this paper published on May 16th 2018, IPFS, was titled \emph{Snowflake to Avalanche: A Novel Metastable Consensus Protocol Family for Cryptocurrencies}.
+%O \textsuperscript{*}\emph{Blasts off at the speed of light!} --- Team Rocket\\
+\textsuperscript{*}\emph{À la vitesse de la lumière !} --- Team Rocket\\
+%O An earlier version of this paper published on May 16th 2018, IPFS, was titled \emph{Snowflake to Avalanche: A Novel Metastable Consensus Protocol Family for Cryptocurrencies}.
+Une version antérieure de cet article, publié le 16 mai 2018, IPFS, était intitulée \emph{Snowflake to Avalanche: A Novel Metastable Consensus Protocol Family for Cryptocurrencies}.
 
@@
 \begin{abstract}
-This paper introduces a family of leaderless Byzantine fault tolerance protocols, built around a metastable mechanism via network subsampling.
-These protocols provide a strong probabilistic safety guarantee in the presence of Byzantine adversaries while their concurrent and leaderless nature enables them to achieve high throughput and scalability.
-Unlike blockchains that rely on proof-of-work, they are quiescent and green.
-Unlike traditional consensus protocols where one or more nodes typically process linear bits in the number of total nodes per decision, no node processes more than logarithmic bits.
+%O This paper introduces a family of leaderless Byzantine fault tolerance protocols, built around a metastable mechanism via network subsampling.
+Cet article présente une famille de protocoles à tolérance de faute byzantine sans dirigeant, bâtis autour d'un mécanisme métastable via sous-échantillonnage réseau.
+%O These protocols provide a strong probabilistic safety guarantee in the presence of Byzantine adversaries while their concurrent and leaderless nature enables them to achieve high throughput and scalability.
+Ces protocoles fournissent une forte garantie de sûreté probabiliste en présence d'adversaires byzantins, tandis que leur nature simultanée et sans dirigeant leur permet d'être performants et extensibles.
+%O Unlike blockchains that rely on proof-of-work, they are quiescent and green.
+Contrairement aux blockchains fondées sur la preuve de travail (\emph{proof-of-work}), ils sont économes et écologiques.
+%O Unlike traditional consensus protocols where one or more nodes typically process linear bits in the number of total nodes per decision, no node processes more than logarithmic bits.
 It does not require accurate knowledge of all participants and exposes new possible tradeoffs and improvements in safety and liveness for building consensus protocols.
+Contrairement aux protocoles de consensus traditionnels où un ou plusieurs nœuds traitent typiquement des bits de manière linéaire dans les nombres de nœuds totaux par décisiono, aucun nœud ne les traite plus que de manière logarithmique.
+HOULA
 %Finally, unlike traditional consensus protocols and similar to longest-chain protocols, our protocols provide increase in both liveness and safety guarantees when the observed adversarial presence decreases below the maximal expected bound.
@@
 
-The paper describes the Snow protocol family, analyzes its guarantees, and describes how it can be used to construct the core of an internet-scale electronic payment system called Avalanche, which is evaluated in a large scale deployment.
-Experiments demonstrate that the system can achieve high throughput (3400 tps), provide low confirmation latency (1.35 sec), and scale well compared to existing systems that deliver similar functionality.
+%O The paper describes the Snow protocol family, analyzes its guarantees, and describes how it can be used to construct the core of an internet-scale electronic payment system called Avalanche, which is evaluated in a large scale deployment.
+Cet article décrit la famille de protocoles Snow, analyse ses garanties et décrit la manière dont il peut être utilisé pour construire le cœur d'un système de paiement électronique à l'échelle de l'internet appelé Avalanche, qui est évalué lors d'un déploiement à grande échelle.
+%O Experiments demonstrate that the system can achieve high throughput (3400 tps), provide low confirmation latency (1.35 sec), and scale well compared to existing systems that deliver similar functionality.
 For our implementation and setup, the bottleneck of the system is in transaction verification.
+Les expériences démontrent que le système peut atteindre un haut début (3400 tps), fournir une faible latence de confirmation (1,35 s) et correctement passer à l'échelle par rapport à d'autres systèmes offrant des fonctionnalités similaires.
+Le goulet d'étranglement du système dans l'implémentation et la configuration retenue est la vérification des transations.
 \end{abstract}
@@
 
-Achieving agreement among a set of distributed hosts lies at the core of countless applications, ranging from Internet-scale services that serve billions of people~\cite{Burrows06,HuntKJR10} to cryptocurrencies worth billions of dollars~\cite{marketcapcryptocurrency}.
-To date, there have been two main families of solutions to this problem.
-Traditional consensus protocols %, exemplified by PBFT~\cite{castro1999practical},
-rely on all-to-all communication to ensure that all correct nodes reach the same decisions with absolute certainty.
-Because they usually require quadratic communication overhead and accurate knowledge of membership, they have been difficult to scale
-to large numbers of participants.
-On the other hand, Nakamoto consensus protocols~\cite{nakamoto2008bitcoin,GarayKL15, PassSS17, SompolinskyZ15, SompolinskyLZ16, SompolinskyZ18, BentovHMN17, EyalGSR16,Kokoris-KogiasJ16,PassS16a, PassS18} have become popular with the rise of Bitcoin.
-These protocols provide a probabilistic safety guarantee: Nakamoto consensus decisions may revert with some probability $\varepsilon$.
-A protocol parameter allows this probability to be rendered arbitrarily small, enabling high-value financial systems to be constructed on this foundation.
-This family is a natural fit for open, permissionless settings where any node can join the system at any time.
-Yet, these protocols are costly, wasteful, and limited in performance.
-By construction, they cannot quiesce: their security relies on constant participation by miners, even when there are no decisions to be made.
-Bitcoin currently consumes around 63.49 TWh/year~\cite{bitcoinpower}, about twice as all of Denmark~\cite{denmarkpower}.
-Moreover, these protocols suffer from an inherent scalability bottleneck that is difficult to overcome through simple reparameterization~\cite{CromanDEGJKMSSS16}.
+%O Achieving agreement among a set of distributed hosts lies at the core of countless applications, ranging from Internet-scale services that serve billions of people~\cite{Burrows06,HuntKJR10} to cryptocurrencies worth billions of dollars~\cite{marketcapcryptocurrency}.
+L'accord parmi un ensemble d'hôtes distribués se trouve au cœur d'innombrables applications allant de services à l'échelle de l'Internet servant des milliards de gens~\cite{Burrows06,HuntKJR10} à des cryptomonnaies valant des milliards de dollars~\cite{marketcapcryptocurrency}.
+%O To date, there have been two main families of solutions to this problem.
+À ce jour, il existe deux familles principales de solutions à ce problème.
+%O Traditional consensus protocols %, exemplified by PBFT~\cite{castro1999practical},
+Les protocoles de consensus traditionnels
+%O rely on all-to-all communication to ensure that all correct nodes reach the same decisions with absolute certainty.
+sont fondés sur une communication entre chaque pair pour s'assurer que tous les nœuds correct arrivent aux mêmes décisions avec une certitude absolue.
+%Because they usually require quadratic communication overhead and accurate knowledge of membership, they have been difficult to scale
+Comme ils impliquent en général un nombre de communication exponentiel et une connaissance exacte des participants, il s'est avéré difficile de l'étendre
+%O to large numbers of participants.
+à de nombreux participants.
+%O On the other hand, Nakamoto consensus protocols~\cite{nakamoto2008bitcoin,GarayKL15, PassSS17, SompolinskyZ15, SompolinskyLZ16, SompolinskyZ18, BentovHMN17, EyalGSR16,Kokoris-KogiasJ16,PassS16a, PassS18} have become popular with the rise of Bitcoin.
+D'un autre côté, les protocoles de consensus Nakamoto~\cite{nakamoto2008bitcoin,GarayKL15, PassSS17, SompolinskyZ15, SompolinskyLZ16, SompolinskyZ18, BentovHMN17, EyalGSR16,Kokoris-KogiasJ16,PassS16a, PassS18} ont été popularisés par Bitcoin.
+%O These protocols provide a probabilistic safety guarantee: Nakamoto consensus decisions may revert with some probability $\varepsilon$.
+Ces protocoles fournissent une garantie de sûreté probabiliste : les décisions par consensus Nakamoto peuvent s'inverser avec une certaine probabilité $\varepsilon$.
+%O A protocol parameter allows this probability to be rendered arbitrarily small, enabling high-value financial systems to be constructed on this foundation.
+Un paramètre du protocole permet d'abaisser à volonté cette probabilité, ce qui permet de bâtir des systèmes financiers de grande valeur sur ces fondations.
+%O This family is a natural fit for open, permissionless settings where any node can join the system at any time.
+Cette famille est faite pour des configurations ouvertes et en libre accès où tout nœud peut rejoindre le système à tout moment.
+%O Yet, these protocols are costly, wasteful, and limited in performance.
+Cependant ces protocoles sont coûteux, gaspilleurs et limités en performance.
+%O By construction, they cannot quiesce: their security relies on constant participation by miners, even when there are no decisions to be made.
+Par construction, ils sont tenus de fonctionner même à vide : leur sécurité se base sur une participation constante des mineurs, même quand aucune décision ne doit être prise.
+%O Bitcoin currently consumes around 63.49 TWh/year~\cite{bitcoinpower}, about twice as all of Denmark~\cite{denmarkpower}.
+Bitcoin consomme actuellement autour de 63.49 TWh/an~\cite{bitcoinpower}, environ deux fois l'énergie pour le Danemark~\cite{denmarkpower}.
+%O Moreover, these protocols suffer from an inherent scalability bottleneck that is difficult to overcome through simple reparameterization~\cite{CromanDEGJKMSSS16}.
 % \Jon{I wanted the paper to expand on this point, but I missed it.}
-
-This paper introduces a new family of consensus protocols called Snow.
-Inspired by gossip algorithms, this family gains its properties through a deliberately metastable mechanism.
-Specifically, the system operates by repeatedly sampling the network at random, and steering correct nodes towards a common outcome.
-Analysis shows that this metastable mechanism is powerful: it can move a large network to an irreversible state quickly, where the irreversibility implies that a sufficiently large portion of the network has accepted a proposal and a conflicting proposal will not be accepted with any higher than negligible ($\varepsilon$) probability.
+De plus, ces protocoles souffrent d'un goulet d'étranglement intrinsèque pour le passage à l'échelle, qui est difficile à dépasser par la simple reparamétrisation~\cite{CromanDEGJKMSSS16}.
+%O This paper introduces a new family of consensus protocols called Snow.
+Cet article présente une nouvelle famille de protocoles de consensus appelés Snow.
+%O Inspired by gossip algorithms, this family gains its properties through a deliberately metastable mechanism.
+Inspirée par les algorithmes de \emph{gossip} ou bavardage, cette famille possède des propriétés grâce à un mécanisme délibérément métastable.
+%O Specifically, the system operates by repeatedly sampling the network at random, and steering correct nodes towards a common outcome.
+En particulier, le système opère en échantillonnant répétitivement le réseau au hasard, et en dirigeant les nœuds corrects vers un résultat commun.
+%O Analysis shows that this metastable mechanism is powerful: it can move a large network to an irreversible state quickly, where the irreversibility implies that a sufficiently large portion of the network has accepted a proposal and a conflicting proposal will not be accepted with any higher than negligible ($\varepsilon$) probability.
+L'analyse montre que ce mécanisme métastable est puissant : il peut rapidement changer l'état d'un vaste réseau de manière irréversible, où l'irréversibilité implique qu'une part suffisamment importante du réseau a accepté une proposition et qu'une proposition différente ne sera acceptée qu'avec une probabilité au maximum négligeable.
 %{\color{red} We should clarify this.  The current statement is too strong.}
@@
 
-Similar to Nakamoto consensus, the Snow protocol family provides a probabilistic safety guarantee, using a tunable security parameter that can render the possibility of a consensus failure arbitrarily small.
-Unlike Nakamoto consensus, the protocols are green, quiescent and efficient; they do not rely on proof-of-work~\cite{DworkN92} and do not consume energy when there are no decisions to be made.
-The efficiency of the protocols stems partly from removing the leader bottleneck: each node requires $\Oh{1}$ communication overhead per round and $\Oh{\log{n}}$ rounds in expectation, whereas classical consensus protocols have one or more nodes that require $\Oh{n}$ communication per round (phase).
-Further, the Snow family tolerates discrepancies in knowledge of membership, as we discuss later.
+%O Similar to Nakamoto consensus, the Snow protocol family provides a probabilistic safety guarantee, using a tunable security parameter that can render the possibility of a consensus failure arbitrarily small.
+De même que le consensus Nakamoto, la famille des protocoles Snow fournissent une garantie de sûreté probabiliste, grâce à un paramètre des sécurité réglable qui peut amenuiser à volonter la possibilité d'un échec du consensus.
+%O Unlike Nakamoto consensus, the protocols are green, quiescent and efficient; they do not rely on proof-of-work~\cite{DworkN92} and do not consume energy when there are no decisions to be made.
+Contrairement au consensus Nakamoto, ces protocoles sont écologiques, économes et efficients ; ils ne dépendent pas d'une preuve de travail\cite{DworkN92} et ne consomment pas d'énergie quand il n'y a pas de décision à prendre.
+%0 The efficiency of the protocols stems partly from removing the leader bottleneck: each node requires $\Oh{1}$ communication overhead per round and $\Oh{\log{n}}$ rounds in expectation, whereas classical consensus protocols have one or more nodes that require $\Oh{n}$ communication per round (phase).
+L'efficacité des protocoles provient en partie de l'élimination du goulet d'étranglement des dirigeants : chaque nœud ne demande qu'un nombre de communications correspondant à $\Oh{1}$ par tour et une prévision de $\Oh{\log{n}}$ tours, alors que, dans les protocoles de consensus classique, un ou plusieurs nœuds requièrent $\Oh{n}$ communications par tour.
+%O Further, the Snow family tolerates discrepancies in knowledge of membership, as we discuss later.
 In contrast, classical consensus protocols require the full and accurate knowledge of $n$ as its safety foundation.
-
-Snow's subsampled voting mechanism has two additional properties that improve on previous approaches for consensus. 
-Whereas the safety of quorum-based approaches breaks down immediately when the predetermined threshold $f$ is exceeded,
-Snow's probabilistic safety guarantee degrades smoothly when Byzantine participants exceed $f$.
-This makes it easier to pick the critical threshold $f$.
-It also exposes new tradeoffs between safety and liveness: the Snow family is more efficient when the fraction of Byzantine nodes is small, and it can be parameterized to tolerate more than a third of the Byzantine
-nodes by trading off liveness.
-
-To demonstrate the potential of this protocol family, we illustrate a practical
-peer-to-peer payment system, Avalanche.
+De plus, la famille Snow tolère des incohérences dans la connaissance des participants, comme nous le verrons plus loin.
+Par contraste, les protocoles de consensus classiques exigent une connaisssance complète et exacte de $n$ comme fondation de la sûreté.
+
+%O Snow's subsampled voting mechanism has two additional properties that improve on previous approaches for consensus. 
+Le mécanisme de vote sous-échantillonné possède deux propriété supplémentaires supérieures par rapport aux approches antérieures du consensus.
+%O Whereas the safety of quorum-based approaches breaks down immediately when the predetermined threshold $f$ is exceeded,
+%O Snow's probabilistic safety guarantee degrades smoothly when Byzantine participants exceed $f$.
+Alors que la sûreté des approches basées sur un quorum s'effondre immédiatement quand le seuil prédéterminé $f$ est dépassé,
+la garantie de sûreté probabiliste de Snow se dégrade progressivement quand les participants byzantins dépassent $f$.
+%O This makes it easier to pick the critical threshold $f$.
+Cela facilite le choix du seuil critique $f$.
+%O It also exposes new tradeoffs between safety and liveness: the Snow family is more efficient when the fraction of Byzantine nodes is small, and it can be parameterized to tolerate more than a third of the Byzantine
+%O nodes by trading off liveness.
+Cela expose également de nouveaux compromis entre sûreté et vitalité : la famille Snow est plus efficace quand la fraction de nœuds byzantins est petite, et cela peut être paramétré pour tolérer plus d'un tiers de nœuds byzantins en sacrifiant la vitalité.
+
+%O To demonstrate the potential of this protocol family, we illustrate a practical
+%O peer-to-peer payment system, Avalanche.
 In effect, Avalanche executes multiple Snowball (one from the Snow family) instances with the aid of a Directed Acyclic Graph (DAG).
 The DAG serves to piggyback multiple instances, reducing the cost from $\Oh{\log{n}}$ to $\Oh{1}$ per node and streamlining the path where there are no conflicting transactions.
+Pour démontrer le potentiel de cette famille de protocoles, nous l'illustrons par un système pratique de paiement pair à pair, Avalanche.
+Dans les faits, Avalanche exécute plusieurs instances de Snowball (l'un des protocoles de la famille Snow) avec l'aide d'un graphe orienté acyclique (\emph{Directed Acyclic Graph} ou DAG).
+Le DAG sert à empiler plusieurs instances en réduisant le coût de $\Oh{\log{n}}$ à $\Oh{1}$ par nœud et en optimisant le chemin où il n'existe pas de transaction contradictoire.
 %This combination of the best features of traditional and Nakamoto consensus involves one significant tradeoff: liveness for \editchange{equivocating proposals}. %conflicting transactions.
@@
 
-Overall, the main contribution of this paper is to introduce a brand new family
-of consensus protocols, based on randomized sampling and metastable decision.
-The next section provides the model, goals and necessary assumptions for the new protocols.
-Section~\ref{sec:protocol} gives intuition behind the protocols, followed by their full specification,
+%O Overall, the main contribution of this paper is to introduce a brand new family
+%O of consensus protocols, based on randomized sampling and metastable decision.
+En résumé, la principale contribution de cet article consiste à présenter une nouvelle famille de protocoles de consensus fondés sur un échantillonnage aléatoire et des décisions métastables.
+%O The next section provides the model, goals and necessary assumptions for the new protocols.
+La section suivante fournit le modèle, les buts et les suppositions nécessaires pour les nouveaux protocoles.
+%O Section~\ref{sec:protocol} gives intuition behind the protocols, followed by their full specification,
+La section~\ref{sec:protocol} donne l'intuition derrière les protocoles, suivis de leurs spécifications complètes,
 %Section~\ref{sec:attack} briefly discusses typical attack vectors and their possible defense,
-Section~\ref{sec:analysis} provides methodology used
-by our formal analysis of safety and liveness in Appendix~\ref{sec:full-analysis},
+%O Section~\ref{sec:analysis} provides methodology used
+%O by our formal analysis of safety and liveness in Appendix~\ref{sec:full-analysis},
+La section~\ref{sec:analysis} fournit la méthodologie employée par notre analyse formelle de sûreté et de vitalité en Appendix~\ref{sec:full-analysis},
 %Section~\ref{sec:implementation} describes a Bitcoin-like payment system,  Section~\ref{sec:evaluation} evaluates the system,
-Section~\ref{sec:implementation} describes Avalanche, a Bitcoin-like payment system,
-Section~\ref{sec:evaluation} evaluates Avalanche,
-Section~\ref{sec:related-work} presents related work, and finally, Section~\ref{sec:conclusions} summarizes our contributions.
+%O Section~\ref{sec:implementation} describes Avalanche, a Bitcoin-like payment system,
+La section~\ref{sec:implementation} décrit Avalanche, un système de paiement similaire à Bitcoin,
+%O Section~\ref{sec:evaluation} evaluates Avalanche,
+La section~\ref{sec:evaluation} évalue Avalanche,
+%O Section~\ref{sec:related-work} presents related work, and finally, Section~\ref{sec:conclusions} summarizes our contributions.
+La section~\ref{sec:related-work} présente les travaux en relation, et, finalement, la section~\ref{sec:conclusions} résume nos contributions.
 
@@
 \label{sec:evaluation}
-\input{sections/evaluation}
+\input{sections/evaluation_fr}
 

### 2020-11-05 paper_fr.tex
S2 - Sureté, Vitalité

@@
 It does not require accurate knowledge of all participants and exposes new possible tradeoffs and improvements in safety and liveness for building consensus protocols.
-Contrairement aux protocoles de consensus traditionnels où un ou plusieurs nœuds traitent typiquement des bits de manière linéaire dans les nombres de nœuds totaux par décisiono, aucun nœud ne les traite plus que de manière logarithmique.
+Contrairement aux protocoles de consensus traditionnels où un ou plusieurs nœuds traitent typiquement des bits de manière linéaire dans les nombres de nœuds totaux par décision, aucun nœud ne les traite plus que de manière logarithmique.
 HOULA
@@
 For our implementation and setup, the bottleneck of the system is in transaction verification.
-Les expériences démontrent que le système peut atteindre un haut début (3400 tps), fournir une faible latence de confirmation (1,35 s) et correctement passer à l'échelle par rapport à d'autres systèmes offrant des fonctionnalités similaires.
+Les expériences démontrent que le système peut atteindre un haut débit (3400~tps), fournir une faible latence de confirmation (1,35~s) et correctement passer à l'échelle par rapport à d'autres systèmes offrant des fonctionnalités similaires.
 Le goulet d'étranglement du système dans l'implémentation et la configuration retenue est la vérification des transations.
@@
 %O Achieving agreement among a set of distributed hosts lies at the core of countless applications, ranging from Internet-scale services that serve billions of people~\cite{Burrows06,HuntKJR10} to cryptocurrencies worth billions of dollars~\cite{marketcapcryptocurrency}.
-L'accord parmi un ensemble d'hôtes distribués se trouve au cœur d'innombrables applications allant de services à l'échelle de l'Internet servant des milliards de gens~\cite{Burrows06,HuntKJR10} à des cryptomonnaies valant des milliards de dollars~\cite{marketcapcryptocurrency}.
+L'accord parmi un ensemble d'hôtes distribués se retrouve au cœur d'innombrables applications allant de services à l'échelle de l'Internet servant des milliards de gens~\cite{Burrows06,HuntKJR10} à des cryptomonnaies valant des milliards de dollars~\cite{marketcapcryptocurrency}.
 %O To date, there have been two main families of solutions to this problem.
@@
 %O Traditional consensus protocols %, exemplified by PBFT~\cite{castro1999practical},
+%O rely on all-to-all communication to ensure that all correct nodes reach the same decisions with absolute certainty.
 Les protocoles de consensus traditionnels
-%O rely on all-to-all communication to ensure that all correct nodes reach the same decisions with absolute certainty.
 sont fondés sur une communication entre chaque pair pour s'assurer que tous les nœuds correct arrivent aux mêmes décisions avec une certitude absolue.
 %Because they usually require quadratic communication overhead and accurate knowledge of membership, they have been difficult to scale
+%O to large numbers of participants.
 Comme ils impliquent en général un nombre de communication exponentiel et une connaissance exacte des participants, il s'est avéré difficile de l'étendre
-%O to large numbers of participants.
 à de nombreux participants.
@@
 %O Bitcoin currently consumes around 63.49 TWh/year~\cite{bitcoinpower}, about twice as all of Denmark~\cite{denmarkpower}.
-Bitcoin consomme actuellement autour de 63.49 TWh/an~\cite{bitcoinpower}, environ deux fois l'énergie pour le Danemark~\cite{denmarkpower}.
+Bitcoin consomme actuellement autour de 63.49 TWh/an~\cite{bitcoinpower}, environ deux fois l'énergie consommée par le Danemark~\cite{denmarkpower}.
 %O Moreover, these protocols suffer from an inherent scalability bottleneck that is difficult to overcome through simple reparameterization~\cite{CromanDEGJKMSSS16}.
@@
 
-\section{Model and Goals}
+%O \section{Model and Goals}
+\section{Modèle et buts}
 \label{sec:model_and_goals}
@@
 
-\tronly{\paragraph{Key Guarantees}}{}
+%O \tronly{\paragraph{Key Guarantees}}{}
+\tronly{\paragraph{Garanties clefs}}{}
 
@@
 
-\paragraph{Safety} Unlike classical consensus protocols, and similar to longest-chain-based consensus protocols such as Nakamoto consensus~\cite{nakamoto2008bitcoin}, we adopt an $\varepsilon$-safety guarantee that is probabilistic.
-In practice, this probabilistic guarantee is as strong as traditional safety guarantees, since appropriately small choices of $\varepsilon$ can render consensus failure negligible, lower than the probability of hardware failure due to random events.
-Figure~\ref{fig:fandepsilon} shows how the portion ($f/n$) of misbehaving participants (or computation power) affects the probability of system safety failure (decision of two conflicting proposals), given a choice of finality.
+%O \paragraph{Safety} Unlike classical consensus protocols, and similar to longest-chain-based consensus protocols such as Nakamoto consensus~\cite{nakamoto2008bitcoin}, we adopt an $\varepsilon$-safety guarantee that is probabilistic.
+\paragraph{Sûreté} À l'inverse des protocoles de consensus classiques, et de même que les protocoles de consensus fondés sur la chaîne la plus longue~\cite{nakamoto2008bitcoin}, nous adoptions une garantie de sûreté-$\varepsilon$ qui est probabiliste.
+%O In practice, this probabilistic guarantee is as strong as traditional safety guarantees, since appropriately small choices of $\varepsilon$ can render consensus failure negligible, lower than the probability of hardware failure due to random events.
+En pratique, cette garantie probabiliste est aussi forte que les garantie de sûreté traditionnelles, puisque le choix d'un petit $\varepsilon$ peut rendre la défaillance du consensus négligeable, inférieurs à la probabilité d'une panne matérielle due à des événements imprévisibles.
+%OFigure~\ref{fig:fandepsilon} shows how the portion ($f/n$) of misbehaving participants (or computation power) affects the probability of system safety failure (decision of two conflicting proposals), given a choice of finality.
+La figure~\ref{fig:fandepsilon} montre comment la portion ($f/n$) de participants (ou de puissance informatique) problématique affecte la probabilité d'une défaillance de la sûreté du système (décision de deux propositions incompatibles), selon un choix de finalité.
 % The maximum tolerated Byzantine presence $f$, is dependent on the choice of $\varepsilon$.
@@
     \includegraphics[width=\linewidth]{figures/safety-f-eps.pdf}
-    \caption{Classical BFT protocols that tolerate $f$ failures will encounter total safety failure when the threshold is exceeded even by one additional node.
+%O    \caption{Classical BFT protocols that tolerate $f$ failures will encounter total safety failure when the threshold is exceeded even by one additional node.
 The Bitcoin curve shows a typical finality choice for Bitcoin where a block is considered final when it is ``buried'' in a branch having 6 additional blocks compared to any other competing forks.
@@
 Snowflake-7,8 uses $\alpha=7$ and $\alpha=8$ respectively (see Section~\ref{sec:protocol} for the definition of $k$, $\alpha$ and $\beta$.}
+    \caption{Les protocoles BFT classiques qui tolèrent $f$ défaillances rencontrent une défaillance totale de leur sûreté quand le seuil est dépaasé même d'un seul nœud supplémentaire.
+La courbe de Bitcoin montre un choix de finalité typique pour Bitcoin où un bloc est considéré final quand il est \guillemotleft\,enterré\,\guillemotright~ dans une branche sous 6 blocs supplémentaires par rapport aux autres branches concurrentes.
+Snowflake appartient à la famille Snow et est configuré avec $k=10$, $\beta=150$.
+Snowflake-7,8 utilise $\alpha=7$ et $\alpha=8$ respectivement (voir en section~\ref{sec:protocol} pour la définition de $k$, $\alpha$ et $\beta$.}
     \label{fig:fandepsilon}
@@
     \includegraphics[width=\linewidth]{figures/safety-f-eps-log.pdf}
-    \caption{Figure~\ref{fig:fandepsilon} with log-scaled y-axis.}
+    \caption{Figure~\ref{fig:fandepsilon} Axe des y en log.}
     \label{fig:fandepsilonlog}
@@
 
-\paragraph{Liveness} All our protocols provide a non-zero probability guarantee of termination within a bounded amount of time. 
-This bounded guarantee is similar to various protocols such as Ben-Or~\cite{ben1983another} and longest-chain protocols.
-In particular, for Nakamoto consensus, the number of required blocks for a transaction increases exponentially with the number of adversarial nodes, with an asymptote at $f = n/2$ wherein the number is infinite.
-In other words, the time required for finality approaches $\infty$ as $f$ approaches $n/2$\tronly{ (Figure~\ref{fig:livenessproperties}).}{.}
-Furthermore, the required number of rounds is calculable ahead of time, as to allow the system designer to tune liveness at the expense of safety.
+%O \paragraph{Liveness} All our protocols provide a non-zero probability guarantee of termination within a bounded amount of time. 
+\paragraph{Vitalité} Tous nos protocoles fournissent une garantie de probabilité supérieure à zéro de terminaison dans un laps de temps donné.
+%O This bounded guarantee is similar to various protocols such as Ben-Or~\cite{ben1983another} and longest-chain protocols.
+Cette garantie de limite est similaire à celle de divers protocoles comme Ben-Or~\cite{ben1983another} et les protocoles de chaîne la plus longue.
+%O In particular, for Nakamoto consensus, the number of required blocks for a transaction increases exponentially with the number of adversarial nodes, with an asymptote at $f = n/2$ wherein the number is infinite.
+En particulier, pour le consensus Nakamoto, le nombre de blocs requis pour une transaction augmente exponentiellement avec le nombre de nœuds malveillants, avec une asymptote à $f = n/2$ où le nombre est infini.
+% wut?
+%O In other words, the time required for finality approaches $\infty$ as $f$ approaches $n/2$\tronly{ (Figure~\ref{fig:livenessproperties}).}{.}
+En d'autres termes, le temps nécessaire à la finalité approche $\infty$ quand $f$ approche $n/2$\tronly{ (Figure~\ref{fig:livenessproperties}).}{.}
+%O Furthermore, the required number of rounds is calculable ahead of time, as to allow the system designer to tune liveness at the expense of safety.
 Lastly, unlike traditional consensus protocols and similar to Nakamoto, our protocols benefit from lower adversarial presence, as discussed in property P3 below.
+De plus, le nombre requis de tours est calculable à l'avance, ce qui permet au concepteur du système de réglere la vitalité aux dépens de la sûreté.
+Enfin, contrairement aux protocoles de consensus traditionnels et semblablement à Nakamoto, nos protocoles bénéficient d'une moindre présence antagoniste, comme le détaille la propriété P3 ci-dessous.
 
@@
     \includegraphics[width=\linewidth]{figures/liveness-f-time.pdf}
-    \caption{The relation between $f/n$ and the convergence speed, given $\varepsilon = 10^{-20}$.
+%O    \caption{The relation between $f/n$ and the convergence speed, given $\varepsilon = 10^{-20}$.
 The left figure shows the expected number of blocks to guarantee $\varepsilon$ in Bitcoin, which, counter to commonly accepted folk wisdom, is not a constant $6$, but depends on adversary size to withhold the same $\varepsilon$.
 The right figure shows the maximum number of rounds required by Snowflake, where being different from Bitcoin, the asymptote is below $0.5$ and varies by the choice of parameters.}
+    \caption{Relation entre $f/n$ et la vitesse de convergence pour $\varepsilon = 10^{-20}$.
+La figure de gauche montre le nombre attendu de blocs pour garantir $\varepsilon$ dans Bitcoin, qui, contrairement à la croyance populaire, n'est pas une constante $6$, mais dépend de la taille de l'adversaire pour maintenir le même $\varepsilon$.
+La figure de droite montre le nombre maximum de tours requis par Snowflake où, étant différent de Bitcoin, l'asymptote se trouve sous $0.5$ et varie selon les paramètres.}
     \label{fig:livenessproperties}

### 2020-11-05 paper_fr.tex
S2 - Déseau, atteinte de la vitalité, adversaire

@@
 %O In particular, for Nakamoto consensus, the number of required blocks for a transaction increases exponentially with the number of adversarial nodes, with an asymptote at $f = n/2$ wherein the number is infinite.
-En particulier, pour le consensus Nakamoto, le nombre de blocs requis pour une transaction augmente exponentiellement avec le nombre de nœuds malveillants, avec une asymptote à $f = n/2$ où le nombre est infini.
+En particulier, pour le consensus Nakamoto, le nombre de blocs requis pour une transaction augmente exponentiellement avec le nombre de nœuds antagonistes (\emph{adversarial}), avec une asymptote à $f = n/2$ où le nombre est infini.
 % wut?
@@
 
-\emph{Formal Guarantees}: Let the system be parameterized for an $\varepsilon$ safety failure probability under a maximum expected $f$ number of adversarial nodes.
+%O \emph{Formal Guarantees}: Let the system be parameterized for an $\varepsilon$ safety failure probability under a maximum expected $f$ number of adversarial nodes.
 Let $\Oh{\log n} < t_{max} < \infty$ be the upper bound of the execution of the protocols.
 The Snow protocols then provide the following guarantees:
+\emph{Garantie formelle} : Soit un paramétrage du système pour une probabilité de défaillance de la sûreté $\varepsilon$ pour un nombre maximal attendu $f$ de nœuds antagonistes.
+Soit $\Oh{\log n} < t_{max} < \infty$ la limite supérieure de l'exécution du protocole.
+Les protocoles Snow fournissent alors les garanties suivantes\,:
 \begin{compactitem}
@@
 
-    \item \textbf{P1. Safety.} When decisions are made by any two correct nodes, they decide on conflicting transactions with negligible probability ($\leq \varepsilon$).
+%O     \item \textbf{P1. Safety.} When decisions are made by any two correct nodes, they decide on conflicting transactions with negligible probability ($\leq \varepsilon$).
+     \item \textbf{P1. Sûreté.} Quand les décisions sont prises par deux nœuds corrects, ils décident des transactions contradictoires avec une probabilité négligeable ($\leq \varepsilon$).
     % \item \textbf{P2.
 Liveness.} Transactions will be accepted by every correct node in $\Oh{\log n}$ rounds if either the observed Byzantine presence $f'$, where $f' < f$, is sufficiently small, or the network is initialized in a univalent state.
-    \item \textbf{P2. Liveness (Upper Bound).} Snow protocols terminate with a strictly positive probability within $t_{max}$ rounds.  
-    \item \textbf{P3. Liveness (Strong Form).} If $f \leq \Oh{\sqrt{n}}$, then the Snow protocols terminate with high probability ($\geq 1 - \varepsilon$) in $\Oh{\log{n}}$ rounds. 
+%O     \item \textbf{P2. Liveness (Upper Bound).} Snow protocols terminate with a strictly positive probability within $t_{max}$ rounds.  
+     \item \textbf{P2. Vitalité (limite supérieure).} Les protocoles Snow se terminent avec une probabilité strictement positive en au plus $t_{max}$ tours.
+%O     \item \textbf{P3. Liveness (Strong Form).} If $f \leq \Oh{\sqrt{n}}$, then the Snow protocols terminate with high probability ($\geq 1 - \varepsilon$) in $\Oh{\log{n}}$ rounds. 
+     \item \textbf{P3. Vitalité (forme forte).} Si $f \leq \Oh{\sqrt{n}}$, alors les protocoles Snow se terminent avec une forte probabilité ($\geq 1 - \varepsilon$) en $\Oh{\log{n}}$ tours.
 \end{compactitem}
@@
 
-\paragraph{Network} 
-In the standard definition of asynchrony~\cite{ben1983another}, message transmission time is finite, but the distribution is unspecified (and thus the delivery time can be unbounded for some messages).
+%O \paragraph{Network} 
+\paragraph{Réseau} 
+%O In the standard definition of asynchrony~\cite{ben1983another}, message transmission time is finite, but the distribution is unspecified (and thus the delivery time can be unbounded for some messages).
 This implies that the scheduling of message transmission itself could behave arbitrarily, and potentially even maliciously (with full asynchrony).
-We use a modified version of this model, which is well-accepted~\cite{banerjee2014epidemic, ganesh2005effect, draief2006thresholds, keeling2011modeling, liggett1997stochastic} in the analysis of epidemic networks and gossip-based stochastic systems.
+Dans la définition standard de l'asynchronie~\cite{ben1983another}, le temps de transmission des messages est fini, mais la distribution n'est pas spécifiée (donc le temps de délivrance peut être sans limite pour certains messages).
+Cela implique que l'échelonnement de la transmission des messages lui-même peut être quelconque, voire potentiellement malveillant (pour une complète asynchronie).
+%O We use a modified version of this model, which is well-accepted~\cite{banerjee2014epidemic, ganesh2005effect, draief2006thresholds, keeling2011modeling, liggett1997stochastic} in the analysis of epidemic networks and gossip-based stochastic systems.
 In particular, we fix the distribution of message delay to that of the exponential distribution.
-We note that, just like in the standard asynchronous model, there is a strictly non-zero probability that any correct node may execute its next local round only after an arbitrarily large amount of time has passed.
-Furthermore, we also note that scheduling only applies to correct nodes, and the adversary may execute arbitrarily, as discussed later.
-
-\paragraph{Achieving Liveness}
-Classical consensus that works with asynchrony does not get stuck in a single phase of voting because the vote initiator always polls votes from all known participants and waits for $n - f$ responses.
+Nous employons one version modifiée de ce modèle, lequel est bien accepté~\cite{banerjee2014epidemic, ganesh2005effect, draief2006thresholds, keeling2011modeling, liggett1997stochastic} dans l'analyse des réseaux épidémiques et des systèmes stochastiques à base de \emph{gossip}.
+En particulier, nous fixons la distribution des délais de messages à celle de la distibution exponentielle.
+%O We note that, just like in the standard asynchronous model, there is a strictly non-zero probability that any correct node may execute its next local round only after an arbitrarily large amount of time has passed.
+Nous notons que, tout comme dans le modèle asynchrone standard, il existe un probabilité strictement non-nulle que tout nœud correct ne peut exécuter son prochain tour local qu'après qu'un laps de temps arbitrairement long a passé.
+%O Furthermore, we also note that scheduling only applies to correct nodes, and the adversary may execute arbitrarily, as discussed later. 
+De plus, nous notons également que l'échelonnement ne s'applique qu'aux nœuds corrects, et que l'adversaire peut exécuter arbitrairement, comme nous le verrons plus loin. % HOULA
+
+%O \paragraph{Achieving Liveness}
+\paragraph{Atteinte de la vitalité} % HOULA
+%O Classical consensus that works with asynchrony does not get stuck in a single phase of voting because the vote initiator always polls votes from all known participants and waits for $n - f$ responses.
+Les consensus classiques qui fonctionnent en asynchronie ne restent pas bloqués lors d'une phase de vote car l'initiateur du vote interroge toujours tous les participants connus et attend $n - f$ réponses.
 %The consequence of these two assumptions is therefore that progress cannot be indefinitely delayed by the adversary since the correct nodes (whose number is sufficient for voting) will eventually deliver the required messages.
-In our system, however, nodes operate via subsampling, hence it is possible for a single sample to select a majority of adversarial nodes, and therefore the node gets stuck waiting for the responses.
+%O In our system, however, nodes operate via subsampling, hence it is possible for a single sample to select a majority of adversarial nodes, and therefore the node gets stuck waiting for the responses.
 To ensure liveness, a node should be able to wait with some timeout.
@@
 \editinsert{RvR pointed this out, we need to carefully word it though.
-
-\paragraph{Adversary}
-The adversarial nodes execute under their own internal scheduler, which is unbounded in speed, meaning that all adversarial nodes can execute at any infinitesimally small point in time, unlike correct nodes.
-The adversary can view the state of every honest node at all times and can instantly modify the state of all adversarial nodes. 
-It cannot, however, schedule or modify communication between correct nodes.
-Finally, we make zero assumptions about the behavior of the adversary, meaning that it can choose any execution strategy of its liking.
-In short, the adversary is computationally bounded (it cannot forge digital signatures) but otherwise is point-to-point informationally unbounded (knows all state) and round-adaptive (can modify its strategy at any time).
+Dans notre système, cependant, les nœuds opèrent par sous-échantillonnage, d'où la possibilité pour un seul échantillon qu'une majorité de nœuds antagonistes soit sélectionner, ce qui contraint le nœud à attendre en vain une réponse.
+Pour s'assurer de la vitalité, un nœud doit pouvoir poser une limite à l'attente.
+Enfin, il faut noter que le consensus Nakamoto est synchrone et que la difficulté requise à la preuve de travail dépend du délai maximum du réseau~\cite{PassSS17}.
+
+%O \paragraph{Achieving Liveness}
+\paragraph{Adversaire}
+%O The adversarial nodes execute under their own internal scheduler, which is unbounded in speed, meaning that all adversarial nodes can execute at any infinitesimally small point in time, unlike correct nodes.
+Les nœuds antagonistes s'exécutent avec un ordonnanceur interne qui leur est propre, non limité en vitesse, ce qui signifie que tout nœud antagoniste peut s'exécuter à un moment précis infiniment petit, au contraire des nœuds corrects.
+%O The adversary can view the state of every honest node at all times and can instantly modify the state of all adversarial nodes. 
+L'adversaire peut voir l'état de chaque nœud honnête à tout moment et peut instantanément modifier l'état de tout nœud antagoniste.
+%O It cannot, however, schedule or modify communication between correct nodes.
+Il ne peut, cependant, ordonnancer ou modifier la communication entre les nœuds corrects.
+%O Finally, we make zero assumptions about the behavior of the adversary, meaning that it can choose any execution strategy of its liking.
+Enfin, nous ne faisons aucune hypothèse sur le comportement de l'adversaire, ce qui signifie qu'il peut choisir la stratégie d'exécution qui lui convient.
+%O In short, the adversary is computationally bounded (it cannot forge digital signatures) but otherwise is point-to-point informationally unbounded (knows all state) and round-adaptive (can modify its strategy at any time).
+En résumé, l'adversaire est limité en puissance informatique (il ne peut pas forger des signatures numériques) mais, hormis cela, il n'est pas limité en matière informationnelle (connaît tout de l'état) et il est adaptatif en matière de tours (peut modifier sa stratégie à tout moment).
 

### 2020-11-05 paper_fr.tex
French translation of section about Slush protocol

@@
 
-\subsection{Slush: Introducing Metastability}
-The core of our approach is a single-decree consensus protocol,
-inspired by epidemic or gossip protocols.
-% In this approach, a state in which all possible outcomes are equally
-% probable is metastable and disturbances are likely to cause the
-% protocol to converge to one of multiple stable states.
-% We will call such consensus protocols metastable themselves.
-%We can model a preference for a decision as a kind of infection: a node catches a disease, and converts other nodes to carry the same disease in epidemic fashion.
-%If a threshold of its interactions are with other infected hosts, it flips its color to the competing agent and joins their ranks.
-%
-The simplest protocol, Slush, is the foundation of this family, shown in Figure~\ref{fig:slush-loop}.
-Slush is \emph{not} tolerant to Byzantine faults, only crash-faults (CFT), but serves as an illustration for the BFT protocols that follow.
-For ease of exposition, we will describe the operation of Slush using a decision between two conflicting colors, red and blue.
-% , though the technique generalizes to any number of competing proposed values.
-% Slush operates as follows.
-
-In Slush, a node starts out initially in an uncolored state.
-Upon receiving a transaction from a client, an uncolored node updates its own color to the one carried in the transaction and initiates a query.
-To perform a query, a node picks a small, constant sized ($k$) sample of the network uniformly at random, and sends a query message.
-Upon receiving a query, an uncolored node adopts the color in the query, responds with that color, and initiates its own query, whereas a colored node simply responds with its current color.
-% If $k$ responses are not received within a time bound, the node picks an additional sample from the remaining nodes uniformly at random and queries them until it collects all responses.
-Once the querying node collects $k$ responses, it checks if a fraction $\ge\alpha$ are for the same color, where $\alpha > \floor{k/2}$ is a protocol parameter.
-If the $\alpha$ threshold is met and the sampled color differs from the node's own color, the node flips to that color.
-It then goes back to the query step, and initiates a subsequent round of query, for a total of $m$ rounds.
-Finally, the node decides the color it ended up with at time $m$.
+%0 \subsection{Slush: Introducing Metastability}
+\subsection{Slush: A propos de métastabililté}
+
+%O The core of our approach is a single-decree consensus protocol,
+%O inspired by epidemic or gossip protocols.
+%O The simplest protocol, Slush, is the foundation of this family, shown in Figure~\ref{fig:slush-loop}.
+%O Slush is \emph{not} tolerant to Byzantine faults, only crash-faults (CFT), but serves as an illustration for the BFT protocols that follow.
+%O For ease of exposition, we will describe the operation of Slush using a decision between two conflicting colors, red and blue.
+
+Le principe de notre approche est un protocole de consensus à décision simple inspiré des protocoles dits "épidémiques"
+ou "gossip". Le protocole le plus simple, Slush, est la fondation de cette famille, comme le montre la
+Figure~\ref{fig:slush-loop}. Slush \emph{n'est pas} tolérant aux fautes byzantines, seulement aux fautes par défaut
+de fonctionnement (CFT), mais sert à mettre en avant les autres protocoles tolérants aux fautes byzantines qui vont
+suivre. Dans un souci de simplicité de présentation, nous décrirons le fonctionnement de Slush en le comparant
+à une décision à prendre entre deux couleurs incompatibles, le rouge et le bleu.
+
+%O In Slush, a node starts out initially in an uncolored state.
+%O Upon receiving a transaction from a client, an uncolored node updates its own color to the one carried in the transaction and initiates a query.
+%O To perform a query, a node picks a small, constant sized ($k$) sample of the network uniformly at random, and sends a query message.
+%O Upon receiving a query, an uncolored node adopts the color in the query, responds with that color, and initiates its own query, whereas a colored node simply responds with its current color.
+%O Once the querying node collects $k$ responses, it checks if a fraction $\ge\alpha$ are for the same color, where $\alpha > \floor{k/2}$ is a protocol parameter.
+%O If the $\alpha$ threshold is met and the sampled color differs from the node's own color, the node flips to that color.
+%O It then goes back to the query step, and initiates a subsequent round of query, for a total of $m$ rounds.
+%O Finally, the node decides the color it ended up with at time $m$.
+
+Avec Slush, un noeud démarre dans un état de couleur neutre. Après réception d'une transaction envoyée par un client,
+un node de couleur neutre adopte la couleur définie dans la transaction et initie une requête. Pour lancer une
+requête, le noeud choisit un échantillon réduit et de taille constante ($k$) de noeuds appartenant au réseau suivant
+une distribution uniformément aléatoire puis initie sa propre requête, tandis qu'un noeud déja coloré répond
+simplement avec sa couleur actuelle. Une fois que le noeud qui fait la requête récupère $k$ réponses, il vérifie si
+une proportion $\ge\alpha$ de l'échantillon retourne la même couleur, $\alpha > \floor{k/2}$ étant un paramètre
+du protocole. Si le seuil $\alpha$ est dépassé et que la couleur dominante revenant de l'échantillon sollicité est
+différente de la couleur propre du node à l'origine de la requête, ce node adopte cette nouvelle couleur. Il revient
+ensuite à l'étape de génération de la requête et lance un nouveau tour de requête jusqu'à un total de $m$ tours.
+En définitive, le node décide de sa couleur finale à l'instant $m$.
 
@@
         \IIf{$\codecolor = \bot$} $\codecolor \assign \codecolor'$
-        \State\Call{respond}{$v, \codecolor$}
+        \State\Call{répondre}{$v, \codecolor$}
     \EndProcedure
     \Procedure{slushLoop}{$u, \codecolor_0 \in \{\texttt{R}, \texttt{B}, \bot\}$}
-        \State $\codecolor \assign \codecolor_0$ \textrm{// initialize with a color}
+        \State $\codecolor \assign \codecolor_0$ \textrm{// initialise avec une couleur}
         \For{$r \in \{1\ldots m\}$}
-            \State \textrm{// if $\bot$, skip until \textsc{onQuery} sets the color}
+            \State \textrm{// Si $\bot$, réitère jusqu'à ce que \textsc{onQuery} applique la couleur}
             \If{$\codecolor = \bot$}
@@
             \EndIf
-            \State \textrm{// randomly sample from the known nodes}
-            \State $\mathcal{K} \assign \Call{sample}{\mathcal{N}\backslash u, k}$
-            \State $P \assign \texttt{[}\Call{query}{v, \codecolor}\quad\textbf{for}\ v \in \mathcal{K}\texttt{]}$
+            \State \textrm{// échantillonne aléatoirement parmi les noeuds connus}
+            \State $\mathcal{K} \assign \Call{échantillon}{\mathcal{N}\backslash u, k}$
+            \State $P \assign \texttt{[}\Call{requête}{v, \codecolor}\quad\textbf{pour}\ v \in \mathcal{K}\texttt{]}$
             \For{$\codecolor' \in \{\texttt{R}, \texttt{B}\}$}
-                \If{$P.\Call{count}{\codecolor'} \ge \alpha$}
+                \If{$P.\Call{nombre}{\codecolor'} \ge \alpha$}
                     \State $\codecolor \assign \codecolor'$
@@
         \EndFor
-        \State \Call{accept}{$\codecolor$}
+        \State \Call{accepte}{$\codecolor$}
     \EndProcedure
-    \captionof{figure}{Slush protocol. Timeouts elided for readability.}\label{fig:slush-loop}
+    \captionof{figure}{Protocole Slush. Les temps d'expiration sont omis dans un souci de lisibilité.}\label{fig:slush-loop}
 \end{algorithmic}
@@
 
-Slush has a few properties of interest. 
-First, it is almost \emph{memoryless}: a node retains no state between rounds other than its current color, and in particular maintains no history of interactions with other peers.
-Second, unlike traditional consensus protocols that query every participant, every round involves sampling just a small, constant-sized slice of the network at random.
-Third, Slush makes progress under any network configuration (even fully bivalent state, i.e.
+%O Slush has a few properties of interest.
+%O First, it is almost \emph{memoryless}: a node retains no state between rounds other than its current color, and in particular maintains no history of interactions with other peers.
+%O Second, unlike traditional consensus protocols that query every participant, every round involves sampling just a small, constant-sized slice of the network at random.
+%O Third, Slush makes progress under any network configuration (even fully bivalent state, i.e.
 50/50 split between colors), since random perturbations in sampling will cause one color to gain a slight edge and repeated samplings afterwards will build upon and amplify that imbalance.
-Finally, if $m$ is chosen high enough, Slush ensures that all nodes will be colored identically with high probability (whp).
-Each node has a constant, predictable communication overhead per round, and $m$ grows logarithmically with $n$.
+%O Finally, if $m$ is chosen high enough, Slush ensures that all nodes will be colored identically with high probability (whp).
+%O Each node has a constant, predictable communication overhead per round, and $m$ grows logarithmically with $n$.
+
+Slush a quelques propriétés intéréssantes. Premièrement, il n'a quasiment \emph{pas besoin de mémoire}: un noeud ne
+stocke aucun état intermédiaire entre chaque tour autre que sa couleur actuelle, et plus particulièrement ne stocke
+aucun historique de ses interactions avec les autres noeuds. Deuxièmement, contrairement aux protocoles de consensus
+traditionnels qui sollicitent tous les participants, chaque tour n'implique qu'un échantillon du réseau réduit et de
+taille constante choisi aléatoirement. Troisièmement, Slush progresse quelque soit la configuration du réseau (même
+depuis un état entièrement bivalent, ex: partage 50/50 entre les deux couleurs) vu que des perturbations aléatoires
+sur l'échantillonnage auront pour conséquence de fournir un léger avantage à l'une des deux couleurs qui sera ensuite
+amplifié au fur et à mesure des échantillonnages.
+Finalement si $m$ est choisi pour être suffisamment élevé, Slush s'assure que tous les noeuds convergeront vers la
+même couleur avec une haute probabilité. Chaque noeud nécéssite un temps de communication par tour constant et
+prévisible, et $m$ croît de manière logarithmique suivant $n$.
 
 \tronly{
-The Slush protocol does not provide a strong safety guarantee in the presence of Byzantine nodes.
-In particular, if the correct nodes develop a preference for one color, a Byzantine adversary can attempt to flip nodes
-to the opposite so as to keep the network in balance, preventing a decision.
-We address this in our first BFT protocol that introduces more state storage at the nodes.
+%O The Slush protocol does not provide a strong safety guarantee in the presence of Byzantine nodes.
+%O In particular, if the correct nodes develop a preference for one color, a Byzantine adversary can attempt to flip nodes
+%O to the opposite so as to keep the network in balance, preventing a decision.
+%O We address this in our first BFT protocol that introduces more state storage at the nodes.
+
+Le protocole Slush ne fournit pas de fortes garanties quant à la sûreté de fonctionnement en la présence de noeuds
+byzantins. En particulier si les noeuds légitimes développent une préférence pour une certaine couleur, un adversaire
+byzantin pourra inverser la couleur des noeuds afin de maintenir le système à l'équilibre, incapable de prendre une
+décision. Nous adressons ce problème dans notre premier protocole tolérant aux fautes byzantines qui introduit un
+meilleur suivi de l'état au niveau des nodes.
 }{
-We next examine how to extend Slush to tolerate Byzantine behavior.
+%O We next examine how to extend Slush to tolerate Byzantine behavior.
+Nous examinons ensuite comment améliorer Slush pour être tolérant au problème des généraux byzantins.
 }

### 2020-11-06 paper_fr.tex
S2 - complet

--- 2020-11-05 paper_fr.tex
@@
 Dans les faits, Avalanche exécute plusieurs instances de Snowball (l'un des protocoles de la famille Snow) avec l'aide d'un graphe orienté acyclique (\emph{Directed Acyclic Graph} ou DAG).
-Le DAG sert à empiler plusieurs instances en réduisant le coût de $\Oh{\log{n}}$ à $\Oh{1}$ par nœud et en optimisant le chemin où il n'existe pas de transaction contradictoire.
+Le DAG sert à empiler plusieurs instances en réduisant le coût de $\Oh{\log{n}}$ à $\Oh{1}$ par nœud et en optimisant le chemin où il n'existe pas de transactions contradictoires.
 %This combination of the best features of traditional and Nakamoto consensus involves one significant tradeoff: liveness for \editchange{equivocating proposals}. %conflicting transactions.
@@
 %O \paragraph{Safety} Unlike classical consensus protocols, and similar to longest-chain-based consensus protocols such as Nakamoto consensus~\cite{nakamoto2008bitcoin}, we adopt an $\varepsilon$-safety guarantee that is probabilistic.
-\paragraph{Sûreté} À l'inverse des protocoles de consensus classiques, et de même que les protocoles de consensus fondés sur la chaîne la plus longue~\cite{nakamoto2008bitcoin}, nous adoptions une garantie de sûreté-$\varepsilon$ qui est probabiliste.
+\paragraph{Sûreté} À l'inverse des protocoles de consensus classiques, et à l'instar des protocoles de consensus fondés sur la chaîne la plus longue~\cite{nakamoto2008bitcoin}, nous adoptons une garantie de sûreté-$\varepsilon$ qui est probabiliste.
 %O In practice, this probabilistic guarantee is as strong as traditional safety guarantees, since appropriately small choices of $\varepsilon$ can render consensus failure negligible, lower than the probability of hardware failure due to random events.
-En pratique, cette garantie probabiliste est aussi forte que les garantie de sûreté traditionnelles, puisque le choix d'un petit $\varepsilon$ peut rendre la défaillance du consensus négligeable, inférieurs à la probabilité d'une panne matérielle due à des événements imprévisibles.
+En pratique, cette garantie probabiliste est aussi forte que les garantie de sûreté traditionnelles puisque le choix d'un petit $\varepsilon$ peut rendre la défaillance du consensus négligeable, inférieurs à la probabilité d'une panne matérielle due à des événements imprévisibles.
 %OFigure~\ref{fig:fandepsilon} shows how the portion ($f/n$) of misbehaving participants (or computation power) affects the probability of system safety failure (decision of two conflicting proposals), given a choice of finality.
@@
 %O \paragraph{Liveness} All our protocols provide a non-zero probability guarantee of termination within a bounded amount of time. 
-\paragraph{Vitalité} Tous nos protocoles fournissent une garantie de probabilité supérieure à zéro de terminaison dans un laps de temps donné.
+\paragraph{Vitalité} Tous nos protocoles fournissent une garantie de probabilité de terminaison dans un laps de temps donné supérieure à zéro.
 %O This bounded guarantee is similar to various protocols such as Ben-Or~\cite{ben1983another} and longest-chain protocols.
@@
 %O In particular, for Nakamoto consensus, the number of required blocks for a transaction increases exponentially with the number of adversarial nodes, with an asymptote at $f = n/2$ wherein the number is infinite.
-En particulier, pour le consensus Nakamoto, le nombre de blocs requis pour une transaction augmente exponentiellement avec le nombre de nœuds antagonistes (\emph{adversarial}), avec une asymptote à $f = n/2$ où le nombre est infini.
+En particulier, pour le consensus Nakamoto, le nombre de blocs requis pour une transaction augmente exponentiellement avec le nombre de nœuds antagonistes, avec une asymptote à $f = n/2$ où le nombre est infini.
 % wut?
@@
 Lastly, unlike traditional consensus protocols and similar to Nakamoto, our protocols benefit from lower adversarial presence, as discussed in property P3 below.
-De plus, le nombre requis de tours est calculable à l'avance, ce qui permet au concepteur du système de réglere la vitalité aux dépens de la sûreté.
-Enfin, contrairement aux protocoles de consensus traditionnels et semblablement à Nakamoto, nos protocoles bénéficient d'une moindre présence antagoniste, comme le détaille la propriété P3 ci-dessous.
+De plus, le nombre requis de tours est calculable à l'avance, ce qui permet au concepteur du système de régler la vitalité aux dépens de la sûreté.
+Enfin, contrairement aux protocoles de consensus traditionnels et similairement à Nakamoto, nos protocoles bénéficient d'une moindre présence antagoniste, comme le détaille la propriété P3 ci-dessous.
 
@@
 In particular, we fix the distribution of message delay to that of the exponential distribution.
-Nous employons one version modifiée de ce modèle, lequel est bien accepté~\cite{banerjee2014epidemic, ganesh2005effect, draief2006thresholds, keeling2011modeling, liggett1997stochastic} dans l'analyse des réseaux épidémiques et des systèmes stochastiques à base de \emph{gossip}.
-En particulier, nous fixons la distribution des délais de messages à celle de la distibution exponentielle.
+Nous employons une version modifiée de ce modèle, lequel est bien accepté~\cite{banerjee2014epidemic, ganesh2005effect, draief2006thresholds, keeling2011modeling, liggett1997stochastic} dans l'analyse des réseaux épidémiques et des systèmes stochastiques à base de \emph{gossip}.
+En particulier, nous fixons la distribution des délais de messages à celle de la distribution exponentielle.
 %O We note that, just like in the standard asynchronous model, there is a strictly non-zero probability that any correct node may execute its next local round only after an arbitrarily large amount of time has passed.
-Nous notons que, tout comme dans le modèle asynchrone standard, il existe un probabilité strictement non-nulle que tout nœud correct ne peut exécuter son prochain tour local qu'après qu'un laps de temps arbitrairement long a passé.
+Nous notons que, tout comme dans le modèle asynchrone standard, il existe une probabilité strictement non-nulle que tout nœud correct ne peut exécuter son prochain tour local qu'après qu'un laps de temps arbitrairement long a passé.
 %O Furthermore, we also note that scheduling only applies to correct nodes, and the adversary may execute arbitrarily, as discussed later. 
-De plus, nous notons également que l'échelonnement ne s'applique qu'aux nœuds corrects, et que l'adversaire peut exécuter arbitrairement, comme nous le verrons plus loin. % HOULA
+De plus, nous notons également que l'échelonnement ne s'applique qu'aux nœuds corrects, et que l'adversaire peut s'exécuter arbitrairement, comme nous le verrons plus loin.
 
 %O \paragraph{Achieving Liveness}
-\paragraph{Atteinte de la vitalité} % HOULA
+\paragraph{Obtention de la vitalité} % HOULA
 %O Classical consensus that works with asynchrony does not get stuck in a single phase of voting because the vote initiator always polls votes from all known participants and waits for $n - f$ responses.
@@
 \editinsert{RvR pointed this out, we need to carefully word it though.
-Dans notre système, cependant, les nœuds opèrent par sous-échantillonnage, d'où la possibilité pour un seul échantillon qu'une majorité de nœuds antagonistes soit sélectionner, ce qui contraint le nœud à attendre en vain une réponse.
+Dans notre système, cependant, les nœuds opèrent par sous-échantillonnage, d'où la possibilité pour un seul échantillon qu'une majorité de nœuds antagonistes soit sélectionné, ce qui contraint le nœud à attendre en vain une réponse.
 Pour s'assurer de la vitalité, un nœud doit pouvoir poser une limite à l'attente.
-Enfin, il faut noter que le consensus Nakamoto est synchrone et que la difficulté requise à la preuve de travail dépend du délai maximum du réseau~\cite{PassSS17}.
-
-%O \paragraph{Achieving Liveness}
-\paragraph{Adversaire}
+Enfin, il faut noter que le consensus Nakamoto est synchrone et que la difficulté requise pour la preuve de travail dépend du délai maximum du réseau~\cite{PassSS17}.
+
+%O \paragraph{Adversary}
+\paragraph{Adversité} % HOULA : Adversaire ? Adversaires ? Adversatif ?
 %O The adversarial nodes execute under their own internal scheduler, which is unbounded in speed, meaning that all adversarial nodes can execute at any infinitesimally small point in time, unlike correct nodes.
@@
 %O In short, the adversary is computationally bounded (it cannot forge digital signatures) but otherwise is point-to-point informationally unbounded (knows all state) and round-adaptive (can modify its strategy at any time).
-En résumé, l'adversaire est limité en puissance informatique (il ne peut pas forger des signatures numériques) mais, hormis cela, il n'est pas limité en matière informationnelle (connaît tout de l'état) et il est adaptatif en matière de tours (peut modifier sa stratégie à tout moment).
-
-\paragraph{Sybil Attacks}
-Consensus protocols provide their guarantees based on assumptions that only a fraction of participants are adversarial.
-These bounds could be violated if the network is naively left open to arbitrary participants.
-In particular, a Sybil attack~\cite{douceur2002sybil}, wherein a large number of identities are generated by an adversary, could be used to exceed the bounds.
-
-A long line of work, including PBFT~\cite{castro1999practical}, treats the Sybil problem separately from consensus, and rightfully so, as Sybil control mechanisms are distinct from the underlying, more complex agreement protocol\footnote{This is not to imply that every consensus protocol can be coupled/decoupled with every Sybil control mechanism.}.
-In fact, to our knowledge, only Nakamoto-style consensus has ``baked-in'' Sybil prevention as part of its consensus, made possible by chained proof-of-work~\cite{aspnes2005exposing}, which requires miners to continuously stake a hardware investment.
-Other protocols, discussed in Section~\ref{sec:related-work}, rely on proof-of-stake (by economic argument) or proof-of-authority (by administrative argument that makes the system ``permissioned'').
-The consensus protocols presented in this paper can adopt any
-Sybil control mechanism, although proof-of-stake is most aligned with their quiescent operation.
-One can use an already established proof-of-stake based mechanism~\cite{GiladHMVZ17}.
-The full deployment of an autonomous P2P payment system incorporating staking mechanism is beyond the scope of this paper, whose focus is on a novel design paradigm of the core consensus algorithm.
+En résumé, l'adversaire est limité en puissance informatique (il ne peut pas forger des signatures numériques) mais, hormis cela, il n'est pas limité en matière informationnelle (il connaît tout de l'état) et il est adaptatif en matière de tours (il peut modifier sa stratégie à tout moment).
+
+\paragraph{Attaques Sybil}
+%O Consensus protocols provide their guarantees based on assumptions that only a fraction of participants are adversarial.
+Les protocoles de consensus fournissent leurs garanties fondées sur la supposition que seule une fraction des participants sont malveillants.
+%O These bounds could be violated if the network is naively left open to arbitrary participants.
+Ces limites peuvent être violées si le réseau est naïvement laissé ouvert à des participants quelconques.
+%O In particular, a Sybil attack~\cite{douceur2002sybil}, wherein a large number of identities are generated by an adversary, could be used to exceed the bounds.
+En particulier, une attaque Sybil~\cite{douceur2002sybil}, où un grand nombre d'identités sont générées par un adversaire, peut être employée pour dépasser ces limites.
+
+%O A long line of work, including PBFT~\cite{castro1999practical}, treats the Sybil problem separately from consensus, and rightfully so, as Sybil control mechanisms are distinct from the underlying, more complex agreement protocol\footnote{This is not to imply that every consensus protocol can be coupled/decoupled with every Sybil control mechanism.}.
+Une longue suite de travaux, comme PBFT~\cite{castro1999practical}, sépare le problème de Sybil de celui du consensus, à juste titre puisque les mécanismes de contrôle de Sybil sont distincts du protocole d'accord plus complexe sous-jacent\footnote{Cela n'implique pas que tout protocole de consensus puisse être couplé/découplé de tout mécanisme de contrôle de Sybil}.
+%O In fact, to our knowledge, only Nakamoto-style consensus has ``baked-in'' Sybil prevention as part of its consensus, made possible by chained proof-of-work~\cite{aspnes2005exposing}, which requires miners to continuously stake a hardware investment.
+En fait, à notre connaissance, le consensus Nakamoto est le seul à avoir intégré la protection anti-Sybil dans son consensus, chose rendue possible par la preuve de travail chaînée~\cite{aspnes2005exposing} qui demande aux mineurs de continuellement mettre en jeu un investissement matériel.
+%O Other protocols, discussed in Section~\ref{sec:related-work}, rely on proof-of-stake (by economic argument) or proof-of-authority (by administrative argument that makes the system ``permissioned'').
+D'autres protocoles, dont il est question en section~\ref{sec:related-work}, se basent sur la preuve d'enjeu (\emph{proof of stake} ou PoS, par un argument économique) ou preuve d'autorité (\emph{proof of authority} ou PoA, par un argument administratif) qui rend le système \guillemotleft\,permissionné\,\guillemotright.
+%O The consensus protocols presented in this paper can adopt any
+%O Sybil control mechanism, although proof-of-stake is most aligned with their quiescent operation.
+Les protocoles de consensus présentés dans cet article peuvent adopter n'importe quel mécanisme de contrôle de Sybil, bien que la preuve d'enjeu soit la plus en ligne avec le caractère économe de leur opération.
+%O One can use an already established proof-of-stake based mechanism~\cite{GiladHMVZ17}.
+On peut utiliser un mécanisme en preuve d'enjeu déjà établi.
+%O The full deployment of an autonomous P2P payment system incorporating staking mechanism is beyond the scope of this paper, whose focus is on a novel design paradigm of the core consensus algorithm.
+Le déploiement complet d'un système de paiement autonome en P2P incorporant des mécanismes d'enjeu dépasse le cadre de cet article, dont l'objet est un nouveau paradigme de conception de l'algorithme de consensus.
 
@@
 
-\paragraph{Flooding Attacks}
-Flooding/spam attacks are a problem for any distributed system. 
-Without a protection mechanism, an attacker can generate large numbers of transactions and flood protocol data structures, consuming storage.
+%O \paragraph{Flooding Attacks}
+\paragraph{Attaques par inondation}
+%O Flooding/spam attacks are a problem for any distributed system. 
+Les attaques par inondation (\emph{flooding}) et spam représentent un problème pour tout système distribué.
+%O Without a protection mechanism, an attacker can generate large numbers of transactions and flood protocol data structures, consuming storage.
+Sans mécanisme de protection, un attaquant peut générer un grand nombre de transactions et inonder les structures de données du protocole, consommant ainsi des ressources de stockage.
 % Although there is no proper, protocol-level methodology to prevent such an attack, there are very effective heuristics that we can employ.
-There are a multitude of techniques to deter such attacks, including network-layer protection, proof-of-authority, local proof-of-work and economic mechanisms. 
-In Avalanche, we use transaction fees, making such attacks costly even if the attacker is sending money back to addresses under its control.
-
-\paragraph{Additional Assumptions}
-We do not assume that all members of the network are known to all participants, but rather may temporarily have some discrepancies in network view.
-We quantify the bounds on the discrepancy in Appendix~\ref{sec:full-analysis-churn}.
-We assume a safe bootstrapping mechanism, similar to that of Bitcoin, that enables a node to connect with sufficiently many correct nodes to acquire a statistically unbiased view of the network.
-We do not assume a PKI\@.
-Finally, we make standard cryptographic assumptions related to digital signatures and hash functions.
+%O There are a multitude of techniques to deter such attacks, including network-layer protection, proof-of-authority, local proof-of-work and economic mechanisms. 
+Il existe une multitude de techniques pour repousser ces attaques comme la protection de la couche réseau, la preuve d'autorité, la preuve de travail local et des mécanismes économiques.
+%O In Avalanche, we use transaction fees, making such attacks costly even if the attacker is sending money back to addresses under its control.
+Dans Avalanche, nous utilisons les frais de transaction, rendant ainsi ces attaques coûteuses même si l'attaquant renvoie l'argent aux adresses qu'il contrôle.
+
+%O \paragraph{Additional Assumptions}
+\paragraph{Hypothèses supplémentaires}
+%O We do not assume that all members of the network are known to all participants, but rather may temporarily have some discrepancies in network view.
+Nous ne supposons pas que tous les membres du réseau sont connus de tous les participants, car la vue du réseau qu'ont ces derniers peut souffrir d'incohérences temporaires.
+%O We quantify the bounds on the discrepancy in Appendix~\ref{sec:full-analysis-churn}.
+Nous quantifions les limites en Annexe~\ref{sec:full-analysis-churn}.
+%O We assume a safe bootstrapping mechanism, similar to that of Bitcoin, that enables a node to connect with sufficiently many correct nodes to acquire a statistically unbiased view of the network.
+Nous supposons un mécanisme de démarrage initial sûr, similaire à celui de Bitcoin, qui permet à un nœud de se connecter à suffisamment de nœuds corrects pour acquérir une vue statistiquement non biaisée du réseau.
+%O We do not assume a PKI\@.
+Nous ne prévoyons pas une PKI\@.
+%O Finally, we make standard cryptographic assumptions related to digital signatures and hash functions.
+Enfin, nous nous reposons sur les hypothèses standard en matière de cryptographie dans le domaine des signatures numériques et des fonctions de hachage.
 %\editinsert{TODO: Talk about synchrony assumption, and justify it by mentioning Bitcoin.}

### 2020-11-07 avalanche_fr.tex
S5 - début DAG, création de avalanche_fr.tex

--- 2020-11-03 avalanche.tex
@@
 
-Avalanche consists of multiple single-decree Snowball instances instantiated as a multi-decree protocol that
-maintains a dynamic, append-only directed acyclic graph (DAG) of all known transactions.
-The DAG has a single sink that is the \emph{genesis vertex}.
-Maintaining a DAG provides two significant benefits.
-First, it improves efficiency, because a single vote on a DAG vertex implicitly votes for all transactions on the path to the genesis vertex.
-Second, it also improves security, because the DAG intertwines the fate of transactions, similar to the Bitcoin blockchain.
-This renders past decisions difficult to undo without the approval of correct nodes.
+%O Avalanche consists of multiple single-decree Snowball instances instantiated as a multi-decree protocol that
+%O maintains a dynamic, append-only directed acyclic graph (DAG) of all known transactions.
+Avalanche consiste en plusieurs instance Snowball à décret unique instanciées en tant que protocole multi-décrets qui maintient un graphe orienté acyclique (\emph{Directed Acyclic Graph},  DAG) dynamique en ajout seul de toutes les transactions connues.
+%O The DAG has a single sink that is the \emph{genesis vertex}.
+Le DAG a un nœud terminal (\emph{sink}) unique qui est le \emph{genesis vertex}, le sommet genèse. % HOULA sink ? 
+%O Maintaining a DAG provides two significant benefits.
+L'emploi d'un DAG procure deux avantages :
+%O First, it improves efficiency, because a single vote on a DAG vertex implicitly votes for all tran
+%O sactions on the path to the genesis vertex.
+D'abord, il est plus efficient car un vote sur un sommet vote implicitement pour toutes les transactions sur le chemin vers le sommet genèse.
+%O Second, it also improves security, because the DAG intertwines the fate of transactions, similar to the Bitcoin blockchain.
+Ensuite, il améliore également la sécurité parce que le entrelace le sort des transactions, similaire en cela à la blockchain de Bitcoin.
+%O This renders past decisions difficult to undo without the approval of correct nodes.
+Cela rend les décisions passées difficiles à défaire sans l'approbation des nœuds corrects.
 
@@
 % The DAG is rooted at a protocol-defined, well-known \emph{genesis} vertex.
-When a client creates a transaction, it names one or more \emph{parents}, which are included inseparably in the transaction and form the edges of the DAG\@.
-The parent-child relationships encoded in the DAG may, but do not need to, correspond to application-specific dependencies; for instance, a child transaction need not spend or have any relationship with the funds received in the parent transaction.
+%O When a client creates a transaction, it names one or more \emph{parents}, which are included inseparably in the transaction and form the edges of the DAG\@.
+Quand un client crée une transaction, il nomme un ou plusieurs \emph{parents}, et ceux-ci sont inclus de façon inséparable dans la transaction et forment les bords du DAG\@.
+%O The parent-child relationships encoded in the DAG may, but do not need to, correspond to application-specific dependencies; for instance, a child transaction need not spend or have any relationship with the funds received in the parent transaction.
+Les relations parent-enfant codées dans le DAG peuvent, sans obligation, correspondre à des dépendances spécifiques aux applications ; par exemple, la transaction d'un enfant ne dépense pas nécessairement, ni n'a forcément de relation avec les fonds reçus lors de la transaction parent.
 % These additional relationships entangle the fate of previous decisions made by the system.
-We use the term \emph{ancestor set} to refer to all transactions reachable via parent edges back in history, and \emph{progeny} to refer to all children transactions and their offspring.
+%O We use the term \emph{ancestor set} to refer to all transactions reachable via parent edges back in history, and \emph{progeny} to refer to all children transactions and their offspring.
+Nous utilisons le terme \emph{ensemble ancêtre} pour nous référer à toutes les transactions accessibles par les arêtes parent historisées, et \emph{progéniture} pour nous référer à toutes les transactions enfant et leur descendance.
 
-The central challenge in the maintenance of the DAG is to choose among \emph{conflicting transactions}.
-The notion of conflict is application-defined and transitive, forming an equivalence relation.
-In our cryptocurrency application, transactions that spend the same funds (\emph{double-spends}) conflict, and form a \emph{conflict set}
+%O The central challenge in the maintenance of the DAG is to choose among \emph{conflicting transactions}.
+Le principal défi de la maintenance du DAG est de choisir parmi des \emph{trasactions conflictuelles}.
+%O The notion of conflict is application-defined and transitive, forming an equivalence relation.
+La notion de conflit est définie par l'application et est transitive, formant une relation d'équivalence.
+%O In our cryptocurrency application, transactions that spend the same funds (\emph{double-spends}) conflict, and form a \emph{conflict set}
+
 %(Figure~\tronly{\ref{fig:dag-conflict-set}}{\ref{fig:dag-cd}})

### 2020-11-07 paper_fr.tex
S5 - début DAG, création de avalanche_fr.tex

@@
 Dans les faits, Avalanche exécute plusieurs instances de Snowball (l'un des protocoles de la famille Snow) avec l'aide d'un graphe orienté acyclique (\emph{Directed Acyclic Graph} ou DAG).
-Le DAG sert à empiler plusieurs instances en réduisant le coût de $\Oh{\log{n}}$ à $\Oh{1}$ par nœud et en optimisant le chemin où il n'existe pas de transactions contradictoires.
+Le DAG sert à empiler plusieurs instances en réduisant le coût de $\Oh{\log{n}}$ à $\Oh{1}$ par nœud et en optimisant le chemin où il n'existe pas de transactions conflictuelles.
 %This combination of the best features of traditional and Nakamoto consensus involves one significant tradeoff: liveness for \editchange{equivocating proposals}. %conflicting transactions.
@@
 %O     \item \textbf{P1. Safety.} When decisions are made by any two correct nodes, they decide on conflicting transactions with negligible probability ($\leq \varepsilon$).
-     \item \textbf{P1. Sûreté.} Quand les décisions sont prises par deux nœuds corrects, ils décident des transactions contradictoires avec une probabilité négligeable ($\leq \varepsilon$).
+     \item \textbf{P1. Sûreté.} Quand les décisions sont prises par deux nœuds corrects, ils décident de transactions conflictuelles avec une probabilité négligeable ($\leq \varepsilon$).
     % \item \textbf{P2.
@@
 
-Using Snowball consensus, we have implemented a bare-bones payment system, Avalanche, which supports Bitcoin transactions.
+%O Using Snowball consensus, we have implemented a bare-bones payment system, Avalanche, which supports Bitcoin transactions.
 In this section, we describe the design and sketch how the implementation can support the value transfer primitive at the center of cryptocurrencies.
-Deploying a full cryptocurrency involves bootstrapping, minting, staking, unstaking,
-and inflation control. While we have solutions for these issues, their full
-discussion is beyond the scope of this paper, whose focus is centered on the
-novel Snow consensus protocol family.
+Grâce au consensus Snowball, nous avons implémenté un système de paiement simpliste, Avalanche, qui supporte les transactions Bitcoin.
+Dans cette section, nous en décrivons la conception et schématison la manière donc l'implémentation peut supporter la primitive de transfert de valeur au centre des cryptomonnaies.
+%O Deploying a full cryptocurrency involves bootstrapping, minting, staking, unstaking,
+%O and inflation control. While we have solutions for these issues, their full
+%O discussion is beyond the scope of this paper, whose focus is centered on the
+%O novel Snow consensus protocol family.
+Le déploiement complet d'une cryptomonnaie implique l'initialisation, l'émission, le dépôt, le retrait et le contrôle de l'inflation.
+Nous disposons de solutions à ces problèmes mais leur description détaillée sort du cadre de cet article, dont l'objet est la nouvelle famille de consensus Snow.
+% HOULA staking unstaking
 %In this section,
@@
 %Second,
-In a cryptocurrency setting, cryptographic signatures enforce that only a key owner is able to create a transaction that spends a particular coin.
+%O In a cryptocurrency setting, cryptographic signatures enforce that only a key owner is able to create a transaction that spends a particular coin.
 Since correct clients follow the protocol as prescribed and never double spend coins, in Avalanche, they are guaranteed both safety and liveness for their \emph{virtuous} transactions.
@@
 Such decisions may stall in the network, but have no safety impact on virtuous transactions.
-We show that this is a sensible tradeoff, and that the resulting system is sufficient for building complex payment systems.
-
-\input{avalanche}
+Dans une configuration de cryptomonnaie, les signatures cryptographiques imposent que seul le possesseur de la clef soit capable de créer une transaction dépensant un dépôt donné.
+Comme les clients corrects suivent le protocole tel qu'il est prescrit et ne peuvent pratiquer la double dépense, ils ont la garantie tant de sûreté de fonctionnement que de vitalité pour leurs transactions \emph{vertueuses}.
+Par contraste, la vitalité n'est pas garantie pour les transactions \emph{malveillantes} soumises par des clients byzantins, qui entrent mutuellement en conflit.
+Ces décisions peuvent arrêter le réseau mais n'ont pas d'impact sur les transactions vertueuses.
+%O We show that this is a sensible tradeoff, and that the resulting system is sufficient for building complex payment systems.
+Nous montrons que ce compromis est sensé et que le système qui en résulte suffit à bâtir des systèmes de paiement complexes.
+
+\input{avalanche_fr}
 

### 2020-11-07 sections/related-work_fr.tex
Création de related-work_fr.tex

\label{section:background}
\tronly{
Bitcoin~\cite{nakamoto2008bitcoin} is a cryptocurrency that uses a blockchain based on
proof-of-work (PoW) to maintain a ledger of UTXO transactions. 
While techniques based on proof-of-work~\cite{DworkN92, aspnes2005exposing}, and even cryptocurrencies with minting based on proof-of-work~\cite{vishnumurthy2003karma,rivest1997payword}, have been explored before, Bitcoin was the first to incorporate PoW into its consensus process.
Unlike more traditional BFT protocols, Bitcoin has a probabilistic safety guarantee
and assumes honest majority computational power rather than a known
membership, which in turn has enabled an internet-scale permissionless protocol. While permissionless and resilient to adversaries,
Bitcoin suffers from low throughput (\textasciitilde{}3 tps) and
high latency (\textasciitilde{}5.6 hours for a network with 20\% Byzantine presence and $2^{-32}$ security guarantee).  Furthermore, PoW requires a substantial amount of computational power that is consumed only for the purpose of maintaining
safety.

% PoW family
Countless cryptocurrencies use PoW~\cite{DworkN92, aspnes2005exposing} to maintain a distributed ledger. 
Like Bitcoin, they suffer from inherent scalability bottlenecks. 
Several proposals for protocols exist that try to
better utilize the effort made by PoW.
Bitcoin-NG~\cite{EyalGSR16} and the permissionless version of
Thunderella~\cite{PassS18} use Nakamoto-like consensus to elect a leader that
dictates writing of the replicated log for a relatively long time so as to
provide higher throughput. Moreover, Thunderella provides an
optimistic bound that, with 3/4 honest computational power and an honest
elected leader, allows transactions to be confirmed rapidly.
ByzCoin~\cite{Kokoris-KogiasJ16} periodically selects a small set of
participants and then runs a PBFT-like protocol within the selected nodes.
%It achieves a throughput of 942 tps with about 35 second latency.

% Byzantine Agreement family
Protocols based on Byzantine agreement~\cite{PeaseSL80, LamportSP82} typically make use of
quorums and require precise knowledge of membership.
PBFT~\cite{castro1999practical, CL02}, a well-known representative, requires a quadratic number of message exchanges in order to
reach agreement. 
%Some variants are able to scale to dozens of nodes~\cite{ClementWADM09, KotlaADCW09}.
The Q/U protocol~\cite{abd2005fault} and HQ replication~\cite{cowling2006hq} use a quorum-based approach to optimize for contention-free cases of operation to achieve consensus in only a single round of communication. However, although these protocols improve on performance, they degrade very poorly under contention. Zyzzyva~\cite{KotlaADCW09} couples BFT with speculative execution to improve the failure-free operation case.
%Aliph~\cite{guerraoui2010next} introduces a protocol with optimized performance under several, rather than just one, cases of execution. In contrast, Ardvark~\cite{ClementWADM09} sacrifices some performance to tolerate worst-case degradation, providing a more uniform execution profile. This work, in particular, sacrifices failure-free optimizations to provide consistent throughput even at high number of failures. 
Past work in permissioned BFT systems typically requires at least $3f+1$ replicas. CheapBFT~\cite{kapitza2012cheapbft} leverages trusted hardware components to construct a protocol that uses $f+1$ replicas.

Other work attempts to introduce new protocols under redefinitions and relaxations of the BFT model. 
Large-scale BFT~\cite{rodrigues2007large} modifies PBFT to allow for arbitrary choice of number of replicas and failure threshold, providing a probabilistic guarantee of liveness for some failure ratio but protecting safety with high probability. 
In another form of relaxation. Zeno~\cite{singh2009zeno} introduces a BFT state machine replication protocol that trades consistency for high availability. More specifically, Zeno guarantees eventual consistency rather than linearizability, meaning that participants can be inconsistent but eventually agree once the network stabilizes. By providing an even weaker consistency guarantee, namely fork-join-causal consistency, Depot~\cite{mahajan2011depot} describes a protocol that guarantees safety under $2f+1$ replicas.

NOW~\cite{guerraoui2013highly} uses sub-quorums to drive smaller instances of consensus. The insight of this paper is that small, logarithmic-sized quorums can be extracted from a potentially large set of nodes in the network, allowing smaller instances of consensus protocols to be run in parallel.

Snow White~\cite{cryptoeprint:2016:919} and
Ouroboros~\cite{KiayiasRDO17} are some of the earliest provably secure PoS
protocols.  Ouroboros uses a secure multiparty coin-flipping protocol to
produce randomness for leader election. The follow-up protocol, Ouroboros
Praos~\cite{DavidGKR18} provides safety in the presence of fully adaptive
adversaries.
HoneyBadger~\cite{MillerXCSS16} provides good liveness in a network with heterogeneous latencies. %and achieves over 819 tps with 5 minute latency on 104 nodes.

Tendermint~\cite{buchman2016tendermint, 1807.04938} rotates the leader for each block
and has been demonstrated with as many as 64 nodes. Ripple~\cite{schwartz2014ripple} has low latency by utilizing collectively-trusted
sub-networks in a large network. The Ripple company provides a
slow-changing default list of trusted nodes, which renders the system essentially centralized.
HotStuff~\cite{hotstuff,hotstuffpodc} improves the communication cost from quadratic to linear and significantly simplifies the protocol specification, although the leader bottleneck still persists. Facebook uses HotStuff as the core consensus for its Libra project.
% Ittai paper
In the synchronous setting, inspired by HotStuff, Sync HotStuff~\cite{synchotstuff} achieves consensus in $2\Delta$ time with quadratic cost and unlike other lock-steped synchronous protocols, it operates as fast as network propagates.
Stellar~\cite{mazieres2015stellar} uses Federated Byzantine Agreement in which \emph{quorum slices}
enable heterogeneous trust for different nodes.  Safety is guaranteed when
transactions can be transitively connected by trusted quorum slices.
Algorand~\cite{GiladHMVZ17} uses a verifiable random function to select a
committee of nodes that participate in a novel Byzantine consensus
protocol.
%It achieves over 874 tps with 50 second latency on
%an emulated network of 2000 committee nodes (500K users in total) distributed among 20 cities. To prevent Sybil
%attacks, it uses a mechanism like proof-of-stake that assigns weights to participants in committee selection based on the money in their accounts.

Some protocols use a Directed Acyclic Graph (DAG) structure instead of a linear chain to achieve
consensus~\cite{SompolinskyZ15,SompolinskyLZ16,SompolinskyZ18,BentovHMN17,baird2016hashgraph}.
Instead of choosing the longest chain as in Bitcoin,
GHOST~\cite{SompolinskyZ15} uses a more efficient chain selection rule that
allows transactions not on the main chain to be taken into consideration, increasing efficiency.
SPECTRE~\cite{SompolinskyLZ16} uses transactions on
the DAG to vote recursively with PoW to achieve consensus, followed up by
PHANTOM~\cite{SompolinskyZ18} that achieves a linear order among all blocks.
Like PHANTOM, Conflux also finalizes a linear order of transactions by PoW
in a DAG structure, with better resistance to liveness attack~\cite{confluxLLXLC18}.
Avalanche
is different in that the voting result is a one-time chit that is determined by
a query without PoW, while the votes in PHANTOM or Conflux are purely determined by PoW in transaction structure.
Similar to Thunderella, Meshcash~\cite{BentovHMN17} combines a slow PoW-based protocol with a fast consensus protocol that allows a high block rate regardless of network latency, offering fast confirmation time.
Hashgraph~\cite{baird2016hashgraph} is a leader-less protocol that builds a DAG via randomized gossip.
It requires full membership knowledge at all times, and, similar to the Ben-Or~\cite{ben1983another}, it requires exponential messages~\cite{aspnes2003randomized,CachinV17} in expectation.
}{
% PoW family
Several proposals for protocols exist that try to
better utilize the effort made by PoW.
Bitcoin-NG~\cite{EyalGSR16} and the permissionless version of
Thunderella~\cite{PassS18} use Nakamoto consensus to elect a leader that
dictates writing of the replicated log for a relatively long time so as to
provide higher throughput. Moreover, Thunderella provides an
optimistic bound that, with 3/4 honest computational power and an honest
elected leader, allows transactions to be confirmed rapidly.
ByzCoin~\cite{Kokoris-KogiasJ16} periodically selects a small set of
participants and then runs a PBFT-like protocol within the selected nodes. 
%It achieves a throughput of 942 tps with about 35 second latency.

% Byzantine Agreement family
Protocols based on Byzantine agreement~\cite{PeaseSL80, LamportSP82} typically make use of
quorums and require precise knowledge of membership.
PBFT~\cite{castro1999practical, CL02}, a well-known representative, requires a quadratic number of message exchanges in order to
reach agreement. Some variants are able to scale to dozens of nodes~\cite{ClementWADM09, KotlaADCW09}.
% However, due to the quadratic message complexity, many variants based on PBFT also inherit this scalability issue.
Snow White~\cite{cryptoeprint:2016:919} and
Ouroboros~\cite{KiayiasRDO17} are some of the earliest provably secure PoS
protocols.  Ouroboros uses a secure multiparty coin-flipping protocol to
produce randomness for leader election. The follow-up protocol, Ouroboros
Praos~\cite{DavidGKR18} provides safety in the presence of fully adaptive
adversaries.
HoneyBadger~\cite{MillerXCSS16} provides good liveness in a network with heterogeneous latencies. %and achieves over 819 tps with 5 minute latency on 104 nodes.
Tendermint~\cite{buchman2016tendermint, 1807.04938} rotates the leader for each block
and has been demonstrated with as many as 64 nodes. Ripple~\cite{schwartz2014ripple} has low latency by utilizing collectively-trusted
sub-networks in a large network. The Ripple company provides a
slow-changing default list of trusted nodes, which renders the system essentially centralized.
HotStuff~\cite{hotstuff,hotstuffpodc} improves the communication cost from quadratic to linear and significantly simplifies the protocol specification, although the leader bottleneck still persists. Facebook uses HotStuff as the core consensus~\cite{librabft} for the Libra project.
% Ittai paper
In the synchronous setting, inspired by HotStuff, Sync HotStuff~\cite{synchotstuff} achieves consensus in $2\Delta$ time with quadratic cost and unlike other lock-steped synchronous protocols, it operates as fast as network propagates.
Stellar~\cite{mazieres2015stellar} uses Federated Byzantine Agreement in which \emph{quorum slices}
enable heterogeneous trust for different nodes.  Safety is guaranteed when
transactions can be transitively connected by trusted quorum slices.

Some protocols use a Directed Acyclic Graph (DAG) structure instead of a linear chain to achieve
consensus~\cite{SompolinskyZ15,SompolinskyLZ16,SompolinskyZ18,BentovHMN17,baird2016hashgraph}.
Instead of choosing the longest chain as in Bitcoin,
GHOST~\cite{SompolinskyZ15} uses a more efficient chain selection rule that
allows transactions not on the main chain to be taken into consideration, increasing efficiency.
SPECTRE~\cite{SompolinskyLZ16} uses transactions on
the DAG to vote recursively with PoW to achieve consensus, followed up by
PHANTOM~\cite{SompolinskyZ18} that achieves a linear order among all blocks.
Avalanche
is different in that the voting result is a one-time chit that is determined by
a query without PoW, while the votes in PHANTOM are purely determined by transaction structure.
Similar to Thunderella, Meshcash~\cite{BentovHMN17} combines a slow PoW-based protocol with a fast consensus protocol that allows a high block rate regardless of network latency, offering fast confirmation time.
Hashgraph~\cite{baird2016hashgraph} is a leader-less protocol that builds a DAG via randomized gossip. 
It requires full membership knowledge at all times, and, similar to the Ben-Or~\cite{ben1983another}, it requires exponential messages~\cite{aspnes2003randomized,CachinV17} in expectation.
}

\begin{comment}
Moreover, in Theorem 5.16, the author mentions if
there is no supermajority virtual vote, all honest nodes will randomly choose
their votes for the next round. Then with non-zero probability, they could
choose the same vote.  The rounds are repeated until all honest nodes
eventually reach the same vote by chance.  This means as the total number of
nodes increases in Hashgraph, the probability of reaching the same vote in a
single trial drops exponentially due to the random voting~\cite{CachinV17}.
\end{comment}
%The paper does not discuss how many rounds of voting are needed in expectation,
%nor\ \ have we been able to obtain open-source code to evaluate Hashgraph.

%\tronly{
%Instead of voting blindly and relying on randomness
%for each agreement, Avalanche creates a positive feedback where votes are
%collected in a separate k-query process determined by current preference
%established locally on each node over time.  We have demonstrated that
%Avalanche is scalable to 2000 nodes distributed world-wide.
%
%The central component of Avalanche is a
%repeated k-query, which serves both as a probabilistic
%broadcast and as a basis for Byzantine agreement.
%As an optimization, these repeated queries are linked
%together via an artificial DAG structure that amortizes the repeated cost of
%querying.  Avalanche does not require overlapping quorums and
%achieves probabilistic consensus with
%$O(kN)$ overall message complexity instead of quadratic
%while allowing for a degree of network churn.
%}

% XXX ADD SAM TOUEG https://zoo.cs.yale.edu/classes/cs426/2017/bib/bracha85asynchronous.pdf

### 2020-11-07 paper_fr.tex
Création de related-work_fr.tex
French translation of sections Snowflake, Snowball, and Analysis

@@
 
-Snowflake augments Slush with a single counter that captures the strength of a node's conviction in its current color.
-This per-node counter stores how many consecutive samples of the network by that node have all yielded the same color.
-A node accepts the current color when its counter reaches $\beta$, another security parameter.
-Figure~\ref{fig:snowflake-loop} shows the amended protocol, which includes
-the following modifications:
+Snowflake améliore Slush en ajoutant un compteur unique qui capture la force de conviction d'un noeud dans sa couleur
+actuelle. Ce compteur associé à chaque noeud stocke le nombre d'échantillons du réseau consécutifs que le noeud a
+sollicité et qui ont retourné la même couleur. Un noeud accepte sa couleur actuelle lorsque son compteur atteint $\beta$,
+un autre paramètre de sécurité. La Figure~\ref{fig:snowflake-loop} montre le protocole amélioré qui inclue les
+modifications suivantes:
+
+%O Snowflake augments Slush with a single counter that captures the strength of a node's conviction in its current color.
+%O This per-node counter stores how many consecutive samples of the network by that node have all yielded the same color.
+%O A node accepts the current color when its counter reaches $\beta$, another security parameter.
+%O Figure~\ref{fig:snowflake-loop} shows the amended protocol, which includes
+%O the following modifications:
 
 \begin{compactenum}
-	\item Each node maintains a counter $\mathit{cnt}$;
-    \item Upon every color change, the node resets $\mathit{cnt}$ to 0;
-    \item Upon every successful query that yields $\ge \alpha$ responses for the same color as the node, the node increments $\mathit{cnt}$.
+	\item Chaque node maintient un compteur $\mathit{cnt}$;
+    \item A chaque changement de couleur, le compteur $\mathit{cnt}$ se réinitialise à 0;
+    \item Après chaque requête réussie qui retourne $\ge \alpha$ réponses en faveur de la même couleur que le noeud,
+      ce dernier incrémente $\mathit{cnt}$.
 \end{compactenum}
+
+%O \begin{compactenum}
+%O 	\item Each node maintains a counter $\mathit{cnt}$;
+%O     \item Upon every color change, the node resets $\mathit{cnt}$ to 0;
+%O     \item Upon every successful query that yields $\ge \alpha$ responses for the same color as the node, the node increments $\mathit{cnt}$.
+%O \end{compactenum}
 
@@
 
-When the protocol is correctly parameterized for a given threshold of Byzantine nodes and a desired $\varepsilon$-guarantee, it can ensure both safety (P1) and liveness (P2, P3).
-As we later show, there exists an irreversible state after which a decision is inevitable.
+Quand le protocole est correctement parametré pour un seuil donné de noeuds byzantins et un niveau de garantie
+$\varepsilon$, il peut garantir à la fois la sécurité de fonctionnement (P1) et la vitalité (P2, P3).
+Comme nous le démontrons plus loin, il existe un état irréversible à partir duquel une décision est inévitable. Les noeuds
+légitimes commencent à se positionner après cet état irréversible et adoptent la même couleur. Une intuition
+supplémentaire, que nous ne détaillons pas dans ce papier, introduit l'existence d'un point de changement de phase à
+partir duquel les nodes byzantins perdent toute capacité à maintenir le réseau dans un état bivalent.
+
+%O When the protocol is correctly parameterized for a given threshold of Byzantine nodes and a desired $\varepsilon$-guarantee, it can ensure both safety (P1) and liveness (P2, P3).
+%O As we later show, there exists an irreversible state after which a decision is inevitable.
 Correct nodes begin to commit past the irreversible state to adopt the same color, whp.
 For additional intuition, which we do not expand in this paper, there also exists a phase-shift point, where the Byzantine nodes lose ability to keep network in a bivalent state.
-%there exists a phase-shift point after which correct nodes are more likely to tend towards a decision than a bivalent state.
-%Further, there exists a point-of-no-return after which a decision is inevitable.
-%The Byzantine nodes lose control past the phase shift, and the correct nodes begin to commit past the point-of-no-return, to adopt the same color, whp.
-
-\subsection{Snowball: Adding Confidence}
-
-Snowflake's notion of state is ephemeral: the counter gets reset with every color flip.
-% While, theoretically, the protocol is able to make strong guarantees with
-% minimal state, we will now improve the protocol to make it harder to
-% attack by incorporating a more permanent notion of belief.
-Snowball augments Snowflake with \emph{confidence counters} that capture the number of queries that have yielded a threshold result for their corresponding color (Figure ~\ref{fig:snowball-loop}).
-A node decides if it gets $\beta$ consecutive chits for a color. However, it only changes preference based on the total accrued confidence.
-The differences between Snowflake and Snowball are as follows:
+
+%O \subsection{Snowball: Adding Confidence}
+
+\subsection{Snowball: Introduire la notion de confiance}
+
+La notion d'état de Snowflake est éphémère: le compteur est réinitialisé à chaque changement de couleur.
+Snowball améliore Snowflake avec des \emph{compteurs de confiance} qui capturent le nombre de requêtes ayant retourné
+un résultat en faveur de la couleur du noeud au dessus d'un certain seuil (Figure ~\ref{fig:snowball-loop}).
+Un node prend une décision s'il obtient $\beta$ réponses consécutives en faveur d'une couleur. Cependant il change
+sa préférence uniquement sur la base de la confiance totale accumulée. Les différences entre Snowflake et Snowball
+sont:
+
 \begin{compactenum}
-    \item Upon every successful query, the node increments its confidence counter for that color.
-    \item A node switches colors when the confidence in its current color becomes lower than the confidence value of the new color.
+    \item Après chaque requête réussie, le noeud incrémente son compteur de confiance associé à la couleur retournée.
+    \item Un noeud change de couleur quand la confiance associée à sa couleur actuelle devient inférieure à la
+      valeur de la confiance associée à  la nouvelle couleur.
 \end{compactenum}
-% Snowball is not only harder to attack than Snowflake, but is more easily generalized to multi-decree protocols.
-
-\section{Analysis}
+
+%O Snowflake's notion of state is ephemeral: the counter gets reset with every color flip.
+%O Snowball augments Snowflake with \emph{confidence counters} that capture the number of queries that have yielded a threshold result for their corresponding color (Figure ~\ref{fig:snowball-loop}).
+%O A node decides if it gets $\beta$ consecutive chits for a color. However, it only changes preference based on the total accrued confidence.
+%O The differences between Snowflake and Snowball are as follows:
+%O \begin{compactenum}
+%O     \item Upon every successful query, the node increments its confidence counter for that color.
+%O     \item A node switches colors when the confidence in its current color becomes lower than the confidence value of the new color.
+%O \end{compactenum}
+
+\section{Analyse}
 \label{sec:analysis}
-Due to space limits, we move some core details to Appendix~\ref{sec:full-analysis}, where we show that under certain independent and distinct assumptions, the Snow family of consensus protocols provide safety (P1) and liveness (P2, P3) properties.
-In this section, we summarize our core results and provide some proof sketches.
-
-\paragraph{Notation} Let the network consist of a set of $n$ nodes (represented by set $\mathcal{N}$), where $c$ are correct nodes (represented by set $\mathcal{C}$) and $f$ are Byzantine nodes (represented by set $\mathcal{B}$).
-Let $u, v \in \mathcal{C}$ refer to any two correct nodes in the network.
+Pour des raisons de manque de place, nous déplaçons certains détails importants vers l'Annexe~\ref{sec:full-analysis},
+dans laquelle nous démontrons que sous certaines hypothèses distinctes et indépendantes, la famille de protocoles de
+consensus Snow possède des propriétés de sûreté de fonctionnement (P1) et de vitalité (P2, P3). Dans cette section,
+nous résumons nos résultats principaux en fournissant des ébauches de preuves.
+
+%O \section{Analysis}
+%O \label{sec:analysis}
+%O Due to space limits, we move some core details to Appendix~\ref{sec:full-analysis}, where we show that under certain independent and distinct assumptions, the Snow family of consensus protocols provide safety (P1) and liveness (P2, P3) properties.
+%O In this section, we summarize our core results and provide some proof sketches.
+
+\paragraph{Notation} En considérant que le réseau consiste d'un ensemble de $n$ noeuds (représentés par l'ensemble
+$\mathcal{N}$), où $c$ sont des noeuds légitimes (représentés par l'ensemble $\mathcal{C}$) et $f$ sont les noeuds
+byzantins (représentés par l'ensemble $\mathcal{B}$). Soient $u, v \in \mathcal{C}$ des références à deux noeuds
+légitimes pris au hasard dans le réseau. Soient $k, \alpha, \beta \in \mathbb{Z}^+$ des entiers positifs où
+$\alpha > \floor{k/2}$. A partir de maintenant, $k$ fera toujours référence à la taille des déchantillons du réseau,
+où $k \leq n$, et $\alpha$ sera le seuil de majorité requis pour considérer l'expérience de vote comme ``réussie''.
+En général, nous définissons $\mathcal{S}$ comme étant l'état (ou la configuration) du réseau à tout moment.
+
+%O \paragraph{Notation} Let the network consist of a set of $n$ nodes (represented by set $\mathcal{N}$), where $c$ are correct nodes (represented by set $\mathcal{C}$) and $f$ are Byzantine nodes (represented by set $\mathcal{B}$).
+%O Let $u, v \in \mathcal{C}$ refer to any two correct nodes in the network.
 Let $k, \alpha, \beta \in \mathbb{Z}^+$ be positive integers where $\alpha > \floor{k/2}$.
@@
 In general, we will refer to $\mathcal{S}$ as the state (or configuration) of the network at any given time.
-
-\paragraph{Modelling Framework} To formally model our protocols, we use continuous-time Markov processes (CTMC). 
-The state space is enumerable (and finite), and state transitions occur in continuous time. 
-CTMCs naturally model our protocols since state transitions do not occur in epochs and in lockstep for every node (at the end of every time unit) but rather occur at any time and independently of each other.
-% CMTCs can also trivially encapsulate synchronous time protocols through the underlying jump chain, whose probabilities of transitions can be extracted from the underlying generator. 
-% This modelling framework, in fact, is the primary way by which stochastic processes that describe systems of interacting particles (epidemic networks, biology, physics, etc) are analyzed. 
-% Analyzing the protocols introduced in this paper
-%differ from classical as well as Nakamoto consensus-style protocols. 
-%Even simple, non-linear dynamical systems often incur complicated analyses as they may not admit analytical solutions.
-% In particular, since the state of a node depends on randomly sampling other nodes, possibly inducing nondeterministic dependency cycles, analyzing all possible state executions is computationally infeasible.
-% Here, we employ stochastic techniques used in the analysis of epidemic networks, which may also be of independent interest to the distributed systems community. 
-%Here, we employ 
-% requires complex stochastic techniques to reason about the state of the network in aggregate.
-
-We focus on binary consensus, although the safety results generalize to more than two values.
+%O
+
+\paragraph{Framework de modélisation} Afin de modéliser nos protocoles de manière formelle, nous utilisons des
+processus Markov à temps continu (CTMC). L'espace contenant les états est dénombrable (et fini), et les transitions
+entre états s'opèrent sur un espace de temps continu. Les CTMCs proposent une modélisation naturelle de nos protocoles
+sachant que les transitions entre états ne se font pas par étapes et en marche ordonnée pour chaque noeud (à la fin de
+chaque unité de temps) mais arrivent plutôt à n'importe quel moment et indépendamment des autres noeuds.
+
+Nous nous concentrons sur un consensus binaire, même si les résultats en term de sûreté de fonctionnement se
+généralisent sur plus de deux valeurs. Nous pouvons imaginer le réseau comme un ensemble de noeuds colorés soit en
+rouge soit en bleu, et nous ferons référence à cette configuration à l'instant $t$ comme étant $\mathcal{S}_t$.
+Nous modélisons nos protocoles à travers un processus à temps continu contenant deux états, dans lequel tous les noeuds
+sont rouges ou tous les noeuds sont bleus. L'espace $\mathcal{S}$ des états du processus stochastique est une version
+condensée de l'espace de configuration complet, où chaque état $\{0, \dots, n\}$ représente le nombre total de noeuds
+bleus dans le système.
+
+La simplification qui nous permet d'analyser ce système consiste à s'affranchir du besoin de garder un historique de
+tous les chemins d'exécution, ainsi que toutes les stratégies d'attaque possibles, et de se concentrer plutôt
+entièrement sur un unique état, sans considérer la manière d'arriver à cet état. Plus spécifiquement, l'idée principale
+à retenir de notre analyse est l'identification d'un état \textit{d'irréversibilité} du système, cet état à partir
+duquel un tel nombre de noeuds légitimes a adopté soit la couleur rouge soit la couleur bleue que la possibilité de
+revenir à la couleur minoritaire est très peu probable.
+
+%O \paragraph{Modelling Framework} To formally model our protocols, we use continuous-time Markov processes (CTMC).
+%O The state space is enumerable (and finite), and state transitions occur in continuous time.
+%O CTMCs naturally model our protocols since state transitions do not occur in epochs and in lockstep for every node (at the end of every time unit) but rather occur at any time and independently of each other.
+
+%O We focus on binary consensus, although the safety results generalize to more than two values.
 We can think of the network as a set of nodes either colored red or blue, and we will refer to this configuration at time $t$ as $\mathcal{S}_t$.
-We model our protocols through a continuous-time process with two absorbing states, where either all nodes are red or all nodes are blue. 
-The state space $\mathcal{S}$ of the stochastic process is a condensed version of the full configuration space, where each state $\{0, \dots, n\}$ represents the total number of blue nodes in the system.
-% At any time, a set of nodes either change their color, or increase their confidence in their latest color. 
-% The dynamics of the system resemble those of epidemic networks, where nodes update their state based on the state of their neighbors, using a function chosen over some probability distribution. 
-% It would be infeasible to keep track of all possible execution paths, since at every round, the possible number of branching executions is equal to the number of possible correct nodes times all possible $k$-sample choices of outcomes that the node may sample.
-Furthermore, the difficulty of this task is greatly amplified due to the Byzantine adversary, whose optimal attack strategy is chosen over an unknown function, possibly itself nondeterministic.
-
-The simplification that allows us to analyze this system is to obviate the need to keep track of all of the execution paths, as well as all possible adversarial strategies, and rather focus entirely on a single state of interest, without regards to how we achieve this state.
-More specifically, the core extractable insight of our analysis is in identifying the \textit{irreversibility} state of the system, the state upon which so many correct nodes have usurped either red or blue that reverting back to the minority color is highly unlikely.
+%O We model our protocols through a continuous-time process with two absorbing states, where either all nodes are red or all nodes are blue.
+%O The state space $\mathcal{S}$ of the stochastic process is a condensed version of the full configuration space, where each state $\{0, \dots, n\}$ represents the total number of blue nodes in the system.
+
+%O The simplification that allows us to analyze this system is to obviate the need to keep track of all of the execution paths, as well as all possible adversarial strategies, and rather focus entirely on a single state of interest, without regards to how we achieve this state.
+%O More specifically, the core extractable insight of our analysis is in identifying the \textit{irreversibility} state of the system, the state upon which so many correct nodes have usurped either red or blue that reverting back to the minority color is highly unlikely.
 
@@
 \label{sec:related-work}
-\input{sections/related-work}
+\input{sections/related-work_fr}
 

### 2020-11-07 paper_fr.tex
French translation of the short 'Protocol Design' introduction

@@
 
-\section{Protocol Design}\label{sec:protocol}
-We start with a non-BFT protocol called Slush and progressively build up to Snowflake and Snowball, all based on the same common majority-based metastable voting mechanism.
-These protocols are single-decree consensus protocols of increasing robustness.
-We provide full specifications for the protocols in this section, and defer the analysis to the next section, and present formal proofs in the appendix.
+\section{Conception du Protocole}\label{sec:protocol}
+Nous partons d'un protocole non résistant au problème des généraux byzantins appelé Slush, en amenant
+progressivement Snowflake et Snowball, tous basés sur le même mécanisme métastable de vote basé sur une majorité.
+Ces derniers sont des protocoles de consensus à simple décret présentant plusieurs niveaux de robustesse. Nous
+détaillons les caractéristiques complètes de ces protocoles dans cette section, suivie de leur analyse en section
+suivante, les preuves formelles étant disponibles en annexe.
+
+%O \section{Protocol Design}\label{sec:protocol}
+%O We start with a non-BFT protocol called Slush and progressively build up to Snowflake and Snowball, all based on the same common majority-based metastable voting mechanism.
+%O These protocols are single-decree consensus protocols of increasing robustness.
+%O We provide full specifications for the protocols in this section, and defer the analysis to the next section, and present formal proofs in the appendix.
 

### 2020-11-07 avalanche_fr.tex
S5 - avalanche_fr.tex terminé

@@
-\subsection{Avalanche: Adding a DAG}%\tronly{}{\vspace{-0.5em}}
+\subsection{Avalanche : ajout d'un DAG}%\tronly{}{\vspace{-0.5em}}
 
@@
 %O In our cryptocurrency application, transactions that spend the same funds (\emph{double-spends}) conflict, and form a \emph{conflict set}
-
+Dans notre application de cryptomonnaie, les transactions qui dépensent les mêmes fonds (\emph{double-dépense}) entrent en conflit et forment un \emph{ensemble de conflit}
 %(Figure~\tronly{\ref{fig:dag-conflict-set}}{\ref{fig:dag-cd}})
-(shaded regions in Figure~\ref{fig:dag-cd}), out of which only a single one can be accepted.
-Note that the conflict set of a virtuous transaction is always a singleton.
+%O (shaded regions in Figure~\ref{fig:dag-cd}), out of which only a single one can be accepted.
+(régions ombrées en Figure~\ref{fig:dag-cd}), duquel une seule peut être acceptée.
+%O Note that the conflict set of a virtuous transaction is always a singleton.
+On note que l'ensemble de conflit d'une transaction vertueuse est toujours un singleton.
 %Note that the graph structure of transactions that spend depend on each other, also known as the UTXO graph, is completely independent of the DAG that Avalanche maintains.
@@
 
-Avalanche instantiates a Snowball instance for each conflict set.
-Whereas Snowball uses repeated queries and multiple counters to capture the amount of confidence built in conflicting transactions (colors),
-Avalanche takes advantage of the DAG structure and uses a transaction's progeny.
-Specifically, when a transaction $T$ is queried, all transactions reachable from $T$ by following the DAG edges are implicitly part of the query.
-A node will only respond positively to the query if $T$ and its entire ancestry are currently the preferred option in their respective conflict sets.
-If more than a threshold of responders vote positively, the transaction is said to collect a \emph{chit}.
-Nodes then compute their \emph{confidence} as the total number of chits in the progeny of that transaction.
-They query a transaction just once and rely on new vertices and possible chits, added to the progeny, to build up their confidence.
-Ties are broken by an initial preference for first-seen transactions.
-Note that chits are decoupled from the DAG structure, making the protocol immune to attacks where
-the attacker generates large, padded subgraphs.
+%O Avalanche instantiates a Snowball instance for each conflict set.
+Avalanche instancie une instance de Snowball pour chaque ensemble de conflit.
+%O Whereas Snowball uses repeated queries and multiple counters to capture the amount of confidence built in conflicting transactions (colors),
+%O Avalanche takes advantage of the DAG structure and uses a transaction's progeny.
+Alors que Snowball utilise des requêtes répétées et plusieurs compteurs pour capturer le degré de confiance associé aux transactions conflictuelles (couleurs),
+Avalanche profite de la structure du DAG et utilise la progéniture d'une transaction.
+%O Specifically, when a transaction $T$ is queried, all transactions reachable from $T$ by following the DAG edges are implicitly part of the query.
+En particulier, quand on interroge une transaction $T$, toutes les transactions accessible depuis $T$ en suivant les arêtes du DAG font implicitement partie de la requête.
+%O A node will only respond positively to the query if $T$ and its entire ancestry are currently the preferred option in their respective conflict sets.
+Un nœud ne répond positivement à la requête que si $T$ et toute son ascendance représente actuellement l'option préférentielle dans leurs ensembles de conflits respectifs.
+%O If more than a threshold of responders vote positively, the transaction is said to collect a \emph{chit}.
+Si un nombre de nœuds répondant positivement est supérieur au seuil, on dit que la transaction récolte un \emph{chit} (un bon).
+%O Nodes then compute their \emph{confidence} as the total number of chits in the progeny of that transaction.
+Les nœuds calculent ensuite leur \emph{confiance} en comptant le nombre de \emph{chits} dans la progéniture de cette transaction.
+%O They query a transaction just once and rely on new vertices and possible chits, added to the progeny, to build up their confidence.
+Ils n'envoient qu'une seule requête et se basent sur les nouveaux sommets les \emph{chits} possibles, ajoutés à la progéniture, pour fonder leur confiance.
+%O Ties are broken by an initial preference for first-seen transactions.
+Les liens sont brisés par une préférence initiale pour les transactions vues en premier.
+%O Note that chits are decoupled from the DAG structure, making the protocol immune to attacks where
+%O the attacker generates large, padded subgraphs.
+On note que les \emph{chits} sont découplés de la structure du DAG, rendant le protocol immune aux attaques où l'attaquant génère de grands sous-graphes pleins de vide.
 
-\subsection{Avalanche: Specification}%\tronly{}{\vspace{-0.5em}}
+%O \subsection{Avalanche: Specification}%\tronly{}{\vspace{-0.5em}}
+\subsection{Avalanche: Spécification}%\tronly{}{\vspace{-0.5em}}
 \label{subsection:specification}
@@
     \Procedure{init}{}
-        \State $\mathcal{T} \assign \emptyset$ \hspace{1ex}\textrm{// the set of known transactions}
-        \State $\mathcal{Q} \assign \emptyset$ \hspace{1ex}\textrm{// the set of queried transactions}
+%O        \State $\mathcal{T} \assign \emptyset$ \hspace{1ex}\textrm{// the set of known transactions}
+%O        \State $\mathcal{Q} \assign \emptyset$ \hspace{1ex}\textrm{// the set of queried transactions}
+        \State $\mathcal{T} \assign \emptyset$ \hspace{1ex}\textrm{// l'ensemble des transactions connues}
+        \State $\mathcal{Q} \assign \emptyset$ \hspace{1ex}\textrm{// l'ensemble des transactions interrogées}
     \EndProcedure
@@
     \EndProcedure
-    \captionof{figure}{Avalanche: transaction generation.}\label{fig:gossipchain-ongen}
+%O    \captionof{figure}{Avalanche: transaction generation.}\label{fig:gossipchain-ongen}
+    \captionof{figure}{Avalanche : génération des transactions.}\label{fig:gossipchain-ongen}
 \end{algorithmic}
@@
         \While {\codetrue}
-            \State\textrm{find  $T$ that satisfies }%\vspace*{-.6\baselineskip}
+%O            \State\textrm{find  $T$ that satisfies }%\vspace*{-.6\baselineskip}
+            \State\textrm{trouver  $T$ satisfaisant }%\vspace*{-.6\baselineskip}
             $T \in \mathcal{T} \land T \notin \mathcal{Q}$
@@
                 \State $c_T \assign 1$
-            \State\textrm{// update the preference for ancestors}
+%O            \State\textrm{// update the preference for ancestors}
+            \State\textrm{// mettre à jour les préférences pour les ancêtres}
             \For{$T' \in \mathcal{T} : T' \stackrel{*}{\gets} T$}
@@
             \EndIf
-            \State\textrm{// otherwise, }$c_T$\textrm{ remains 0 forever}
-            \State $\mathcal{Q} \assign \mathcal{Q} \cup \{T\}$ \hspace {1ex} \textrm{// mark T as queried}
+%O            \State\textrm{// otherwise, }$c_T$\textrm{ remains 0 forever}
+            \State\textrm{// sinon, }$c_T$\textrm{ reste éternellement à 0}
+%O            \State $\mathcal{Q} \assign \mathcal{Q} \cup \{T\}$ \hspace {1ex} \textrm{// mark T as queried}
+            \State $\mathcal{Q} \assign \mathcal{Q} \cup \{T\}$ \hspace {1ex} \textrm{// marquer T comme interrogé}
         \EndWhile
     \EndProcedure
-    \captionof{figure}{Avalanche: the main loop.}\label{fig:gossipchain-main}
+%O    \captionof{figure}{Avalanche: the main loop.}\label{fig:gossipchain-main}
+    \captionof{figure}{Avalanche : boucle principale.}\label{fig:gossipchain-main}
 \end{algorithmic}
@@
 
-Each correct node $u$
-keeps track of all transactions it has learned about in set $\mathcal{T}_u$,
-partitioned into mutually exclusive conflict sets $\mathcal{P}_T$, $T \in \mathcal{T}_u$.
+%O Each correct node $u$
+%O keeps track of all transactions it has learned about in set $\mathcal{T}_u$,
+%O partitioned into mutually exclusive conflict sets $\mathcal{P}_T$, $T \in \mathcal{T}_u$.
+Chaque nœud correct $u$ garde la trace de toutes les transactions dont il a été informé dans l'ensemble $\mathcal{T}_u$, partitionné en ensembles de conflits mutuellement exclusifs $\mathcal{P}_T$, $T \in \mathcal{T}_u$.
 % Transactions are determined to be conflicting based on a deterministic function known to every node.
-Since conflicts are transitive, if $T_i$ and $T_j$ are conflicting, then they belong to the same conflict set, i.e.
+%O Since conflicts are transitive, if $T_i$ and $T_j$ are conflicting, then they belong to the same conflict set, i.e.
 $\mathcal{P}_{T_i} = \mathcal{P}_{T_j}$.
 This relation may sound counter-intuitive: conflicting transitions have the \emph{equivalence} relation, because they are equivocations spending the \emph{same} funds.
+Comme les conflits sont transitifs, is $T_i$ et $T_j$ sont en conflit, ils appartiennent au même ensemble de conflit, c'est-à-dire $\mathcal{P}_{T_i} = \mathcal{P}_{T_j}$.
+Cette relation peut sembler contre-intuitive : les transactions en conflit ont la même relation d'\emph{équivalence} car ce sont des équivoques dépensant les \emph{mêmes} fonds.
 %\Jon{By $=$ surely you mean $\neq$?}
 
-We write $T' \gets T$ if $T$ has a parent edge to transaction $T'$,
-The ``$\stackrel{*}{\gets}$''-relation is its reflexive transitive closure, indicating a path from $T$ to $T'$.
-DAGs built by different nodes are guaranteed to be compatible, though at any one time, the two nodes may not have a complete view of all vertices in the system.
-Specifically, if $T' \gets T$, then every node in the system that has $T$ will also have $T'$ and the same relation $T' \gets T$; and conversely, if $T' \cancel{\gets} T$, then no nodes will end up with $T' \gets T$.
+%O We write $T' \gets T$ if $T$ has a parent edge to transaction $T'$,
+Nous écrivons $T' \gets T$ si $T$ a une arête commune avev la transaction $T'$.
+%O The ``$\stackrel{*}{\gets}$''-relation is its reflexive transitive closure, indicating a path from $T$ to $T'$.
+La relation ``$\stackrel{*}{\gets}$'' est sa fermeture réflexive transitive, indiquant un chemin de $T$ vers $T'$.
+%O DAGs built by different nodes are guaranteed to be compatible, though at any one time, the two nodes may not have a complete view of all vertices in the system.
+Les DAG construits par des nœuds différents offrent la garantie d'être compatibles, bien qu'à un moment donné les nœuds puissent ne pas avoir une vue complète de tous les sommets dans le système.
+%O Specifically, if $T' \gets T$, then every node in the system that has $T$ will also have $T'$ and the same relation $T' \gets T$; and conversely, if $T' \cancel{\gets} T$, then no nodes will end up with $T' \gets T$.
+En particulier, si $T' \gets T$, alors tout nœud dans le système qui possède $T$ aura également $T'$ et la même relation $T' \gets T$ ; inversement, si f $T' \cancel{\gets} T$, alors aucun nœud n'arrivera à $T' \gets T$.
 
-Each node $u$ can compute a confidence value, $d_u(T)$, from the progeny as follows:
+%O Each node $u$ can compute a confidence value, $d_u(T)$, from the progeny as follows:
+Chaque nœud $u$ peut calculer une valeur de confiance $d_u(T)$ de sa progéniture comme suit :
 \[ d_u(T) = \sum_{T' \in \mathcal{T}_u, T \stackrel{*}{\gets} T'}c_{uT'}\]
-where $c_{uT'}$ stands for the chit value of $T'$ for node $u$. Each transaction initially has a chit value of $0$ before the node gets
-the query results. If the node collects a threshold of $\alpha$ yes-votes after the query, the value $c_{uT'}$ is set to 1, otherwise remains $0$ forever.
-Therefore, a chit value reflects the result from the one-time query of its associated transaction and becomes immutable afterwards, while $d(T)$ can increase as the DAG grows by collecting more chits in its progeny.
+%O where $c_{uT'}$ stands for the chit value of $T'$ for node $u$. Each transaction initially has a chit value of $0$ before the node gets
+%O the query results. If the node collects a threshold of $\alpha$ yes-votes after the query, the value $c_{uT'}$ is set to 1, otherwise remains $0$ forever.
+où $c_{uT'}$ représente la valeur de \emph{chit} de $T'$ pour le nœud $u$.
+Chaque transaction a un valeur de \emph{chit} à $0$ avant que le nœud obtienne les résultats des requêtes.
+Si le nœud récolte un seuil de $\alpha$ votes «\,oui\,» après la requête, la valeur $c_{uT'}$ est mise à 1, sinon elle reste éternellement à $0$.
+%O Therefore, a chit value reflects the result from the one-time query of its associated transaction and becomes immutable afterwards, while $d(T)$ can increase as the DAG grows by collecting more chits in its progeny.
+Donc, une valeur de \emph{chit} reflète le résultat de la requête à un moment donné de ses transactions associées et devient immuable après cela, alors que $d(T)$ peut augmenter au fur et à mesure de la croissance du DAG en collectant d'autres \emph{chit} dans sa progéniture.
 Because $c_T \in \{0, 1\}$, confidence values are monotonic.
@@
 
-In addition, node $u$ maintains its own local list of known nodes $\mathcal{N}_u \subseteq \mathcal{N}$ that comprise the system.
-For simplicity, we assume for now $\mathcal{N}_u = \mathcal{N}$, and elide subscript $u$ in contexts without ambiguity.
+%O In addition, node $u$ maintains its own local list of known nodes $\mathcal{N}_u \subseteq \mathcal{N}$ that comprise the system.
+De plus, le nœud $u$ maintient sa propre liste locale de nœuds connus $\mathcal{N}_u \subseteq \mathcal{N}$ qui constituent le système.
+%O For simplicity, we assume for now $\mathcal{N}_u = \mathcal{N}$, and elide subscript $u$ in contexts without ambiguity.
+Dans un souci de simplicité, nous supposons pour l'instant que $\mathcal{N}_u = \mathcal{N}$, et nous éludons le $u$ souscrit dans les contextes sans ambiguïté.
 %
@@
 
-Each node implements an event-driven state machine, centered around a query that serves both to solicit votes on each transaction and to notify other nodes of the existence of newly discovered transactions.
-In particular, when node $u$ discovers a transaction $T$ through a query, it starts a one-time query process by sampling $k$ random peers and sending a message to them, after $T$ is delivered via $\textsc{onReceiveTx}$.
+%O Each node implements an event-driven state machine, centered around a query that serves both to solicit votes on each transaction and to notify other nodes of the existence of newly discovered transactions.
+Chaque nœud implémente une machine à état orientée événements, centrée autour d'une requête qui sert à la fois à solliciter des votes pour chaque transaction et à notifier d'autres nœuds de l'existence de transactions nouvellement découvertes.
+%O In particular, when node $u$ discovers a transaction $T$ through a query, it starts a one-time query process by sampling $k$ random peers and sending a message to them, after $T$ is delivered via $\textsc{onReceiveTx}$.
+En particulier, quand le nœud $u$ découvre une transaction $T$ au cours d'une requête, il commence un processus de requête unique % HOULA one-time query
+en échantillonnant $k$ pairs au hasard et en leur envoyant un message après que $T$ a été délivré par $\textsc{onReceiveTx}$.
 % determining the set of conflicting transactions $\mathcal{P}_T$,
 
-Node $u$ answers a query by checking whether each $T'$ such that $T' \stackrel{*}{\gets} T$ is currently preferred among competing transactions $\forall T'' \in \mathcal{P}_{T'}$.
-If every single ancestor $T'$ fulfills this criterion, the transaction is said to be \emph{strongly preferred}, and receives a yes-vote (1).
+%O Node $u$ answers a query by checking whether each $T'$ such that $T' \stackrel{*}{\gets} T$ is currently preferred among competing transactions $\forall T'' \in \mathcal{P}_{T'}$.
+Le nœud $u$ répond à une requête en vérifiant si chaque $T'$ telle que $T' \stackrel{*}{\gets} T$ est actuellement préférée par rapport aux transactions concurrentes $\forall T'' \in \mathcal{P}_{T'}$.
+%O If every single ancestor $T'$ fulfills this criterion, the transaction is said to be \emph{strongly preferred}, and receives a yes-vote (1).
 A failure of this criterion at any $T'$ yields a no-vote (0).
-When $u$ accumulates $k$ responses, it checks whether there are $\alpha$ yes-votes for $T$, and if so grants the chit (chit value $c_T \assign 1$) for $T$.
+Si chacun des ancêtres $T'$ répond à ce critère, la transaction est dite \emph{fortement préférée}, et reçoit un vote oui (1). Un manquement à ce critère pour toute $T'$ engendre un vote non (0).
+%O When $u$ accumulates $k$ responses, it checks whether there are $\alpha$ yes-votes for $T$, and if so grants the chit (chit value $c_T \assign 1$) for $T$.
+Quand $u$ accumule $k$ réponses, il vérifie s'il existe
 The above process will yield a labeling of the DAG with a chit value and associated confidence for each transaction~$T$.
@@
     \input{figures/dag-cd-compact}
-    \captionof{figure}{Example of $\langle \textrm{chit}, \textrm{confidence}\rangle$ values.
+%O     \captionof{figure}{Example of $\langle \textrm{chit}, \textrm{confidence}\rangle$ values.
 Darker boxes indicate transactions with higher confidence values.
 At most one transaction in each shaded region will be accepted.}
+    \captionof{figure}{Exemple de valeurs pour $\langle \textrm{chit}, \textrm{confidence}\rangle$.
+Les boîtes sombres indiquent des transactions dont l'indice de confiance est supérieur.
+Au plus une transaction dans chaque région ombrée sera acceptée.}
     \label{fig:dag-cd}
@@
 
-Figure~\ref{fig:dag-cd} illustrates a sample DAG built by Avalanche.
-Similar to Snowball, sampling in Avalanche will create a positive feedback for the preference of a single transaction in its conflict set.
-For example, because $T_2$ has larger confidence than $T_3$, its descendants are more likely collect chits in the future compared to $T_3$.
+%O Figure~\ref{fig:dag-cd} illustrates a sample DAG built by Avalanche.
+La Figure~\ref{fig:dag-cd} illustre un exemple de DAG construit par Avalanche.
+%O Similar to Snowball, sampling in Avalanche will create a positive feedback for the preference of a single transaction in its conflict set.
+De même que pour Snowball, l'échantillonnage dans Avalanche créer une rétroaction positive pour la préférence d'une seule transaction dans sont ensemble de conflit.
+%O For example, because $T_2$ has larger confidence than $T_3$, its descendants are more likely collect chits in the future compared to $T_3$.
+Par exemple, comme $T_2$ a un indice de confiance supérieur à celui de $T_3$, ses descendants seront plus à même de récolter des \emph{chits} à l'avenir que $T_3$.
 
-\tronly{Similar to Bitcoin, Avalanche leaves determining the acceptance point of a transaction to the application.
+%O \tronly{Similar to Bitcoin, Avalanche leaves determining the acceptance point of a transaction to the application.
 An application supplies an \textsc{isAccepted} predicate that can take into account the value at risk in the transaction and the chances of a decision being reverted to determine when to decide.}{}
+\tronly{Comme Bitcoin, Avalanche laisse à l'application le soin de déterminer le point d'acceptation d'une transaction.
+Une application fournit un prédicat \textsc{isAccepted} qui peut prendre en compte la valeur en jeu dans la transaction et les chances de réversion d'une décision pour déterminer le moment de cette décision.}{}
 
-Committing a transaction can be performed through a \emph{safe early commitment}.
+%O Committing a transaction can be performed through a \emph{safe early commitment}.
 For virtuous transactions, $T$ is accepted when it is the only transaction in its conflict set and has a confidence not less than threshold $\beta_1$.
-As in Snowball, $T$ can also be accepted after a $\beta_2$ number of consecutive successful queries.
-If a virtuous transaction fails to get accepted due to a problem
-with parents, it could be accepted if reissued with different parents.
-Figure~\ref{fig:gossipchain-ongen} shows how Avalanche entangles transactions.
+L'engagement concernant une transaction peut être effectié par un \emph{engagement préalable sûr}.
+Pour les transactions vertueuses, $T$ est acceptée quand elle est la seule transaction dans son ensemble de conflit et que sa confiance n'est pas inférieure à un seuil $\beta_1$.
+%O As in Snowball, $T$ can also be accepted after a $\beta_2$ number of consecutive successful queries.
+Comme dans Snowball, $T$ peut aussi être acceptée après un nombre de requêtes réussies consécutives $\beta_2$.
+%O If a virtuous transaction fails to get accepted due to a problem
+%O with parents, it could be accepted if reissued with different parents.
+Si une transaction vertueuse échoue à être acceptée en raison d'un problème avec ses parents elle peut être acceptée si elle est réémise avec des parents différents.
+%O Figure~\ref{fig:gossipchain-ongen} shows how Avalanche entangles transactions.
 Because transactions that consume and generate the same UTXO do not conflict with each other, any transaction can be reissued with different parents.
+La Figure~\ref{fig:gossipchain-ongen} montre comment Avalanche lie les transactions entre elles.
+Puisque les transactions qui consomment et génèrent la même UTXO n'entrent pas en conflit entre elles, toute transaction peut être réémise avec des parents différents.
 
 \tronly{
-Figure~\ref{fig:gossipchain-main} illustrates the protocol main loop
-executed by each node.
-In each iteration, the node attempts to select a transaction $T$ that has not
-yet been queried.  If no such
-transaction exists, the loop will stall until a new transaction is
-added to $\mathcal{T}$.
-It then selects $k$ peers and queries those peers.
-If more than $\alpha$ of those peers return a positive response, the chit value is set to~1.
-After that, it updates the preferred transaction of each conflict set
-of the transactions in its ancestry.
-Next, $T$ is added to the set $\mathcal{Q}$
-so it will never be queried again by the node.
-The code that selects additional peers if some of the $k$ peers are
-unresponsive is omitted for simplicity.
+%O Figure~\ref{fig:gossipchain-main} illustrates the protocol main loop
+%O executed by each node.
+La igure~\ref{fig:gossipchain-main} illustre la boucle principale du protocole exécutée par chaque nœud.
+%O In each iteration, the node attempts to select a transaction $T$ that has not
+%O yet been queried.  If no such
+%O transaction exists, the loop will stall until a new transaction is
+%O added to $\mathcal{T}$.
+À chaque itération, le nœud tente de sélectionner une transaction $T$ qui n'a pas encore été interrogée. 
+S'il n'existe pas de telle transaction, la boucle se met en pause jusqu'à ce qu'une nouvelle transaction soit ajoutée à $\mathcal{T}$.
+%O It then selects $k$ peers and queries those peers.
+Il sélectionne ensuite $k$ pairs et les interroge.
+%O If more than $\alpha$ of those peers return a positive response, the chit value is set to~1.
+Si plus que $\alpha$ de ces pairs reviennent à une réponse positive, la valeur de \emph{chit} est mise à~1.
+%O After that, it updates the preferred transaction of each conflict set
+%O of the transactions in its ancestry.
+Après cela, il met à jour la transaction qu'il préfère de chaque ensemble de conflits des transactions dans son ascendance.
+%O Next, $T$ is added to the set $\mathcal{Q}$
+%O so it will never be queried again by the node.
+Ensuite, $T$ est ajoutée à l'ensemble $\mathcal{Q}$ afin de ne plus jamais être interrogée par le nœud.
+%O The code that selects additional peers if some of the $k$ peers are
+%O unresponsive is omitted for simplicity.
+Le code qui sélectionne des pairs supplémentaires sur certains des $k$ pairs ne répondent pas est omis dans un but de simplicité.
 
-Figure~\ref{fig:gossipchain-onquery} shows what happens when a node
-receives a query for transaction $T$ from peer $j$.
-First it adds $T$ to $\mathcal{T}$, unless it already has it.
-Then it determines if $T$ is currently strongly preferred.
-If so, the node returns a positive response to peer $j$.
-Otherwise, it returns a negative response.
-Notice that in the pseudocode, we assume when a node knows $T$, it also
-recursively knows the entire ancestry of $T$. This can be achieved by
-postponing the delivery of $T$ until its entire ancestry is recursively
-fetched.
-In practice, an additional gossip process that disseminates
-transactions is used in parallel, but is not shown in pseudocode for simplicity.
+%O Figure~\ref{fig:gossipchain-onquery} shows what happens when a node
+%O receives a query for transaction $T$ from peer $j$.
+La Figure~\ref{fig:gossipchain-onquery} montre ce qui se produit quand un nœud reçoir une requête pour la transaction $T$ du pair $j$.
+%O First it adds $T$ to $\mathcal{T}$, unless it already has it.
+Il ajoute d'abord $T$ à $\mathcal{T}$, à moins qu'il ne l'ait déjà.
+%O Then it determines if $T$ is currently strongly preferred.
+Il détermine ensuite si $T$ est actuellement fortement préféré.
+%O If so, the node returns a positive response to peer $j$.
+Dans ce cas, le nœud renvoie une réponse positive au pair $j$.
+%O Otherwise, it returns a negative response.
+Sinon, il renvoie une réponse négative.
+%O Notice that in the pseudocode, we assume when a node knows $T$, it also
+%O recursively knows the entire ancestry of $T$. This can be achieved by
+%O postponing the delivery of $T$ until its entire ancestry is recursively
+%O fetched.
+On remarque que, dans le pseudocode, nous assumons que quand un nœud connaît $T$, il connaît aussi récursivement toute son ascendance.
+Cela peut être réalisé en retardant la délivrance de $T$ jusqu'à ce que l'intégralité de son ascendance soit récursivement récupérée.
+%O In practice, an additional gossip process that disseminates
+%O transactions is used in parallel, but is not shown in pseudocode for simplicity.
+En pratique, un processus \emph{gossip} supplémentaire qui dissémine des transaction est utilisé en parallèle, mais n'est pas montré dans le pseudocode dans un but de simplicité.
 }{}

### 2020-11-07 paper_fr.tex
S5 - avalanche_fr.tex terminé

--- 2020-11-07 paper_fr.tex
@@
 
-\section{Peer-to-Peer Payment System}
+\section{Système de paiement pair à pair}
 \label{sec:implementation}
@@
 Grâce au consensus Snowball, nous avons implémenté un système de paiement simpliste, Avalanche, qui supporte les transactions Bitcoin.
-Dans cette section, nous en décrivons la conception et schématison la manière donc l'implémentation peut supporter la primitive de transfert de valeur au centre des cryptomonnaies.
+Dans cette section, nous en décrivons la conception et schématison la manière dont l'implémentation peut supporter la primitive de transfert de valeur au cœur des cryptomonnaies.
 %O Deploying a full cryptocurrency involves bootstrapping, minting, staking, unstaking,

### 2020-11-07 paper_fr.tex
S5 - titre de schéma etc.

@@
 Comme les clients corrects suivent le protocole tel qu'il est prescrit et ne peuvent pratiquer la double dépense, ils ont la garantie tant de sûreté de fonctionnement que de vitalité pour leurs transactions \emph{vertueuses}.
-Par contraste, la vitalité n'est pas garantie pour les transactions \emph{malveillantes} soumises par des clients byzantins, qui entrent mutuellement en conflit.
+Par contraste, la vitalité n'est pas garantie pour les transactions \emph{malveillantes} soumises par des clients byzantins, lesquelles entrent mutuellement en conflit.
 Ces décisions peuvent arrêter le réseau mais n'ont pas d'impact sur les transactions vertueuses.

### 2020-11-08 avalanche_fr.tex
S5 - titre de schéma etc.
S5 (P2P payment system) complet

@@
-\subsection{Avalanche : ajout d'un DAG}%\tronly{}{\vspace{-0.5em}}
+\subsection{Avalanche\,: ajout d'un DAG}%\tronly{}{\vspace{-0.5em}}
 
@@
 %O The DAG has a single sink that is the \emph{genesis vertex}.
-Le DAG a un nœud terminal (\emph{sink}) unique qui est le \emph{genesis vertex}, le sommet genèse. % HOULA sink ? 
+Le DAG possède un nœud terminal (\emph{sink}) unique qui est le \emph{genesis vertex}, le sommet genèse. % HOULA sink ? 
 %O Maintaining a DAG provides two significant benefits.
@@
 %O Second, it also improves security, because the DAG intertwines the fate of transactions, similar to the Bitcoin blockchain.
-Ensuite, il améliore également la sécurité parce que le entrelace le sort des transactions, similaire en cela à la blockchain de Bitcoin.
+Ensuite, il améliore également la sécurité parce que le DAG lie fortement le sort des transactions, similaire en cela à la blockchain de Bitcoin.
 %O This renders past decisions difficult to undo without the approval of correct nodes.
@@
 %O The central challenge in the maintenance of the DAG is to choose among \emph{conflicting transactions}.
-Le principal défi de la maintenance du DAG est de choisir parmi des \emph{trasactions conflictuelles}.
+Le principal défi de la maintenance du DAG est de choisir parmi des \emph{transactions conflictuelles}.
 %O The notion of conflict is application-defined and transitive, forming an equivalence relation.
@@
 %O In our cryptocurrency application, transactions that spend the same funds (\emph{double-spends}) conflict, and form a \emph{conflict set}
-Dans notre application de cryptomonnaie, les transactions qui dépensent les mêmes fonds (\emph{double-dépense}) entrent en conflit et forment un \emph{ensemble de conflit}
 %(Figure~\tronly{\ref{fig:dag-conflict-set}}{\ref{fig:dag-cd}})
 %O (shaded regions in Figure~\ref{fig:dag-cd}), out of which only a single one can be accepted.
+Dans notre application de cryptomonnaie, les transactions qui dépensent les mêmes fonds (\emph{double-dépense}) entrent en conflit et forment un \emph{ensemble de conflits}
 (régions ombrées en Figure~\ref{fig:dag-cd}), duquel une seule peut être acceptée.
 %O Note that the conflict set of a virtuous transaction is always a singleton.
-On note que l'ensemble de conflit d'une transaction vertueuse est toujours un singleton.
+On note que l'ensemble de conflits d'une transaction vertueuse est toujours un singleton.
 %Note that the graph structure of transactions that spend depend on each other, also known as the UTXO graph, is completely independent of the DAG that Avalanche maintains.
@@
 %O Avalanche takes advantage of the DAG structure and uses a transaction's progeny.
-Alors que Snowball utilise des requêtes répétées et plusieurs compteurs pour capturer le degré de confiance associé aux transactions conflictuelles (couleurs),
+Alors que Snowball utilise des requêtes répétées et plusieurs compteurs pour obtenir le degré de confiance associé aux transactions conflictuelles (couleurs),
 Avalanche profite de la structure du DAG et utilise la progéniture d'une transaction.
@@
 %O the attacker generates large, padded subgraphs.
-On note que les \emph{chits} sont découplés de la structure du DAG, rendant le protocol immune aux attaques où l'attaquant génère de grands sous-graphes pleins de vide.
+On note que les \emph{chits} sont découplés de la structure du DAG, rendant le protocol immune aux attaques où un attaquant génère de grands sous-graphes faussement remplis.
 
 %O \subsection{Avalanche: Specification}%\tronly{}{\vspace{-0.5em}}
-\subsection{Avalanche: Spécification}%\tronly{}{\vspace{-0.5em}}
+\subsection{Avalanche\,: Spécification}%\tronly{}{\vspace{-0.5em}}
 \label{subsection:specification}
@@
 %O    \captionof{figure}{Avalanche: transaction generation.}\label{fig:gossipchain-ongen}
-    \captionof{figure}{Avalanche : génération des transactions.}\label{fig:gossipchain-ongen}
+    \captionof{figure}{Avalanche\,: génération des transactions.}\label{fig:gossipchain-ongen}
 \end{algorithmic}
@@
 %O    \captionof{figure}{Avalanche: the main loop.}\label{fig:gossipchain-main}
-    \captionof{figure}{Avalanche : boucle principale.}\label{fig:gossipchain-main}
+    \captionof{figure}{Avalanche\,: boucle principale.}\label{fig:gossipchain-main}
 \end{algorithmic}
@@
             (&(\forall T' \in \mathcal{T}, T' \gets T: \Call{isAccepted}{T'}) \\
-                &\land |\mathcal{P}_T| = 1 \land \mathcal{P}_T\mathit{.cnt} \ge \beta_1) \texttt{\hspace{.1in}// safe early commitment} \\
+                &\land |\mathcal{P}_T| = 1 \land \mathcal{P}_T\mathit{.cnt} \ge \beta_1) \texttt{\hspace{.1in}// engagement préalable sûr} \\
             %& \left(\frac{d(T)}{d'} > \gamma \land d' \ge \beta_2\right)
             \lor &(%\mathcal{P}_T\textrm{.pref} = \mathcal{P}_T\textrm{.last} \land
-            \mathcal{P}_T\mathit{.cnt} \ge \beta_2)\texttt{\hspace{.1in}// consecutive counter}
+            \mathcal{P}_T\mathit{.cnt} \ge \beta_2)\texttt{\hspace{.1in}// compteur consécutif}
         \end{align*}
@@
     \EndProcedure
-    \captionof{figure}{Avalanche: voting and decision primitives.}\label{fig:gossipchain-onquery}
+    \captionof{figure}{Avalanche\,: primitives de vote et de décision.}\label{fig:gossipchain-onquery}
 \end{algorithmic}
@@
 This relation may sound counter-intuitive: conflicting transitions have the \emph{equivalence} relation, because they are equivocations spending the \emph{same} funds.
-Comme les conflits sont transitifs, is $T_i$ et $T_j$ sont en conflit, ils appartiennent au même ensemble de conflit, c'est-à-dire $\mathcal{P}_{T_i} = \mathcal{P}_{T_j}$.
+Comme les conflits sont transitifs, si $T_i$ et $T_j$ sont en conflit, ils appartiennent au même ensemble de conflits, c'est-à-dire $\mathcal{P}_{T_i} = \mathcal{P}_{T_j}$.
 Cette relation peut sembler contre-intuitive : les transactions en conflit ont la même relation d'\emph{équivalence} car ce sont des équivoques dépensant les \emph{mêmes} fonds.
@@
 Donc, une valeur de \emph{chit} reflète le résultat de la requête à un moment donné de ses transactions associées et devient immuable après cela, alors que $d(T)$ peut augmenter au fur et à mesure de la croissance du DAG en collectant d'autres \emph{chit} dans sa progéniture.
-Because $c_T \in \{0, 1\}$, confidence values are monotonic.
+%O Because $c_T \in \{0, 1\}$, confidence values are monotonic.
+Comme $c_T \in \{0, 1\}$, les valeurs de confiance sont monotones.
 %\Jon{the notation $c_{uT}$ shows up before it's definet. I guess it's the number of chits counted for $T$ by participant $u$?}
@@
 Quand $u$ accumule $k$ réponses, il vérifie s'il existe
-The above process will yield a labeling of the DAG with a chit value and associated confidence for each transaction~$T$.
+%O The above process will yield a labeling of the DAG with a chit value and associated confidence for each transaction~$T$.
+Le processus ci-dessus ramène un étiquetage du DAG avec une valeur de \emph{chit} et une confiance associée pour chaque transaction~$T$.
 
@@
 %O Similar to Snowball, sampling in Avalanche will create a positive feedback for the preference of a single transaction in its conflict set.
-De même que pour Snowball, l'échantillonnage dans Avalanche créer une rétroaction positive pour la préférence d'une seule transaction dans sont ensemble de conflit.
+De même que pour Snowball, l'échantillonnage dans Avalanche crée une rétroaction positive pour la préférence d'une seule transaction dans sont ensemble de conflit.
 %O For example, because $T_2$ has larger confidence than $T_3$, its descendants are more likely collect chits in the future compared to $T_3$.
@@
 %O executed by each node.
-La igure~\ref{fig:gossipchain-main} illustre la boucle principale du protocole exécutée par chaque nœud.
+La Figure~\ref{fig:gossipchain-main} illustre la boucle principale du protocole exécutée par chaque nœud.
 %O In each iteration, the node attempts to select a transaction $T$ that has not
@@
 %O transactions is used in parallel, but is not shown in pseudocode for simplicity.
-En pratique, un processus \emph{gossip} supplémentaire qui dissémine des transaction est utilisé en parallèle, mais n'est pas montré dans le pseudocode dans un but de simplicité.
+En pratique, un processus \emph{gossip} supplémentaire qui dissémine des transactions est utilisé en parallèle, mais n'est pas montré dans le pseudocode dans un but de simplicité.
 }{}

### 2020-11-08 paper_fr.tex
S5 (P2P payment system) complet

@@
 Grâce au consensus Snowball, nous avons implémenté un système de paiement simpliste, Avalanche, qui supporte les transactions Bitcoin.
-Dans cette section, nous en décrivons la conception et schématison la manière dont l'implémentation peut supporter la primitive de transfert de valeur au cœur des cryptomonnaies.
+Dans cette section, nous en décrivons la conception et schématisons la manière dont l'implémentation peut supporter la primitive de transfert de valeur au cœur des cryptomonnaies.
 %O Deploying a full cryptocurrency involves bootstrapping, minting, staking, unstaking,
@@
 %O We show that this is a sensible tradeoff, and that the resulting system is sufficient for building complex payment systems.
-Nous montrons que ce compromis est sensé et que le système qui en résulte suffit à bâtir des systèmes de paiement complexes.
+Nous montrons que ce compromis est raisnnable et que le système qui en résulte suffit à bâtir des systèmes de paiement complexes.
 
@@
 
-\subsection{Multi-Input UTXO Transactions}
-In addition to the DAG structure in Avalanche, an \emph{unspent transaction output} (UTXO)~\cite{nakamoto2008bitcoin} graph that captures
-spending dependency is used to realize the ledger for the payment system. To
-avoid ambiguity, we denote the transactions that encode the data for money
-transfer \emph{transactions}, while we call the
-transactions ($T \in \mathcal{T}$) in Avalanche's DAG \emph{vertices}.
+%O \subsection{Multi-Input UTXO Transactions}
+\subsection{Transactions à UTXO multi-entrées}
+%O In addition to the DAG structure in Avalanche, an \emph{unspent transaction output} (UTXO)~\cite{nakamoto2008bitcoin} graph that captures
+%O spending dependency is used to realize the ledger for the payment system. To
+%O avoid ambiguity, we denote the transactions that encode the data for money
+%O transfer \emph{transactions}, while we call the
+%O transactions ($T \in \mathcal{T}$) in Avalanche's DAG \emph{vertices}.
+En addition à la structure de DAG dans Avalanche, un graphe de \emph{unspent transaction output} (UTXO)~\cite{nakamoto2008bitcoin}, ou sortie de transaction non dépensée, qui capture les dépendances des dépenses est utilisé pour réaliser le registre du système de paiement.
+Pour lever toute ambiguïté, nous appelons les transactions qui codent les données de transfert d'argent \emph{transactions} alors que nous appelons les transactions ($T \in \mathcal{T}$) dans le DAG d'Avalanche des \emph{sommets}.
 
@@
 %{
-We inherit the transaction and address mechanisms from Bitcoin. At their simplest, transactions consist of multiple inputs and outputs, with corresponding redeem scripts.
-Addresses are identified by the hash of their public keys, and signatures are generated by corresponding private keys.
-The full scripting language is used to ensure that a redeem script is authenticated to spend a UTXO\@.
-UTXOs are fully consumed by a valid transaction, and may generate new UTXOs spendable by named recipients.
-Multi-input transactions consume multiple UTXOs, and in Avalanche, may appear in multiple conflict sets.
-To account for these correctly, we represent \emph{transaction-input} pairs (e.g. $\texttt{In}_{a1}$) as Avalanche vertices.
-The conflict relation of transaction-input pairs is transitive because of each pair only spends one unspent output.
-Then, we use the conjunction of \textsc{isAccepted} for all inputs of a transaction to ensure that no transaction will be accepted unless all its inputs are accepted (Figure~\ref{fig:cash-system-b}).
+%O We inherit the transaction and address mechanisms from Bitcoin. At their simplest, transactions consist of multiple inputs and outputs, with corresponding redeem scripts.
+Nous héritons les mécanismes de transaction et d'adresse de Bitcoin. En simplifiant à l'extrême, les transactions consistent en multiples entrées et sorties, avec les scripts d'échange correspondants.
+%O Addresses are identified by the hash of their public keys, and signatures are generated by corresponding private keys.
+Les adresses sont identifiées par l'empreinte de leurs clefs publiques et les signatures sont générées par les clefs privées.
+%0 The full scripting language is used to ensure that a redeem script is authenticated to spend a UTXO\@.
+Le langage de script est utilisé pour s'assurer qu'un script d'échange est authentifié pour dépenser une UTXO\@.
+%O UTXOs are fully consumed by a valid transaction, and may generate new UTXOs spendable by named recipients.
+Les UTXO sont totalement consommées par une transaction valide et peut engendrer de nouvelles UTXO que des destinataires nommés peuvent dépenser.
+%O Multi-input transactions consume multiple UTXOs, and in Avalanche, may appear in multiple conflict sets.
+Les transactions multi-entrées consomment plusieurs UTXO et, dans Avalanche, peuvent apparaître dans plusieurs ensembles de conflits.
+%O To account for these correctly, we represent \emph{transaction-input} pairs (e.g. $\texttt{In}_{a1}$) as Avalanche vertices.
+Pour les prendre correctement en compte, nous représentons les paires d'entrées de transactions (\emph{transaction-input}), par exemple $\texttt{In}_{a1}$, comme des sommets Avalanche.
+%O The conflict relation of transaction-input pairs is transitive because of each pair only spends one unspent output.
+La relation des conflits de paires de sorties de transation est transitive car une seule paire ne dépense qu'une sortie non dépensée.
+%O Then, we use the conjunction of \textsc{isAccepted} for all inputs of a transaction to ensure that no transaction will be accepted unless all its inputs are accepted (Figure~\ref{fig:cash-system-b}).
 In other words, a transaction is accepted only if all its transaction-input pairs are accepted in their respective Snowball conflict sets.
-Following this idea, we finally implement the DAG of transaction-input pairs such that multiple
-transactions can be batched together per query.
+Nous utilisons la conjonction de \textsc{isAccepted} pour toutes les entrées d'une transaction pour nous assurer qu'aucune transaction ne sera acceptée à moins que toutes ses entrées ne soient acceptées (Figure~\ref{fig:cash-system-b}).
+En d'autres termes, une transaction n'est acceptée que si toutes ses paires d'entrées de transaction sont acceptées dans leurs ensembles de conflits Snowball respectifs.
+%O Following this idea, we finally implement the DAG of transaction-input pairs such that multiple
+%O transactions can be batched together per query.
+Selon cette idée, nous pouvons finalement implémenter le DAG des paires d'entrées de transactions tel que plusieurs transactions peuvent être regroupées par requête.
 
@@
     \input{figures/cash-system-b}
-    \captionof{figure}{The underlying logical DAG structure used by Avalanche.
-    The tiny squares with shades are dummy vertices which just help form the
-    DAG topology for the purpose of clarity, and can be replaced by direct
-    edges. The rounded gray regions are the conflict sets.}\label{fig:cash-system-b}
+%O     \captionof{figure}{The underlying logical DAG structure used by Avalanche.
+%O     The tiny squares with shades are dummy vertices which just help form the
+%O     DAG topology for the purpose of clarity, and can be replaced by direct
+%O     edges. The rounded gray regions are the conflict sets.}\label{fig:cash-system-b}
+    \captionof{figure}{La structure de DAG logique sous-jacente utilisée par Avalanche.
+    Les petits carrés avec des ombres sont des sommets fictifs qui ne servent qu'à former
+    la topologie du DAG dans un but de clarté, et peuvent être remplacés par des arêtes
+    directes. Les régions grises arrondies sont les ensembles de conflits.}\label{fig:cash-system-b}
 \end{center}
@@
 
-\paragraph{Optimizations}
-We implement some optimizations to help the system scale.
-First, we use \emph{lazy updates} to the DAG, because the recursive definition for confidence may otherwise require a costly DAG traversal.
-We maintain the current $d(T)$ value for each active vertex on the DAG, and update it only when a descendant vertex gets a chit.
-Since the search path can be pruned at accepted vertices, the cost for an update is constant if the rejected vertices have a limited number of descendants and the undecided region of
-the DAG stays at constant size.
-Second, the conflict set could be large in practice, because a rogue client can generate a large volume of conflicting transactions.
-Instead of keeping a container data structure for each conflict set, we create a mapping from each UTXO to the preferred transaction that stands as the representative for the entire conflict set.
-This enables a node to quickly determine future conflicts, and the appropriate response to queries.
-Finally, we speed up the query process by terminating early as soon as the $\alpha$ threshold is met, without waiting for $k$ responses.
+%O \paragraph{Optimizations}
+\paragraph{Optimisations}
+%O We implement some optimizations to help the system scale.
+Nous implémentons quelques optimisations pour que le système puisse passer à l'échelle.
+%O First, we use \emph{lazy updates} to the DAG, because the recursive definition for confidence may otherwise require a costly DAG traversal.
+D'abord, nous utilisons des \emph{lazy updates}, des mises à jour retardées, du DAG, car la définition récursive de la confiance pourrait engendrer une traversée du DAG coûteuse.
+
+%O We maintain the current $d(T)$ value for each active vertex on the DAG, and update it only when a descendant vertex gets a chit.
+Nous maintenons la valeurs $d(T)$ courante pour chaque sommet actif sur le DAG, et nous ne le mettons à jour que quand un sommet de la progéniture obtient un \emph{chit}.
+%O Since the search path can be pruned at accepted vertices, the cost for an update is constant if the rejected vertices have a limited number of descendants and the undecided region of
+%O the DAG stays at constant size.
+Comme le chemin de recherche peut être purgé sur des sommets acceptés, le coût d'une mise à jour est constant si les sommets rejetés ont un nombre limité de descendant et si la région non décidée du DAG reste de taille constante.
+%O Second, the conflict set could be large in practice, because a rogue client can generate a large volume of conflicting transactions.
+Ensuite, l'ensemble de conflits peut être vaste en pratique, car un client malveillant peut générer un volume important de transactions conflictuelles.
+%O Instead of keeping a container data structure for each conflict set, we create a mapping from each UTXO to the preferred transaction that stands as the representative for the entire conflict set.
+Au lieu de conserver une structure de données comme un conteneur pour chaque ensemble de conflits, nous créons une correspondance à partir de chaque UTXO vers la transaction préférée qui peut représenter tout l'ensemble de conflits.
+%O This enables a node to quickly determine future conflicts, and the appropriate response to queries.
+Cela permet à un nœud de déterminer rapidement les conflits à venir ainsi que la réponse appropriée aux requêtes.
+%O Finally, we speed up the query process by terminating early as soon as the $\alpha$ threshold is met, without waiting for $k$ responses.
+Enfin, nous accélérons le processus de requête en terminant au plus tôt, dès que le seuil $\alpha$ est atteint, sans attendre les $k$ réponses.
 
@@
 
-\paragraph{DAG} Compared to Snowball, Avalanche introduces a DAG structure that entangles the fate of unrelated conflict sets, each of which is a single-decree instance.
-This entanglement embodies a tension: attaching a virtuous transaction to undecided parents helps propel transactions towards a decision, while it puts transactions at risk of suffering liveness failures when parents turn out to be rogue.
-We can resolve this tension and provide a liveness guarantee with the aid of two mechanisms.
-
-First we adopt an adaptive parent selection strategy, where transactions are attached at the live edge of the DAG, and are retried with new parents closer to the genesis vertex.
+%O \paragraph{DAG} Compared to Snowball, Avalanche introduces a DAG structure that entangles the fate of unrelated conflict sets, each of which is a single-decree instance.
+\paragraph{DAG} Comparé à Snowball, Avalanche introduit une structure de DAG qui lie fortement le sort d'ensembles de conflits sans relation entre eux, chacun étant une instance à décret unique.
+%O This entanglement embodies a tension: attaching a virtuous transaction to undecided parents helps propel transactions towards a decision, while it puts transactions at risk of suffering liveness failures when parents turn out to be rogue.
+Cette intrication concrétise une tension : l'attachement d'une transaction à des parents non décidés contribue à la décision concernant les transactions, tandis que les transactions risquent des défaillances de vitalité quand les parents s'avèrent malveillants.
+%O We can resolve this tension and provide a liveness guarantee with the aid of two mechanisms.
+Nous pouvons résoudre cette tension et fournir une garantie de vitalité à l'aide de deux mécanismes.
+
+%O First we adopt an adaptive parent selection strategy, where transactions are attached at the live edge of the DAG, and are retried with new parents closer to the genesis vertex.
 This procedure is guaranteed to terminate with uncontested, decided parents, ensuring that a transaction cannot suffer liveness failure due to contested, rogue transactions.
-A secondary mechanism ensures that virtuous transactions with decided ancestry will receive sufficient chits.
+D'abord nous adoptons une stratégie adaptative de sélection des parents, où les transactions sont attachées sur le bord actif du DAG et sont réessayées avec de nouveaux parents plus proche du sommet genèse.
+Cette procédure garantit une terminaison avec des parents incontestés, assurant ainsi qu'une transaction ne peut souffrir de défaillance de vitalité en raison de transactions contestées et malveillantes.
+%O A secondary mechanism ensures that virtuous transactions with decided ancestry will receive sufficient chits.
 Correct nodes examine the DAG for virtuous transactions that lack sufficient progeny and emit no-op transactions to help increase their confidence.
-With these two mechanisms in place, it is easy to see that, at worst, Avalanche will degenerate into separate instances of Snowball, and thus provide the same liveness guarantee for virtuous transactions.
-
-Unlike other cryptocurrencies~\cite{IOTA} that use graph vertices
-directly as votes, Avalanche only uses a DAG for the purpose of batching queries
-in the underlying Snowball instances.
-Because confidence is built by collected chits, and not by just the presence of
-a vertex, simply flooding the network with vertices attached to the rejected
-side of a subgraph will not subvert the protocol.
-
-\subsection{Communication Complexity}
-Let the DAG induced by Avalanche have an expected branching factor of $p$, corresponding to the width of the DAG, and determined by the parent selection algorithm.
-Given the $\beta_1$ and $\beta_2$ decision threshold, a transaction that has just reached the point of decision will have an associated progeny $\mathcal{Y}$.
-Let $m$ be the expected depth of $\mathcal{Y}$.
-If we were to let the Avalanche network make progress and then freeze the DAG at a depth $y$,
-then it will have roughly $py$ vertices/transactions, of which $p(y - m)$ are decided in expectation.
-Only $pm$ recent transactions would lack the progeny required for a decision.
-For each node, each query requires $k$ samples, and therefore the total message cost per transaction is in expectation $(pky) / (p(y - m)) = ky/(y-m)$.
-Since $m$ is a constant determined by the undecided region of the DAG as the system constantly makes progress, message complexity per node is $O(k)$, while the total complexity is $O(kn)$.
+Un mécanisme secondaire s'assure que les transactions vertueuses avec une ascendance décidée reçoit suffisamment de \emph{chits}.
+Les nœuds corrects examinent le DAG pour y trouver des transactions vertueuses manquant d'une progéniture suffisante et émettent des transactions \emph{no-op} pour accroître leur confiance.
+%O With these two mechanisms in place, it is easy to see that, at worst, Avalanche will degenerate into separate instances of Snowball, and thus provide the same liveness guarantee for virtuous transactions.
+Ces deux mécanismes mis en place, il est facile de voir qu'au pire Avalanche dégénérera en instances distinctes de Snowball, fournissant ainsi la même garantie de vitalité pour les transactions vertueuses.
+
+%O Unlike other cryptocurrencies~\cite{IOTA} that use graph vertices
+%O directly as votes, Avalanche only uses a DAG for the purpose of batching queries
+%O in the underlying Snowball instances.
+Contrairement à d'autres cryptomonnaies~\cite{IOTA} qui utilisent directement les sommets du graphe comme votes, Avalanche n'utilise un DAG que pour regrouper les requêtes dans les instances Snowball sous-jacentes.
+
+%O Because confidence is built by collected chits, and not by just the presence of
+%O a vertex, simply flooding the network with vertices attached to the rejected
+%O side of a subgraph will not subvert the protocol.
+Comme la confiance est bâtie sur les \emph{chits} récoltés, et non simplement par la présence d'un sommet, la simple inondation du réseau par des sommets attachés à une partie rejetée du sous-graphe ne subvertit pas le protocole.
+
+%O \subsection{Communication Complexity}
+\subsection{Complexité des communications}
+%O Let the DAG induced by Avalanche have an expected branching factor of $p$, corresponding to the width of the DAG, and determined by the parent selection algorithm.
+Soit un facteur de ramification attendu de $p$ dans Avalanche, correspondant à la largeur du DAG, et déterminé par l'algorithme de sélection du parent.
+%O Given the $\beta_1$ and $\beta_2$ decision threshold, a transaction that has just reached the point of decision will have an associated progeny $\mathcal{Y}$.
+Étant donné le seuil de décision $\beta_1$ et $\beta_2$, une transaction qui vient d'atteindre le point de décision aura une progéniture associée $\mathcal{Y}$.
+%O Let $m$ be the expected depth of $\mathcal{Y}$.
+Soit $m$ la profondeur attendue de $\mathcal{Y}$.
+%O If we were to let the Avalanche network make progress and then freeze the DAG at a depth $y$,
+%O then it will have roughly $py$ vertices/transactions, of which $p(y - m)$ are decided in expectation.
+Si nous devions permettre au réseau Avalanche de progresser puis de geler le DAG à une profondeur $y$, alors il aurait à peu près $py$ sommets/transactions, desquels on pourrait attendre que $p(y - m)$ soient décidés.
+%O Only $pm$ recent transactions would lack the progeny required for a decision.
+Seules $pm$ transactions récentes manqueraient de la progéniture nécessaires à une prise de décision.
+%O For each node, each query requires $k$ samples, and therefore the total message cost per transaction is in expectation $(pky) / (p(y - m)) = ky/(y-m)$.
+Pour chaque nœud, chaque requête demande $k$ échantillons, et donc on attend un coût total des messages par transaction de $(pky) / (p(y - m)) = ky/(y-m)$.
+%O Since $m$ is a constant determined by the undecided region of the DAG as the system constantly makes progress, message complexity per node is $O(k)$, while the total complexity is $O(kn)$.
+Comme $m$ est une constante déterminée par la région non décidée du DAG pendant que le système progresse constamment, la complexité en messages par nœuds est $O(k)$ alors que la complexité totale est $O(kn)$.
 

### 2020-11-08 paper_fr.tex
Finish french translation of section 4

@@
 
-\subsection{Safety}
+\subsection{Sûreté de fonctionnement}
+
+%O \subsection{Safety}
 
 \paragraph{Slush}
-We assume that
-%$\mathcal{L}(u) = \mathcal{N}$ for all $u \in \mathcal{N}$.
-all nodes share the same $\mathcal{N}$, and in
-Appendix~\ref{sec:full-analysis-churn}, we sketch how to relax the requirement of the membership knowledge.
-We model the dynamics of the system through a continuous-time process where two states are absorbing, namely the all-red or all-blue state.
+En partant du principe que tous les noeuds partagent le même $\mathcal{N}$, et en Annexe~\ref{sec:full-analysis-churn},
+nous expliquons comment relâcher le besoin pour les noeuds d'avoir connaissance de leur appartenance. Nous modélisons
+les dynamiques du système à travers un processus à temps continu contenant deux états, tous les noeuds rouges ou tous
+les noeuds bleus. Prenons $\{X_{t \geq 0}\}$ la variable aléatoire qui décrit l'état du système à l'instant $t$, où
+$X_0 \in \{0, \dots, c\}$.
+Nous commençons par détailler le résultat le plus important des dynamiques de sûreté de fonctionnement de notre
+processus: les probabilités de \emph{réversibilité} du processus \textbf{Slush}. Tous les autres résultats formels de
+ce papier sont, en simplifiant, des dérivations et améliorations intuitives de ce résultat.
+
+%0 \paragraph{Slush}
+%0 We assume that
+%0 %$\mathcal{L}(u) = \mathcal{N}$ for all $u \in \mathcal{N}$.
+%0 all nodes share the same $\mathcal{N}$, and in
+%0 Appendix~\ref{sec:full-analysis-churn}, we sketch how to relax the requirement of the membership knowledge.
+%0 We model the dynamics of the system through a continuous-time process where two states are absorbing, namely the all-red or all-blue state.
 Let $\{X_{t \geq 0}\}$ be the random variable that describes the state of the system at time $t$, where $X_0 \in \{0, \dots, c\}$.
-We begin by immediately discussing the most important result of the safety dynamics of our processes: the \emph{reversibility} probabilities of the \textbf{Slush} process.
+%0 We begin by immediately discussing the most important result of the safety dynamics of our processes: the \emph{reversibility} probabilities of the \textbf{Slush} process.
 All the other formal results in this paper are, informally speaking, intuitive derivations and augmentations of this result.
@@
 \begin{theorem}
-Let the configuration of the system at time $t$ be $\mathcal{S}_t = n/2 + \delta$, meaning that the network has drifted to $2\delta$ more blue nodes than red nodes ($\delta = 0$ means that red and blue are equal).
-Let $\xi_\delta$ be the probability of absorption to the all-red state (minority).
-Then, for all $0 \leq \delta \leq n/2$, we have
+Prenons la configuration du système à l'instant $t$ comme étant $\mathcal{S}_t = n/2 + \delta$, ce qui signifie que le
+réseau a dérivé de $2\delta$ noeuds bleus de plus que les noeuds rouges ($\delta = 0$ signifie que rouges et bleus
+sont en nombre égal). $\xi_\delta$ étant la probabilité d'un basculement vers l'état où tous les noeuds sont rouges,
+nous avons donc pour tout $0 \leq \delta \leq n/2$
 \begin{equation}
@@
 \begin{proof}
-This bound follows from the Hoeffding-derived tail bounds of the hypergeometric distribution by Chvatal~\cite{chvatal1979tail}. 
+Cette borne suit les bornes de queue dérivées de Hoeffding de la distribution hypergéométrique définie par
+Chvatal~\cite{chvatal1979tail}.
 \end{proof}
 
-We note that Chvatal's bounds are introduced for simplicity of exposition and are extremely weak.  
-We leave the full closed-form expression in Theorem~\ref{theorem:slush_prob_convergence_minority} to the appendix, which is also significantly stronger than the Chvatal bound. 
-Nonetheless, using the loose Chvatal bound, we make the key observation that as the drift $\delta$ increases, given fixed $\alpha$ and $k$, the probability of moving towards the minority value decreases \emph{exponentially fast} (in fact, even faster, since there is a quadratic term in the inverse exponent).
+%O \begin{theorem}
+%O Let the configuration of the system at time $t$ be $\mathcal{S}_t = n/2 + \delta$, meaning that the network has drifted to $2\delta$ more blue nodes than red nodes ($\delta = 0$ means that red and blue are equal).
+Let $\xi_\delta$ be the probability of absorption to the all-red state (minority).
+Then, for all $0 \leq \delta \leq n/2$, we have
+%O \begin{equation}
+%O \begin{split}
+%O     \xi_\delta &\leq \left(\dfrac{1/2 - \delta/n}{\alpha/k}\right)^{\alpha}\left(\dfrac{1/2 + \delta/n}{1- \alpha/k}\right)^{k-\alpha}\\
+%O     &\leq e^{-2((\alpha/k) - (1/2) + (\delta/n))^2 k}
+%O \end{split}
+%O \end{equation}
+%O \end{theorem}
+
+%O \begin{proof}
+%O This bound follows from the Hoeffding-derived tail bounds of the hypergeometric distribution by Chvatal~\cite{chvatal1979tail}.
+%O \end{proof}
+
+Nous notons que les bornes de Chvatal sont introduites ici dans un but de simplification et sont extrêmement faibles.
+Nous laissons l'expression dans sa forme complète dans le Théorème~\ref{theorem:slush_prob_convergence_minority}
+disponible en annexe, qui est aussi bien plus fort que la borne de Chvatal.
+Néanmoins, l'utilisation de la borne de Chvatal nous permet de faire l'observation qui consiste à dire qu'au fur et à
+mesure que la dérivation $\delta$ augmente, pour un $\alpha$ et un $k$ constants, la probabilité de basculer vers la
+valeur minoritaire décroît \emph{de manière exponentielle} (en fait encore plus rapidement vu qu'il y a un terme
+quadratique dans l'exposant inverse). De plus, la même observation est valable si l'on fait croître $\alpha$ en gardant
+$k$ constant.
+
+Les conclusions de ce théorème démontreint une propriété clé du système: une fois que le réseau perd sa bivalence
+(ex: $\delta > 0$), il tend à se renverser et converger rapidement vers la couleur majoritaire et il devient alors
+suffisamment improbable qu'il rebascule vers la couleur minoritaire. C'est cette propriété fondamentale que nous
+exploitons dans nos protocoles, et ce qui les rend sûrs malgré le fait que l'échantillonnage se fait sur une petite
+partie du réseau de taille fixe. Le résultat principal qui s'en suit pour les garanties de sûreté de Snowflake consiste
+à trouver des régions (à partir de paramètres spécifiquement choisis) où la réversibilité n'est possible qu'à partir
+d'une probabilité maximale $\varepsilon$ même en conditions d'adversité.
+
+%O We note that Chvatal's bounds are introduced for simplicity of exposition and are extremely weak.
+%O We leave the full closed-form expression in Theorem~\ref{theorem:slush_prob_convergence_minority} to the appendix, which is also significantly stronger than the Chvatal bound.
+%O Nonetheless, using the loose Chvatal bound, we make the key observation that as the drift $\delta$ increases, given fixed $\alpha$ and $k$, the probability of moving towards the minority value decreases \emph{exponentially fast} (in fact, even faster, since there is a quadratic term in the inverse exponent).
 Additionally, the same result holds for increasing $\alpha$ given a fixed $k$.
-
-The outcomes of this theorem demonstrate a key property: once the network loses full bivalency (i.e.
+%O
+%O The outcomes of this theorem demonstrate a key property: once the network loses full bivalency (i.e.
 $\delta > 0$), it tends to topple and converge rapidly towards the majority color, unable to revert back to the minority with significant probability.
@@
 \paragraph{Snowflake}
-For Snowflake, we assume that some fraction of nodes are adversarial.
+
+Dans le cas de Snowflake, nous prenons comme hypothèse qu'une partie des noeuds est antagoniste. Avec Slush, une
+fois que le réseau obtient une majorité significative pour l'une des propositions (ex: la couleur bleue), il devient
+improbable qu'une proposition minoritaire (ex: la couleur rouge) puisse être adoptée dans le futur (irréversibilité).
+De plus, les noeuds Slush ont simplement besoin d'exécuter le protocole pour un nombre déterministe $m$ de tours qui
+est connu en amont de l'exécution du protocole. Lorsqu'on introduit des noeuds antagonistes qui suivent des stratégies
+arbitraires, les noeuds ne peuvent simplement plus exécuter le protocole pour un nombre déterministe de tours vu que
+les adversaires peuvent affecter la valeur de $m$ de manière non déterministe. A la place, les noeuds légitimes doivent
+implémenter un mécanisme pour explicitement détecter que l'irréversibilité a été atteinte. Dans ce but, dans Snowflake
+chaque noeud légitime implémente une fonction de décision, $\mathcal{D}(u, \mathcal{S}_t, blue) \rightarrow \{0, 1\}$,
+qui est une variable aléatoire qui retourne $1$ si le noeud $u$ détecte que l'état d'irréversibilité a été atteint à
+l'instant $t$ pour bleu. Ce mécanisme de décision est probabiliste, ce qui veut dire qu'il peut échouer, même s'il est
+conçu pour que la probabilité de cet échec soit négligeable. Nous pouvons maintenant ébaucher la preuve de Snowflake.
+
+\noindent \emph{Ebauche de preuve}. Nous définissons la stratégie d'échec comme étant l'évênement lors duquel deux
+noeuds légitimes $u$ et $v$ décident entre le bleu et le rouge, ex: $\mathcal{D}(u, \mathcal{S}_t, bleu) \rightarrow 1$
+et $\mathcal{D}(v, \mathcal{S}_{t'}, rouge) \rightarrow 1$, pour deux instants $t$ et $t'$. Nous modélisons une fois de
+plus le système comme un processus à temps aléatoire. L'espace des états est défini de la même manière que pour Slush.
+Néanmoins, nous observons des subtilités critiques. Tout d'abord, même si tous les noeuds légitimes acceptent une
+couleur, les noeuds byzantins peuvent toujours basculer vers l'autre couleur. Ensuite nous devons aussi prendre en
+considération le mécanisme de décision $\mathcal{D}(*)$. Pour cette analyse, nous nous affranchissons du besoin de
+garder l'historique de toutes les configurations du réseau soumises à toutes les stratégies antagonistes et prenons pour
+hypothèse qu'un noeud $u$ décide d'abord d'adopter la couleur bleue. Ensuite en fonction de l'état du réseau suite à la
+décision de $u$, nous calculons la probabilité qu'un autre noeud $v$ adopte le rouge, qui est fonction à la fois de la
+probabilité que le réseau bascule vers l'état minoritaire bleu et que $v$ bascule vers cet état.
+Nous démontrons qu'en choisissant correctement $k$, $\alpha$, et $\beta$, nous pouvons construire des instances
+extrêmement sécurisées de Snowflake (ex: échec de sécurité avec une probabilité de $\leq \varepsilon$) quand le réseau
+atteint un biais de $\delta$, comme le montre la Figure~\ref{fig:states_feasible_solutions}. Un exemple concret est
+présenté dans la Figure~\ref{fig:fandepsilon}.
+
+\begin{figure}[h]
+\begin{center}
+\input{figures/analysis2}
+\caption{Représentation de l'état d'irréversibilité qui existe quand -- même en la présence de $f$ noeuds byzantins --
+  le nombre de noeuds légitimes bleus dépasse le nombre de noeuds légitimes rouges de plus de $2\delta$.
+}
+\label{fig:states_feasible_solutions}
+\end{center}
+\end{figure}
+
+%O \paragraph{Snowflake}
+%O For Snowflake, we assume that some fraction of nodes are adversarial.
 In Slush, once the network gains significant majority support for one proposal (e.g., the color blue), it becomes unlikely for a minority proposal (e.g., the color red) to ever become decided in the future (irreversibility).
@@
 We now sketch the proof of Snowflake.
-
-\noindent \emph{Proof Sketch}.
+%O
+%O \noindent \emph{Proof Sketch}.
 We define safety failure to be the event wherein any two correct nodes $u$ and $v$ decide on blue and red, i.e.
@@
 Then, conditioned on the state of the network upon $u$ deciding, we calculate the probability that another node $v$ decides red, which is a function of both the probability that the network reverts towards a minority blue state and that $v$ decides at that state.
-We show that under appropriate choices of $k$, $\alpha$, and $\beta$, we can construct highly secure instances of Snowflake (i.e.
+%O We show that under appropriate choices of $k$, $\alpha$, and $\beta$, we can construct highly secure instances of Snowflake (i.e.
 safety failure with probability $\leq \varepsilon$) when the network reaches some bias of $\delta$, as shown in Figure~\ref{fig:states_feasible_solutions}.
 A concrete example is provided in Figure~\ref{fig:fandepsilon}.
-
-\begin{figure}[h]
-\begin{center}
-\input{figures/analysis2}
-\caption{Representation of the irreversibility state, which exists when -- even under $f$ Byzantine nodes -- the number of blue correct nodes exceeds that of red correct nodes by more than $2\delta$.
-}
-\label{fig:states_feasible_solutions}
-\end{center}
-\end{figure}
+%O
+%O \begin{figure}[h]
+%O \begin{center}
+%O \input{figures/analysis2}
+%O \caption{Representation of the irreversibility state, which exists when -- even under $f$ Byzantine nodes -- the number of blue correct nodes exceeds that of red correct nodes by more than $2\delta$.
+%O }
+%O \label{fig:states_feasible_solutions}
+%O \end{center}
+%O \end{figure}
 
@@
 \paragraph{Snowball}
-Snowball is an improvement in security over Snowflake, where random perturbations in network samples are reduced by introducing a limited form of history, which we refer to as confidence. 
-% Modeling Snowball with a Markov chain is difficult because of a state space explosion problem.
+
+Snowball est une amélioration en terme de sécurité par rapport à Snowflake, où les perturbations aléatoires dans les
+échantillons du réseau sont réduites par l'introduction d'une forme limitée d'historique que l'on appelle confiance.
+
+\noindent \emph{Ebauche de preuve}. Nous appliquons des inégalités de concentration pour prouver qu'une fois le système
+ayant atteint l'état d'irréversibilité, alors l'évolution de la confiance envers la couleur majoritaire adoptée va
+augmenter constamment et dériver de plus en plus loin des noeuds de couleur minoritaire, jusqu'à rendre la réversibilité
+de plus en plus improbable avec le temps. Si cette dérive finit par se retourner, alors l'analyse de la réversibilité
+devient identique à celle de Snowflake.
+
+%O \paragraph{Snowball}
+%O Snowball is an improvement in security over Snowflake, where random perturbations in network samples are reduced by introducing a limited form of history, which we refer to as confidence.
+%O % Modeling Snowball with a Markov chain is difficult because of a state space explosion problem.
 In particular, it is not sufficient to simply keep track of color preferences of each correct node, the analysis must also maintain information about their confidence.
-
-\noindent \emph{Proof Sketch}.
+%O
+%O \noindent \emph{Proof Sketch}.
 We apply martingale concentration inequalities to prove that once the system has reached the irreversibility state, then the growth of the confidence of the majority decided color will perpetually grow and drift further away from those of the minority color, effectively rendering reversibility less likely over time.
@@
 
-\subsection{Liveness}
-
-We assume that the observed adversarial presence $0 \leq f' \leq n(k - \alpha - \psi)/k \leq f$, where we refer to $\psi$ as the buffer zone. 
-The bigger $\psi$, the quicker the ability of the decision mechanism to finalize a value. 
-If, of course, $\psi$ approaches zero or becomes negative, then we violate the upper bound of adversarial tolerance for the parameterized system, and thus the adversary can, with high probability, stall termination by simply choosing to not respond, although the safety guarantees may still hold.
-
-Assuming that $\psi$ is strictly positive, termination is strictly finite under all network configurations where a proposal has at least $\alpha$ support.
+\subsection{Vitalité}
+
+Nous prenons pour hypothèse la présence antagoniste observée $0 \leq f' \leq n(k - \alpha - \psi)/k \leq f$, où $\psi$
+représente la zone tampon.
+Plus $\psi$ est grand, plus la capacité du mécanisme de décision à tendre vers une valeur est rapide. Si bien sûr $\psi$
+tend vers zéro ou devient négatif alors nous dépassons la limite supérieure de tolérance antagoniste pour le système
+paramétré, en conséquence l'attaquant peut, avec une haute probabilité, bloquer la convergence en choisissant simplement
+de ne pas répondre, et ceci même si les garanties de sûreté s'appliquent toujours.
+
+En partant de $\psi$ strictement positif, la convergence est strictement finie pour toutes les configurations où une
+proposition a un support d'au moins $\alpha$. De plus, non seulement la convergence est finie avec une probabilité de
+un, mais nous avons aussi une probabilité strictement positive de convergence jusqu'à un instant borné $t_{max}$, comme
+décrit dans la Proposition~\ref{lemma:finitetermination}, qui découle du Théorème~\ref{theorem:mean-convergence-time}.
+Cela précise la propriété de vitalité P2.
+
+\noindent \emph{Ebauche de preuve}. En utilisant la construction du système pour prouver l'irréversibilité, nous
+caractérisons la distribution de la moyenne du temps passé (les périodes "sojourn") dans chaque état avant que le
+système ne termine son exécution en adoptant l'un des états. Ce temps de convergence est l'union de ces périodes.
+
+Pour les transactions non conflictuelles, vu que l'attaquant est incapable de générer un conflit, le temps de décision
+est simplement le temps de mélange, qui est $\Oh{\log{n}}$. La vitalité guarantit qu'une configuration de réseau
+entièrement bivalent atteint un temps de convergence optimal de $\Oh{\log{n}}$ tours si le nombre d'attaquants est au
+plus $\Oh{\sqrt{n}}$, pour $\alpha = \floor{k/2} + 1$. Nous donnons plus de détails dans la
+Proposition~\ref{lemma:centrallimit}.
+Quand le nombre de noeuds antagonistes dépasse $\Oh{\sqrt{n}}$, le nombre le plus pessimiste de tours grandit de manière
+polynomiale, et quand $f$ tend vers $n/2$ ce nombre de tours tend vers des taux de convergence exponentiels.
+
+\noindent \emph{Ebauche de preuve}. Nous modifions le Théorème~\ref{theorem:mean-convergence-time} en y incluant le
+nombre de nodes antagonistes, ce qui annule toute déséquilibre dans le réseau en le gardant entièrement bivalent.
+
+%O \subsection{Liveness}
+%O
+%O We assume that the observed adversarial presence $0 \leq f' \leq n(k - \alpha - \psi)/k \leq f$, where we refer to $\psi$ as the buffer zone.
+%O The bigger $\psi$, the quicker the ability of the decision mechanism to finalize a value.
+%O If, of course, $\psi$ approaches zero or becomes negative, then we violate the upper bound of adversarial tolerance for the parameterized system, and thus the adversary can, with high probability, stall termination by simply choosing to not respond, although the safety guarantees may still hold.
+%O
+%O Assuming that $\psi$ is strictly positive, termination is strictly finite under all network configurations where a proposal has at least $\alpha$ support.
 Furthermore, not only is termination finite with probability one, we also have a strictly positive probability of termination within any bounded amount of time $t_{max}$, as discussed in Lemma~\ref{lemma:finitetermination}, which follows from Theorem~\ref{theorem:mean-convergence-time}.
 This captures liveness property P2.
-
-\noindent\emph{Proof Sketch.} Using the construction of the system to prove irreversibility, we characterize the distribution of the average time spent (sojourn times) at each state before the system terminates execution by absorption at either absorbing state.
+%O
+%O \noindent\emph{Proof Sketch.} Using the construction of the system to prove irreversibility, we characterize the distribution of the average time spent (sojourn times) at each state before the system terminates execution by absorption at either absorbing state.
 The termination time is then a union of these times.
-
-For non-conflicting transactions, since the adversary is unable to forge a conflict, the time to decision is simply the mixing time , which is $\Oh{\log{n}}$.
-Liveness guarantees under a fully bivalent network configuration reduce to an optimal convergence time of $\Oh{\log{n}}$ rounds if the adversary is at most $\Oh{\sqrt{n}}$, for $\alpha = \floor{k/2} + 1$.
+%O
+%O For non-conflicting transactions, since the adversary is unable to forge a conflict, the time to decision is simply the mixing time , which is $\Oh{\log{n}}$.
+%O Liveness guarantees under a fully bivalent network configuration reduce to an optimal convergence time of $\Oh{\log{n}}$ rounds if the adversary is at most $\Oh{\sqrt{n}}$, for $\alpha = \floor{k/2} + 1$.
 We leave additional detains to Lemma~\ref{lemma:centrallimit}.
-% This result is independently supported by Ben-Or~\cite{ben1983another} and Doerr et al~\cite{doerr2011stabilizing}. 
-When the adversary surpasses $\Oh{\sqrt{n}}$ nodes, the worst-case number of rounds increases polynomially, and as $f$ approaches $n/2$ it approaches exponential convergence rates.
-
-\noindent\emph{Proof Sketch.} We modify Theorem~\ref{theorem:mean-convergence-time} to include the adversary, which reverts any imbalances in the network by keeping network fully bivalent.
-
-\paragraph{Multi-Value Consensus}
-Our binary consensus protocol could support multi-value consensus by running logarithmic binary instances, one for each bit of the proposed value.
+%O % This result is independently supported by Ben-Or~\cite{ben1983another} and Doerr et al~\cite{doerr2011stabilizing}.
+%O When the adversary surpasses $\Oh{\sqrt{n}}$ nodes, the worst-case number of rounds increases polynomially, and as $f$ approaches $n/2$ it approaches exponential convergence rates.
+%O
+%O \noindent\emph{Proof Sketch.} We modify Theorem~\ref{theorem:mean-convergence-time} to include the adversary, which reverts any imbalances in the network by keeping network fully bivalent.
+
+\paragraph{Consensus multi-valeur}
+
+Notre protocole de consensus binaire pourrait supporter des consensus multi-valeur en exécutant des instances binaires
+logarithmiques, une pour chaque bit de la valeur proposée. Cependant, une telle réduction théorique n'est pas
+forcément efficace en pratique. A la place, nous pourrions directement incorporer des valeurs multiples comme des
+couleurs multiples dans le protocole en permettant toujours la généralisation de l'analyse de sûreté.
+
+Pour ce qui est de la vitalité nous ébauchons en Annexe~\ref{sec:sync-heuristic} un mécanisme d'initialisation sans
+leader, qui de toute attente nécessite $\Oh{\log{n}}$ tours dans l'hypothèse que le réseau soit synchronisé. Bien que
+la conception de ce mécanisme soit intéréssante, il est à noter qu'un tel mécanisme n'est pas nécéssaire pour un système
+de paiement décentralisé, comme nous le démontrons en Section~\ref{sec:implementation}.
+Finalement, nous développons les divergences entre vues et pertes en Annexe~\ref{sec:full-analysis-churn}.
+
+%0 \paragraph{Multi-Value Consensus}
+%0 Our binary consensus protocol could support multi-value consensus by running logarithmic binary instances, one for each bit of the proposed value.
 However, such theoretical reduction might not be efficient in practice.
 Instead, we could directly incorporate multi-values as multi-colors in the protocol, where safety analysis could still be generalized.
-
-For liveness, we sketch a leaderless initialization mechanism, which in expectation uses $\Oh{\log{n}}$ rounds under the assumption that the network is synchronized in the Appendix~\ref{sec:sync-heuristic}.
-While the design of initialization mechanisms is interesting, note that it is not necessary for a decentralized payment system, as we show in Section~\ref{sec:implementation}.
-%We leave additional research into the initialization mechanisms to future work. 
-%However, we finally note that when building a decentralized payment system, an initialization mechanism is not necessary, as we show in the implementation of Avalanche in Section~\ref{sec:implementation}.
-Finally, in the Appendix~\ref{sec:full-analysis-churn}, we discuss churn and view discrepancies.
+%0
+%0 For liveness, we sketch a leaderless initialization mechanism, which in expectation uses $\Oh{\log{n}}$ rounds under the assumption that the network is synchronized in the Appendix~\ref{sec:sync-heuristic}.
+%0 While the design of initialization mechanisms is interesting, note that it is not necessary for a decentralized payment system, as we show in Section~\ref{sec:implementation}.
+%0 %We leave additional research into the initialization mechanisms to future work.
+%0 %However, we finally note that when building a decentralized payment system, an initialization mechanism is not necessary, as we show in the implementation of Avalanche in Section~\ref{sec:implementation}.
+%0 Finally, in the Appendix~\ref{sec:full-analysis-churn}, we discuss churn and view discrepancies.
 

### 2020-11-08 paper_fr.tex
Annexe : Analyse - A1 A2

--- 2020-11-08 paper_fr.tex
@@
 
-\newtheorem{theorem}{Theorem}
-\newtheorem{lemma}[theorem]{Lemma}
+%O \newtheorem{theorem}{Theorem}
+\newtheorem{theorem}{Théorème}
+%O \newtheorem{lemma}[theorem]{Lemma}
+\newtheorem{lemma}[theorem]{Lemme}
 \newtheorem{conjecture}[theorem]{Conjecture}
+%O \theoremstyle{definition}
 \theoremstyle{definition}
-\newtheorem{definition}{Definition}[section]
+%O \newtheorem{definition}{Definition}[section]
+\newtheorem{definition}{Définition}[section]
 
@@
 
+% Default section titles for French version
+\renewcommand{\abstractname}{Résumé}
+\renewcommand{\refname}{Références}
+\renewcommand{\proofname}{Preuve}
 \settimeformat{hhmmsstime}
@@
 D'abord nous adoptons une stratégie adaptative de sélection des parents, où les transactions sont attachées sur le bord actif du DAG et sont réessayées avec de nouveaux parents plus proche du sommet genèse.
-Cette procédure garantit une terminaison avec des parents incontestés, assurant ainsi qu'une transaction ne peut souffrir de défaillance de vitalité en raison de transactions contestées et malveillantes.
+Cette procédure garantit un achèvement avec des parents incontestés, assurant ainsi qu'une transaction ne peut souffrir de défaillance de vitalité en raison de transactions contestées et malveillantes.
 %O A secondary mechanism ensures that virtuous transactions with decided ancestry will receive sufficient chits.
@@
 \begin{appendices}
-\section{Analysis}
+%O \section{Analysis}
+\section{Analyse}
 \label{sec:full-analysis}
-In this appendix, we provide an analysis of Slush, Snowflake and Snowball.
+%O In this appendix, we provide an analysis of Slush, Snowflake and Snowball.
+Cette annexe contient une analyse de Slush, Snowflake et Snowball.
 %For lack of space, we leave additional details to an accompanying report.
 
-\subsection{Preliminaries}
-We assume the network model as discussed in Section~\ref{sec:model_and_goals}. We let $\mathtt{R}$ (``red'') and $\mathtt{B}$ (``blue'') represent two generic conflicting choices.
-Without loss of generality, we focus our attention on counts of $\mathtt{B}$, i.e.\ the total number of nodes that prefer blue.
-
-\paragraph{Hypergeometric Distribution} Each network query of $k$ peers corresponds to a sample without replacement out of a network of $n$ nodes, also referred to as a hypergeometric sample.
-We let the random variable $\mathcal{H}(\mathcal{N}, x, k) \rightarrow \{0, \dots, k\}$ denote the resulting counts of $\mathtt{B}$ in the sample (unless otherwise stated), where $x$ is the total count of $\mathtt{B}$ in the population.
+%O \subsection{Preliminaries}
+\subsection{Préliminaires}
+%O We assume the network model as discussed in Section~\ref{sec:model_and_goals}. We let $\mathtt{R}$ (``red'') and $\mathtt{B}$ (``blue'') represent two generic conflicting choices.
+Nous supposons le modèle réseau discuté en section~\ref{sec:model_and_goals}. Soient $\mathtt{R}$ (``rouge'') et $\mathtt{B}$ (``bleu'') représentant deux choix conflictuels génériques.
+%O Without loss of generality, we focus our attention on counts of $\mathtt{B}$, i.e.\ the total number of nodes that prefer blue.
+Sans perdre en généralité, nous portons notre attention sur les décomptes de $\mathtt{B}$, c'est-à-dire le nombre de nœuds qui préfèrent le bleu.
+
+%O \paragraph{Hypergeometric Distribution} Each network query of $k$ peers corresponds to a sample without replacement out of a network of $n$ nodes, also referred to as a hypergeometric sample.
+\paragraph{Distribution hypergéométrique} Chaque requête réseau de $k$ pairs correspond à un échantillon sans remplacement depuis un réseau de $n$ nœuds, également appelé un échantillon hypergéométrique.
+%O We let the random variable $\mathcal{H}(\mathcal{N}, x, k) \rightarrow \{0, \dots, k\}$ denote the resulting counts of $\mathtt{B}$ in the sample (unless otherwise stated), where $x$ is the total count of $\mathtt{B}$ in the population.
 The probability that the query achieves the required threshold of $\alpha$ or more votes is given by:
+La variable aléatoire $\mathcal{H}(\mathcal{N}, x, k) \rightarrow \{0, \dots, k\}$ note les décomptes résultant de $\mathtt{B}$ dans l'échantillon (sauf mention spécifique), où $x$ est le nombre total de $\mathtt{B}$ dans la population.
+La probabilité que la requête atteigne le seuil requis d'au moins $\alpha$ votes est donnée par :
 \begin{equation}
@@
 \end{equation}
-For ease of notation, we overload $\mathcal{H}(*)$ by implicitly referring to $P(\mathcal{H}(\mathcal{N}, x, k) \geq \alpha)$ as $\mathcal{H}(\mathcal{N}, x, k, \alpha)$.
-
-\paragraph{Tail Bounds On Hypergeometric Distribution} We can reduce some of the complexity in Equation~\ref{eq:hypergeometric} by introducing a bound on the hypergeometric distribution induced by $\mathcal{H}^k_{\mathcal{N},x}$.
-Let $p=x/n$ be the ratio of support for $\mathtt{B}$ in the population.
-The expectation of $\mathcal{H}(\mathcal{N}, x, k)$ is exactly $kp$.
-Then, the probability that $\mathcal{H}(\mathcal{N}, x, k)$ will deviate from the mean by more than some small constant $\psi$ is given by the Hoeffding tail bound \cite{hoeffding1963probability}, as follows,
+%O For ease of notation, we overload $\mathcal{H}(*)$ by implicitly referring to $P(\mathcal{H}(\mathcal{N}, x, k) \geq \alpha)$ as $\mathcal{H}(\mathcal{N}, x, k, \alpha)$. 
+Pour faciliter la notation, nous surchargeons $\mathcal{H}(*)$ en nous référant implicitement à $P(\mathcal{H}(\mathcal{N}, x, k) \geq \alpha)$ par $\mathcal{H}(\mathcal{N}, x, k, \alpha)$.
+
+%O \paragraph{Tail Bounds On Hypergeometric Distribution} We can reduce some of the complexity in Equation~\ref{eq:hypergeometric} by introducing a bound on the hypergeometric distribution induced by $\mathcal{H}^k_{\mathcal{N},x}$.
+\paragraph{Bornes de queue sur la distribution hypergéométrique} Nous pouvons un peu réduire la complexité de l'équation~\ref{eq:hypergeometric} en introduisant une borne sur la distribution hypergéométrique induite par $\mathcal{H}^k_{\mathcal{N},x}$.
+%O Let $p=x/n$ be the ratio of support for $\mathtt{B}$ in the population.
+Soit $p=x/n$ la proportion du support pour $\mathtt{B}$ dans la population.
+%O The expectation of $\mathcal{H}(\mathcal{N}, x, k)$ is exactly $kp$.
+Nous attendons exactement $kp$ pour $\mathcal{H}(\mathcal{N}, x, k)$.
+%O Then, the probability that $\mathcal{H}(\mathcal{N}, x, k)$ will deviate from the mean by more than some small constant $\psi$ is given by the Hoeffding tail bound \cite{hoeffding1963probability}, as follows,
+Donc, la probabilité que $\mathcal{H}(\mathcal{N}, x, k)$ dévie de la moyenne de plus d'une petite constante $\psi$ est donnée par la borne de queue de Hoeffding~\cite{hoeffding1963probability}, comme suit,
 \begin{equation}
@@
 \end{equation}
-where $\mathcal{D}(p-\psi, p)$ is the Kullback-Leibler divergence, measured as
+where $\mathcal{D}(p-\psi, p)$ est la divergence de Kullback-Leibler, measurée par
 \begin{equation}
@@
 
-\paragraph{Concentration of Sub-Martingales}
-Let $\{X_{\{t \geq 0\}}\}$ be a sub-martingale and $|X_t - X_{t-1}| < c_t$ almost surely. Then, for all positive reals $\psi$ and all positive integers $t$, 
+\paragraph{Concentration de sous-martingales}
+%O Let $\{X_{\{t \geq 0\}}\}$ be a sub-martingale and $|X_t - X_{t-1}| < c_t$ almost surely. Then, for all positive reals $\psi$ and all positive integers $t$, 
+Soit $\{X_{\{t \geq 0\}}\}$ une sous-martingale et $|X_t - X_{t-1}| < c_t$ presque sûrement. Alors, pour tout réel positif $\psi$ et tout entier positif $t$, 
 \begin{equation}
@@
 %\label{sec:analysis}
-Slush operates in a non-Byzantine setting; that is, $f = 0$, $c = n$.
-In this section, we will characterize the irreversibility properties of Slush (which appear in Snowflake and Snowball), as well as the precise converge rate distribution.
+%O Slush operates in a non-Byzantine setting; that is, $f = 0$, $c = n$.
+Slush opère dans une configuration non-byzantine, c'est-à-dire $f = 0$, $c = n$.
+%O In this section, we will characterize the irreversibility properties of Slush (which appear in Snowflake and Snowball), as well as the precise converge rate distribution.
 The distribution of of both safety and liveness of Slush translate well to the Byzantine setting.
+Dans cette section, nous caractériserons les propriétés d'irréversibilité de Slush (qui apparaissent dans Snowflake et Snowball), ainsi que la distribution de taux de convergence précise.
+La distribution de la sûreté de fonctionnement et de vitalité de Slush se traduisent bien dans une configuration byzantine.
 
@@
 
-The procedural version of Slush in Figure~\ref{fig:slush-loop} made use of a parameter $m$, the number of rounds that a node executes Slush queries. 
+%O The procedural version of Slush in Figure~\ref{fig:slush-loop} made use of a parameter $m$, the number of rounds that a node executes Slush queries. 
+La version procédurale de Slush en figure~\ref{fig:slush-loop} faisait usage d'un paramètre $m$, le nombre de tours d'un nœud exécutant des requêtes Slush.
 % To derive this parameter, we transform the protocol execution from a procedural and concurrent version to one carried out by a scheduler, shown in Figure~\ref{fig:slush_protocol_simulator}.
-What we ultimately want to extract is the total number of rounds $\phi$ that the scheduler will need to execute in order to guarantee that the entire network is the same color, whp.
-
-We analyze the system mainly using a continuous time process. Let $\{X_{\{t \geq 0\}}\}$ be a CTMC.
-The state space $\mathcal{S}$ of the stochastic process is a condensed version of the full configuration space, where each state $\{0, \dots, n\}$ represents the total number of blue nodes in the system.
-
-Let $\mathcal{F}_{X_s}$ be the filtration, or the history pertaining to the process, up to time $s$. This process is Markovian and time-homogeneous, conforming to 
+%O What we ultimately want to extract is the total number of rounds $\phi$ that the scheduler will need to execute in order to guarantee that the entire network is the same color, whp.
+Nous voulons en définitive extraire le nombre total de tours $\phi$ que l'ordonnanceur doit exécuter afin de garantir que l'entièreté du réseau se retrouve de la même couleur, ceci avec une forte probabilité.
+
+%O We analyze the system mainly using a continuous time process. Let $\{X_{\{t \geq 0\}}\}$ be a CTMC.
+Nous analysons le système en utilisant principalement un processus à temps continu. Soit $\{X_{\{t \geq 0\}}\}$ un processus de Markov à temps continu.
+%O The state space $\mathcal{S}$ of the stochastic process is a condensed version of the full configuration space, where each state $\{0, \dots, n\}$ represents the total number of blue nodes in the system.
+L'espace d'état $\mathcal{S}$ du processus stochastique est une version condensée de tout l'espace de configuration, où chaque état $\{0, \dots, n\}$ représente le nombre total de nœuds bleus dans le système.
+
+%O Let $\mathcal{F}_{X_s}$ be the filtration, or the history pertaining to the process, up to time $s$. This process is Markovian and time-homogeneous, conforming to 
+Soit $\mathcal{F}_{X_s}$ la filtration, ou l'historique afférent au processus, jusqu'au temps $s$. Ce processus est Markovien et homogène dans le temps, se conformant à
 \[
@@
 \]
-Throughout the paper, we use $Q \equiv (q_{ij}, i, j \in \mathcal{S})$ notation to refer to the infinitesimal generator of the process, where death ($i \rightarrow i-1$) and birth ($i \rightarrow i+1$) rates of configuration transitions are denoted via $\mu_i$ and $\lambda_i$ ($\lambda_ i$ is distinct from the clock parameter $\lambda$, and will be clear from context).
+%O Throughout the paper, we use $Q \equiv (q_{ij}, i, j \in \mathcal{S})$ notation to refer to the infinitesimal generator of the process, where death ($i \rightarrow i-1$) and birth ($i \rightarrow i+1$) rates of configuration transitions are denoted via $\mu_i$ and $\lambda_i$ ($\lambda_ i$ is distinct from the clock parameter $\lambda$, and will be clear from context).
 These rates are
+Dans cet article, nous utilisons la notation $Q \equiv (q_{ij}, i, j \in \mathcal{S})$ pour nous référer au générateur infinitésimal du processus, où les taux de mort ($i \rightarrow i-1$) et de naissance ($i \rightarrow i+1$) de transitions de configuration sont notées via $\mu_i$, et$\lambda_i$ ($\lambda_ i$ est disctint du paramètre d'horloge $\lambda$ et sera sorti du contexte).
+Ces taux sont
 \[
@@
 \]
-for $1\leq i\leq c-1$, and where $i = 0$ and $i = c$ are absorbing. Let $p_{ij}(t)$ refer to the probability of transitioning from state $i$ to $j$ at time $t$. 
-We always assume that 
+%O for $1\leq i\leq c-1$, and where $i = 0$ and $i = c$ are absorbing. Let $p_{ij}(t)$ refer to the probability of transitioning from state $i$ to $j$ at time $t$. 
+pour $1\leq i\leq c-1$, et où $i = 0$ et $i = c$ sont absorbants. Soit $p_{ij}(t)$ une référence à la probabilité de transition de l'état $i$ à $j$ à l'instant $t$.
+%O We always assume that 
+Nous supposons toujours que
 \[
@@
     \begin{cases}
-      \lambda_it + o(t), & \text{for } j = i + 1 \\
-      \mu_it + o(t), & \text{for } j = i - 1 \\
-      1 - (\lambda_i + \mu_i)t + o(t), & \text{for } j = i \\
-      o(t), & \text{otherwise }\\
+      \lambda_it + o(t), & \text{pour } j = i + 1 \\
+      \mu_it + o(t), & \text{pour } j = i - 1 \\
+      1 - (\lambda_i + \mu_i)t + o(t), & \text{pour } j = i \\
+      o(t), & \text{sinon }\\
     \end{cases}
 \]
-where all $o(t)$ are uniform in $i$.
-
-\paragraph{Irreversibility}
+%O where all $o(t)$ are uniform in $i$. 
+où tous les $o(t)$ sont uniformes en $i$.
+
+%O \paragraph{Irreversibility}
+\paragraph{Irreversibilité}
 %We now discuss the core results of irreversibility properties of our family of protocols.
-In Section~\ref{sec:analysis}, we discussed the loose Chvatal bound which provided intuitive understanding into the strong irreversibility dynamics of our core subsampling mechanism.
+%O In Section~\ref{sec:analysis}, we discussed the loose Chvatal bound which provided intuitive understanding into the strong irreversibility dynamics of our core subsampling mechanism.
 In particular, once the network drifts to some majority value, it tends to revert back with only an exponentially small probability.
 We compute the closed-form expression for reversibility, and show that it is exponentially small.
+En section ~\ref{sec:analysis}, nous avons traité de la borne \emph{loose} % HOULA
+de Chtaval qui a fourni une compréhension intuitive de la dynamique de forte irréversibilité au cœur de notre mécanisme de sous-échantillonnage.
+En particulier, une fois que le réseau tend vers une valeur majoritaire donnée, il ne peut s'inverser qu'avec une probabilité exponentiellement petite.
+Nous calculons l'expression de forme fermée pour la réversibilité, et nous montrons qu'elle est exponentiellement petite.
 \begin{theorem}
 \label{theorem:slush_prob_convergence_minority}
-Let $\xi_\delta$ be the probability of absorption into the all-red state ($s_0$), starting from a drift of $\delta$ (i.e. $\delta$ drift away from $n/2$). Then, assuming $\delta > 1$, 
+%O Let $\xi_\delta$ be the probability of absorption into the all-red state ($s_0$), starting from a drift of $\delta$ (i.e. $\delta$ drift away from $n/2$). Then, assuming $\delta > 1$, 
+Soit $\xi_\delta$ la probabilité d'absorption dans l'état totalement rouge ($s_0$), commençant par un drift de $\delta$ (c'est-à-dire que $\delta$ tendent à s'éloigner de $n/2$).
+Alors, en supposant $\delta > 1$,
 \begin{equation}
@@
 \end{equation}
-and
+et
 % \begin{equation}
@@
 \end{equation}
-where from now on we refer to $\mathcal{u}_{\delta+1}$ as the drift of the process. 
+%O where from now on we refer to $\mathcal{u}_{\delta+1}$ as the drift of the process. 
+où à partir de maintenant nous nous référons à $\mathcal{u}_{\delta+1}$ par le drift du processus.
 \end{theorem}
@@
 \begin{proof}
-Our results are derived based on constructions from Tan~\cite{tan1976absorption}. We construct a sub-matrix of $Q$, denoted $B$, as shown in Figure~\ref{fig:matrixB}.
+%O Our results are derived based on constructions from Tan~\cite{tan1976absorption}. We construct a sub-matrix of $Q$, denoted $B$, as shown in Figure~\ref{fig:matrixB}.
+Nos résultats sont dérivés en se fondant sur les constructions de Tan~\cite{tan1976absorption}. Nous construisons une sous-matrice de $Q$, notée $B$, comme on le voit en figure~\ref{fig:matrixB}.
 \begin{figure*}
@@
 \]
-\caption{Matrix $B$.}
+\caption{Matrice $B$.}
 \label{fig:matrixB}
@@
 \]
-As a reminder, the stationary distribution can be found via $\lim_{t \rightarrow \infty} P(t) = e^{Qt}$, where we have
+%O As a reminder, the stationary distribution can be found via $\lim_{t \rightarrow \infty} P(t) = e^{Qt}$, where we have
+Rappelons que la distribution stationnaire peut être trouvée \emph{via} $\lim_{t \rightarrow \infty} P(t) = e^{Qt}$, où nous avons
 \[
@@
 As Tan (eq. 2.3) shows, we have
+Comme le montre Tan (eq. 2.3), nous avons
 \[
@@
 \]
-Since we want the ultimate probabilities, we have that 
+%O Since we want the ultimate probabilities, we have that 
+Comme nous voulons les probabilités ultimes, nous avons 
 \[
@@
 \]
-We can explicitly compute $\xi_\delta$ in terms of our rates $\mu_i$ and $\lambda_i$, getting 
+%O We can explicitly compute $\xi_\delta$ in terms of our rates $\mu_i$ and $\lambda_i$, getting 
+Nous pouvons explicitement calculer $\xi_\delta$ en termes de nos taux $\mu_i$ et $\lambda_i$, obtenant ainsi
 \[
@@
 \]
-However, we note that $u_{i} = \lambda_{n-i}$.
+%O However, we note that $u_{i} = \lambda_{n-i}$.
 Algebraic manipulation from this observation leads to the two equations in the theorem.
 This expression is strictly lower than the Chvatal bounds used in Section~\ref{sec:analysis}.
+Cependant, notons que $u_{i} = \lambda_{n-i}$.
+La manipulation algébrique à partir de cette observation mène aux deux équations du théorème.
+Cette expression est strictement inférieure aux bornes de Chtaval utilisées en section~\ref{sec:analysis}.
 \end{proof}
 
-Using the construction for the absorption (and (ir)reversibility) probabilities as discussed previously, a natural follow up computation is in regards to \emph{mean convergence time}. 
-Let $T_{z}(t) = \inf \{t \geq 0 : X_t = \{0, n\} | X_0 = z\}$, and let $\tau_z = \mathbb{E}[T_{z}(t)]$.
+%O Using the construction for the absorption (and (ir)reversibility) probabilities as discussed previously, a natural follow up computation is in regards to \emph{mean convergence time}. 
+En utilisant la construction pour les probabilités d'absorption (et d'(ir)réversibilité) précédemment traitées, il est naturel de poursuivre par un calcul concernant le \emph{temps de convergence moyen}.
+%O Let $T_{z}(t) = \inf \{t \geq 0 : X_t = \{0, n\} | X_0 = z\}$, and let $\tau_z = \mathbb{E}[T_{z}(t)]$.
 $\tau_z$ is the mean time to reach either absorbing state, starting from state $z$, which corresponds to the mean convergence time.
 The next theorem characterizes this distribution.
+Soit $T_{z}(t) = \inf \{t \geq 0 : X_t = \{0, n\} | X_0 = z\}$, et soit $\tau_z = \mathbb{E}[T_{z}(t)]$.
+$\tau_z$ est le temps moyen pour atteindre l'un ou l'autre des états absorbants, à partir de l'état $z$, qui correspond au temps moyen de convergence.
+Le théorème suivant caractérise cette distribution.
 
@@
 \label{theorem:mean-convergence-time}
-Let $\tau_z$ be the expected time to convergence, starting from state $z > n/2$, to any of the two converging states in the network (all-red or all-blue). Then, 
+%O Let $\tau_z$ be the expected time to convergence, starting from state $z > n/2$, to any of the two converging states in the network (all-red or all-blue). Then, 
+Soit $\tau_z$ le temps attendu de convergence, à partir de l'état $z > n/2$, vers l'un des deux états convergents dans le réseau (totalement rouge ou totalement bleu). Alors,
 \begin{equation}
@@
 \end{equation}
-where $x(d)$ and $y(d)$ are
+%O where $x(d)$ and $y(d)$ are
+où $x(d)$ et $y(d)$ sont
 \begin{equation}
@@
 \begin{proof}
-Following the calculations from before, $-B^{-1}$ at row $z$ provides the number of traversals to each other state starting from $z$.
+%O Following the calculations from before, $-B^{-1}$ at row $z$ provides the number of traversals to each other state starting from $z$.
 Calculating their sum, we have our result.
 The above equation is the full expression of the matrix row sum.
+D'après les calculs qui précèdent, $-B^{-1}$ à la rangée $z$ fournit le nombre de traversées vers chaque autre état en partant de $z$.
+En calculant leur somme, nous avons notre résultat.
+L'équation ci-dessus est l'expression complète de la somme des rangées des matrices.
 \end{proof}
 
-Theorem~\ref{theorem:mean-convergence-time} leads to the next lemma that captures property P2, under the assumption that at the beginning of the protocol, one proposal has at least $\alpha$ support in the network.
+%O Theorem~\ref{theorem:mean-convergence-time} leads to the next lemma that captures property P2, under the assumption that at the beginning of the protocol, one proposal has at least $\alpha$ support in the network.
+Le théorème~\ref{theorem:mean-convergence-time} mène au lemme suivant qui capture la propriété P2, en suppposant qu'au début du protocole, une proposition a un support d'au moins $\alpha$ dans le réseau.
 \begin{lemma}
-Slush reaches an absorbing state in finite time almost surely.
+%O Slush reaches an absorbing state in finite time almost surely.
+Slush atteint presque sûrement un état absorbant dans un temps fini.
 \label{lemma:finitetermination}
@@
 \begin{proof}
-Starting from any non-absorbing, transient state, there is a non-zero probability of being absorbed.
+%O Starting from any non-absorbing, transient state, there is a non-zero probability of being absorbed.
 Additionally, since termination is finite and everywhere differentiable, Theorem~\ref{theorem:mean-convergence-time} also implies that the probability of termination of any network configuration where a proposal has $\geq \alpha$ support in bounded time $t_{max}$ is strictly positive.
+En partant d'un quelconque état non-absorbant et transitoire, il y a une probabilité non nulle d'absorbtion.
+De plus, comme la terminaison est finie et différentiable partout, le théorème~\ref{theorem:mean-convergence-time} implique aussi que la probabilité de terminaison de toute configuration réseau où une proposition a un support $\geq \alpha$ dans un temps borné $t_{max}$ est strictement positive.
 \end{proof}

### 2020-11-09 paper_fr.tex
Typos A.1 A.2
Analysis - A3

@@
 %O Consensus protocols provide their guarantees based on assumptions that only a fraction of participants are adversarial.
-Les protocoles de consensus fournissent leurs garanties fondées sur la supposition que seule une fraction des participants sont malveillants.
+Les protocoles de consensus fournissent leurs garanties fondées sur la supposition que seule une fraction des participants sont antagonistes.
 %O These bounds could be violated if the network is naively left open to arbitrary participants.
@@
 %O \paragraph{Irreversibility}
-\paragraph{Irreversibilité}
+\paragraph{Irréversibilité}
 %We now discuss the core results of irreversibility properties of our family of protocols.
@@
 \]
-As Tan (eq. 2.3) shows, we have
+%O As Tan (eq. 2.3) shows, we have
 Comme le montre Tan (eq. 2.3), nous avons
@@
 %O Theorem~\ref{theorem:mean-convergence-time} leads to the next lemma that captures property P2, under the assumption that at the beginning of the protocol, one proposal has at least $\alpha$ support in the network.
-Le théorème~\ref{theorem:mean-convergence-time} mène au lemme suivant qui capture la propriété P2, en suppposant qu'au début du protocole, une proposition a un support d'au moins $\alpha$ dans le réseau.
+Le théorème~\ref{theorem:mean-convergence-time} mène au lemme suivant qui reflète % HOULA
+la propriété P2, en suppposant qu'au début du protocole, une proposition possède un support d'au moins $\alpha$ dans le réseau.
 \begin{lemma}
@@
 \label{subsection:appendix_snowflake}
-In Snowflake, the sampled set of nodes includes Byzantine nodes.
-We introduce the decision function $\mathcal{D}(*)$, which is constructed by having each node also keep track of the total number of consecutive times it has sampled a majority of the same color ($\beta$).
-Finally, we introduce a function called $\mathcal{A}(\mathcal{S}_t)$, the adversarial strategy, that takes as parameters the entire configuration of the network at time $t$, as well as the next set of nodes chosen by the scheduler to execute, and as a side-effect, modifies the set of nodes $\mathcal{B}$ to some arbitrary configuration of colors.
-
-In order for our prior framework to apply to Snowflake, we must deal with a key subtlety. 
-Unlike in Slush, where it is clear that once the network has reached one of the converging states and therefore may not revert back, this no longer applies to Snowflake, since any adversary $f \geq \alpha$ has strictly positive probability of reverting the system, albeit this probability may be infinitesimally small.
-The CTMC is flexible enough to deal with a system where there is only one absorbing state, but the long-term behavior of the system is no longer meaningful since, after an infinite amount of time, the system is guaranteed to revert, violating safety.
-We could trivially bound the amount of time, and show safety using this bounded time assumption by simply characterizing the distribution of $e^{tQ}$, where $Q$ is the generator. 
-However, we can make the following observation: if the probability of going from state $c$ (all-blue) to $c-1$ is exponentially small, then it will take the attacker exponential time (in expectation; note, this is a lower bound, and in reality it will take much longer) to succeed in reverting the system.
-Hence, we can assume that once all correct nodes are the same color, the attack from the adversary will terminate since it is impractical to continue an attack. 
-In fact, under reasonably bounded timeframes, the variational distance between the exact approach and the approximation is very small. 
-We leave details to the accompanying paper, but we briefly discuss how analysis proceeds for Snowflake.
-
-As stated in Section~\ref{sec:analysis}, the way to analyze the adversary using the same construction as in Slush is to condition reversibility on the first node $u$ deciding on blue, which can happen at any state (as specified by $\mathcal{D}(*)$).
+%O In Snowflake, the sampled set of nodes includes Byzantine nodes.
+Dans Snowflake, l'ensemble des nœuds échantillonnés comprend des nœuds byzantins.
+%O We introduce the decision function $\mathcal{D}(*)$, which is constructed by having each node also keep track of the total number of consecutive times it has sampled a majority of the same color ($\beta$).
+Nous introduisons finalement une fonction appelée $\mathcal{A}(\mathcal{S}_t)$ qui est construite en faisant conserver à chaque nœud la trace du nombre total de fois consécutives qu'il a échantillonné une majorité de la couleur d'échantillon ($\beta$).
+%O Finally, we introduce a function called $\mathcal{A}(\mathcal{S}_t)$, the adversarial strategy, that takes as parameters the entire configuration of the network at time $t$, as well as the next set of nodes chosen by the scheduler to execute, and as a side-effect, modifies the set of nodes $\mathcal{B}$ to some arbitrary configuration of colors.
+Enfin, nous introduisons une fonction appelée $\mathcal{A}(\mathcal{S}_t)$, la stratégie contradictoire, qui prend comme paramètres toute la configuration du réseau à l'instant $t$, ainsi que l'ensemble suivant de nœuds choisi par l'ordonnanceur pour exécution, avec pour effet secondaire la modification de l'ensemble de nœuds $\mathcal{B}$ vers une configuration arbitraire de couleurs.
+
+%O In order for our prior framework to apply to Snowflake, we must deal with a key subtlety. 
+Afin d'appliquer notre framework antérieur à Snowflake, nous devons porter notre attention sur une subtilité essentielle.
+%O Unlike in Slush, where it is clear that once the network has reached one of the converging states and therefore may not revert back, this no longer applies to Snowflake, since any adversary $f \geq \alpha$ has strictly positive probability of reverting the system, albeit this probability may be infinitesimally small.
+Contrairement à Slush, où il est clair qu'une fois que le réseau a atteint l'un des états convergents et ne reviendra donc pas en arrière, cela ne s'applique pas à Snowflake, puisque tout adversaire $f \geq \alpha$ a une probabilité strictement positive d'inverser la tendance du système, fût-elle infinitésimale.
+%O The CTMC is flexible enough to deal with a system where there is only one absorbing state, but the long-term behavior of the system is no longer meaningful since, after an infinite amount of time, the system is guaranteed to revert, violating safety.
+Le processus de Markov à temps continu est assez souple pour s'accommoder d'un système où il n'existe qu'un seul état absorbant, mais le comportement du système sur le long terme n'est plus significatif puisque, après un laps de temps infini, le système doit s'inverser, violant ainsi la sécurité de fonctionnement.
+%O We could trivially bound the amount of time, and show safety using this bounded time assumption by simply characterizing the distribution of $e^{tQ}$, where $Q$ is the generator. 
+Nous pouvons borner trivialement l'intervalle de temps et montrer la sûreté de fonctionnement grâce à ce bornage en nous contentant de caractériser la distribution de $e^{tQ}$, où $Q$ est le générateur.
+%O However, we can make the following observation: if the probability of going from state $c$ (all-blue) to $c-1$ is exponentially small, then it will take the attacker exponential time (in expectation; note, this is a lower bound, and in reality it will take much longer) to succeed in reverting the system.
+Nous pouvons cependant faire l'observation suivante : si la probabilité d'aller de l'état $c$ (totalement bleu) à $c-1$ est exponentiellement petite, alors il faudra à l'attaquant un temps exponentiel (attendu\,; on note que c'est une borne inférieure et, en réalité, il faudra beaucoup plus longtemps) pour réussir à inverser l'état du système.
+%O Hence, we can assume that once all correct nodes are the same color, the attack from the adversary will terminate since it is impractical to continue an attack. 
+Nous pouvons donc suppposer qu'une fois que tous les nœuds corrects sont de la même couleur, l'attaque de l'adversaire se termine puisqu'il serait irréaliste de la poursuivre.
+
+%O In fact, under reasonably bounded timeframes, the variational distance between the exact approach and the approximation is very small. 
+En fait, dans un cadre temporel raisonnablement borné, la distance variationnelle entre l'approche exacte et l'approximation est très petite.
+%O We leave details to the accompanying paper, but we briefly discuss how analysis proceeds for Snowflake. 
+Nous laissons les détails à l'article qui accompagne celui-ci, mais nous discutons brièvement du déroulement de l'analyse pour Snowflake.
+
+%O As stated in Section~\ref{sec:analysis}, the way to analyze the adversary using the same construction as in Slush is to condition reversibility on the first node $u$ deciding on blue, which can happen at any state (as specified by $\mathcal{D}(*)$).
 At that point, the adversarial strategy collapses to a single function, which is to continually vote for red.
@@
 The birth and death rates are transformed as follows:
+Comme il en a été fait mention en section~\ref{sec:analysis}, la manière d'analyser l'adversaire en employant la même construction que dans Slush consiste à conditionner la réversibilité sur le premier nœud $u$ se décidant pour le bleu, ce qui peut se produire à n'importe quel état (comme le spécifie $\mathcal{D}(*)$).
+À ce point, la stratégie contradictoire se réduit à une seule fonction, qui est de continuellement voter pour rouge.
+Les probabilités de réversibilité, pour tous les états $\{1, \dots, c-1\}$ doivent coder la probabilité que des nœuds bleus s'engagent ansi que la fonction unique de l'adversaire.
+Les taux de naissance et de mort sont transformés comme suit\,:
 \[
@@
 \]
-From here on, the analysis is the same as in Slush. Under various $k$ and $\beta$, we can find the minimal $\alpha$ that provides the system strong irreversibility properties.
-
-The next lemma captures P3, and the proof follows from central limit theorem. 
+%O From here on, the analysis is the same as in Slush. Under various $k$ and $\beta$, we can find the minimal $\alpha$ that provides the system strong irreversibility properties. 
+À partir de maintenant, l'analyse est la même que pour Slush. Pour divers $k$ et $\beta$, nous pouvons trouver le $\alpha$ minimal qui fournit au système des propriétés de forte irréversibilité.
+
+%O The next lemma captures P3, and the proof follows from central limit theorem. 
+Le lemme suivant reflète P3 et la preuve procède du théorème central limite.
 \begin{lemma}
-If $f < \Oh{\sqrt{n}}$, and $\alpha = \floor{k/2} + 1$, then Snowflake terminates in $\Oh{\log n}$ rounds with high probability. 
+%O If $f < \Oh{\sqrt{n}}$, and $\alpha = \floor{k/2} + 1$, then Snowflake terminates in $\Oh{\log n}$ rounds with high probability. 
+Si $f < \Oh{\sqrt{n}}$, et $\alpha = \floor{k/2} + 1$, alors Snowflake se termine en $\Oh{\log n}$ tours avec une forte probabilité. 
 \label{lemma:centrallimit}
@@
 \begin{proof}
-The results follows from central limit theorem, wherein for $\alpha = \floor{k/2} + 1$, the expected bias in the network after sampling will be $\Oh{\sqrt{n}}$.
+%O The results follows from central limit theorem, wherein for $\alpha = \floor{k/2} + 1$, the expected bias in the network after sampling will be $\Oh{\sqrt{n}}$.
 An adversary smaller than this bias will be unable to keep the network in a fully-bivalent state for more than a constant number of rounds.
 The logarithmic factor remains from the mixing time lower bound.
+Le résultat procède du théorème central limite, où pour $\alpha = \floor{k/2} + 1$, la polarisation % HOULA
+attendue dans le réseau après échantillonnage sera $\Oh{\sqrt{n}}$.
+Un adversaire plus petit que ce résultat sera incapable de conserver le réseau dans un état totalement bivalent pendant plus qu'un nombre constant de tours.
+Le facteur logarithmique provient de la limite inférieure du temps de mélange.
+% HOULA
 \end{proof}

### 2020-11-10 paper_fr.tex
Analysis - complet

@@
 \end{figure*}
-Let $W_1'$ = $(\mu_1, 0, \dots, 0)$, $W_{n-1}'$ = $(0, \dots, 0, \lambda_{n-1})$. Then, we can express $Q$ as 
+%O Let $W_1'$ = $(\mu_1, 0, \dots, 0)$, $W_{n-1}'$ = $(0, \dots, 0, \lambda_{n-1})$. Then, we can express $Q$ as 
+Soit $W_1'$ = $(\mu_1, 0, \dots, 0)$, $W_{n-1}'$ = $(0, \dots, 0, \lambda_{n-1})$. Alors nous pouvons exprimer $Q$ par
 \[
@@
 %O Hence, we can assume that once all correct nodes are the same color, the attack from the adversary will terminate since it is impractical to continue an attack. 
-Nous pouvons donc suppposer qu'une fois que tous les nœuds corrects sont de la même couleur, l'attaque de l'adversaire se termine puisqu'il serait irréaliste de la poursuivre.
+Nous pouvons donc suppposer qu'une fois tous les nœuds corrects de la même couleur, l'attaque de l'adversaire se termine puisqu'il serait irréaliste de la poursuivre.
 
@@
 À ce point, la stratégie contradictoire se réduit à une seule fonction, qui est de continuellement voter pour rouge.
-Les probabilités de réversibilité, pour tous les états $\{1, \dots, c-1\}$ doivent coder la probabilité que des nœuds bleus s'engagent ansi que la fonction unique de l'adversaire.
+Les probabilités de réversibilité, pour tous les états $\{1, \dots, c-1\}$ doivent coder la probabilité que des nœuds bleus s'engagent, anisi que la fonction unique de l'adversaire.
 Les taux de naissance et de mort sont transformés comme suit\,:
@@
 \subsection{Snowball}
-We make the following observation: if the confidences between red and blue are equal, then the adversary has the same identical leverage in the irreversibility of the system as in Snowflake, regardless of network configuration.
+%O We make the following observation: if the confidences between red and blue are equal, then the adversary has the same identical leverage in the irreversibility of the system as in Snowflake, regardless of network configuration.
 In fact, Snowflake can be viewed as Snowball but where drifts in confidences never exceed one.
@@
 The guarantees that can be extracted hereon are that the confidences of the majority committed value (in our frame of reference is always blue), grow always more than those of the minority value, with high probability, drifting away as $t \rightarrow t_{max}$.
-\subsection{Safe Early Commitment}
-As we reasoned previously, each conflict set in Avalanche can be viewed as an instance of Snowball, where each progeny instance iteratively votes for the entire path of the ancestry.
-This feature provides various benefits; however, it also can lead to some virtuous transactions that depend on a rogue transaction to suffer the fate of the latter.
+Nous faisons l'observation suivante : si la confiance entre rouge et bleu est équivalente, alors l'adversaire a même levier % HOULA
+sur l'irréversibilité du système que dans Snowflake, quelle que soit la configuration du réseau. En fait, Snowflake peut être vu comme Snowball où les drifts % HOULA
+de confiance n'excèdent jamais un.
+La même analyse s'applique à Snowball comme à Snowflake, avec la contrainte supplémentaire de bornage des comportements des confiances à long terme sur le réseau.
+L'analyse suit à cette fin en emplouant des inégalités de concentration de martingales, notamment celle qui a été introduite en équation~\ref{eq:submartingale}.
+Snowball peut être vu comme un système à deux urnes où chaque urne est une sous-martingale.
+Les garanties qui peuvent en être extraites ci-après sont que les confiances de la valeur engagée par la majorité (toujours bleu dans notre cadre de référence) croissent toujours plus que celles de la valeur minoritaire, avec une forte probabilité, tendant à s'écarter alors que $t \rightarrow t_{max}$.
+% HOULA TOUT LE PARAGRAPHE
+%O \subsection{Safe Early Commitment}
+\subsection{Engagement préalable sûr}
+%O As we reasoned previously, each conflict set in Avalanche can be viewed as an instance of Snowball, where each progeny instance iteratively votes for the entire path of the ancestry.
+Comme nous l'avons précédemment déduit, chaque ensemble de conflits dans Avalanche peut être vu comme une instance de Snowball où chaque instance de la progéniture vote itérativement pour tout le chemin de l'ascendance.
+%O This feature provides various benefits; however, it also can lead to some virtuous transactions that depend on a rogue transaction to suffer the fate of the latter.
+Cette caractéristique offre plusieurs avantages\,; cependant, elle peut aussi mener à ce que des transactions vertueuses dépendant d'une transaction malveillante souffrent le sort de celle-ci.
 %has the drawback that it can entangle the fate of some unfortunate virtuous transactions with the fate of rogue ones.
-In particular, rogue transactions can interject in-between virtuous transactions and reduce the ability of the virtuous transactions to ever reach the required $\textsc{isAccepted}$ predicate.
-As a thought experiment, suppose that a transaction $T_i$ names a set of parent transactions that are all decided, as per local view.
-If $T_i$ is sampled over a large enough set of successful queries without discovering any conflicts, then, since by assumption the entire ancestry of $T_i$ is decided, it must be the case (probabilistically) that we have achieved irreversibility.
-
-To then statistically measure the assuredness that $T_i$ has been accepted by a large percentage of correct nodes without any conflicts, we make use of a one-way birth process, where a birth occurs when a new correct node discovers the conflict of $T_i$.
+%O In particular, rogue transactions can interject in-between virtuous transactions and reduce the ability of the virtuous transactions to ever reach the required $\textsc{isAccepted}$ predicate.
+En particulier, les transactions malveillantes peuvent s'insérer entre des transactions vertueuses et réduire la possibilité que celles-ci atteignent le prédicat $\textsc{isAccepted}$ nécessaire.
+%O As a thought experiment, suppose that a transaction $T_i$ names a set of parent transactions that are all decided, as per local view.
+En guise d'expérience mentale, supposons qu'une transaction $T_i$ nomme un ensemble de transactions parent qui sont toutes décidées, selon la vue locale.
+%O If $T_i$ is sampled over a large enough set of successful queries without discovering any conflicts, then, since by assumption the entire ancestry of $T_i$ is decided, it must be the case (probabilistically) that we have achieved irreversibility.
+Si $T_i$ est échantillonnée sur un ensemble suffisamment grand de requêtes réussies découvrir aucun conflit, alors, puisque par supposition toute l'ascendance de $T_i$ est décidée, il s'ensuit (de manière probabiliste) que nous sommes arrivés à l'irréversibilité.
+
+%O To then statistically measure the assuredness that $T_i$ has been accepted by a large percentage of correct nodes without any conflicts, we make use of a one-way birth process, where a birth occurs when a new correct node discovers the conflict of $T_i$.
 Necessarily, deaths cannot exist in this model, because a conflicting transaction cannot be unseen once a correct node discovers it.
+Pour ensuite mesurer statistiquement l'assurance que $T_i$ a été acceptée par un grand pourcentage de nœuds corrects sans aucun conflit, nous nous servons d'un processus de naissance à sens unique, où une naissance se produit quand une nouveau nœud correct découvre le conflit de $T_i$.
+Les morts ne peuvent pas exister dans ce modèle, car une transaction conflictuelle ne peut pas être escamotée une fois qu'un nœud correct la découvre.
 % Let $t = 0$ be the time when $T_j$, which conflicts with $T_i$, is introduced to a single correct node $u$.
 % Let $s_x$, for $x = 1$ to $c$, be the state where the number of correct nodes that know about $T_j$ is $x$, and let $p(s_x)$ be the probability of birth at state $s_x$. Then, we have:
-Our births are as follows:
+%O Our births are as follows:
+Nos naissances se font comme suit\,:
 \begin{equation}
@@
 \end{equation}
-Solving for the expected time to reach the final birth state provides a lower bound to the $\beta_1$ parameter in the $\textsc{isAccepted}$ fast-decision branch.
+%O Solving for the expected time to reach the final birth state provides a lower bound to the $\beta_1$ parameter in the $\textsc{isAccepted}$ fast-decision branch.
 The table below shows an example of the analysis for $n = 2000$, $\alpha = 0.8k$, and various $k$, where $\varepsilon \ll 10^{-9}$, and where $\beta$ is the minimum required value before deciding.
+La résolution du temps attendu pour atteindre l'état final de naissance fournit une borne inférieure au paramètre $\beta_1$ dans la branche de décision rapide $\textsc{isAccepted}$.
+La table ci-dessous montre un exemple d'analyse pour $n = 2000$, $\alpha = 0.8k$ et divers $k$, où $\varepsilon \ll 10^{-9}$, et où $\beta$ est la valeur minimale requise avant la décision.
 \begin{table}[h!]
@@
 % \end{figure}
-\noindent Overall, a very small number of iterations are sufficient for the safe early commitment predicate. This supports the choice of $\beta$ in our evaluation.
-
-\subsection{Initialization Heuristic}
+%O \noindent Overall, a very small number of iterations are sufficient for the safe early commitment predicate. This supports the choice of $\beta$ in our evaluation.
+
+\noindent Globalement, un très petit nombre d'itérations est suffisant pour le prédicat d'engagement préalable sûr. Cela confirme le choix de $\beta$ dans notre évaluation.
+
+%O \subsection{Initialization Heuristic}
+\subsection{Heuristique d'initialisation}
 \label{sec:sync-heuristic}
-To improve liveness, we can use strong synchrony assumptions.
+%O To improve liveness, we can use strong synchrony assumptions.
 The heuristic works as follows.
@@
 It then selects the first value by the order as its initial state when it starts the actual consensus protocol.
-In a cryptocurrency setting, the deterministic ordering function would incorporate fees paid out for every new proposal, which means that the adversary is financially limited in its ability to launch a fairness attack against the initialization.
-
-\subsection{Churn and View Discrepancy}\label{sec:full-analysis-churn}
-
-Realistic systems need to accommodate the departure and arrival of nodes.
-Up to now, we simplified our analysis by assuming a precise knowledge of network membership, i.e. $\mathcal{L}(u) = \mathcal{N}$.
-We now demonstrate that correct nodes can admit a well-characterized amount of churn, by showing how to pick parameters such that Avalanche nodes can differ in their view of the network and still safely make decisions.
+Pour améiorer la vitalité, nous pouvons utiliser des hypothèses de forte synchronie.
+L'heuristique fonctionne comme suit.
+Chaque nœud opère en deux phases\,: dans la première, elle diffuse et collecte des propositions pendant $\Oh{\log{n}}$ tours, où chaque tour dure le temps maximum d'un message, ce qui garantit que la proposition d'un nœud correct se propage vers presque tous les autres nœuds corrects\,; dans la seconde phase, chaque nœud arrête de collecter les propositions dont il a localement la connaissance, en recherchant ceux qui ont une majorité $\alpha$, ordonnée de manière déterministe, comme par la valeur de leur empreinte cryptographique.
+Il sélectionne ensuite la première valeur dans l'ordre pour en faire son état initial quand il commence vraiment le protocole de consensus.
+%O In a cryptocurrency setting, the deterministic ordering function would incorporate fees paid out for every new proposal, which means that the adversary is financially limited in its ability to launch a fairness attack against the initialization.
+Dans une configuration de cryptomonnaie, la fonction de tri déterministe incorporerait les frais payés pour toute nouvelle proposition, ce qui signifie que l'adversaire serait financièrement limité pour lancer une attaque d'impartialité contre l'initialisation.
+
+%O \subsection{Churn and View Discrepancy}\label{sec:full-analysis-churn}
+\subsection{Roulement et incohérence de vues}\label{sec:full-analysis-churn}
+
+%O Realistic systems need to accommodate the departure and arrival of nodes.
+Les systèmes réalistes doivent s'accommoder du départ et de l'arrivée des nœuds.
+%O Up to now, we simplified our analysis by assuming a precise knowledge of network membership, i.e. $\mathcal{L}(u) = \mathcal{N}$.
+Jusqu'à présent, nous avons simplifié notre analyse en supposant une connaissance précise des membres du réseau, c'est-à-dire $\mathcal{L}(u) = \mathcal{N}$.
+%O We now demonstrate that correct nodes can admit a well-characterized amount of churn, by showing how to pick parameters such that Avalanche nodes can differ in their view of the network and still safely make decisions.
+Nous allons maintenant démontrer que les nœuds corrects peuvent admettre un roulement caractérisé d'entrants et de sortants en montrant comment choisir des paramètres tels que les nœuds Avalanche puissent différer dans leurs vues respectives du réseau tout en maintenant leur capacité à prendre des décisions en toute sécurité.
 
@@
 
-To characterize churn we use a generalized set intersection construction that allows us to make arguments about worst-case network view splits.
+%O To characterize churn we use a generalized set intersection construction that allows us to make arguments about worst-case network view splits.
 Before formalizing it, we provide the intuition: suppose we split the network into two entirely independent, but fully connected, subsets.
@@
 This represents the worst case view split.
+Pour caractériser le roulement, nous utilisons une construction d'intersections d'ensembles généralisés qui nous permet d'établir des arguments concernant les éclatements de vues du réseau.
+Avant de le formaliser, nous fournissons cette intuition\,: supposons que nous éclations le réseau en deux sous-ensembles entièrement indépendants mais parfaitement connectés.
+Il est clair que l'adversaire byzantin gagne avec une probabilité de un puisqu'il peut envoyer deux transactions conflictuelles à chaque réseau indépendant, et chacun finaliserait immédiatement les transactions.
+Cela représente le cas d'éclatement le plus défavorable.
 \begin{figure}
@@
 \includegraphics[width=0.8\linewidth]{figures/network_view.pdf}
-\caption{Changes in network view based on $S_d$'s size.
+%O \caption{Changes in network view based on $S_d$'s size.
 All prior proofs up to now represent the variant where $S_a = S_b = \emptyset$.
 With the new construction, the probability of safety violation is simply a direct application of the prior sets of proofs under the new subsets.}
+\caption{Changements de la vue du réseau en se basant sur la taille de $S_d$.
+Toutes les preuves antérieures jusqu'à présent représentent la variante où $S_a = S_b = \emptyset$.
+Avec la nouvelle construction, la probabilité de violation de la sécurité de fonctionnement n'est qu'une application directe des ensembles antérieurs de preuves pour les nouveaux sous-ensembles.}
 \label{fig:network_view}
 \end{figure}
-We can generalize this to arbitrary network splits, which can even be applied recursively in each subset.
+%O We can generalize this to arbitrary network splits, which can even be applied recursively in each subset.
 The proofs of safety then are a matter of characterizing the probability of red and blue committing into the two (possibly independent) subsets.
-
-Suppose we divide the set of correct nodes into three subsets, $S_a$, $S_d$, $S_b$.
+Nous pouvons généraliser ceci à d'autres éclatements du réseau, et même récursivement dans chaque sous-ensemble.
+Les preuves de sûreté de fonctionnement sont alors une question de caractérisation de la probabilité que les rouges et les bleus s'engagent dans les deux sous-ensembles (éventuellement indépendants).
+
+%O Suppose we divide the set of correct nodes into three subsets, $S_a$, $S_d$, $S_b$.
 We overload $\mathcal{L}(S_{*})$ to represent the views of any node within the input set.
@@
 This construction is visually demonstrated in Figure~\ref{fig:network_view}.
+Supposons que nous divisions l'ensemble de nœuds corrects en trois sous-ensembles, $S_a$, $S_d$, $S_b$.
+Nous surchargeons $\mathcal{L}(S_{*})$ pour représenter les vues de tout nœud dans l'ensemble d'entrées.
+La vue de tous les nœuds dans $S_a$ est $\mathcal{L}(S_a) = S_a \cup S_d \cup \mathcal{B}$, celle de tous les nœuds dans $S_b$ est $\mathcal{L}(S_b) = S_b \cup S_d \cup \mathcal{B}$, et la vue de $S_d$ est $\mathcal{N}$.
+Nous supposons le cas le plus défavorable, ce qui signifie que les nœuds de l'adversaire sont communs à tous les sous-ensembles.
+Quand $S_d = \emptyset$, cela représente une division du réseau en deux sous-ensembles égaux où $|S_a| = |S_b| = n/2$~\footnote{$|S_a|$ et $|S_b|$ ne sont pas nécessairement égaux, nous posons ceci pour la démonstration.}.
+Si $S_d$ est uniquement composé de nœuds corrects, alors $|S_a| = |S_b| = n$.
+Cette construction est démontrée visuellement en figure~\ref{fig:network_view}.
 
 \begin{lemma}
-Let $\tau \in \mathbb{Z}^{+}$.
+%O Let $\tau \in \mathbb{Z}^{+}$.
 Let $|S_d| = n - \tau$, and thus $|S_{\{a, b\}}| = \tau/2$.
 There exists some maximal size of $\tau$ such that probability of any two nodes $u, v \in S_a,\ S_b,\ S_d$ finalizing equivocating transactions is less than $\epsilon$.
+Soit $\tau \in \mathbb{Z}^{+}$.
+Soit $|S_d| = n - \tau$, et donc $|S_{\{a, b\}}| = \tau/2$.
+Il existe un taille maximale de $\tau$ telle que la probabilité pour que deux nœuds quelconques $u, v \in S_a,\ S_b,\ S_d$ finalisant des transactions équivoques soit inférieuire à $\epsilon$.
 \end{lemma}
 \begin{proof}
-We assume that the adversary has full control of the network view splits, meaning that they choose in full how to create $S_a,\ S_b,\ S_d$.
+%O We assume that the adversary has full control of the network view splits, meaning that they choose in full how to create $S_a,\ S_b,\ S_d$.
 To prove safety, we simply reuse the same exact construction as in Subsection~\ref{subsection:appendix_snowflake}, but we replace the original set $\mathcal{N}$ with a new set of interest, namely $S_a \cup S_d \cup \mathcal{B}$ (i.e., we exclude $S_b$)~\footnote{Sets are symmetric in this example.}.
 To thus find the maximal $\tau$, we simply replace $u_i$ and $\lambda_i$ with
+Nous supposons que l'adversaire a le contrôle total des éclatements de la vue du réseau, ce qui signifie qu'il peut choisir sans contrainte de créer $S_a,\ S_b,\ S_d$.
+Pour prouver la sécurité de fonctionnement, il nous suffit de réutiliser exactement la même construction qu'en section~\ref{subsection:appendix_snowflake}, mais nous remplaçons l'ensemble originel $\mathcal{N}$ avec le nouvel ensemble qui nous intéresse, nommément $S_a \cup S_d \cup \mathcal{B}$ (c'est-à-dire que nous excluons $S_b$)~\footnote{Les ensembles sont symétriques dans cet exemple.}.
+Donc, pour trouver les $\tau$ maximal, nous nous contentons de remplacer $u_i$ et $\lambda_i$ par
 \begin{equation}
@@
 \end{equation}
-where
+où
 \begin{equation}
@@
 \end{equation}
-and apply the same construction as in Subsection~\ref{subsection:appendix_snowflake}.
+%O and apply the same construction as in Subsection~\ref{subsection:appendix_snowflake}.
 As $\tau$ increases, it is clear that $S_a$ (and conversely, $S_b$), become more independent and less reliant on the values proposed by members in $S_d$, thus incrementing the ability of the adversaries.
+et nous appliquons la même construction qu'en section~\ref{subsection:appendix_snowflake}.
+Alors que $\tau$ augmente, il est clair que $S_a$ (et, réciproquement, $S_b$) deviennent plus indépendant et dépendent moins des valeurs par les membres de $S_d$, augmentant ainsi les capacités de l'adversaire.
 \end{proof}

### 2020-11-14 paper_fr.tex
Conclusion

@@
 \label{sec:conclusions}
-This paper introduced a novel family of consensus protocols, coupled with the appropriate mathematical tools for analyzing them.
-\tronly{These protocols are highly efficient and robust, combining the best features of classical and Nakamoto consensus.}
-They scale well, achieve high throughput and quick finality, work without precise membership knowledge, and degrade gracefully under catastrophic adversarial attacks.
-
-There is much work to do to improve this line of research. \tronly{
-One such improvement could be the introduction of an adversarial network scheduler.
-Another}{One} improvement would be to characterize the system's guarantees under an adversary whose powers are realistically limited, whereupon performance would improve even further.
+%O This paper introduced a novel family of consensus protocols, coupled with the appropriate mathematical tools for analyzing them.
+Cet article a présenté une nouvelle famille de protocoles de consensus, accompagnée des outils mathématiques appropriés pour leur analyse.
+%O \tronly{These protocols are highly efficient and robust, combining the best features of classical and Nakamoto consensus.}
+\tronly{Ces protocoles sont extrêmement efficaces et robustes, et rassemblent les meilleures fonctionnalités des consensus classiques et Nakamoto.}{} 
+%O They scale well, achieve high throughput and quick finality, work without precise membership knowledge, and degrade gracefully under catastrophic adversarial attacks.
+Ils passent bien à l'échelle, permettent un haut débit et une finalité rapide, ne nécessitent pas une connaissance précise des participants et leur fonctionnement ne se dégrade que peu à peu lors d'attaques catastrophiques.
+
+%O There is much work to do to improve this line of research. \tronly{
+Il reste beaucoup à faire dans ce domaine de recherche. \tronly{
+%O One such improvement could be the introduction of an adversarial network scheduler.
+Une amélioration possible consisterait à introduire un ordonnanceur de réseaux adverses. % HOULA
+%O Another}{One} improvement would be to characterize the system's guarantees under an adversary whose powers are realistically limited, whereupon performance would improve even further.
 \tronly{Finally, more}{More} sophisticated initialization mechanisms would bear fruitful in improving liveness of multi-value consensus.
-Overall, we hope that the protocols and analysis techniques presented here add to the arsenal of the distributed system developers and provide a foundation for new lightweight and scalable mechanisms.
+Une autre}{Une} serait de caractériser les garanties du système face à un adversaire dont les pouvoirs seraient limités de manière réaliste, à la suite de quoi les performances se verraient encode améliorées.
+\tronly{D'autres}{Plus} mécanismes d'initialisation sophistiqués seraient souhaitables pour améliorer la vitalité d'un consensus multi-valeurs.
+%O Overall, we hope that the protocols and analysis techniques presented here add to the arsenal of the distributed system developers and provide a foundation for new lightweight and scalable mechanisms.
+En résumé, nous espérons que les protocoles et les techniques d'analyse présentés ici renforceront l'arsenal des développeurs de systèmes distribués et permettront de bâtir de nouveaux mécanismes légers et évolutifs.
 
@@
 La même analyse s'applique à Snowball comme à Snowflake, avec la contrainte supplémentaire de bornage des comportements des confiances à long terme sur le réseau.
-L'analyse suit à cette fin en emplouant des inégalités de concentration de martingales, notamment celle qui a été introduite en équation~\ref{eq:submartingale}.
+L'analyse suit à cette fin en employant des inégalités de concentration de martingales, notamment celle qui a été introduite en équation~\ref{eq:submartingale}.
 Snowball peut être vu comme un système à deux urnes où chaque urne est une sous-martingale.

### 2020-11-14 sections/related-work_fr.tex
French translation of the related work part

@@
 \tronly{
-Bitcoin~\cite{nakamoto2008bitcoin} is a cryptocurrency that uses a blockchain based on
-proof-of-work (PoW) to maintain a ledger of UTXO transactions. 
-While techniques based on proof-of-work~\cite{DworkN92, aspnes2005exposing}, and even cryptocurrencies with minting based on proof-of-work~\cite{vishnumurthy2003karma,rivest1997payword}, have been explored before, Bitcoin was the first to incorporate PoW into its consensus process.
-Unlike more traditional BFT protocols, Bitcoin has a probabilistic safety guarantee
-and assumes honest majority computational power rather than a known
-membership, which in turn has enabled an internet-scale permissionless protocol. While permissionless and resilient to adversaries,
-Bitcoin suffers from low throughput (\textasciitilde{}3 tps) and
-high latency (\textasciitilde{}5.6 hours for a network with 20\% Byzantine presence and $2^{-32}$ security guarantee).
+Bitcoin~\cite{nakamoto2008bitcoin} est une cryptomonnaie qui utilise une blockchain basée sur la preuve de travail
+(PoW pour "Proof Of Work") et qui tient un livre de comptes des "sorties de transaction non dépensées", ou UTXO
+("Unspent Transaction Output") en anglais. Bien que des techniques basées sur la preuve de
+travail~\cite{DworkN92, aspnes2005exposing}, et même d'autres cryptomonnaies dont la génération de pièces est basée
+sur la preuve de travail~\cite{vishnumurthy2003karma,rivest1997payword}, ont été explorées avant l'apparition de
+Bitcoin, ce dernier fut le premier à incorporer la preuve de travail dans son processus de consensus. Contrairement
+aux protocoles résistants aux fautes byzantines traditionnels, Bitcoin apporte une garantie de sûreté probabiliste et
+considère que la majorité de la puissance de calcul agit de manière honnête plutôt que de contrôler les membres du
+réseau, ce qui a permis de faire émerger un protocole sans permission à l'échelle d'internet. Bien qu'il soit sans
+permission et résilient face aux attaques, Bitcoin souffre d'un débit de transaction faible (\textasciitilde{}3 tps)
+et d'une latence élevée (\textasciitilde{}5.6 heures pour un réseau comportant 20\% de noeuds byzantins et une garantie
+de sûreté de $2^{-32}$). De plus la preuve de travail nécéssite une puissance de calcul considérable qui est consommée
+dans le seul et unique but de maintenir la sûreté de fonctionnement.
+
+%0 Bitcoin~\cite{nakamoto2008bitcoin} is a cryptocurrency that uses a blockchain based on
+%0 proof-of-work (PoW) to maintain a ledger of UTXO transactions.
+%0 While techniques based on proof-of-work~\cite{DworkN92, aspnes2005exposing}, and even cryptocurrencies with minting based on proof-of-work~\cite{vishnumurthy2003karma,rivest1997payword}, have been explored before, Bitcoin was the first to incorporate PoW into its consensus process.
+%0 Unlike more traditional BFT protocols, Bitcoin has a probabilistic safety guarantee
+%0 and assumes honest majority computational power rather than a known
+%0 membership, which in turn has enabled an internet-scale permissionless protocol. While permissionless and resilient to adversaries,
+%0 Bitcoin suffers from low throughput (\textasciitilde{}3 tps) and
+%0 high latency (\textasciitilde{}5.6 hours for a network with 20\% Byzantine presence and $2^{-32}$ security guarantee).
 Furthermore, PoW requires a substantial amount of computational power that is consumed only for the purpose of maintaining
-safety.
+%0 safety.
 
-% PoW family
-Countless cryptocurrencies use PoW~\cite{DworkN92, aspnes2005exposing} to maintain a distributed ledger. 
-Like Bitcoin, they suffer from inherent scalability bottlenecks. 
-Several proposals for protocols exist that try to
-better utilize the effort made by PoW.
-Bitcoin-NG~\cite{EyalGSR16} and the permissionless version of
-Thunderella~\cite{PassS18} use Nakamoto-like consensus to elect a leader that
-dictates writing of the replicated log for a relatively long time so as to
-provide higher throughput. Moreover, Thunderella provides an
-optimistic bound that, with 3/4 honest computational power and an honest
-elected leader, allows transactions to be confirmed rapidly.
-ByzCoin~\cite{Kokoris-KogiasJ16} periodically selects a small set of
-participants and then runs a PBFT-like protocol within the selected nodes.
-%It achieves a throughput of 942 tps with about 35 second latency.
+D'innombrables cryptomonnaies utilisent la preuve de travail~\cite{DworkN92, aspnes2005exposing} dans le but de tenir
+un livre de comptes distribué. Tout comme Bitcoin, elles souffrent des mêmes limitations lorsqu'il s'agit de monter en
+charge. Il existe plusieurs propositions de protocoles essayant de faire meilleur usage des progrès apportés par la
+preuve de travail. Bitcoin-NG~\cite{EyalGSR16} et la version sans permission de Thunderella~\cite{PassS18} utilise
+un consensus similaire à celui de Nakamoto pour élire un leader qui va dicter les écritures dans le journal répliqué
+pour un temps donné assez long dans le but de fournir un meilleur débit de transactions. En outre, Thunderella apporte
+une borne optimiste qui, en considérant 3/4 de puissance de calcul et un leader élu honnêtes, permet de confimer les
+transactions rapidement. ByzCoin~\cite{Kokoris-KogiasJ16} sélectionne un échantillon réduit de participants de manière
+périodique et exécute ensuite un protocole résistant aux fautes byzantines en pratique (PBFT) entre les noeuds
+sélectionnés.
 
-% Byzantine Agreement family
-Protocols based on Byzantine agreement~\cite{PeaseSL80, LamportSP82} typically make use of
-quorums and require precise knowledge of membership.
-PBFT~\cite{castro1999practical, CL02}, a well-known representative, requires a quadratic number of message exchanges in order to
-reach agreement. 
-%Some variants are able to scale to dozens of nodes~\cite{ClementWADM09, KotlaADCW09}.
-The Q/U protocol~\cite{abd2005fault} and HQ replication~\cite{cowling2006hq} use a quorum-based approach to optimize for contention-free cases of operation to achieve consensus in only a single round of communication.
+%O % PoW family
+%O Countless cryptocurrencies use PoW~\cite{DworkN92, aspnes2005exposing} to maintain a distributed ledger.
+%O Like Bitcoin, they suffer from inherent scalability bottlenecks.
+%O Several proposals for protocols exist that try to
+%O better utilize the effort made by PoW.
+%O Bitcoin-NG~\cite{EyalGSR16} and the permissionless version of
+%O Thunderella~\cite{PassS18} use Nakamoto-like consensus to elect a leader that
+%O dictates writing of the replicated log for a relatively long time so as to
+%O provide higher throughput. Moreover, Thunderella provides an
+%O optimistic bound that, with 3/4 honest computational power and an honest
+%O elected leader, allows transactions to be confirmed rapidly.
+%O ByzCoin~\cite{Kokoris-KogiasJ16} periodically selects a small set of
+%O participants and then runs a PBFT-like protocol within the selected nodes.
+%O %It achieves a throughput of 942 tps with about 35 second latency.
+
+Les protocoles basés sur un accord byzantin~\cite{PeaseSL80, LamportSP82} se basent généralement sur un quorum et
+nécéssitent une connaissance précise des membres du réseau. PBFT~\cite{castro1999practical, CL02}, l'un de ces
+protocoles les plus connus, nécessite un nombre quadratique d'échanges de messages avant de parvenir à un accord.
+Le protocole Q/U~\cite{abd2005fault} et la réplication HQ~\cite{cowling2006hq} utilisent une approche basée sur un
+quorum afin d'optimiser le protocole pour obtenir des cas de fonctionnement sans contention et parvenir à un consensus
+en un seul tour de communication. Malgré tout, même si ces protocoles améliorent les performances, ils se dégradent
+fortement lorsqu'ils sont soumis à une forte contention. Zyzzyva~\cite{KotlaADCW09} ajoute au protocole BFT une
+exécution spéculative pour tendre vers un cas de fonctionnement sans erreurs. Les travaux passés portant sur les
+systèmes résistants aux fautes byzantines avec contrôle d'accès nécessitent typiquement au moins $f+1$ répliques.
+CheapBFT~\cite{kapitza2012cheapbft} tire profit de composants matériels certifiés pour construire un protocole
+nécessitant exactement $f+1$ répliques.
+
+%O % Byzantine Agreement family
+%O Protocols based on Byzantine agreement~\cite{PeaseSL80, LamportSP82} typically make use of
+%O quorums and require precise knowledge of membership.
+%O PBFT~\cite{castro1999practical, CL02}, a well-known representative, requires a quadratic number of message exchanges in order to
+%O reach agreement.
+%O %Some variants are able to scale to dozens of nodes~\cite{ClementWADM09, KotlaADCW09}.
+%O The Q/U protocol~\cite{abd2005fault} and HQ replication~\cite{cowling2006hq} use a quorum-based approach to optimize for contention-free cases of operation to achieve consensus in only a single round of communication.
 However, although these protocols improve on performance, they degrade very poorly under contention.
 Zyzzyva~\cite{KotlaADCW09} couples BFT with speculative execution to improve the failure-free operation case.
-%Aliph~\cite{guerraoui2010next} introduces a protocol with optimized performance under several, rather than just one, cases of execution.
+%O %Aliph~\cite{guerraoui2010next} introduces a protocol with optimized performance under several, rather than just one, cases of execution.
 In contrast, Ardvark~\cite{ClementWADM09} sacrifices some performance to tolerate worst-case degradation, providing a more uniform execution profile.
 This work, in particular, sacrifices failure-free optimizations to provide consistent throughput even at high number of failures.
-Past work in permissioned BFT systems typically requires at least $3f+1$ replicas.
+%O Past work in permissioned BFT systems typically requires at least $3f+1$ replicas.
 CheapBFT~\cite{kapitza2012cheapbft} leverages trusted hardware components to construct a protocol that uses $f+1$ replicas.
 
-Other work attempts to introduce new protocols under redefinitions and relaxations of the BFT model. 
-Large-scale BFT~\cite{rodrigues2007large} modifies PBFT to allow for arbitrary choice of number of replicas and failure threshold, providing a probabilistic guarantee of liveness for some failure ratio but protecting safety with high probability.
-In another form of relaxation.
+D'autres travaux tentent d'introduire de nouveaux protocoles qui redéfinissent et relâchent les contraintes du modèle
+de résistance aux fautes byzantines. La résistance byzantine à grande échelle~\cite{rodrigues2007large} modifie la
+résistance byzantine pratique (PBFT) pour autoriser un nombre arbitraire de répliques et un seuil d'échec, en
+fournissant une garantie de vitalité probabilitse pour un taux d'échec donné tout en protégeant la sûreté de
+fonctionnement avec une probabilité élevée. Dans une autre forme de relâchement des contraintes,
+Zeno~\cite{singh2009zeno} introduit un protocole de réplication de machine d'état résistante aux fautes byzantines qui
+échange la cohérence pour une meilleure disponibilité. Plus spécifiquement, Zeno garantit une cohérence assurée au final
+plutôt qu'une linéarisation de celle-ci, ce qui veut dire que les participants n'ont pas besoin d'être conhérents mais
+doivent se mettrent d'accord au final une fois le réseau stabilisé. En fournissant une garantie de cohérence encore
+plus faible, appelée "cohérence causale par embranchement-raccord", Depot~\cite{mahajan2011depot} définit un protocole
+garantissant la sûreté de fonctionnement pour $2f+1$ répliques.
+
+NOW~\cite{guerraoui2013highly} uttilise des sous-quorums pour diriger des instances de consensus réduites.
+L'enseignement de ce papier est que de petits quorums de taille logarithmique peuvent être extraits à partir d'un
+échantillon potentiellement large de noeuds du réseau, ce qui permet à des instances de protocole de consensus plus
+petites de tourner en parallèle.
+
+Snow White~\cite{cryptoeprint:2016:919} et Ouroboros~\cite{KiayiasRDO17} font partie des premiers protocoles à la
+sûreté prouvée basés sur la preuve d'enjeu (PoS - ou "Proof Of Stake" en anglais). Ouroboros utilise un protocole de
+bascule de jeton multi-parties sécurisé dans le but d'incorporer un caractère aléatoire dans l'élection du leader.
+Son évolution Ouroboros Praos~\cite{DavidGKR18} apporte la sûreté en présence d'attaquants entièrement adaptatifs.
+HoneyBadger~\cite{MillerXCSS16} démontre une bonne vitalité dans un réseau comportant des latences hétérogènes.
+
+Tendermint~\cite{buchman2016tendermint, 1807.04938} change de leader à chaque bloc et a démontré ses qualités dans une
+configuration à 64 noeuds. Ripple~\cite{schwartz2014ripple} a une faible latence grâce à l'utilisation de sous-réseaux
+certifiés collectivement au sein d'un réseau plus large. La société Ripple génère une liste de noeuds vérifiés par
+défaut renouvelée peu souvent, qui en fait un réseau essentiellement centralisé.
+
+HotStuff~\cite{hotstuff,hotstuffpodc} améliore les coûts de communication d'une distribution quadratique vers une
+distribution linéaire et simplifie drastiquement les caractéristiques du protocole, même si les limitations liées
+à l'élection d'un leader persistent. Facebook utilise Hotstuff comme le consensus principal de son projet Libra.
+
+Dans une configuration synchrone, inspirée par Hoststuff, Sync HotStuff~\cite{synchotstuff} atteint un consensus en
+un temps $2\Delta$ avec un coût quadratique et contrairement à d'autres protocoles synchrones à étapes vérouillées, il
+opère aussi rapidement que le réseau se propage. Stellar~\cite{mazieres2015stellar} se base sur un accord byzantin
+fédéré dans lequel des \emph{tranches de quorum} permettent une confiance hétérogène entre des noeuds différents. La
+sûreté de fonctionnement est garantie lorsque les transactions peuvent être reliées entre elles de manière transitive
+par des tranches de quorum. Algorand~\cite{GiladHMVZ17} utilise une fonction aléatoire vérifiable pour sélectionner un
+comité de noeuds qui participent à un protocole de consensus byzantin innovant.
+
+%0 Other work attempts to introduce new protocols under redefinitions and relaxations of the BFT model.
+%0 Large-scale BFT~\cite{rodrigues2007large} modifies PBFT to allow for arbitrary choice of number of replicas and failure threshold, providing a probabilistic guarantee of liveness for some failure ratio but protecting safety with high probability.
+%0 In another form of relaxation.
 Zeno~\cite{singh2009zeno} introduces a BFT state machine replication protocol that trades consistency for high availability.
@@
 By providing an even weaker consistency guarantee, namely fork-join-causal consistency, Depot~\cite{mahajan2011depot} describes a protocol that guarantees safety under $2f+1$ replicas.
-
-NOW~\cite{guerraoui2013highly} uses sub-quorums to drive smaller instances of consensus.
+%0
+%0 NOW~\cite{guerraoui2013highly} uses sub-quorums to drive smaller instances of consensus.
 The insight of this paper is that small, logarithmic-sized quorums can be extracted from a potentially large set of nodes in the network, allowing smaller instances of consensus protocols to be run in parallel.
 
-Snow White~\cite{cryptoeprint:2016:919} and
-Ouroboros~\cite{KiayiasRDO17} are some of the earliest provably secure PoS
-protocols.  Ouroboros uses a secure multiparty coin-flipping protocol to
-produce randomness for leader election. The follow-up protocol, Ouroboros
-Praos~\cite{DavidGKR18} provides safety in the presence of fully adaptive
-adversaries.
-HoneyBadger~\cite{MillerXCSS16} provides good liveness in a network with heterogeneous latencies. %and achieves over 819 tps with 5 minute latency on 104 nodes.
+%O Snow White~\cite{cryptoeprint:2016:919} and
+%O Ouroboros~\cite{KiayiasRDO17} are some of the earliest provably secure PoS
+%O protocols.  Ouroboros uses a secure multiparty coin-flipping protocol to
+%O produce randomness for leader election. The follow-up protocol, Ouroboros
+%O Praos~\cite{DavidGKR18} provides safety in the presence of fully adaptive
+%O adversaries.
+%O HoneyBadger~\cite{MillerXCSS16} provides good liveness in a network with heterogeneous latencies. %and achieves over 819 tps with 5 minute latency on 104 nodes.
 
-Tendermint~\cite{buchman2016tendermint, 1807.04938} rotates the leader for each block
-and has been demonstrated with as many as 64 nodes. Ripple~\cite{schwartz2014ripple} has low latency by utilizing collectively-trusted
-sub-networks in a large network. The Ripple company provides a
-slow-changing default list of trusted nodes, which renders the system essentially centralized.
-HotStuff~\cite{hotstuff,hotstuffpodc} improves the communication cost from quadratic to linear and significantly simplifies the protocol specification, although the leader bottleneck still persists.
+%O Tendermint~\cite{buchman2016tendermint, 1807.04938} rotates the leader for each block
+%O and has been demonstrated with as many as 64 nodes. Ripple~\cite{schwartz2014ripple} has low latency by utilizing collectively-trusted
+%O sub-networks in a large network. The Ripple company provides a
+%O slow-changing default list of trusted nodes, which renders the system essentially centralized.
+%O HotStuff~\cite{hotstuff,hotstuffpodc} improves the communication cost from quadratic to linear and significantly simplifies the protocol specification, although the leader bottleneck still persists.
 Facebook uses HotStuff as the core consensus for its Libra project.
-% Ittai paper
-In the synchronous setting, inspired by HotStuff, Sync HotStuff~\cite{synchotstuff} achieves consensus in $2\Delta$ time with quadratic cost and unlike other lock-steped synchronous protocols, it operates as fast as network propagates.
-Stellar~\cite{mazieres2015stellar} uses Federated Byzantine Agreement in which \emph{quorum slices}
-enable heterogeneous trust for different nodes.  Safety is guaranteed when
-transactions can be transitively connected by trusted quorum slices.
-Algorand~\cite{GiladHMVZ17} uses a verifiable random function to select a
-committee of nodes that participate in a novel Byzantine consensus
-protocol.
+%O % Ittai paper
+%O In the synchronous setting, inspired by HotStuff, Sync HotStuff~\cite{synchotstuff} achieves consensus in $2\Delta$ time with quadratic cost and unlike other lock-steped synchronous protocols, it operates as fast as network propagates.
+%O Stellar~\cite{mazieres2015stellar} uses Federated Byzantine Agreement in which \emph{quorum slices}
+%O enable heterogeneous trust for different nodes.  Safety is guaranteed when
+%O transactions can be transitively connected by trusted quorum slices.
+%O Algorand~\cite{GiladHMVZ17} uses a verifiable random function to select a
+%O committee of nodes that participate in a novel Byzantine consensus
+%O protocol.
 %It achieves over 874 tps with 50 second latency on
@@
 
-Some protocols use a Directed Acyclic Graph (DAG) structure instead of a linear chain to achieve
-consensus~\cite{SompolinskyZ15,SompolinskyLZ16,SompolinskyZ18,BentovHMN17,baird2016hashgraph}.
-Instead of choosing the longest chain as in Bitcoin,
-GHOST~\cite{SompolinskyZ15} uses a more efficient chain selection rule that
-allows transactions not on the main chain to be taken into consideration, increasing efficiency.
-SPECTRE~\cite{SompolinskyLZ16} uses transactions on
-the DAG to vote recursively with PoW to achieve consensus, followed up by
-PHANTOM~\cite{SompolinskyZ18} that achieves a linear order among all blocks.
-Like PHANTOM, Conflux also finalizes a linear order of transactions by PoW
-in a DAG structure, with better resistance to liveness attack~\cite{confluxLLXLC18}.
-Avalanche
-is different in that the voting result is a one-time chit that is determined by
-a query without PoW, while the votes in PHANTOM or Conflux are purely determined by PoW in transaction structure.
-Similar to Thunderella, Meshcash~\cite{BentovHMN17} combines a slow PoW-based protocol with a fast consensus protocol that allows a high block rate regardless of network latency, offering fast confirmation time.
-Hashgraph~\cite{baird2016hashgraph} is a leader-less protocol that builds a DAG via randomized gossip.
-It requires full membership knowledge at all times, and, similar to the Ben-Or~\cite{ben1983another}, it requires exponential messages~\cite{aspnes2003randomized,CachinV17} in expectation.
+Certains protocoles utilisent une structure de graphe orienté acyclique (DAG) en lieu et place d'une chaîne linéaire
+pour atteindre un consensus~\cite{SompolinskyZ15,SompolinskyLZ16,SompolinskyZ18,BentovHMN17,baird2016hashgraph}.
+Au lieu de choisir la chaîne la plus longue comme dans Bitcoin, GHOST~\cite{SompolinskyZ15} définit des règles de
+sélection de la chaîne qui permettent de prendre en considération des transactions non présentes sur la chaîne
+principale, ce qui augmente son efficacité. SPECTRE~\cite{SompolinskyLZ16} utilise des transactions sur le DAG pour
+voter de manière récursive via la preuve de travail pour atteindre un consensus, suivi par
+SPECTRE~\cite{SompolinskyLZ16} qui obtient un ordre linéaire à partir de tous les blocs. Tout comme PHANTOM, Conflux
+obtient également un ordre linéaire entre les transactions via la preuve de travail utilisée sur une structure DAG,
+avec une meilleure résistance aux attaques à la vitalité~\cite{confluxLLXLC18}.
+
+Avalanche est différent dans son approche par laquelle le résultat du vote est un reçu instantané déterminé par une
+requête sans preuve de travail, tandis que les votes dans PHANTOM ou Conflux sont purement déterminés par la preuve de
+travail dans la structure de transactions. A la manière de Thunderella, Meshcash~\cite{BentovHMN17} combine un protocole
+lent basé sur la preuve de travail et un protocole de consensus rapide qui apporte une fréquence de génération de blocs
+élevée quelque soit la latence du réseau, offrant un temps de confirmation rapide. Hashgraph~\cite{baird2016hashgraph}
+est un protocole sans leader qui construit un DAG sur la base d'un "bavardage rendu aléatoire". Il nécessite une
+connaissance complète des membres du réseau à tout moment, et comme Ben-Or~\cite{ben1983another}, il s'attend à un
+nombre de messages exponentiel~\cite{aspnes2003randomized,CachinV17} pour fonctionner.
+
+%O Some protocols use a Directed Acyclic Graph (DAG) structure instead of a linear chain to achieve
+%O consensus~\cite{SompolinskyZ15,SompolinskyLZ16,SompolinskyZ18,BentovHMN17,baird2016hashgraph}.
+%O Instead of choosing the longest chain as in Bitcoin,
+%O GHOST~\cite{SompolinskyZ15} uses a more efficient chain selection rule that
+%O allows transactions not on the main chain to be taken into consideration, increasing efficiency.
+%O SPECTRE~\cite{SompolinskyLZ16} uses transactions on
+%O the DAG to vote recursively with PoW to achieve consensus, followed up by
+%O PHANTOM~\cite{SompolinskyZ18} that achieves a linear order among all blocks.
+%O Like PHANTOM, Conflux also finalizes a linear order of transactions by PoW
+%O in a DAG structure, with better resistance to liveness attack~\cite{confluxLLXLC18}.
+%O Avalanche
+%O is different in that the voting result is a one-time chit that is determined by
+%O a query without PoW, while the votes in PHANTOM or Conflux are purely determined by PoW in transaction structure.
+%O Similar to Thunderella, Meshcash~\cite{BentovHMN17} combines a slow PoW-based protocol with a fast consensus protocol that allows a high block rate regardless of network latency, offering fast confirmation time.
+%O Hashgraph~\cite{baird2016hashgraph} is a leader-less protocol that builds a DAG via randomized gossip.
+%O It requires full membership knowledge at all times, and, similar to the Ben-Or~\cite{ben1983another}, it requires exponential messages~\cite{aspnes2003randomized,CachinV17} in expectation.
 }{

### 2020-11-14 avalanche_fr.tex
Relecture jusqu'à l'annexe

@@
 %O The DAG has a single sink that is the \emph{genesis vertex}.
-Le DAG possède un nœud terminal (\emph{sink}) unique qui est le \emph{genesis vertex}, le sommet genèse. % HOULA sink ? 
+Le DAG possède un nœud terminal (\emph{sink}) unique qui est le \emph{genesis vertex}, le sommet genèse.
 %O Maintaining a DAG provides two significant benefits.
-L'emploi d'un DAG procure deux avantages :
+L'emploi d'un DAG procure deux avantages\,:
 %O First, it improves efficiency, because a single vote on a DAG vertex implicitly votes for all tran
 %O sactions on the path to the genesis vertex.
-D'abord, il est plus efficient car un vote sur un sommet vote implicitement pour toutes les transactions sur le chemin vers le sommet genèse.
+D'abord, il est plus efficient car un vote sur un sommet vote implicitement pour toutes les transactions se trouvant sur le chemin du sommet genèse.
 %O Second, it also improves security, because the DAG intertwines the fate of transactions, similar to the Bitcoin blockchain.
@@
 %O The parent-child relationships encoded in the DAG may, but do not need to, correspond to application-specific dependencies; for instance, a child transaction need not spend or have any relationship with the funds received in the parent transaction.
-Les relations parent-enfant codées dans le DAG peuvent, sans obligation, correspondre à des dépendances spécifiques aux applications ; par exemple, la transaction d'un enfant ne dépense pas nécessairement, ni n'a forcément de relation avec les fonds reçus lors de la transaction parent.
+Les relations parent-enfant codées dans le DAG peuvent, sans obligation, correspondre à des dépendances spécifiques aux applications\,; par exemple, la transaction d'un enfant ne dépense pas nécessairement, ni n'a forcément de relation avec les fonds reçus lors de la transaction parent.
 % These additional relationships entangle the fate of previous decisions made by the system.
@@
 Dans notre application de cryptomonnaie, les transactions qui dépensent les mêmes fonds (\emph{double-dépense}) entrent en conflit et forment un \emph{ensemble de conflits}
-(régions ombrées en Figure~\ref{fig:dag-cd}), duquel une seule peut être acceptée.
+(régions ombrées en figure~\ref{fig:dag-cd}), duquel une seule peut être acceptée.
 %O Note that the conflict set of a virtuous transaction is always a singleton.
@@
 %O Nodes then compute their \emph{confidence} as the total number of chits in the progeny of that transaction.
-Les nœuds calculent ensuite leur \emph{confiance} en comptant le nombre de \emph{chits} dans la progéniture de cette transaction.
+Les nœuds calculent ensuite leur \emph{confiance} en comptant le nombre de chits dans la progéniture de cette transaction.
 %O They query a transaction just once and rely on new vertices and possible chits, added to the progeny, to build up their confidence.
-Ils n'envoient qu'une seule requête et se basent sur les nouveaux sommets les \emph{chits} possibles, ajoutés à la progéniture, pour fonder leur confiance.
+Ils n'envoient qu'une seule requête et se basent sur les nouveaux sommets les chits possibles, ajoutés à la progéniture, pour fonder leur confiance.
 %O Ties are broken by an initial preference for first-seen transactions.
@@
 %O the attacker generates large, padded subgraphs.
-On note que les \emph{chits} sont découplés de la structure du DAG, rendant le protocol immune aux attaques où un attaquant génère de grands sous-graphes faussement remplis.
+On note que les chits sont découplés de la structure du DAG, rendant le protocole immune aux attaques où un attaquant génère de grands sous-graphes remplis de fausses valeurs.
 
@@
 %O partitioned into mutually exclusive conflict sets $\mathcal{P}_T$, $T \in \mathcal{T}_u$.
-Chaque nœud correct $u$ garde la trace de toutes les transactions dont il a été informé dans l'ensemble $\mathcal{T}_u$, partitionné en ensembles de conflits mutuellement exclusifs $\mathcal{P}_T$, $T \in \mathcal{T}_u$.
+Chaque nœud correct $u$ conserve la trace de toutes les transactions dont il a été informé dans l'ensemble $\mathcal{T}_u$, partitionné en ensembles de conflits mutuellement exclusifs $\mathcal{P}_T$, $T \in \mathcal{T}_u$.
 % Transactions are determined to be conflicting based on a deterministic function known to every node.
@@
 Comme les conflits sont transitifs, si $T_i$ et $T_j$ sont en conflit, ils appartiennent au même ensemble de conflits, c'est-à-dire $\mathcal{P}_{T_i} = \mathcal{P}_{T_j}$.
-Cette relation peut sembler contre-intuitive : les transactions en conflit ont la même relation d'\emph{équivalence} car ce sont des équivoques dépensant les \emph{mêmes} fonds.
+Cette relation peut sembler contre-intuitive\,: les transactions en conflit ont la même relation d'\emph{équivalence} car ce sont des équivoques dépensant les \emph{mêmes} fonds.
 %\Jon{By $=$ surely you mean $\neq$?}
@@
 %O Each node $u$ can compute a confidence value, $d_u(T)$, from the progeny as follows:
-Chaque nœud $u$ peut calculer une valeur de confiance $d_u(T)$ de sa progéniture comme suit :
+Chaque nœud $u$ peut calculer une valeur de confiance $d_u(T)$ de sa progéniture comme suit\,:
 \[ d_u(T) = \sum_{T' \in \mathcal{T}_u, T \stackrel{*}{\gets} T'}c_{uT'}\]
@@
 %O the query results. If the node collects a threshold of $\alpha$ yes-votes after the query, the value $c_{uT'}$ is set to 1, otherwise remains $0$ forever.
-où $c_{uT'}$ représente la valeur de \emph{chit} de $T'$ pour le nœud $u$.
-Chaque transaction a un valeur de \emph{chit} à $0$ avant que le nœud obtienne les résultats des requêtes.
+où $c_{uT'}$ représente la valeur de chit de $T'$ pour le nœud $u$.
+Chaque transaction a un valeur de chit à $0$ avant que le nœud obtienne les résultats des requêtes.
 Si le nœud récolte un seuil de $\alpha$ votes «\,oui\,» après la requête, la valeur $c_{uT'}$ est mise à 1, sinon elle reste éternellement à $0$.
 %O Therefore, a chit value reflects the result from the one-time query of its associated transaction and becomes immutable afterwards, while $d(T)$ can increase as the DAG grows by collecting more chits in its progeny.
-Donc, une valeur de \emph{chit} reflète le résultat de la requête à un moment donné de ses transactions associées et devient immuable après cela, alors que $d(T)$ peut augmenter au fur et à mesure de la croissance du DAG en collectant d'autres \emph{chit} dans sa progéniture.
+Donc, une valeur de chit reflète le résultat de la requête à un moment donné de ses transactions associées et devient immuable après cela, alors que $d(T)$ peut augmenter au fur et à mesure de la croissance du DAG en collectant d'autres chits dans sa progéniture.
 %O Because $c_T \in \{0, 1\}$, confidence values are monotonic.
@@
 %O The above process will yield a labeling of the DAG with a chit value and associated confidence for each transaction~$T$.
-Le processus ci-dessus ramène un étiquetage du DAG avec une valeur de \emph{chit} et une confiance associée pour chaque transaction~$T$.
+Le processus ci-dessus ramène un étiquetage du DAG avec une valeur de chit et une confiance associée pour chaque transaction~$T$.
 
@@
 %O Figure~\ref{fig:dag-cd} illustrates a sample DAG built by Avalanche.
-La Figure~\ref{fig:dag-cd} illustre un exemple de DAG construit par Avalanche.
+La figure~\ref{fig:dag-cd} illustre un exemple de DAG construit par Avalanche.
 %O Similar to Snowball, sampling in Avalanche will create a positive feedback for the preference of a single transaction in its conflict set.
-De même que pour Snowball, l'échantillonnage dans Avalanche crée une rétroaction positive pour la préférence d'une seule transaction dans sont ensemble de conflit.
+De même que pour Snowball, l'échantillonnage dans Avalanche crée une rétroaction positive pour la préférence d'une seule transaction dans son ensemble de conflits.
 %O For example, because $T_2$ has larger confidence than $T_3$, its descendants are more likely collect chits in the future compared to $T_3$.
-Par exemple, comme $T_2$ a un indice de confiance supérieur à celui de $T_3$, ses descendants seront plus à même de récolter des \emph{chits} à l'avenir que $T_3$.
+Par exemple, comme $T_2$ a un indice de confiance supérieur à celui de $T_3$, ses descendants seront plus à même de récolter des chits à l'avenir que $T_3$.
 
@@
 For virtuous transactions, $T$ is accepted when it is the only transaction in its conflict set and has a confidence not less than threshold $\beta_1$.
-L'engagement concernant une transaction peut être effectié par un \emph{engagement préalable sûr}.
-Pour les transactions vertueuses, $T$ est acceptée quand elle est la seule transaction dans son ensemble de conflit et que sa confiance n'est pas inférieure à un seuil $\beta_1$.
+L'engagement concernant une transaction peut être effectué par un \emph{engagement préalable sûr}.
+Pour les transactions vertueuses, $T$ est acceptée quand elle est la seule transaction dans son ensemble de conflits et que sa confiance n'est pas inférieure à un seuil $\beta_1$.
 %O As in Snowball, $T$ can also be accepted after a $\beta_2$ number of consecutive successful queries.
@@
 %O with parents, it could be accepted if reissued with different parents.
-Si une transaction vertueuse échoue à être acceptée en raison d'un problème avec ses parents elle peut être acceptée si elle est réémise avec des parents différents.
+Si une transaction vertueuse échoue à être acceptée en raison d'un problème avec ses parents, elle peut être acceptée si elle est réémise avec des parents différents.
 %O Figure~\ref{fig:gossipchain-ongen} shows how Avalanche entangles transactions.
 Because transactions that consume and generate the same UTXO do not conflict with each other, any transaction can be reissued with different parents.
-La Figure~\ref{fig:gossipchain-ongen} montre comment Avalanche lie les transactions entre elles.
+La figure~\ref{fig:gossipchain-ongen} montre comment Avalanche lie les transactions entre elles.
 Puisque les transactions qui consomment et génèrent la même UTXO n'entrent pas en conflit entre elles, toute transaction peut être réémise avec des parents différents.
@@
 %O executed by each node.
-La Figure~\ref{fig:gossipchain-main} illustre la boucle principale du protocole exécutée par chaque nœud.
+La figure~\ref{fig:gossipchain-main} illustre la boucle principale du protocole exécutée par chaque nœud.
 %O In each iteration, the node attempts to select a transaction $T$ that has not
@@
 %O If more than $\alpha$ of those peers return a positive response, the chit value is set to~1.
-Si plus que $\alpha$ de ces pairs reviennent à une réponse positive, la valeur de \emph{chit} est mise à~1.
+Si plus que $\alpha$ de ces pairs reviennent à une réponse positive, la valeur de chit est mise à~1.
 %O After that, it updates the preferred transaction of each conflict set
@@
 %O receives a query for transaction $T$ from peer $j$.
-La Figure~\ref{fig:gossipchain-onquery} montre ce qui se produit quand un nœud reçoir une requête pour la transaction $T$ du pair $j$.
+La figure~\ref{fig:gossipchain-onquery} montre ce qui se produit quand un nœud reçoir une requête pour la transaction $T$ du pair $j$.
 %O First it adds $T$ to $\mathcal{T}$, unless it already has it.

### 2020-11-14 paper_fr.tex
French translation of the related work part
Relecture jusqu'à l'annexe

@@
 %O \Large\bf Scalable and Probabilistic Leaderless BFT Consensus through Metastability}
-\Large\bf Consensus BFT sans dirigeant, extensible et probabiliste, par métastabilité}
+\Large\bf Consensus BFT sans dirigeant extensible et probabiliste par métastabilité}
 \author{%
@@
 It does not require accurate knowledge of all participants and exposes new possible tradeoffs and improvements in safety and liveness for building consensus protocols.
-Contrairement aux protocoles de consensus traditionnels où un ou plusieurs nœuds traitent typiquement des bits de manière linéaire dans les nombres de nœuds totaux par décision, aucun nœud ne les traite plus que de manière logarithmique.
-HOULA
+Contrairement aux protocoles de consensus traditionnels où, typiquement, un ou plusieurs nœuds traitent linéairement l'information par décision, aucun nœud ne traite l'information plus que de manière logarithmique.
+% HOULA
 %Finally, unlike traditional consensus protocols and similar to longest-chain protocols, our protocols provide increase in both liveness and safety guarantees when the observed adversarial presence decreases below the maximal expected bound.
@@
 %O The paper describes the Snow protocol family, analyzes its guarantees, and describes how it can be used to construct the core of an internet-scale electronic payment system called Avalanche, which is evaluated in a large scale deployment.
-Cet article décrit la famille de protocoles Snow, analyse ses garanties et décrit la manière dont il peut être utilisé pour construire le cœur d'un système de paiement électronique à l'échelle de l'internet appelé Avalanche, qui est évalué lors d'un déploiement à grande échelle.
+Cet article décrit la famille de protocoles Snow, analyse ses garanties et décrit la manière dont il peut être employé pour bâtir le cœur d'un système de paiement électronique à l'échelle de l'internet appelé Avalanche, lequel est évalué lors d'un déploiement à grande échelle.
 %O Experiments demonstrate that the system can achieve high throughput (3400 tps), provide low confirmation latency (1.35 sec), and scale well compared to existing systems that deliver similar functionality.
 For our implementation and setup, the bottleneck of the system is in transaction verification.
-Les expériences démontrent que le système peut atteindre un haut débit (3400~tps), fournir une faible latence de confirmation (1,35~s) et correctement passer à l'échelle par rapport à d'autres systèmes offrant des fonctionnalités similaires.
-Le goulet d'étranglement du système dans l'implémentation et la configuration retenue est la vérification des transations.
+Les expériences montrent que le système peut atteindre un haut débit (3400~tps), fournir une faible latence de confirmation (1,35~s) et correctement passer à l'échelle par rapport à d'autres systèmes offrant des fonctionnalités similaires.
+Le goulet d'étranglement du système dans l'implémentation et la configuration retenues est la vérification des transactions.
 \end{abstract}
@@
 %O Achieving agreement among a set of distributed hosts lies at the core of countless applications, ranging from Internet-scale services that serve billions of people~\cite{Burrows06,HuntKJR10} to cryptocurrencies worth billions of dollars~\cite{marketcapcryptocurrency}.
-L'accord parmi un ensemble d'hôtes distribués se retrouve au cœur d'innombrables applications allant de services à l'échelle de l'Internet servant des milliards de gens~\cite{Burrows06,HuntKJR10} à des cryptomonnaies valant des milliards de dollars~\cite{marketcapcryptocurrency}.
+L'obtention d'un accord au sein d'un ensemble d'hôtes distribués se retrouve au cœur d'innombrables applications allant de services à l'échelle de l'Internet servant des milliards de gens~\cite{Burrows06,HuntKJR10} à des cryptomonnaies valant des milliards de dollars~\cite{marketcapcryptocurrency}.
 %O To date, there have been two main families of solutions to this problem.
@@
 %O to large numbers of participants.
-Comme ils impliquent en général un nombre de communication exponentiel et une connaissance exacte des participants, il s'est avéré difficile de l'étendre
+Comme ils impliquent en général un nombre de communications quadratique et une connaissance exacte des membres du réseau, il s'est avéré difficile de l'étendre
 à de nombreux participants.
@@
 %O These protocols provide a probabilistic safety guarantee: Nakamoto consensus decisions may revert with some probability $\varepsilon$.
-Ces protocoles fournissent une garantie de sûreté probabiliste : les décisions par consensus Nakamoto peuvent s'inverser avec une certaine probabilité $\varepsilon$.
+Ces protocoles fournissent une garantie de sûreté probabiliste\,: les décisions par consensus Nakamoto peuvent s'inverser avec une certaine probabilité $\varepsilon$.
 %O A protocol parameter allows this probability to be rendered arbitrarily small, enabling high-value financial systems to be constructed on this foundation.
@@
 %O This family is a natural fit for open, permissionless settings where any node can join the system at any time.
-Cette famille est faite pour des configurations ouvertes et en libre accès où tout nœud peut rejoindre le système à tout moment.
+Cette famille est conçue pour des configurations ouvertes et en libre accès où tout nœud peut rejoindre le système à tout moment.
 %O Yet, these protocols are costly, wasteful, and limited in performance.
@@
 %O By construction, they cannot quiesce: their security relies on constant participation by miners, even when there are no decisions to be made.
-Par construction, ils sont tenus de fonctionner même à vide : leur sécurité se base sur une participation constante des mineurs, même quand aucune décision ne doit être prise.
+Par construction, ils sont tenus de fonctionner même à vide\,: leur sécurité se base sur une participation constante des mineurs, même quand aucune décision ne doit être prise.
 %O Bitcoin currently consumes around 63.49 TWh/year~\cite{bitcoinpower}, about twice as all of Denmark~\cite{denmarkpower}.
@@
 % \Jon{I wanted the paper to expand on this point, but I missed it.}
-De plus, ces protocoles souffrent d'un goulet d'étranglement intrinsèque pour le passage à l'échelle, qui est difficile à dépasser par la simple reparamétrisation~\cite{CromanDEGJKMSSS16}.
+De plus, ces protocoles souffrent d'un goulet d'étranglement intrinsèque pour le passage à l'échelle, qui s'avère difficile à dépasser par la simple reparamétrisation~\cite{CromanDEGJKMSSS16}.
 %O This paper introduces a new family of consensus protocols called Snow.
@@
 %O Analysis shows that this metastable mechanism is powerful: it can move a large network to an irreversible state quickly, where the irreversibility implies that a sufficiently large portion of the network has accepted a proposal and a conflicting proposal will not be accepted with any higher than negligible ($\varepsilon$) probability.
-L'analyse montre que ce mécanisme métastable est puissant : il peut rapidement changer l'état d'un vaste réseau de manière irréversible, où l'irréversibilité implique qu'une part suffisamment importante du réseau a accepté une proposition et qu'une proposition différente ne sera acceptée qu'avec une probabilité au maximum négligeable.
+L'analyse montre que ce mécanisme métastable est puissant\,: il peut rapidement changer l'état d'un vaste réseau de manière irréversible, où l'irréversibilité implique qu'une part suffisamment importante du réseau a accepté une proposition et qu'une proposition différente ne sera acceptée qu'avec une probabilité au maximum négligeable.
 %{\color{red} We should clarify this.  The current statement is too strong.}
@@
 %O Similar to Nakamoto consensus, the Snow protocol family provides a probabilistic safety guarantee, using a tunable security parameter that can render the possibility of a consensus failure arbitrarily small.
-De même que le consensus Nakamoto, la famille des protocoles Snow fournissent une garantie de sûreté probabiliste, grâce à un paramètre des sécurité réglable qui peut amenuiser à volonter la possibilité d'un échec du consensus.
+De même que le consensus Nakamoto, la famille des protocoles Snow fournissent une garantie de sûreté probabiliste, grâce à un paramètre de sécurité configurable qui peut amenuiser à volonté la possibilité d'un échec du consensus.
 %O Unlike Nakamoto consensus, the protocols are green, quiescent and efficient; they do not rely on proof-of-work~\cite{DworkN92} and do not consume energy when there are no decisions to be made.
@@
 %0 The efficiency of the protocols stems partly from removing the leader bottleneck: each node requires $\Oh{1}$ communication overhead per round and $\Oh{\log{n}}$ rounds in expectation, whereas classical consensus protocols have one or more nodes that require $\Oh{n}$ communication per round (phase).
-L'efficacité des protocoles provient en partie de l'élimination du goulet d'étranglement des dirigeants : chaque nœud ne demande qu'un nombre de communications correspondant à $\Oh{1}$ par tour et une prévision de $\Oh{\log{n}}$ tours, alors que, dans les protocoles de consensus classique, un ou plusieurs nœuds requièrent $\Oh{n}$ communications par tour.
+L'efficacité des protocoles provient en partie de l'élimination du goulet d'étranglement des dirigeants\,: chaque nœud ne demande qu'un nombre de communications correspondant à $\Oh{1}$ par tour et une prévision de $\Oh{\log{n}}$ tours, alors que, dans les protocoles de consensus classiques, un ou plusieurs nœuds requièrent $\Oh{n}$ communications par tour.
 %O Further, the Snow family tolerates discrepancies in knowledge of membership, as we discuss later.
@@
 De plus, la famille Snow tolère des incohérences dans la connaissance des participants, comme nous le verrons plus loin.
-Par contraste, les protocoles de consensus classiques exigent une connaisssance complète et exacte de $n$ comme fondation de la sûreté.
+Par contraste, les protocoles de consensus classiques exigent une connaisssance complète et exacte de $n$ comme fondation de la sûreté de fonctionnement.
 
@@
 %O Snow's probabilistic safety guarantee degrades smoothly when Byzantine participants exceed $f$.
-Alors que la sûreté des approches basées sur un quorum s'effondre immédiatement quand le seuil prédéterminé $f$ est dépassé,
+Alors que la sûreté de fonctionnement dans les approches basées sur un quorum s'effondre immédiatement quand le seuil prédéterminé $f$ est dépassé,
 la garantie de sûreté probabiliste de Snow se dégrade progressivement quand les participants byzantins dépassent $f$.
@@
 %O nodes by trading off liveness.
-Cela expose également de nouveaux compromis entre sûreté et vitalité : la famille Snow est plus efficace quand la fraction de nœuds byzantins est petite, et cela peut être paramétré pour tolérer plus d'un tiers de nœuds byzantins en sacrifiant la vitalité.
+Cela expose également de nouveaux compromis entre sûreté et vitalité\,: la famille Snow est plus efficace quand la fraction de nœuds byzantins est petite, et cela peut être paramétré pour tolérer plus d'un tiers de nœuds byzantins en sacrifiant la vitalité.
 
@@
 Pour démontrer le potentiel de cette famille de protocoles, nous l'illustrons par un système pratique de paiement pair à pair, Avalanche.
-Dans les faits, Avalanche exécute plusieurs instances de Snowball (l'un des protocoles de la famille Snow) avec l'aide d'un graphe orienté acyclique (\emph{Directed Acyclic Graph} ou DAG).
-Le DAG sert à empiler plusieurs instances en réduisant le coût de $\Oh{\log{n}}$ à $\Oh{1}$ par nœud et en optimisant le chemin où il n'existe pas de transactions conflictuelles.
+Dans les faits, Avalanche exécute plusieurs instances de Snowball (l'un des protocoles de la famille Snow) à l'aide d'un graphe orienté acyclique (\emph{Directed Acyclic Graph} ou DAG).
+Le DAG sert à empiler plusieurs instances en faisant passer le coût de $\Oh{\log{n}}$ à $\Oh{1}$ par nœud et en optimisant le chemin là où il n'existe pas de transactions conflictuelles.
 %This combination of the best features of traditional and Nakamoto consensus involves one significant tradeoff: liveness for \editchange{equivocating proposals}. %conflicting transactions.
@@
 %O The next section provides the model, goals and necessary assumptions for the new protocols.
-La section suivante fournit le modèle, les buts et les suppositions nécessaires pour les nouveaux protocoles.
+La section suivante fournit le modèle, les buts et les hypothèses nécessaires pour les nouveaux protocoles.
 %O Section~\ref{sec:protocol} gives intuition behind the protocols, followed by their full specification,
-La section~\ref{sec:protocol} donne l'intuition derrière les protocoles, suivis de leurs spécifications complètes,
+La section~\ref{sec:protocol} expose l'intuition ayant mené à ces protocoles, suivie de leurs spécifications complètes,
 %Section~\ref{sec:attack} briefly discusses typical attack vectors and their possible defense,
@@
 %O by our formal analysis of safety and liveness in Appendix~\ref{sec:full-analysis},
-La section~\ref{sec:analysis} fournit la méthodologie employée par notre analyse formelle de sûreté et de vitalité en Appendix~\ref{sec:full-analysis},
+La section~\ref{sec:analysis} fournit la méthodologie employée par notre analyse formelle de sûreté de fonctionnement et de vitalité en annexe~\ref{sec:full-analysis},
 %Section~\ref{sec:implementation} describes a Bitcoin-like payment system,  Section~\ref{sec:evaluation} evaluates the system,
@@
 %O Section~\ref{sec:related-work} presents related work, and finally, Section~\ref{sec:conclusions} summarizes our contributions.
-La section~\ref{sec:related-work} présente les travaux en relation, et, finalement, la section~\ref{sec:conclusions} résume nos contributions.
+La section~\ref{sec:related-work} présente les travaux en rapport, et, finalement, la section~\ref{sec:conclusions} résume nos contributions.
 
@@
 %O \paragraph{Safety} Unlike classical consensus protocols, and similar to longest-chain-based consensus protocols such as Nakamoto consensus~\cite{nakamoto2008bitcoin}, we adopt an $\varepsilon$-safety guarantee that is probabilistic.
-\paragraph{Sûreté} À l'inverse des protocoles de consensus classiques, et à l'instar des protocoles de consensus fondés sur la chaîne la plus longue~\cite{nakamoto2008bitcoin}, nous adoptons une garantie de sûreté-$\varepsilon$ qui est probabiliste.
+\paragraph{Sûreté de fonctionnement} À l'inverse des protocoles de consensus classiques, et à l'instar des protocoles de consensus fondés sur la chaîne la plus longue~\cite{nakamoto2008bitcoin}, nous adoptons une garantie de sûreté de fonctionnement~$\varepsilon$ qui est probabiliste.
 %O In practice, this probabilistic guarantee is as strong as traditional safety guarantees, since appropriately small choices of $\varepsilon$ can render consensus failure negligible, lower than the probability of hardware failure due to random events.
-En pratique, cette garantie probabiliste est aussi forte que les garantie de sûreté traditionnelles puisque le choix d'un petit $\varepsilon$ peut rendre la défaillance du consensus négligeable, inférieurs à la probabilité d'une panne matérielle due à des événements imprévisibles.
+En pratique, cette garantie probabiliste est aussi forte que les garanties de sûreté traditionnelles puisque le choix d'un petit $\varepsilon$ peut rendre la défaillance du consensus négligeable, inférieure à la probabilité d'une panne matérielle due à des événements imprévisibles.
 %OFigure~\ref{fig:fandepsilon} shows how the portion ($f/n$) of misbehaving participants (or computation power) affects the probability of system safety failure (decision of two conflicting proposals), given a choice of finality.
-La figure~\ref{fig:fandepsilon} montre comment la portion ($f/n$) de participants (ou de puissance informatique) problématique affecte la probabilité d'une défaillance de la sûreté du système (décision de deux propositions incompatibles), selon un choix de finalité.
+La figure~\ref{fig:fandepsilon} montre comment la portion ($f/n$) de participants (ou de puissance informatique) problématique affecte la probabilité d'une défaillance de la sûreté du système (décision de deux propositions incompatibles), étant donné un choix de finalités.
 % The maximum tolerated Byzantine presence $f$, is dependent on the choice of $\varepsilon$.
@@
 Snowflake-7,8 uses $\alpha=7$ and $\alpha=8$ respectively (see Section~\ref{sec:protocol} for the definition of $k$, $\alpha$ and $\beta$.}
-    \caption{Les protocoles BFT classiques qui tolèrent $f$ défaillances rencontrent une défaillance totale de leur sûreté quand le seuil est dépaasé même d'un seul nœud supplémentaire.
+    \caption{Les protocoles BFT classiques qui tolèrent $f$ défaillances rencontrent une défaillance totale de leur sûreté de fonctionnement quand le seuil est dépassé même d'un seul nœud supplémentaire.
 La courbe de Bitcoin montre un choix de finalité typique pour Bitcoin où un bloc est considéré final quand il est \guillemotleft\,enterré\,\guillemotright~ dans une branche sous 6 blocs supplémentaires par rapport aux autres branches concurrentes.
@@
 %O In other words, the time required for finality approaches $\infty$ as $f$ approaches $n/2$\tronly{ (Figure~\ref{fig:livenessproperties}).}{.}
-En d'autres termes, le temps nécessaire à la finalité approche $\infty$ quand $f$ approche $n/2$\tronly{ (Figure~\ref{fig:livenessproperties}).}{.}
+En d'autres termes, le temps nécessaire à la finalité approche $\infty$ quand $f$ approche $n/2$\tronly{ (figure~\ref{fig:livenessproperties}).}{.}
 %O Furthermore, the required number of rounds is calculable ahead of time, as to allow the system designer to tune liveness at the expense of safety.
 Lastly, unlike traditional consensus protocols and similar to Nakamoto, our protocols benefit from lower adversarial presence, as discussed in property P3 below.
-De plus, le nombre requis de tours est calculable à l'avance, ce qui permet au concepteur du système de régler la vitalité aux dépens de la sûreté.
+De plus, le nombre requis de tours est calculable à l'avance, ce qui permet au concepteur du système de régler la vitalité aux dépens de la sûreté de fonctionnement.
 Enfin, contrairement aux protocoles de consensus traditionnels et similairement à Nakamoto, nos protocoles bénéficient d'une moindre présence antagoniste, comme le détaille la propriété P3 ci-dessous.
@@
 The Snow protocols then provide the following guarantees:
-\emph{Garantie formelle} : Soit un paramétrage du système pour une probabilité de défaillance de la sûreté $\varepsilon$ pour un nombre maximal attendu $f$ de nœuds antagonistes.
+\emph{Garantie formelle}\,: Soit un paramétrage du système pour une probabilité de défaillance de la sûreté de fonctionnement $\varepsilon$ pour un nombre maximal attendu $f$ de nœuds antagonistes.
 Soit $\Oh{\log n} < t_{max} < \infty$ la limite supérieure de l'exécution du protocole.
@@
 %O     \item \textbf{P1. Safety.} When decisions are made by any two correct nodes, they decide on conflicting transactions with negligible probability ($\leq \varepsilon$).
-     \item \textbf{P1. Sûreté.} Quand les décisions sont prises par deux nœuds corrects, ils décident de transactions conflictuelles avec une probabilité négligeable ($\leq \varepsilon$).
+     \item \textbf{P1.
+Sûreté de fonctionnement.} Quand les décisions sont prises par deux nœuds corrects, ils décident des transactions conflictuelles avec une probabilité négligeable ($\leq \varepsilon$).
     % \item \textbf{P2.
@@
 This implies that the scheduling of message transmission itself could behave arbitrarily, and potentially even maliciously (with full asynchrony).
-Dans la définition standard de l'asynchronie~\cite{ben1983another}, le temps de transmission des messages est fini, mais la distribution n'est pas spécifiée (donc le temps de délivrance peut être sans limite pour certains messages).
-Cela implique que l'échelonnement de la transmission des messages lui-même peut être quelconque, voire potentiellement malveillant (pour une complète asynchronie).
+Dans la définition standard de l'asynchronie~\cite{ben1983another}, le temps de transmission des messages est fini, mais la distribution n'est pas spécifiée (donc le temps de délivrance peut être non borné pour certains messages).
+Cela implique que l'ordonnancement de la transmission des messages lui-même peut être quelconque, voire potentiellement malveillant (pour une complète asynchronie).
 %O We use a modified version of this model, which is well-accepted~\cite{banerjee2014epidemic, ganesh2005effect, draief2006thresholds, keeling2011modeling, liggett1997stochastic} in the analysis of epidemic networks and gossip-based stochastic systems.
@@
 %O Furthermore, we also note that scheduling only applies to correct nodes, and the adversary may execute arbitrarily, as discussed later. 
-De plus, nous notons également que l'échelonnement ne s'applique qu'aux nœuds corrects, et que l'adversaire peut s'exécuter arbitrairement, comme nous le verrons plus loin.
+De plus, notons également que l'ordonnancement ne s'applique qu'aux nœuds corrects, et que l'adversaire peut s'exécuter arbitrairement, comme nous le verrons plus loin.
 
@@
 Dans notre système, cependant, les nœuds opèrent par sous-échantillonnage, d'où la possibilité pour un seul échantillon qu'une majorité de nœuds antagonistes soit sélectionné, ce qui contraint le nœud à attendre en vain une réponse.
-Pour s'assurer de la vitalité, un nœud doit pouvoir poser une limite à l'attente.
+Pour s'assurer de la vitalité, un nœud doit pouvoir limiter son attente.
 Enfin, il faut noter que le consensus Nakamoto est synchrone et que la difficulté requise pour la preuve de travail dépend du délai maximum du réseau~\cite{PassSS17}.
@@
 %O Consensus protocols provide their guarantees based on assumptions that only a fraction of participants are adversarial.
-Les protocoles de consensus fournissent leurs garanties fondées sur la supposition que seule une fraction des participants sont antagonistes.
+Les protocoles de consensus fournissent leurs garanties en se fondant sur l'hypothèse que seule une fraction des participants sont antagonistes.
 %O These bounds could be violated if the network is naively left open to arbitrary participants.
@@
 %O Other protocols, discussed in Section~\ref{sec:related-work}, rely on proof-of-stake (by economic argument) or proof-of-authority (by administrative argument that makes the system ``permissioned'').
-D'autres protocoles, dont il est question en section~\ref{sec:related-work}, se basent sur la preuve d'enjeu (\emph{proof of stake} ou PoS, par un argument économique) ou preuve d'autorité (\emph{proof of authority} ou PoA, par un argument administratif) qui rend le système \guillemotleft\,permissionné\,\guillemotright.
+D'autres protocoles, dont il est question en section~\ref{sec:related-work}, se basent sur la preuve d'enjeu (\emph{proof of stake} ou PoS, par un argument économique) ou la preuve d'autorité (\emph{proof of authority} ou PoA, par un argument administratif) qui rend le système \guillemotleft\,permissionné\,\guillemotright.
 %O The consensus protocols presented in this paper can adopt any
@@
 %O Flooding/spam attacks are a problem for any distributed system. 
-Les attaques par inondation (\emph{flooding}) et spam représentent un problème pour tout système distribué.
+Les attaques par inondation et par spam représentent un problème pour tout système distribué.
 %O Without a protection mechanism, an attacker can generate large numbers of transactions and flood protocol data structures, consuming storage.
@@
 %O There are a multitude of techniques to deter such attacks, including network-layer protection, proof-of-authority, local proof-of-work and economic mechanisms. 
-Il existe une multitude de techniques pour repousser ces attaques comme la protection de la couche réseau, la preuve d'autorité, la preuve de travail local et des mécanismes économiques.
+Il existe une multitude de techniques pour repousser ces attaques comme la protection de la couche réseau, la preuve d'autorité, la preuve de travail local, ainsi que des mécanismes économiques.
 %O In Avalanche, we use transaction fees, making such attacks costly even if the attacker is sending money back to addresses under its control.
@@
 %O We quantify the bounds on the discrepancy in Appendix~\ref{sec:full-analysis-churn}.
-Nous quantifions les limites en Annexe~\ref{sec:full-analysis-churn}.
+Nous en quantifions les bornes en annexe~\ref{sec:full-analysis-churn}.
 %O We assume a safe bootstrapping mechanism, similar to that of Bitcoin, that enables a node to connect with sufficiently many correct nodes to acquire a statistically unbiased view of the network.
@@
 %0 \subsection{Slush: Introducing Metastability}
-\subsection{Slush: A propos de métastabililté}
+\subsection{Slush\,: présentation de la métastabililté}
 
@@
 
-Le principe de notre approche est un protocole de consensus à décision simple inspiré des protocoles dits "épidémiques"
-ou "gossip". Le protocole le plus simple, Slush, est la fondation de cette famille, comme le montre la
-Figure~\ref{fig:slush-loop}. Slush \emph{n'est pas} tolérant aux fautes byzantines, seulement aux fautes par défaut
+Le principe de notre approche est un protocole de consensus à simple décret inspiré des protocoles dits épidémiques
+ou \emph{gossip}. Le protocole le plus simple, Slush, est la fondation de cette famille, comme le montre la
+figure~\ref{fig:slush-loop}. Slush \emph{n'est pas} tolérant aux fautes byzantines, seulement aux fautes par défaut % HOULA : faute ou défaillance ?
 de fonctionnement (CFT), mais sert à mettre en avant les autres protocoles tolérants aux fautes byzantines qui vont
@@
 
-Avec Slush, un noeud démarre dans un état de couleur neutre. Après réception d'une transaction envoyée par un client,
-un node de couleur neutre adopte la couleur définie dans la transaction et initie une requête. Pour lancer une
-requête, le noeud choisit un échantillon réduit et de taille constante ($k$) de noeuds appartenant au réseau suivant
-une distribution uniformément aléatoire puis initie sa propre requête, tandis qu'un noeud déja coloré répond
-simplement avec sa couleur actuelle. Une fois que le noeud qui fait la requête récupère $k$ réponses, il vérifie si
+Avec Slush, un nœud démarre dans un état de couleur neutre. Après réception d'une transaction envoyée par un client,
+un nœud de couleur neutre adopte la couleur définie dans la transaction et initie une requête. Pour lancer une
+requête, le nœud choisit un échantillon réduit et de taille constante ($k$) de nœuds appartenant au réseau suivant
+une distribution uniformément aléatoire puis initie sa propre requête, tandis qu'un nœud déja coloré répond
+simplement avec sa couleur actuelle. Une fois que le nœud qui fait la requête récupère $k$ réponses, il vérifie si
 une proportion $\ge\alpha$ de l'échantillon retourne la même couleur, $\alpha > \floor{k/2}$ étant un paramètre
 du protocole. Si le seuil $\alpha$ est dépassé et que la couleur dominante revenant de l'échantillon sollicité est
-différente de la couleur propre du node à l'origine de la requête, ce node adopte cette nouvelle couleur. Il revient
+différente de la couleur propre du nœud à l'origine de la requête, ce nœud adopte cette nouvelle couleur. Il revient
 ensuite à l'étape de génération de la requête et lance un nouveau tour de requête jusqu'à un total de $m$ tours.
-En définitive, le node décide de sa couleur finale à l'instant $m$.
+En définitive, le nœud décide de sa couleur finale à l'instant $m$.
 
@@
         \For{$r \in \{1\ldots m\}$}
-            \State \textrm{// Si $\bot$, réitère jusqu'à ce que \textsc{onQuery} applique la couleur}
+            \State \textrm{// Si $\bot$, saute jusqu'à ce que \textsc{onQuery} applique la couleur}
             \If{$\codecolor = \bot$}
@@
             \EndIf
-            \State \textrm{// échantillonne aléatoirement parmi les noeuds connus}
+            \State \textrm{// échantillonne aléatoirement parmi les nœuds connus}
             \State $\mathcal{K} \assign \Call{échantillon}{\mathcal{N}\backslash u, k}$
@@
     \EndProcedure
-    \captionof{figure}{Protocole Slush. Les temps d'expiration sont omis dans un souci de lisibilité.}\label{fig:slush-loop}
+    \captionof{figure}{Protocole Slush. Temps d'expiration omis dans un souci de lisibilité.}\label{fig:slush-loop}
 \end{algorithmic}
@@
 
-Slush a quelques propriétés intéréssantes. Premièrement, il n'a quasiment \emph{pas besoin de mémoire}: un noeud ne
+Slush possède quelques propriétés intéressantes. Premièrement, il n'a quasiment \emph{pas besoin de mémoire}\,: un nœud ne
 stocke aucun état intermédiaire entre chaque tour autre que sa couleur actuelle, et plus particulièrement ne stocke
-aucun historique de ses interactions avec les autres noeuds. Deuxièmement, contrairement aux protocoles de consensus
+aucun historique de ses interactions avec les autres nœuds. Deuxièmement, contrairement aux protocoles de consensus
 traditionnels qui sollicitent tous les participants, chaque tour n'implique qu'un échantillon du réseau réduit et de
 taille constante choisi aléatoirement. Troisièmement, Slush progresse quelque soit la configuration du réseau (même
-depuis un état entièrement bivalent, ex: partage 50/50 entre les deux couleurs) vu que des perturbations aléatoires
-sur l'échantillonnage auront pour conséquence de fournir un léger avantage à l'une des deux couleurs qui sera ensuite
+depuis un état entièrement bivalent, c'est-à-dire un partage 50/50 entre les deux couleurs) puisque des perturbations aléatoires
+sur l'échantillonnage auront pour conséquence de fournir un léger avantage à l'une des deux couleurs, lequel sera ensuite
 amplifié au fur et à mesure des échantillonnages.
-Finalement si $m$ est choisi pour être suffisamment élevé, Slush s'assure que tous les noeuds convergeront vers la
-même couleur avec une haute probabilité. Chaque noeud nécéssite un temps de communication par tour constant et
+Finalement, si $m$ est choisi pour être suffisamment élevé, Slush s'assure que tous les nœuds convergent vers la
+même couleur avec une forte probabilité. Chaque nœud nécessite un nombre de communications par tour constant et   % HOULA communication overhead
 prévisible, et $m$ croît de manière logarithmique suivant $n$.
@@
 
-Le protocole Slush ne fournit pas de fortes garanties quant à la sûreté de fonctionnement en la présence de noeuds
-byzantins. En particulier si les noeuds légitimes développent une préférence pour une certaine couleur, un adversaire
-byzantin pourra inverser la couleur des noeuds afin de maintenir le système à l'équilibre, incapable de prendre une
-décision. Nous adressons ce problème dans notre premier protocole tolérant aux fautes byzantines qui introduit un
-meilleur suivi de l'état au niveau des nodes.
+Le protocole Slush ne fournit pas de fortes garanties quant à la sûreté de fonctionnement en présence de nœuds
+byzantins. En particulier, si les nœuds légitimes développent une préférence pour une certaine couleur, un adversaire
+byzantin pourra inverser la couleur des nœuds afin de maintenir le système à l'équilibre, empêchant ainsi de prendre une
+décision. Nous étudions ce problème dans notre premier protocole BFT qui ajoute un stockage de l'état sur les nœuds.
 }{
 %O We next examine how to extend Slush to tolerate Byzantine behavior.
-Nous examinons ensuite comment améliorer Slush pour être tolérant au problème des généraux byzantins.
+Nous examinons ensuite la manière de fournir à Slush une tolérance aux comporements byzantins.
 }
 
-\subsection{Snowflake: BFT}
-
-Snowflake améliore Slush en ajoutant un compteur unique qui capture la force de conviction d'un noeud dans sa couleur
-actuelle. Ce compteur associé à chaque noeud stocke le nombre d'échantillons du réseau consécutifs que le noeud a
-sollicité et qui ont retourné la même couleur. Un noeud accepte sa couleur actuelle lorsque son compteur atteint $\beta$,
-un autre paramètre de sécurité. La Figure~\ref{fig:snowflake-loop} montre le protocole amélioré qui inclue les
-modifications suivantes:
+\subsection{Snowflake\,: BFT}
+
+Snowflake améliore Slush en ajoutant un compteur unique qui capture la force de conviction d'un nœud concernant sa couleur
+actuelle. Ce compteur associé à chaque nœud stocke le nombre d'échantillons du réseau consécutifs que le nœud a
+sollicité et qui ont retourné la même couleur. Un nœud accepte la couleur actuelle lorsque son compteur atteint $\beta$,
+un autre paramètre de sécurité. La figure~\ref{fig:snowflake-loop} montre le protocole amélioré qui inclut les
+modifications suivantes\,:
 
@@
 \begin{compactenum}
-	\item Chaque node maintient un compteur $\mathit{cnt}$;
-    \item A chaque changement de couleur, le compteur $\mathit{cnt}$ se réinitialise à 0;
-    \item Après chaque requête réussie qui retourne $\ge \alpha$ réponses en faveur de la même couleur que le noeud,
+	\item Chaque nœud maintient un compteur $\mathit{cnt}$\,;
+    \item À chaque changement de couleur, le compteur $\mathit{cnt}$ se réinitialise à 0\,;
+    \item Après chaque requête réussie qui renvoie $\ge \alpha$ réponses en faveur de la même couleur que le nœud,
       ce dernier incrémente $\mathit{cnt}$.
@@
 
-Quand le protocole est correctement parametré pour un seuil donné de noeuds byzantins et un niveau de garantie
+Quand le protocole est correctement paramétré pour un seuil donné de nœuds byzantins et un niveau de garantie
 $\varepsilon$, il peut garantir à la fois la sécurité de fonctionnement (P1) et la vitalité (P2, P3).
-Comme nous le démontrons plus loin, il existe un état irréversible à partir duquel une décision est inévitable. Les noeuds
+Comme nous le démontrons plus loin, il existe un état irréversible à partir duquel une décision est inévitable. Les nœuds
 légitimes commencent à se positionner après cet état irréversible et adoptent la même couleur. Une intuition
-supplémentaire, que nous ne détaillons pas dans ce papier, introduit l'existence d'un point de changement de phase à
-partir duquel les nodes byzantins perdent toute capacité à maintenir le réseau dans un état bivalent.
+supplémentaire, que nous ne détaillerons pas dans ce papier, indique l'existence d'un point de changement de phase à
+partir duquel les nœuds byzantins perdent toute capacité à maintenir le réseau dans un état bivalent.
 
@@
 
-\subsection{Snowball: Introduire la notion de confiance}
-
-La notion d'état de Snowflake est éphémère: le compteur est réinitialisé à chaque changement de couleur.
+\subsection{Snowball\,: ajout de la confiance}
+
+La notion d'état de Snowflake est éphémère\,: le compteur est réinitialisé à chaque changement de couleur.
 Snowball améliore Snowflake avec des \emph{compteurs de confiance} qui capturent le nombre de requêtes ayant retourné
-un résultat en faveur de la couleur du noeud au dessus d'un certain seuil (Figure ~\ref{fig:snowball-loop}).
-Un node prend une décision s'il obtient $\beta$ réponses consécutives en faveur d'une couleur. Cependant il change
+un résultat en faveur de la couleur du nœud au dessus d'un certain seuil (figure ~\ref{fig:snowball-loop}).
+Un nœud prend une décision s'il obtient $\beta$ réponses consécutives en faveur d'une couleur. Cependant, il change
 sa préférence uniquement sur la base de la confiance totale accumulée. Les différences entre Snowflake et Snowball
-sont:
+sont les suivantes\,:
 
 \begin{compactenum}
-    \item Après chaque requête réussie, le noeud incrémente son compteur de confiance associé à la couleur retournée.
-    \item Un noeud change de couleur quand la confiance associée à sa couleur actuelle devient inférieure à la
+    \item Après chaque requête réussie, le nœud incrémente son compteur de confiance associé à la couleur retournée\,;
+    \item Un nœud change de couleur quand la confiance associée à sa couleur actuelle devient inférieure à la
       valeur de la confiance associée à la nouvelle couleur.
@@
 \label{sec:analysis}
-Pour des raisons de manque de place, nous déplaçons certains détails importants vers l'Annexe~\ref{sec:full-analysis},
+Pour des raisons de manque de place, nous déplaçons certains détails importants vers l'annexe~\ref{sec:full-analysis},
 dans laquelle nous démontrons que sous certaines hypothèses distinctes et indépendantes, la famille de protocoles de
@@
 
-\paragraph{Notation} En considérant que le réseau consiste d'un ensemble de $n$ noeuds (représentés par l'ensemble
-$\mathcal{N}$), où $c$ sont des noeuds légitimes (représentés par l'ensemble $\mathcal{C}$) et $f$ sont les noeuds
-byzantins (représentés par l'ensemble $\mathcal{B}$). Soient $u, v \in \mathcal{C}$ des références à deux noeuds
+\paragraph{Notation} En considérant que le réseau consiste en un ensemble de $n$ nœuds (représentés par l'ensemble
+$\mathcal{N}$), où $c$ sont des nœuds légitimes (représentés par l'ensemble $\mathcal{C}$) et $f$ sont les nœuds
+byzantins (représentés par l'ensemble $\mathcal{B}$). Soient $u, v \in \mathcal{C}$ des références à deux nœuds
 légitimes pris au hasard dans le réseau. Soient $k, \alpha, \beta \in \mathbb{Z}^+$ des entiers positifs où
 $\alpha > \floor{k/2}$. A partir de maintenant, $k$ fera toujours référence à la taille des déchantillons du réseau,
-où $k \leq n$, et $\alpha$ sera le seuil de majorité requis pour considérer l'expérience de vote comme ``réussie''.
+où $k \leq n$, et $\alpha$ sera le seuil de majorité requis pour considérer l'expérience de vote comme «\,réussie\,».
 En général, nous définissons $\mathcal{S}$ comme étant l'état (ou la configuration) du réseau à tout moment.
@@
 \paragraph{Framework de modélisation} Afin de modéliser nos protocoles de manière formelle, nous utilisons des
-processus Markov à temps continu (CTMC). L'espace contenant les états est dénombrable (et fini), et les transitions
-entre états s'opèrent sur un espace de temps continu. Les CTMCs proposent une modélisation naturelle de nos protocoles
-sachant que les transitions entre états ne se font pas par étapes et en marche ordonnée pour chaque noeud (à la fin de
-chaque unité de temps) mais arrivent plutôt à n'importe quel moment et indépendamment des autres noeuds.
-
-Nous nous concentrons sur un consensus binaire, même si les résultats en term de sûreté de fonctionnement se
-généralisent sur plus de deux valeurs. Nous pouvons imaginer le réseau comme un ensemble de noeuds colorés soit en
-rouge soit en bleu, et nous ferons référence à cette configuration à l'instant $t$ comme étant $\mathcal{S}_t$.
-Nous modélisons nos protocoles à travers un processus à temps continu contenant deux états, dans lequel tous les noeuds
-sont rouges ou tous les noeuds sont bleus. L'espace $\mathcal{S}$ des états du processus stochastique est une version
-condensée de l'espace de configuration complet, où chaque état $\{0, \dots, n\}$ représente le nombre total de noeuds
+processus de Markov à temps continu (CTMC). L'espace contenant les états est dénombrable (et fini), et les transitions
+entre états s'opèrent sur un espace de temps continu. Les CTMC proposent une modélisation naturelle de nos protocoles
+sachant que les transitions entre états ne se font pas par étapes et en marche ordonnée pour chaque nœud (à la fin de
+chaque unité de temps) mais arrivent plutôt à n'importe quel moment et indépendamment des autres nœuds.
+
+Nous concentrons l'étude sur un consensus binaire, même si les résultats en termes de sûreté de fonctionnement se
+généralisent pour plus de deux valeurs. Nous pouvons imaginer le réseau comme un ensemble de nœuds colorés, soit en
+rouge, soit en bleu, et nous ferons référence à cette configuration à l'instant $t$ comme étant $\mathcal{S}_t$.
+Nous modélisons nos protocoles à travers un processus à temps continu contenant deux états, dans lequel tous les nœuds
+sont rouges ou tous les nœuds sont bleus. L'espace $\mathcal{S}$ des états du processus stochastique est une version
+condensée de l'espace de configuration complet, où chaque état $\{0, \dots, n\}$ représente le nombre total de nœuds
 bleus dans le système.
@@
 La simplification qui nous permet d'analyser ce système consiste à s'affranchir du besoin de garder un historique de
-tous les chemins d'exécution, ainsi que toutes les stratégies d'attaque possibles, et de se concentrer plutôt
-entièrement sur un unique état, sans considérer la manière d'arriver à cet état. Plus spécifiquement, l'idée principale
+tous les chemins d'exécution, ainsi que de toutes les stratégies d'attaque possibles, et de se concentrer 
+entièrement sur un unique état quelle que soit la manière d'arriver à celui-ci. Plus spécifiquement, l'idée principale
 à retenir de notre analyse est l'identification d'un état \textit{d'irréversibilité} du système, cet état à partir
-duquel un tel nombre de noeuds légitimes a adopté soit la couleur rouge soit la couleur bleue que la possibilité de
-revenir à la couleur minoritaire est très peu probable.
+duquel un tel nombre de nœuds légitimes a adopté soit la couleur rouge soit la couleur bleue que la probabilité de
+revenir à la couleur minoritaire est très faible.
 
@@
 \paragraph{Slush}
-En partant du principe que tous les noeuds partagent le même $\mathcal{N}$, et en Annexe~\ref{sec:full-analysis-churn},
-nous expliquons comment relâcher le besoin pour les noeuds d'avoir connaissance de leur appartenance. Nous modélisons
-les dynamiques du système à travers un processus à temps continu contenant deux états, tous les noeuds rouges ou tous
-les noeuds bleus. Prenons $\{X_{t \geq 0}\}$ la variable aléatoire qui décrit l'état du système à l'instant $t$, où
+Nous partons du principe que tous les nœuds partagent le même $\mathcal{N}$, et 
+nous expliquons en annexe~\ref{sec:full-analysis-churn} comment assouplir l'exigence de la connaissance des participants. Nous modélisons
+les aspects dynamiques du système grâce à un processus à temps continu contenant deux états, tous les nœuds étant rouges ou tous  % HOULA dynamics → dynamique, comme economics
+les nœuds étant bleus. Soit $\{X_{t \geq 0}\}$ la variable aléatoire qui décrit l'état du système à l'instant $t$, où
 $X_0 \in \{0, \dots, c\}$.
-Nous commençons par détailler le résultat le plus important des dynamiques de sûreté de fonctionnement de notre
-processus: les probabilités de \emph{réversibilité} du processus \textbf{Slush}. Tous les autres résultats formels de
+Nous commençons par détailler le résultat le plus important de la dynamique de la sûreté de fonctionnement de notre
+processus\,: les probabilités de \emph{réversibilité} du processus \textbf{Slush}. Tous les autres résultats formels de
 ce papier sont, en simplifiant, des dérivations et améliorations intuitives de ce résultat.
@@
 Prenons la configuration du système à l'instant $t$ comme étant $\mathcal{S}_t = n/2 + \delta$, ce qui signifie que le
-réseau a dérivé de $2\delta$ noeuds bleus de plus que les noeuds rouges ($\delta = 0$ signifie que rouges et bleus
-sont en nombre égal). $\xi_\delta$ étant la probabilité d'un basculement vers l'état où tous les noeuds sont rouges,
+réseau a dérivé de $2\delta$ nœuds bleus de plus que les nœuds rouges ($\delta = 0$ signifie que rouges et bleus
+sont en nombre égal). $\xi_\delta$ étant la probabilité d'un basculement vers l'état où tous les nœuds sont rouges,
 nous avons donc pour tout $0 \leq \delta \leq n/2$
@@
 Nous laissons l'expression dans sa forme complète dans le Théorème~\ref{theorem:slush_prob_convergence_minority}
-disponible en annexe, qui est aussi bien plus fort que la borne de Chvatal.
+disponible en annexe, qui est aussi significativement plus fort que la borne de Chvatal.
 Néanmoins, l'utilisation de la borne de Chvatal nous permet de faire l'observation qui consiste à dire qu'au fur et à
-mesure que la dérivation $\delta$ augmente, pour un $\alpha$ et un $k$ constants, la probabilité de basculer vers la
-valeur minoritaire décroît \emph{de manière exponentielle} (en fait encore plus rapidement vu qu'il y a un terme
-quadratique dans l'exposant inverse). De plus, la même observation est valable si l'on fait croître $\alpha$ en gardant
+mesure que la dérivation $\delta$ augmente, pour un $\alpha$ et un $k$ constants, la probabilité de basculer vers la   % HOULA drift
+valeur minoritaire décroît \emph{de manière exponentielle} (en fait encore plus rapidement puisqu'il y a un terme
+quadratique dans l'exposant inverse). De plus, cette même observation est valable si l'on fait croître $\alpha$ en gardant
 $k$ constant.
 
-Les conclusions de ce théorème démontreint une propriété clé du système: une fois que le réseau perd sa bivalence
-(ex: $\delta > 0$), il tend à se renverser et converger rapidement vers la couleur majoritaire et il devient alors
+Les conclusions de ce théorème démontrent une propriété clef du système: une fois que le réseau perd sa bivalence
+(ex: $\delta > 0$), il tend à se renverser et à converger rapidement vers la couleur majoritaire, et il devient alors
 suffisamment improbable qu'il rebascule vers la couleur minoritaire. C'est cette propriété fondamentale que nous
-exploitons dans nos protocoles, et ce qui les rend sûrs malgré le fait que l'échantillonnage se fait sur une petite
-partie du réseau de taille fixe. Le résultat principal qui s'en suit pour les garanties de sûreté de Snowflake consiste
+exploitons dans nos protocoles, et c'est ce qui les rend sûrs malgré le fait que l'échantillonnage se fait sur une petite
+partie du réseau de taille fixe. Le résultat principal qui s'ensuit pour les garanties de sûreté de fonctionnement de Snowflake consiste
 à trouver des régions (à partir de paramètres spécifiquement choisis) où la réversibilité n'est possible qu'à partir
@@
 
-Dans le cas de Snowflake, nous prenons comme hypothèse qu'une partie des noeuds est antagoniste. Avec Slush, une
+Dans le cas de Snowflake, nous prenons comme hypothèse qu'une partie des nœuds est antagoniste. Avec Slush, une
 fois que le réseau obtient une majorité significative pour l'une des propositions (ex: la couleur bleue), il devient
 improbable qu'une proposition minoritaire (ex: la couleur rouge) puisse être adoptée dans le futur (irréversibilité).
-De plus, les noeuds Slush ont simplement besoin d'exécuter le protocole pour un nombre déterministe $m$ de tours qui
-est connu en amont de l'exécution du protocole. Lorsqu'on introduit des noeuds antagonistes qui suivent des stratégies
-arbitraires, les noeuds ne peuvent simplement plus exécuter le protocole pour un nombre déterministe de tours vu que
-les adversaires peuvent affecter la valeur de $m$ de manière non déterministe. A la place, les noeuds légitimes doivent
-implémenter un mécanisme pour explicitement détecter que l'irréversibilité a été atteinte. Dans ce but, dans Snowflake
-chaque noeud légitime implémente une fonction de décision, $\mathcal{D}(u, \mathcal{S}_t, blue) \rightarrow \{0, 1\}$,
-qui est une variable aléatoire qui retourne $1$ si le noeud $u$ détecte que l'état d'irréversibilité a été atteint à
+De plus, les nœuds Slush ont simplement besoin d'exécuter le protocole pour un nombre déterministe $m$ de tours qui
+est connu en amont de l'exécution du protocole. Lorsqu'on introduit des nœuds antagonistes qui suivent des stratégies
+arbitraires, les nœuds ne peuvent tout simplement plus exécuter le protocole pour un nombre déterministe de tours puisque
+les adversaires peuvent affecter la valeur de $m$ de manière non déterministe. En revanche, les nœuds légitimes doivent
+implémenter un mécanisme pour explicitement détecter que l'irréversibilité a été atteinte. Dans ce but, 
+chaque nœud légitime de Snowflake implémente une fonction de décision, $\mathcal{D}(u, \mathcal{S}_t, blue) \rightarrow \{0, 1\}$,
+qui est une variable aléatoire qui retourne $1$ si le nœud $u$ détecte que l'état d'irréversibilité a été atteint à
 l'instant $t$ pour bleu. Ce mécanisme de décision est probabiliste, ce qui veut dire qu'il peut échouer, même s'il est
@@
 
-\noindent \emph{Ebauche de preuve}. Nous définissons la stratégie d'échec comme étant l'évênement lors duquel deux
-noeuds légitimes $u$ et $v$ décident entre le bleu et le rouge, ex: $\mathcal{D}(u, \mathcal{S}_t, bleu) \rightarrow 1$
+\noindent \emph{Ébauche de preuve}. Nous définissons la stratégie d'échec comme étant l'événement lors duquel deux
+nœuds légitimes $u$ et $v$ décident entre le bleu et le rouge, ex: $\mathcal{D}(u, \mathcal{S}_t, bleu) \rightarrow 1$
 et $\mathcal{D}(v, \mathcal{S}_{t'}, rouge) \rightarrow 1$, pour deux instants $t$ et $t'$. Nous modélisons une fois de
 plus le système comme un processus à temps aléatoire. L'espace des états est défini de la même manière que pour Slush.
-Néanmoins, nous observons des subtilités critiques. Tout d'abord, même si tous les noeuds légitimes acceptent une
-couleur, les noeuds byzantins peuvent toujours basculer vers l'autre couleur. Ensuite nous devons aussi prendre en
+Néanmoins, nous observons des subtilités critiques. Tout d'abord, même si tous les nœuds légitimes acceptent une
+couleur, les nœuds byzantins peuvent toujours basculer vers l'autre couleur. Puis, nous devons aussi prendre en
 considération le mécanisme de décision $\mathcal{D}(*)$. Pour cette analyse, nous nous affranchissons du besoin de
 garder l'historique de toutes les configurations du réseau soumises à toutes les stratégies antagonistes et prenons pour
-hypothèse qu'un noeud $u$ décide d'abord d'adopter la couleur bleue. Ensuite en fonction de l'état du réseau suite à la
-décision de $u$, nous calculons la probabilité qu'un autre noeud $v$ adopte le rouge, qui est fonction à la fois de la
+hypothèse qu'un nœud $u$ décide d'abord d'adopter la couleur bleue. Ensuite, en fonction de l'état du réseau suite à la
+décision de $u$, nous calculons la probabilité qu'un autre nœud $v$ adopte le rouge, qui est fonction à la fois de la
 probabilité que le réseau bascule vers l'état minoritaire bleu et que $v$ bascule vers cet état.
@@
 extrêmement sécurisées de Snowflake (ex: échec de sécurité avec une probabilité de $\leq \varepsilon$) quand le réseau
-atteint un biais de $\delta$, comme le montre la Figure~\ref{fig:states_feasible_solutions}. Un exemple concret est
-présenté dans la Figure~\ref{fig:fandepsilon}.
+atteint un biais de $\delta$, comme le montre la figure~\ref{fig:states_feasible_solutions}. Un exemple concret est
+présenté en figure~\ref{fig:fandepsilon}.
 
@@
 \input{figures/analysis2}
-\caption{Représentation de l'état d'irréversibilité qui existe quand -- même en la présence de $f$ noeuds byzantins --
-  le nombre de noeuds légitimes bleus dépasse le nombre de noeuds légitimes rouges de plus de $2\delta$.
+\caption{Représentation de l'état d'irréversibilité qui existe quand -- même en la présence de $f$ nœuds byzantins --
+  le nombre de nœuds légitimes bleus dépasse le nombre de nœuds légitimes rouges de plus de $2\delta$.
 }
@@
 ayant atteint l'état d'irréversibilité, alors l'évolution de la confiance envers la couleur majoritaire adoptée va
-augmenter constamment et dériver de plus en plus loin des noeuds de couleur minoritaire, jusqu'à rendre la réversibilité
+augmenter constamment et dériver de plus en plus loin des nœuds de couleur minoritaire, jusqu'à rendre la réversibilité
 de plus en plus improbable avec le temps. Si cette dérive finit par se retourner, alors l'analyse de la réversibilité
@@
 représente la zone tampon.
-Plus $\psi$ est grand, plus la capacité du mécanisme de décision à tendre vers une valeur est rapide. Si bien sûr $\psi$
-tend vers zéro ou devient négatif alors nous dépassons la limite supérieure de tolérance antagoniste pour le système
-paramétré, en conséquence l'attaquant peut, avec une haute probabilité, bloquer la convergence en choisissant simplement
-de ne pas répondre, et ceci même si les garanties de sûreté s'appliquent toujours.
+Plus $\psi$ est grand, plus la capacité du mécanisme de décision à tendre vers une valeur est rapide. Bien sûr, si $\psi$
+tend vers zéro ou devient négatif, alors nous dépassons la limite supérieure de tolérance antagoniste pour le système
+paramétré, et en conséquence l'attaquant peut, avec une haute probabilité, bloquer la convergence en choisissant simplement
+de ne pas répondre, ceci même si les garanties de sûreté de fonctionnement s'appliquent toujours.
 
@@
 un, mais nous avons aussi une probabilité strictement positive de convergence jusqu'à un instant borné $t_{max}$, comme
-décrit dans la Proposition~\ref{lemma:finitetermination}, qui découle du Théorème~\ref{theorem:mean-convergence-time}.
+décrit dans la proposition~\ref{lemma:finitetermination}, qui découle du théorème~\ref{theorem:mean-convergence-time}.
 Cela précise la propriété de vitalité P2.
 
-\noindent \emph{Ebauche de preuve}. En utilisant la construction du système pour prouver l'irréversibilité, nous
-caractérisons la distribution de la moyenne du temps passé (les périodes "sojourn") dans chaque état avant que le
+\noindent \emph{Ébauche de preuve}. En utilisant la construction du système pour prouver l'irréversibilité, nous
+caractérisons la distribution de la moyenne du temps passé (les périodes de séjour) dans chaque état avant que le   % HOULA sojourn
 système ne termine son exécution en adoptant l'un des états. Ce temps de convergence est l'union de ces périodes.
 
-Pour les transactions non conflictuelles, vu que l'attaquant est incapable de générer un conflit, le temps de décision
-est simplement le temps de mélange, qui est $\Oh{\log{n}}$. La vitalité guarantit qu'une configuration de réseau
+Pour les transactions non conflictuelles, l'attaquant étant incapable de générer un conflit, le temps de décision
+est simplement le temps de mélange $\Oh{\log{n}}$. La vitalité guarantit qu'une configuration de réseau
 entièrement bivalent atteint un temps de convergence optimal de $\Oh{\log{n}}$ tours si le nombre d'attaquants est au
-plus $\Oh{\sqrt{n}}$, pour $\alpha = \floor{k/2} + 1$. Nous donnons plus de détails dans la
-Proposition~\ref{lemma:centrallimit}.
-Quand le nombre de noeuds antagonistes dépasse $\Oh{\sqrt{n}}$, le nombre le plus pessimiste de tours grandit de manière
-polynomiale, et quand $f$ tend vers $n/2$ ce nombre de tours tend vers des taux de convergence exponentiels.
-
-\noindent \emph{Ebauche de preuve}. Nous modifions le Théorème~\ref{theorem:mean-convergence-time} en y incluant le
-nombre de nodes antagonistes, ce qui annule toute déséquilibre dans le réseau en le gardant entièrement bivalent.
+plus $\Oh{\sqrt{n}}$, pour $\alpha = \floor{k/2} + 1$. Nous donnons plus de détails en
+lemme~\ref{lemma:centrallimit}.
+Quand le nombre de nœuds antagonistes dépasse $\Oh{\sqrt{n}}$, le nombre le moins favorable de tours grandit de manière
+polynomiale, et quand $f$ tend vers $n/2$, ce nombre de tours tend vers des taux de convergence exponentiels.
+
+\noindent \emph{Ebauche de preuve}. Nous modifions le théorème~\ref{theorem:mean-convergence-time} en y incluant le
+nombre de nœuds antagonistes, ce qui annule tout déséquilibre dans le réseau en le gardant entièrement bivalent.
 
@@
 
-\paragraph{Consensus multi-valeur}
-
-Notre protocole de consensus binaire pourrait supporter des consensus multi-valeur en exécutant des instances binaires
+\paragraph{Consensus multi-valeurs}
+
+Notre protocole de consensus binaire pourrait supporter des consensus multi-valeurs en exécutant des instances binaires
 logarithmiques, une pour chaque bit de la valeur proposée. Cependant, une telle réduction théorique n'est pas
-forcément efficace en pratique. A la place, nous pourrions directement incorporer des valeurs multiples comme des
-couleurs multiples dans le protocole en permettant toujours la généralisation de l'analyse de sûreté.
-
-Pour ce qui est de la vitalité nous ébauchons en Annexe~\ref{sec:sync-heuristic} un mécanisme d'initialisation sans
-leader, qui de toute attente nécessite $\Oh{\log{n}}$ tours dans l'hypothèse que le réseau soit synchronisé. Bien que
-la conception de ce mécanisme soit intéréssante, il est à noter qu'un tel mécanisme n'est pas nécéssaire pour un système
-de paiement décentralisé, comme nous le démontrons en Section~\ref{sec:implementation}.
-Finalement, nous développons les divergences entre vues et pertes en Annexe~\ref{sec:full-analysis-churn}.
+forcément efficace en pratique. À la place, nous pourrions directement incorporer des valeurs multiples comme des
+couleurs multiples dans le protocole en permettant toujours la généralisation de l'analyse de sûreté de fonctionnement.
+
+Pour ce qui est de la vitalité, nous ébauchons en annexe~\ref{sec:sync-heuristic} un mécanisme d'initialisation sans
+dirigean, qui nécessite $\Oh{\log{n}}$ tours attendus dans l'hypothèse que le réseau est synchronisé. Bien que
+la conception de ce mécanisme soit intéressante, il est à noter qu'un tel mécanisme n'est pas nécessaire pour un système
+de paiement décentralisé, comme nous le démontrons en section~\ref{sec:implementation}.
+Finalement, nous développons le roulement, dans le sens de renouvellement des nouds, et les divergences de vues en annexe~\ref{sec:full-analysis-churn}.
 
@@
 In this section, we describe the design and sketch how the implementation can support the value transfer primitive at the center of cryptocurrencies.
-Grâce au consensus Snowball, nous avons implémenté un système de paiement simpliste, Avalanche, qui supporte les transactions Bitcoin.
+Grâce au consensus Snowball, nous avons implémenté un système de paiement élémentaire, Avalanche, qui supporte les transactions Bitcoin.
 Dans cette section, nous en décrivons la conception et schématisons la manière dont l'implémentation peut supporter la primitive de transfert de valeur au cœur des cryptomonnaies.
@@
 Dans une configuration de cryptomonnaie, les signatures cryptographiques imposent que seul le possesseur de la clef soit capable de créer une transaction dépensant un dépôt donné.
-Comme les clients corrects suivent le protocole tel qu'il est prescrit et ne peuvent pratiquer la double dépense, ils ont la garantie tant de sûreté de fonctionnement que de vitalité pour leurs transactions \emph{vertueuses}.
+Comme les clients corrects suivent le protocole tel qu'il est prescrit et ne peuvent pratiquer la double dépense, ils ont la garantie tant de la sûreté de fonctionnement que de la vitalité pour leurs transactions \emph{vertueuses}.
 Par contraste, la vitalité n'est pas garantie pour les transactions \emph{malveillantes} soumises par des clients byzantins, lesquelles entrent mutuellement en conflit.
@@
 %O UTXOs are fully consumed by a valid transaction, and may generate new UTXOs spendable by named recipients.
-Les UTXO sont totalement consommées par une transaction valide et peut engendrer de nouvelles UTXO que des destinataires nommés peuvent dépenser.
+Les UTXO sont totalement consommées par une transaction valide et peuvent engendrer de nouvelles UTXO que des destinataires nommés peuvent dépenser.
 %O Multi-input transactions consume multiple UTXOs, and in Avalanche, may appear in multiple conflict sets.
@@
 In other words, a transaction is accepted only if all its transaction-input pairs are accepted in their respective Snowball conflict sets.
-Nous utilisons la conjonction de \textsc{isAccepted} pour toutes les entrées d'une transaction pour nous assurer qu'aucune transaction ne sera acceptée à moins que toutes ses entrées ne soient acceptées (Figure~\ref{fig:cash-system-b}).
+Nous utilisons la conjonction de \textsc{isAccepted} pour toutes les entrées d'une transaction pour nous assurer qu'aucune transaction ne sera acceptée à moins que toutes ses entrées ne soient acceptées (figure~\ref{fig:cash-system-b}).
 En d'autres termes, une transaction n'est acceptée que si toutes ses paires d'entrées de transaction sont acceptées dans leurs ensembles de conflits Snowball respectifs.
@@
 Correct nodes examine the DAG for virtuous transactions that lack sufficient progeny and emit no-op transactions to help increase their confidence.
-Un mécanisme secondaire s'assure que les transactions vertueuses avec une ascendance décidée reçoit suffisamment de \emph{chits}.
+Un mécanisme secondaire s'assure que les transactions vertueuses avec une ascendance décidée reçoivent suffisamment de chits.
 Les nœuds corrects examinent le DAG pour y trouver des transactions vertueuses manquant d'une progéniture suffisante et émettent des transactions \emph{no-op} pour accroître leur confiance.
@@
 %O side of a subgraph will not subvert the protocol.
-Comme la confiance est bâtie sur les \emph{chits} récoltés, et non simplement par la présence d'un sommet, la simple inondation du réseau par des sommets attachés à une partie rejetée du sous-graphe ne subvertit pas le protocole.
+Comme la confiance se fonde sur les chits récoltés, et non simplement par la présence d'un sommet, la simple inondation du réseau par des sommets attachés à une partie rejetée du sous-graphe ne subvertit pas le protocole.
 
@@
 %O They scale well, achieve high throughput and quick finality, work without precise membership knowledge, and degrade gracefully under catastrophic adversarial attacks.
-Ils passent bien à l'échelle, permettent un haut débit et une finalité rapide, ne nécessitent pas une connaissance précise des participants et leur fonctionnement ne se dégrade que peu à peu lors d'attaques catastrophiques.
+Ils passent bien à l'échelle, permettent un haut débit et une finalité rapide, ne nécessitent pas une connaissance précise des participants et leur fonctionnement se dégrade progressivement lors d'attaques catastrophiques.
 

### 2020-11-14 sections/evaluation_fr.tex
Relecture jusqu'à l'annexe

@@
 des instances \texttt{c5.large}, chacune simulant un noeud individuel. AWS fournit une bande passante
-allant jusqu'à 2 Gigabits par seconde, sachant que le protocole {\sysname} utilise au maximum 100 Megabits
+allant jusqu'à 2 gigabits par seconde, sachant que le protocole {\sysname} utilise au maximum 100 mégabits
 par seconde environ.
 
-Notre implémentation supporte deux types de transactions: la première est le format d'UTXO
+Notre implémentation supporte deux types de transactions\,: la première est le format d'UTXO
 customisé, tandis que l'autre repose directement sur le code de Bitcoin 0.16. Ces deux formats
@@
 batch de 40. Notre système sature plus vite avec des petites tailles de batch en comparaison avec d'autres
-blockchains dont les performances sont connues: Bitcoin regroupe des batchs de plusieurs milliers de
-transactions par bloc, Algorand~\cite{GiladHMVZ17} gère des blocks de 2-10 MégaOctets, i.e., 8.4--41.9K
-transactions par batch et Conflux~\cite{confluxLLXLC18} gère des blocs de 4 MégaOctets, i.e., 16.8K tx/batch.
+blockchains dont les performances sont connues\,: Bitcoin regroupe des batchs de plusieurs milliers de
+transactions par bloc, Algorand~\cite{GiladHMVZ17} gère des blocks de 2-10 mégaoctets, c'est-à-dire 8.4--41.9K
+transactions par batch et Conflux~\cite{confluxLLXLC18} gère des blocs de 4 mégaoctets, c'est-à-dire 16.8K tx/batch.
 Ces systèmes sont relativement lents pour prendre une simple décision, et demandent de surcroît une taille
@@
 
-Avalanche tire sa bonne mise à l'échelle de plusieurs sources: premièrement en maintenant un ordre partiel qui
+Avalanche tire sa bonne mise à l'échelle de plusieurs sources\,: premièrement en maintenant un ordre partiel qui
 capture uniquement les relations entre les dépenses ce qui permet une meilleure parallélisation qu'un journal
-BFT répliqué classique qui traite toutes les transactions de manière linéaire; deuxièmement l'absence de leader
-permet d'éviter les goulots d'étranglement; et finalement le nombre de messages que chaque noeud doit gérer par
+BFT répliqué classique qui traite toutes les transactions de manière linéaire\,; deuxièmement l'absence de dirigeant
+permet d'éviter les goulets d'étranglement\,; et finalement le nombre de messages que chaque noeud doit gérer par
 décision est $O(k)$ et n'augmente pas au fur et à mesure que le réseau s'agrandit.
@@
 
-Nous examinons ensuite où se situent les goulots d'étranglement dans notre implémentation.
+Nous examinons ensuite où se situent les goulets d'étranglement dans notre implémentation.
 La barre violette à la droite de chaque groupe dans la Figure~\ref{fig:eval-thr-raw} montre le débit
@@
 par rapport à la barre bleue sur la gauche.
-Cela révèle que le surplus de calcul dû à la vérification cryptographique est le principal goulot d'étranglement
-de l'implémentation de notre système. Ce problème peut être mitigé en déléguant la vérification des transactions
+Cela révèle que le surplus de calcul dû à la vérification cryptographique est le principal goulet d'étranglement
+de l'implémentation de notre système. On peut pallier ce problème en déléguant la vérification des transactions
 à un GPU\@. Même en l'absence d'une telle optimisation, 7K tps dépasse de loin ce qui existe en matière de blockchain.
@@
 
-\subsection{Clients malicieux}
+\subsection{Clients malveillants}
 \label{sec:evaluation-misbehaving}
@@
 \includegraphics[width=\linewidth]{figures/lat3.pdf}
-\captionof{figure}{Latence en fonction du taux de transactions véreuses.}
+\captionof{figure}{Latence en fonction du taux de transactions malveillantes.}
 \label{fig:eval-lat3}
@@
 
-Nous évaluons ensuite comment des transactions véreuses générées par des clients malicieux qui effectuent des
-double dépenses sur des sorties non dépensées peuvent affecter la latence des transactions légitimes crées par
-des clients honnêtes. Nous adoptons une stratégie qui simule des clients malicieux où une fraction (de $0\%$ à $25\%$)
+Nous évaluons ensuite comment des transactions conflictuelles générées par des clients malveillants effectuant des
+double dépenses sur des sorties non dépensées peuvent affecter la latence des transactions légitimes créées par
+des clients honnêtes. Nous adoptons une stratégie qui simule des clients malveillants où une fraction (de $0\%$ à $25\%$)
 des transactions en attente sont en conflit avec d'autres transactions existantes.
@@
 
-La latence d'{\sysname} est seulement légèrement influencée par les clients malicieux, comme le montre la
-Figure~\ref{fig:eval-lat3}. Etonnement la latence maximale baisse légèrement lorsque le pourcentage de transactions
-véreuses augmente. Ce comportement s'explique par le fait que lorsqu'on introduit des transactions véreuses, le débit
+La latence d'{\sysname} est seulement légèrement influencée par les clients malveillants, comme le montre la
+Figure~\ref{fig:eval-lat3}. Étonnamment, la latence maximale baisse légèrement lorsque le pourcentage de transactions
+conflictuelles augmente. Ce comportement s'explique par le fait que lorsqu'on introduit de telles transactions, le débit
 \emph{effectif} est réduit et soulage donc la charge du système.
 
-Cette observation est confirmée par la Figure~\ref{fig:eval-thr2}, qui montre que le débit (de transactions légitimes)
-decroît avec le taux de transactions véreuses. De plus, la baisse du débit apparaît proportionnelle au nombre de
-clients malicieux ce qui démontre que les attaques potentielles ne bénéficient pas d'un effet de levier.
+Cette observation est confirmée par la figure~\ref{fig:eval-thr2} qui montre que le débit (de transactions légitimes)
+decroît avec le taux de transactions conflictuelles. De plus, la baisse du débit apparaît proportionnelle au nombre de
+clients malveillants, ce qui démontre que les attaques potentielles ne bénéficient pas d'un effet de levier.
 
@@
 \includegraphics[width=\linewidth]{figures/thr-byz.pdf}
-\captionof{figure}{Débit de transactions en fonction du taux de transactions véreuses.}
+\captionof{figure}{Débit de transactions en fonction du taux de transactions conflictuelles.}
 \label{fig:eval-thr2}
@@
 détaillé dans un travail préliminaire~\cite{GiladHMVZ17}.
-Nous avons selectionné 20 villes majeures qui semblent être localisées à proximité d'un nombre conséquent de noeuds
-Bitcoin joignables, d'après ~\cite{bitnodes2018}. Ces villes couvrent l'Amérique du nord, l'Europe, l'Asie occidentale,
+Nous avons sélectionné 20 villes majeures qui semblent être localisées à proximité d'un nombre conséquent de noeuds
+Bitcoin joignables, d'après ~\cite{bitnodes2018}. Ces villes couvrent l'Amérique du Nord, l'Europe, l'Asie occidentale,
 l'Asie orientale, l'Océanie, et incluent également les 10 villes où l'on trouve le plus grand nombre de noeuds
@@
 noeuds d'une même ville. Comme pour l'évaluation d'Algorand, nous limitons aussi la bande passante par processus
-à 20 Mégabits par seconde dans le but de simuler des paramètres à l'échelle d'internet d'un réseau comportant de
+à 20 mégabits par seconde dans le but de simuler des paramètres à l'échelle d'internet d'un réseau comportant de
 nombreux liens. Nous assignons ensuite un processus client à chaque ville, qui génère 400 transactions sortantes
@@
 
-En conséquent, nous choisissons Algorand and Conflux pour notre étude comparative. Algorand, Conflux, et Avalanche
-sont tous fondamentalement différents dans leur conception. L'algorithme de consensus à l'échelle d'un comité
-d'Algorand repose sur un accord byzantin basé sur un quorum, et Conflux supplémente le consensus de Nakamoto d'une
-structure type Graphe Acyclique Dirigé (DAG) pour permettre un meilleur débit, tandis qu'Avalanche appartient à
+En conséquent, nous choisissons Algorand et Conflux pour notre étude comparative. Algorand, Conflux et Avalanche
+sont tous fondamentalement différents dans leur conception. L'algorithme de consensus par comité
+d'Algorand repose sur un accord byzantin basé sur un quorum, et Conflux supplémente le consensus de Nakamoto par une
+structure de type Graphe Acyclique Orienté (DAG) pour permettre un meilleur débit, tandis qu'Avalanche appartient à
 une nouvelle famille de protocoles basés sur la méta-stabilité. Par ailleurs, nous utilisons
@@
 
-Les évaluations d'Algorand et Avalanche reposent toutes deux sur un réseau décisionnel de 2000 noeuds sur EC2.
+Les évaluations d'Algorand et d'Avalanche reposent toutes deux sur un réseau décisionnel de 2000 noeuds sur EC2.
 Notre évaluation se fait sur des instances \texttt{c5.large} partagées, tandis qu'Algorand utilise des instances
 \texttt{m4.2xlarge}\@. Ces deux plateformes sont très similaires à l'exception d'un léger avantage en terme de
-fréquence CPU pour le \texttt{c5.large}, qui au final n'est que partiellement utilisé vu que nos processus ne
+fréquence CPU pour le \texttt{c5.large}, qui au final n'est que partiellement utilisé puisque nos processus ne
 consomment que $30\%$ de temps processeur lors de ces expériences. Les paramètres de sécurité choisis pour nos
@@
 $20\%$ de noeuds byzantins, en comparaison de l'évaluation Algorand qui garantit une probabilité de violation en dessous
-de $5 \times 10^{-9}$ avec $20\%$ noeuds byzantins.
+de $5 \times 10^{-9}$ avec $20\%$ de noeuds byzantins.
 
 Ni Algorand ni Conflux ne prennent en compte le surplus de calcul lié à la vérification cryptographique dans leurs
-études. Leurs évaluations utilisent des blocs qui contiennent des MégaOctets de données factices et présentent les
-mesures de débit en MégaOctets ou GigaOctets par heure. Nous utilisons donc la taille moyenne d'une transaction
-Bitcoin, 250 octets, pour en déduire leurs débits de transactions. Par contraste nos expériences transportent des
+études. Leurs évaluations utilisent des blocs qui contiennent des mégaoctets de données factices et présentent les
+mesures de débit en mégaoctets ou gigaoctets par heure. Nous utilisons donc la taille moyenne d'une transaction
+Bitcoin, 250 octets, pour en déduire leurs débits de transactions. Par contraste, nos expériences transportent des
 données de transactions réelles et prennent entièrement en compte le temps de calcul des vérifications cryptographiques.
 
-Le débit de transactions est de 3-7 tps pour Bitcoin, 874 tps pour Algorand (avec des blocs de 10 MégaOctets), 3355 tps
+Le débit de transactions est de 3-7 tps pour Bitcoin, 874 tps pour Algorand (avec des blocs de 10 mégaoctets), 3355 tps
 pour Conflux (le papier fait mention d'un débit 3.84 fois supérieur au débit d'Algorand dans les mêmes conditions).
 
-En comparaison, {\sysname} obtient plus de 3400 tps de manière consistante sur un réseau allant jusqu'à 2000 noeuds
-sans aucun comité de décision ou preuve de travail. Pour ce qui est de la latence, une transaction est confirmée
+En comparaison, {\sysname} obtient constamment plus de 3400 tps sur un réseau allant jusqu'à 2000 noeuds
+sans aucun comité de décision ni preuve de travail. Pour ce qui est de la latence, une transaction est confirmée
 après 10--60 minutes pour Bitcoin, environ 50 secondes pour Algorand, 7.6--13.8 minutes avec Conflux, et 1.35 secondes
@@
 
-Avalanche se comporte bien mieux qu'Algorand à la fois en terme de débit et de latence vu qu'Algorand utilise une
+Avalanche se comporte bien mieux qu'Algorand à la fois en terme de débit et de latence étant donné qu'Algorand utilise une
 fonction aléatoire vérifiable pour élire ses comités, et maintient un journal ordonné pendant qu'{\sysname} établit
 un ordre partiel uniquement. Bien que l'entité à la tête d'Algorand soit anonyme et change continuellement, elle reste
-néanmoins basée sur un leader unique qui pourrait être un facteur limitant pour la mise à l'échelle, tandis
+néanmoins basée sur un dirigeant unique qui pourrait être un facteur limitant pour le passage à l'échelle, tandis
 qu'{\sysname} n'est pas contrôlé par une entité unique.

### 2020-11-14 paper_fr.tex
Corrections after review of the appendix

@@
 It does not require accurate knowledge of all participants and exposes new possible tradeoffs and improvements in safety and liveness for building consensus protocols.
-Contrairement aux protocoles de consensus traditionnels où, typiquement, un ou plusieurs nœuds traitent linéairement l'information par décision, aucun nœud ne traite l'information plus que de manière logarithmique.
-% HOULA
+Contrairement aux protocoles de consensus traditionnels où, typiquement, un ou plusieurs nœuds traitent linéairement l'information pour chaque décision, aucun nœud ne traite l'information d'une manière plus complexe que logarithmique.
+Il ne requiert pas de connaissance précise de tous les participants et propose de nouveaux compromis et améliorations en termes de sûreté de fonctionnement et de vitalité pour la construction de nouveaux protocoles de consensus.
 %Finally, unlike traditional consensus protocols and similar to longest-chain protocols, our protocols provide increase in both liveness and safety guarantees when the observed adversarial presence decreases below the maximal expected bound.
@@
 %O \paragraph{Achieving Liveness}
-\paragraph{Obtention de la vitalité} % HOULA
+\paragraph{Objectifs de vitalité}
 %O Classical consensus that works with asynchrony does not get stuck in a single phase of voting because the vote initiator always polls votes from all known participants and waits for $n - f$ responses.
@@
 %O The adversary can view the state of every honest node at all times and can instantly modify the state of all adversarial nodes. 
-L'adversaire peut voir l'état de chaque nœud honnête à tout moment et peut instantanément modifier l'état de tout nœud antagoniste.
+L'adversaire peut voir l'état de chaque nœud correct à tout moment et peut instantanément modifier l'état de tout nœud antagoniste.
 %O It cannot, however, schedule or modify communication between correct nodes.
@@
 Nous partons du principe que tous les nœuds partagent le même $\mathcal{N}$, et 
-nous expliquons en annexe~\ref{sec:full-analysis-churn} comment assouplir l'exigence de la connaissance des participants. Nous modélisons
-les aspects dynamiques du système grâce à un processus à temps continu contenant deux états, tous les nœuds étant rouges ou tous  % HOULA dynamics → dynamique, comme economics
-les nœuds étant bleus. Soit $\{X_{t \geq 0}\}$ la variable aléatoire qui décrit l'état du système à l'instant $t$, où
-$X_0 \in \{0, \dots, c\}$.
+nous expliquons en annexe~\ref{sec:full-analysis-churn} comment assouplir l'exigence de la connaissance des participants.
+Nous modélisons la dynamique du système grâce à un processus à temps continu contenant deux états, tous les nœuds étant
+rouges ou tous les nœuds étant bleus. Soit $\{X_{t \geq 0}\}$ la variable aléatoire qui décrit l'état du système à
+l'instant $t$, où $X_0 \in \{0, \dots, c\}$.
 Nous commençons par détailler le résultat le plus important de la dynamique de la sûreté de fonctionnement de notre
@@
 Néanmoins, l'utilisation de la borne de Chvatal nous permet de faire l'observation qui consiste à dire qu'au fur et à
-mesure que la dérivation $\delta$ augmente, pour un $\alpha$ et un $k$ constants, la probabilité de basculer vers la   % HOULA drift
+mesure que la déviation $\delta$ augmente, pour un $\alpha$ et un $k$ constants, la probabilité de basculer vers la
 valeur minoritaire décroît \emph{de manière exponentielle} (en fait encore plus rapidement puisqu'il y a un terme
@@
 
-\section{Related Work}
+\section{Travaux associés}
 \label{sec:related-work}
@@
 \end{equation}
-where $\mathcal{D}(p-\psi, p)$ est la divergence de Kullback-Leibler, measurée par
+où $\mathcal{D}(p-\psi, p)$ est la divergence de Kullback-Leibler, measurée par
 \begin{equation}
@@
 These rates are
-Dans cet article, nous utilisons la notation $Q \equiv (q_{ij}, i, j \in \mathcal{S})$ pour nous référer au générateur infinitésimal du processus, où les taux de mort ($i \rightarrow i-1$) et de naissance ($i \rightarrow i+1$) de transitions de configuration sont notées via $\mu_i$, et$\lambda_i$ ($\lambda_ i$ est disctint du paramètre d'horloge $\lambda$ et sera sorti du contexte).
+Dans cet article, nous utilisons la notation $Q \equiv (q_{ij}, i, j \in \mathcal{S})$ pour nous référer au générateur infinitésimal du processus, où les taux de mort ($i \rightarrow i-1$) et de naissance ($i \rightarrow i+1$) de transitions de configuration sont notées via $\mu_i$, et $\lambda_i$ ($\lambda_ i$ est disctint du paramètre d'horloge $\lambda$ et sera sorti du contexte).
 Ces taux sont
@@
 We compute the closed-form expression for reversibility, and show that it is exponentially small.
-En section ~\ref{sec:analysis}, nous avons traité de la borne \emph{loose} % HOULA
-de Chtaval qui a fourni une compréhension intuitive de la dynamique de forte irréversibilité au cœur de notre mécanisme de sous-échantillonnage.
+En section ~\ref{sec:analysis}, nous avons traité la borne de Chtaval qui fournit une compréhension intuitive de la dynamique de forte irréversibilité au cœur de notre mécanisme de sous-échantillonnage.
 En particulier, une fois que le réseau tend vers une valeur majoritaire donnée, il ne peut s'inverser qu'avec une probabilité exponentiellement petite.
@@
 %O where from now on we refer to $\mathcal{u}_{\delta+1}$ as the drift of the process. 
-où à partir de maintenant nous nous référons à $\mathcal{u}_{\delta+1}$ par le drift du processus.
+où à partir de maintenant nous nous référons à $\mathcal{u}_{\delta+1}$ comme étant le drift du processus.
 \end{theorem}
@@
 Additionally, since termination is finite and everywhere differentiable, Theorem~\ref{theorem:mean-convergence-time} also implies that the probability of termination of any network configuration where a proposal has $\geq \alpha$ support in bounded time $t_{max}$ is strictly positive.
-En partant d'un quelconque état non-absorbant et transitoire, il y a une probabilité non nulle d'absorbtion.
+En partant d'un quelconque état non-absorbant et transitoire, il y a une probabilité non nulle d'absorption.
 De plus, comme la terminaison est finie et différentiable partout, le théorème~\ref{theorem:mean-convergence-time} implique aussi que la probabilité de terminaison de toute configuration réseau où une proposition a un support $\geq \alpha$ dans un temps borné $t_{max}$ est strictement positive.
@@
 %O Hence, we can assume that once all correct nodes are the same color, the attack from the adversary will terminate since it is impractical to continue an attack. 
-Nous pouvons donc suppposer qu'une fois tous les nœuds corrects de la même couleur, l'attaque de l'adversaire se termine puisqu'il serait irréaliste de la poursuivre.
+Nous pouvons donc suppposer qu'une fois que tous les nœuds légitimes ont adopté la même couleur, l'attaque de l'adversaire se termine puisqu'il serait irréaliste de la poursuivre.
 
@@
 The guarantees that can be extracted hereon are that the confidences of the majority committed value (in our frame of reference is always blue), grow always more than those of the minority value, with high probability, drifting away as $t \rightarrow t_{max}$.
-Nous faisons l'observation suivante : si la confiance entre rouge et bleu est équivalente, alors l'adversaire a même levier % HOULA
-sur l'irréversibilité du système que dans Snowflake, quelle que soit la configuration du réseau. En fait, Snowflake peut être vu comme Snowball où les drifts % HOULA
-de confiance n'excèdent jamais un.
+Nous faisons l'observation suivante : si la confiance entre rouge et bleu est équivalente, alors l'adversaire a la même influence sur l'irréversibilité du système que dans Snowflake, quelle que soit la configuration du réseau.
+En fait, Snowflake peut être vu comme Snowball où les déviations des valeurs de confiance n'excèdent jamais un.
 La même analyse s'applique à Snowball comme à Snowflake, avec la contrainte supplémentaire de bornage des comportements des confiances à long terme sur le réseau.
@@
 %O This feature provides various benefits; however, it also can lead to some virtuous transactions that depend on a rogue transaction to suffer the fate of the latter.
-Cette caractéristique offre plusieurs avantages\,; cependant, elle peut aussi mener à ce que des transactions vertueuses dépendant d'une transaction malveillante souffrent le sort de celle-ci.
+Cette caractéristique offre plusieurs avantages\,; cependant, elle peut aussi mener à ce que des transactions vertueuses dépendantes d'une transaction malveillante subissent le même sort que celle-ci.
 %has the drawback that it can entangle the fate of some unfortunate virtuous transactions with the fate of rogue ones.
@@
 %O As a thought experiment, suppose that a transaction $T_i$ names a set of parent transactions that are all decided, as per local view.
-En guise d'expérience mentale, supposons qu'une transaction $T_i$ nomme un ensemble de transactions parent qui sont toutes décidées, selon la vue locale.
+En guise d'expérience mentale, supposons qu'une transaction $T_i$ nomme un ensemble de transactions parentes qui sont toutes décidées, selon leur point de vue local.
 %O If $T_i$ is sampled over a large enough set of successful queries without discovering any conflicts, then, since by assumption the entire ancestry of $T_i$ is decided, it must be the case (probabilistically) that we have achieved irreversibility.
-Si $T_i$ est échantillonnée sur un ensemble suffisamment grand de requêtes réussies découvrir aucun conflit, alors, puisque par supposition toute l'ascendance de $T_i$ est décidée, il s'ensuit (de manière probabiliste) que nous sommes arrivés à l'irréversibilité.
+Si $T_i$ est échantillonnée sur un ensemble suffisamment grand de requêtes réussies sans découvrir aucun conflit, alors, puisque par supposition toute l'ascendance de $T_i$ est décidée, il s'ensuit (de manière probabiliste) que nous sommes arrivés à l'irréversibilité.
 
@@
 It then selects the first value by the order as its initial state when it starts the actual consensus protocol.
-Pour améiorer la vitalité, nous pouvons utiliser des hypothèses de forte synchronie.
+Pour améliorer la vitalité, nous pouvons utiliser des hypothèses de forte synchronie.
 L'heuristique fonctionne comme suit.
@@
 %O In a cryptocurrency setting, the deterministic ordering function would incorporate fees paid out for every new proposal, which means that the adversary is financially limited in its ability to launch a fairness attack against the initialization.
-Dans une configuration de cryptomonnaie, la fonction de tri déterministe incorporerait les frais payés pour toute nouvelle proposition, ce qui signifie que l'adversaire serait financièrement limité pour lancer une attaque d'impartialité contre l'initialisation.
+Dans le cas d'une cryptomonnaie, la fonction de tri déterministe incorporerait les frais payés pour toute nouvelle proposition, ce qui signifie que l'adversaire serait financièrement limité pour lancer une attaque d'impartialité contre l'initialisation.
 
@@
 This represents the worst case view split.
-Pour caractériser le roulement, nous utilisons une construction d'intersections d'ensembles généralisés qui nous permet d'établir des arguments concernant les éclatements de vues du réseau.
-Avant de le formaliser, nous fournissons cette intuition\,: supposons que nous éclations le réseau en deux sous-ensembles entièrement indépendants mais parfaitement connectés.
+Pour caractériser le roulement, nous utilisons une construction généralisée d'intersection d'ensembles qui nous permet d'établir des arguments concernant les éclatements de vues du réseau.
+Avant de le formaliser, nous partons de cette intuition\,: supposons que nous éclations le réseau en deux sous-ensembles entièrement indépendants mais parfaitement connectés.
 Il est clair que l'adversaire byzantin gagne avec une probabilité de un puisqu'il peut envoyer deux transactions conflictuelles à chaque réseau indépendant, et chacun finaliserait immédiatement les transactions.
@@
 Pour prouver la sécurité de fonctionnement, il nous suffit de réutiliser exactement la même construction qu'en section~\ref{subsection:appendix_snowflake}, mais nous remplaçons l'ensemble originel $\mathcal{N}$ avec le nouvel ensemble qui nous intéresse, nommément $S_a \cup S_d \cup \mathcal{B}$ (c'est-à-dire que nous excluons $S_b$)~\footnote{Les ensembles sont symétriques dans cet exemple.}.
-Donc, pour trouver les $\tau$ maximal, nous nous contentons de remplacer $u_i$ et $\lambda_i$ par
+Donc, pour trouver le $\tau$ maximal, nous nous contentons de remplacer $u_i$ et $\lambda_i$ par
 \begin{equation}
@@
 et nous appliquons la même construction qu'en section~\ref{subsection:appendix_snowflake}.
-Alors que $\tau$ augmente, il est clair que $S_a$ (et, réciproquement, $S_b$) deviennent plus indépendant et dépendent moins des valeurs par les membres de $S_d$, augmentant ainsi les capacités de l'adversaire.
+Alors que $\tau$ augmente, il est clair que $S_a$ (et, réciproquement, $S_b$) deviennent plus indépendants et moins dépendents des valeurs proposées par les membres de $S_d$, augmentant ainsi les capacités de l'adversaire.
 \end{proof}

### 2020-11-14 sections/evaluation_fr.tex
Corrections after review of the appendix

@@
 double dépenses sur des sorties non dépensées peuvent affecter la latence des transactions légitimes créées par
-des clients honnêtes. Nous adoptons une stratégie qui simule des clients malveillants où une fraction (de $0\%$ à $25\%$)
+des clients corrects. Nous adoptons une stratégie qui simule des clients malveillants où une fraction (de $0\%$ à $25\%$)
 des transactions en attente sont en conflit avec d'autres transactions existantes.

### 2020-11-14 avalanche_fr.tex
Première relecture

@@
 %O Avalanche takes advantage of the DAG structure and uses a transaction's progeny.
-Alors que Snowball utilise des requêtes répétées et plusieurs compteurs pour obtenir le degré de confiance associé aux transactions conflictuelles (couleurs),
+Alors que Snowball utilise des requêtes répétées et plusieurs compteurs pour refléter le degré de confiance associé aux transactions conflictuelles (couleurs),
 Avalanche profite de la structure du DAG et utilise la progéniture d'une transaction.
@@
 %O partitioned into mutually exclusive conflict sets $\mathcal{P}_T$, $T \in \mathcal{T}_u$.
-Chaque nœud correct $u$ conserve la trace de toutes les transactions dont il a été informé dans l'ensemble $\mathcal{T}_u$, partitionné en ensembles de conflits mutuellement exclusifs $\mathcal{P}_T$, $T \in \mathcal{T}_u$.
+Chaque nœud correct $u$ conserve la trace de toutes les tran\-sactions dont il a été informé dans l'ensemble $\mathcal{T}_u$, partitionné en ensembles de conflits mutuellement exclusifs $\mathcal{P}_T$, $T \in \mathcal{T}_u$.
 % Transactions are determined to be conflicting based on a deterministic function known to every node.
@@
 %O The ``$\stackrel{*}{\gets}$''-relation is its reflexive transitive closure, indicating a path from $T$ to $T'$.
-La relation ``$\stackrel{*}{\gets}$'' est sa fermeture réflexive transitive, indiquant un chemin de $T$ vers $T'$.
+La relation ``$\stackrel{*}{\gets}$'' est sa fermeture réfle\-xive transitive, indiquant un chemin de $T$ vers $T'$.
 %O DAGs built by different nodes are guaranteed to be compatible, though at any one time, the two nodes may not have a complete view of all vertices in the system.

### 2020-11-14 paper_fr.tex
Première relecture

@@
-\documentclass[letterpaper,twocolumn,10pt]{article}
+\documentclass[a4,twocolumn,10pt]{article}
 \usepackage{usenix2019_v3}
@@
 \renewcommand{\proofname}{Preuve}
+
 \settimeformat{hhmmsstime}
@@
 Emin~G\"un~Sirer\\
-Cornell University\textsuperscript{*}
+Cornell University\textsuperscript{*}\\
+\small{Traduit par Grégory Lemercier et Jean Zundel}
 }{%
@@
 %O This paper introduces a family of leaderless Byzantine fault tolerance protocols, built around a metastable mechanism via network subsampling.
-Cet article présente une famille de protocoles à tolérance de faute byzantine sans dirigeant, bâtis autour d'un mécanisme métastable via sous-échantillonnage réseau.
+Cet article présente une famille de protocoles à tolérance de faute byzantine sans dirigeant, bâtis autour d'un mécanisme métastable \emph{via} sous-échantillonnage réseau.
 %O These protocols provide a strong probabilistic safety guarantee in the presence of Byzantine adversaries while their concurrent and leaderless nature enables them to achieve high throughput and scalability.
@@
 %O to large numbers of participants.
-Comme ils impliquent en général un nombre de communications quadratique et une connaissance exacte des membres du réseau, il s'est avéré difficile de l'étendre
+Comme ils impliquent en général un nombre de communications croissant de façon quadratique et une connaissance exacte des membres du réseau, il s'est avéré difficile de l'étendre
 à de nombreux participants.
@@
 % \Jon{I wanted the paper to expand on this point, but I missed it.}
-De plus, ces protocoles souffrent d'un goulet d'étranglement intrinsèque pour le passage à l'échelle, qui s'avère difficile à dépasser par la simple reparamétrisation~\cite{CromanDEGJKMSSS16}.
+De plus, ces protocoles souffrent d'un goulet d'étranglement intrinsèque pour le passage à l'échelle qui s'avère difficile à dépasser par la simple reparamétrisation~\cite{CromanDEGJKMSSS16}.
 %O This paper introduces a new family of consensus protocols called Snow.
@@
 %O Unlike Nakamoto consensus, the protocols are green, quiescent and efficient; they do not rely on proof-of-work~\cite{DworkN92} and do not consume energy when there are no decisions to be made.
-Contrairement au consensus Nakamoto, ces protocoles sont écologiques, économes et efficients ; ils ne dépendent pas d'une preuve de travail\cite{DworkN92} et ne consomment pas d'énergie quand il n'y a pas de décision à prendre.
+Contrairement au consensus Nakamoto, ces protocoles sont écologiques, économes et efficients\,; ils ne dépendent pas d'une preuve de travail\cite{DworkN92} et ne consomment pas d'énergie quand il n'y a pas de décision à prendre.
 %0 The efficiency of the protocols stems partly from removing the leader bottleneck: each node requires $\Oh{1}$ communication overhead per round and $\Oh{\log{n}}$ rounds in expectation, whereas classical consensus protocols have one or more nodes that require $\Oh{n}$ communication per round (phase).
@@
 Snowflake appartient à la famille Snow et est configuré avec $k=10$, $\beta=150$.
-Snowflake-7,8 utilise $\alpha=7$ et $\alpha=8$ respectivement (voir en section~\ref{sec:protocol} pour la définition de $k$, $\alpha$ et $\beta$.}
+Snowflake-7,8 utilise $\alpha=7$ et $\alpha=8$ res\-pectivement (voir en section~\ref{sec:protocol} pour la définition de $k$, $\alpha$ et $\beta$.}
     \label{fig:fandepsilon}
@@
     \includegraphics[width=\linewidth]{figures/safety-f-eps-log.pdf}
-    \caption{Figure~\ref{fig:fandepsilon} Axe des y en log.}
+%O    \caption{Figure~\ref{fig:fandepsilon}, axe des y en log.}
+    \caption{Figure~\ref{fig:fandepsilon} avec l'axe des y en log.}
     \label{fig:fandepsilonlog}
@@
 In particular, we fix the distribution of message delay to that of the exponential distribution.
-Nous employons une version modifiée de ce modèle, lequel est bien accepté~\cite{banerjee2014epidemic, ganesh2005effect, draief2006thresholds, keeling2011modeling, liggett1997stochastic} dans l'analyse des réseaux épidémiques et des systèmes stochastiques à base de \emph{gossip}.
+Nous employons une version mo\-difiée de ce modèle, lequel est bien accepté~\cite{banerjee2014epidemic, ganesh2005effect, draief2006thresholds, keeling2011modeling, liggett1997stochastic} dans l'analyse des réseaux épidémiques et des systèmes stochastiques à base de \emph{gossip}.
 En particulier, nous fixons la distribution des délais de messages à celle de la distribution exponentielle.
@@
 %O \paragraph{Adversary}
-\paragraph{Adversité} % HOULA : Adversaire ? Adversaires ? Adversatif ?
+\paragraph{Adversité}
 %O The adversarial nodes execute under their own internal scheduler, which is unbounded in speed, meaning that all adversarial nodes can execute at any infinitesimally small point in time, unlike correct nodes.
@@
 %O These bounds could be violated if the network is naively left open to arbitrary participants.
-Ces limites peuvent être violées si le réseau est naïvement laissé ouvert à des participants quelconques.
+Ces limites peuvent être violées si le réseau est naïvement laissé ouvert à des parti\-cipants quelconques.
 %O In particular, a Sybil attack~\cite{douceur2002sybil}, wherein a large number of identities are generated by an adversary, could be used to exceed the bounds.
@@
 %O A long line of work, including PBFT~\cite{castro1999practical}, treats the Sybil problem separately from consensus, and rightfully so, as Sybil control mechanisms are distinct from the underlying, more complex agreement protocol\footnote{This is not to imply that every consensus protocol can be coupled/decoupled with every Sybil control mechanism.}.
-Une longue suite de travaux, comme PBFT~\cite{castro1999practical}, sépare le problème de Sybil de celui du consensus, à juste titre puisque les mécanismes de contrôle de Sybil sont distincts du protocole d'accord plus complexe sous-jacent\footnote{Cela n'implique pas que tout protocole de consensus puisse être couplé/découplé de tout mécanisme de contrôle de Sybil}.
+Une longue suite de travaux, comme PBFT~\cite{castro1999practical}, sépare le problème de Sybil de celui du consensus, à juste titre puisque les mécanismes de contrôle de Sybil sont distincts du protocole d'accord plus complexe sous-jacent\footnote{Cela n'implique pas que tout protocole de consensus puisse être couplé ou découplé de tout mécanisme de contrôle de Sybil.}.
 %O In fact, to our knowledge, only Nakamoto-style consensus has ``baked-in'' Sybil prevention as part of its consensus, made possible by chained proof-of-work~\cite{aspnes2005exposing}, which requires miners to continuously stake a hardware investment.
@@
 ou \emph{gossip}. Le protocole le plus simple, Slush, est la fondation de cette famille, comme le montre la
-figure~\ref{fig:slush-loop}. Slush \emph{n'est pas} tolérant aux fautes byzantines, seulement aux fautes par défaut % HOULA : faute ou défaillance ?
+figure~\ref{fig:slush-loop}. Slush \emph{n'est pas} tolérant aux fautes byzantines, seulement aux défauts
 de fonctionnement (CFT), mais sert à mettre en avant les autres protocoles tolérants aux fautes byzantines qui vont
@@
 Finalement, si $m$ est choisi pour être suffisamment élevé, Slush s'assure que tous les nœuds convergent vers la
-même couleur avec une forte probabilité. Chaque nœud nécessite un nombre de communications par tour constant et   % HOULA communication overhead
+même couleur avec une forte probabilité. Chaque nœud nécessite un nombre de communications par tour constant et
 prévisible, et $m$ croît de manière logarithmique suivant $n$.
@@
 
-Snowflake améliore Slush en ajoutant un compteur unique qui capture la force de conviction d'un nœud concernant sa couleur
+Snowflake améliore Slush en ajoutant un compteur unique qui reflète la force de conviction d'un nœud concernant sa couleur
 actuelle. Ce compteur associé à chaque nœud stocke le nombre d'échantillons du réseau consécutifs que le nœud a
@@
 La notion d'état de Snowflake est éphémère\,: le compteur est réinitialisé à chaque changement de couleur.
-Snowball améliore Snowflake avec des \emph{compteurs de confiance} qui capturent le nombre de requêtes ayant retourné
+Snowball améliore Snowflake avec des \emph{compteurs de confiance} qui reflètent le nombre de requêtes ayant renvoyé
 un résultat en faveur de la couleur du nœud au dessus d'un certain seuil (figure ~\ref{fig:snowball-loop}).
@@
 Nous notons que les bornes de Chvatal sont introduites ici dans un but de simplification et sont extrêmement faibles.
-Nous laissons l'expression dans sa forme complète dans le Théorème~\ref{theorem:slush_prob_convergence_minority}
+Nous laissons l'expression dans sa forme complète dans le théorème~\ref{theorem:slush_prob_convergence_minority}
 disponible en annexe, qui est aussi significativement plus fort que la borne de Chvatal.
@@
 décrit dans la proposition~\ref{lemma:finitetermination}, qui découle du théorème~\ref{theorem:mean-convergence-time}.
-Cela précise la propriété de vitalité P2.
+Cela formalise la propriété de vitalité P2.
 
 \noindent \emph{Ébauche de preuve}. En utilisant la construction du système pour prouver l'irréversibilité, nous
-caractérisons la distribution de la moyenne du temps passé (les périodes de séjour) dans chaque état avant que le   % HOULA sojourn
+caractérisons la distribution de la moyenne du temps passé (les périodes de séjour) dans chaque état avant que le
 système ne termine son exécution en adoptant l'un des états. Ce temps de convergence est l'union de ces périodes.
@@
 Nous disposons de solutions à ces problèmes mais leur description détaillée sort du cadre de cet article, dont l'objet est la nouvelle famille de consensus Snow.
-% HOULA staking unstaking
 %In this section,
@@
 %O transactions ($T \in \mathcal{T}$) in Avalanche's DAG \emph{vertices}.
-En addition à la structure de DAG dans Avalanche, un graphe de \emph{unspent transaction output} (UTXO)~\cite{nakamoto2008bitcoin}, ou sortie de transaction non dépensée, qui capture les dépendances des dépenses est utilisé pour réaliser le registre du système de paiement.
+En addition à la structure de DAG dans Avalanche, un graphe de \emph{unspent transaction output} (UTXO)~\cite{nakamoto2008bitcoin}, ou sortie de transaction non dépensée, qui reflète les dépendances des dépenses est utilisé pour réaliser le registre du système de paiement.
 Pour lever toute ambiguïté, nous appelons les transactions qui codent les données de transfert d'argent \emph{transactions} alors que nous appelons les transactions ($T \in \mathcal{T}$) dans le DAG d'Avalanche des \emph{sommets}.
@@
 %O One such improvement could be the introduction of an adversarial network scheduler.
-Une amélioration possible consisterait à introduire un ordonnanceur de réseaux adverses. % HOULA
+Une amélioration possible consisterait à introduire un coordinateur d'attaques sur le réseau.
 %O Another}{One} improvement would be to characterize the system's guarantees under an adversary whose powers are realistically limited, whereupon performance would improve even further.
@@
 %O Theorem~\ref{theorem:mean-convergence-time} leads to the next lemma that captures property P2, under the assumption that at the beginning of the protocol, one proposal has at least $\alpha$ support in the network.
-Le théorème~\ref{theorem:mean-convergence-time} mène au lemme suivant qui reflète % HOULA
+Le théorème~\ref{theorem:mean-convergence-time} mène au lemme suivant qui formalise
 la propriété P2, en suppposant qu'au début du protocole, une proposition possède un support d'au moins $\alpha$ dans le réseau.
@@
 %O The next lemma captures P3, and the proof follows from central limit theorem. 
-Le lemme suivant reflète P3 et la preuve procède du théorème central limite.
+Le lemme suivant formalise P3 et la preuve procède du théorème central limite.
 \begin{lemma}
@@
 The logarithmic factor remains from the mixing time lower bound.
-Le résultat procède du théorème central limite, où pour $\alpha = \floor{k/2} + 1$, la polarisation % HOULA
+Le résultat procède du théorème central limite, où pour $\alpha = \floor{k/2} + 1$, la polarisation
 attendue dans le réseau après échantillonnage sera $\Oh{\sqrt{n}}$.
@@
 Le facteur logarithmique provient de la limite inférieure du temps de mélange.
-% HOULA
 \end{proof}
@@
 Snowball peut être vu comme un système à deux urnes où chaque urne est une sous-martingale.
-Les garanties qui peuvent en être extraites ci-après sont que les confiances de la valeur engagée par la majorité (toujours bleu dans notre cadre de référence) croissent toujours plus que celles de la valeur minoritaire, avec une forte probabilité, tendant à s'écarter alors que $t \rightarrow t_{max}$.
-% HOULA TOUT LE PARAGRAPHE
+Les garanties qui peuvent en être extraites ci-après sont que les confiances en la valeur engagée par la majorité (toujours bleu dans notre cadre de référence) croissent toujours plus que celles de la valeur minoritaire, avec une forte probabilité, tendant à s'écarter alors que $t \rightarrow t_{max}$.
 %O \subsection{Safe Early Commitment}
@@
 %O This feature provides various benefits; however, it also can lead to some virtuous transactions that depend on a rogue transaction to suffer the fate of the latter.
-Cette caractéristique offre plusieurs avantages\,; cependant, elle peut aussi mener à ce que des transactions vertueuses dépendantes d'une transaction malveillante subissent le même sort que celle-ci.
+Cette caractéristique offre plusieurs avantages\,; cependant, elle peut aussi mener à ce que des transactions vertueuses dépendantes d'une tran\-saction malveillante subissent le même sort que celle-ci.
 %has the drawback that it can entangle the fate of some unfortunate virtuous transactions with the fate of rogue ones.
@@
 L'heuristique fonctionne comme suit.
-Chaque nœud opère en deux phases\,: dans la première, elle diffuse et collecte des propositions pendant $\Oh{\log{n}}$ tours, où chaque tour dure le temps maximum d'un message, ce qui garantit que la proposition d'un nœud correct se propage vers presque tous les autres nœuds corrects\,; dans la seconde phase, chaque nœud arrête de collecter les propositions dont il a localement la connaissance, en recherchant ceux qui ont une majorité $\alpha$, ordonnée de manière déterministe, comme par la valeur de leur empreinte cryptographique.
+Chaque nœud opère en deux phases\,: dans la première, elle diffuse et collecte des propositions pendant $\Oh{\log{n}}$ tours, où chaque tour dure le temps maximum d'un message, ce qui garantit que la proposition d'un nœud correct se propage vers presque tous les autres nœuds corrects\,; dans la se\-conde phase, chaque nœud arrête de collecter les propositions dont il a localement la connaissance, en recherchant ceux qui ont une majorité $\alpha$, ordonnée de manière déterministe, comme par la valeur de leur empreinte cryptographique.
 Il sélectionne ensuite la première valeur dans l'ordre pour en faire son état initial quand il commence vraiment le protocole de consensus.
@@
 Avant de le formaliser, nous partons de cette intuition\,: supposons que nous éclations le réseau en deux sous-ensembles entièrement indépendants mais parfaitement connectés.
-Il est clair que l'adversaire byzantin gagne avec une probabilité de un puisqu'il peut envoyer deux transactions conflictuelles à chaque réseau indépendant, et chacun finaliserait immédiatement les transactions.
+Il est clair que l'adversaire byzantin gagne avec une probabilité de un puisqu'il peut envoyer deux transactions conflictuelles à chaque réseau indépendant, et chacun fina\-liserait immédiatement les transactions.
 Cela représente le cas d'éclatement le plus défavorable.
@@
 Nous supposons que l'adversaire a le contrôle total des éclatements de la vue du réseau, ce qui signifie qu'il peut choisir sans contrainte de créer $S_a,\ S_b,\ S_d$.
-Pour prouver la sécurité de fonctionnement, il nous suffit de réutiliser exactement la même construction qu'en section~\ref{subsection:appendix_snowflake}, mais nous remplaçons l'ensemble originel $\mathcal{N}$ avec le nouvel ensemble qui nous intéresse, nommément $S_a \cup S_d \cup \mathcal{B}$ (c'est-à-dire que nous excluons $S_b$)~\footnote{Les ensembles sont symétriques dans cet exemple.}.
+Pour prouver la sécurité de fonctionnement, il nous suffit de réutiliser exac\-tement la même construction qu'en section~\ref{subsection:appendix_snowflake}, mais nous remplaçons l'ensemble originel $\mathcal{N}$ avec le nouvel ensemble qui nous intéresse, nommément $S_a \cup S_d \cup \mathcal{B}$ (c'est-à-dire que nous excluons $S_b$)~\footnote{Les ensembles sont symétriques dans cet exemple.}.
 Donc, pour trouver le $\tau$ maximal, nous nous contentons de remplacer $u_i$ et $\lambda_i$ par

### 2020-11-14 sections/evaluation_fr.tex
Première relecture

@@
 
-Comme le montre le premier groupe de barres dans la Figure~\ref{fig:eval-thr}, le système atteint 6851
+Comme le montre le premier groupe de barres dans la figure~\ref{fig:eval-thr}, le système atteint 6851
 transactions par seconde (tps) pour une taille de batch de 20 et plus de 7002 tps pour une taille de
@@
 
-La Figure~\ref{fig:eval-thr} montre que le débit global se dégrade d'environ $1.34\%$ à 6909 tps lorsque le
+La figure~\ref{fig:eval-thr} montre que le débit global se dégrade d'environ $1.34\%$ à 6909 tps lorsque le
 réseau augmente d'un facteur 16 pour atteindre $n = 2000$. Cette dégradation est mineure si on la compare au
@@
 Nous examinons ensuite où se situent les goulets d'étranglement dans notre implémentation.
-La barre violette à la droite de chaque groupe dans la Figure~\ref{fig:eval-thr-raw} montre le débit
+La barre violette à la droite de chaque groupe dans la figure~\ref{fig:eval-thr-raw} montre le débit
 d'Avalanche lorsque la vérification des signatures est désactivée. Les débits s'améliorent d'environ 2.6x, par
@@
 
-La Figure~\ref{fig:eval-lat1} représente un histogramme montrant la distribution des latences en utilisant la
+La figure~\ref{fig:eval-lat1} représente un histogramme montrant la distribution des latences en utilisant la
 même configuration que pour les mesures de débits à partir de 2000 noeuds. L'axe X représente le temps en secondes
@@
 
-La Figure~\ref{fig:eval-lat2} montre la latence de transaction pour un nombre variable de noeuds. Les arêtes
+La figure~\ref{fig:eval-lat2} montre la latence de transaction pour un nombre variable de noeuds. Les arêtes
 horizontales des boîtes représentent respectivement du bas vers le haut le minimum, le premier quartile, la médiane,
@@
 La latence d'{\sysname} est seulement légèrement influencée par les clients malveillants, comme le montre la
-Figure~\ref{fig:eval-lat3}. Étonnamment, la latence maximale baisse légèrement lorsque le pourcentage de transactions
+figure~\ref{fig:eval-lat3}. Étonnamment, la latence maximale baisse légèrement lorsque le pourcentage de transactions
 conflictuelles augmente. Ce comportement s'explique par le fait que lorsqu'on introduit de telles transactions, le débit
@@
 Nous avons sélectionné 20 villes majeures qui semblent être localisées à proximité d'un nombre conséquent de noeuds
-Bitcoin joignables, d'après ~\cite{bitnodes2018}. Ces villes couvrent l'Amérique du Nord, l'Europe, l'Asie occidentale,
+Bitcoin joignables, d'après~\cite{bitnodes2018}. Ces villes couvrent l'Amérique du Nord, l'Europe, l'Asie occidentale,
 l'Asie orientale, l'Océanie, et incluent également les 10 villes où l'on trouve le plus grand nombre de noeuds
@@
 Dans ce scénario, Avalanche atteint un débit de transactions moyen de 3401 tps, avec un écart type de 39 tps. Comme
-le montre la Figure~\ref{fig:eval-lat4}, la latence de transaction médiane est de 1.35 secondes, avec une latence
+le montre la figure~\ref{fig:eval-lat4}, la latence de transaction médiane est de 1.35 secondes, avec une latence
 maximale de 4.25 secondes. Nous supportons également le code initial de Bitcoin pour les transactions, qui dans ce

### 2020-11-14 sections/related-work_fr.tex
Première relecture

@@
 Bitcoin~\cite{nakamoto2008bitcoin} est une cryptomonnaie qui utilise une blockchain basée sur la preuve de travail
-(PoW pour "Proof Of Work") et qui tient un livre de comptes des "sorties de transaction non dépensées", ou UTXO
-("Unspent Transaction Output") en anglais. Bien que des techniques basées sur la preuve de
-travail~\cite{DworkN92, aspnes2005exposing}, et même d'autres cryptomonnaies dont la génération de pièces est basée
-sur la preuve de travail~\cite{vishnumurthy2003karma,rivest1997payword}, ont été explorées avant l'apparition de
+(PoW pour \emph{Proof Of Work}) et qui tient un livre de comptes des «\,sorties de transaction non dépensées\,», ou UTXO
+(\emph{Unspent Transaction Output}). Bien que des techniques basées sur la preuve de
+travail~\cite{DworkN92, aspnes2005exposing}, et même d'autres cryptomonnaies dont l'émission est basée
+sur la preuve de travail~\cite{vishnumurthy2003karma,rivest1997payword}, aient été explorées avant l'apparition de
 Bitcoin, ce dernier fut le premier à incorporer la preuve de travail dans son processus de consensus. Contrairement
 aux protocoles résistants aux fautes byzantines traditionnels, Bitcoin apporte une garantie de sûreté probabiliste et
-considère que la majorité de la puissance de calcul agit de manière honnête plutôt que de contrôler les membres du
+considère que la majorité de la puissance de calcul agit de manière honnête plutôt que de contrôler les participants au
 réseau, ce qui a permis de faire émerger un protocole sans permission à l'échelle d'internet. Bien qu'il soit sans
 permission et résilient face aux attaques, Bitcoin souffre d'un débit de transaction faible (\textasciitilde{}3 tps)
-et d'une latence élevée (\textasciitilde{}5.6 heures pour un réseau comportant 20\% de noeuds byzantins et une garantie
-de sûreté de $2^{-32}$). De plus la preuve de travail nécéssite une puissance de calcul considérable qui est consommée
+et d'une latence élevée (\textasciitilde{}5.6 heures pour un réseau comportant 20\% de nœuds byzantins et une garantie
+de sûreté de $2^{-32}$). De plus la preuve de travail nécessite une puissance de calcul considérable qui est consommée
 dans le seul et unique but de maintenir la sûreté de fonctionnement.
@@
 preuve de travail. Bitcoin-NG~\cite{EyalGSR16} et la version sans permission de Thunderella~\cite{PassS18} utilise
-un consensus similaire à celui de Nakamoto pour élire un leader qui va dicter les écritures dans le journal répliqué
+un consensus similaire à celui de Nakamoto pour élire un dirigeant qui va dicter les écritures dans le journal répliqué
 pour un temps donné assez long dans le but de fournir un meilleur débit de transactions. En outre, Thunderella apporte
-une borne optimiste qui, en considérant 3/4 de puissance de calcul et un leader élu honnêtes, permet de confimer les
+une borne optimiste qui, en considérant 3/4 de puissance de calcul et un dirigeant élu honnête, permet de confirmer les
 transactions rapidement. ByzCoin~\cite{Kokoris-KogiasJ16} sélectionne un échantillon réduit de participants de manière
-périodique et exécute ensuite un protocole résistant aux fautes byzantines en pratique (PBFT) entre les noeuds
+périodique et exécute ensuite un protocole résistant aux fautes byzantines en pratique (PBFT) entre les nœuds
 sélectionnés.
@@
 quorum afin d'optimiser le protocole pour obtenir des cas de fonctionnement sans contention et parvenir à un consensus
-en un seul tour de communication. Malgré tout, même si ces protocoles améliorent les performances, ils se dégradent
+en un seul tour de communications. Malgré tout, même si ces protocoles améliorent les performances, ils se dégradent
 fortement lorsqu'ils sont soumis à une forte contention. Zyzzyva~\cite{KotlaADCW09} ajoute au protocole BFT une
-exécution spéculative pour tendre vers un cas de fonctionnement sans erreurs. Les travaux passés portant sur les
+exécution spéculative pour tendre vers un cas de fonctionnement sans erreurs. Les travaux antérieurs portant sur les
 systèmes résistants aux fautes byzantines avec contrôle d'accès nécessitent typiquement au moins $f+1$ répliques.
@@
 D'autres travaux tentent d'introduire de nouveaux protocoles qui redéfinissent et relâchent les contraintes du modèle
-de résistance aux fautes byzantines. La résistance byzantine à grande échelle~\cite{rodrigues2007large} modifie la
-résistance byzantine pratique (PBFT) pour autoriser un nombre arbitraire de répliques et un seuil d'échec, en
-fournissant une garantie de vitalité probabilitse pour un taux d'échec donné tout en protégeant la sûreté de
+de résistance aux fautes byzantines. La résistance byzantine à grande échelle~\cite{rodrigues2007large} modifie
+PBFT pour autoriser un nombre quelconque de répliques et un seuil d'échec en
+fournissant une garantie de vi\-talité probabilitse pour un taux d'échec donné tout en protégeant la sûreté de
 fonctionnement avec une probabilité élevée. Dans une autre forme de relâchement des contraintes,
@@
 plutôt qu'une linéarisation de celle-ci, ce qui veut dire que les participants n'ont pas besoin d'être conhérents mais
-doivent se mettrent d'accord au final une fois le réseau stabilisé. En fournissant une garantie de cohérence encore
-plus faible, appelée "cohérence causale par embranchement-raccord", Depot~\cite{mahajan2011depot} définit un protocole
+doivent se mettrent finalement d'accord une fois le réseau stabilisé. En fournissant une garantie de cohérence encore
+plus faible, appelée «\,cohérence causale par \emph{fork-join}\,», Depot~\cite{mahajan2011depot} définit un protocole
 garantissant la sûreté de fonctionnement pour $2f+1$ répliques.
 
-NOW~\cite{guerraoui2013highly} uttilise des sous-quorums pour diriger des instances de consensus réduites.
-L'enseignement de ce papier est que de petits quorums de taille logarithmique peuvent être extraits à partir d'un
-échantillon potentiellement large de noeuds du réseau, ce qui permet à des instances de protocole de consensus plus
+NOW~\cite{guerraoui2013highly} uttilise des sous-quorums pour diriger des ins\-tances de consensus réduites.
+Cet article enseigne que de petits quorums de taille logarithmique peuvent être extraits à partir d'un
+échantillon potentiellement large de nœuds du réseau, ce qui permet à des instances de protocole de consensus plus
 petites de tourner en parallèle.
@@
 Snow White~\cite{cryptoeprint:2016:919} et Ouroboros~\cite{KiayiasRDO17} font partie des premiers protocoles à la
-sûreté prouvée basés sur la preuve d'enjeu (PoS - ou "Proof Of Stake" en anglais). Ouroboros utilise un protocole de
-bascule de jeton multi-parties sécurisé dans le but d'incorporer un caractère aléatoire dans l'élection du leader.
+sûreté prouvée basés sur la preuve d'enjeu (PoS ou \emph{Proof of Stake}). Ouroboros utilise un protocole de
+bascule de jeton multi-parties sécurisé dans le but d'incorporer un caractère aléatoire dans l'élection du dirigeant.
 Son évolution Ouroboros Praos~\cite{DavidGKR18} apporte la sûreté en présence d'attaquants entièrement adaptatifs.
-HoneyBadger~\cite{MillerXCSS16} démontre une bonne vitalité dans un réseau comportant des latences hétérogènes.
+HoneyBadger~\cite{MillerXCSS16} fournit une bonne vitalité dans un réseau comportant des latences hétérogènes.
 
-Tendermint~\cite{buchman2016tendermint, 1807.04938} change de leader à chaque bloc et a démontré ses qualités dans une
-configuration à 64 noeuds. Ripple~\cite{schwartz2014ripple} a une faible latence grâce à l'utilisation de sous-réseaux
-certifiés collectivement au sein d'un réseau plus large. La société Ripple génère une liste de noeuds vérifiés par
+Tendermint~\cite{buchman2016tendermint, 1807.04938} change de dirigeant à chaque bloc et a démontré ses qualités dans une
+configuration à 64 nœuds. Ripple~\cite{schwartz2014ripple} a une faible latence grâce à l'utilisation de sous-réseaux
+certifiés collectivement au sein d'un réseau plus large. La société Ripple génère une liste de nœuds vérifiés par
 défaut renouvelée peu souvent, qui en fait un réseau essentiellement centralisé.
@@
 distribution linéaire et simplifie drastiquement les caractéristiques du protocole, même si les limitations liées
-à l'élection d'un leader persistent. Facebook utilise Hotstuff comme le consensus principal de son projet Libra.
+à l'élection d'un dirigeant persistent. Facebook utilise Hotstuff comme le consensus principal de son projet Libra.
 
@@
 opère aussi rapidement que le réseau se propage. Stellar~\cite{mazieres2015stellar} se base sur un accord byzantin
-fédéré dans lequel des \emph{tranches de quorum} permettent une confiance hétérogène entre des noeuds différents. La
+fédéré dans lequel des \emph{tranches de quorum} permettent une confiance hétérogène entre des nœuds différents. La
 sûreté de fonctionnement est garantie lorsque les transactions peuvent être reliées entre elles de manière transitive
 par des tranches de quorum. Algorand~\cite{GiladHMVZ17} utilise une fonction aléatoire vérifiable pour sélectionner un
-comité de noeuds qui participent à un protocole de consensus byzantin innovant.
+comité de nœuds qui participent à un protocole de consensus byzantin innovant.
 
@@
 avec une meilleure résistance aux attaques à la vitalité~\cite{confluxLLXLC18}.
-
-Avalanche est différent dans son approche par laquelle le résultat du vote est un reçu instantané déterminé par une
-requête sans preuve de travail, tandis que les votes dans PHANTOM ou Conflux sont purement déterminés par la preuve de
-travail dans la structure de transactions. A la manière de Thunderella, Meshcash~\cite{BentovHMN17} combine un protocole
+Avalanche est différent en ce que le résultat du vote est un chit à usage unique déterminé par une
+requête sans preuve de travail, alors que les votes dans PHANTOM ou Conflux sont purement déterminés par le PoW
+dans la structure des transactions. A la manière de Thunderella, Meshcash~\cite{BentovHMN17} combine un protocole
 lent basé sur la preuve de travail et un protocole de consensus rapide qui apporte une fréquence de génération de blocs
 élevée quelque soit la latence du réseau, offrant un temps de confirmation rapide. Hashgraph~\cite{baird2016hashgraph}
-est un protocole sans leader qui construit un DAG sur la base d'un "bavardage rendu aléatoire". Il nécessite une
-connaissance complète des membres du réseau à tout moment, et comme Ben-Or~\cite{ben1983another}, il s'attend à un
+est un protocole sans dirigeant qui construit un DAG sur la base d'un \emph{gossip} aléatoire. Il nécessite une
+connaissance complète des membres du réseau à tout moment, et comme Ben-Or~\cite{ben1983another}, il demande un
 nombre de messages exponentiel~\cite{aspnes2003randomized,CachinV17} pour fonctionner.

### 2020-11-15 paper_fr.tex
Relecture finale

@@
 %O This paper introduces a family of leaderless Byzantine fault tolerance protocols, built around a metastable mechanism via network subsampling.
-Cet article présente une famille de protocoles à tolérance de faute byzantine sans dirigeant, bâtis autour d'un mécanisme métastable \emph{via} sous-échantillonnage réseau.
+Cet article présente une famille de protocoles à tolérance de faute byzantine sans dirigeant, bâtis autour d'un méca\-nisme métastable \emph{via} sous-échantillonnage réseau.
 %O These protocols provide a strong probabilistic safety guarantee in the presence of Byzantine adversaries while their concurrent and leaderless nature enables them to achieve high throughput and scalability.
@@
 It does not require accurate knowledge of all participants and exposes new possible tradeoffs and improvements in safety and liveness for building consensus protocols.
-Contrairement aux protocoles de consensus traditionnels où, typiquement, un ou plusieurs nœuds traitent linéairement l'information pour chaque décision, aucun nœud ne traite l'information d'une manière plus complexe que logarithmique.
+Contrairement aux protocoles de consensus traditionnels où, typiquement, un ou plusieurs nœuds traitent linéairement l'information pour chaque décision, aucun nœud ne traite l'information d'une manière plus complexe que loga\-rithmiquement.
 Il ne requiert pas de connaissance précise de tous les participants et propose de nouveaux compromis et améliorations en termes de sûreté de fonctionnement et de vitalité pour la construction de nouveaux protocoles de consensus.
@@
 %O Inspired by gossip algorithms, this family gains its properties through a deliberately metastable mechanism.
-Inspirée par les algorithmes de \emph{gossip} ou bavardage, cette famille possède des propriétés grâce à un mécanisme délibérément métastable.
+Ins\-pirée par les algorithmes de \emph{gossip} ou bavardage, cette famille possède des propriétés grâce à un mécanisme délibérément métastable.
 %O Specifically, the system operates by repeatedly sampling the network at random, and steering correct nodes towards a common outcome.
@@
 %O Similar to Nakamoto consensus, the Snow protocol family provides a probabilistic safety guarantee, using a tunable security parameter that can render the possibility of a consensus failure arbitrarily small.
-De même que le consensus Nakamoto, la famille des protocoles Snow fournissent une garantie de sûreté probabiliste, grâce à un paramètre de sécurité configurable qui peut amenuiser à volonté la possibilité d'un échec du consensus.
+De même que le consensus Nakamoto, la famille des protocoles Snow fournit une garantie de sûreté probabiliste, grâce à un paramètre de sécurité configurable qui peut amenuiser à volonté la possibilité d'un échec du consensus.
 %O Unlike Nakamoto consensus, the protocols are green, quiescent and efficient; they do not rely on proof-of-work~\cite{DworkN92} and do not consume energy when there are no decisions to be made.
@@
 %O Snow's subsampled voting mechanism has two additional properties that improve on previous approaches for consensus. 
-Le mécanisme de vote sous-échantillonné possède deux propriété supplémentaires supérieures par rapport aux approches antérieures du consensus.
+Le mécanisme de vote sous-échantillonné possède deux propriétés additionnelles supérieures par rapport aux approches antérieures du consensus.
 %O Whereas the safety of quorum-based approaches breaks down immediately when the predetermined threshold $f$ is exceeded,
@@
 %O Section~\ref{sec:related-work} presents related work, and finally, Section~\ref{sec:conclusions} summarizes our contributions.
-La section~\ref{sec:related-work} présente les travaux en rapport, et, finalement, la section~\ref{sec:conclusions} résume nos contributions.
+La section~\ref{sec:related-work} présente les travaux existants en rapport avec cette étude et, finalement, la section~\ref{sec:conclusions} résume nos contributions.
 
@@
 %OFigure~\ref{fig:fandepsilon} shows how the portion ($f/n$) of misbehaving participants (or computation power) affects the probability of system safety failure (decision of two conflicting proposals), given a choice of finality.
-La figure~\ref{fig:fandepsilon} montre comment la portion ($f/n$) de participants (ou de puissance informatique) problématique affecte la probabilité d'une défaillance de la sûreté du système (décision de deux propositions incompatibles), étant donné un choix de finalités.
+La figure~\ref{fig:fandepsilon} montre comment la portion ($f/n$) de participants (ou de puissance informatique) problématique affecte la probabilité d'une défaillance de la sûreté du système (décision concernant deux propositions incompatibles), étant donné un choix de finalités.
 % The maximum tolerated Byzantine presence $f$, is dependent on the choice of $\varepsilon$.
@@
 Les conclusions de ce théorème démontrent une propriété clef du système: une fois que le réseau perd sa bivalence
-(ex: $\delta > 0$), il tend à se renverser et à converger rapidement vers la couleur majoritaire, et il devient alors
+(par exem\-ple\,: $\delta > 0$), il tend à se renverser et à converger rapidement vers la couleur majoritaire, et il devient alors
 suffisamment improbable qu'il rebascule vers la couleur minoritaire. C'est cette propriété fondamentale que nous
@@
 Dans le cas de Snowflake, nous prenons comme hypothèse qu'une partie des nœuds est antagoniste. Avec Slush, une
-fois que le réseau obtient une majorité significative pour l'une des propositions (ex: la couleur bleue), il devient
-improbable qu'une proposition minoritaire (ex: la couleur rouge) puisse être adoptée dans le futur (irréversibilité).
+fois que le réseau obtient une majorité significative pour l'une des propositions (par exemple\,: la couleur bleue), il devient
+improbable qu'une proposition minoritaire (par exemple\,: la couleur rouge) puisse être adoptée dans le futur (irréversibilité).
 De plus, les nœuds Slush ont simplement besoin d'exécuter le protocole pour un nombre déterministe $m$ de tours qui
@@
 \noindent \emph{Ébauche de preuve}. Nous définissons la stratégie d'échec comme étant l'événement lors duquel deux
-nœuds légitimes $u$ et $v$ décident entre le bleu et le rouge, ex: $\mathcal{D}(u, \mathcal{S}_t, bleu) \rightarrow 1$
+nœuds légitimes $u$ et $v$ décident entre le bleu et le rouge, par exemple\,: $\mathcal{D}(u, \mathcal{S}_t, bleu) \rightarrow 1$
 et $\mathcal{D}(v, \mathcal{S}_{t'}, rouge) \rightarrow 1$, pour deux instants $t$ et $t'$. Nous modélisons une fois de
@@
 Nous démontrons qu'en choisissant correctement $k$, $\alpha$, et $\beta$, nous pouvons construire des instances
-extrêmement sécurisées de Snowflake (ex: échec de sécurité avec une probabilité de $\leq \varepsilon$) quand le réseau
+extrêmement sécurisées de Snowflake (par exemple\,: échec de sécurité avec une probabilité de $\leq \varepsilon$) quand le réseau
 atteint un biais de $\delta$, comme le montre la figure~\ref{fig:states_feasible_solutions}. Un exemple concret est
@@
 de paiement décentralisé, comme nous le démontrons en section~\ref{sec:implementation}.
-Finalement, nous développons le roulement, dans le sens de renouvellement des nouds, et les divergences de vues en annexe~\ref{sec:full-analysis-churn}.
+Finalement, nous développons le roulement, dans le sens de renouvellement des nœuds, et les divergences de vues en annexe~\ref{sec:full-analysis-churn}.
 
@@
 %O We inherit the transaction and address mechanisms from Bitcoin. At their simplest, transactions consist of multiple inputs and outputs, with corresponding redeem scripts.
-Nous héritons les mécanismes de transaction et d'adresse de Bitcoin. En simplifiant à l'extrême, les transactions consistent en multiples entrées et sorties, avec les scripts d'échange correspondants.
+Nous héritons des mécanismes de transaction et d'adresse de Bitcoin. En simplifiant à l'extrême, les transactions consistent en multiples entrées et sorties, avec les scripts d'échange correspondants.
 %O Addresses are identified by the hash of their public keys, and signatures are generated by corresponding private keys.
@@
 %O Unlike in Slush, where it is clear that once the network has reached one of the converging states and therefore may not revert back, this no longer applies to Snowflake, since any adversary $f \geq \alpha$ has strictly positive probability of reverting the system, albeit this probability may be infinitesimally small.
-Contrairement à Slush, où il est clair qu'une fois que le réseau a atteint l'un des états convergents et ne reviendra donc pas en arrière, cela ne s'applique pas à Snowflake, puisque tout adversaire $f \geq \alpha$ a une probabilité strictement positive d'inverser la tendance du système, fût-elle infinitésimale.
+Contrairement à Slush, où il est clair qu'une fois que le réseau a atteint l'un des états convergents, il ne reviendra pas en arrière, cela ne s'applique pas à Snowflake, puisque tout adversaire $f \geq \alpha$ a une probabilité strictement positive d'inverser la tendance du système, fût-elle infinitésimale.
 %O The CTMC is flexible enough to deal with a system where there is only one absorbing state, but the long-term behavior of the system is no longer meaningful since, after an infinite amount of time, the system is guaranteed to revert, violating safety.

### 2020-11-15 sections/evaluation_fr.tex
Relecture finale

@@
 détaillé dans un travail préliminaire~\cite{GiladHMVZ17}.
-Nous avons sélectionné 20 villes majeures qui semblent être localisées à proximité d'un nombre conséquent de noeuds
+Nous avons sélectionné 20 grandes villes qui semblent être localisées à proximité d'un nombre conséquent de noeuds
 Bitcoin joignables, d'après~\cite{bitnodes2018}. Ces villes couvrent l'Amérique du Nord, l'Europe, l'Asie occidentale,
@@
 d'Algorand repose sur un accord byzantin basé sur un quorum, et Conflux supplémente le consensus de Nakamoto par une
-structure de type Graphe Acyclique Orienté (DAG) pour permettre un meilleur débit, tandis qu'Avalanche appartient à
+structure de type graphe orienté acyclique (DAG) pour permettre un meilleur débit, tandis qu'Avalanche appartient à
 une nouvelle famille de protocoles basés sur la méta-stabilité. Par ailleurs, nous utilisons

### 2020-11-15 sections/related-work_fr.tex
Relecture finale

@@
 transactions rapidement. ByzCoin~\cite{Kokoris-KogiasJ16} sélectionne un échantillon réduit de participants de manière
-périodique et exécute ensuite un protocole résistant aux fautes byzantines en pratique (PBFT) entre les nœuds
+périodique et exécute ensuite un protocole similaire à PBFT entre les nœuds
 sélectionnés.
@@
 D'autres travaux tentent d'introduire de nouveaux protocoles qui redéfinissent et relâchent les contraintes du modèle
-de résistance aux fautes byzantines. La résistance byzantine à grande échelle~\cite{rodrigues2007large} modifie
+de résistance aux fautes byzantines. Large-scale BFT~\cite{rodrigues2007large} modifie
 PBFT pour autoriser un nombre quelconque de répliques et un seuil d'échec en
@@
 fonctionnement avec une probabilité élevée. Dans une autre forme de relâchement des contraintes,
-Zeno~\cite{singh2009zeno} introduit un protocole de réplication de machine d'état résistante aux fautes byzantines qui
+Zeno~\cite{singh2009zeno} introduit un protocole de réplication de machine à état résistante aux fautes byzantines qui
 échange la cohérence pour une meilleure disponibilité. Plus spécifiquement, Zeno garantit une cohérence assurée au final
-plutôt qu'une linéarisation de celle-ci, ce qui veut dire que les participants n'ont pas besoin d'être conhérents mais
+plutôt qu'une linéarisation de celle-ci, ce qui veut dire que les participants n'ont pas besoin d'être cohérents mais
 doivent se mettrent finalement d'accord une fois le réseau stabilisé. En fournissant une garantie de cohérence encore
@@
 Dans une configuration synchrone, inspirée par Hoststuff, Sync HotStuff~\cite{synchotstuff} atteint un consensus en
-un temps $2\Delta$ avec un coût quadratique et contrairement à d'autres protocoles synchrones à étapes vérouillées, il
+un temps $2\Delta$ avec un coût quadratique et contrairement à d'autres protocoles synchrones à étapes verrouillées, il
 opère aussi rapidement que le réseau se propage. Stellar~\cite{mazieres2015stellar} se base sur un accord byzantin
@@
 lent basé sur la preuve de travail et un protocole de consensus rapide qui apporte une fréquence de génération de blocs
-élevée quelque soit la latence du réseau, offrant un temps de confirmation rapide. Hashgraph~\cite{baird2016hashgraph}
+élevée quelle que soit la latence du réseau, offrant un temps de confirmation rapide. Hashgraph~\cite{baird2016hashgraph}
 est un protocole sans dirigeant qui construit un DAG sur la base d'un \emph{gossip} aléatoire. Il nécessite une

### 2020-11-15 avalanche_fr.tex
Relecture finale
Après relecture par Nicolas

@@
 %O We write $T' \gets T$ if $T$ has a parent edge to transaction $T'$,
-Nous écrivons $T' \gets T$ si $T$ a une arête commune avev la transaction $T'$.
+Nous écrivons $T' \gets T$ si $T$ a une arête commune avec la transaction $T'$.
 %O The ``$\stackrel{*}{\gets}$''-relation is its reflexive transitive closure, indicating a path from $T$ to $T'$.
@@
 %O The above process will yield a labeling of the DAG with a chit value and associated confidence for each transaction~$T$.
-Le processus ci-dessus ramène un étiquetage du DAG avec une valeur de chit et une confiance associée pour chaque transaction~$T$.
+Le processus ci-dessus étiquette le DAG avec une valeur de chit et une confiance associée pour chaque transaction~$T$.
 
@@
 At most one transaction in each shaded region will be accepted.}
-    \captionof{figure}{Exemple de valeurs pour $\langle \textrm{chit}, \textrm{confidence}\rangle$.
+    \captionof{figure}{Exemple de valeurs pour $\langle \textrm{chit}, \textrm{confiance}\rangle$.
 Les boîtes sombres indiquent des transactions dont l'indice de confiance est supérieur.

### 2020-11-15 paper_fr.tex
Après relecture par Nicolas

@@
 Les protocoles de consensus traditionnels
-sont fondés sur une communication entre chaque pair pour s'assurer que tous les nœuds correct arrivent aux mêmes décisions avec une certitude absolue.
+sont fondés sur une communication entre chaque pair pour s'assurer que tous les nœuds corrects arrivent aux mêmes décisions avec une certitude absolue.
 %Because they usually require quadratic communication overhead and accurate knowledge of membership, they have been difficult to scale
@@
 Avec Slush, un nœud démarre dans un état de couleur neutre. Après réception d'une transaction envoyée par un client,
-un nœud de couleur neutre adopte la couleur définie dans la transaction et initie une requête. Pour lancer une
-requête, le nœud choisit un échantillon réduit et de taille constante ($k$) de nœuds appartenant au réseau suivant
-une distribution uniformément aléatoire puis initie sa propre requête, tandis qu'un nœud déja coloré répond
+un nœud de couleur neutre adopte la couleur définie dans la transaction et initie une requête. 
+Pour lancer une requête, le nœud choisit un échantillon réduit et de taille constante ($k$) de nœuds appartenant au réseau selon une distribution uniformément aléatoire, puis lance un message de requête.
+À réception d'une requête, un nœud non coloré adopte la couleur de la requête, répond avec cette couleur et initie sa propre requête,
+tandis qu'un nœud déja coloré répond
 simplement avec sa couleur actuelle. Une fois que le nœud qui fait la requête récupère $k$ réponses, il vérifie si
@@
 \tronly{Finally, more}{More} sophisticated initialization mechanisms would bear fruitful in improving liveness of multi-value consensus.
-Une autre}{Une} serait de caractériser les garanties du système face à un adversaire dont les pouvoirs seraient limités de manière réaliste, à la suite de quoi les performances se verraient encode améliorées.
+Une autre}{Une} serait de caractériser les garanties du système face à un adversaire dont les pouvoirs seraient limités de manière réaliste, à la suite de quoi les performances se verraient encore améliorées.
 \tronly{D'autres}{Plus} mécanismes d'initialisation sophistiqués seraient souhaitables pour améliorer la vitalité d'un consensus multi-valeurs.
@@
 %O We introduce the decision function $\mathcal{D}(*)$, which is constructed by having each node also keep track of the total number of consecutive times it has sampled a majority of the same color ($\beta$).
-Nous introduisons finalement une fonction appelée $\mathcal{A}(\mathcal{S}_t)$ qui est construite en faisant conserver à chaque nœud la trace du nombre total de fois consécutives qu'il a échantillonné une majorité de la couleur d'échantillon ($\beta$).
+Nous introduisons finalement la fonction 
+de décision $\mathcal{D}(*)$,
+%appelée $\mathcal{A}(\mathcal{S}_t)$ qui est 
+construite en faisant conserver à chaque nœud la trace du nombre total de fois consécutives qu'il a échantillonné une majorité de la couleur d'échantillon ($\beta$).
 %O Finally, we introduce a function called $\mathcal{A}(\mathcal{S}_t)$, the adversarial strategy, that takes as parameters the entire configuration of the network at time $t$, as well as the next set of nodes chosen by the scheduler to execute, and as a side-effect, modifies the set of nodes $\mathcal{B}$ to some arbitrary configuration of colors.

### 2020-11-15 sections/evaluation_fr.tex
Après relecture par Nicolas

@@
 sont tous fondamentalement différents dans leur conception. L'algorithme de consensus par comité
-d'Algorand repose sur un accord byzantin basé sur un quorum, et Conflux supplémente le consensus de Nakamoto par une
+d'Algorand repose sur un consensus BFT classique basé sur un quorum, et Conflux supplémente le consensus de Nakamoto par une
 structure de type graphe orienté acyclique (DAG) pour permettre un meilleur débit, tandis qu'Avalanche appartient à
